### 2016-10-17 Analysis/CharactersCorrelation.Rmd
Rmd file (INIT draft from BS)

---
title: "Characters Correlation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

Characters Correlation
===============

What is the effect of morphological characters correlation on phylogenetic inference?

<!-- Before starting
===============

Loading the package
-----------------------------------

```{r, eval = FALSE}
# The latest version of devtools
if(!require(devtools)) install.packages("devtools")

## The latest version of dispRity 
install_github("TGuillerme/dispRity")
```

Loading the data
-----------------------------------

We're just going to use the iris data set here.

```{r}
set.seed(123)

## Loading the data
data(iris)
head(iris, 3)

## Isolating the variables
iris_variables <- iris[, 1:4]
iris_species <- iris[, 5]
```

Ordination
===============

Simple (classic ordination).

```{r}
## Simple basic ordination
iris_ord_simple <- prcomp(iris_variables)

## Relative variance explained by axis
iris_ord_simple_var <- iris_ord_simple[[1]]/sum(iris_ord_simple[[1]])

```

Plotting the ordination axis variance and the ordination

```{r, fig.width=8, fig.height=4}
## Graphical parameters 
op <- par(mfrow = c(1,2), bty = "n")

## Plotting the axis explained variance
barplot(iris_ord_simple_var, main = "explained variance")

## Plotting the ordination (axis 1 and 2)
plot(iris_ord_simple$x[,1], iris_ord_simple$x[,2], main = "2D ordination plot",
     xlab = paste("PC1 (", round(iris_ord_simple_var[1], digit = 2), ")", sep = ""),
     ylab = paste("PC2 (", round(iris_ord_simple_var[2], digit = 2), ")", sep = ""),
     pch = 19, col = palette()[as.factor(iris_species)])

## Adding the eigen vectors
for (rotation in 1:nrow(iris_ord_simple$rotation)) {
    # Isolating the coordinates
    coord_x <- iris_ord_simple$rotation[rotation,1]
    coord_y <- iris_ord_simple$rotation[rotation,2]
    # Eigen vectors
    lines(c(0, coord_x), c(0, coord_y), col = palette()[4])
    # Contributors
    text(x = coord_x*1.1, y = coord_y*1.1,
         rownames(iris_ord_simple$rotation)[rotation], cex =0.8)
}
par(op)
```

Note that here, Length and Width are kind of collinear and Sepal/Petal are orthogonal.

"Jack-knifing the ordination"
===============

Removing one variable each time for performing the ordination

```{r}
## Create a list of matrices minus one variable
matrix_list <- list(iris_variables[,-1],
                    iris_variables[,-2],
                    iris_variables[,-3],
                    iris_variables[,-4])
names(matrix_list) <- c("No_Sep.Len", "No_Sep.Wid", "No_Pet.Len", "No_Pet.Wid")

## Simple basic ordination
iris_ord_JN <- lapply(matrix_list, prcomp)

## Relative variance explained by axis
iris_ord_JN_var <- lapply(iris_ord_JN, function(X) X[[1]]/sum(X[[1]]))

```

Plotting the ordination axis variances

```{r, fig.width=8, fig.height=8}
## Graphical parameters 
op <- par(mfrow = c(2,2), bty = "n")
## Plotting the axis explained variance
for(JN in 1:length(matrix_list)) {
    barplot(iris_ord_JN_var[[JN]], main = paste("Var.", names(iris_ord_JN_var)[JN]))
}
par(op)
```

```{r, fig.width=8, fig.height=8}
## Graphical parameters 
op <- par(mfrow = c(2,2), bty = "n")

for(JN in 1:length(matrix_list)) {
    ## Plotting the ordination (axis 1 and 2)
    plot(iris_ord_JN[[JN]]$x[,1], iris_ord_JN[[JN]]$x[,2], main = names(iris_ord_JN)[JN],
         xlab = paste("PC1 (", round(iris_ord_JN[[JN]][[1]][1]/sum(iris_ord_JN[[JN]][[1]]),
         digit = 2), ")", sep = ""), ylab = paste("PC2 (",
         round(iris_ord_JN[[JN]][[1]][2]/sum(iris_ord_JN[[JN]][[1]]), digit = 2), ")", sep = ""),
         pch = 19, col = palette()[as.factor(iris_species)])

    ## Adding the eigen vectors
    for (rotation in 1:nrow(iris_ord_JN[[JN]]$rotation)) {
        # Isolating the coordinates
        coord_x <- iris_ord_JN[[JN]]$rotation[rotation,1]
        coord_y <- iris_ord_JN[[JN]]$rotation[rotation,2]
        # Eigen vectors
        lines(c(0, coord_x), c(0, coord_y), col = palette()[4])
        # Contributors
        text(x = coord_x*1.1, y = coord_y*1.1,
             rownames(iris_ord_JN[[JN]]$rotation)[rotation], cex =0.8)
    }
}
par(op)
```

And note here that collinearity and orthogonality is not quiet as much conserved!

A more complex example (from Anderson et al 2011 Nature)
===============

Loading the data
-----------------------------------

```{r}
## Loading the data
anderson_data <- read.csv("Anderson.et.al2011-Nature.csv", header = TRUE)
## Selecting the data subset
and_subdata <- anderson_data[,c(1,2,6:16)]
## Removing NAs (for the example here, shouldn't be a problem normaly)
data_clean <- and_subdata[-which(apply(apply(and_subdata, 1, is.na), 2, any) == TRUE), ]
## And now it looks something like that
head(data_clean[,1:5])
## Make the data a matrix with taxa as rownames (scaled matrix)
data_matrix <- scale(as.matrix(data_clean[-c(1:2)]))
rownames(data_matrix) <- data_clean[,2]
## Separating the phylogenetic groups
data_familly <- as.character(data_clean[,1])
## Fixing typo
data_familly[which(data_familly == "Placoderma")] <- "Placodermi"
```

Ordination
-----------------------------------

Simple ordination.

```{r}
## Simple basic ordination
ord_anderson <- prcomp(data_matrix)
## Relative variance explained by axis
ord_anderson_var <- ord_anderson[[1]]/sum(ord_anderson[[1]])
```

Plotting the ordination axis variance and the ordination.

```{r, fig.width=8, fig.height=4}
## Graphical parameters 
op <- par(mfrow = c(1,2), bty = "n")

## Plotting the axis explained variance
barplot(ord_anderson_var, main = "explained variance")

## Plotting the ordination (axis 1 and 2)
plot(ord_anderson$x[,1], ord_anderson$x[,2], main = "2D ordination plot",
     xlab = paste("PC1 (", round(ord_anderson_var[1], digit = 2), ")", sep = ""),
     ylab = paste("PC2 (", round(ord_anderson_var[2], digit = 2), ")", sep = ""),
     pch = 19, col = palette()[as.factor(data_familly)])

## Adding the eigen vectors
for (rotation in 1:nrow(ord_anderson$rotation)) {
    arrows_scaling = 5
    # Isolating the coordinates
    coord_x <- ord_anderson$rotation[rotation,1]*arrows_scaling
    coord_y <- ord_anderson$rotation[rotation,2]*arrows_scaling
    # Eigen vectors
    lines(c(0, coord_x), c(0, coord_y), col = palette()[length(levels(as.factor(data_familly)))+1])
    # Contributors
    text(x = coord_x*1.1, y = coord_y*1.1,
         rownames(ord_anderson$rotation)[rotation], cex =0.8)
}
par(op)
```

The variables are length and ratios of the jaw ([details here](http://www.nature.com/nature/journal/v476/n7359/extref/nature10207-s1.pdf)).

"Jack-knifing the ordination"
===============

Removing one variable each time for performing the ordination

```{r}
## Create a list of matrices minus one variable
matrix_list <- list(data_matrix[,-1],
                    data_matrix[,-2],
                    data_matrix[,-3],
                    data_matrix[,-4],
                    data_matrix[,-5],
                    data_matrix[,-6],
                    data_matrix[,-7],
                    data_matrix[,-8],
                    data_matrix[,-9],
                    data_matrix[,-10],
                    data_matrix[,-11])
names(matrix_list) <- c("-C1", "-C2", "-C3", "-C4", "-C5", "-C6", "-C7", "-C8", "-C9", "-C10",
    "-C11")

## Simple basic ordination
ord_anderson_JN <- lapply(matrix_list, prcomp)

## Relative variance explained by axis
ord_anderson_JN_var <- lapply(ord_anderson_JN, function(X) X[[1]]/sum(X[[1]]))
```

Plotting the ordination axis variances

```{r, fig.width=8, fig.height=8}
## Plot part 1
## Graphical parameters 
op <- par(mfrow = c(2,2), bty = "n")
## Plotting the axis explained variance
for(JN in 1:4) {
    barplot(ord_anderson_JN_var[[JN]], main = paste("Var.", names(ord_anderson_JN_var)[JN]))
}
par(op)
```
```{r, fig.width=8, fig.height=8}
## Plot part 2
## Graphical parameters 
op <- par(mfrow = c(2,2), bty = "n")
## Plotting the axis explained variance
for(JN in 5:8) {
    barplot(ord_anderson_JN_var[[JN]], main = paste("Var.", names(ord_anderson_JN_var)[JN]))
}
par(op)
```
```{r, fig.width=8, fig.height=8}
## Plot part 3
## Graphical parameters 
op <- par(mfrow = c(2,2), bty = "n")
## Plotting the axis explained variance
for(JN in 9:11) {
    barplot(ord_anderson_JN_var[[JN]], main = paste("Var.", names(ord_anderson_JN_var)[JN]))
}
par(op)
```

```{r, fig.width=8, fig.height=8}
## Plot part 1
## Graphical parameters 
op <- par(mfrow = c(2,2), bty = "n")

for(JN in 1:4) {
    ## Plotting the ordination (axis 1 and 2)
    plot(ord_anderson_JN[[JN]]$x[,1], ord_anderson_JN[[JN]]$x[,2],
         main = names(ord_anderson_JN)[JN], xlab = paste("PC1 (",
         round(ord_anderson_JN[[JN]][[1]][1]/sum(ord_anderson_JN[[JN]][[1]]), digit = 2), ")",
         sep = ""), ylab = paste("PC2 (",
         round(ord_anderson_JN[[JN]][[1]][2]/sum(ord_anderson_JN[[JN]][[1]]), digit = 2), ")",
         sep = ""), pch = 19, col = palette()[as.factor(data_familly)])

    ## Adding the eigen vectors
    for (rotation in 1:nrow(ord_anderson_JN[[JN]]$rotation)) {
        arrows_scaling = 5
        # Isolating the coordinates
        coord_x <- ord_anderson_JN[[JN]]$rotation[rotation,1]*arrows_scaling
        coord_y <- ord_anderson_JN[[JN]]$rotation[rotation,2]*arrows_scaling
        # Eigen vectors
        lines(c(0, coord_x), c(0, coord_y),
            col = palette()[length(levels(as.factor(data_familly)))+1])
        # Contributors
        text(x = coord_x*1.1, y = coord_y*1.1,
             rownames(ord_anderson_JN[[JN]]$rotation)[rotation], cex =0.8)
    }
}
par(op)
```

```{r, fig.width=8, fig.height=8}
## Plot part 2
## Graphical parameters 
op <- par(mfrow = c(2,2), bty = "n")

for(JN in 5:8) {
    ## Plotting the ordination (axis 1 and 2)
    plot(ord_anderson_JN[[JN]]$x[,1], ord_anderson_JN[[JN]]$x[,2],
         main = names(ord_anderson_JN)[JN], xlab = paste("PC1 (",
         round(ord_anderson_JN[[JN]][[1]][1]/sum(ord_anderson_JN[[JN]][[1]]), digit = 2), ")",
         sep = ""), ylab = paste("PC2 (",
         round(ord_anderson_JN[[JN]][[1]][2]/sum(ord_anderson_JN[[JN]][[1]]), digit = 2), ")",
         sep = ""), pch = 19, col = palette()[as.factor(data_familly)])

    ## Adding the eigen vectors
    for (rotation in 1:nrow(ord_anderson_JN[[JN]]$rotation)) {
        arrows_scaling = 5
        # Isolating the coordinates
        coord_x <- ord_anderson_JN[[JN]]$rotation[rotation,1]*arrows_scaling
        coord_y <- ord_anderson_JN[[JN]]$rotation[rotation,2]*arrows_scaling
        # Eigen vectors
        lines(c(0, coord_x), c(0, coord_y),
            col = palette()[length(levels(as.factor(data_familly)))+1])
        # Contributors
        text(x = coord_x*1.1, y = coord_y*1.1,
             rownames(ord_anderson_JN[[JN]]$rotation)[rotation], cex =0.8)
    }
}
par(op)
```

```{r, fig.width=8, fig.height=8}
## Plot part 3
## Graphical parameters 
op <- par(mfrow = c(2,2), bty = "n")

for(JN in 9:11) {
    ## Plotting the ordination (axis 1 and 2)
    plot(ord_anderson_JN[[JN]]$x[,1], ord_anderson_JN[[JN]]$x[,2],
         main = names(ord_anderson_JN)[JN], xlab = paste("PC1 (",
         round(ord_anderson_JN[[JN]][[1]][1]/sum(ord_anderson_JN[[JN]][[1]]), digit = 2), ")",
         sep = ""), ylab = paste("PC2 (",
         round(ord_anderson_JN[[JN]][[1]][2]/sum(ord_anderson_JN[[JN]][[1]]), digit = 2), ")",
         sep = ""), pch = 19, col = palette()[as.factor(data_familly)])

    ## Adding the eigen vectors
    for (rotation in 1:nrow(ord_anderson_JN[[JN]]$rotation)) {
        arrows_scaling = 5
        # Isolating the coordinates
        coord_x <- ord_anderson_JN[[JN]]$rotation[rotation,1]*arrows_scaling
        coord_y <- ord_anderson_JN[[JN]]$rotation[rotation,2]*arrows_scaling
        # Eigen vectors
        lines(c(0, coord_x), c(0, coord_y),
            col = palette()[length(levels(as.factor(data_familly)))+1])
        # Contributors
        text(x = coord_x*1.1, y = coord_y*1.1,
             rownames(ord_anderson_JN[[JN]]$rotation)[rotation], cex =0.8)
    }
}
par(op)
```
-->

### 2016-10-19 Writing/draft.tex
CD formula

\documentclass[12pt,letterpaper]{article}
\usepackage{natbib}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{dcolumn}
\usepackage{mathtools}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\newcolumntype{d}[1]{D{.}{.}{#1}}

\pagenumbering{arabic}

%Pagination style and stuff
\linespread{2}
\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}
%\bibpunct{(}{)}{;}{a}{}{,}

%---------------------------------------------
%
%       START
%
%---------------------------------------------

\begin{document}

%Running head
\begin{flushright}
Version dated: \today
\end{flushright}
\bigskip
\noindent RH: Characters correlation

\bigskip
\medskip
%\begin{center}

%\noindent{\Large \bf Tree rearrangements rearranged} %TG: I'm bad at puns.
\bigskip

% \noindent {\normalsize \sc Thomas Guillerme$^1$$^*$, and Martin D. Brazeau$^1$}\\ %TG: Author order can be swapped of course! There's only a finite combination of 2 elements anyway!
% \noindent {\small \it 
% $^1$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\}
% \end{center}
% \medskip
% \noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\  %TG: Same as above
\vspace{1in}

%Line numbering
\modulolinenumbers[1]
\linenumbers

%---------------------------------------------
%
%       ABSTRACT
%
%---------------------------------------------

\newpage
\begin{abstract}
blablabla
\end{abstract}

\noindent (Keywords: )\\

\vspace{1.5in}

\newpage

%---------------------------------------------
% LaTeX tips for modifying/editing the document:
%---------------------------------------------
% - You can comment using the percentage sign. I suggest you use the % sign alone for commenting out sections of the text:
%       e.g. "This is a really long sentence %because this sentence is very long." Here the % is used for ignoring the end of the sentence (but for some reason you want to keep track of it).%
%       For comments as in verbose comments, I suggest you use "%MB:":
%       e.g. "This is a really long sentence %because this sentence is very long. %MB: yeah, no shit!" 
% - For optimal version control, write only one sentence per line (for more precise track changes)
% - To build the pdf, use command+B in Sublime.
% - Because of the bibliography, the pdf needs to be build in the same folder that contains "References.bib" and "sysbio.bst"
% - For citing papers, you must put their bibtex reference in the "References.bib" file and then you can use the following sysbio tags:
%        \cite{bibtexBob2000} for citing within a sentence: "Bob (2000)"
%        \citep{bibtexBob2000} for citing within brackets: "(Bob, 2000)"
%        \citep[Before:][-After]{bibtexBob2000} for citing within brackets with additional text: "(Before: Bob, 2000 -After)"
%        \citealt{bibtexBob2000} for citing without brackets: "Bob, 2000"
%        You can put more cites in each \cite tag by separating them with commas.
% - For equation, find every details here: https://en.wikibooks.org/wiki/LaTeX/Mathematics
% - For titles and stuff, the hierarchy goes \section{}, \subsection{}, \subsubsection{} and so forth...
% - For bullet points or enumerations you can use:
%       \begin{itemize}
%           \item my first bullet point/enumeration
%       \end{itemize}
%       With replacing "itemize" by "enumerate" for enumeration.

%---------------------------------------------
%
%       INTRODUCTION
%
%---------------------------------------------
\section{Introduction}

\section{Methods}

\subsection{Characters difference (CD)}

\begin{equation}
    CD = 1 - \left(\frac{\abs*{\frac{\abs*{\sum_{i}^{n}(x_{i} - y_{i})}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
\end{equation}

\noindent Where $n$ is the number of taxa and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.

% This operation can be repeated for all the internal edges in the original tree, thus the maximum TBR can be:
% \begin{equation}
%     \text{Maximum TBR} = \overbrace{(n-3)}^{\text{internal edges}} \sum_{i=2}^{n-2} (2i-3)(2(n-i)-3) %TG: Check! Summation might be incorrect
% \end{equation}
% Where $i$ and $n-i$ corresponds respectively to the size of the smallest and biggest trees resulting from the bisection the original tree.

\section{Acknowledgments}
European Research Council under the European Union’s Seventh Framework Programme (FP/2007–2013)/ERC Grant Agreement number 311092.

% \bibliographystyle{sysbio}
% \bibliography{References}

\end{document}

### 2016-10-19 Analysis/Sandbox.R
Counting characters differences

@@ -64,14 +64,5 @@ set.correlation <- function(matrix, nchar, cor.dis) {
 
 
 char.diff <- function(X,Y) {
-    if(length(X) != length(Y)) {
-        stop("Characters have not the same length!")
-    }
-    if(class(X) != "numeric" | class(Y) != "numeric") {
-        stop("Characters are not numeric!")
-    }
-
-    abs(sum(X-Y))
-
-    return(1 - ( (abs(sum(X-Y))/length(X)-0.5)/0.5 ))
+    return(1 - ( abs(sum(abs(X-Y))/length(X)-0.5)/0.5 )) #NOT GOOD!
 }

### 2016-11-09 Analysis/Character_diff_empirical.Rmd
Character difference on empirical data

---
title: "Characters differences in empirical matrices"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 12
---

Before starting
===============

Loading the package
-----------------------------------

```{r, eval = FALSE}
## The latest version of devtools
if(!require(devtools)) install.packages("devtools")

## The latest version of dispRity 
install_github("TGuillerme/dispRity")

## The latest version of Claddis
install_github("graemetlloyd/Claddis")
```

Loading the functions
-----------------------------------

```{r}
library(Claddis) ; library(dispRity)
source("functions.R") ; load.functions(test = FALSE)

## Function for helping loading the matrices
find.non.numerics <- function(x) {
    options(warn = -1)
    if(is.na(as.numeric(x))) {
        return(TRUE)
    } else {
        return(FALSE)
    }
}

## Functions for helping the density plots
get.max.x <- function(density) {
    return(max(density$x))
}

get.max.y <- function(density) {
    return(max(density$y))
}

get.min.x <- function(density) {
    return(min(density$x))
}

get.min.y <- function(density) {
    return(min(density$y))
}
```

Characters differences in empirical matrices
===============

Loading the matrices
-----------------------------------

We are going to use 100 matrices from TreeBase.
The matrices are the ones from Guillerme & Cooper 2016 (MPE) with different matrices sizes and different focal groups.

```{r}
## Set up the matrices path
matrix_path = "../Data/100_Matrices"
files <- list.files(path = matrix_path, pattern = ".nex")

## Read the matrices
matrices <- list()
for(matrix in 1:length(files)) {
    matrices[[matrix]] <- ReadMorphNexus(paste(matrix_path, files[matrix], sep ="/"))$matrix
    matrices[[matrix]][apply(matrices[[matrix]], c(1,2), find.non.numerics)] <- "?"
}
```

Measuring the correlations distributions
-----------------------------------

We are using the character differences metric (see the other vignette for more details) for comparing how different characters information is.
A difference of 0 means the characters convey the same phylogenetic information and a difference of 1 means they convey the most different information (i.e. not the opposite).

```{r}
## Calculate the differences
differences <- lapply(matrices, cor.matrix, char.diff)

## Estimate the distributions densities
densities <- lapply(differences, density, na.rm = TRUE)
```

We can then plot the distributions of the differences as so:

```{r, fig.width=10, fig.height=10}
## Plot the the densities distribution
op <- par(bty = "n")
## Empty plot
plot(1,1, col = "white",
    xlim = c(min(unlist(lapply(densities, get.min.x))), max(unlist(lapply(densities, get.max.x)))),
    ylim = c(min(unlist(lapply(densities, get.min.y))), max(unlist(lapply(densities, get.max.y)))),
    main = "Observed characters differences in 100 matrices",
    xlab = "Character differences", ylab = "Density")
## Plotting the density curves
silent <- lapply(densities, lines, col = "grey")
## Plotting the cumulative density
lines(density(na.omit(unlist(differences))), col = "black")
## Adding the legend
legend("topleft", legend = c("Combined", "Individual"), lty = 1, col = c("black", "grey"))
par(op)
```

### 2016-11-09 Functions/char.diff.R
Missing data implementation

@@ -3,9 +3,11 @@
 #' @description Plots the correlation matrix.
 #'
 #' @param X,Y Two morphological characters.
+#' @param missing The algorithm to deal with missing characters. Either \code{"Max"} or \code{"Gower"}. See details.
 #' 
 #' @details
-#' \emph{unknown differences}: when unknown tokens are found ("?"), character state are replaced by a new state not observed (this maximises difference between characters).
+#' \code{missing = "Max"} maximises the difference between two characters by counting the missing character states tokens (\code{"?"}) as a separate state. In this scenario the difference between \code{A = c(0,1,1,1)} and \code{B = c(0,0,?,?)} is always maximise.
+#' \code{missing = "Gower"} uses a Gower (1966) like character comparisons: only the comparable characters states are used. In the same example as above, only \code{A = c(0,1)} and \code{B = c(0,0)} will be used.
 #' 
 #' @examples
 #' ##
@@ -13,23 +15,36 @@
 #' @author Thomas Guillerme
 #' @export
 #' 
+#Gower J.C. 1966. Some distance properties of latent root and vector methods used in multivariate analysis. Biometrika 53:325–338.
+
 
 # TODO: deal with '&' characters
 # TODO: deal with '?' and a minimising algorithm
 
 
-char.diff <- function(X,Y){ 
+char.diff <- function(X,Y, missing = "Max"){ 
 
     # Convert character
     convert.character <- function(X) {
         if(class(X) == "numeric") {
             X <- LETTERS[X+1]
         } else {
-            #if(unknown.difference == "max") {
+            if(missing == "Max") {
                 X <- as.factor(X)
                 levels(X) <- 1:length(levels(X))
                 X <- as.numeric(X)
-            #}
+            } else {
+                options(warn = -1)
+                X <- as.numeric(X)
+                X_NA <- which(is.na(X))
+                if(length(X_NA) != 0) {
+                    NAs <<- X_NA
+                    X <- X[-X_NA]
+                } else {
+                    NAs <<- NULL
+                }
+                options(warn = 0)
+            }
         }
         return(X)
     }
@@ -52,12 +67,35 @@ char.diff <- function(X,Y){
         return(X)
     }
 
+    # Set up missing (if needed)
+    if(missing != "Max") {
+        NA_remove <- NULL
+    }
+
     # Convert the characters to numeric (if needed)
     if(class(X) != "numeric") {
[…]

### 2016-11-09 Functions/cor.matrix.R
Allowing optional arguments

@@ -5,6 +5,7 @@
 #' @param matrix A discrete morphological matrix.
 #' @param diff.fun A function for computing the difference (must intake at least two arguments \code{X} and \code{Y}).
 #' @param verbose A logical value stating whether to be verbose or not (default = \code{FALSE}).
+#' @param ... Any additional arguments to be passed to \code{diff.fun}.
 #'
 #' @return
 #' A pairwise comparison matrix.
@@ -16,12 +17,12 @@
 #' @export
 #' 
 
-cor.matrix <- function(matrix, diff.fun, verbose = FALSE) {
+cor.matrix <- function(matrix, diff.fun, verbose = FALSE, ...) {
 
     # Apply function for the loop
-    apply.diff.fun <- function(x, matrix, verbose, diff.fun) {
+    apply.diff.fun <- function(x, matrix, verbose, diff.fun, ...) {
         # Calculates the difference between two characters taken from the combination matrix (x)
-        difference <- diff.fun(matrix[,x[1]], matrix[,x[2]])
+        difference <- diff.fun(matrix[,x[1]], matrix[,x[2]], ...)
         if(verbose != FALSE) {cat(".")}
         return(difference)
     }
@@ -32,14 +33,14 @@ cor.matrix <- function(matrix, diff.fun, verbose = FALSE) {
     comb_matrix <- combn(1:ncol(matrix), 2)
 
     # Calculating the pairwise differences
-    differences <- apply(comb_matrix, 2, apply.diff.fun, matrix = matrix, verbose = verbose, diff.fun = diff.fun)
+    differences <- apply(comb_matrix, 2, apply.diff.fun, matrix = matrix, verbose = verbose, diff.fun = diff.fun, ...)
 
     if(verbose) {cat("Done.")}
 
     # Create the empty matrix
     differences_matrix <- matrix(data = NA, ncol = ncol(matrix), nrow = ncol(matrix))
     # Set up the diagonal (no difference)
-    diag(differences_matrix) <- diff.fun(matrix[,1], matrix[,1])
+    diag(differences_matrix) <- diff.fun(matrix[,1], matrix[,1], ...)
     # Fill up the lower triangle
     differences_matrix[lower.tri(differences_matrix, diag = FALSE)] <- differences
 

### 2016-11-12 Analysis/char.diff.performance.R
char.diff function finalised

@@ -44,7 +44,7 @@ mul_C <- as.matrix(char.diff.C(matrix_multi))
 
 mul_C == mul_R
 
-matrix <- sim.morpho(rtree(100), 400, states = c(0.85, 0.15), rates = c(rgamma, 10, 5), invariant = FALSE)
+matrix <- sim.morpho(rtree(200), 800, states = c(0.85, 0.15), rates = c(rgamma, 10, 5), invariant = FALSE)
 
 time_R <- system.time(results_R <- cor.matrix(matrix, char.diff))
 time_C <- system.time(results_C <- as.matrix(char.diff.C(matrix)))

### 2016-11-12 Functions/char.diff.R

@@ -22,38 +22,16 @@
 # TODO: deal with '?' and a minimising algorithm
 
 
-dyn.load("char.diff.so")
+#dyn.load("char.diff.so")
 
 
-char.diff.C <- function (x, method = "Gower")  {
+char.diff <- function (x, method = "Gower")  {
 
-    ## Convert character
-    convert.character <- function(X) {
-        if(class(X) == "numeric") {
-            X <- LETTERS[X+1]
-        } else {
-            X <- as.factor(X)
-            levels(X) <- 1:length(levels(X))
-            X <- as.numeric(X)
-        }
-        return(X)
-    }
-
-    ## Transform states into similar values
-    normalise.character <- function(X) {
-        ## Get the present states
-        states <- as.numeric(levels(as.factor(X)))
-        ## Get the states of X
-        states_match <- sort(match(states, X))
-
-        ## Replacing the original states
-        for(state in 1:length(states)) {
-            X <- gsub(X[states_match[state]], LETTERS[state], X)
-        }
-
-        X <- convert.character(X)
-        
-        return(X)
+    ## Convert matrix (if not numeric)
+    if(!all(apply(x, 2, class) == "numeric")) {
+        options(warn = -1)
+        x <- apply(x, 2, as.numeric)
+        options(warn = 0)
     }
 
     ## Options to remove:
@@ -80,7 +58,7 @@ char.diff.C <- function (x, method = "Gower")  {
 
 
 
-char.diff <- function(X,Y){ 
+char.diff_R <- function(X,Y){ 
 
     # Convert character
     convert.character <- function(X) {

### 2016-11-12 Analysis/Comparing_characters.Rmd
Comparing characters

---
title: "Comparing characters"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 12
---

Before starting
===============

Loading the package
-----------------------------------

```{r, eval = FALSE}
# The latest version of devtools
if(!require(devtools)) install.packages("devtools")

## The latest version of dispRity 
install_github("TGuillerme/dispRity")
```

Comparing characters?
===============

The first step is to look at how to compare morphological characters.
How can we efficiently measure the correlation between two characters?
Of course, in some examples, it is pretty easy: a character `A {0,1,1,1,1}` is completely correlated to a character `B {0,1,1,1,1}`.
Moreover, `A {0,1,1,1,1}` must also be completely correlated with `C {1,0,0,0,0}`.
But for two other characters `D {0,0,1,1,1}` and `E {0,1,0,1,1}`, which one is more correlated to `A` (if any)?

Here we are going to explore the inverse absolute Spearman correlation (ranging from 0 - full correlation - to 1 - no correlation) as follows:

```{r}
cor.spearman <- function(X, Y) {
    return(1-round(abs(cor(X,Y, method = "spearman")), digit = 10))
}
## Note that to avoid rounding issues, we will round the correlation value up to the last 10 digits
```

We will also introduce a new difference simply called "Character difference" that counts the difference in states pattern (ranging from 0 - no difference - to 1 - complete difference) as follows:

![equation](http://www.sciweavers.org/tex2img.php?eq=1%20-%20%5Cleft%28%20%5Cfrac%7B%5Cmid%20%7B%5Cfrac%7B%5Cabs%7B%5Csum_%7Bi%7D%5E%7Bn%7D%28%5Cmid%20x_%7Bi%7D%20-%20y_%7Bi%7D%5Cmid%20%29%7D%7D%20%7Bn%7D-%5Cfrac%7B1%7D%7B2%7D%7D%20%5Cmid%20%20%7D%7B%5Cfrac%7B1%7D%7B2%7D%7D%5Cright%29&bc=White&fc=Black&im=jpg&fs=12&ff=arev&edit=0)
<!-- \begin{equation}
    1 - \left( \frac{\mid {\frac{\abs{\sum_{i}^{n}(\mid x_{i} - y_{i}\mid )}} {n}-\frac{1}{2}} \mid  }{\frac{1}{2}}\right)
\end{equation} -->

Where *n* is the number of taxa with comparable data and *x_i* and *y_i* are each characters states for the characters *x* and *y* and the taxon *i*.

Note that in practice, the characters are first normalised and the Gower distance is then calculated.
The normalisation works by changing the characters tokens labels to always start from `0` being the first state.
For example, a character `{8,8,2,1,3,0}` will be normalised by changing all the occurences of the first state (`8`) by `0`, then all the occurences of the second one (`2`) by `1` and so on resulting in the character being normalised as `{0,0,1,2,3,4}`.
Note that from a phylogenetic point of view, if the character is Fitch-like, both characters convey the exact same information.

```{r}
source("../Functions/char.diff.R")
dyn.load("../Functions/char.diff.so")
```

A simple case
----------------
Let's create a bunch of binary characters first

```{r}
A <- c(1,0,0,0,0)
B <- c(0,1,1,1,1)
C <- c(1,1,0,0,0)
D <- c(0,1,1,0,0)
E <- c(1,0,0,1,1)
```

Straight away, we expect `A` and `B` to be totally correlated as well as `D` and `E`.
Also, will will check if our difference metric is triangular (i.e. if `A` = `B` then the difference `(A,X)` must be equal to `(B,X)`).

```{r}
## As expected:
cor.spearman(A,B)
cor.spearman(D,E)
## Note that the order doesn't matters
cor.spearman(B,A)
cor.spearman(E,D)
## And concerning the triangularity
cor.spearman(A,C) == cor.spearman(B,C)
cor.spearman(D,C) == cor.spearman(E,C)
```

This works as well for our character difference metric

```{r}
char.diff(matrix(c(A,B), nrow = 5))[2,1]
char.diff(matrix(c(D,E), nrow = 5))[2,1]
char.diff(matrix(c(B,A), nrow = 5))[2,1]
char.diff(matrix(c(E,D), nrow = 5))[2,1]
char.diff(matrix(c(A,C), nrow = 5))[2,1] ==
char.diff(matrix(c(B,C), nrow = 5))[2,1]
char.diff(matrix(c(D,C), nrow = 5))[2,1] ==
char.diff(matrix(c(E,C), nrow = 5))[2,1]
```

Let's have a look at the differences between `A` and the other characters:

```{r, fig.width=6, fig.height=6}
## Calculating the differences/correlations
character_correlations <- unlist(lapply(list(B,C,D,E), cor.spearman, A))
character_differences <- char.diff(matrix(c(A,B,C,D,E), nrow = 5))[1,-1]

## Plotting the results
plot(character_differences, type = "l", col = "blue", ylim = c(0,1), xaxt = "n", xlab = "",
    ylab = "", main = "Characters comparisons (binary)")
axis(1, at = 1:4, labels = c("A:B", "A:C", "A:D", "A:E"))
lines(character_correlations, col = "orange")
legend("topleft", legend = c("Difference", "Correlation"), lty = 1, xjust = 1, yjust = 1,
    col = c("blue", "orange"))
```

Although both metrics show different results, the magnitudes of changes seems consistent between them.
In other words, the "Character difference" seems to pick up slightly higher differences than the correlation one but overall the same trend is observed (increase differences and then no differences).

A non-binary case
----------------
However, in reality, some characters might have more than two state.

```{r}
A <- c(1,2,0,0,0)
B <- c(2,3,4,4,4)
C <- c(1,2,1,0,0)
D <- c(2,0,2,1,1)
E <- c(1,2,2,0,0)
F <- c(0,1,2,0,0)
G <- c(0,2,1,0,0)
```

In this case again, we expect `A` and

Straight away, we expect `A` and `B` to be totally correlated/not different as well as well as `C` and `D`.
However, the spearman correlation does pick up some difference (i.e. not totally correlated)

```{r}
cor.spearman(A,B)
cor.spearman(C,D)
## And triangularity is not conserved!
cor.spearman(C,F) == cor.spearman(D,F)
```
Thankfully though, the "Character Difference" metric has no problems!

```{r}
char.diff(matrix(c(A,B), nrow = 5))[2,1]
char.diff(matrix(c(C,D), nrow = 5))[2,1]
char.diff(matrix(c(A,F), nrow = 5))[2,1] ==
char.diff(matrix(c(B,F), nrow = 5))[2,1]
char.diff(matrix(c(C,F), nrow = 5))[2,1] ==
char.diff(matrix(c(D,F), nrow = 5))[2,1]
```

Let's have a look at the differences between `A` and the other characters:

```{r, fig.width=6, fig.height=6}
## Calculating the differences/correlations
character_correlations <- unlist(lapply(list(B,C,D,E,F,G), cor.spearman, A))
character_differences <- char.diff(matrix(c(A,B,C,D,E,F,G), nrow = 5))[1,-1]

## Plotting the results
plot(character_differences, type = "l", col = "blue", ylim = c(0,1), xaxt = "n", xlab = "",
    ylab = "", main = "Characters comparisons (non-binary)")
axis(1, at = 1:6, labels = c("A:B", "A:C", "A:D", "A:E", "A:F", "A:G"))
lines(character_correlations, col = "orange")
legend("topleft", legend = c("Difference", "Correlation"), lty = 1, xjust = 1, yjust = 1,
    col = c("blue", "orange"))
```

The correlation metric does not match the actual differences between characters anymore!

Some more complex example
===============

Making up some data
-----------------------------------
First let's make up some discrete morphological matrices using the `dispRity` package.

```{r}
library(dispRity) ; set.seed(123)
# A random tree (coalescent)
tree <- rcoal(15)
# A morphological matrix (Mk model)
matrix <- sim.morpho(tree, 50, states = c(0.85, 0.15), rates = c(rgamma, rate = 10, shape = 5),
    invariant = FALSE)
# Checking matrix
check.morpho(matrix, tree)

## This produces a fairly "realistic" matrix
# Transforming the characters to be numeric
matrix <- apply(matrix, 2, as.numeric)
```

Let's now run the pairwise comparisons for both metrics. We will use the `cor.matrix` function for calculating the spearman correlation (slow).

```{r}
source("../Functions/cor.matrix.R")
pairwise_matrix_correlation <- cor.matrix(matrix, cor.spearman)
pairwise_matrix_differences <- char.diff(matrix)
```

We can now visualise a same single pairwise comparison (character `1` *versus* the others):

```{r, fig.width=6, fig.height=6}
source("../Functions/plot.cor.matrix.R")
## Extracting the first character differences
character_differences <- pairwise_matrix_differences[,1]
character_correlation <- pairwise_matrix_correlation[,1]

## Sorting the values following the characters differences
sorting <- sort(character_differences, index.return = TRUE)[[2]]
character_differences <- character_differences[sorting]
character_correlation <- character_correlation[sorting]

## Plotting the results
plot(character_differences, type = "l", col = "blue", ylim = c(0,1), xlab = "", ylab = "",
    main = "Characters comparisons (non-binary)")
lines(character_correlation, col = "orange")
legend("topleft", legend = c("Difference", "Correlation"), lty = 1, xjust = 1, yjust = 1,
    col = c("blue", "orange"))
```
Note how the correlations quickly seem to plateau around 0.6-0.8 with a decrease in variance as the character difference increase (see below for replicated simulations).

We can also visualise the whole pairwise character differences/correlation distributions:

```{r, fig.width = 12, fig.height = 12}
op <- par(mfrow = c(2,2), bty = "n")
plot.cor.matrix(pairwise_matrix_differences, main = "Pairwise Characters Differences")
plot.cor.matrix(pairwise_matrix_correlation, main = "Pairwise Correlations score")
hist(na.omit(as.vector(pairwise_matrix_differences)), breaks = 15,
    main = "Differences distribution", xlab = "Character differences")
hist(na.omit(as.vector(pairwise_matrix_correlation)), breaks = 15,
    main = "Correlations distribution", xlab = "Character correlations")
par(op)
```

Two interesting points to note for the character differences metric:
 * The metric is not continuous: some values are never achieved!
 * The distribution looks like "half" of a normal distribution. This is easily explainable since we consider "mirror" character (e.g. `A {0,1,1}` and `B {1,0,0}`) to be identical and thus only one side of the distribution is explored. 
We can also have a look at the difference between both methods.
Ideally this difference (the residuals) should be normally distributed.

```{r, fig.width = 6, fig.height = 6}
## Calculating the differences
pairwise_differences <- pairwise_matrix_differences-pairwise_matrix_correlation
## Plotting the differences
hist(na.omit(as.vector(pairwise_differences)), breaks = 15,
    main = "Methods Differences distribution", xlab = "Residuals")

## Are the residuals normal?
shapiro.test(na.omit(pairwise_differences))
```

Appendix
===============

Some more simulations
-----------------------------------
Because here we are affected by the random seed and the size of the matrix, here is a longer simulation running past more matrices:

For speed improvement, we will "functionise" the different procedures (i.e. making matrices, etc...).

```{r}
# Function for looping the first step (matrix generation)
make.a.matrix <- function(taxa = 80, characters = 400) {
    ## Generate a tree
    tree <- rcoal(taxa)
    ## Generate a matrix
    matrix <- sim.morpho(tree, characters, states = c(0.85, 0.15),
        rates = c(rgamma, rate = 10, shape = 5), invariant = FALSE, verbose = FALSE)
    # Make the matrix output numeric
    matrix <- apply(matrix, 2, as.numeric)

    return(matrix)
}
```

```{r}
# Number of replicates:
replicates = 100
# Generate the matrices
matrices <- replicate(replicates, make.a.matrix(), simplify = FALSE)

# Get the character differences
character_differences <- lapply(matrices, char.diff)
character_correlation <- lapply(matrices, cor.matrix, cor.spearman)

# Calculating the residuals
characters_differences_distribution <- NULL
characters_correlation_distribution <- NULL
for (replicate in 1:replicates) {
    characters_differences_distribution <- c(characters_differences_distribution,
        as.vector(na.omit(character_differences[[replicate]])))
    characters_correlation_distribution <- c(characters_correlation_distribution,
        as.vector(na.omit(character_correlation[[replicate]])))
}
residuals <- characters_differences_distribution - characters_correlation_distribution

# Sorting the first characters
first_character_differences <- matrix(ncol = replicates, nrow = nrow(character_differences[[1]]))
first_character_correlation <- first_character_differences
for (replicate in 1:replicates) {
    sorting <- sort(character_differences[[replicate]][,1], index.return = TRUE)[[2]]
    first_character_differences[,replicate] <- character_differences[[replicate]][,1][sorting]
    first_character_correlation[,replicate] <- character_correlation[[replicate]][,1][sorting]
}

# Calculating the medians of the first characters differences
median_first_character_differences <- apply(first_character_differences, 1, median)
median_first_character_correlation <- apply(first_character_correlation, 1, median)
```

Now let's plot all these results:

```{r, fig.width = 12, fig.height = 12}
op <- par(mfrow = c(2,2), bty = "n")
hist(characters_differences_distribution, breaks = 40, main = "Characters Difference distribution",
    xlab = "Characters Differences")
hist(characters_correlation_distribution, breaks = 40, main = "Characters Correlation distribution",
    xlab = "Characters Correlation")
plot(density(residuals), main = "Difference between both methods",
    xlab = "Residuals")
par(op)
```

And finally the example for the first character again:

```{r, fig.width = 10, fig.height = 10}
plot(first_character_differences[,1], type = "l", col = "lightblue", ylim = c(0,1), xlab = "",
    ylab = "", main = "Characters comparison (matrix 80x400)")
lines(first_character_correlation[,1], col = "yellow", lty = 3)
for(replicate in 2:replicates) {
    lines(first_character_differences[,replicate], col = "lightblue")
    lines(first_character_correlation[,replicate], col = "yellow", lty = 3)
}
# Adding the medians
lines(median_first_character_differences, col = "blue")
lines(median_first_character_correlation, col = "orange")
legend("topleft", legend = c("Differences", "Correlation"), lty = 1, xjust = 1,
    yjust = 1, col = c("blue", "orange"))
```

Digression on missing data
-----------------------------------

In practice some (all?) matrices contain a number of missing data entries. In the character difference metric, because we use the Gower distance, only comparable sites are taken into account.
For example, the difference between two characters `A = {0,0,?,?,?}` and `B = {0,1,1,?,?}` will be obtained by comparing only `A' = {0,0}` and `B' = {0,1}` (resulting into a difference of 1).
If a third character `C = {?,?,0,0,1}`, the difference between `A` and `C` is typically inapplicable (`NA`).

### 2016-11-14 Functions/modify.matrix.R
Modifying matrix finalised wrapper function

@@ -1,0 +1,71 @@
+#' @title Modify matrix
+#'
+#' @description Modifies a matrix to minimise/maximise differences among characters
+#'
+#' @param matrix A discrete morphological matrix.
+#' @param type Whether to \code{"maximise"} or \code{"minimise"} character differences
+#' @param threshold The threshold value beyond/above which to maximise/minimise.
+#' @param character.differences Optional, a pre-calculated character difference matrix (if missing, the matrix is calculated from \code{matrix}).
+#'
+#' @return
+#' A modified matrix.
+#' 
+#' @examples
+#' ##
+#' 
+#' @author Thomas Guillerme
+#' @export
+#' 
+
+modify.matrix <- function(matrix, type = "maximise", threshold = 0.25, character.differences) {
+    ## Backup the matrix
+    matrix_modified <- matrix
+
+    if(missing(character.differences)) {
+        ## Calculate the character differences
+        character_differences <- char.diff(matrix)
+    } else {
+        character_differences <- character.differences
+    }
+
+    ## Get the worst characters
+    get.worst.characters <- function(matrix, type, threshold, character.differences = character_differences) {
+
+        ## Set up the quantile threshold (i.e. for maximising difference, remove characters that have 5% of their differences lower than the threshold; for minimising, remove characters that have 95% of their differences higher than the threshold).
+        if(type == "maximise") {
+            quantile.threshold = 0.05
+        } else {
+            quantile.threshold = 0.75
+        }
+
+        ## Calculate the quantiles
+        quantiles <- apply(character_differences, 2, quantile, probs = c(quantile.threshold), na.rm = TRUE)
+
+        ## Select the worst characters
+        if(type == "maximise") {
+            worst_characters <- which(quantiles < threshold)
+        } else {
+            worst_characters <- which(quantiles > threshold)
+        }
+    }
+    
+
+    ## Get the worst characters
+    worst_characters <- get.worst.characters(matrix_modified, type, threshold)
+    ## Replace the worst characters (randomly)
+    if(length(worst_characters) !=  ncol(matrix)) {
+        # If any characters are left to be replace:
+        matrix_modified[,worst_characters] <- matrix_modified[,sample(seq(1:ncol(matrix_modified))[-worst_characters], length(worst_characters), replace = TRUE)]
+    } else {
+        # Randomly reshuffle the matrix for consistency:
[…]

### 2016-11-14 Functions/plot.char.diff.density.R
plot character difference profiles function

@@ -1,0 +1,76 @@
+#' @title Plots the character difference density
+#'
+#' @description Plots the character difference density profile
+#'
+#' @param matrix A square correlation matrix.
+#' @param legend The text of the legend. Default is \code{c("Combined", "Individual")}.
+#' @param col Two colors for the different densities: the first color is used for cumulative density and the second for individual characters' ones. Default is \code{col = c("black", "grey")}.
+#' @param legend.pos The position of the legend. Can be two \code{numeric}. Default is \code{"topleft"}.
+#' @param ... Any additional graphical arguments to be passed to \code{plot}. See details.
+#' 
+#' @details
+#' Note that \code{main} will use \code{"Character differences profile"} as default; \code{xlab} uses \code{"Character differences"} and \code{ylab} uses \code{"Density"}.
+#' 
+#' @examples
+#' ##
+#' 
+#' @author Thomas Guillerme
+#' @export
+#' 
+
+
+plot.char.diff.density <- function(matrix, main, legend, col, xlim, ylim, legend.pos, xlab, ylab) {
+
+    ## Functions for getting the the density plot limits
+    get.max.x <- function(density) return(max(density$x))
+    get.max.y <- function(density) return(max(density$y))
+    get.min.x <- function(density) return(min(density$x))
+    get.min.y <- function(density) return(min(density$y))
+
+    ## Default options:
+    if(missing(main)) {
+        main = "Character differences profile"
+    }
+    if(missing(legend)) {
+        legend = c("Combined", "Individual")
+    }
+    if(missing(col)) {
+        col = c("black", "grey")
+    }
+    if(missing(xlab)) {
+        xlab = "Character differences"
+    }
+    if(missing(ylab)) {
+        ylab = "Density"
+    }
+    if(missing(legend.pos)) {
+        legend.pos = "topleft"
+    }
+    ## Calculating the character difference matrix
+    chara_diff <- char.diff(matrix)
+
+    ## Measuring the densities
+    densities <- apply(chara_diff, 2, density, na.rm = TRUE)
+
+    if(missing(xlim)) {
+        xlim = c(min(unlist(lapply(densities, get.min.x))), max(unlist(lapply(densities, get.max.x))))
+    }
+    if(missing(ylim)) {
+        ylim = c(min(unlist(lapply(densities, get.min.y))), max(unlist(lapply(densities, get.max.y))))
+    }
+
[…]

### 2016-11-14 Writing/draft.tex
Updated methods

@@
 \medskip
-%\begin{center}
+\begin{center}
 
-%\noindent{\Large \bf Tree rearrangements rearranged} %TG: I'm bad at puns.
+\noindent{\Large \bf Effect of coding correlated discrete character correlation on tree inference}
 \bigskip
 
-% \noindent {\normalsize \sc Thomas Guillerme$^1$$^*$, and Martin D. Brazeau$^1$}\\ %TG: Author order can be swapped of course! There's only a finite combination of 2 elements anyway!
-% \noindent {\small \it 
-% $^1$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\}
-% \end{center}
-% \medskip
-% \noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\  %TG: Same as above
+\noindent {\normalsize \sc Thomas Guillerme$^1$$^*$, and Martin D. Brazeau$^1$}\\
+\noindent {\small \it 
+$^1$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\}
+\end{center}
+\medskip
+\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\ 
 \vspace{1in}
@@
 
+Discrete characters correlation have been studied in the context of Phylogenetic Comparative Methods to test whether two or more characters could have coevolved (Maddison 1990, Evolution; Pagel 1994, Proc B).
+However, nobody (that we know of) has looked at what is the effect of actually coding characters that coevolved in a phylogenetic matrix.
+
+-Morphological characters are more and more used: phenomics, tip-dating, Total-Evidence.
+
+-More accurate methods (parsimony vs Bayesian) and better models (Mk, Mk variable from Wright).
+
+-But question remains on which characters to code!
+
+-Don't code correlated characters but what is the actual effect of that? And how many characters are ``secretly'' correlated?
+
+\cite{Davalos01072014}: character correlation + signal in bats
+\cite{Grabowski2016}: amount of data needed for testing correlations? CHECK
+\cite{Lande1983}: measure correlation CHECK
+\cite{Pagel1994}: measure correlation
+\cite{Pagel2006}: measure correlation Bayesian
+\cite{Maddison1990}: measure binary state character correlation for testing coevol
+
 \section{Methods}
 
-\subsection{Characters difference (CD)}
+We define characters as being correlated if they give the same phylogenetic information.
 
+\subsection{Measuring correlation between characters}
+
+In order to measure, the difference between two characters, we propose a new distance metric defined as follows:
+
+\subsubsection{Characters difference (CD)}
 \begin{equation}
-    CD = 1 - \left(\frac{\abs*{\frac{\abs*{\sum_{i}^{n}(x_{i} - y_{i})}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
+    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\abs*{\sum_{i}^{n}(x_{i} - y_{i})}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
 \end{equation}
 
-\noindent Where $n$ is the number of taxa and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
+\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$ \citep[i.e.
+the Gower distance;][]{GowerDist}.
+Since we are considering differences as being only Fitch-like, we ignore the actual distance between two different character states tokens: two same character states tokens have a difference of $0$ and two different, a difference of $1$ (e.g.
+$0 - 0 = 0$ or $1 - 8 = 1$).
+In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
+In other words, we replaced all the occurences of the first token to be $0$, the second to be $1$, etc.
+This way, a character \texttt{A = {2,2,3,0,0,3}} would be standardised as \texttt{A' = {0,0,1,2,2,1}}.
+Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
+When the character difference is equal to $0$, it means that characters convey the same phylogenetic signal.
+When the character difference is equal to 1 it means it conveys the most different signal.
+For example with three characters \texttt{A = {0,1,1,1}}, \texttt{B = {1,0,0,0}} and \texttt{C = {0,1,2,3}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
 
-% This operation can be repeated for all the internal edges in the original tree, thus the maximum TBR can be:
-% \begin{equation}
-%     \text{Maximum TBR} = \overbrace{(n-3)}^{\text{internal edges}} \sum_{i=2}^{n-2} (2i-3)(2(n-i)-3) %TG: Check! Summation might be incorrect
-% \end{equation}
-% Where $i$ and $n-i$ corresponds respectively to the size of the smallest and biggest trees resulting from the bisection the original tree.
+This metric respects triangular equality [CHECK THE PROPER NAME] (i.e. $CD_{(A,B)} = CD_{(B,A)}$ and $CD_{(A,C)} = CD_{(B,C)}$).
 
-\section{Acknowledgments}
+\subsection{Simulating discrete morphological matrices}
+
+\subsection{Modifying the matrices}
+
+\subsection{Inferring the phylogenies}
+
+\subsection{Comparing the topologies}
+
+\cite{Guillerme2016146}
+
+\subsection{Empirical morphological matrices}
+
+\section{Acknowledgements}
 European Research Council under the European Union’s Seventh Framework Programme (FP/2007–2013)/ERC Grant Agreement number 311092.
 
-% \bibliographystyle{sysbio}
-% \bibliography{References}
+% Suggested reviewers:
+%- Liliana Davalos
+
+\bibliographystyle{sysbio}
+\bibliography{References}
 

### 2016-11-14 Analysis/Modify_character_difference.Rmd
Character difference modification Rmd

---
title: "Modifying character differences"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 12
---

Before starting
===============

Loading the package
-----------------------------------

```{r, eval = FALSE}
## The latest version of devtools
if(!require(devtools)) install.packages("devtools")

## The latest version of dispRity 
install_github("TGuillerme/dispRity")

## The latest version of Claddis
install_github("graemetlloyd/Claddis")
```

Loading the functions
-----------------------------------

```{r}
library(Claddis) ; library(dispRity)
source("functions.R") ; load.functions(test = FALSE)

## Load the C libraries
dyn.load("../Functions/char.diff.so")

## Function for helping loading the matrices
find.non.numerics <- function(x) {
    options(warn = -1)
    if(is.na(as.numeric(x))) {
        return(TRUE)
    } else {
        return(FALSE)
    }
}

## Functions for helping the density plots
get.max.x <- function(density) return(max(density$x))
get.max.y <- function(density) return(max(density$y))
get.min.x <- function(density) return(min(density$x))
get.min.y <- function(density) return(min(density$y))
```
And loading one matrix from TreeBase.

```{r}
## Set up the matrices path
matrix_path = "../Data/100_Matrices"
files <- list.files(path = matrix_path, pattern = ".nex")

## Read one matrix
matrix <- ReadMorphNexus(paste(matrix_path, files[39], sep ="/"))$matrix
matrix[apply(matrix, c(1,2), find.non.numerics)] <- "?"

```

Modifying character differences in a matrix
===============

Maximising and minimising the differences
-----------------------------------

The maximisation or minimisation of the character differences works in a fairly straightforward way:

1. the quantiles for each character compared to all other characters is calculated (e.g. first character against all the other)
2. characters that have their last/first quantile lower/higher than the threshold (0.25 and 0.75) are then selected
3. the selected characters are randomly replaced by the characters left in the matrix

```{r}
## Maximising the character differences
matrix_max <- modify.matrix(matrix, "maximise", threshold = 0.25)
## Minimising the character differences
matrix_min <- modify.matrix(matrix, "minimise", threshold = 0.75)
```

Plotting the differences
-----------------------------------

Let's first plot the normal density profile of the matrix' character differences:

```{r, fig.width=6, fig.height=18, fig.align = 'center'}
## Graphical parameters
op <- par(mfrow = c(3,1), bty = "n")

## Plotting the density profile un-modified
plot.char.diff.density(matrix, main = "Original differences")

## Plotting the maximised matrix' density
plot.char.diff.density(matrix_max, main = "Maximised differences")

## Plotting the minimised matrix' density
plot.char.diff.density(matrix_min, main = "Minimised differences")
```

### 2016-11-15 Writing/draft.tex
Added references and updated structure

@@
 
-Discrete characters correlation have been studied in the context of Phylogenetic Comparative Methods to test whether two or more characters could have coevolved (Maddison 1990, Evolution; Pagel 1994, Proc B).
-However, nobody (that we know of) has looked at what is the effect of actually coding characters that coevolved in a phylogenetic matrix.
+The last five years have seen a drastic expansion (resurgence?) of the use of discrete morphological data in macroevolutionary studies.
+In fact, traditionally confined to classic cladistics \citep{felsenstein2004inferring} these methods have recently been highly improved whether based on the models and methods for inferring phylogenies (\citealt{lewisa2001} + BAPSTinReview; with some debate on which method to use: \citealt{spencerefficacy2013,wrightbayesian2014,OReilly20160081}) or the methods to date them \citep[e.g.][]{bapsta2013} or to combine them with molecular data \citep{pyrondivergence2011,ronquista2012}.
+In parallel, much development has also been done on pre-phylogenetic analysis \citep[e.g.
+data collection;][]{morphobank} as well as on post-phylogenetic analysis \citep[e.g.
+morphological disparity analysis;][]{Close2015,Claddis}.
 
--Morphological characters are more and more used: phenomics, tip-dating, Total-Evidence.
+However, using discrete morphological data comes with several caveats, like the effect of missing data \citep{Guillerme2016146} or of the model \citep{OReilly20160081}.
+One specific assumption is that coded characters should not be correlated \citep{Brazeau2011}[FIND BETTER CITE?].
 
--More accurate methods (parsimony vs Bayesian) and better models (Mk, Mk variable from Wright).
+Character correlation can be due to
+- co-evolution of two traits (e.g.
+the character ``number of digits on a limb'' co-evolves with the character ``presence/absence of limb'' - this correlation can be sometimes fairly obvious, sometimes less)
+- modularity of morphological features [CITE GOSWAMI!]
+- convergent evolution (e.g. the character ``flight'') [???]
+- error from the coder (e.g.
+a coder can unintentionally code the same feature multiple times, thus effectively waiting it, for example: ``presence/absence of a proximal end on a limb'' and ``presence/absence of a distal end on a limb'' - effectively weighting the character ``presence/absence of a limb'' by two).
+%-Don't code correlated characters but what is the actual effect of that? And how many characters are ``secretly'' correlated?
 
--But question remains on which characters to code!
+A lot of work has been done on character correlation \textit{per se}, but focussing mainly on how to measure correlation between characters to detect patterns of coevolution \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
+However, only several attempts have been made on the actual effect of characters correlation on phylogenetic inference with two notable expections of \cite{Davalos01072014} on empirical data and \cite{ZouConvergence} on molecular data.
+It is thus still unclear what will be the effect of characters correlation on phylogenetic inference.
 
--Don't code correlated characters but what is the actual effect of that? And how many characters are ``secretly'' correlated?
+% \cite{Davalos01072014}: character correlation + signal in bats
+% \cite{Grabowski2016}: amount of data needed for testing correlations? CHECK
+% \cite{Lande1983}: measure correlation CHECK
+% \cite{Pagel1994}: measure correlation
+% \cite{Pagel2006}: measure correlation Bayesian
+% \cite{Maddison1990}: measure binary state character correlation for testing coevol
+% \cite{ZouConvergence}: convergence measured as characters correlation (Removing convergence prone characters improves accuracy)
 
-\cite{Davalos01072014}: character correlation + signal in bats
-\cite{Grabowski2016}: amount of data needed for testing correlations? CHECK
-\cite{Lande1983}: measure correlation CHECK
-\cite{Pagel1994}: measure correlation
-\cite{Pagel2006}: measure correlation Bayesian
-\cite{Maddison1990}: measure binary state character correlation for testing coevol
+In this study, we propose a new way to measure character correlation (more adapted to discrete morphological characters) and test on simulated and empirical data what is the effect of character correlation on topology.
+%TG: check if we do just topology?
+In other words, we generate matrices and maximise or minimise character correlation and see how that affects topology.
 
@@
 
-We define characters as being correlated if they give the same phylogenetic information.
-
 \subsection{Measuring correlation between characters}
 
+We define characters as being correlated if they give the same phylogenetic information.
 In order to measure, the difference between two characters, we propose a new distance metric defined as follows:
@@
 
+We simulated discrete morphological matrices based on the \texttt{dispRity R} package \citep[][; \url{https://github.com/TGuillerme/dispRity}]{thomas_guillerme_2016_55646}.
+
 \subsection{Modifying the matrices}
@@
 
+We applied the same procedure described above to 100 empirical matrices downloaded from TreeBASE (\url{http://treebase.org/}).
+The 100 matrices where selected to be the most representative of empirical work with discrete morphological data: with randomly selected matrices with more than 100 each, published between 1985 and 2013 and covering 19 taxonomic classes within 6 phyla (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
+The full list of matrices is available in the supplementary materials @@@.
+
 \section{Acknowledgements}
 European Research Council under the European Union’s Seventh Framework Programme (FP/2007–2013)/ERC Grant Agreement number 311092.
+
+\section{Results}
+
+\section{Discussion}
+
+\subsection{Conclusion}
 

### 2016-12-14 Writing/draft.tex
Updated methods

@@
 
-We simulated discrete morphological matrices based on the \texttt{dispRity R} package \citep[][; \url{https://github.com/TGuillerme/dispRity}]{thomas_guillerme_2016_55646}.
+To simulate the matrices we applied a protocol similar to \citep{Guillerme2016146}.
+First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform ($0$,$1$) distribution maintaining $\lambda$ $>$ $\mu$ using the diversitree \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
+For each tree, we arbitrarily set the outgroup to be the first taxa (alphabetically) thus effectively rooting the trees on this taxa.
+These trees are hereafter called the ``true'' trees.
+We then simulated the characters using the either of the two following models:
+\begin{itemize}
+    \item The HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution an scale to sum to one) and using a transition/transvertion rate of two \citep{douadycomparison2003} but where the purines (AG) were changed into state $0$ and the pyrimidines (CT) in state $1$.
+    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][]{OReilly20160081} but can only generate binary state characters.
+    \item To generate more that two states characters, we used the M$k$ model \citep{lewisa2001}.
+    We draw the number of character states with a probability of $0.85$ for two-state characters (binary) and $0.15$ for three state characters \citep{Guillerme2016146}.
+    Assuming the same transition rate between character states might seem unrealistic and ignoring characters following, for example, a Dollo pattern, \cite{Wright01072016} have shown that an equal rate transition is still strictly majority in empirical data.
+\end{itemize}
+
+For each character, both models (HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $3$).
+We resimulated every invariant characters to obtain a matrix with no invariant characters.
+Choosing these low evolutionary rate values allowed us to obtain discrete characters matrices with realistic values of Consistency Indices that where superior that $0.26$ \citep{OReilly20160081}.
+For each trees with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
+The matrices where generated using the \texttt{dispRity R} package \citep[][; \url{https://github.com/TGuillerme/dispRity}]{thomas_guillerme_2016_55646}.
+To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 50 times resulting in 450 morphological matrices (hereafter called the ``normal'' matrices).
 
@@
 
+We then calculated the pairwise character differences for each generated matrices. 
+We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
+For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e.
+with an average character difference $<$$0.25$) and replaced them randomly by any of the remaining characters.
+This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
+Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e.
+with an average character difference $<$$0.75$) and randomly replaced them with the remaining ones.
+Finally, because this operation effectively changes the weight of characters (i.e.
+giving the characters $<$$0.25$ or $>$$0.75$ a weight of $0$ and giving the randomly selected characters to replace them a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a null expectation modification (i.e.
+randomly weighting characters).
+This step resulted in a total of 1800 matrices (hereafter called the ``maximised'', ``minimised'' and ``random'' matrices).
+The algorithms for the three modifications is available on GitHub@@@. %TG: "@@@" is my "to replace" sign.
+
 \subsection{Inferring the phylogenies}
+
+We inferred the topologies in both Bayesian and Parsimony using MrBayes and PAUP* respectively.
+Bayesian settings.
+PAUP ``settings''.
 
@@
 
-\cite{Guillerme2016146}
+We compared the topologies using the same approach as in \cite{Guillerme2016146}.
 
@@
 
-We applied the same procedure described above to 100 empirical matrices downloaded from TreeBASE (\url{http://treebase.org/}).
+We applied the same procedure described above (without the morphological matrices simulation) to 100 empirical matrices downloaded from TreeBASE (\url{http://treebase.org/}).
 The 100 matrices where selected to be the most representative of empirical work with discrete morphological data: with randomly selected matrices with more than 100 each, published between 1985 and 2013 and covering 19 taxonomic classes within 6 phyla (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).

### 2016-12-15 Writing/draft.tex
Updates on the draft

@@
 
-\noindent{\Large \bf Effect of coding correlated discrete character correlation on tree inference}
+\noindent{\Large \bf Effect of discrete character correlation on tree inference}
 \bigskip
@@
 For each character, both models (HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $3$).
+This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices.
 We resimulated every invariant characters to obtain a matrix with no invariant characters.
-Choosing these low evolutionary rate values allowed us to obtain discrete characters matrices with realistic values of Consistency Indices that where superior that $0.26$ \citep{OReilly20160081}.
+To ensure the that our simulation where reflecting the same parameters as real data, we only selected matrices with Consistency Indices superior to $0.26$ \citep{OReilly20160081}.
+
 For each trees with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
@@
 We inferred the topologies in both Bayesian and Parsimony using MrBayes and PAUP* respectively.
+
 Bayesian settings.
+
 PAUP ``settings''.

### 2016-12-20 Analysis/Character_difference/Character_diff_empirical.Rmd
Cleaned Analysis folder

--- 2016-11-09 Analysis/Character_diff_empirical.Rmd
@@
 library(Claddis) ; library(dispRity)
-source("functions.R") ; load.functions(test = FALSE)
+source("../../Functions/cor.matrix.R")
+source("../../Functions/modify.matrix.R")
+source("../../Functions/plot.char.diff.density.R")
+source("../../Functions/plot.cor.matrix.R")
+source("../../Functions/write.nexus.std.R")
+source("../../Functions/char.diff.R")
+
+## Load the C libraries
+dyn.load("../../Functions/char.diff.so")
 
@@
 ## Set up the matrices path
-matrix_path = "../Data/100_Matrices"
+matrix_path = "../../Data/100_Matrices"
 files <- list.files(path = matrix_path, pattern = ".nex")
@@
 ## Calculate the differences
-differences <- lapply(matrices, cor.matrix, char.diff)
+differences <- lapply(matrices, char.diff)
 

### 2016-12-20 Analysis/Character_difference/Comparing_characters.Rmd
Cleaned Analysis folder

--- 2016-11-12 Analysis/Comparing_characters.Rmd
@@
 }
-## Note that to avoid rounding issues, we will round the correlation value up to the last 10 digits
 ```
+> Note that to avoid rounding issues, we will round the correlation value up to the last 10 digits
 
-We will also introduce a new difference simply called "Character difference" that counts the difference in states pattern (ranging from 0 - no difference - to 1 - complete difference) as follows:
+We will also introduce a new difference called "Character Difference" that counts the difference in states pattern (ranging from 0 - no difference - to 1 - complete difference) as follows:
 
@@
 
-Note that in practice, the characters are first normalised and the Gower distance is then calculated.
-The normalisation works by changing the characters tokens labels to always start from `0` being the first state.
-For example, a character `{8,8,2,1,3,0}` will be normalised by changing all the occurences of the first state (`8`) by `0`, then all the occurences of the second one (`2`) by `1` and so on resulting in the character being normalised as `{0,0,1,2,3,4}`.
+Note that in practice, the characters are first normalised and the Gower distance is then calculated:
+ * The normalisation works by changing the characters tokens labels to always start from `0` being the first state.
+For example, a character `{8,8,2,1,3,0}` will be normalised by changing all the occurrences of the first state (`8`) by `0`, then all the occurrences of the second one (`2`) by `1` and so on resulting in the character being normalised as `{0,0,1,2,3,4}`.
 Note that from a phylogenetic point of view, if the character is Fitch-like, both characters convey the exact same information.
+ * The Gower distance is simply the distance divided by the number of observation (i.e. to take into account the inapplicable or missing data).
+
+The function to calculate the character difference is a modification of the `stats::dist` function's source code (in `C`):
 
 ```{r}
-source("../Functions/char.diff.R")
-dyn.load("../Functions/char.diff.so")
+source("../../Functions/char.diff.R")
+dyn.load("../../Functions/char.diff.so")
 ```
@@
 ----------------
-Let's create a bunch of binary characters first
+Let's create a bunch of binary characters first:
 
@@
 Straight away, we expect `A` and `B` to be totally correlated as well as `D` and `E`.
-Also, will will check if our difference metric is triangular (i.e. if `A` = `B` then the difference `(A,X)` must be equal to `(B,X)`).
+Also, we will check if our difference metric is triangular (i.e. if `A` = `B` then the difference `(A,X)` must be equal to `(B,X)`).
 
@@
 ----------------
-However, in reality, some characters might have more than two state.
+However, in reality, some characters might have more than two state:
 
@@
 
-In this case again, we expect `A` and
-
-Straight away, we expect `A` and `B` to be totally correlated/not different as well as well as `C` and `D`.
-However, the spearman correlation does pick up some difference (i.e. not totally correlated)
+Similarly, we expect `A` and `B` to be totally totally correlated as well as `C` and `D`.
+However, the Spearman correlation does pick up some difference (i.e. not totally correlated)
 
@@
 ```{r}
-source("../Functions/cor.matrix.R")
+source("../../Functions/cor.matrix.R")
 pairwise_matrix_correlation <- cor.matrix(matrix, cor.spearman)
@@
 ```{r, fig.width=6, fig.height=6}
-source("../Functions/plot.cor.matrix.R")
+source("../../Functions/plot.cor.matrix.R")
 ## Extracting the first character differences

### 2016-12-20 Analysis/Character_difference/Modifying_matrices.Rmd
Cleaned Analysis folder

---
title: "Generating and modifying matrices"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 12
---

#### Loading the package and functions

```{r, eval = FALSE}
## The latest version of devtools
if(!require(devtools)) install.packages("devtools")

## The latest version of dispRity 
install_github("TGuillerme/dispRity")

## The latest version of Claddis
install_github("graemetlloyd/Claddis")

## The cran version of diversitree
if(!require(diversitree)) install.packages("diversitree")
```

```{r, results="hide", message=FALSE}
library(Claddis) ; library(dispRity) ; library(diversitree)
source("functions.R") ; load.functions(test = FALSE)

## Load the C libraries
dyn.load("../Functions/char.diff.so")

## Seed for repeatability
set.seed(0)
```

Generating the matrix
===============

Before anything, let's set our simulation variables: the number of taxa and the number of characters:

```{r}
## Number of taxa
ntaxa <- 25

## Number of characters
ncharacters <- 100
```

Generating the tree
-----------------------------------

First, we need to generate a realistic birth death tree upon which to simulate our characters.
For that we will need a function for randomly sampling our birth/death rates from a uniform (0,1) distribution while making sure that the death rate is lower that the birth rate:

```{r}
sample.birth.death <- function() {
    birth <- runif(min = 0, max = 1, 1)
    death <- runif(min = 0, max = birth, 1)
    return(c(birth, death))
}
```

Then let's generate the tree.
We add a while loop to the algorithm to make sure that the tree actually generate properly.
For example, if the birth rate sampled is 0, no tree will be generated!

```{r}
tree <- NULL
## The birth-death parameters
(birth_death <- sample.birth.death())
## Generate the tree
while(is.null(tree)) {
    tree <- tree.bd(birth_death, max.taxa = ntaxa)
}
```

We can then remove the tip labels (that conflict somehow with `phyclust` functions used downstream) and arbitrarily set the first species as the outgroup (thus rooting the tree).

```{r}
## Remove the $node labels (bugging for some reason)
tree$node.label <- NULL

## Make the first taxa (arbitrarily) the outgroup
tree$tip.label[[1]] <- "outgroup"
tree <- root(tree, "outgroup", resolve.root = TRUE)
```

And this is what our tree looks like

```{r, fig.width=6, fig.height=6}
plot(tree, cex = 0.8)

```
<!-- ```{r}
## Save the tree
simulationID = "0001"
write.nexus(tree, file = paste(ntaxa,"t_", ncharacters, "c_", simulationID, "_truetree.nex", sep = ""))
``` -->

Generating the matrix
-----------------------------------

We can now generate the matrix by using a `"mixed"` model (using both HKY-binary and Mk) with a state distribution of 0.85 binary characters and 0.15 three states characters and an evolutionary rate sampled from a gamma distribution with alpha = 5 and beta = 10.
We will also make sure there are no invariant characters and that the consistency index of the matrix is higher than 0.26.

```{r}
## Simulate a matrix
matrix_norm <- sim.morpho(tree, ncharacters, states = c(0.85,0.15),
    model = "mixed", rates = c(rgamma, rate = 100, shape = 3),
    invariant = FALSE)

## Resimulate it if the consistency index is too low
while(check.morpho(matrix_norm)[2,] < 0.26) {
    matrix_norm <- sim.morpho(tree, ncharacters, states = c(0.85,0.15),
        model = "mixed", rates = c(rgamma, rate = 100, shape = 3),
        invariant = FALSE)
}
```

We can also check how realistic our simulated matrix is:

```{r}
check.morpho(matrix_norm, tree)
```

Modifying the matrix
-----------------------------------

We can now maximise, minimise and randomise the character differences from the original matrix (`matrix_norm`).

```{r}
## Maximising the character difference
matrix_maxi <- modify.matrix(matrix_norm, type = "maximise", threshold = 0.25)
## Minimising the character difference
matrix_mini <- modify.matrix(matrix_norm, type = "minimise", threshold = 0.75)
## Randomising the character difference
matrix_rand <- modify.matrix(matrix_norm, type = "randomise")
```

We can then visualise the modifications by looking at the pairwise character differences:

```{r, fig.width=12, fig.height=6}
par(mfrow = c(1, 2))
## The character difference density (per character)
plot.char.diff.density(char.diff(matrix_norm), "Normal matrix")
## The character difference matrix (pairwise)
plot.cor.matrix(char.diff(matrix_norm), main = "")
```

```{r, fig.width=12, fig.height=6}
par(mfrow = c(1, 2))
## The character difference density (per character)
plot.char.diff.density(char.diff(matrix_maxi), "Maximised matrix")
## The character difference matrix (pairwise)
plot.cor.matrix(char.diff(matrix_maxi), main = "")
```

```{r, fig.width=12, fig.height=6}
par(mfrow = c(1, 2))
## The character difference density (per character)
plot.char.diff.density(char.diff(matrix_mini), "Minimised matrix")
## The character difference matrix (pairwise)
plot.cor.matrix(char.diff(matrix_mini), main = "")
```

```{r, fig.width=12, fig.height=6}
par(mfrow = c(1, 2))
## The character difference density (per character)
plot.char.diff.density(char.diff(matrix_rand), "Randomised matrix")
## The character difference matrix (pairwise)
plot.cor.matrix(char.diff(matrix_rand), main = "")
```

### 2016-12-20 Analysis/Gamma_parameter_testing/Gamma_parameter_testing.R

@@ -2,8 +2,13 @@
 library(Claddis)
 library(dispRity)
 library(diversitree)
-source("functions.R") ; load.functions(test = FALSE)
-dyn.load("../Functions/char.diff.so")
+source("../../Functions/cor.matrix.R")
+source("../../Functions/modify.matrix.R")
+source("../../Functions/plot.char.diff.density.R")
+source("../../Functions/plot.cor.matrix.R")
+source("../../Functions/write.nexus.std.R")
+source("../../Functions/char.diff.R")
+dyn.load("../../Functions/char.diff.so")
 write.nexus.std <- ape::write.nexus.data
 body(write.nexus.std)[[2]] <- substitute(format <- match.arg(toupper(format), c("DNA", "PROTEIN", "STANDARD")))
 body(write.nexus.std)[[26]][[3]][[4]] <- substitute(fcat(indent, "FORMAT", " ", DATATYPE, " ", MISSING, " ", GAP, 
@@ -46,7 +51,7 @@ get.CI <- function(X) {
 
 ## Testing parameter (CI distribution)
 gamma <- list(c(rgamma, rate = 1, shape = 1), c(rgamma, rate = 10, shape = 1), c(rgamma, rate = 100, shape = 1), c(rgamma, rate = 1, shape = 5), c(rgamma, rate = 10, shape = 5), c(rgamma, rate = 100, shape = 5))
-runs <- 50
+runs <- 30
 
 for(param in 1:length(gamma)) {
     t25_c100 <- replicate(runs, morpho.wrapper(25, 100, gamma[[param]]), simplify = FALSE)
@@ -64,11 +69,11 @@ for(param in 1:length(gamma)) {
     hist(unlist(lapply(t25_c100, get.CI)), breaks = 20, main = "25t_100c", xlab = "Consistency Index", xlim = c(0,1)) ; abline(v = 0.26, lty = 3)
     hist(unlist(lapply(t25_c350, get.CI)), breaks = 20, main = "25t_350c", xlab = "Consistency Index", xlim = c(0,1)) ; abline(v = 0.26, lty = 3)
     hist(unlist(lapply(t25_c1000, get.CI)), breaks = 20, main = "25t_1000c", xlab = "Consistency Index", xlim = c(0,1)) ; abline(v = 0.26, lty = 3)
[…]

### 2016-12-20 Analysis/Sandbox.R

@@ -1,199 +1,90 @@
-# Header
+## Packages and functions
 library(Claddis)
 library(dispRity)
+library(diversitree)
 source("functions.R") ; load.functions(test = FALSE)
 dyn.load("../Functions/char.diff.so")
-
-## Modify the write.nexus.data function
 write.nexus.std <- ape::write.nexus.data
 body(write.nexus.std)[[2]] <- substitute(format <- match.arg(toupper(format), c("DNA", "PROTEIN", "STANDARD")))
 body(write.nexus.std)[[26]][[3]][[4]] <- substitute(fcat(indent, "FORMAT", " ", DATATYPE, " ", MISSING, " ", GAP, 
     " ", INTERLEAVE, " symbols=\"0123456789\";\n"))
-
-
-###############
-# Parameters for generating the matrix
-###############
-
-# Variables that can be changed here are:
-# The tree size ?
-# The length of the matrix (50, 100, 200, 1000) ?
-# The models (Mk or HKYbin)
-# The rates (log normal or gamma)
-
-###############
-# Loading the matrices
-###############
-
-matrix_path = "../Data/100_Matrices"
-files <- list.files(path = matrix_path, pattern = ".nex")
-
-## Function for finding non-numeric values
-find.non.numerics <- function(x) {
-    options(warn = -1)
-    if(is.na(as.numeric(x))) {
-        return(TRUE)
+get.digit <- function(column) {
+    if(max(nchar(round(column)), na.rm = TRUE) <= 4) {
+       return(4-max(nchar(round(column)), na.rm = TRUE))
     } else {
-        return(FALSE)
+        return(0)
     }
 }
-
-## Read the matrices
-# matrices <- list()
-# for(matrix in 1:length(files)) {
-#     cat(files[[matrix]],"\n")
-#     matrices[[matrix]] <- ReadMorphNexus(paste(matrix_path, files[matrix], sep ="/"))$matrix
-#     matrices[[matrix]][apply(matrices[[matrix]], c(1,2), find.non.numerics)] <- "?"
-# }
-
-
-###############
-# Modifying the matrices
-###############
-
-
-matrices <- list()
-for(matrix in 1:10) {
-    cat(files[[matrix]],"\n")
-    matrices[[matrix]] <- ReadMorphNexus(paste(matrix_path, files[matrix], sep ="/"))$matrix
[…]

### 2016-12-20 Functions/char.diff.R
Commented out R bit

@@ -57,89 +57,86 @@ char.diff <- function (x, method = "Gower")  {
 }
 
 
+## R version of the code (SLOW!)
+# char.diff_R <- function(X,Y){ 
 
-char.diff_R <- function(X,Y){ 
+#     # Convert character
+#     convert.character <- function(X) {
+#         if(class(X) == "numeric") {
+#             X <- LETTERS[X+1]
+#         } else {
+#             X <- as.factor(X)
+#             levels(X) <- 1:length(levels(X))
+#             X <- as.numeric(X)
+#         }
+#         return(X)
+#     }
 
-    # Convert character
-    convert.character <- function(X) {
-        if(class(X) == "numeric") {
-            X <- LETTERS[X+1]
-        } else {
-            X <- as.factor(X)
-            levels(X) <- 1:length(levels(X))
-            X <- as.numeric(X)
-        }
-        return(X)
-    }
+#     # Transform states into similar values
+#     normalise.character <- function(X, states) {
+#         # Convert X to character
+#         if(class(X) != "numeric") {
+#             X <- convert.character(X)
+#         }
+#         X <- as.character(X)
+#         # Get the states of X
+#         states_match <- sort(match(states, X))
 
-    # Transform states into similar values
-    normalise.character <- function(X, states) {
-        # Convert X to character
-        if(class(X) != "numeric") {
-            X <- convert.character(X)
-        }
-        X <- as.character(X)
-        # Get the states of X
-        states_match <- sort(match(states, X))
+#         # Replacing the original states
+#         for(state in 1:length(states)) {
+#             X <- gsub(X[states_match[state]], LETTERS[state], X)
+#         }
+#         X <- convert.character(X)
+#         return(X)
+#     }
 
-        # Replacing the original states
-        for(state in 1:length(states)) {
-            X <- gsub(X[states_match[state]], LETTERS[state], X)
-        }
-        X <- convert.character(X)
-        return(X)
-    }
+#     # Convert the characters to numeric (if needed)
+#     if(class(X) != "numeric") {
+#         X <- convert.character(X)
+#     }
+#     if(class(Y) != "numeric") {
+#         Y <- convert.character(Y)
+#     }
 
-    # Convert the characters to numeric (if needed)
[…]

### 2016-12-20 Analysis/Matrix_generation_pipeline.R
Updated log file name

@@ -26,7 +26,7 @@ sample.birth.death <- function() {
 
 ## Seed
 set.seed(0)
-system("echo \"\" >> Matrix_gen.log")
+system("echo \"\" > Matrix_generation.log")
 
 ## Variables
 path <- "../Data/Simulations/"
@@ -64,21 +64,21 @@ for(simulation in 1:length(simulationID)) {
 
     ## Verbose
     cat(paste("RUN: ", ntaxa,"t_", ncharacters, "c_", simulationID[simulation], " - ", sep = ""))
-    system(paste("printf \"RUN: ", ntaxa,"t_", ncharacters, "c_", simulationID[simulation], " - \" >> Matrix_gen.log", sep = ""))
+    system(paste("printf \"RUN: ", ntaxa,"t_", ncharacters, "c_", simulationID[simulation], " - \" >> Matrix_generation.log", sep = ""))
 
 
     ## 1 - Generating the tree:
     tree <- NULL
     birth_death <- sample.birth.death()
     cat("Tree:")
-    system("printf \"Tree:\" >> Matrix_gen.log")
+    system("printf \"Tree:\" >> Matrix_generation.log")
     while(is.null(tree)) {
         tree <- tree.bd(birth_death, max.taxa = ntaxa)
         cat(".")
-        system("printf \".\" >> Matrix_gen.log")
+        system("printf \".\" >> Matrix_generation.log")
     }
     cat("Done - ")
-    system("printf \"Done - \" >> Matrix_gen.log")
+    system("printf \"Done - \" >> Matrix_generation.log")
 
     tree$node.label <- NULL
     tree$tip.label[[1]] <- "outgroup"
@@ -87,30 +87,27 @@ for(simulation in 1:length(simulationID)) {
 
     ## 2 - Generating the matrices
     cat("Matrix:")
-    system("printf \"Matrix:\" >> Matrix_gen.log")
-    matrix_norm <- NULL
-    try(matrix_norm <- sim.morpho(tree, ncharacters, states = c(0.85,0.15), model = "mixed", rates = c(rgamma, rate = 100, shape = 3), invariant = FALSE), silent = TRUE)
-    if(is.null(matrix_norm)) {
-        cat("FAILURE.\n")
-        system("printf \"FAILURE.\n\" >> Matrix_gen.log")
-    } else {
+    system("printf \"Matrix:\" >> Matrix_generation.log")
+    matrix_norm <- NULL ; counter <- 0 
[…]

### 2016-12-20 Analysis/Character_difference/Modify_character_difference.Rmd
Modifying vignettes for new compilation

--- 2016-11-14 Analysis/Modify_character_difference.Rmd
@@
 library(Claddis) ; library(dispRity)
-source("functions.R") ; load.functions(test = FALSE)
+source("../../Functions/cor.matrix.R")
+source("../../Functions/modify.matrix.R")
+source("../../Functions/plot.char.diff.density.R")
+source("../../Functions/plot.cor.matrix.R")
+source("../../Functions/write.nexus.std.R")
+source("../../Functions/char.diff.R")
 
 ## Load the C libraries
-dyn.load("../Functions/char.diff.so")
+dyn.load("../../Functions/char.diff.so")
 

### 2016-12-20 Analysis/Character_difference/Modifying_matrices.Rmd
Modifying vignettes for new compilation

@@
 library(Claddis) ; library(dispRity) ; library(diversitree)
-source("functions.R") ; load.functions(test = FALSE)
+source("../../Functions/cor.matrix.R")
+source("../../Functions/modify.matrix.R")
+source("../../Functions/plot.char.diff.density.R")
+source("../../Functions/plot.cor.matrix.R")
+source("../../Functions/write.nexus.std.R")
+source("../../Functions/char.diff.R")
 
 ## Load the C libraries
-dyn.load("../Functions/char.diff.so")
+dyn.load("../../Functions/char.diff.so")
 

### 2016-12-22 Writing/draft.tex
Updated rate parameter in methods

@@
 
-For each character, both models (HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $3$).
-This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices.
+For each character, both models (HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
+This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices (see supplementary materials @@@).
 We resimulated every invariant characters to obtain a matrix with no invariant characters.

### 2017-02-14 Writing/draft.tex
Methods update

@@
 
-However, using discrete morphological data comes with several caveats, like the effect of missing data \citep{Guillerme2016146} or of the model \citep{OReilly20160081}.
+However, using discrete morphological data comes with several caveats, like the effect of missing data \citep{Guillerme2016146} or of the method \citep{OReilly20160081}.
+
+Common wisdom says that characters should not be correlated
+
 One specific assumption is that coded characters should not be correlated \citep{Brazeau2011}[FIND BETTER CITE?].
@@
 We define characters as being correlated if they give the same phylogenetic information.
-In order to measure, the difference between two characters, we propose a new distance metric defined as follows:
+In order to measure the difference between two characters, we propose a new distance metric defined as follows:
 
@@
 the Gower distance;][]{GowerDist}.
-Since we are considering differences as being only Fitch-like, we ignore the actual distance between two different character states tokens: two same character states tokens have a difference of $0$ and two different, a difference of $1$ (e.g.
+Since we are considering differences as being only Fitch-like (i.e.
+non-weighted and non-ordinate), we ignore the actual distance between two different character states tokens: two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g.
 $0 - 0 = 0$ or $1 - 8 = 1$).
 In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
-In other words, we replaced all the occurences of the first token to be $0$, the second to be $1$, etc.
-This way, a character \texttt{A = {2,2,3,0,0,3}} would be standardised as \texttt{A' = {0,0,1,2,2,1}}.
+In other words, we replaced all the occurrences of the first token to be $0$, the second to be $1$, etc.
+This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{0,0,1,2,2,1\}}.
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
 When the character difference is equal to $0$, it means that characters convey the same phylogenetic signal.
-When the character difference is equal to 1 it means it conveys the most different signal.
-For example with three characters \texttt{A = {0,1,1,1}}, \texttt{B = {1,0,0,0}} and \texttt{C = {0,1,2,3}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
-
-This metric respects triangular equality [CHECK THE PROPER NAME] (i.e. $CD_{(A,B)} = CD_{(B,A)}$ and $CD_{(A,C)} = CD_{(B,C)}$).
+When the character difference is equal to $1$ it means it conveys the most different signal.
+For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
+This metric is thus a distance metric since it respects distances equality (i.e. $CD_{(A,B)} = CD_{(B,A)}$ and $CD_{(A,C)} = CD_{(B,C)}$).
 
 \subsection{Simulating discrete morphological matrices}
-
-To simulate the matrices we applied a protocol similar to \citep{Guillerme2016146}.
+To simulate the matrices we applied a protocol very similar to \citep{Guillerme2016146}.
 First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform ($0$,$1$) distribution maintaining $\lambda$ $>$ $\mu$ using the diversitree \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
@@
 These trees are hereafter called the ``true'' trees.
-We then simulated the characters using the either of the two following models:
+We then simulated discrete morphological characters using the either of the two following models:
 \begin{itemize}
-    \item The HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution an scale to sum to one) and using a transition/transvertion rate of two \citep{douadycomparison2003} but where the purines (AG) were changed into state $0$ and the pyrimidines (CT) in state $1$.
-    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][]{OReilly20160081} but can only generate binary state characters.
-    \item To generate more that two states characters, we used the M$k$ model \citep{lewisa2001}.
-    We draw the number of character states with a probability of $0.85$ for two-state characters (binary) and $0.15$ for three state characters \citep{Guillerme2016146}.
-    Assuming the same transition rate between character states might seem unrealistic and ignoring characters following, for example, a Dollo pattern, \cite{Wright01072016} have shown that an equal rate transition is still strictly majority in empirical data.
+    \item The HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ an scale to sum to one) and using a transition/transvertion rate of two \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
+    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][; see below]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
+    \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
+    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146}.
+    This model assumes a equal transition rate between character states that might seem overly simplifying and excludes other observed transition patterns \citep[e.g.
+Dollo characters;][]{Dollo,wright2015came}.
+    Recently however, \cite{Wright01072016} have shown that an equal rate transition is still strictly majority in empirical data.
 \end{itemize}
 
-For each character, both models (HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
+\noindent For each character, both models (HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
 This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices (see supplementary materials @@@).
-We resimulated every invariant characters to obtain a matrix with no invariant characters.
-To ensure the that our simulation where reflecting the same parameters as real data, we only selected matrices with Consistency Indices superior to $0.26$ \citep{OReilly20160081}.
+We re-simulated every invariant characters to obtain a matrix with no invariant characters.
+To ensure that our simulation where reflecting realistic observed parameters, we only selected matrices with Consistency Indices superior to $0.26$ \citep{OReilly20160081}.
 
@@
 
-We then calculated the pairwise character differences for each generated matrices. 
+We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep[][; \url{https://github.com/TGuillerme/dispRity}]{thomas_guillerme_2016_55646}.
+%TG: TO IMPLEMENT!
 We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
@@
 Finally, because this operation effectively changes the weight of characters (i.e.
-giving the characters $<$$0.25$ or $>$$0.75$ a weight of $0$ and giving the randomly selected characters to replace them a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a null expectation modification (i.e.
+giving the characters $<$$0.25$ or $>$$0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a null expectation modification (i.e.
 randomly weighting characters).
-This step resulted in a total of 1800 matrices (hereafter called the ``maximised'', ``minimised'' and ``random'' matrices).
+This step resulted in a total of 1800 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``random'' matrices).
 The algorithms for the three modifications is available on GitHub@@@. %TG: "@@@" is my "to replace" sign.
@@
 
-We inferred the topologies in both Bayesian and Parsimony using MrBayes and PAUP* respectively.
+We inferred the topologies in both Bayesian and Maximum Parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
+The Maximum Parsimony inference was run over @@@ CPU hours using a heuristic search with random sequence addition replicate 10 times (\texttt{hsearch addseq=random nreps=10}).
+The trees where also bootstrapped 100 times using the same search parameters mentioned above.
 
-Bayesian settings.
+The Bayesian inference was run over @@@ CPU hours
 
-PAUP ``settings''.
+% lset nst=1 rates=gamma Ngammacat=4;
+% prset ratepr=variable Shapepr=Exponential(0.5);
+
+% mcmc nruns=2 Nchains=6 ngen=100000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES file=150t_100c_0304_maxi;
+% sump Filename=150t_100c_0304_maxi Relburnin=YES Burninfrac=0.25;
+% sumt Filename=150t_100c_0304_maxi Relburnin=YES Burninfrac=0.25;
 
@@
 
-We compared the topologies using the same approach as in \cite{Guillerme2016146}.
+We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``random'' matrices and the tree inferred from the ``normal'' matrix.
+The metrics scores where obtained using the TreeCmp java script \citep{Bogdanowicz2012}.
+The measurements where then standardised using the Normalised Tree Similarity metric \citep[i.e.
+centering the metrics scores using the mean metric score for 1000 pairwise comparisons between random trees with $n$ taxa;][]{Bogdanowicz2012,Guillerme2016146}.
+When the normalised metric has a score of $1$ it means both trees are identical, when it has a score of $0$ it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
+The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Fould Similarity reflects the conservation of clades (i.e.
+a score close to $1$ indicates that most clades are identical in both trees); and (2) the Normalised Triplets Similarity reflects the position of taxa (i.e.
+a score close to $1$ indicates that most taxa have the same neighbours in both trees).
 
@@
 European Research Council under the European Union’s Seventh Framework Programme (FP/2007–2013)/ERC Grant Agreement number 311092.
+Cluster acknowledgements
+Bertie
 

### 2017-02-27 Writing/draft.tex
Draft equations proofs

@@
 \begin{equation}
-    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\abs*{\sum_{i}^{n}(x_{i} - y_{i})}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
+    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
 \end{equation}
@@
 
-% mcmc nruns=2 Nchains=6 ngen=100000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES file=150t_100c_0304_maxi;
+% mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES file=150t_100c_0304_maxi;
 % sump Filename=150t_100c_0304_maxi Relburnin=YES Burninfrac=0.25;
@@
 % Suggested reviewers:
-%- Liliana Davalos
+%- Liliana Davalos for MPE
+
+\section{Appendix A - Character difference}
+
+The Character Difference ($CD$) metric is a distance $d: CD \times CD \rightarrow [0;\infty)$ with any character $x$, $y$ and $z$ $\in CD$ if an only if the followings are satisfied:
+
+\begin{equation}
+    \label{Positivity}
+    CD_{(x,y)} \geq 0
+\end{equation}
+\begin{equation}
+    \label{Identity}
+    CD_{(x,y)} = 0 \Leftrightarrow x = y
+\end{equation}
+\begin{equation}
+    \label{Equality}
+    CD_{x,y} = CD_{y,x}
+\end{equation}
+\begin{equation}
+    \label{Subadditivity}
+    CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}
+\end{equation}
+
+\subsection{Character difference positivity}
+
+For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ (because the difference between character states is considered as Fitch-like, i.e.
+the difference between two characters states \texttt{0} and \texttt{8} is equal to $1$ and the difference between two identical character states is equal to $0$:
+
+\begin{equation}
+    \label{positive_demo}
+    \begin{gathered}
+    0 \leq  \abs{x_{n}-y_{n}} \leq 1\\
+    0 \leq  \sum_{i=1}^{n}\abs{x_{i}-y_{i}} \leq n\\
+    0 \leq  \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} \leq 1\\
+    0 \leq  \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} \leq \frac{1}{2}\\
+    0 \leq  \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} \leq 1\\
+    0 \leq  1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) \leq 1\\
+    \end{gathered}
+\end{equation}
+
+Thus equation \ref{Positivity} is true.
+
+\subsection{Character difference identity}
+
+For any $n \in [1;\infty)$ and $x_{n} = y_{n}$ then:
+%TG: Is that valid for a demonstration?
+\begin{equation}
+    \label{identitydemo}
+    \begin{aligned}
+    \abs{x_{n}-y_{n}} &= 0\\
+    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= 0\\
+    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= 0\\
+    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= 0\\
+    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= 0\\
+    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 0\\
+    \end{aligned}
+\end{equation}
+
+Thus equation \ref{Identity} is true.
+
+\subsection{Character difference equality}
+
+For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$, similarly to equation \ref{positive_demo}:
+
+\begin{equation}
+    \label{equality_demo}
+    \begin{aligned}
+    \abs{x_{n}-y_{n}} &= \abs{y_{n}-x_{n}}\\
+    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= \sum_{i}^{n}\abs{y_{i}-x_{i}}\\
+    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= \frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}\\
+    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\\
+    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= \frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\\
+    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)\\
+    \end{aligned}
+\end{equation}
+
+Thus equation \ref{Equality} is true.
+
+\subsection{Character difference subadditivity}
+
+As described in equations \ref{Positivity}, \ref{Identity} and \ref{Equality},
+
+\begin{equation}
+    \label{subadditivity_demo}
+    \begin{aligned}
+    &0\leq CD_{x,z} \leq 1\\
+    &\text{and}\\
+    &0\leq CD_{z,y} \leq 1\\
+    &\text{thus}\\
+    &0\leq CD_{x,z} + CD_{z,y}  \leq 2\\
+    \end{aligned}
+\end{equation}
+
+\subsubsection{Case 1}
+If $CD_{x,z} = 0$, then, following \ref{Identity}:
+
+\begin{equation}
+    \label{subadditivity_demo_case1}
+    \begin{aligned}
+    &CD_{x,z} = CD_{z,y} = CD_{x,y}\\
+    &\text{then}\\
+    &CD_{x,y} \leq CD_{x,z} + CD_{z,y}\\
+    \end{aligned}
+\end{equation}
+
+\subsubsection{Case 2}
+If $CD_{x,z} > 0$, then:
+
+\begin{equation}
+    \label{subadditivity_demo_case2}
+    \begin{aligned}
+    &CD_{x,z} = CD_{z,y} = CD_{x,y}\\
+    &\text{then}\\
+    &CD_{x,y} \leq CD_{x,z} + CD_{z,y}\\
+    \end{aligned}
+\end{equation}
+
+% \subsubsection{Character difference metric is ontological (identity of indiscernibles)}
+
+% \begin{equation}
+%     \label{CDnull}
+%     CD_{(x,y)} = 0 \Leftrightarrow x = y
+% \end{equation}
+% This is because of equation \ref{equality}; in fact $x$ and $y$ is actually expressed in terms of phylogenetic information per character (i.e. normalised for character state token).
+
+% \subsubsection{Character difference metric is sub-additive (triangle inequality)}
+
+% \begin{equation}
+%     \label{subadditive}
+%     CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}
+% \end{equation}
+% Following equations \ref{CDpositive}, \ref{equality}, \ref{CDnull}.
+
+% Thus:
+
+% \begin{equation}
+%     d: \frac{\abs*{\sum_{i}^{n}(x_{i} - y_{i})}{n} \rightarrow [0;1]
+% \end{equation}
+
+% Thus:
+
+% \begin{equation}
+%     d: \frac{\sum_{i}^{n}(x_{i} - y_{i})}{n} \rightarrow [0;1]
+% \end{equation}
 

### 2017-02-27 Functions/shells/CC_TreeCmp.sh
Updated TreeCmp shell (now intakes specific trees)

@@ -4,6 +4,7 @@
 #SYNTAX: TreeCmp -c <chain> -o <output> -f <treeformat> -d <draws> -r <rooted>
 #with:
 #-c <chain> the name of the chain
+#-r <ref> the reference tree number (1 = normal, 2 = maximum, 3 = minimum, 4 = null)
 #-o <output> OPTIONAL the output name (if missing, is set to <chain>)
 #-f <treeformat> OPTIONAL the tree in and out format (if missing, is set to '.nex')
 #-d <draws> OPTIONAL the number of draws for the comparison (if missing, is set to '1000')
@@ -28,6 +29,10 @@ key="$1"
 case $key in
     -c|--chain)
         CHAIN="$2"
+        shift
+        ;;
+    -r|--ref)
+        REF="$2"
         shift
         ;;
     -o|--output)
@@ -57,6 +62,11 @@ if [ "$CHAIN" == "" ]
 then
     echo "ERROR: --chain is missing!"
     exit
+fi
+
+if [ "$REF" == "" ]
+then
+    REF=1
 fi
 
 if [ "$OUTPUT" == "" ]
@@ -97,6 +107,7 @@ else
     fi
 fi
 
+# ##DEBUG
 # echo "chain name = ${CHAIN}"
 # echo "output name = ${OUTPUT}"
 # echo "format out = ${FORMAT}"
@@ -116,7 +127,23 @@ else
 fi
 
 ## Get the Chain folder name
-chain_folder=$(echo $CHAIN | sed 's/c_.*/c_cmp/g')
+if [ "$REF" == 1 ]
+then
+    chain_folder=$(echo $CHAIN | sed 's/c_.*/c_cmp_norm/g')
+fi
+if [ "$REF" == 2 ]
+then
+    chain_folder=$(echo $CHAIN | sed 's/c_.*/c_cmp_maxi/g')
+fi
+if [ "$REF" == 3 ]
+then
+    chain_folder=$(echo $CHAIN | sed 's/c_.*/c_cmp_mini/g')
+fi
+if [ "$REF" == 4 ]
+then
+    chain_folder=$(echo $CHAIN | sed 's/c_.*/c_cmp_null/g')
+fi
+
 
 ## Create the folder if needed
 if ls -d | grep ${chain_folder} > /dev/null
@@ -136,8 +163,26 @@ then
         sh Nex2New.sh ${CHAIN} "newick"
 
         ## Save the three trees in a same file
[…]

### 2017-03-06 Writing/draft.tex
corrected proof for equation 3

@@
     \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= 0\\
-    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= 0\\
-    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= 0\\
+    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \frac{1}{2}\\
+    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= 1\\
     1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 0\\
@@
 
-As described in equations \ref{Positivity}, \ref{Identity} and \ref{Equality},
+% As described in equations \ref{Positivity}, \ref{Identity} and \ref{Equality},
 
-\begin{equation}
-    \label{subadditivity_demo}
-    \begin{aligned}
-    &0\leq CD_{x,z} \leq 1\\
-    &\text{and}\\
-    &0\leq CD_{z,y} \leq 1\\
-    &\text{thus}\\
-    &0\leq CD_{x,z} + CD_{z,y}  \leq 2\\
-    \end{aligned}
-\end{equation}
+% \begin{equation}
+%     \label{subadditivity_demo}
+%     \begin{aligned}
+%     &0\leq CD_{x,z} \leq 1\\
+%     &\text{and}\\
+%     &0\leq CD_{z,y} \leq 1\\
+%     &\text{thus}\\
+%     &0\leq CD_{x,z} + CD_{z,y}  \leq 2\\
+%     \end{aligned}
+% \end{equation}
 
-\subsubsection{Case 1}
-If $CD_{x,z} = 0$, then, following \ref{Identity}:
+% \subsubsection{Case 1}
+% If $CD_{x,z} = 0$, then, following \ref{Identity}:
 
-\begin{equation}
-    \label{subadditivity_demo_case1}
-    \begin{aligned}
-    &CD_{x,z} = CD_{z,y} = CD_{x,y}\\
-    &\text{then}\\
-    &CD_{x,y} \leq CD_{x,z} + CD_{z,y}\\
-    \end{aligned}
-\end{equation}
+% \begin{equation}
+%     \label{subadditivity_demo_case1}
+%     \begin{aligned}
+%     &CD_{x,z} = CD_{z,y} = CD_{x,y}\\
+%     &\text{then}\\
+%     &CD_{x,y} \leq CD_{x,z} + CD_{z,y}\\
+%     \end{aligned}
+% \end{equation}
 
-\subsubsection{Case 2}
-If $CD_{x,z} > 0$, then:
+% \subsubsection{Case 2}
+% If $CD_{x,z} > 0$, then:
 
-\begin{equation}
-    \label{subadditivity_demo_case2}
-    \begin{aligned}
-    &CD_{x,z} = CD_{z,y} = CD_{x,y}\\
-    &\text{then}\\
-    &CD_{x,y} \leq CD_{x,z} + CD_{z,y}\\
-    \end{aligned}
-\end{equation}
+% \begin{equation}
+%     \label{subadditivity_demo}
+%     \begin{aligned}
+    % CD_{x,y} &\leq CD_{x,z} + CD_{z,y}\\
+    % 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) 
+    %  &\leq 
+    %  1 - \left(\frac{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}}}{\frac{1}{2}}\right) 
+    %  +
+    %  1 - \left(\frac{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}}}{\frac{1}{2}}\right) \\
+    % \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}} 
+    %  &\geq 
+    %  \abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}}}} 
+    %  +
+    %  \abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}}}} - 2\\
+
+     % Add brackets symbols
+
+     % \in[0,\frac{1}{2}] &\geq (-2) \in[0,\frac{1}{2}] \in[0,\frac{1}{2}]\\
+     % \in[0,\frac{1}{2}] &\geq \in[-2,-1]\\
+
+%     \end{aligned}
+% \end{equation}
 

### 2017-03-06 Writing/draft.tex
Added proof for CD metric equation 5

--- 2017-02-27 Writing/draft.tex
@@
 Bertie
+Guillermo Herraiz Yebes for helping with the CD metric proof.
 
@@
 
-As described in equations \ref{Positivity}, \ref{Identity} and \ref{Equality},
-
 \begin{equation}
@@
     \begin{aligned}
-    &0\leq CD_{x,z} \leq 1\\
-    &\text{and}\\
-    &0\leq CD_{z,y} \leq 1\\
-    &\text{thus}\\
-    &0\leq CD_{x,z} + CD_{z,y}  \leq 2\\
+    CD_{x,y} &\leq CD_{x,z} + CD_{z,y}\\
+    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
+     &\leq 
+     1 - \left(\frac{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}}}{\frac{1}{2}}\right) 
+     +
+     1 - \left(\frac{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}}}{\frac{1}{2}}\right) \\
+    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} 
+     &\geq 
+     \abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}}
+     +
+     \abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}} - 2\\
+          % Add brackets symbols
+     \in[0,\frac{1}{2}] &\geq (-2) \in[0,\frac{1}{2}] \in[0,\frac{1}{2}]\\
+     \in[0,\frac{1}{2}] &\geq \in[-2,-1]\\
     \end{aligned}
 \end{equation}
-
-\subsubsection{Case 1}
-If $CD_{x,z} = 0$, then, following \ref{Identity}:
-
-\begin{equation}
-    \label{subadditivity_demo_case1}
-    \begin{aligned}
-    &CD_{x,z} = CD_{z,y} = CD_{x,y}\\
-    &\text{then}\\
-    &CD_{x,y} \leq CD_{x,z} + CD_{z,y}\\
-    \end{aligned}
-\end{equation}
-
-\subsubsection{Case 2}
-If $CD_{x,z} > 0$, then:
-
-\begin{equation}
-    \label{subadditivity_demo_case2}
-    \begin{aligned}
-    &CD_{x,z} = CD_{z,y} = CD_{x,y}\\
-    &\text{then}\\
-    &CD_{x,y} \leq CD_{x,z} + CD_{z,y}\\
-    \end{aligned}
-\end{equation}
-
-% \subsubsection{Character difference metric is ontological (identity of indiscernibles)}
-
-% \begin{equation}
-%     \label{CDnull}
-%     CD_{(x,y)} = 0 \Leftrightarrow x = y
-% \end{equation}
-% This is because of equation \ref{equality}; in fact $x$ and $y$ is actually expressed in terms of phylogenetic information per character (i.e. normalised for character state token).
-
-% \subsubsection{Character difference metric is sub-additive (triangle inequality)}
-
-% \begin{equation}
-%     \label{subadditive}
-%     CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}
-% \end{equation}
-% Following equations \ref{CDpositive}, \ref{equality}, \ref{CDnull}.
-
-% Thus:
-
-% \begin{equation}
-%     d: \frac{\abs*{\sum_{i}^{n}(x_{i} - y_{i})}{n} \rightarrow [0;1]
-% \end{equation}
-
-% Thus:
-
-% \begin{equation}
-%     d: \frac{\sum_{i}^{n}(x_{i} - y_{i})}{n} \rightarrow [0;1]
-% \end{equation}
 

### 2017-03-20 Functions/shells/rerun.missing.sh
Updated path in rerun shell

@@ -58,7 +58,7 @@ then
 
     ## Modify the job file
     # sed -i -e 's/pbsexec mpiexec mb \$HOME\/CharSim\/Bayesian\/'"${chain}"'_norm.mbcmd/cp \$HOME\/CharSim\/Bayesian\/'"${chain}"'_norm.append \$WORK\/CharSim\/Bayesian\/'"${chain}"'\/'"${chain}"'_norm ; pbsexec mpiexec mb \$WORK\/CharSim\/Bayesian\/'"${chain}"'_norm/' ${chain}.mbjob-bis
-    sed -i -e 's/pbsexec mpiexec mb \$HOME\/CharSim\/Bayesian\/'"${chain}"'_norm.mbcmd/CURRENT=pwd ; cd \$WORK\/CharSim\/Bayesian\/'"${chain}"'\/ ; cp \$HOME\/CharSim\/Bayesian\/'"${chain}"'_norm.append '"${chain}"'_norm ; pbsexec mpiexec mb \$WORK\/CharSim\/Bayesian\/'"${chain}"'_norm ; cd \$CURRENT/' ${chain}.mbjob-bis
+    sed -i -e 's/pbsexec mpiexec mb \$HOME\/CharSim\/Bayesian\/'"${chain}"'_norm.mbcmd/cd \$WORK\/CharSim\/Bayesian\/'"${chain}"'\/ ; cp \$HOME\/CharSim\/Bayesian\/'"${chain}"'_norm.append '"${chain}"'_norm ; pbsexec mpiexec mb '"${chain}"'_norm ; cd \$HOME\/CharSim\/Bayesian\//' ${chain}.mbjob-bis
 fi
 
 if echo $broken | grep '2' > /dev/null
@@ -71,7 +71,7 @@ then
     sed -i -e 's/mcmc nruns=2/mcmc append=yes nruns=2/' ${chain}_maxi.append
 
     ## Modify the job file
[…]

### 2017-03-20 Writing/draft.tex
polished up equations
Added felsenstein notation to the character correlation explanation

@@
 
-\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$ \citep[i.e.
-the Gower distance;][]{GowerDist}.
+\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
+$CD$ is a continuous distance metric valid in a mathematical sense (see Appendix 1).
 Since we are considering differences as being only Fitch-like (i.e.
-non-weighted and non-ordinate), we ignore the actual distance between two different character states tokens: two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g.
-$0 - 0 = 0$ or $1 - 8 = 1$).
+non-weighted and non-ordinate), we ignore the actual distance between two different character states tokens \citep[i.e.
+a Gower distance;][]{GowerDist}.
+Two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
 In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
 In other words, we replaced all the occurrences of the first token to be $0$, the second to be $1$, etc.
-This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{0,0,1,2,2,1\}}.
+This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{0,0,1,2,2,1\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
@@
 For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
-This metric is thus a distance metric since it respects distances equality (i.e. $CD_{(A,B)} = CD_{(B,A)}$ and $CD_{(A,C)} = CD_{(B,C)}$).
 
@@
 For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ (because the difference between character states is considered as Fitch-like, i.e.
-the difference between two characters states \texttt{0} and \texttt{8} is equal to $1$ and the difference between two identical character states is equal to $0$:
+the difference between two characters states \texttt{0} and \texttt{8} is equal to $1$ and the difference between two identical character states is equal to $0$):
 
@@
 
+For any $n, j, b \in [1;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:
+
 \begin{equation}
@@
     CD_{x,y} &\leq CD_{x,z} + CD_{z,y}\\
+    \text{if:}\\
     1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
@@
      1 - \left(\frac{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}}}{\frac{1}{2}}\right) \\
-    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} 
+     \text{then, the opposite is:}\\
+    \underbrace{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} }_{\in[0,1/2]}
      &\geq 
-     \abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}}
+     \underbrace{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}} }_{\in[0,1/2]}
      +
-     \abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}} - 2\\
-          % Add brackets symbols
-     \in[0,\frac{1}{2}] &\geq (-2) \in[0,\frac{1}{2}] \in[0,\frac{1}{2}]\\
+     \underbrace{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}} }_{\in[0,1/2]} - 2\\
+     \text{thus:}\\
+     \in[0,\frac{1}{2}] &\geq \in[0,\frac{1}{2}] \in[0,\frac{1}{2}] (-2) \\
      \in[0,\frac{1}{2}] &\geq \in[-2,-1]\\
@@
 \end{equation}
+
+Thus equation \ref{Subadditivity} is true.
 

### 2017-03-22 Writing/draft.tex
Updated manuscript (intro)

@@
 
-The last five years have seen a drastic expansion (resurgence?) of the use of discrete morphological data in macroevolutionary studies.
-In fact, traditionally confined to classic cladistics \citep{felsenstein2004inferring} these methods have recently been highly improved whether based on the models and methods for inferring phylogenies (\citealt{lewisa2001} + BAPSTinReview; with some debate on which method to use: \citealt{spencerefficacy2013,wrightbayesian2014,OReilly20160081}) or the methods to date them \citep[e.g.][]{bapsta2013} or to combine them with molecular data \citep{pyrondivergence2011,ronquista2012}.
+The last five years have seen a drastic expansion of the use of discrete morphological data in macroevolutionary studies.
+In fact, traditionally confined to ``classic'' cladistics \citep{felsenstein2004inferring}, these methods have recently been expended and improved whether based on the models and methods for inferring phylogenies \citep{lewisa2001,Wright01072016} or the methods to date them \citep[e.g.][]{bapsta2013} or to combine them with molecular data \citep{pyrondivergence2011,ronquista2012}.
+% Debate on which inference method to use \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081}.
+% Debate on dating methods \citep{Arcila2015131}.
+% Debate on data sources (BAPST in review)
 In parallel, much development has also been done on pre-phylogenetic analysis \citep[e.g.
@@
 morphological disparity analysis;][]{Close2015,Claddis}.
+However, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146}(+BAPST); the coding of this data \citep{Brazeau2011}; the inference \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081} or the dating methods \citep{Arcila2015131}.
 
-However, using discrete morphological data comes with several caveats, like the effect of missing data \citep{Guillerme2016146} or of the method \citep{OReilly20160081}.
+However, one aspect of the data is often neglected: the correlation that exists between discrete morphological characters.
+In fact, because of their nature, these characters are prone to many sources of correlation such as: (1) the co-evolution between traits (e.g.
+the number of digits on a limb co-evolves with the actual presence/absence of the limb); (2) morphological modularity \citep[i.e.
+the unity of entire morphological regions][]{goswami2006morphological}; (3) convergent evolution between traits (e.g.
+wings in many vertebrates); or (4), more down-to-earth, to errors from the coder \citep[e.g.
+the repetition of the same character described slighlty differently each case][]{Brazeau2011}.
+These sources of correlation can alter the results of phylogenetic inference by effectively weighting the correlated characters against their non-correlated counter parts leading to biases in phylogenetic inference by supporting clades with artificial synapomorphies.
+This fact have led to the popular wisdom rule of avoiding coding correlated characters for macroevolutionary studies.
 
-Common wisdom says that characters should not be correlated
+Up to this date, a lot of work has been done on character correlation \textit{per se}, but focussing mainly on how to measure correlation between characters to detect patterns of coevolution \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
+However, only several attempts have been made on the actual effect of characters correlation on phylogenetic inference with two notable exceptions of \cite{Davalos01072014} on empirical data and \cite{ZouConvergence} on molecular data.
+It is thus still unclear what will be the effect of characters correlation on phylogenetic inference: how does these correlation really affect topology?
+Additionally, because of the nature of discrete morphological characters and organisms morphology, correlation is expected to rise through both combinatorics properties of the characters or simply through evolutionary processes.
+What is the effect of these two sources of correlations and can we (should we) avoid them?
 
-One specific assumption is that coded characters should not be correlated \citep{Brazeau2011}[FIND BETTER CITE?].
-
-Character correlation can be due to
-- co-evolution of two traits (e.g.
-the character ``number of digits on a limb'' co-evolves with the character ``presence/absence of limb'' - this correlation can be sometimes fairly obvious, sometimes less)
-- modularity of morphological features [CITE GOSWAMI!]
-- convergent evolution (e.g. the character ``flight'') [???]
-- error from the coder (e.g.
-a coder can unintentionally code the same feature multiple times, thus effectively waiting it, for example: ``presence/absence of a proximal end on a limb'' and ``presence/absence of a distal end on a limb'' - effectively weighting the character ``presence/absence of a limb'' by two).
-%-Don't code correlated characters but what is the actual effect of that? And how many characters are ``secretly'' correlated?
-
-A lot of work has been done on character correlation \textit{per se}, but focussing mainly on how to measure correlation between characters to detect patterns of coevolution \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
-However, only several attempts have been made on the actual effect of characters correlation on phylogenetic inference with two notable expections of \cite{Davalos01072014} on empirical data and \cite{ZouConvergence} on molecular data.
-It is thus still unclear what will be the effect of characters correlation on phylogenetic inference.
-
-% \cite{Davalos01072014}: character correlation + signal in bats
-% \cite{Grabowski2016}: amount of data needed for testing correlations? CHECK
-% \cite{Lande1983}: measure correlation CHECK
-% \cite{Pagel1994}: measure correlation
-% \cite{Pagel2006}: measure correlation Bayesian
-% \cite{Maddison1990}: measure binary state character correlation for testing coevol
-% \cite{ZouConvergence}: convergence measured as characters correlation (Removing convergence prone characters improves accuracy)
-
-In this study, we propose a new way to measure character correlation (more adapted to discrete morphological characters) and test on simulated and empirical data what is the effect of character correlation on topology.
-%TG: check if we do just topology?
-In other words, we generate matrices and maximise or minimise character correlation and see how that affects topology.
+In this study, we formally assess the effect of discrete character's correlation using simulated and empirical data.
+We propose a new metric to measure the correlation (or difference) between two discrete characters and a protocol to modify discrete data matrices to increase/decrease the correlation between characters within/
+We found that blabalbal
 
@@
 We define characters as being correlated if they give the same phylogenetic information.
-In order to measure the difference between two characters, we propose a new distance metric defined as follows:
+In order to measure this similitude in phylogenetic information, we propose a new distance metric to measure the difference between two characters:
 
@@
 
-We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep[][; \url{https://github.com/TGuillerme/dispRity}]{thomas_guillerme_2016_55646}.
-%TG: TO IMPLEMENT!
+We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}. %TG: TO IMPLEMENT! 
 We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
@@
 This step resulted in a total of 1800 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``random'' matrices).
-The algorithms for the three modifications is available on GitHub@@@. %TG: "@@@" is my "to replace" sign.
+The algorithms for the three modifications are available on GitHub@@@. %TG: "@@@" is my "to replace" sign.
 
@@
 
-\section{Acknowledgements}
-European Research Council under the European Union’s Seventh Framework Programme (FP/2007–2013)/ERC Grant Agreement number 311092.
-Cluster acknowledgements
-Bertie
-Guillermo Herraiz Yebes for helping with the CD metric proof.
-
 \section{Results}
@@
 \subsection{Conclusion}
+
+\section{Funding}
+European Research Council under the European Union’s Seventh Framework Programme (FP/2007–2013)/ERC Grant Agreement number 311092.
+
+\section{Acknowledgements}
+Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
+We thank Alberto Pascual Garcia and April Wright for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD proof.
 

### 2017-04-24 Functions/shells/CC_TreeCmp.sh
Updated tree comparisons

@@ -141,7 +141,7 @@ then
 fi
 if [ "$REF" == 4 ]
 then
-    chain_folder=$(echo $CHAIN | sed 's/c_.*/c_cmp_null/g')
+    chain_folder=$(echo $CHAIN | sed 's/c_.*/c_cmp_rand/g')
 fi
 
 
@@ -170,15 +170,15 @@ then
         fi
         if [ "$REF" == 2 ]
         then
-            cat ${CHAIN}_norm.con.tre.newick > ${CHAIN}_ref.tre
+            cat ${CHAIN}_maxi.con.tre.newick > ${CHAIN}_ref.tre
         fi
         if [ "$REF" == 3 ]
         then
-            cat ${CHAIN}_norm.con.tre.newick > ${CHAIN}_ref.tre
+            cat ${CHAIN}_mini.con.tre.newick > ${CHAIN}_ref.tre
         fi
         if [ "$REF" == 4 ]
         then
-            cat ${CHAIN}_norm.con.tre.newick > ${CHAIN}_ref.tre
+            cat ${CHAIN}_rand.con.tre.newick > ${CHAIN}_ref.tre
         fi
 
         cat ${CHAIN}_norm.con.tre.newick > ${CHAIN}_comb.con.tre
@@ -197,15 +197,15 @@ then
         fi
         if [ "$REF" == 2 ]
         then
-            cat ${CHAIN}_norm.con.tre > ${CHAIN}_ref.tre
+            cat ${CHAIN}_maxi.con.tre > ${CHAIN}_ref.tre
         fi
         if [ "$REF" == 3 ]
         then
-            cat ${CHAIN}_norm.con.tre > ${CHAIN}_ref.tre
+            cat ${CHAIN}_mini.con.tre > ${CHAIN}_ref.tre
         fi
         if [ "$REF" == 4 ]
         then
-            cat ${CHAIN}_norm.con.tre > ${CHAIN}_ref.tre
+            cat ${CHAIN}_rand.con.tre > ${CHAIN}_ref.tre
         fi
         
         cat ${CHAIN}_norm.con.tre > ${CHAIN}_comb.con.tre

### 2017-04-24 Writing/draft.tex
Updated methods (signal/noise)

@@
 randomly weighting characters).
-This step resulted in a total of 1800 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``random'' matrices).
+This step resulted in a total of 1800 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``null'' matrices).
 The algorithms for the three modifications are available on GitHub@@@. %TG: "@@@" is my "to replace" sign.
@@
 
-We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``random'' matrices and the tree inferred from the ``normal'' matrix.
+We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``null'' matrices and the tree inferred from the ``normal'' matrix.
+
+We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the different inferred trees.
+We ran two types of comparisons, (1) the ability to recover the correct tree (``normal'' trees \textit{vs} ``null'' and ``true'' trees and (2) the effect of character differences by measuring the noise to signal ratio between the ``maximised'' and ``minimised'' trees and the ``null'' and ``normal'' trees.
+
 The metrics scores where obtained using the TreeCmp java script \citep{Bogdanowicz2012}.

### 2017-04-25 Analysis/Matrix_generation_pipeline.R
Cleaned pipeline analysis

@@ -1,16 +1,20 @@
 ## Pipeline for generating the matrices.
 ## Refer to Matrix modification vignette for more details.
 
-## Packages and functions
+## Packages
 library(Claddis)
 library(dispRity)
 library(diversitree)
+
+## Load the functions
 source("functions.R") ; load.functions(test = FALSE)
-dyn.load("../Functions/char.diff.so")
+## Modifying ape::write.nexus function
 write.nexus.std <- ape::write.nexus.data
 body(write.nexus.std)[[2]] <- substitute(format <- match.arg(toupper(format), c("DNA", "PROTEIN", "STANDARD")))
 body(write.nexus.std)[[26]][[3]][[4]] <- substitute(fcat(indent, "FORMAT", " ", DATATYPE, " ", MISSING, " ", GAP, 
     " ", INTERLEAVE, " symbols=\"0123456789\";\n"))
+
+## Counting digits in a column
 get.digit <- function(column) {
     if(max(nchar(round(column)), na.rm = TRUE) <= 4) {
        return(4-max(nchar(round(column)), na.rm = TRUE))
@@ -18,6 +22,8 @@ get.digit <- function(column) {
         return(0)
     }
 }
+
+## Birth death sampler with birth > death
 sample.birth.death <- function() {
     birth <- runif(min = 0, max = 1, 1)
     death <- runif(min = 0, max = birth, 1)
@@ -41,7 +47,6 @@ simulationID <- unlist(mapply(paste.ID, as.list(digits), as.list(simulationID)))
 
 
 for(simulation in 1:length(simulationID)) {
-    ## Set the variables
     ## Set the variables
     if(ceiling(simulation/150) == 1) {
         ntaxa <- ntaxa_list[1]

### 2017-04-25 Writing/draft.tex
Updated outline in draft and remove supplementaries

@@
 
+We applied the following simulation protocol to assess the effect of characters correlation (Fig.\ref{Fig:outline}):
+\begin{enumerate}
+    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
+    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by duplicating characters in order to maximise and minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
+We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximised or minimised character difference to create null expectancy matrices (hereafter called the ``null'' matrices).
+    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``null'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``null'' trees).
+    \item \textbf{Comparing topologies}: finally, we compared to topologies in two ways: (1) first we compared the ``normal'' to the ``null'' trees to assess our simulation protocol; (2) second, we compared the ``maximised'' and ``minimised'' trees to the ``null'' and ``normal'' trees respectively to measure the signal to noise ratio and thus measure the effect of character correlation on topology.
+\end{enumerate}
+Each steps are described below in more details, as well as our proposed definition of the Character Difference metric.
+
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
+\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using Maximum Parsimony and Bayesian inference (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).}
+\label{Fig:outline}
+\end{figure}
+
 \subsection{Measuring correlation between characters}
-
 We define characters as being correlated if they give the same phylogenetic information.
@@
 \noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
-$CD$ is a continuous distance metric valid in a mathematical sense (see Appendix 1).
+$CD$ is a continuous distance metric valid in a mathematical sense (see Supplementary).
 Since we are considering differences as being only Fitch-like (i.e.
@@
 
-\subsection{Inferring the phylogenies}
+\subsection{Inferring topologies}
 
@@
 
-\subsection{Comparing the topologies}
+\subsection{Comparing topologies}
 
@@
 
-\section{Appendix A - Character difference}
-
-The Character Difference ($CD$) metric is a distance $d: CD \times CD \rightarrow [0;\infty)$ with any character $x$, $y$ and $z$ $\in CD$ if an only if the followings are satisfied:
-
-\begin{equation}
-    \label{Positivity}
-    CD_{(x,y)} \geq 0
-\end{equation}
-\begin{equation}
-    \label{Identity}
-    CD_{(x,y)} = 0 \Leftrightarrow x = y
-\end{equation}
-\begin{equation}
-    \label{Equality}
-    CD_{x,y} = CD_{y,x}
-\end{equation}
-\begin{equation}
-    \label{Subadditivity}
-    CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}
-\end{equation}
-
-\subsection{Character difference positivity}
-
-For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ (because the difference between character states is considered as Fitch-like, i.e.
-the difference between two characters states \texttt{0} and \texttt{8} is equal to $1$ and the difference between two identical character states is equal to $0$):
-
-\begin{equation}
-    \label{positive_demo}
-    \begin{gathered}
-    0 \leq  \abs{x_{n}-y_{n}} \leq 1\\
-    0 \leq  \sum_{i=1}^{n}\abs{x_{i}-y_{i}} \leq n\\
-    0 \leq  \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} \leq 1\\
-    0 \leq  \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} \leq \frac{1}{2}\\
-    0 \leq  \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} \leq 1\\
-    0 \leq  1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) \leq 1\\
-    \end{gathered}
-\end{equation}
-
-Thus equation \ref{Positivity} is true.
-
-\subsection{Character difference identity}
-
-For any $n \in [1;\infty)$ and $x_{n} = y_{n}$ then:
-%TG: Is that valid for a demonstration?
-\begin{equation}
-    \label{identitydemo}
-    \begin{aligned}
-    \abs{x_{n}-y_{n}} &= 0\\
-    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= 0\\
-    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= 0\\
-    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \frac{1}{2}\\
-    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= 1\\
-    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 0\\
-    \end{aligned}
-\end{equation}
-
-Thus equation \ref{Identity} is true.
-
-\subsection{Character difference equality}
-
-For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$, similarly to equation \ref{positive_demo}:
-
-\begin{equation}
-    \label{equality_demo}
-    \begin{aligned}
-    \abs{x_{n}-y_{n}} &= \abs{y_{n}-x_{n}}\\
-    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= \sum_{i}^{n}\abs{y_{i}-x_{i}}\\
-    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= \frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}\\
-    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\\
-    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= \frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\\
-    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)\\
-    \end{aligned}
-\end{equation}
-
-Thus equation \ref{Equality} is true.
-
-\subsection{Character difference subadditivity}
-
-For any $n, j, b \in [1;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:
-
-\begin{equation}
-    \label{subadditivity_demo}
-    \begin{aligned}
-    CD_{x,y} &\leq CD_{x,z} + CD_{z,y}\\
-    \text{if:}\\
-    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
-     &\leq 
-     1 - \left(\frac{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}}}{\frac{1}{2}}\right) 
-     +
-     1 - \left(\frac{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}}}{\frac{1}{2}}\right) \\
-     \text{then, the opposite is:}\\
-    \underbrace{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} }_{\in[0,1/2]}
-     &\geq 
-     \underbrace{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}} }_{\in[0,1/2]}
-     +
-     \underbrace{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}} }_{\in[0,1/2]} - 2\\
-     \text{thus:}\\
-     \in[0,\frac{1}{2}] &\geq \in[0,\frac{1}{2}] \in[0,\frac{1}{2}] (-2) \\
-     \in[0,\frac{1}{2}] &\geq \in[-2,-1]\\
-    \end{aligned}
-\end{equation}
-
-Thus equation \ref{Subadditivity} is true.
-
 \bibliographystyle{sysbio}

### 2017-04-25 Writing/supplementary/Sup1-CharDiff.tex
Character Difference sup. mat

\documentclass[12pt,letterpaper]{article}
\usepackage{natbib}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{dcolumn}
\usepackage{mathtools}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\newcolumntype{d}[1]{D{.}{.}{#1}}

\pagenumbering{arabic}

%Pagination style and stuff
\linespread{2}
\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}
%\bibpunct{(}{)}{;}{a}{}{,}

%---------------------------------------------
%
%       START
%
%---------------------------------------------

\begin{document}

%Running head
\begin{flushright}
Version dated: \today
\end{flushright}
\bigskip
\noindent RH: Characters correlation

\bigskip
\medskip
\begin{center}

\noindent{\Large \bf Effect of discrete character correlation on tree inference}
\noindent{\Large \bf Supplementary material 1 - Character difference}
\bigskip

\noindent {\normalsize \sc Thomas Guillerme$^1$$^*$, and Martin D. Brazeau$^1$}\\
\noindent {\small \it 
$^1$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\}
\end{center}
\medskip
\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\

%Line numbering
% \modulolinenumbers[1]
% \linenumbers

\section{Definition}

\begin{equation}
    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
\end{equation}
\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.

\newpage
\section{Distance metric proof}

The Character Difference ($CD$) metric is a distance $d: CD \times CD \rightarrow [0;\infty)$ with any character $x$, $y$ and $z$ $\in CD$ if an only if the followings are satisfied:

\begin{equation}
    \label{Positivity}
    CD_{(x,y)} \geq 0
\end{equation}
\begin{equation}
    \label{Identity}
    CD_{(x,y)} = 0 \Leftrightarrow x = y
\end{equation}
\begin{equation}
    \label{Equality}
    CD_{x,y} = CD_{y,x}
\end{equation}
\begin{equation}
    \label{Subadditivity}
    CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}
\end{equation}

\subsection{Character difference positivity}

For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ (because the difference between character states is considered as Fitch-like, i.e. the difference between two characters states \texttt{0} and \texttt{8} is equal to $1$ and the difference between two identical character states is equal to $0$):

\begin{equation}
    \label{positive_demo}
    \begin{gathered}
    0 \leq  \abs{x_{n}-y_{n}} \leq 1\\
    0 \leq  \sum_{i=1}^{n}\abs{x_{i}-y_{i}} \leq n\\
    0 \leq  \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} \leq 1\\
    0 \leq  \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} \leq \frac{1}{2}\\
    0 \leq  \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} \leq 1\\
    0 \leq  1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) \leq 1\\
    \end{gathered}
\end{equation}

Thus equation \ref{Positivity} is true.

\subsection{Character difference identity}

For any $n \in [1;\infty)$ and $x_{n} = y_{n}$ then:
%TG: Is that valid for a demonstration?
\begin{equation}
    \label{identitydemo}
    \begin{aligned}
    \abs{x_{n}-y_{n}} &= 0\\
    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= 0\\
    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= 0\\
    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \frac{1}{2}\\
    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= 1\\
    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 0\\
    \end{aligned}
\end{equation}

Thus equation \ref{Identity} is true.

\subsection{Character difference equality}

For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$, similarly to equation \ref{positive_demo}:

\begin{equation}
    \label{equality_demo}
    \begin{aligned}
    \abs{x_{n}-y_{n}} &= \abs{y_{n}-x_{n}}\\
    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= \sum_{i}^{n}\abs{y_{i}-x_{i}}\\
    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= \frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}\\
    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\\
    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= \frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\\
    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)\\
    \end{aligned}
\end{equation}

Thus equation \ref{Equality} is true.

\subsection{Character difference subadditivity}

For any $n, j, b \in [1;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:

\begin{equation}
    \label{subadditivity_demo}
    \begin{aligned}
    CD_{x,y} &\leq CD_{x,z} + CD_{z,y}\\
    \text{if:}\\
    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
     &\leq 
     1 - \left(\frac{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}}}{\frac{1}{2}}\right) 
     +
     1 - \left(\frac{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}}}{\frac{1}{2}}\right) \\
     \text{then, the opposite is:}\\
    \underbrace{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} }_{\in[0,1/2]}
     &\geq 
     \underbrace{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}} }_{\in[0,1/2]}
     +
     \underbrace{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}} }_{\in[0,1/2]} - 2\\
     \text{thus:}\\
     \in[0,\frac{1}{2}] &\geq \in[0,\frac{1}{2}] \in[0,\frac{1}{2}] (-2) \\
     \in[0,\frac{1}{2}] &\geq \in[-2,-1]\\
    \end{aligned}
\end{equation}

Thus equation \ref{Subadditivity} is true.

\end{document}

### 2017-04-25 Writing/supplementary/Sup2-SimParam.tex
Simulation parameters sup. mat

--- 2017-04-25 Writing/draft.tex
@@
 \bigskip
+\noindent{\Large \bf Supplementary material 2 - Simulation parameters}
 
@@
 \noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\
-\vspace{1in}
 
 %Line numbering
-\modulolinenumbers[1]
-\linenumbers
+% \modulolinenumbers[1]
+% \linenumbers
 
-%---------------------------------------------
-%
-%       ABSTRACT
-%
-%---------------------------------------------
+\section{Morphological matrices simulations}
 
-\newpage
-\begin{abstract}
-blablabla
-\end{abstract}
-
-\noindent (Keywords: )\\
-
-\vspace{1.5in}
-
-\newpage
-
-%---------------------------------------------
-% LaTeX tips for modifying/editing the document:
-%---------------------------------------------
-% - You can comment using the percentage sign. I suggest you use the % sign alone for commenting out sections of the text:
-%       e.g. "This is a really long sentence %because this sentence is very long." Here the % is used for ignoring the end of the sentence (but for some reason you want to keep track of it).%
-%       For comments as in verbose comments, I suggest you use "%MB:":
-%       e.g. "This is a really long sentence %because this sentence is very long. %MB: yeah, no shit!" 
-% - For optimal version control, write only one sentence per line (for more precise track changes)
-% - To build the pdf, use command+B in Sublime.
-% - Because of the bibliography, the pdf needs to be build in the same folder that contains "References.bib" and "sysbio.bst"
-% - For citing papers, you must put their bibtex reference in the "References.bib" file and then you can use the following sysbio tags:
-%        \cite{bibtexBob2000} for citing within a sentence: "Bob (2000)"
-%        \citep{bibtexBob2000} for citing within brackets: "(Bob, 2000)"
-%        \citep[Before:][-After]{bibtexBob2000} for citing within brackets with additional text: "(Before: Bob, 2000 -After)"
-%        \citealt{bibtexBob2000} for citing without brackets: "Bob, 2000"
-%        You can put more cites in each \cite tag by separating them with commas.
-% - For equation, find every details here: https://en.wikibooks.org/wiki/LaTeX/Mathematics
-% - For titles and stuff, the hierarchy goes \section{}, \subsection{}, \subsubsection{} and so forth...
-% - For bullet points or enumerations you can use:
-%       \begin{itemize}
-%           \item my first bullet point/enumeration
-%       \end{itemize}
-%       With replacing "itemize" by "enumerate" for enumeration.
-
-%---------------------------------------------
-%
-%       INTRODUCTION
-%
-%---------------------------------------------
-\section{Introduction}
-
-The last five years have seen a drastic expansion of the use of discrete morphological data in macroevolutionary studies.
-In fact, traditionally confined to ``classic'' cladistics \citep{felsenstein2004inferring}, these methods have recently been expended and improved whether based on the models and methods for inferring phylogenies \citep{lewisa2001,Wright01072016} or the methods to date them \citep[e.g.][]{bapsta2013} or to combine them with molecular data \citep{pyrondivergence2011,ronquista2012}.
-% Debate on which inference method to use \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081}.
-% Debate on dating methods \citep{Arcila2015131}.
-% Debate on data sources (BAPST in review)
-In parallel, much development has also been done on pre-phylogenetic analysis \citep[e.g.
-data collection;][]{morphobank} as well as on post-phylogenetic analysis \citep[e.g.
-morphological disparity analysis;][]{Close2015,Claddis}.
-However, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146}(+BAPST); the coding of this data \citep{Brazeau2011}; the inference \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081} or the dating methods \citep{Arcila2015131}.
-
-However, one aspect of the data is often neglected: the correlation that exists between discrete morphological characters.
-In fact, because of their nature, these characters are prone to many sources of correlation such as: (1) the co-evolution between traits (e.g.
-the number of digits on a limb co-evolves with the actual presence/absence of the limb); (2) morphological modularity \citep[i.e.
-the unity of entire morphological regions][]{goswami2006morphological}; (3) convergent evolution between traits (e.g.
-wings in many vertebrates); or (4), more down-to-earth, to errors from the coder \citep[e.g.
-the repetition of the same character described slighlty differently each case][]{Brazeau2011}.
-These sources of correlation can alter the results of phylogenetic inference by effectively weighting the correlated characters against their non-correlated counter parts leading to biases in phylogenetic inference by supporting clades with artificial synapomorphies.
-This fact have led to the popular wisdom rule of avoiding coding correlated characters for macroevolutionary studies.
-
-Up to this date, a lot of work has been done on character correlation \textit{per se}, but focussing mainly on how to measure correlation between characters to detect patterns of coevolution \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
-However, only several attempts have been made on the actual effect of characters correlation on phylogenetic inference with two notable exceptions of \cite{Davalos01072014} on empirical data and \cite{ZouConvergence} on molecular data.
-It is thus still unclear what will be the effect of characters correlation on phylogenetic inference: how does these correlation really affect topology?
-Additionally, because of the nature of discrete morphological characters and organisms morphology, correlation is expected to rise through both combinatorics properties of the characters or simply through evolutionary processes.
-What is the effect of these two sources of correlations and can we (should we) avoid them?
-
-In this study, we formally assess the effect of discrete character's correlation using simulated and empirical data.
-We propose a new metric to measure the correlation (or difference) between two discrete characters and a protocol to modify discrete data matrices to increase/decrease the correlation between characters within/
-We found that blabalbal
-
-\section{Methods}
-
-We applied the following simulation protocol to assess the effect of characters correlation (Fig.\ref{Fig:outline}):
-\begin{enumerate}
-    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
-    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by duplicating characters in order to maximise and minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
-We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximised or minimised character difference to create null expectancy matrices (hereafter called the ``null'' matrices).
-    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``null'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``null'' trees).
-    \item \textbf{Comparing topologies}: finally, we compared to topologies in two ways: (1) first we compared the ``normal'' to the ``null'' trees to assess our simulation protocol; (2) second, we compared the ``maximised'' and ``minimised'' trees to the ``null'' and ``normal'' trees respectively to measure the signal to noise ratio and thus measure the effect of character correlation on topology.
-\end{enumerate}
-Each steps are described below in more details, as well as our proposed definition of the Character Difference metric.
-
-\begin{figure}[!htbp]
-\centering
-   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
-\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using Maximum Parsimony and Bayesian inference (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).}
-\label{Fig:outline}
-\end{figure}
-
-\subsection{Measuring correlation between characters}
-We define characters as being correlated if they give the same phylogenetic information.
-In order to measure this similitude in phylogenetic information, we propose a new distance metric to measure the difference between two characters:
-
-\subsubsection{Characters difference (CD)}
-\begin{equation}
-    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
-\end{equation}
-
-\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
-$CD$ is a continuous distance metric valid in a mathematical sense (see Supplementary).
-Since we are considering differences as being only Fitch-like (i.e.
-non-weighted and non-ordinate), we ignore the actual distance between two different character states tokens \citep[i.e.
-a Gower distance;][]{GowerDist}.
-Two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
-In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
-In other words, we replaced all the occurrences of the first token to be $0$, the second to be $1$, etc.
-This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{0,0,1,2,2,1\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
-Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
-When the character difference is equal to $0$, it means that characters convey the same phylogenetic signal.
-When the character difference is equal to $1$ it means it conveys the most different signal.
-For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
-
-\subsection{Simulating discrete morphological matrices}
-To simulate the matrices we applied a protocol very similar to \citep{Guillerme2016146}.
-First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform ($0$,$1$) distribution maintaining $\lambda$ $>$ $\mu$ using the diversitree \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
-For each tree, we arbitrarily set the outgroup to be the first taxa (alphabetically) thus effectively rooting the trees on this taxa.
-These trees are hereafter called the ``true'' trees.
-We then simulated discrete morphological characters using the either of the two following models:
+\subsection{``True'' Tree simulation - using diversitree \texttt{R} package v 0.9-8 \cite{fitzjohndiversitree2012}}
 \begin{itemize}
-    \item The HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ an scale to sum to one) and using a transition/transvertion rate of two \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
-    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][; see below]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
-    \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
-    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146}.
-    This model assumes a equal transition rate between character states that might seem overly simplifying and excludes other observed transition patterns \citep[e.g.
-Dollo characters;][]{Dollo,wright2015came}.
-    Recently however, \cite{Wright01072016} have shown that an equal rate transition is still strictly majority in empirical data.
+  \item Model
+  \begin{itemize}
+    \item Birth-death tree (\texttt{diversitree::tree.bd})
+  \end{itemize}
+  \item Parameters
+  \begin{itemize}
+    \item Birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform distribution (0,1) with $\lambda > \mu$.
+    \item Stopped when the tree reached 25, 75 or 150 taxa (extinct taxa excluded).
+  \end{itemize}
 \end{itemize}
 
-\noindent For each character, both models (HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
-This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices (see supplementary materials @@@).
-We re-simulated every invariant characters to obtain a matrix with no invariant characters.
-To ensure that our simulation where reflecting realistic observed parameters, we only selected matrices with Consistency Indices superior to $0.26$ \citep{OReilly20160081}.
+\subsection{``Normal'' matrix simulation - using dispRity \texttt{R} package v 0.4 \cite{thomas_guillerme_2016_55646}}
+\begin{itemize}
+  \item Model
+  \begin{itemize}
+    \item Mk or HKY-binary selected randomly (\texttt{dispRity::sim.morpho})
+  \end{itemize}
+  \item Parameters
+  \begin{itemize}
+    \item States frequency: 0.85 binary characters and 0.15 three states (for Mk only)
+    \item Transition/transversion ratio (for HKY-binary): 2 
+    \item Character rate (for both models): gamma distribution with a rate $\alpha$ = 100 and a shape $\beta$ = 5.
+    \item No invariant characters.
+  \end{itemize}
+\end{itemize}
 
-For each trees with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
-The matrices where generated using the \texttt{dispRity R} package \citep[][; \url{https://github.com/TGuillerme/dispRity}]{thomas_guillerme_2016_55646}.
-To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 50 times resulting in 450 morphological matrices (hereafter called the ``normal'' matrices).
+\subsection{``Maximised'', ``Minimised'' and ``Null'' matrices}
+\begin{itemize}
+  \item Maximised
+  \begin{itemize}
+    \item Replacing the $n$ characters with a character difference (CD) to all other characters $< 0.25$.
+  \end{itemize}
 
-\subsection{Modifying the matrices}
+  \item Minimised
+  \begin{itemize}
+    \item Replacing the $m$ character with a CD to all other characters $> 0.75$.
+  \end{itemize}
 
-We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}. %TG: TO IMPLEMENT! 
-We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
-For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e.
-with an average character difference $<$$0.25$) and replaced them randomly by any of the remaining characters.
-This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
-Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e.
-with an average character difference $<$$0.75$) and randomly replaced them with the remaining ones.
-Finally, because this operation effectively changes the weight of characters (i.e.
-giving the characters $<$$0.25$ or $>$$0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a null expectation modification (i.e.
-randomly weighting characters).
-This step resulted in a total of 1800 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``null'' matrices).
-The algorithms for the three modifications are available on GitHub@@@. %TG: "@@@" is my "to replace" sign.
+  \item Null
+  \begin{itemize}
+    \item Randomly replacing $\frac{n+m}{2}$ characters.
+  \end{itemize}
 
-\subsection{Inferring topologies}
+\end{itemize}
 
-We inferred the topologies in both Bayesian and Maximum Parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
-The Maximum Parsimony inference was run over @@@ CPU hours using a heuristic search with random sequence addition replicate 10 times (\texttt{hsearch addseq=random nreps=10}).
-The trees where also bootstrapped 100 times using the same search parameters mentioned above.
+\section{Tree inference parameters}
 
-The Bayesian inference was run over @@@ CPU hours
+\subsection{Maximum parsimony: PAUP* version 4.0a151 \cite{swofford2001paup}}
 
-% lset nst=1 rates=gamma Ngammacat=4;
-% prset ratepr=variable Shapepr=Exponential(0.5);
+\begin{itemize}
+  \item Search settings
+  \begin{itemize}
+    \item Max. trees = 500 with auto-increase of 500.
+    \item Heuristic search with random sequence addition replicated 10 times
+  \end{itemize}
+  \item Summary settings
+  \begin{itemize}
+    \item Summarising up to $1\times10^9$ trees in a strict majority rule consensus tree (cut off = 50).
+  \end{itemize}
+\end{itemize}
 
-% mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES file=150t_100c_0304_maxi;
-% sump Filename=150t_100c_0304_maxi Relburnin=YES Burninfrac=0.25;
-% sumt Filename=150t_100c_0304_maxi Relburnin=YES Burninfrac=0.25;
+\subsection{Bayesian: MrBayes version 3.2.7 \cite{Ronquist2012mrbayes}}
 
-\subsection{Comparing topologies}
+\begin{itemize}
+  \item Models
+  \begin{itemize}
+    \item M\textit{kv} + $\Gamma_4$
+  \end{itemize}
 
-We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``null'' matrices and the tree inferred from the ``normal'' matrix.
+  \item Priors
+  \begin{itemize}
+    \item variable rate prior with an exponential shape prior set to 0.5
+  \end{itemize}
 
-We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the different inferred trees.
-We ran two types of comparisons, (1) the ability to recover the correct tree (``normal'' trees \textit{vs} ``null'' and ``true'' trees and (2) the effect of character differences by measuring the noise to signal ratio between the ``maximised'' and ``minimised'' trees and the ``null'' and ``normal'' trees.
+  \item MCMC
+  \begin{itemize}
+    \item Two runs
+    \item Six chains per run
+    \item Generations $<$ $1\times10^9$
+    \item Sample frequency = 200
+    \item ASDS diagnosis frequency = 10000
+    \item ASDS $<$ 0.01
+    \item ESS $>>$ 200
+  \end{itemize}
 
-The metrics scores where obtained using the TreeCmp java script \citep{Bogdanowicz2012}.
-The measurements where then standardised using the Normalised Tree Similarity metric \citep[i.e.
-centering the metrics scores using the mean metric score for 1000 pairwise comparisons between random trees with $n$ taxa;][]{Bogdanowicz2012,Guillerme2016146}.
-When the normalised metric has a score of $1$ it means both trees are identical, when it has a score of $0$ it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
-The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Fould Similarity reflects the conservation of clades (i.e.
-a score close to $1$ indicates that most clades are identical in both trees); and (2) the Normalised Triplets Similarity reflects the position of taxa (i.e.
-a score close to $1$ indicates that most taxa have the same neighbours in both trees).
+  \item Summary settings
+  \begin{itemize}
+    \item Strict majority consensus tree (cut off = 50) with a relative burning of 25\%.
+  \end{itemize}
 
-\subsection{Empirical morphological matrices}
+\end{itemize}
 
-We applied the same procedure described above (without the morphological matrices simulation) to 100 empirical matrices downloaded from TreeBASE (\url{http://treebase.org/}).
-The 100 matrices where selected to be the most representative of empirical work with discrete morphological data: with randomly selected matrices with more than 100 each, published between 1985 and 2013 and covering 19 taxonomic classes within 6 phyla (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
-The full list of matrices is available in the supplementary materials @@@.
-
-\section{Results}
-
-\section{Discussion}
-
-\subsection{Conclusion}
-
-\section{Funding}
-European Research Council under the European Union’s Seventh Framework Programme (FP/2007–2013)/ERC Grant Agreement number 311092.
-
-\section{Acknowledgements}
-Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
-We thank Alberto Pascual Garcia and April Wright for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD proof.
-
-% Suggested reviewers:
-%- Liliana Davalos for MPE
-
-\bibliographystyle{sysbio}
-\bibliography{References}
+\bibliographystyle{../sysbio}
+\bibliography{../References}
 

### 2017-05-18 Functions/shells/set.runs.sh
PAUP: added rearrangement limit to 5M but increased reps to 500

@@ -129,7 +129,7 @@ then
 
     echo "" >> base_ppcmd.tmp
     echo "[Parsimony search]" >> base_ppcmd.tmp
-    echo "hsearch addseq=random nreps=100 rseed=01234;" >> base_ppcmd.tmp
+    echo "hsearch addseq=random nreps=500 rseed=01234 rearrlimit=5000000;" >> base_ppcmd.tmp #rearrlimit=XXXX
     # echo "bootstrap bseed=12345 nreps=100 ConLevel=50 KeepAll=yes TreeFile = <CHAIN>.bs;" >> base_ppcmd.tmp
 
 

### 2017-05-29 Analysis/Comparing.trees.R
Cleaned analysis folder

@@ -1,157 +1,0 @@
-## Analysis for comparing the trees
-
-## Load the functions
-source("functions.R") ; load.functions(test = FALSE)
-dyn.load("../Functions/char.diff.so")
-
-############
-## Load the tree comparisons
-###########
-current <- getwd()
-
-## ~~~~
-## Bayesian
-## ~~~~
-
-## Normal trees
-setwd("../Data/Trees_out/Single_trees/Bayesian/Consensus_trees/")
-bay_25t_100c <- read.TreeCmp("25t_100c")
-bay_25t_350c <- read.TreeCmp("25t_350c")
-bay_25t_1000c <- read.TreeCmp("25t_1000c")
-
-# bay_75t_100c <- read.TreeCmp("75t_100c")
-# bay_75t_350c <- read.TreeCmp("75t_350c")
-# bay_75t_1000c <- read.TreeCmp("75t_1000c")
-
-# bay_150t_100c <- read.TreeCmp("150t_100c")
-# bay_150t_350c <- read.TreeCmp("150t_350c")
-# bay_150t_1000c <- read.TreeCmp("150t_1000c")
-setwd(current)
-
-## True trees
-setwd("../Data/Trees_out/True_trees/Bayesian/")
-bay_true_25t_100c <- read.TreeCmp("25t_100c", true = TRUE)
-bay_true_25t_350c <- read.TreeCmp("25t_350c", true = TRUE)
-bay_true_25t_1000c <- read.TreeCmp("25t_1000c", true = TRUE)
-
-# bay_true_75t_100c <- read.TreeCmp("75t_100c", true = TRUE)
-# bay_true_75t_350c <- read.TreeCmp("75t_350c", true = TRUE)
-# bay_true_75t_1000c <- read.TreeCmp("75t_1000c", true = TRUE)
-
-# bay_true_150t_100c <- read.TreeCmp("150t_100c", true = TRUE)
-# bay_true_150t_350c <- read.TreeCmp("150t_350c", true = TRUE)
-# bay_true_150t_1000c <- read.TreeCmp("150t_1000c", true = TRUE)
-setwd(current)
-
-## ~~~~
-## Parsimony
-## ~~~~
-
-## Normal trees
-setwd("../Data/Trees_out/Single_trees/Parsimony/Consensus_trees/")
-par_25t_100c <- read.TreeCmp("25t_100c")
[…]

### 2017-05-29 Analysis/FlippingData.R

@@ -1,2 +1,0 @@
-## What is the difference in information between the characters and the taxa in terms of disparity?
-

### 2017-05-29 Analysis/Sandbox.R

@@ -1,90 +1,0 @@
-## Packages and functions
-library(Claddis)
-library(dispRity)
-library(diversitree)
-source("functions.R") ; load.functions(test = FALSE)
-dyn.load("../Functions/char.diff.so")
-write.nexus.std <- ape::write.nexus.data
-body(write.nexus.std)[[2]] <- substitute(format <- match.arg(toupper(format), c("DNA", "PROTEIN", "STANDARD")))
-body(write.nexus.std)[[26]][[3]][[4]] <- substitute(fcat(indent, "FORMAT", " ", DATATYPE, " ", MISSING, " ", GAP, 
-    " ", INTERLEAVE, " symbols=\"0123456789\";\n"))
-get.digit <- function(column) {
-    if(max(nchar(round(column)), na.rm = TRUE) <= 4) {
-       return(4-max(nchar(round(column)), na.rm = TRUE))
-    } else {
-        return(0)
-    }
-}
-sample.birth.death <- function() {
-    birth <- runif(min = 0, max = 1, 1)
-    death <- runif(min = 0, max = birth, 1)
-    return(c(birth, death))
-}
-
-tree.wrapper <- function(ntaxa) {
-    tree <- NULL
-    birth_death <- sample.birth.death()
-    while(is.null(tree)) {
-        tree <- tree.bd(birth_death, max.taxa = ntaxa)
-    }
-    tree$node.label <- NULL
-    tree$tip.label[[1]] <- "outgroup"
-    tree <- root(tree, "outgroup", resolve.root = TRUE)
-
-    return(tree)
-}
-
-morpho.wrapper <- function(ntaxa, ncharacters, gamma) {
-    tree <- tree.wrapper(ntaxa)
-    cat(".")
-    return(check.morpho(sim.morpho(tree, ncharacters, states = c(0.85,0.15), model = "mixed", rates = gamma, invariant = FALSE), tree))
-}
-
-## Testing parameter (CI distribution)
-gamma <- c(rgamma, rate = 100, shape = 3)
-runs = 20
-
-cat("25t_100c:")
-t25_c100 <- replicate(runs, morpho.wrapper(25, 100, gamma), simplify = FALSE)
-cat("Done.\n")
-cat("25t_350c:")
[…]

### 2017-05-29 Analysis/character.modularity.R

@@ -1,10 +1,0 @@
-## Checking network structure between characters!
-
-## Loading the functions/libraries
-library(Claddis)
-library(reshape2)
-source("functions.R") ; load.functions(test = FALSE)
-dyn.load("../Functions/char.diff.so")
-
-
-export.ntw("../Data/Network/Coates et al Dwykaselachus.nex")

### 2017-05-29 Analysis/utilities/Getting.NTS.R

@@ -1,12 +1,12 @@
 ## Analysis for comparing the trees
 
 ## Load the functions
-source("functions.R") ; load.functions(test = FALSE)
-dyn.load("../Functions/char.diff.so")
+source("../functions.R") ; load.functions(test = FALSE)
+dyn.load("../../Functions/char.diff.so")
 
 ## Loading the tree comparisons
 current <- getwd()
-setwd("../Data/NTS/Comparisons/")
+setwd("../../Data/NTS/Comparisons/")
 nts_25t <- read.table("25t_NTS.Cmp", sep = "\t", header = TRUE)
 nts_75t <- read.table("75t_NTS.Cmp", sep = "\t", header = TRUE)
 nts_150t <- read.table("150t_NTS.Cmp", sep = "\t", header = TRUE)
@@ -19,4 +19,4 @@ mean_150t <- apply(nts_150t[,4:7], 2, mean)
 
 ## Saving the NTS data as a table
 matrix_nts <- matrix(c(mean_25t, mean_75t, mean_150t), ncol = 4, byrow = TRUE, dimnames = list(c(25, 75, 150), names(mean_25t)))
-save(matrix_nts, file = "../Data/NTS/matrix.nts.Rda")
+save(matrix_nts, file = "../../Data/NTS/matrix.nts.Rda")

### 2017-05-29 Functions/NTS.R
Streamlined pipeline functions

@@ -37,5 +37,5 @@ NTS <- function(data, taxa) {
     rand <- matrix_nts[which(rownames(matrix_nts) == taxa), ]
 
     ## Calculate the NTS
-    return(lapply(data, apply.NTS, rand))
+    return(lapply(data, lapply, apply.NTS, rand))
 }

### 2017-05-29 Functions/plot.results.R

@@ -1,39 +1,177 @@
-plot.results.single <- function(data_list, metric, main, legend = FALSE) {
+#' @title Plotting TreeCmp results
+#'
+#' @description Simple wrapper for plotting the results from TreeCmp
+#'
+#' @param taxa_list a list of tree metrics from \code{read.TreeCmp} for all the scenario for a number of taxa
+#' @param metric a numeric of which metric to use (\code{1} = \code{MatchingCluster}, \code{2} = \code{R.F_Cluster}, \code{3} = \code{NodalSplitted}, \code{4} = \code{Triples})
+#' @param what which type of plot? Can be \code{"null"}, \code{"real"}, \code{"true"}, \code{"check"}. See details.
+#' @param ylim \code{ylim} parameter from \code{plot} (if missing is automatic)
+#' @param NTS \code{logical}, whether NTS was used
+#'
+#' @details
+#' \code{"null"} = "maxi", "mini", "norm" tree vs. "rand"
+#' \code{"best"} = "maxi", "mini", "rand" tree vs. "norm"
+#' \code{"true"} = "norm", "rand" tree vs. "true"
+#' \code{"check"} = "norm", vs "rand"
+#' 
+#' @return
+#' A list of metrics for the comparisons.
+#' 
+#' @examples
+#' ##
+#' 
+#' @author Thomas Guillerme
+#' @export
+#' 
+
+
+plot.results.single <- function(taxa_list, metric, what, ylim, main, legend = FALSE, NTS = TRUE) {
     
     ## Graphic parameters
+    ## Colours
     cols <- c("red", "orange", "green3", "lightgreen", "blue", "lightblue")
-    ylab <- ifelse(metric == 2, "Similarity (Robinson-Foulds)", "Similarity (Triplets)")
-    ylim <- c(0,1)
-    ylim[1] <- ifelse(metric == 2, 0, -0.5)
-    line <- ifelse(metric == 2, -1, 0)
-    par(bty = "n")
+    
+    ## Y lab
+    if(metric == 1) ylab <- "Similarity (Matching Cluster)"
+    if(metric == 2) ylab <- "Similarity (Robinson-Foulds)"
+    if(metric == 3) ylab <- "Similarity (Nodal Split)"
+    if(metric == 4) ylab <- "Similarity (Triplets)"
 
-    ## Boxplots
-    boxplot(xlab = "Characters", ylab = ylab, xaxt = "n", main = main, ylim = ylim,
-        ## 100c
[…]

### 2017-05-29 Functions/read.TreeCmp.R

@@ -3,6 +3,7 @@
 #' @description Read in as a list the results from the different tree comparisons
 #'
 #' @param chain the general name of the chain (e.g. 25t_100c)
+#' @param path the path to where the files are stored
 #'
 #' @return
 #' A list of metrics for the comparisons.
@@ -14,54 +15,88 @@
 #' @export
 #' 
 
-read.TreeCmp <- function(chain, true = FALSE) {
+#path_alt <- "/Users/TGuillerme/Projects/CharactersCorrelation/Data/Trees_out/Consensus_trees/Parsimony"
+
+read.TreeCmp <- function(chain, path = "/Users/TGuillerme/Projects/CharactersCorrelation/Data/Trees_out/Consensus_trees", type = "Bayesian") {
+
+    path <- paste(path, type, sep = "/")
+
     ## Sanitizing
     if(class(chain) !='character') {
         stop('No files has been found within the given chain name.')
     }
 
     ## Get the different files within the path
-    files <- list.files(paste(chain, "_cmp/", sep =""))
-    if(length(files) == 0) {
-        stop('No files has been found within the given chain name.')
+    files_norm <- list.files(paste(paste(path, chain, sep = "/"), "_cmp_norm/", sep =""))
+    if(length(files_norm) == 0) {
+        stop('No files for norm has been found within the given chain name.')
     }
 
+    files_rand <- list.files(paste(paste(path, chain, sep = "/"), "_cmp_rand/", sep =""))
+    if(length(files_rand) == 0) {
+        stop('No files for rand has been found within the given chain name.')
+    }
+
+
     ## Load all the files
-    matrices <- list()
-    message("Reading the TreeCmp results:", appendLF = FALSE)
-    for(file in 1:length(files)) {
-        matrices[[file]] <- as.data.frame(read.table(paste(chain, "_cmp/",files[[file]], sep = ""), header = TRUE, row.names = 1))
+    matrices_norm <- list()
+    message("Reading the TreeCmp results (norm):", appendLF = FALSE)
+    for(file in 1:length(files_norm)) {
[…]

### 2017-05-29 Analysis/EffectCorrelation.Rmd
Markdown analysis

---
title: "Effect of character correlation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Effect of character correlation
===============

Before starting
===============

Loading the functions
-----------------------------------

Getting the functions

```{r, eval = TRUE}
source("functions.R") ; load.functions(test = FALSE)
dyn.load("../Functions/char.diff.so")
load("../Data/NTS/matrix.nts.Rda")
```

Loading the data
-----------------------------------

Loading the tree comparisons from TreeCmp

```{r, message = FALSE}
## 25t_100c
bay_25t_100c <- read.TreeCmp("25t_100c", type = "Bayesian")
par_25t_100c <- read.TreeCmp("25t_100c", type = "Parsimony")

## 25t_350c
bay_25t_350c <- read.TreeCmp("25t_350c", type = "Bayesian")
par_25t_350c <- read.TreeCmp("25t_350c", type = "Parsimony")

## 25t_1000c
bay_25t_1000c <- read.TreeCmp("25t_1000c", type = "Bayesian")
par_25t_1000c <- read.TreeCmp("25t_1000c", type = "Parsimony")

# ## 75t_100c
# bay_75t_100c <- read.TreeCmp("75t_100c", type = "Bayesian")
# par_75t_100c <- read.TreeCmp("75t_100c", type = "Parsimony")

# ## 75t_350c
# bay_75t_350c <- read.TreeCmp("75t_350c", type = "Bayesian")
# par_75t_350c <- read.TreeCmp("75t_350c", type = "Parsimony")

# ## 75t_1000c
# bay_75t_1000c <- read.TreeCmp("75t_1000c", type = "Bayesian")
# par_75t_1000c <- read.TreeCmp("75t_1000c", type = "Parsimony")

```

Analysis the differences
===============

Normalise the score
-----------------------------------

Normalising the scores using NTS

```{r, message = FALSE}
## 25t_100c
bay_25t_100c <- NTS(bay_25t_100c, 25)
par_25t_100c <- NTS(par_25t_100c, 25)

## 25t_350c
bay_25t_350c <- NTS(bay_25t_350c, 25)
par_25t_350c <- NTS(par_25t_350c, 25)

## 25t_1000c
bay_25t_1000c <- NTS(bay_25t_1000c, 25)
par_25t_1000c <- NTS(par_25t_1000c, 25)

## Combining into a list
t25_list <- list(
    "c100" = list("bayesian" = bay_25t_100c,
                  "parsimony" = par_25t_100c
                  ),
    "c350" = list("bayesian" = bay_25t_350c,
                  "parsimony" = par_25t_350c
                  ),
    "c1000" = list("bayesian" = bay_25t_1000c,
                   "parsimony" = par_25t_1000c
                   )
                )
#

# ## 75t_100c
# bay_75t_100c <- NTS(bay_75t_100c, 75)
# par_75t_100c <- NTS(par_75t_100c, 75)

# ## 75t_350c
# bay_75t_350c <- NTS(bay_75t_350c, 75)
# par_75t_350c <- NTS(par_75t_350c, 75)

# ## 75t_1000c
# bay_75t_1000c <- NTS(bay_75t_1000c, 75)
# par_75t_1000c <- NTS(par_75t_1000c, 75)

```

Plotting the results

```{r, fig.width=12, fig.height=6}
## Comparisons to the best tree (RF)
plot.results.single(t25_list, metric = 2, what = "best",
    main = "comparison to 'best' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the best tree (Triplets)
plot.results.single(t25_list, metric = 4, what = "best",
    main = "comparison to 'best' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the null tree (RF)
plot.results.single(t25_list, metric = 2, what = "null",
    main = "comparison to 'null' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the null tree (Triplets)
plot.results.single(t25_list, metric = 4, what = "null",
    main = "comparison to 'null' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (RF)
plot.results.single(t25_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (Triplets)
plot.results.single(t25_list, metric = 4, what = "true",
    main = "comparison to 'true' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Checking comparisons (RF)
plot.results.single(t25_list, metric = 2, what = "check",
    main = "Check comparison (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Checking comparisons (Triplets)
plot.results.single(t25_list, metric = 4, what = "check",
    main = "Check comparison (25 taxa)", legend = TRUE, NTS = TRUE)
```

### 2017-06-05 Analysis/EffectCorrelation.Rmd
Pipelined analysis

@@
 
-# ## 75t_100c
-# bay_75t_100c <- read.TreeCmp("75t_100c", type = "Bayesian")
-# par_75t_100c <- read.TreeCmp("75t_100c", type = "Parsimony")
+## 75t_100c
+bay_75t_100c <- read.TreeCmp("75t_100c", type = "Bayesian")
+par_75t_100c <- read.TreeCmp("75t_100c", type = "Parsimony")
 
-# ## 75t_350c
-# bay_75t_350c <- read.TreeCmp("75t_350c", type = "Bayesian")
-# par_75t_350c <- read.TreeCmp("75t_350c", type = "Parsimony")
+## 75t_350c
+bay_75t_350c <- read.TreeCmp("75t_350c", type = "Bayesian")
+par_75t_350c <- read.TreeCmp("75t_350c", type = "Parsimony")
 
-# ## 75t_1000c
-# bay_75t_1000c <- read.TreeCmp("75t_1000c", type = "Bayesian")
-# par_75t_1000c <- read.TreeCmp("75t_1000c", type = "Parsimony")
+## 75t_1000c 
+bay_75t_1000c <- read.TreeCmp("75t_1000c", type = "Bayesian")
+par_75t_1000c <- read.TreeCmp("75t_1000c", type = "Parsimony")
+
+## 150t_100c
+bay_150t_100c <- read.TreeCmp("150t_100c", type = "Bayesian")
+par_150t_100c <- read.TreeCmp("150t_100c", type = "Parsimony")
+
+## 150t_350c
+bay_150t_350c <- read.TreeCmp("150t_350c", type = "Bayesian")
+par_150t_350c <- read.TreeCmp("150t_350c", type = "Parsimony")
+
+## 150t_1000c 
+bay_150t_1000c <- read.TreeCmp("150t_1000c", type = "Bayesian")
+par_150t_1000c <- read.TreeCmp("150t_1000c", type = "Parsimony")
 
@@
                 )
-#
 
-# ## 75t_100c
-# bay_75t_100c <- NTS(bay_75t_100c, 75)
-# par_75t_100c <- NTS(par_75t_100c, 75)
+## 75t_100c
+bay_75t_100c <- NTS(bay_75t_100c, 25)
+par_75t_100c <- NTS(par_75t_100c, 25)
 
-# ## 75t_350c
-# bay_75t_350c <- NTS(bay_75t_350c, 75)
-# par_75t_350c <- NTS(par_75t_350c, 75)
+## 75t_350c
+bay_75t_350c <- NTS(bay_75t_350c, 25)
+par_75t_350c <- NTS(par_75t_350c, 25)
 
-# ## 75t_1000c
-# bay_75t_1000c <- NTS(bay_75t_1000c, 75)
-# par_75t_1000c <- NTS(par_75t_1000c, 75)
+## 75t_1000c
+bay_75t_1000c <- NTS(bay_75t_1000c, 25)
+par_75t_1000c <- NTS(par_75t_1000c, 25)
+
+## Combining into a list
+t75_list <- list(
+    "c100" = list("bayesian" = bay_75t_100c,
+                  "parsimony" = par_75t_100c
+                  ),
+    "c350" = list("bayesian" = bay_75t_350c,
+                  "parsimony" = par_75t_350c
+                  ),
+    "c1000" = list("bayesian" = bay_75t_1000c,
+                   "parsimony" = par_75t_1000c
+                   )
+                )
+
+## 150t_100c
+bay_150t_100c <- NTS(bay_150t_100c, 25)
+par_150t_100c <- NTS(par_150t_100c, 25)
+
+## 150t_350c
+bay_150t_350c <- NTS(bay_150t_350c, 25)
+par_150t_350c <- NTS(par_150t_350c, 25)
+
+## 150t_1000c
+bay_150t_1000c <- NTS(bay_150t_1000c, 25)
+par_150t_1000c <- NTS(par_150t_1000c, 25)
+
+## Combining into a list
+t150_list <- list(
+    "c100" = list("bayesian" = bay_150t_100c,
+                  "parsimony" = par_150t_100c
+                  ),
+    "c350" = list("bayesian" = bay_150t_350c,
+                  "parsimony" = par_150t_350c
+                  ),
+    "c1000" = list("bayesian" = bay_150t_1000c,
+                   "parsimony" = par_150t_1000c
+                   )
+                )
 
@@
 
-Plotting the results
+Plotting the results for 25 taxa
 
@@
 
+Plotting the results for 75 taxa
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the best tree (RF)
+plot.results.single(t75_list, metric = 2, what = "best",
+    main = "comparison to 'best' tree (75 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the best tree (Triplets)
+plot.results.single(t75_list, metric = 4, what = "best",
+    main = "comparison to 'best' tree (75 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the null tree (RF)
+plot.results.single(t75_list, metric = 2, what = "null",
+    main = "comparison to 'null' tree (75 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the null tree (Triplets)
+plot.results.single(t75_list, metric = 4, what = "null",
+    main = "comparison to 'null' tree (75 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+Plotting the results for 150 taxa
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the best tree (RF)
+plot.results.single(t150_list, metric = 2, what = "best",
+    main = "comparison to 'best' tree (150 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the best tree (Triplets)
+plot.results.single(t150_list, metric = 4, what = "best",
+    main = "comparison to 'best' tree (150 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the null tree (RF)
+plot.results.single(t150_list, metric = 2, what = "null",
+    main = "comparison to 'null' tree (150 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the null tree (Triplets)
+plot.results.single(t150_list, metric = 4, what = "null",
+    main = "comparison to 'null' tree (150 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+Checking the results for 25 taxa
+
 ```{r, fig.width=12, fig.height=6}
@@
 ```
+
+Checking the results for 75 taxa
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the true tree (RF)
+plot.results.single(t75_list, metric = 2, what = "true",
+    main = "comparison to 'true' tree (75 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the true tree (Triplets)
+plot.results.single(t75_list, metric = 4, what = "true",
+    main = "comparison to 'true' tree (75 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Checking comparisons (RF)
+plot.results.single(t75_list, metric = 2, what = "check",
+    main = "Check comparison (75 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Checking comparisons (Triplets)
+plot.results.single(t75_list, metric = 4, what = "check",
+    main = "Check comparison (75 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+Checking the results for 150 taxa
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the true tree (RF)
+plot.results.single(t150_list, metric = 2, what = "true",
+    main = "comparison to 'true' tree (150 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Comparisons to the true tree (Triplets)
+plot.results.single(t150_list, metric = 4, what = "true",
+    main = "comparison to 'true' tree (150 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Checking comparisons (RF)
+plot.results.single(t150_list, metric = 2, what = "check",
+    main = "Check comparison (150 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, fig.width=12, fig.height=6}
+## Checking comparisons (Triplets)
+plot.results.single(t150_list, metric = 4, what = "check",
+    main = "Check comparison (150 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+Comparison to 'best' tree (RF normalised distance)
+
+```{r, fig.width=12, fig.height=12}
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 2, what = "best",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 2, what = "best",
+    ylab = "RF for 150 taxa", NTS = TRUE)
+par(op)
+```
+
+Comparison to 'best' tree (Triplets normalised distance)
+
+```{r, fig.width=12, fig.height=12}
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 4, what = "best",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 4, what = "best", ylab = "Tr for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 4, what = "best",
+    ylab = "Tr for 150 taxa", NTS = TRUE)
+par(op)
+```
+
+Comparison to 'null' tree (RF normalised distance)
+
+```{r, fig.width=12, fig.height=12}
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 2, what = "null",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 2, what = "null", ylab = "RF for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 2, what = "null",
+    ylab = "RF for 150 taxa", NTS = TRUE)
+par(op)
+```
+
+Comparison to 'null' tree (Triplets normalised distance)
+
+```{r, fig.width=12, fig.height=12}
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 4, what = "null",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 4, what = "null", ylab = "Tr for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 4, what = "null",
+    ylab = "Tr for 150 taxa", NTS = TRUE)
+par(op)
+```

### 2017-06-05 Functions/plot.results.R
Minor functions corrections

@@ -6,7 +6,12 @@
 #' @param metric a numeric of which metric to use (\code{1} = \code{MatchingCluster}, \code{2} = \code{R.F_Cluster}, \code{3} = \code{NodalSplitted}, \code{4} = \code{Triples})
 #' @param what which type of plot? Can be \code{"null"}, \code{"real"}, \code{"true"}, \code{"check"}. See details.
 #' @param ylim \code{ylim} parameter from \code{plot} (if missing is automatic)
+#' @param legend \code{logical}, whether to display the legend
 #' @param NTS \code{logical}, whether NTS was used
+#' @param ylab a \code{character} string to be passed to \code{boxplot(..., ylab)}
+#' @param xlab a \code{character} string to be passed to \code{boxplot(..., xlab)}
+#' @param axislab a \code{character} string to be passed to \code{axis(..., labels)}
+#' @param ... any additional argument to be passed to \code{boxplot}
 #'
 #' @details
 #' \code{"null"} = "maxi", "mini", "norm" tree vs. "rand"
@@ -25,17 +30,19 @@
 #' 
 
 
-plot.results.single <- function(taxa_list, metric, what, ylim, main, legend = FALSE, NTS = TRUE) {
+plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, NTS = TRUE, ylab, xlab, axislab, ...) {
     
     ## Graphic parameters
     ## Colours
     cols <- c("red", "orange", "green3", "lightgreen", "blue", "lightblue")
     
     ## Y lab
-    if(metric == 1) ylab <- "Similarity (Matching Cluster)"
-    if(metric == 2) ylab <- "Similarity (Robinson-Foulds)"
-    if(metric == 3) ylab <- "Similarity (Nodal Split)"
-    if(metric == 4) ylab <- "Similarity (Triplets)"
+    if(missing(ylab)) {
+        if(metric == 1) ylab <- "Similarity (Matching Cluster)"
+        if(metric == 2) ylab <- "Similarity (Robinson-Foulds)"
+        if(metric == 3) ylab <- "Similarity (Nodal Split)"
+        if(metric == 4) ylab <- "Similarity (Triplets)"
+    }
 
     ## Y lim
     get.min.list <- function(X, metric) {
[…]

### 2017-06-05 Functions/read.TreeCmp.R

@@ -89,7 +89,7 @@ read.TreeCmp <- function(chain, path = "/Users/TGuillerme/Projects/CharactersCor
     message(".", appendLF = FALSE)
 
     for(file in 2:length(matrices_rand)) {
-        norm_rand <- cbind(norm_norm, as.matrix(lapply(as.list(matrices_rand[[file]]), `[[`, 1)))
+        norm_rand <- cbind(norm_rand, as.matrix(lapply(as.list(matrices_rand[[file]]), `[[`, 1)))
         maxi_rand <- cbind(maxi_rand, as.matrix(lapply(as.list(matrices_rand[[file]]), `[[`, 2)))
         mini_rand <- cbind(mini_rand, as.matrix(lapply(as.list(matrices_rand[[file]]), `[[`, 3)))
         rand_rand <- cbind(rand_rand, as.matrix(lapply(as.list(matrices_rand[[file]]), `[[`, 4)))

### 2017-07-17 Analysis/EffectCorrelation.Rmd
Simplified vignette

@@
 
-Before starting
-===============
-
-Loading the functions
+Loading the functions and the data
 -----------------------------------
 
-Getting the functions
+Getting the functions and the data
 
-```{r, eval = TRUE}
+```{r, eval = TRUE, message = FALSE}
 source("functions.R") ; load.functions(test = FALSE)
@@
 load("../Data/NTS/matrix.nts.Rda")
+source("../Data/load.data.R")
 ```
 
-Loading the data
+Results
 -----------------------------------
 
-Loading the tree comparisons from TreeCmp
+Comparison to 'best' tree (RF normalised distance)
 
-```{r, message = FALSE}
-## 25t_100c
-bay_25t_100c <- read.TreeCmp("25t_100c", type = "Bayesian")
-par_25t_100c <- read.TreeCmp("25t_100c", type = "Parsimony")
-
-## 25t_350c
-bay_25t_350c <- read.TreeCmp("25t_350c", type = "Bayesian")
-par_25t_350c <- read.TreeCmp("25t_350c", type = "Parsimony")
-
-## 25t_1000c
-bay_25t_1000c <- read.TreeCmp("25t_1000c", type = "Bayesian")
-par_25t_1000c <- read.TreeCmp("25t_1000c", type = "Parsimony")
-
-## 75t_100c
-bay_75t_100c <- read.TreeCmp("75t_100c", type = "Bayesian")
-par_75t_100c <- read.TreeCmp("75t_100c", type = "Parsimony")
-
-## 75t_350c
-bay_75t_350c <- read.TreeCmp("75t_350c", type = "Bayesian")
-par_75t_350c <- read.TreeCmp("75t_350c", type = "Parsimony")
-
-## 75t_1000c 
-bay_75t_1000c <- read.TreeCmp("75t_1000c", type = "Bayesian")
-par_75t_1000c <- read.TreeCmp("75t_1000c", type = "Parsimony")
-
-## 150t_100c
-bay_150t_100c <- read.TreeCmp("150t_100c", type = "Bayesian")
-par_150t_100c <- read.TreeCmp("150t_100c", type = "Parsimony")
-
-## 150t_350c
-bay_150t_350c <- read.TreeCmp("150t_350c", type = "Bayesian")
-par_150t_350c <- read.TreeCmp("150t_350c", type = "Parsimony")
-
-## 150t_1000c 
-bay_150t_1000c <- read.TreeCmp("150t_1000c", type = "Bayesian")
-par_150t_1000c <- read.TreeCmp("150t_1000c", type = "Parsimony")
-
+```{r, fig.width=12, fig.height=12}
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 2, what = "best",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 2, what = "best",
+    ylab = "RF for 150 taxa", NTS = TRUE)
+par(op)
 ```
 
-Analysis the differences
-===============
+Comparison to 'best' tree (Triplets normalised distance)
 
-Normalise the score
------------------------------------
+```{r, fig.width=12, fig.height=12}
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 4, what = "best",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 4, what = "best", ylab = "Tr for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 4, what = "best",
+    ylab = "Tr for 150 taxa", NTS = TRUE)
+par(op)
+```
 
-Normalising the scores using NTS
+Comparison to 'null' tree (RF normalised distance)
 
-```{r, message = FALSE}
-## 25t_100c
-bay_25t_100c <- NTS(bay_25t_100c, 25)
-par_25t_100c <- NTS(par_25t_100c, 25)
+```{r, fig.width=12, fig.height=12}
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 2, what = "null",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 2, what = "null", ylab = "RF for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 2, what = "null",
+    ylab = "RF for 150 taxa", NTS = TRUE)
+par(op)
+```
 
-## 25t_350c
-bay_25t_350c <- NTS(bay_25t_350c, 25)
-par_25t_350c <- NTS(par_25t_350c, 25)
+Comparison to 'null' tree (Triplets normalised distance)
 
-## 25t_1000c
-bay_25t_1000c <- NTS(bay_25t_1000c, 25)
-par_25t_1000c <- NTS(par_25t_1000c, 25)
-
-## Combining into a list
-t25_list <- list(
-    "c100" = list("bayesian" = bay_25t_100c,
-                  "parsimony" = par_25t_100c
-                  ),
-    "c350" = list("bayesian" = bay_25t_350c,
-                  "parsimony" = par_25t_350c
-                  ),
-    "c1000" = list("bayesian" = bay_25t_1000c,
-                   "parsimony" = par_25t_1000c
-                   )
-                )
-
-## 75t_100c
-bay_75t_100c <- NTS(bay_75t_100c, 25)
-par_75t_100c <- NTS(par_75t_100c, 25)
-
-## 75t_350c
-bay_75t_350c <- NTS(bay_75t_350c, 25)
-par_75t_350c <- NTS(par_75t_350c, 25)
-
-## 75t_1000c
-bay_75t_1000c <- NTS(bay_75t_1000c, 25)
-par_75t_1000c <- NTS(par_75t_1000c, 25)
-
-## Combining into a list
-t75_list <- list(
-    "c100" = list("bayesian" = bay_75t_100c,
-                  "parsimony" = par_75t_100c
-                  ),
-    "c350" = list("bayesian" = bay_75t_350c,
-                  "parsimony" = par_75t_350c
-                  ),
-    "c1000" = list("bayesian" = bay_75t_1000c,
-                   "parsimony" = par_75t_1000c
-                   )
-                )
-
-## 150t_100c
-bay_150t_100c <- NTS(bay_150t_100c, 25)
-par_150t_100c <- NTS(par_150t_100c, 25)
-
-## 150t_350c
-bay_150t_350c <- NTS(bay_150t_350c, 25)
-par_150t_350c <- NTS(par_150t_350c, 25)
-
-## 150t_1000c
-bay_150t_1000c <- NTS(bay_150t_1000c, 25)
-par_150t_1000c <- NTS(par_150t_1000c, 25)
-
-## Combining into a list
-t150_list <- list(
-    "c100" = list("bayesian" = bay_150t_100c,
-                  "parsimony" = par_150t_100c
-                  ),
-    "c350" = list("bayesian" = bay_150t_350c,
-                  "parsimony" = par_150t_350c
-                  ),
-    "c1000" = list("bayesian" = bay_150t_1000c,
-                   "parsimony" = par_150t_1000c
-                   )
-                )
-
+```{r, fig.width=12, fig.height=12}
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 4, what = "null",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 4, what = "null", ylab = "Tr for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 4, what = "null",
+    ylab = "Tr for 150 taxa", NTS = TRUE)
+par(op)
 ```
@@
 ```
-
-Comparison to 'best' tree (RF normalised distance)
-
-```{r, fig.width=12, fig.height=12}
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
-plot.results.single(t25_list, metric = 2, what = "best",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
-    axislab = FALSE)
-plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
-    xlab = "", axislab = FALSE, NTS = TRUE)
-plot.results.single(t150_list, metric = 2, what = "best",
-    ylab = "RF for 150 taxa", NTS = TRUE)
-par(op)
-```
-
-Comparison to 'best' tree (Triplets normalised distance)
-
-```{r, fig.width=12, fig.height=12}
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
-plot.results.single(t25_list, metric = 4, what = "best",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
-    axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "best", ylab = "Tr for 75 taxa",
-    xlab = "", axislab = FALSE, NTS = TRUE)
-plot.results.single(t150_list, metric = 4, what = "best",
-    ylab = "Tr for 150 taxa", NTS = TRUE)
-par(op)
-```
-
-Comparison to 'null' tree (RF normalised distance)
-
-```{r, fig.width=12, fig.height=12}
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
-plot.results.single(t25_list, metric = 2, what = "null",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
-    axislab = FALSE)
-plot.results.single(t75_list, metric = 2, what = "null", ylab = "RF for 75 taxa",
-    xlab = "", axislab = FALSE, NTS = TRUE)
-plot.results.single(t150_list, metric = 2, what = "null",
-    ylab = "RF for 150 taxa", NTS = TRUE)
-par(op)
-```
-
-Comparison to 'null' tree (Triplets normalised distance)
-
-```{r, fig.width=12, fig.height=12}
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
-plot.results.single(t25_list, metric = 4, what = "null",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
-    axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "null", ylab = "Tr for 75 taxa",
-    xlab = "", axislab = FALSE, NTS = TRUE)
-plot.results.single(t150_list, metric = 4, what = "null",
-    ylab = "Tr for 150 taxa", NTS = TRUE)
-par(op)
-```

### 2017-07-17 Analysis/Character_difference/Comparing_characters.Rmd
Added formula in LaTeX

@@
 
-```{r, eval = FALSE}
-# The latest version of devtools
-if(!require(devtools)) install.packages("devtools")
-
-## The latest version of dispRity 
-install_github("TGuillerme/dispRity")
+```{r, eval = TRUE}
+library(dispRity) 
 ```
@@
 
-![equation](http://www.sciweavers.org/tex2img.php?eq=1%20-%20%5Cleft%28%20%5Cfrac%7B%5Cmid%20%7B%5Cfrac%7B%5Cabs%7B%5Csum_%7Bi%7D%5E%7Bn%7D%28%5Cmid%20x_%7Bi%7D%20-%20y_%7Bi%7D%5Cmid%20%29%7D%7D%20%7Bn%7D-%5Cfrac%7B1%7D%7B2%7D%7D%20%5Cmid%20%20%7D%7B%5Cfrac%7B1%7D%7B2%7D%7D%5Cright%29&bc=White&fc=Black&im=jpg&fs=12&ff=arev&edit=0)
-<!-- \begin{equation}
-    1 - \left( \frac{\mid {\frac{\abs{\sum_{i}^{n}(\mid x_{i} - y_{i}\mid )}} {n}-\frac{1}{2}} \mid  }{\frac{1}{2}}\right)
-\end{equation} -->
+$$1 - \left( \frac{\mid {\frac{\abs{\sum_{i}^{n}(\mid x_{i} - y_{i}\mid )}} {n}-\frac{1}{2}} \mid  }{\frac{1}{2}}\right)$$
 
@@
 The function to calculate the character difference is a modification of the `stats::dist` function's source code (in `C`):
-
-```{r}
-source("../../Functions/char.diff.R")
-dyn.load("../../Functions/char.diff.so")
-```
 
@@
 ```{r}
-library(dispRity) ; set.seed(123)
+set.seed(0)
 # A random tree (coalescent)

### 2017-07-17 Functions/NTS.R
Corrected functions

@@ -19,7 +19,7 @@ get.NTS <- function(obs, rand) {
 apply.NTS <- function(matrix, rand) {
     mat_out <- matrix
     for(col in 1:4) {
-        mat_out[,col] <- get.NTS(unlist(matrix[,col]), rand[col])
+        mat_out[,col] <- get.NTS(matrix[,col], rand[col])
     }
     return(mat_out)
 }

### 2017-07-17 Functions/read.TreeCmp.R

@@ -22,7 +22,7 @@ read.TreeCmp <- function(chain, path = "/Users/TGuillerme/Projects/CharactersCor
     path <- paste(path, type, sep = "/")
 
     ## Sanitizing
-    if(class(chain) !='character') {
+    if(class(chain) != 'character') {
         stop('No files has been found within the given chain name.')
     }
 
@@ -58,45 +58,22 @@ read.TreeCmp <- function(chain, path = "/Users/TGuillerme/Projects/CharactersCor
 
     ## Set up the list for the combined results
     message("Combining the TreeCmp results (norm):", appendLF = FALSE)
-    file = 1
-    norm_norm <- as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 1))
-    maxi_norm <- as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 2))
-    mini_norm <- as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 3))
-    rand_norm <- as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 4))
-    true_norm <- as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 5))
-    
-    message(".", appendLF = FALSE)
 
-    for(file in 2:length(matrices_norm)) {
-        norm_norm <- cbind(norm_norm, as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 1)))
-        maxi_norm <- cbind(maxi_norm, as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 2)))
-        mini_norm <- cbind(mini_norm, as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 3)))
-        rand_norm <- cbind(rand_norm, as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 4)))
-        true_norm <- cbind(true_norm, as.matrix(lapply(as.list(matrices_norm[[file]]), `[[`, 5)))
-        message(".", appendLF = FALSE)
+    flip.list <- function(matrices_list, ncol, nrow) {
+        splitted_matrices <- lapply(matrices_list, function (X) split(t(X), rep(1:ncol, each = nrow)))
+        names <- colnames(matrices_list[[1]])
+            
+        flipidy.flop <- function(element, splitted_matrices, nrow, names) {
[…]

### 2017-07-17 Functions/shells/make.con.tre.sh

@@ -10,7 +10,8 @@ echo "${chain}_${tree}"
 cd ${chain}/
 
 ## Check if the chains are finished
-if grep -q "end;" ${chain}_${tree}.run1.t
+tail_end=$(tail -1 ${chain}_${tree}.run1.t)
+if echo ${tail_end} | grep -q "end;"
 then
     run1_finished="TRUE"
 else
@@ -20,7 +21,8 @@ else
     echo "Done."
 fi
 
-if grep -q "end;" ${chain}_${tree}.run2.t
+tail_end=$(tail -1 ${chain}_${tree}.run2.t)
+if echo ${tail_end} | grep -q "end;"
 then
     run2_finished="TRUE"
 else
@@ -44,6 +46,7 @@ then
     ## Remove all the MrBayes commands and replace by sumt
     mbcmd_start=$(grep -n "begin mrbayes;" ${chain}_${tree} | sed 's/:begin mrbayes;//g')
     sed ''"$mbcmd_start"',$d' ${chain}_${tree} > tmp
+    cp ${chain}_${tree} ${chain}_${tree}.bkp
     mv tmp ${chain}_${tree}
     echo "begin mrbayes;" >> ${chain}_${tree}
     echo "set autoclose=yes nowarn=yes;" >> ${chain}_${tree}
@@ -80,6 +83,7 @@ else
         let "run1_length -=1"
         echo "Trimming ${chain}_${tree}.run1.t..."
         sed ''"$run2_length"','"$run1_length"'d' ${chain}_${tree}.run1.t > tmp
+        cp ${chain}_${tree}.run1.t ${chain}_${tree}.run1.bkp
         mv tmp ${chain}_${tree}.run1.t
         echo "Done."
     else
@@ -88,6 +92,7 @@ else
             let "run2_length -=1"
             echo "Trimming ${chain}_${tree}.run2.t..."
             sed ''"$run1_length"','"$run2_length"'d' ${chain}_${tree}.run2.t > tmp
+            cp ${chain}_${tree}.run2.t ${chain}_${tree}.run2.bkp
             mv tmp ${chain}_${tree}.run2.t
             echo "Done."
         fi
@@ -114,6 +119,12 @@ else
 
     ## Run the script
     mb ${chain}_${tree}
+
+    ## Remove the backup (if consensus tree exists)
+    if ls | grep -q ${chain}_${tree}.con.tre
+    then
+        rm ${chain}_${tree}*.bkp
+    fi
 
     ## Return to parent dir
     cd ..

### 2017-07-17 Writing/draft.tex
PAUP: added rearrangement limit to 5M but increased reps to 500
Updated draft

@@
 Up to this date, a lot of work has been done on character correlation \textit{per se}, but focussing mainly on how to measure correlation between characters to detect patterns of coevolution \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
+
+Also \cite{goswami2010}.
+
 However, only several attempts have been made on the actual effect of characters correlation on phylogenetic inference with two notable exceptions of \cite{Davalos01072014} on empirical data and \cite{ZouConvergence} on molecular data.
@@
 
+ % "hsearch addseq=random nreps=500 rseed=01234 rearrlimit=5000000;"
+
+Due to cluster hardware requirements an to save some time, when chains didn't converged and the runs exceeded 10GB, we aborted the MCMC and computed the consensus tree from the uncoverged chains.
+In practice, these chains got stuck at an ASDS around (but not below) 0.01.
+
 \subsection{Comparing topologies}
@@
 
-\subsection{Empirical morphological matrices}
+% \subsection{Empirical morphological matrices}
 
-We applied the same procedure described above (without the morphological matrices simulation) to 100 empirical matrices downloaded from TreeBASE (\url{http://treebase.org/}).
-The 100 matrices where selected to be the most representative of empirical work with discrete morphological data: with randomly selected matrices with more than 100 each, published between 1985 and 2013 and covering 19 taxonomic classes within 6 phyla (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
-The full list of matrices is available in the supplementary materials @@@.
+% We applied the same procedure described above (without the morphological matrices simulation) to 100 empirical matrices downloaded from TreeBASE (\url{http://treebase.org/}).
+% The 100 matrices where selected to be the most representative of empirical work with discrete morphological data: with randomly selected matrices with more than 100 each, published between 1985 and 2013 and covering 19 taxonomic classes within 6 phyla (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
+% The full list of matrices is available in the supplementary materials @@@.
 

### 2017-07-19 Writing/draft.tex
Updated draft (added discussion bits)

@@
 \medskip
-\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\ 
+\noindent{*\bf Corresponding author.} \textit{guillert@tcd.ie}\\ 
 \vspace{1in}
@@
 
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=0.9\textwidth]{Figures/RF_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
+\caption{Effect of Character Difference on recovering the ``best'' topology.
+The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
+The x axis represents the different Character Difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}
+\label{Fig:RF_results_best}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=0.9\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
+\caption{Effect of Character Difference on recovering the ``best'' topology.
+The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
+\label{Fig:Tr_results_best}
+\end{figure}
+
 \section{Discussion}
 
+\subsection{Effect of character differences on topology}
+As expected, reducing character difference reduces the ability of recovering the best tree topology since the ``assumption'' of character independence is ``violated''.
+Conversely increasing character difference increases the ability of recover the best tree topology.
+However this has not more effect of topology than simply bootstrapping the matrix.
+
+\subsection{Effect of character differences on the inference method}
+Scenarios where the characters are most different to each other could favor Bayesian over Parsimony.
+In fact, the only way to deal with it in parsimony analysis is to generate really homoplasic trees (since most character gives a different signal).
+On the other hand probabilistic methods can better deal with this problem by increasing branch length to accommodate more changes in characters between taxa.
+
+\subsection{Distinction between different character correlations}
+
+Throughout literature, the concept of character correlation (as a requirement for avoiding characters dependence in phylogenetic analysis) can vary along four major points based on the origin of the character correlation:
+\begin{itemize}
+    \item Biological induce correlation between characters: linkage through development or biological function (e.g. upper and lower teeth).
+    These links are readily understood through developmental biology or understanding of shared function between features.
+    This is linked to organisms morphological modules \citep[\textit{sensu}][]{goswami2006morphological}.
+
+    \item Evolutionary induced correlation between characters: linkage through shared evolutionary history (e.g. character co-evolving along branches).
+    These links are the direct cause of the biological correlation highlighted above.
+    In practice, however, these linkages can only be inferred based on a phylogeny and will thus be dependant on the phylogenetic methods used.
+    For example, if a series of characters co-evol along a branch, this can be due to actual accumulation of evolutionary changes along this branch (e.g.
+characters linked to volancy along the Chiroptera branch) but could also be the result of inference biases (e.g.
+when long branch attraction occurs).
+
+    \item Coding induced correlation between characters: linkage through coding similar characters on a feature that share a similar pattern.
+    These links can be due to researcher inadvertently defining a character twice or the researcher having to define as many characters as possible on a small feature due to the fragmentary nature of the fossil record (e.g.
+only isolated teeth are available to describe a group of taxa).
+
+    \item Software induced correlation between characters: this designates the way software actually interprets the differences between characters.
+    I.e.
+the vast majority of phylogenetic software ignore both the character's definition, the different states signification (simply treated as different/similar tokens).
+For example, if we have a good assumption that two binary characters coded for four taxa are not correlated (by a prior analysis of the correlations described above), if these two characters both coded `{0,0,0,1}', the software will treat them strictly identically.
+
+    How actually softwares do not care about the character’s definition and whether its in such or such module or arose from such or such evolutionary history (plus what the tokens mean to the researchers).
+It just looks whether the tokens are different or not between species per characters.
+
+\end{itemize}
+
+Although both biological and evolutionary induce correlation is what we're interested, unless they are studied explicitly (through evo-devo hypothesis testing for example), coding and software induced correlation are the correlations we end up having to deal with.
+In our study, we do not take into account biological correlation (because our characters are randomly simulated, we can not propose developmental or functional links between them), however, because they are simulated using evolutionary models based on Birth-Death trees, evolutionary induced correlation is present in our simulation (though not analysed specifically).
+The coding induce correlation is simulated through our protocol of maximising/minimising the Character Difference metric from the ``normal'' matrix.
+Finally, the software induced correlation is what is actually measured through these simulations.
+
 \subsection{Conclusion}
+Character differences as a proxy for character correlation is not a problem \textit{per se} as long as characters are not actively coded to be similar.
+On the other hand, actively selecting characters based to favor independence does not increase the ability of recovering tree topology more than bootstrapping the characters.
+
+\section{Data availability and repeatability}
+The consensus trees are available on figshare at \url{@@@}.
+The code for reproducing the simulation is available and explained at \url{https://github.com/TGuillerme/CharactersCorrelation}.
+The vignette for reproducing the analysis and the figure is available at \url{http://htmlpreview.github.io/?https://github.com/TGuillerme/CharactersCorrelation/Analysis/EffectCorrelation.html}.
 
@@
 Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
-We thank Alberto Pascual Garcia and April Wright for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD proof.
+We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD proof.
 
 % Suggested reviewers:
-%- Liliana Davalos for MPE
+%- Liliana Davalos for MPE / Joe O'Reilly, April Wright
 

### 2017-07-19 Functions/run.aov.R
run aov INIT

@@ -1,0 +1,48 @@
+#' @title Runs an anova on the results of the analysis
+#'
+#' @description Runs an anova on the results of the analysis
+#'
+#' @param data the global dataset (as a list)
+#' @param metric which metric to use
+#' @param best which best tree (rand or norm)
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+run.aov <- function(data, metric, best) {
+
+    get.metric <- function(list, metric) {
+        return(list[,metric])
+    }
+
+    lapply.recursive <- function(data, metric) {
+        return(lapply(data, lapply, lapply, lapply, get.metric, metric))
+    }
+
+    ## Set up which scenario to use
+    best <- ifelse(best == "norm", 1, 2)
+    if(best == 1) {
+        scenarios <- c(2,3,4)
+    } else {
+        scenarios <- c(1,2,3)
+    }
+
+    ## Getting the metric
+    if(metric == "RF") metric <- 2
+    if(metric == "Triples") metric <- 4
+
+    ## Getting all the data for one metric
+    NTS_all <- lapply(data, lapply.recursive, metric)
+
+    ## Extra the right base (norm and rand)
+    NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, best)
+
+    ## Extra the right scenarios (i.e. drop rand/norm and true)
+    NTS <- lapply(NTS_base, lapply, lapply, `[`, scenarios)
+
+    return(NTS)
+}

### 2017-07-20 Writing/draft.tex
Restructured introduction and global pitch

@@
 
-The last five years have seen a drastic expansion of the use of discrete morphological data in macroevolutionary studies.
-In fact, traditionally confined to ``classic'' cladistics \citep{felsenstein2004inferring}, these methods have recently been expended and improved whether based on the models and methods for inferring phylogenies \citep{lewisa2001,Wright01072016} or the methods to date them \citep[e.g.][]{bapsta2013} or to combine them with molecular data \citep{pyrondivergence2011,ronquista2012}.
-% Debate on which inference method to use \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081}.
-% Debate on dating methods \citep{Arcila2015131}.
-% Debate on data sources (BAPST in review)
+The last five years have seen a fantastic expansion of the use of discrete morphological data in macroevolutionary studies.
+In fact, traditionally confined to ``classic'' cladistic studies \citep{felsenstein2004inferring}, these methods have recently been widely expended and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep{pyrondivergence2011,ronquista2012}.
 In parallel, much development has also been done on pre-phylogenetic analysis \citep[e.g.
@@
 morphological disparity analysis;][]{Close2015,Claddis}.
-However, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146}(+BAPST); the coding of this data \citep{Brazeau2011}; the inference \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081} or the dating methods \citep{Arcila2015131}.
+Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011}; the inference \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.
 
-However, one aspect of the data is often neglected: the correlation that exists between discrete morphological characters.
-In fact, because of their nature, these characters are prone to many sources of correlation such as: (1) the co-evolution between traits (e.g.
-the number of digits on a limb co-evolves with the actual presence/absence of the limb); (2) morphological modularity \citep[i.e.
-the unity of entire morphological regions][]{goswami2006morphological}; (3) convergent evolution between traits (e.g.
-wings in many vertebrates); or (4), more down-to-earth, to errors from the coder \citep[e.g.
-the repetition of the same character described slighlty differently each case][]{Brazeau2011}.
-These sources of correlation can alter the results of phylogenetic inference by effectively weighting the correlated characters against their non-correlated counter parts leading to biases in phylogenetic inference by supporting clades with artificial synapomorphies.
-This fact have led to the popular wisdom rule of avoiding coding correlated characters for macroevolutionary studies.
+One aspect of the data, however, is often neglected: the correlation between morphological characters.
+In fact, the assumption of character independence is at the base of phylogenetic inference and as been mentioned or implied throughout the history of phylogenetic methods \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
+Nowadays, this assumption of character independence is even often accepted as a fact and rarely tested or questioned in empirical datasets.
+However, specifically in the case of discrete morphological data, this assumption of independence is probably violated more often than not due to the sheer nature of the data.
+In practice, dependence between characters can be induced in at least three different ways:
 
-Up to this date, a lot of work has been done on character correlation \textit{per se}, but focussing mainly on how to measure correlation between characters to detect patterns of coevolution \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
+\begin{itemize}
+    \item \textbf{Biological induce dependence between characters:} this is the result of a link between two characters through development or biological function (e.g. upper and lower teeth).
+    These links are readily understood through developmental biology or understanding of shared function between features \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
+    One trivial example would be the link between lower and upper molar characters in mammals: one character describing a feature of a lower molar will be expected to be correlated to another character describing a feature of the associated upper molar.
 
-Also \cite{goswami2010}.
+    \item \textbf{Evolutionary induced dependence between characters:} this is the result of sets of character co-evolving along branches in the phylogeny.
+    The character's correlation is thus due to shared evolutionary history of these characters.
+    Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
+    For example, in gnatosthomes, the apparition of the jaw in the gnatosthomes clade correlates with a whole sweep of characters present in modern gnatosthomes (e.g.
+beaks, tribosphenic molars, jaw protrusion, etc.).
+%TG: this links to inapplicable data!
 
-However, only several attempts have been made on the actual effect of characters correlation on phylogenetic inference with two notable exceptions of \cite{Davalos01072014} on empirical data and \cite{ZouConvergence} on molecular data.
+    \item \textbf{Coding induced dependence between characters:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep[see][and reference therein]{Brazeau2011}.
+    These dependences could be due to methodological errors. 
+    For example one could use a character describing a particular feature on a left molar as well as another character describing the same feature on a right molar.
+    However, these dependences could also be due to the nature of the available data, especially in palaeontology.
+    For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
+\end{itemize}
+
+\noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induce correlation due to the modularity of the molars, their shared history and the duplicated coding!
+
+As mentioned above, these sources of dependence between characters are well studied in biology.
+Biological and evolutionary dependences are inherent parts to evo-devo and maroevolution and best practices to avoid coding induced dependences are commonly known and applied.
+However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inferences softwares that are blind to these distinctions.
+This introduces a new, less studied, source of character dependence:
+
+\noindent \textbf{Software induced correlation between characters:} this is the result how software actually interprets the differences between characters.
+I.e.
+the vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treated as different/similar tokens) leading to dependence between characters.
+For example, in a matrix containing four cetartiodactyles - say a pig (e.g.
+\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four characters - say (\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley - resulting in the following matrix:
+
+\begin{table}
+\center
+    \begin{tabular}{r|cccc}
+            & C1 & C2 & C3 & C4\\
+        \hline
+        \textit{Sus} & 1 & 1 & 1 & 1\\
+        \textit{Cervus} & 1 & 1 & 1 & 1\\
+        \textit{Hippopotamus} & 1 & 1 & 1 & 1\\
+        \textit{Balaenoptera} & 0 & 0 & 0 & 0\\
+    \end{tabular}
+    \caption{Example of a matrix with software induced character correlation.
+\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley.}
+    \label{Tab:example_matrix}
+\end{table}
+
+In the example in Table \ref{Tab:example_matrix}, the characters C1 and C2 are the most likely to be truly independent; characters C3 and C4 have a coding induced dependency; characters C1 and C3/C4 have an evolutionary induce dependency and again, characters C2 and C3/C4 are likely to be independent.
+Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the different character states tokens will be used in order to infer a phylogenetic tree.
+
+To our knowledge, this effect of software induced character correlation
+
 It is thus still unclear what will be the effect of characters correlation on phylogenetic inference: how does these correlation really affect topology?
@@
 We found that blabalbal
+
+% However, only several attempts have been made on the actual effect of characters correlation on phylogenetic inference with two notable exceptions of \cite{Davalos01072014} on empirical data and \cite{ZouConvergence} on molecular data.
+%TG: put that somwhere
 
@@
 \subsection{Distinction between different character correlations}
+%TG: this is linked to the story in the intro
+Although both biological and evolutionary induce correlation is what we're interested, unless they are studied explicitly (through evo-devo hypothesis testing for example), coding and software induced correlation are the correlations we end up having to deal with.
+In our study, we do not take into account biological correlation (because our characters are randomly simulated, we can not propose developmental or functional links between them), however, because they are simulated using evolutionary models based on Birth-Death trees, evolutionary induced correlation is present in our simulation (though not analysed specifically).
+The coding induce correlation is simulated through our protocol of maximising/minimising the Character Difference metric from the ``normal'' matrix.
+Finally, the software induced correlation is what is actually measured through these simulations.
 
-Throughout literature, the concept of character correlation (as a requirement for avoiding characters dependence in phylogenetic analysis) can vary along four major points based on the origin of the character correlation:
-\begin{itemize}
-    \item Biological induce correlation between characters: linkage through development or biological function (e.g. upper and lower teeth).
-    These links are readily understood through developmental biology or understanding of shared function between features.
-    This is linked to organisms morphological modules \citep[\textit{sensu}][]{goswami2006morphological}.
-
-    \item Evolutionary induced correlation between characters: linkage through shared evolutionary history (e.g. character co-evolving along branches).
-    These links are the direct cause of the biological correlation highlighted above.
+About evolutionary induced correlation:
 In practice, however, these linkages can only be inferred based on a phylogeny and will thus be dependant on the phylogenetic methods used.
@@
 when long branch attraction occurs).
-
-    \item Coding induced correlation between characters: linkage through coding similar characters on a feature that share a similar pattern.
-    These links can be due to researcher inadvertently defining a character twice or the researcher having to define as many characters as possible on a small feature due to the fragmentary nature of the fossil record (e.g.
-only isolated teeth are available to describe a group of taxa).
-
-    \item Software induced correlation between characters: this designates the way software actually interprets the differences between characters.
-    I.e.
-the vast majority of phylogenetic software ignore both the character's definition, the different states signification (simply treated as different/similar tokens).
-For example, if we have a good assumption that two binary characters coded for four taxa are not correlated (by a prior analysis of the correlations described above), if these two characters both coded `{0,0,0,1}', the software will treat them strictly identically.
-
-    How actually softwares do not care about the character’s definition and whether its in such or such module or arose from such or such evolutionary history (plus what the tokens mean to the researchers).
-It just looks whether the tokens are different or not between species per characters.
-
-\end{itemize}
-
-Although both biological and evolutionary induce correlation is what we're interested, unless they are studied explicitly (through evo-devo hypothesis testing for example), coding and software induced correlation are the correlations we end up having to deal with.
-In our study, we do not take into account biological correlation (because our characters are randomly simulated, we can not propose developmental or functional links between them), however, because they are simulated using evolutionary models based on Birth-Death trees, evolutionary induced correlation is present in our simulation (though not analysed specifically).
-The coding induce correlation is simulated through our protocol of maximising/minimising the Character Difference metric from the ``normal'' matrix.
-Finally, the software induced correlation is what is actually measured through these simulations.
 

### 2017-07-25 Analysis/Character_difference/Modifying_matrices.Rmd
Updated modifying matrix script to fit with the new version of dispRity

--- 2016-12-20 Analysis/Character_difference/Modifying_matrices.Rmd
@@
 
-```{r, eval = FALSE}
+```{r, eval = TRUE}
 ## The latest version of devtools
@@
 ## The latest version of dispRity 
-install_github("TGuillerme/dispRity")
-
-## The latest version of Claddis
-install_github("graemetlloyd/Claddis")
+if(!require(dispRity)) install_github("TGuillerme/dispRity")
 
@@
 ```{r, results="hide", message=FALSE}
-library(Claddis) ; library(dispRity) ; library(diversitree)
-source("functions.R") ; load.functions(test = FALSE)
-
-## Load the C libraries
-dyn.load("../Functions/char.diff.so")
+## Loading the functions not included in dispRity
+source("../../Functions/modify.matrix.R")
 
@@
 ```{r, fig.width=12, fig.height=6}
+## Character difference
+char_diff_matrix_norm <- char.diff(matrix_norm)
 par(mfrow = c(1, 2))
 ## The character difference density (per character)
-plot.char.diff.density(char.diff(matrix_norm), "Normal matrix")
+plot(char_diff_matrix_norm, type = "density", main = "Normal matrix")
 ## The character difference matrix (pairwise)
-plot.cor.matrix(char.diff(matrix_norm), main = "")
+plot(char_diff_matrix_norm, type = "matrix")
 ```
@@
 ```{r, fig.width=12, fig.height=6}
+## Character difference
+char_diff_matrix_maxi <- char.diff(matrix_maxi)
 par(mfrow = c(1, 2))
 ## The character difference density (per character)
-plot.char.diff.density(char.diff(matrix_maxi), "Maximised matrix")
+plot(char_diff_matrix_maxi, type = "density", main = "Maximised matrix")
 ## The character difference matrix (pairwise)
-plot.cor.matrix(char.diff(matrix_maxi), main = "")
+plot(char_diff_matrix_maxi, type = "matrix")
 ```
@@
 ```{r, fig.width=12, fig.height=6}
+## Character difference
+char_diff_matrix_mini <- char.diff(matrix_mini)
 par(mfrow = c(1, 2))
 ## The character difference density (per character)
-plot.char.diff.density(char.diff(matrix_mini), "Minimised matrix")
+plot(char_diff_matrix_mini, type = "density", main = "Minimised matrix")
 ## The character difference matrix (pairwise)
-plot.cor.matrix(char.diff(matrix_mini), main = "")
+plot(char_diff_matrix_mini, type = "matrix")
 ```
@@
 ```{r, fig.width=12, fig.height=6}
+## Character difference
+char_diff_matrix_rand <- char.diff(matrix_rand)
 par(mfrow = c(1, 2))
 ## The character difference density (per character)
-plot.char.diff.density(char.diff(matrix_rand), "Randomised matrix")
+plot(char_diff_matrix_rand, type = "density", main = "Randomised matrix")
 ## The character difference matrix (pairwise)
-plot.cor.matrix(char.diff(matrix_rand), main = "")
+plot(char_diff_matrix_rand, type = "matrix")
 ```
+
+<!-- Paper figure -->
+
+<!-- ```{r, fig.width=12, fig.height=12}
+par(mfrow = c(2, 3), bty = "n", mar = c(1,1,1,1))
+## Tree
+plot(tree, cex = 0.8, main = "A - Simulated tree")
+## Normal matrix
+plot(char_diff_matrix_norm, type = "matrix", main = "B - Normal matrix", legend = FALSE)
+## Empty plot
+plot(0,0, col = "white", xaxt = 'n', yaxt = 'n', xlab = "", ylab = "")
+## Maximised matrix
+plot(char_diff_matrix_maxi, type = "matrix", main = "C - Maximised matrix", legend = FALSE)
+## Minimised matrix
+plot(char_diff_matrix_mini, type = "matrix", main = "D - Minimised matrix", legend = FALSE)
+## Randomised matrix
+plot(char_diff_matrix_rand, type = "matrix", main = "E - Randomised matrix", legend = FALSE)
+```
+ -->

### 2017-07-25 Analysis/EffectCorrelation.Rmd
Added aov analysis

@@
 ```{r, eval = TRUE, message = FALSE}
+library(dispRity)
 source("functions.R") ; load.functions(test = FALSE)
@@
 ```
+
+Some stats
+-----------------------------------
+
+```{r}
+## Getting the whole data
+whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)
+
+## getting the aov data
+aov_data <- get.aov.data(whole_data, "RF", "norm", combined = TRUE)
+
+## with lme4
+library(lme4) ; library(lmerTest)
+
+## NTS function of scenario (random = method + taxa + character)
+model <- lmer(NTS ~ scenario + (1|method/taxa/character), data = aov_data$data)
+anova(model)
+difflsmeans(model, test.effs = "scenario")
+
+## Plotting the residuals
+normal_test <- shapiro.test(residuals(model))
+hist(residuals(model),
+    main = paste("Normal (W=", round(normal_test$statistic, digit = 2) ,
+        "; p=", round(normal_test$p.value, digit = 2) ,")", sep = ""))
+
+## Residuals vs predicted plot
+plot(fitted(model), residuals(model))
+
+## The residuals seem
+```
+
+```{r, fig.width=12, fig.height=6}
+op <- par(mfrow=c(1,2))
+## Plotting the residuals
+normal_test <- shapiro.test(residuals(model))
+hist(residuals(model),
+    main = paste("non-normal (W=", round(normal_test$statistic, digit = 2) ,
+        "; p=", round(normal_test$p.value, digit = 2) ,")", sep = ""))
+
+## Residuals vs predicted plot
+homo_test <- bartlett.test(NTS ~ scenario, data = aov_data$data)
+plot(fitted(model), residuals(model),
+    main = paste("heterodastic (Bartlett K^2=", round(homo_test$statistic, digit = 2) ,
+        "; p=", round(homo_test$p.value, digit = 2) ,")", sep = ""))
+par(op)
+```
+
+The results seem to be heterodastic
+
+Supp
+-----------------------------------
 

### 2017-07-25 Functions/get.aov.data.R
aov.table function

@@ -1,0 +1,124 @@
+#' @title Runs an anova on the results of the analysis
+#'
+#' @description Runs an anova on the results of the analysis
+#'
+#' @param data the global dataset (as a list)
+#' @param metric which metric to use
+#' @param best which best tree (rand or norm)
+#' @param combined whether to combine all the scenarios together (for posthoc TuckeyHSD)
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+get.aov.data <- function(data, metric, best, combined = FALSE) {
+
+    ## Getting the metric
+    if(metric == "RF") metric <- 2
+    if(metric == "Triples") metric <- 4
+
+    ## Getting all the data for one metric
+    NTS_all <- lapply(data, lapply.get.metric, metric)
+
+    ## Extract the right scenarios (i.e. drop rand/norm and true)
+    if(best == "norm") {
+        ## Extra the right base (norm and rand)
+        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 1)
+        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(2,3,4))
+    } else {
+        ## Extra the right base (norm and rand)
+        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 2)
+        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(1,2,3))
+    }
+
+    ## Getting the formula
+    if(combined) {
+        formula <- NTS ~ taxa + character + method + scenario
+    } else {
+        if(best == "norm") {
+            formula <- NTS ~ taxa + character + method + maximised + minimised + randomised
+        } else {
+            formula <- NTS ~ taxa + character + method + normal + maximised + minimised
+        }
+    }
+
+    ## Create the common aov table
+    aov_data <- data.frame("taxa" = count.taxa(NTS),
+                           "character" = count.character(NTS),
+                           "method" = count.method(NTS),
+                           "NTS" = as.vector(unlist(NTS)))
+
+    ## Add the scenarios
+    if(combined) {
+        if(best == "norm") {
+            aov_data <- cbind(aov_data, data.frame("scenario" = count.scenario(NTS, 1, combined = TRUE, best = "norm")))
+        } else {
+            aov_data <- cbind(aov_data, data.frame("scenario" = count.scenario(NTS, 1, combined = TRUE, best = "null")))
+        }
+    } else {
+        if(best == "norm") {
[…]

### 2017-07-25 Functions/run.aov.R

@@ -1,105 +1,0 @@
-#' @title Runs an anova on the results of the analysis
-#'
-#' @description Runs an anova on the results of the analysis
-#'
-#' @param data the global dataset (as a list)
-#' @param metric which metric to use
-#' @param best which best tree (rand or norm)
-#' 
-#' @examples
-#'
-#' @seealso
-#' 
-#' @author Thomas Guillerme
-#' @export
-
-run.aov <- function(data, metric, best) {
-
-    ## Return the right metric from a table
-    get.metric <- function(list, metric) {
-        return(list[,metric])
-    }
-
-    ## Lapply get.metric in a recursive way
-    lapply.get.metric <- function(data, metric) {
-        return(lapply(data, lapply, lapply, lapply, get.metric, metric))
-    }
-
-    count.scenario <- function(NTS, which) {
-        ## Getting the number of counts for each methods
-        counts <- lapply(unlist(unlist(unlist(lapply(NTS, lapply, lapply, `[`, c(1,2,3)), recursive = FALSE), recursive = FALSE), recursive = FALSE), length)
-
-        ## Initialise the scenario selector (binary)
-        names <- rep(0,length(NTS[[1]][[1]][[1]]))
-        names[which] <- 1
-
-        ## Replicate the scenario for the number of counts
-        return(unlist(mapply(rep, as.list(rep(names, 18)), counts)))
-    }
-
-    count.method <- function(NTS) {
-        ## Getting the number of counts for each methods
-        counts <- lapply(unlist(unlist(lapply(NTS, lapply, lapply, unlist), recursive = FALSE), recursive = FALSE), length)
-        ## Replicate the method for the number of counts
-        return(unlist(mapply(rep, as.list(rep(names(NTS[[1]][[1]]), 18)), counts)))
-    }
-
-    count.character <- function(NTS) {
-        ## Getting the number of counts for each character
-        counts <- lapply(unlist(lapply(NTS, lapply, unlist), recursive = FALSE), length)
-        ## Replicate the character number for the number of counts
-        return(unlist(mapply(rep, as.list(rep(names(NTS[[1]]), 3)), counts)))
-    }
-
-    count.taxa <- function(NTS) {
-        ## Getting the number of counts for each taxa
-        counts <- lapply(lapply(NTS, unlist), length)
-        ## Replicate the taxa number for the number of counts
-        return(unlist(mapply(rep, as.list(names(NTS)), counts)))
-    }
-
-    ## Getting the metric
-    if(metric == "RF") metric <- 2
[…]

### 2017-07-25 Functions/run.bhatt.coeff.R
bhatt.coeff function

@@ -1,0 +1,75 @@
+#' @title run a Bhattacharrya Coefficient test on the data
+#'
+#' @description run a Bhattacharrya Coefficient test on the data
+#'
+#' @param data the global dataset
+#' @param metric which metric to use
+#' @param best which best tree (rand or norm)
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+run.bhatt.coeff <- function(data, metric, best) {
+
+    ## Combined
+    combined <- TRUE
+
+    ## Getting the metric
+    if(metric == "RF") metric <- 2
+    if(metric == "Triples") metric <- 4
+
+    ## Getting all the data for one metric
+    NTS_all <- lapply(data, lapply.get.metric, metric)
+
+    ## Extract the right scenarios (i.e. drop rand/norm and true)
+    if(best == "norm") {
+        ## Extra the right base (norm and rand)
+        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 1)
+        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(2,3,4))
+    } else {
+        ## Extra the right base (norm and rand)
+        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 2)
+        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(1,2,3))
+    }
+
+    ## Getting the NTS values as a list
+    NTS_list <- unlist(unlist(unlist(NTS, recursive = FALSE), recursive = FALSE), recursive = FALSE)
+
+    ## Getting the levels of parameters
+    list_names <- names(NTS_list)
+    parameter_names <- c("t25.", "t75.", "t150.", "c100.", "c350.", "c1000.", "bayesian.", "parsimony.")
+    if(best == "norm") {
+        parameter_names <- c(parameter_names, c("maxi", "mini", "rand"))
+    } else {
+        parameter_names <- c(parameter_names, c("norm", "maxi", "mini"))
+    }
+    parameters <- lapply(as.list(parameter_names), grep, list_names, fixed = TRUE)
+    names(parameters) <- parameter_names
+
+    ## Selecting the parameters combinations
+    combo_matrix <- matrix(c(rep(1, 3), rep(2, 3), rep(3, 3), rep(c(4,5,6), 3)), nrow = 2, byrow = TRUE)
[…]

### 2017-07-25 Writing/draft.tex
rewritten intro + methods

@@
 
-The last five years have seen a fantastic expansion of the use of discrete morphological data in macroevolutionary studies.
-In fact, traditionally confined to ``classic'' cladistic studies \citep{felsenstein2004inferring}, these methods have recently been widely expended and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep{pyrondivergence2011,ronquista2012}.
+The last six years have seen a fantastic expansion of the use of discrete morphological data in macroevolutionary studies.
+Traditionally confined to ``classic'' cladistic studies, these methods have recently been widely expanded and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep[e.g.][]{pyrondivergence2011,ronquista2012}.
 In parallel, much development has also been done on pre-phylogenetic analysis \citep[e.g.
@@
 morphological disparity analysis;][]{Close2015,Claddis}.
-Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011}; the inference \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.
+Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011,simoes2017giant}; the inference method \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.
 
@@
 In fact, the assumption of character independence is at the base of phylogenetic inference and as been mentioned or implied throughout the history of phylogenetic methods \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-Nowadays, this assumption of character independence is even often accepted as a fact and rarely tested or questioned in empirical datasets.
-However, specifically in the case of discrete morphological data, this assumption of independence is probably violated more often than not due to the sheer nature of the data.
+Nowadays, this assumption of character independence is nearly always accepted as a fact and rarely tested or questioned in empirical datasets.
+However - especially for discrete morphological data - this assumption of independence is probably violated more often than not due to the sheer nature of phylogenetic data.
 In practice, dependence between characters can be induced in at least three different ways:
@@
 \begin{itemize}
-    \item \textbf{Biological induce dependence between characters:} this is the result of a link between two characters through development or biological function (e.g. upper and lower teeth).
+    \item \textbf{Biological induce dependence between characters:} this is the result of a link between two characters through development and/or biological function.
     These links are readily understood through developmental biology or understanding of shared function between features \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
-    One trivial example would be the link between lower and upper molar characters in mammals: one character describing a feature of a lower molar will be expected to be correlated to another character describing a feature of the associated upper molar.
+    One trivial example would be the link between lower and upper molar characters in mammals: one character describing a feature of a lower molar will be expected to be correlated to another character describing another feature of the associated upper molar.
 
-    \item \textbf{Evolutionary induced dependence between characters:} this is the result of sets of character co-evolving along branches in the phylogeny.
+    \item \textbf{Evolutionary induced dependence between characters:} this is the result of sets of characters co-evolving along branches in the phylogeny.
     The character's correlation is thus due to shared evolutionary history of these characters.
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
-    For example, in gnatosthomes, the apparition of the jaw in the gnatosthomes clade correlates with a whole sweep of characters present in modern gnatosthomes (e.g.
+    For example, the apparition of the jaw as a character in gnatosthomes correlates with a whole sweep of characters present in modern gnatosthomes (e.g.
 beaks, tribosphenic molars, jaw protrusion, etc.).
@@
 
-    \item \textbf{Coding induced dependence between characters:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep[see][and reference therein]{Brazeau2011}.
+    \item \textbf{Coding induced dependence between characters:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
     These dependences could be due to methodological errors. 
-    For example one could use a character describing a particular feature on a left molar as well as another character describing the same feature on a right molar.
-    However, these dependences could also be due to the nature of the available data, especially in palaeontology.
+    For example one could define a character describing a particular feature on a left molar as well as another character describing the same feature on a right molar.
+    It is worth noting that these dependences could also be due to the nature of the available data, especially in palaeontology.
     For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
@@
 
-\noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induce correlation due to the modularity of the molars, their shared history and the duplicated coding!
+\noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding!
 
@@
 I.e.
-the vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treated as different/similar tokens) leading to dependence between characters.
+the vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different/similar tokens) leading to dependence between characters.
 For example, in a matrix containing four cetartiodactyles - say a pig (e.g.
@@
 
-In the example in Table \ref{Tab:example_matrix}, the characters C1 and C2 are the most likely to be truly independent; characters C3 and C4 have a coding induced dependency; characters C1 and C3/C4 have an evolutionary induce dependency and again, characters C2 and C3/C4 are likely to be independent.
+In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
 Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the different character states tokens will be used in order to infer a phylogenetic tree.
+It is thus still unclear what will be the effect of characters correlation on phylogenetic inference: how does these correlation really affect topology?
 
-To our knowledge, this effect of software induced character correlation
-
-It is thus still unclear what will be the effect of characters correlation on phylogenetic inference: how does these correlation really affect topology?
-Additionally, because of the nature of discrete morphological characters and organisms morphology, correlation is expected to rise through both combinatorics properties of the characters or simply through evolutionary processes.
-What is the effect of these two sources of correlations and can we (should we) avoid them?
-
+If the three induced sources of correlation are fairly well studied, their actual interpretation in phylogenetic inference, has, to our knowledge, never been studied through a simulation protocol.
 In this study, we formally assess the effect of discrete character's correlation using simulated and empirical data.
-We propose a new metric to measure the correlation (or difference) between two discrete characters and a protocol to modify discrete data matrices to increase/decrease the correlation between characters within/
+We propose a new metric to measure the difference between characters (as a proxy for these three sources of correlation) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences/similarities between characters.
 We found that blabalbal
@@
     \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
-    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by duplicating characters in order to maximise and minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
-We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximised or minimised character difference to create null expectancy matrices (hereafter called the ``null'' matrices).
-    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``null'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``null'' trees).
-    \item \textbf{Comparing topologies}: finally, we compared to topologies in two ways: (1) first we compared the ``normal'' to the ``null'' trees to assess our simulation protocol; (2) second, we compared the ``maximised'' and ``minimised'' trees to the ``null'' and ``normal'' trees respectively to measure the signal to noise ratio and thus measure the effect of character correlation on topology.
+    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by duplicating characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
+    We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character difference to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
+    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
+    \item \textbf{Comparing topologies}: finally, we compared to topologies in two ways: (1) first we compared the ``normal'' to the ``randomised'' trees to assess our simulation protocol; (2) second, we compared the ``maximised'' and ``minimised'' trees to the ``randomised'' and ``normal'' trees respectively to measure the effect of character correlation on topology.
 \end{enumerate}
@@
 
-\subsection{Measuring correlation between characters}
+\subsection{Measuring differences between characters}
 We define characters as being correlated if they give the same phylogenetic information.
@@
 
-\subsubsection{Characters difference (CD)}
+\subsubsection{Characters Difference (CD)}
 \begin{equation}
@@
 \noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
-$CD$ is a continuous distance metric valid in a mathematical sense (see Supplementary).
-Since we are considering differences as being only Fitch-like (i.e.
-non-weighted and non-ordinate), we ignore the actual distance between two different character states tokens \citep[i.e.
-a Gower distance;][]{GowerDist}.
+$CD$ is a continuous distance metric valid in a mathematical sense (see Supplementary Material 1).
+Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we calculated the difference between character states in a qualitative way.
 Two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
+Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
 In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
-In other words, we replaced all the occurrences of the first token to be $0$, the second to be $1$, etc.
-This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{0,0,1,2,2,1\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
+In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
+This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
+
 When the character difference is equal to $0$, it means that characters convey the same phylogenetic signal.
@@
 These trees are hereafter called the ``true'' trees.
-We then simulated discrete morphological characters using the either of the two following models:
+We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
 \begin{itemize}
-    \item The HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ an scale to sum to one) and using a transition/transvertion rate of two \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
+    \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ and scaled to sum to $1$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
     This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][; see below]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
@@
 Dollo characters;][]{Dollo,wright2015came}.
-    Recently however, \cite{Wright01072016} have shown that an equal rate transition is still strictly majority in empirical data.
+    Recently however, \cite{Wright01072016} have shown that an equal rate transition is still the most present in empirical data.
 \end{itemize}
 
-\noindent For each character, both models (HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
-This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices (see supplementary materials @@@).
+\noindent For each character, both models (morphological HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
+This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices.
 We re-simulated every invariant characters to obtain a matrix with no invariant characters.
@@
 
+It can be noted that this Consistency Index threshold might favor Maximum Parsimony inference (since we measured this index using quick parsimony trees) yet the evolutionary models of each character is slightly in favor of Bayesian inference (since, on average, half the characters evolved following a M$k$ model).
+
 For each trees with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
-The matrices where generated using the \texttt{dispRity R} package \citep[][; \url{https://github.com/TGuillerme/dispRity}]{thomas_guillerme_2016_55646}.
-To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 50 times resulting in 450 morphological matrices (hereafter called the ``normal'' matrices).
+The matrices where generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
+To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 40 times resulting in 360 ``normal'' morphological matrices.
 
 \subsection{Modifying the matrices}
-
-We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}. %TG: TO IMPLEMENT! 
+We calculated the pairwise Character Differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}. %TG: TO IMPLEMENT! 
 We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
@@
 Finally, because this operation effectively changes the weight of characters (i.e.
-giving the characters $<$$0.25$ or $>$$0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a null expectation modification (i.e.
+giving the characters $<$$0.25$ or $>$$0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification (i.e.
 randomly weighting characters).
-This step resulted in a total of 1800 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``null'' matrices).
-The algorithms for the three modifications are available on GitHub@@@. %TG: "@@@" is my "to replace" sign.
+This step resulted in a total of 1440 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
+The algorithms for the three modifications are available on GitHub@@@. %(Modifying_matrices.html) %TG: "@@@" is my "to replace" sign.
+
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=1\textwidth]{Figures/Modif_matrix.pdf}
+\caption{Example illustration of the protocol for modifying matrices.
+The matrices represent the pairwise character differences for 100 characters.
+Blue colours correspond to low character differences and orange colours correspond to high character differences.
+\textbf{A} - a random Birth-Death tree is simulated and used for generating the normal matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised character difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).}
+\label{Fig:modif_matrix}
+\end{figure}
 
 \subsection{Inferring topologies}
+We inferred the topologies in both Bayesian and Maximum Parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
+For both methods, we set the arbitrarily chosen outgroup as an outgroup in the models.
+The Maximum Parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
 
-We inferred the topologies in both Bayesian and Maximum Parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
-The Maximum Parsimony inference was run over @@@ CPU hours using a heuristic search with random sequence addition replicate 10 times (\texttt{hsearch addseq=random nreps=10}).
-The trees where also bootstrapped 100 times using the same search parameters mentioned above.
+The Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$ - \texttt{lset nst=1 rates=gamma Ngammacat=4}) with an variable rate prior an exponential (0.5) shape (\texttt{prset ratepr=variable Shapepr=Exponential(0.5)}).
+We ran two runs of 6 chains each (2 hot, 4 cold) for a maximum of $1\times10^9$ generations with a sampling every 200 generations.
+We automatically stopped the MCMC when the average standard deviation of split frequencies (ASDSF) between both runs fell below 0.01 (with a diagnosis every $1\times10^4$ generations - \texttt{mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES}).
+Due to cluster hardware requirements an to save some time, when chains didn't converged and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the uncoverged chains.
+In practice, these few MCMC got stuck at an ASDSF around (but not below) 0.01.
 
-The Bayesian inference was run over @@@ CPU hours
-
-% lset nst=1 rates=gamma Ngammacat=4;
-% prset ratepr=variable Shapepr=Exponential(0.5);
-
-% mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES file=150t_100c_0304_maxi;
-% sump Filename=150t_100c_0304_maxi Relburnin=YES Burninfrac=0.25;
-% sumt Filename=150t_100c_0304_maxi Relburnin=YES Burninfrac=0.25;
-
- % "hsearch addseq=random nreps=500 rseed=01234 rearrlimit=5000000;"
-
-Due to cluster hardware requirements an to save some time, when chains didn't converged and the runs exceeded 10GB, we aborted the MCMC and computed the consensus tree from the uncoverged chains.
-In practice, these chains got stuck at an ASDS around (but not below) 0.01.
+A strict majority rule tree was then calculated for both Bayesian an Maximum Parsimony trees.
+For the Bayesian consensus trees, the 25\% first trees of the posterior tree distribution were excluded as a burnin.
+The 2880 tree inferences took around 1 CPU century on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC}.
 
 \subsection{Comparing topologies}
-
-We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``null'' matrices and the tree inferred from the ``normal'' matrix.
-
-We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the different inferred trees.
-We ran two types of comparisons, (1) the ability to recover the correct tree (``normal'' trees \textit{vs} ``null'' and ``true'' trees and (2) the effect of character differences by measuring the noise to signal ratio between the ``maximised'' and ``minimised'' trees and the ``null'' and ``normal'' trees.
+We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
+We ran two types of comparisons, (1) the effect of Character Difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}); and (2) the signal to noise ratio induced by the simulation protocol by comparing the ``randomised'' trees to the ``normal'', ``maximised'' and ``minimised'' trees (Figs \ref{Fig:RF_results_rand} and \ref{Fig:Tr_results_rand}).
 
@@
 
-% \subsection{Empirical morphological matrices}
-
-% We applied the same procedure described above (without the morphological matrices simulation) to 100 empirical matrices downloaded from TreeBASE (\url{http://treebase.org/}).
-% The 100 matrices where selected to be the most representative of empirical work with discrete morphological data: with randomly selected matrices with more than 100 each, published between 1985 and 2013 and covering 19 taxonomic classes within 6 phyla (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
-% The full list of matrices is available in the supplementary materials @@@.
-
 \section{Results}
@@
 \centering
-   \includegraphics[width=0.9\textwidth]{Figures/RF_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
+   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
 \caption{Effect of Character Difference on recovering the ``best'' topology.
@@
 \centering
-   \includegraphics[width=0.9\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
+   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
 \caption{Effect of Character Difference on recovering the ``best'' topology.
@@
 
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=1\textwidth]{Figures/RF_results_null.pdf} %TG: These figures will be cleaned up once I have the final results.
+\caption{Effect of Character Difference on recovering the ``random'' topology.
+The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Ronbinson-Fould distance.}
+\label{Fig:RF_results_rand}
+\end{figure}
+
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=1\textwidth]{Figures/Tr_results_null.pdf} %TG: These figures will be cleaned up once I have the final results.
+\caption{Effect of Character Difference on recovering the ``random'' topology.
+The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
+\label{Fig:Tr_results_rand}
+\end{figure}
+
 \section{Discussion}
+
+This goes with former results (see Parins-Fukuchi in review.)
 
@@
 Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
-We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD proof.
+We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD metric proof.
 

### 2017-07-26 Analysis/EffectCorrelation.Rmd
Separated stats bit for supplementaries

@@
 ## getting the aov data
-aov_data <- get.aov.data(whole_data, "RF", "norm", combined = TRUE)
+aov_data <- get.aov.data(whole_data, "RF", "best", combined = TRUE)
 
@@
 model <- lmer(NTS ~ scenario + (1|method/taxa/character), data = aov_data$data)
-anova(model)
-difflsmeans(model, test.effs = "scenario")
+#anova(model)
+#difflsmeans(model, test.effs = "scenario")
+```
 
-## Plotting the residuals
-normal_test <- shapiro.test(residuals(model))
-hist(residuals(model),
-    main = paste("Normal (W=", round(normal_test$statistic, digit = 2) ,
-        "; p=", round(normal_test$p.value, digit = 2) ,")", sep = ""))
-
-## Residuals vs predicted plot
-plot(fitted(model), residuals(model))
-
-## The residuals seem
-```
+First let's test our model assumptions
 
@@
 
-The results seem to be heterodastic
+Data is heterodastic for Triples norm + RF norm
+Data is homoscedastic for Triples best + RF best
+Data is non-normal for Triples norm + RF norm
+
+The test assumptions are not validated.
+We can instead use a non-parametric nested anova and a Bhattacharrya Coefficient (probability of overlap)
+
+```{r}
+##
+run.bhatt.coeff(data, metric = "RF", best = "norm")
+
+```
 

### 2017-07-26 Analysis/EffectCorrelationStats.Rmd
Separated stats bit for supplementaries

---
title: "Effect of character correlation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Effect of character correlation
===============

Loading the functions and the data
-----------------------------------

Getting the functions and the data

```{r, eval = TRUE, message = FALSE}
library(dispRity)
library(lme4)
library(lmerTest)
library(ordinal)
library(MASS)
source("functions.R") ; load.functions(test = FALSE)
dyn.load("../Functions/char.diff.so")
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")
```

Nested anova
-----------------------------------

Let's analysis the data using a nested anova.
The models look at the effect of our scenarios ("maximised", "minimised", etc...) on the normalised tree score (NTS - using both the Robinson-Fould and Triplets scores) with a nested effect of the method (parsimony or bayesian), the number of taxa (25, 75, 150) and the number of characters (100, 150, 1000).

```{r}
## Getting the whole data
whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)

## getting the aov data
aov_data_RF_best <- get.aov.data(whole_data, "RF", "norm", combined = TRUE)
aov_data_Tr_best <- get.aov.data(whole_data, "Triples", "norm", combined = TRUE)
aov_data_RF_null <- get.aov.data(whole_data, "RF", "null", combined = TRUE)
aov_data_Tr_null <- get.aov.data(whole_data, "Triples", "null", combined = TRUE)

## Models with normalised tree score function of scenario (with nestedness)
model_RF_best <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_RF_best)
model_Tr_best <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_Tr_best)
model_RF_null <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_RF_null)
model_Tr_null <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_Tr_null)

## Testing model assumptions
## Normality of the residuals
norm_RF_best <- shapiro.test(residuals(model_RF_best))
norm_Tr_best <- shapiro.test(residuals(model_Tr_best))
norm_RF_null <- shapiro.test(residuals(model_RF_null))
norm_Tr_null <- shapiro.test(residuals(model_Tr_null))
normality <- list(norm_RF_best, norm_Tr_best, norm_RF_null, norm_Tr_null)

## Variance homoscedasticity
homo_RF_best <- bartlett.test(NTS ~ scenario, data = aov_data_RF_best)
homo_Tr_best <- bartlett.test(NTS ~ scenario, data = aov_data_Tr_best)
homo_RF_null <- bartlett.test(NTS ~ scenario, data = aov_data_RF_null)
homo_Tr_null <- bartlett.test(NTS ~ scenario, data = aov_data_Tr_null)
homoscedasticity <- list(homo_RF_best, homo_Tr_best, homo_RF_null, homo_Tr_null)

## Summarise the results
results <- matrix(nrow = 4, data = c(
                round(unlist(lapply(normality, `[[`, 1)), digit = 3),
                round(unlist(lapply(normality, `[[`, 2)), digit = 3),
                round(unlist(lapply(homoscedasticity, `[[`, 1)), digit = 3),
                round(unlist(lapply(homoscedasticity, `[[`, 2)), digit = 3),
                round(unlist(lapply(homoscedasticity, `[[`, 3)), digit = 3)),
            dimnames = list(
                c("RF best", "Triples best", "RF null", "Triples null"),
                c("W", "p.value", "Bartlett's K-squared", "df", "p.value")))

## Check assumption's validities
validity <- apply(results, 1,
    function(X) ifelse((X[2] >= 0.05 && X[5] >= 0.05), TRUE, FALSE))

## Report results
(results <- cbind(as.data.frame(results), validity))

```

All our models violate a least one of the two models assumptions (normality of the residuals and homoscedasticity of variance) and thus cannot be used to compare the scenarios.

Distributions probability overlap (Bhattacharrya Coefficient)
-----------------------------------

We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don't overlap (significant effect of one scenario on topology).
Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).

```{r}
## Running the pairwise Bhattacharrya Coefficient analysis:
(bc_RF_best <- run.bhatt.coeff(whole_data, "RF", "norm"))
(bc_Tr_best <- run.bhatt.coeff(whole_data, "Triples", "norm"))
(bc_RF_null <- run.bhatt.coeff(whole_data, "RF", "null"))
(bc_Tr_null <- run.bhatt.coeff(whole_data, "Triples", "null"))
```

Non-parametric anova
-----------------------------------

We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don't overlap (significant effect of one scenario on topology).
Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).

```{r}
# ## Cumulative Link Mixed Models
# mm1 <- clmm(NTS ~ scenario + (1|taxa/method) + (1|character/method),
#             data = aov_data_RF_best)
# mm1
# summary(mm1)

# ## Ordinated logistic regression
# polr(scenario ~ NTS + (1) ,
#                       data = aov_data_RF_best)

```

### 2017-07-26 Functions/get.aov.data.R
removed formula

@@ -35,15 +35,15 @@ get.aov.data <- function(data, metric, best, combined = FALSE) {
     }
 
     ## Getting the formula
-    if(combined) {
-        formula <- NTS ~ taxa + character + method + scenario
-    } else {
-        if(best == "norm") {
-            formula <- NTS ~ taxa + character + method + maximised + minimised + randomised
-        } else {
-            formula <- NTS ~ taxa + character + method + normal + maximised + minimised
-        }
-    }
+    # if(combined) {
+    #     formula <- NTS ~ taxa + character + method + scenario
+    # } else {
+    #     if(best == "norm") {
+    #         formula <- NTS ~ taxa + character + method + maximised + minimised + randomised
+    #     } else {
+    #         formula <- NTS ~ taxa + character + method + normal + maximised + minimised
+    #     }
+    # }
 
     ## Create the common aov table
     aov_data <- data.frame("taxa" = count.taxa(NTS),
@@ -69,7 +69,7 @@ get.aov.data <- function(data, metric, best, combined = FALSE) {
                                                    "minimised" = count.scenario(NTS, 3, combined = FALSE)))
         }
     }
-    return(list("data" = aov_data, "formula" = formula))
+    return(aov_data)
 }
 
 ## Return the right metric from a table

### 2017-07-26 Functions/run.bhatt.coeff.R
running pairwise bhatt coifs

@@ -41,7 +41,7 @@ run.bhatt.coeff <- function(data, metric, best) {
 
     ## Getting the levels of parameters
     list_names <- names(NTS_list)
-    parameter_names <- c("t25.", "t75.", "t150.", "c100.", "c350.", "c1000.", "bayesian.", "parsimony.")
+    parameter_names <- c("t25.", "t75.", "t150.", "c100.", "c350.", "c1000.", "bayesian", "parsimony")
     if(best == "norm") {
         parameter_names <- c(parameter_names, c("maxi", "mini", "rand"))
     } else {
@@ -68,8 +68,10 @@ run.bhatt.coeff <- function(data, metric, best) {
     bhatt_coeffs <- apply(combo_list, 2, function(X) pairwise.bhatt.coeff(NTS_list[X]))
 
     ## Get the column names (scenario)
+    colnames(bhatt_coeffs) <- apply(combo_matrix, 2, function(X) paste(parameter_names[X], collapse = ""))
     
     ## Get the row names (min:max, etc..)
+    rownames(bhatt_coeffs) <- c("max-min", "max-rand", "min-rand")
 
     return(t(bhatt_coeffs))
 }

### 2017-07-27 Functions/apply.test.R
Applying tests

@@ -1,0 +1,115 @@
+#' @title apply test
+#'
+#' @description applies pairwise tests to the dataset
+#'
+#' @param data the global dataset
+#' @param metric which metric to use
+#' @param best which best tree (rand or norm)
+#' @param test which test (if NULL, simply summarises the distribution)
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+apply.test <- function(data, metric, best, test) {
+
+    ## Combined
+    combined <- TRUE
+
+    ## Getting the metric
+    if(metric == "RF") metric <- 2
+    if(metric == "Triples") metric <- 4
+
+    ## Getting all the data for one metric
+    NTS_all <- lapply(data, lapply.get.metric, metric)
+
+    ## Extract the right scenarios (i.e. drop rand/norm and true)
+    if(best == "norm") {
+        ## Extra the right base (norm and rand)
+        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 1)
+        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(2,3,4))
+    } else {
+        ## Extra the right base (norm and rand)
+        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 2)
+        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(1,2,3))
+    }
+
+    ## Getting the NTS values as a list
+    NTS_list <- unlist(unlist(unlist(NTS, recursive = FALSE), recursive = FALSE), recursive = FALSE)
+
+    ## Getting the levels of parameters
+    list_names <- names(NTS_list)
+    parameter_names <- c("t25.", "t75.", "t150.", "c100.", "c350.", "c1000.", "bayesian", "parsimony")
+    if(best == "norm") {
+        parameter_names <- c(parameter_names, c("maxi", "mini", "rand"))
+    } else {
+        parameter_names <- c(parameter_names, c("norm", "maxi", "mini"))
+    }
+    parameters <- lapply(as.list(parameter_names), grep, list_names, fixed = TRUE)
+    names(parameters) <- parameter_names
+
+    ## Selecting the parameters combinations
+    combo_matrix <- matrix(c(rep(1, 3), rep(2, 3), rep(3, 3), rep(c(4,5,6), 3)), nrow = 2, byrow = TRUE)
[…]

### 2017-07-27 Functions/run.bhatt.coeff.R

@@ -1,77 +1,0 @@
-#' @title run a Bhattacharrya Coefficient test on the data
-#'
-#' @description run a Bhattacharrya Coefficient test on the data
-#'
-#' @param data the global dataset
-#' @param metric which metric to use
-#' @param best which best tree (rand or norm)
-#' 
-#' @examples
-#'
-#' @seealso
-#' 
-#' @author Thomas Guillerme
-#' @export
-
-run.bhatt.coeff <- function(data, metric, best) {
-
-    ## Combined
-    combined <- TRUE
-
-    ## Getting the metric
-    if(metric == "RF") metric <- 2
-    if(metric == "Triples") metric <- 4
-
-    ## Getting all the data for one metric
-    NTS_all <- lapply(data, lapply.get.metric, metric)
-
-    ## Extract the right scenarios (i.e. drop rand/norm and true)
-    if(best == "norm") {
-        ## Extra the right base (norm and rand)
-        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 1)
-        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(2,3,4))
-    } else {
-        ## Extra the right base (norm and rand)
-        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 2)
-        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(1,2,3))
-    }
-
-    ## Getting the NTS values as a list
-    NTS_list <- unlist(unlist(unlist(NTS, recursive = FALSE), recursive = FALSE), recursive = FALSE)
-
-    ## Getting the levels of parameters
-    list_names <- names(NTS_list)
-    parameter_names <- c("t25.", "t75.", "t150.", "c100.", "c350.", "c1000.", "bayesian", "parsimony")
-    if(best == "norm") {
-        parameter_names <- c(parameter_names, c("maxi", "mini", "rand"))
-    } else {
-        parameter_names <- c(parameter_names, c("norm", "maxi", "mini"))
-    }
-    parameters <- lapply(as.list(parameter_names), grep, list_names, fixed = TRUE)
-    names(parameters) <- parameter_names
-
-    ## Selecting the parameters combinations
-    combo_matrix <- matrix(c(rep(1, 3), rep(2, 3), rep(3, 3), rep(c(4,5,6), 3)), nrow = 2, byrow = TRUE)
[…]

### 2017-07-27 Writing/draft.tex
Updated methods

@@
 The metrics scores where obtained using the TreeCmp java script \citep{Bogdanowicz2012}.
-The measurements where then standardised using the Normalised Tree Similarity metric \citep[i.e.
+The measurements where then standardised using the Normalised Tree Similarity metric \citep[$NTS$; i.e.
 centering the metrics scores using the mean metric score for 1000 pairwise comparisons between random trees with $n$ taxa;][]{Bogdanowicz2012,Guillerme2016146}.
 When the normalised metric has a score of $1$ it means both trees are identical, when it has a score of $0$ it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
-The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Fould Similarity reflects the conservation of clades (i.e.
-a score close to $1$ indicates that most clades are identical in both trees); and (2) the Normalised Triplets Similarity reflects the position of taxa (i.e.
+The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Fould ($NTS_{RF}$)Similarity reflects the conservation of clades (i.e.
+a score close to $1$ indicates that most clades are identical in both trees); and (2) the Normalised Triplets Similarity ($NTS_{Tr}$) reflects the position of taxa (i.e.
 a score close to $1$ indicates that most taxa have the same neighbours in both trees).
+
+Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at 1, the resulting residuals distribution of $NTS$ scores was not normal thus preventing the use of parametric nested anova for comparisons (see Supplementary @@@).
+Instead, we used the a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
+We considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
+Comparisons falling within these range can not be designated as strictly similar/different but can still be ranked (e.g.
+for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).
+
+% TG: Data not normal so no fancy test yet
 

### 2017-07-27 Analysis/EffectCorrelationStats.Rmd
Stats vignette

@@
 
-Getting the functions and the data
+Getting the functions and the data.
 
@@
 library(lmerTest)
-library(ordinal)
-library(MASS)
 source("functions.R") ; load.functions(test = FALSE)
@@
 source("../Data/load.data.R")
+
+## Getting the whole data
+whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)
+```
+
+Summary statistics
+-----------------------------------
+
+We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
+Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
+The p-values are adjusted for multiple tests using Bonferroni-Holm correction.
+
+```{r}
+## Running the pairwise Wilcox tests:
+(sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL))
+(sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL))
+(sum_RF_null <- apply.test(whole_data, "RF", "null", NULL))
+(sum_Tr_null <- apply.test(whole_data, "Triples", "null", NULL))
+
 ```
@@
 ```{r}
-## Getting the whole data
-whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)
-
 ## getting the aov data
@@
 All our models violate a least one of the two models assumptions (normality of the residuals and homoscedasticity of variance) and thus cannot be used to compare the scenarios.
+This is due to the fact that the metric is bounded to 1.
 
@@
 ## Running the pairwise Bhattacharrya Coefficient analysis:
-(bc_RF_best <- run.bhatt.coeff(whole_data, "RF", "norm"))
-(bc_Tr_best <- run.bhatt.coeff(whole_data, "Triples", "norm"))
-(bc_RF_null <- run.bhatt.coeff(whole_data, "RF", "null"))
-(bc_Tr_null <- run.bhatt.coeff(whole_data, "Triples", "null"))
+(bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff))
+(bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff))
+(bc_RF_null <- apply.test(whole_data, "RF", "null", bhatt.coeff))
+(bc_Tr_null <- apply.test(whole_data, "Triples", "null", bhatt.coeff))
 ```
 
-Non-parametric anova
+Non-parametric pairwise comparisons
 -----------------------------------
 
-We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
-A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don't overlap (significant effect of one scenario on topology).
-Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).
+We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
+Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
+The p-values are adjusted for multiple tests using Bonferroni-Holm correction.
 
 ```{r}
-# ## Cumulative Link Mixed Models
-# mm1 <- clmm(NTS ~ scenario + (1|taxa/method) + (1|character/method),
-#             data = aov_data_RF_best)
-# mm1
-# summary(mm1)
+## Running the pairwise Wilcox tests:
+wx_RF_best <- apply.test(whole_data, "RF", "norm", wilcox.test)
+wx_Tr_best <- apply.test(whole_data, "Triples", "norm", wilcox.test)
+wx_RF_null <- apply.test(whole_data, "RF", "null", wilcox.test)
+wx_Tr_null <- apply.test(whole_data, "Triples", "null", wilcox.test)
 
-# ## Ordinated logistic regression
-# polr(scenario ~ NTS + (1) ,
-#                       data = aov_data_RF_best)
+## Applying Bonferroni correction
+wx_RF_best[,c(2,4,6)] <- p.adjust(wx_RF_best[,c(2,4,6)], method = "bonferroni")
+wx_Tr_best[,c(2,4,6)] <- p.adjust(wx_Tr_best[,c(2,4,6)], method = "bonferroni")
+wx_RF_null[,c(2,4,6)] <- p.adjust(wx_RF_null[,c(2,4,6)], method = "bonferroni")
+wx_Tr_null[,c(2,4,6)] <- p.adjust(wx_Tr_null[,c(2,4,6)], method = "bonferroni")
 
+## Print the results
+wx_RF_best
+wx_Tr_best
+wx_RF_null
+wx_Tr_null
 ```

### 2017-07-28 Writing/draft.tex
methods

@@
 \usepackage{mathtools}
+\usepackage{pdflscape}
 
@@
 \subsection{Measuring differences between characters}
+\label{CD_description}
 We define characters as being correlated if they give the same phylogenetic information.
@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We ran two types of comparisons, (1) the effect of Character Difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}); and (2) the signal to noise ratio induced by the simulation protocol by comparing the ``randomised'' trees to the ``normal'', ``maximised'' and ``minimised'' trees (Figs \ref{Fig:RF_results_rand} and \ref{Fig:Tr_results_rand}).
+We explored the effect of Character Difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Tables @@@).
 
@@
 
-% TG: Data not normal so no fancy test yet
+\section{Results}
 
-\section{Results}
+@@@ NOTE: THESE RESULTS ARE NOT FINAL - I thus didn't spend too much time describing them @@@
+
+\subsection{Effect of character differences on topology}
+The overall amount of pairwise $CD$ in a matrix has an effect of the ability to recover the correct topology.
+An overall decrease in $CD$ leads to an decrease in $NTS_{RF}$ (as a proxy for clade conservation - Fig.
+\ref{Fig:RF_results_best}) and in $NTS_{Tr}$ (as a proxy for taxa position - Fig.
+\ref{Fig:Tr_results_best}).
+Conversely, an increase in $CD$ leads to an increase in in $NTS_{RF}$ (Fig. \ref{Fig:RF_results_best}) and $NTS_{Tr}$ (Fig. \ref{Fig:Tr_results_best}).
+This beneficial effect can be contrasted however with a similar beneficial effect when simply randomising the character differences (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
+
+\subsubsection{Effect of the number of characters}
+This negative effect of decreased $CD$ (and positive of increased $CD$) on topology is increased with the number of characters.
+% Test pulled difference between character numbers.
+
+\subsubsection{Effect of the number of taxa}
+Conversely, the effect of variation in $CD$ doesn't seem to drastically increase with the number of taxa after 75 taxa in our simulations.
+% Test pulled difference between number of taxa.
+
+\subsection{Effect of character differences on the inference method}
+There seems to be a strong effect on the ability to recover the correct topology depending on the tree inference method.
+Parsimony trees seems to be more sensible than Bayesian ones.
+% Test pulled difference between methods.
 
@@
    \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
-\caption{Effect of Character Difference on recovering the ``best'' topology.
+\caption{\small{Effect of Character Difference on recovering the ``best'' topology.
 The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
-The x axis represents the different Character Difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}
+The x axis represents the different Character Difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
 \label{Fig:RF_results_best}
@@
 
-\begin{figure}[!htbp]
+% latex table generated in R 3.4.0 by xtable 1.8-2 package
+% Fri Jul 28 10:29:26 2017
+\begin{table}[ht]
 \centering
-   \includegraphics[width=1\textwidth]{Figures/RF_results_null.pdf} %TG: These figures will be cleaned up once I have the final results.
-\caption{Effect of Character Difference on recovering the ``random'' topology.
-The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Ronbinson-Fould distance.}
-\label{Fig:RF_results_rand}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-   \includegraphics[width=1\textwidth]{Figures/Tr_results_null.pdf} %TG: These figures will be cleaned up once I have the final results.
-\caption{Effect of Character Difference on recovering the ``random'' topology.
-The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
-\label{Fig:Tr_results_rand}
-\end{figure}
+\begin{tabular}{rlllrrrrrrrrr}
+  \hline
+Taxa & Char. & Method & \multicolumn{3}{c}{max \textit{vs}. min} & \multicolumn{3}{c}{max \textit{vs}. rand} & \multicolumn{3}{c}{min \textit{vs}. rand} \\
+ & & & BC & W & p & BC & W & p & BC & W & p \\ 
+  \hline
+25 & 100 & Bay. & 0.67 & 1130.00 & \textbf{0.00} & 0.85 & 856.00 & \textbf{1.00} & 0.74 & 320.00 & \textbf{0.00} \\ 
+   & 350 &   & 0.71 & 1236.50 & \textbf{0.00} & \textbf{0.96} & 844.00 & \textbf{1.00} & 0.73 & 338.00 & \textbf{0.00} \\ 
+   & 1000 &   & 0.71 & 1321.50 & \textbf{0.00} & 0.89 & 802.00 & 1.00 & 0.67 & 226.00 & \textbf{0.00} \\ 
+75 & 100 &   & 0.48 & 1366.50 & \textbf{0.00} & 0.66 & 1134.00 & \textbf{0.01} & 0.69 & 264.00 & \textbf{0.00} \\ 
+   & 350 &   & 0.40 & 999.50 & \textbf{0.00} & 0.59 & 878.50 & \textbf{0.00} & 0.71 & 188.00 & \textbf{0.00} \\ 
+   & 1000 &   & 0.53 & 390.00 & \textbf{0.00} & 0.80 & 341.50 & 0.13 & 0.69 & 100.00 & 0.14 \\ 
+150 & 100 &   & \textbf{0.00} & 256.00 & \textbf{0.00} & \textbf{0.00} & 256.00 & \textbf{0.00} & 0.42 & 24.00 & \textbf{0.01} \\ 
+    & 350 &   & 0.07 & 361.00 & \textbf{0.00} & \textbf{0.05} & 361.00 & \textbf{0.00} & 0.39 & 56.50 & \textbf{0.02} \\ 
+    & 1000 &   & \textbf{0.00} & 121.00 & \textbf{0.00} & 0.13 & 119.00 & \textbf{0.01} & \textbf{0.00} & 0.00 & \textbf{0.00} \\ 
+25 & 100 & Par. & 0.63 & 357.00 & \textbf{0.00} & 0.91 & 234.00 & 1.00 & 0.71 & 85.00 & 0.10 \\ 
+   & 350 &   & 0.34 & 387.00 & \textbf{0.00} & 0.79 & 287.50 & 0.94 & 0.49 & 61.50 & \textbf{0.01} \\ 
+   & 1000 &  & 0.43 & 384.00 & \textbf{0.00} & 0.79 & 289.50 & 0.62 & 0.71 & 65.00 & \textbf{0.01} \\ 
+75 & 100 &   & 0.38 & 379.50 & \textbf{0.00} & 0.40 & 359.00 & \textbf{0.00} & 0.75 & 157.50 & 1.00 \\ 
+   & 350 &   & 0.25 & 393.00 & \textbf{0.00} & 0.76 & 311.00 & 0.13 & 0.65 & 55.50 & \textbf{0.01} \\ 
+   & 1000 &   & 0.15 & 393.00 & \textbf{0.00} & 0.46 & 362.00 & \textbf{0.00} & 0.46 & 58.00 & \textbf{0.01} \\ 
+150 & 100 &   & 0.34 & 357.50 & \textbf{0.00} & 0.49 & 356.50 & \textbf{0.00} & 0.73 & 197.00 & 1.00 \\ 
+    & 350 &   & 0.17 & 390.00 & \textbf{0.00} & 0.60 & 348.00 & \textbf{0.00} & 0.63 & 63.50 & \textbf{0.01} \\ 
+    & 1000 &   & \textbf{0.00} & 400.00 & \textbf{0.00} & 0.71 & 328.50 &\textbf{0.03} & 0.34 & 24.00 & \textbf{0.00} \\ 
+   \hline
+\end{tabular}
+\caption{Example of a result table: BC = Bhattacharyya Coefficient, W = wilcox test statistic and p = associated p-value} 
+\label{LABEL}
+\end{table}
 
@@
 \subsection{Effect of character differences on topology}
-As expected, reducing character difference reduces the ability of recovering the best tree topology since the ``assumption'' of character independence is ``violated''.
-Conversely increasing character difference increases the ability of recover the best tree topology.
-However this has not more effect of topology than simply bootstrapping the matrix.
+One way to look at the $CD$ metric is as a inverse of character correlation (see Methods \ref{CD_description}): a high $CD$ approximates a low level of character correlation and vice versa.
+When characters are correlated, one could expect the matrices to convey a really strong (but probably incorrect) phylogenetic signal (since every characters agrees with each other) and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasies.
+Intuitively this would lead to badly resolved trees in our scenarios where the $CD$ was high.
+
+One first noteworthy point is that, on average, the matrices with the highest $CD$ where the ones that performed the best at recovering the ``normal'' topology which in turns confirms the intuition that discrete morphological characters should be independent.
+This is also true when the characters are more similar to each other (low $CD$) resulting in worth trees due to the software treating all these characters as correlated.
+
+One second noteworthy point is that compared to simply randomising $CD$ in the matrices (effectively bootstrapping), maximising the $CD$ was, at least in small sized matrices, not more effective to recover the ``normal'' topology.
+Since many matrices are in the range size \citep{guillerme2016assessment} it is actually encouraging that actively coding characters to have high $CD$ values will not significantly improve the quality of the matrix compared to actually randomly coding characters.
+Coding characters with low $CD$ however will decrease the quality of the matrix.
+
+\subsubsection{Effect of the number of characters}
+As expected by the sheer design of this research project, the number of character has a nested effect on character correlation: increasing the number of characters in our simulation scenarios (i.e.
+``maximised'', ``minimised'' and ``randomised'') effectively increased the amount of high/low $CD$ values per matrix.
+This increase in character numbers thus increases the effect of $CD$ on tree topology with $CD$
+
+\subsubsection{Effect of the number of taxa}
+Same blabalbal.
 
 \subsection{Effect of character differences on the inference method}
-Scenarios where the characters are most different to each other could favor Bayesian over Parsimony.
+The rather strong nested effect of the method for recovering can be due to the sheer difference in both methods.
+These difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted}.
+In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[those resulting in a $CI>0.26$ in a quick parsimony search;]{OReilly20160081} could also favour Parsimony analysis.
+It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
+
+That being said, it is interesting to analysis this difference in tree inference method in the light of the effect of high/low $CD$.
+In fact the differences observed here could be due to the inherent mechanisms of each method:
+(1) when many characters are homoplasic (high $CD$) parsimony will cope with this information by generating longer trees (which are in turn disfavoured in the tree search).
+Conversely Bayesian can cope with homoplasy by making longer branches and also saves sub-optimal trees.
+
+Scenarios where the characters are most different to each other could favour Bayesian over Parsimony.
 In fact, the only way to deal with it in parsimony analysis is to generate really homoplasic trees (since most character gives a different signal).

### 2017-07-28 Writing/draft.tex
Updated draft (kind of full)

@@
 
-\noindent{\Large \bf Effect of discrete character correlation on tree inference}
+\noindent{\Large \bf Effect of violating character independence on tree topology}
 \bigskip
@@
 %---------------------------------------------
+
+% SystBiol
+% Palaeobiology
+% JEB
+% Biological Journal of the Linnean Society
+% Royal Society Open Science - but rewriting needed (more story style)
+% Palaeontology
 
@@
 In parallel, much development has also been done on pre-phylogenetic analysis \citep[e.g.
-data collection;][]{morphobank} as well as on post-phylogenetic analysis \citep[e.g.
+data collection;][]{morphobank} as well as on post-phylogenetic ones \citep[e.g.
 morphological disparity analysis;][]{Close2015,Claddis}.
@@
     \item \textbf{Coding induced dependence between characters:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
-    These dependences could be due to methodological errors. 
-    For example one could define a character describing a particular feature on a left molar as well as another character describing the same feature on a right molar.
-    It is worth noting that these dependences could also be due to the nature of the available data, especially in palaeontology.
+    These links can be due to methodological errors: for example one could define a character describing a particular feature on a left molar as well as another character describing the same feature on a right molar.
+    It is worth noting, however, that these dependences could also be due to the nature of the available data, especially in palaeontology.
     For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
@@
 the vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different/similar tokens) leading to dependence between characters.
-For example, in a matrix containing four cetartiodactyles - say a pig (e.g.
-\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four characters - say (\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley - resulting in the following matrix:
+For example, consider the following matrix containing four cetartiodactyles - say a pig (e.g.
+\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four characters - say (\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley:
 
@@
 We propose a new metric to measure the difference between characters (as a proxy for these three sources of correlation) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences/similarities between characters.
-We found that blabalbal
+We found that @@@
 
@@
     \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
-    \item \textbf{Comparing topologies}: finally, we compared to topologies in two ways: (1) first we compared the ``normal'' to the ``randomised'' trees to assess our simulation protocol; (2) second, we compared the ``maximised'' and ``minimised'' trees to the ``randomised'' and ``normal'' trees respectively to measure the effect of character correlation on topology.
+    \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on topology.
 \end{enumerate}
@@
 
-\subsubsection{Characters Difference (CD)}
+\subsubsection{Characters Difference ($CD$)}
 \begin{equation}
@@
 Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
-In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
+We standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
 In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
@@
 We re-simulated every invariant characters to obtain a matrix with no invariant characters.
-To ensure that our simulation where reflecting realistic observed parameters, we only selected matrices with Consistency Indices superior to $0.26$ \citep{OReilly20160081}.
-
-It can be noted that this Consistency Index threshold might favor Maximum Parsimony inference (since we measured this index using quick parsimony trees) yet the evolutionary models of each character is slightly in favor of Bayesian inference (since, on average, half the characters evolved following a M$k$ model).
+To ensure that our simulation where reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{OReilly20160081}.
 
@@
 We inferred the topologies in both Bayesian and Maximum Parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
-For both methods, we set the arbitrarily chosen outgroup as an outgroup in the models.
+For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
 The Maximum Parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
@@
 We automatically stopped the MCMC when the average standard deviation of split frequencies (ASDSF) between both runs fell below 0.01 (with a diagnosis every $1\times10^4$ generations - \texttt{mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES}).
-Due to cluster hardware requirements an to save some time, when chains didn't converged and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the uncoverged chains.
+Due to cluster hardware requirements an to save some time, when chains didn't converged and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the unconverged chains.
 In practice, these few MCMC got stuck at an ASDSF around (but not below) 0.01.
@@
 When the normalised metric has a score of $1$ it means both trees are identical, when it has a score of $0$ it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
-The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Fould ($NTS_{RF}$)Similarity reflects the conservation of clades (i.e.
+The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Fould ($NTS_{RF}$) Similarity reflects the conservation of clades (i.e.
 a score close to $1$ indicates that most clades are identical in both trees); and (2) the Normalised Triplets Similarity ($NTS_{Tr}$) reflects the position of taxa (i.e.
@@
 
-\subsection{Effect of character differences on topology}
+\subsection{Effect of Character Differences on topology}
 The overall amount of pairwise $CD$ in a matrix has an effect of the ability to recover the correct topology.
@@
 
-\subsection{Effect of character differences on the inference method}
+\subsection{Effect of Character Differences on the inference method}
 There seems to be a strong effect on the ability to recover the correct topology depending on the tree inference method.
@@
 
-This goes with former results (see Parins-Fukuchi in review.)
+% This goes with former results (see Parins-Fukuchi in review.)
 
-\subsection{Effect of character differences on topology}
+\subsection{Effect of Character Differences on topology}
 One way to look at the $CD$ metric is as a inverse of character correlation (see Methods \ref{CD_description}): a high $CD$ approximates a low level of character correlation and vice versa.
@@
 
-One first noteworthy point is that, on average, the matrices with the highest $CD$ where the ones that performed the best at recovering the ``normal'' topology which in turns confirms the intuition that discrete morphological characters should be independent.
+One first noteworthy point is that, on average, the matrices with the highest $CD$ where the ones that performed the best at recovering the ``normal'' topology which in turns confirms the intuition that discrete morphological characters should be independent in phylogenetics.
 This is also true when the characters are more similar to each other (low $CD$) resulting in worth trees due to the software treating all these characters as correlated.
-
 One second noteworthy point is that compared to simply randomising $CD$ in the matrices (effectively bootstrapping), maximising the $CD$ was, at least in small sized matrices, not more effective to recover the ``normal'' topology.
@@
 \subsubsection{Effect of the number of taxa}
-Same blabalbal.
+Same blabla (probably).
 
-\subsection{Effect of character differences on the inference method}
+\subsection{Effect of Character Differences on the inference method}
 The rather strong nested effect of the method for recovering can be due to the sheer difference in both methods.
-These difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted}.
-In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[those resulting in a $CI>0.26$ in a quick parsimony search;]{OReilly20160081} could also favour Parsimony analysis.
+These difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will more readily recover the ``right'' topology when analysing discrete morphological alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have been made in favour of parsimony: \citealt{spencerefficacy2013,goloboff2017weighted}).
+In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e.
+those with in a $CI>0.26$ in a quick parsimony search;]{OReilly20160081} could also favour Parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
@@
 That being said, it is interesting to analysis this difference in tree inference method in the light of the effect of high/low $CD$.
-In fact the differences observed here could be due to the inherent mechanisms of each method:
-(1) when many characters are homoplasic (high $CD$) parsimony will cope with this information by generating longer trees (which are in turn disfavoured in the tree search).
-Conversely Bayesian can cope with homoplasy by making longer branches and also saves sub-optimal trees.
-
-Scenarios where the characters are most different to each other could favour Bayesian over Parsimony.
-In fact, the only way to deal with it in parsimony analysis is to generate really homoplasic trees (since most character gives a different signal).
-On the other hand probabilistic methods can better deal with this problem by increasing branch length to accommodate more changes in characters between taxa.
+In fact the differences observed here could be due to the inherent mechanisms of each method.
+When many characters are homoplasic (high $CD$) parsimony will cope with this information by generating longer trees (which will be more likely to be discarded in the tree search).
+Conversely Bayesian can cope with homoplasy more easily by increasing some branch length to accommodate more changes and will also incorporate sub-optimal trees in the final results.
 
 \subsection{Distinction between different character correlations}
-%TG: this is linked to the story in the intro
-Although both biological and evolutionary induce correlation is what we're interested, unless they are studied explicitly (through evo-devo hypothesis testing for example), coding and software induced correlation are the correlations we end up having to deal with.
-In our study, we do not take into account biological correlation (because our characters are randomly simulated, we can not propose developmental or functional links between them), however, because they are simulated using evolutionary models based on Birth-Death trees, evolutionary induced correlation is present in our simulation (though not analysed specifically).
-The coding induce correlation is simulated through our protocol of maximising/minimising the Character Difference metric from the ``normal'' matrix.
-Finally, the software induced correlation is what is actually measured through these simulations.
+Although both biological and evolutionary induce correlation are what most evolutionary biologists are genuinely interested in and that coding induced correlation is readily avoided through best practices, software induced character correlation would be the main source of character dependence in systematic biology.
 
-About evolutionary induced correlation:
-In practice, however, these linkages can only be inferred based on a phylogeny and will thus be dependant on the phylogenetic methods used.
-For example, if a series of characters co-evol along a branch, this can be due to actual accumulation of evolutionary changes along this branch (e.g.
-characters linked to volancy along the Chiroptera branch) but could also be the result of inference biases (e.g.
-when long branch attraction occurs).
+As we mentioned in the introduction of this study, we identified three different major sources of character correlation (biological, evolutionary and coding induced correlation).
+In practice, however, we argue that software induced character correlation would be the main source of character dependence in systematic biology (see Table \ref{Tab:example_matrix}).
+Therefore, in this study, we mainly focused on this source of correlation between characters.
+It is important to note, however, that because the characters were simulated using evolutionary models based on Birth-Death trees, evolutionary induced correlation is present in our simulation (though not analysed specifically).
+Conversely, biological induced correlation would be more complex to simulate and are out of the scope of this analysis.
+
+Additionally, the two types of correlations that most evolutionary biologists are genuinely interested in (biological and evolutionary) can only be studied \textit{a posteriori} using phylogenetic trees and should not be used as an \textit{a priori}.
+In fact, the links between specific characters should studied explicitly through hypothesis testing based on an underlying phylogenetic framework - for example in evo-devo \citep[e.g.][]{goswami2006morphological} or in macroevolution \citep[e.g.][]{fitzjohn2014much} - and should thus be independent of the phylogenetic inference itself.
 
 \subsection{Conclusion}
-Character differences as a proxy for character correlation is not a problem \textit{per se} as long as characters are not actively coded to be similar.
-On the other hand, actively selecting characters based to favor independence does not increase the ability of recovering tree topology more than bootstrapping the characters.
+As we presented in the introduction, dependence between characters can be induced through three main phenomenon: the biological, evolutionary or coding induced correlation between characters.
+Although one could argue (righteously!) that coding induced correlation is easily avoidable and that biological and evolutionary correlations are actually at the core of evolutionary biology, in the age of ever expanding morphological matrices \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the dependence between characters to increase stochastically.
+Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
+
+We found that Character Differences as a proxy for character correlation is not a problem \textit{per se} as long as characters are not actively coded to be similar.
+On the other hand, actively selecting characters based to favour independence does not highly increase the ability of recovering tree topology (compared to simply bootstrapping a matrix - at least for modest size matrices).
+We hope that these results will encourage researchers working on discrete morphological characters to continue exploring morphology through discrete characters to better understand the Tree of Life.
 

### 2017-08-09 Analysis/EffectCorrelation.Rmd
Updated analysis

--- 2017-07-17 Analysis/EffectCorrelation.Rmd
@@
 ```
+
+Some stats
+-----------------------------------
+
+Supp
+-----------------------------------
 

### 2017-09-12 Analysis/EffectCorrelationStats.Rmd
updated analysis

@@
 ## Print the results
-wx_RF_best
-wx_Tr_best
-wx_RF_null
-wx_Tr_null
+round(wx_RF_best, digit = 3)
+round(wx_Tr_best, digit = 3)
+round(wx_RF_null, digit = 3)
+round(wx_Tr_null, digit = 3)
 ```
+
+```{r}
+library(xtable)
+## Combining both tables (for paper)
+col_description <- matrix(ncol = 3,
+    data = c(rep(rep(c("25", "75", "150"), each = 3), 2),
+             rep(c("100", "350", "1000"), 6),
+             rep(c("Bay.", "Par."), each = 9)), dimnames = NULL)
+
+matrix_out <- cbind(as.data.frame(col_description), as.data.frame(cbind(bc_RF_best[,1], wx_RF_best[,1:2], bc_RF_best[,2], wx_RF_best[,3:4], bc_RF_best[,3], wx_RF_best[,5:6])))
+
+rownames(matrix_out) <- NULL
+
+xtable(matrix_out, caption = "CAPTION", label = "LABEL", round = 3)
+
+(bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff))
+(bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff))
+(bc_RF_null <- apply.test(whole_data, "RF", "null", bhatt.coeff))
+(bc_Tr_null <- apply.test(whole_data, "Triples", "null", bhatt.coeff))
+
+```

### 2017-09-14 Writing/draft.tex
Updated results
Abstract thoughts

@@
 
+The project implies inferring a total of 2880 phylogenetic trees (40 replicates of 9 data combination scenarios generating 4 trees each; both inferred in Bayesian and under Maximum Parsimony).
+The calculations for the project is ongoing since December 2016 a and is now 95.27\% completed (with a current cumulated computational time of around 80 CPU years that generated 3.2 TB of trees).
+The pipelines for analysing the data and producing the publishable figures and tables is already implemented and most of the manuscript is already written ()
+
+Overall we found a positive effect of reducing character correlations on recovering the correct topology.
+However, this effect is less important for matrices with a small number of taxa (~25) were reducing character correlation was not more effective than randomly drawing characters.
+Furthermore, in bigger matrices (~350 taxa), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
+
+These results suggest that ignoring the problem of character correlation or independence can easily impact topology in phylogenetic analysis.
+However, encouragingly, they also suggest that, unless correlation is effectively maximised or minimised, probabilistic methods can easily accommodate for random correlation between characters.
+
 \noindent (Keywords: )\\
@@
 \newpage
-
-%---------------------------------------------
-% LaTeX tips for modifying/editing the document:
-%---------------------------------------------
-% - You can comment using the percentage sign. I suggest you use the % sign alone for commenting out sections of the text:
-%       e.g. "This is a really long sentence %because this sentence is very long." Here the % is used for ignoring the end of the sentence (but for some reason you want to keep track of it).%
-%       For comments as in verbose comments, I suggest you use "%MB:":
-%       e.g. "This is a really long sentence %because this sentence is very long. %MB: yeah, no shit!" 
-% - For optimal version control, write only one sentence per line (for more precise track changes)
-% - To build the pdf, use command+B in Sublime.
-% - Because of the bibliography, the pdf needs to be build in the same folder that contains "References.bib" and "sysbio.bst"
-% - For citing papers, you must put their bibtex reference in the "References.bib" file and then you can use the following sysbio tags:
-%        \cite{bibtexBob2000} for citing within a sentence: "Bob (2000)"
-%        \citep{bibtexBob2000} for citing within brackets: "(Bob, 2000)"
-%        \citep[Before:][-After]{bibtexBob2000} for citing within brackets with additional text: "(Before: Bob, 2000 -After)"
-%        \citealt{bibtexBob2000} for citing without brackets: "Bob, 2000"
-%        You can put more cites in each \cite tag by separating them with commas.
-% - For equation, find every details here: https://en.wikibooks.org/wiki/LaTeX/Mathematics
-% - For titles and stuff, the hierarchy goes \section{}, \subsection{}, \subsubsection{} and so forth...
-% - For bullet points or enumerations you can use:
-%       \begin{itemize}
-%           \item my first bullet point/enumeration
-%       \end{itemize}
-%       With replacing "itemize" by "enumerate" for enumeration.
 
@@
 \begin{itemize}
+
+    %TG: Change this induced thing
+
     \item \textbf{Biological induce dependence between characters:} this is the result of a link between two characters through development and/or biological function.
@@
 
+    %TG: This one can be reworded. Get a dichotomie a priori/a posteriori correlation. + add trade off in coding (more characters = more info but less independence)
     \item \textbf{Coding induced dependence between characters:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.

### 2017-09-18 Analysis/EffectCorrelation.Rmd
Updated results (95% done)

@@
 
+<!-- For generating the figure for the paper (without the legend) -->
+
+```{r, eval = FALSE, echo = FALSE}
+pdf("../Writing/Figures/RF_results_best.pdf",width = 8,height = 8)
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 2, what = "best",
+    legend = FALSE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 2, what = "best",
+    ylab = "RF for 150 taxa", NTS = TRUE)
+par(op)
+dev.off()
+```
+
 Comparison to 'best' tree (Triplets normalised distance)
@@
 par(op)
+```
+
+<!-- For generating the figure for the paper (without the legend) -->
+
+```{r, eval = FALSE, echo = FALSE}
+pdf("../Writing/Figures/Tr_results_best.pdf", width = 8, height = 8)
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 4, what = "best",
+    legend = FALSE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 4, what = "best", ylab = "Tr for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 4, what = "best",
+    ylab = "Tr for 150 taxa", NTS = TRUE)
+par(op)
+dev.off()
 ```
@@
 
+<!-- For generating the figure for the paper (without the legend) -->
+
+```{r, eval = FALSE, echo = FALSE}
+pdf("../Writing/Figures/RF_results_null.pdf", width = 8, height = 8)
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 2, what = "null",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 2, what = "null", ylab = "RF for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 2, what = "null",
+    ylab = "RF for 150 taxa", NTS = TRUE)
+par(op)
+dev.off()
+```
+
 Comparison to 'null' tree (Triplets normalised distance)
@@
 par(op)
+```
+
+<!-- For generating the figure for the paper (without the legend) -->
+
+```{r, eval = FALSE, echo = FALSE}
+pdf("../Writing/Figures/Tr_results_null.pdf", width = 8, height = 8)
+op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+plot.results.single(t25_list, metric = 4, what = "null",
+    legend = FALSE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 4, what = "null", ylab = "Tr for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 4, what = "null",
+    ylab = "Tr for 150 taxa", NTS = TRUE)
+par(op)
+dev.off()
 ```

### 2017-09-18 Functions/apply.test.R
updated apply test function (for proper publication standard outputs)

@@ -6,6 +6,7 @@
 #' @param metric which metric to use
 #' @param best which best tree (rand or norm)
 #' @param test which test (if NULL, simply summarises the distribution)
+#' @param convert.row.names whether to convert the row names into columns data (TRUE) or not (FALSE)
 #' 
 #' @examples
 #'
@@ -14,7 +15,7 @@
 #' @author Thomas Guillerme
 #' @export
 
-apply.test <- function(data, metric, best, test) {
+apply.test <- function(data, metric, best, test, convert.row.names = TRUE) {
 
     ## Combined
     combined <- TRUE
@@ -69,12 +70,30 @@ apply.test <- function(data, metric, best, test) {
         ## Make the names matrix
         names_matrix <- cbind(rep(parameter_names[1:3], each = 18), # taxa
                               rep(rep(parameter_names[4:6], each = 6), 3),  # character
-                              rep(parameter_names[7:8], 27),
-                              rep(parameter_names[9:11], 18)) # method
+                              rep(rep(parameter_names[7:8], each = 3), 9),
+                              rep(parameter_names[9:11], 18)) # scenario
         ## Combine all that
-        return(cbind(as.data.frame(names_matrix), as.data.frame(summary_mat)))
-    }
+        if(convert.row.names) {
+            ## Remove the t/c
+            names_matrix <- gsub("t", "", names_matrix)
+            names_matrix <- gsub("c", "", names_matrix)
+            ## Sorting in the Method/taxa/character/scenario format
+            results_out <- cbind(as.data.frame(names_matrix[,c(3,1,2,4)], stringsAsFactors= FALSE), as.data.frame(summary_mat))
+            ## Sorting the rows by method
+            results_out <- results_out[c(as.vector(combo_list[,1:9]), as.vector(combo_list[,10:18])),]
+            rownames(results_out) <- 1:54
+            ## Remove duplicated names
+            results_out[-c(1,28), 1] <- "" #method
+            results_out[-c(1,10,19,28,37,46), 2] <- "" #taxa
+            results_out[-seq(from = 1, to = 52, by = 3), 3] <- "" #characters
+            ## Adding the column names
+            colnames(results_out)[1:4] <- c("method", "taxa", "characters", "scenario")
 
+            return(results_out)
+        } else {
+            return(cbind(as.data.frame(names_matrix), as.data.frame(summary_mat)))
+        }
+    }
 
[…]

### 2017-09-18 Analysis/EffectCorrelationStats.Rmd
Updated stats part

--- 2017-07-27 Analysis/EffectCorrelationStats.Rmd
@@
 ```{r, eval = TRUE, message = FALSE}
+library(knitr)
 library(dispRity)
@@
 
+Pooled results
+=================
+
+Individual results
+=================
+
+We can also analyse the effect of each parameters combination separately:
+
 Summary statistics
------------------------------------
+-----------------
 
@@
 ## Running the pairwise Wilcox tests:
-(sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL))
-(sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL))
-(sum_RF_null <- apply.test(whole_data, "RF", "null", NULL))
-(sum_Tr_null <- apply.test(whole_data, "Triples", "null", NULL))
+sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL)
+sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL)
+sum_RF_null <- apply.test(whole_data, "RF", "null", NULL)
+sum_Tr_null <- apply.test(whole_data, "Triples", "null", NULL)
+```
 
+```{r, echo = FALSE}
+## Setting the table captions
+captions_summary <- paste("Summary statistics of the normalised", rep(c("Robinson-Fould", "Triplets"), 2), "distance to the", c(rep("best", 2), rep("null", 2)), "tree.")
+
+## Printing the tables
+kable(sum_RF_best, row.names = FALSE, caption = captions_summary[1])
+kable(sum_Tr_best, row.names = FALSE, caption = captions_summary[2])
+kable(sum_RF_null, row.names = FALSE, caption = captions_summary[3])
+kable(sum_Tr_null, row.names = FALSE, caption = captions_summary[4])
+```
+
+<!-- Table for publication -->
+
+```{r, echo = FALSE, eval = FALSE}
+library(xtable)
+## Wrapper function
+make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE) {
+
+  table <- xtable(table, digit = digit, caption = caption, label = label)
+
+  if(longtable == TRUE) {
+    print(table, tabular.environment = 'longtable', floating = FALSE, include.rownames = FALSE)
+  } else {
+    print(table, include.rownames = FALSE)
+  }
+}
+
+## Labels
+labels_summary <- paste0("Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
+
+## Tables
+make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1])
+make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2])
+make.xtable(sum_RF_null, caption = captions_summary[3], label = labels_summary[3])
+make.xtable(sum_Tr_null, caption = captions_summary[4], label = labels_summary[4])
 ```
@@
 
-## Summarise the results
-results <- matrix(nrow = 4, data = c(
-                round(unlist(lapply(normality, `[[`, 1)), digit = 3),
-                round(unlist(lapply(normality, `[[`, 2)), digit = 3),
-                round(unlist(lapply(homoscedasticity, `[[`, 1)), digit = 3),
-                round(unlist(lapply(homoscedasticity, `[[`, 2)), digit = 3),
-                round(unlist(lapply(homoscedasticity, `[[`, 3)), digit = 3)),
-            dimnames = list(
-                c("RF best", "Triples best", "RF null", "Triples null"),
-                c("W", "p.value", "Bartlett's K-squared", "df", "p.value")))
+# ## Summarise the results
+results <- t(rbind(sapply(normality, function(x) x[c(1,2)]), 
+                   sapply(homoscedasticity, function(x) x[c(1,2,3)])))
+results <- matrix(as.numeric(results), ncol = 5)
+colnames(results) <- c("W", "p.value", "Bartlett's K^2", "df", "p.value")
+rownames(results) <- c("RF best", "Triples best", "RF null", "Triples null")
+results <- round(results, 3)
 
-## Check assumption's validities
-validity <- apply(results, 1,
-    function(X) ifelse((X[2] >= 0.05 && X[5] >= 0.05), TRUE, FALSE))
+# ## Check assumption's validities
+check.validity <- function(X) {ifelse(X[2] >= 0.05 && X[5] >= 0.05, TRUE, FALSE)}
+validity <- apply(results, 1, check.validity)
 
 ## Report results
-(results <- cbind(as.data.frame(results), validity))
+results <- cbind(as.data.frame(results), validity)
+```
 
+```{r, echo = FALSE}
+kable(results, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
 ```
@@
 We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
-A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don't overlap (significant effect of one scenario on topology).
+A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don t overlap (significant effect of one scenario on topology).
 Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).
@@
 ## Running the pairwise Bhattacharrya Coefficient analysis:
-(bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff))
-(bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff))
-(bc_RF_null <- apply.test(whole_data, "RF", "null", bhatt.coeff))
-(bc_Tr_null <- apply.test(whole_data, "Triples", "null", bhatt.coeff))
+bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff)
+bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff)
+bc_RF_null <- apply.test(whole_data, "RF", "null", bhatt.coeff)
+bc_Tr_null <- apply.test(whole_data, "Triples", "null", bhatt.coeff)
+
+## Tables
+kable(bc_RF_best, row.names = FALSE, digit = 3)
+kable(bc_RF_best, row.names = FALSE, digit = 3)
+kable(bc_RF_best, row.names = FALSE, digit = 3)
+kable(bc_RF_best, row.names = FALSE, digit = 3)
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_bc <- paste0("Bhattacharrya Coefficient (",rep(c("Robinson-Fould distance", "Triplets distance"), 2) ," from the ",c(rep("best", 2), rep("null", 2)), " tree)" , " between the different scenarios.")
+
+## Printing the tables
+kable(bc_RF_best, row.names = FALSE, caption = captions_bc[1], digit = 3)
+kable(bc_Tr_best, row.names = FALSE, caption = captions_bc[2], digit = 3)
+kable(bc_RF_null, row.names = FALSE, caption = captions_bc[3], digit = 3)
+kable(bc_Tr_null, row.names = FALSE, caption = captions_bc[4], digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, echo = FALSE, eval = FALSE}
+## Labels
+labels_bc <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
+
+## Tables
+make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1])
+make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2])
+make.xtable(bc_RF_null, caption = captions_bc[3], label = labels_bc[3])
+make.xtable(bc_Tr_null, caption = captions_bc[4], label = labels_bc[4])
 ```
@@
 
-```{r}
+```{r, warning = FALSE}
 ## Running the pairwise Wilcox tests:
@@
 ## Applying Bonferroni correction
-wx_RF_best[,c(2,4,6)] <- p.adjust(wx_RF_best[,c(2,4,6)], method = "bonferroni")
-wx_Tr_best[,c(2,4,6)] <- p.adjust(wx_Tr_best[,c(2,4,6)], method = "bonferroni")
-wx_RF_null[,c(2,4,6)] <- p.adjust(wx_RF_null[,c(2,4,6)], method = "bonferroni")
-wx_Tr_null[,c(2,4,6)] <- p.adjust(wx_Tr_null[,c(2,4,6)], method = "bonferroni")
+for(column in c(5,7,9)) {
+  wx_RF_best[,column] <- p.adjust(wx_RF_best[,column], method = "bonferroni")
+  wx_Tr_best[,column] <- p.adjust(wx_Tr_best[,column], method = "bonferroni")
+  wx_RF_null[,column] <- p.adjust(wx_RF_null[,column], method = "bonferroni")
+  wx_Tr_null[,column] <- p.adjust(wx_Tr_null[,column], method = "bonferroni")
+}
+```
 
-## Print the results
-wx_RF_best
-wx_Tr_best
-wx_RF_null
-wx_Tr_null
+```{r, echo = FALSE}
+## Setting the table captions
+captions_wx <- paste0("Non-parametric Wilcoxon test (",rep(c("Robinson-Fould distance", "Triplets distance"), 2) ," from the ",c(rep("best", 2), rep("null", 2)), " tree)" , " between the different scenarios (p-values corrected using Bonferroni-Holm correction).")
+
+## Printing the tables
+kable(wx_RF_best, row.names = FALSE, caption = captions_wx[1], digit = 3)
+kable(wx_Tr_best, row.names = FALSE, caption = captions_wx[2], digit = 3)
+kable(wx_RF_null, row.names = FALSE, caption = captions_wx[3], digit = 3)
+kable(wx_Tr_null, row.names = FALSE, caption = captions_wx[4], digit = 3)
 ```
+
+<!-- Table for publication -->
+
+```{r, echo = FALSE, eval = FALSE}
+## Labels
+labels_wx <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
+
+## Tables
+make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1])
+make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2])
+make.xtable(wx_RF_null, caption = captions_wx[3], label = labels_wx[3])
+make.xtable(wx_Tr_null, caption = captions_wx[4], label = labels_wx[4])
+```

### 2017-09-18 Functions/pool.distribution.R
pooling a distribution

@@ -1,0 +1,73 @@
+#' @title Pool a distribution for one parameter
+#'
+#' @description Pool a distribution for one parameter
+#'
+#' @param data the whole data
+#' @param param which parameter
+#' @param metric which metric
+#' @param best which comparison
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+pool.distribution <- function(data, param, metric, best) {
+
+    ## Get the data for the right metric and best
+
+    ## Getting the metric
+    if(metric == "RF") metric <- 2
+    if(metric == "Tr") metric <- 4
+
+    ## Getting all the data for one metric
+    NTS_all <- lapply(data, lapply.get.metric, metric)
+
+    ## Extract the right scenarios (i.e. drop rand/norm and true)
+    if(best == "norm") {
+        ## Extra the right base (norm and rand)
+        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 1)
+        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(2,3,4))
+    } else {
+        ## Extra the right base (norm and rand)
+        NTS_base <- lapply(NTS_all, lapply, lapply, `[[`, 2)
+        NTS <- lapply(NTS_base, lapply, lapply, `[`, c(1,2,3))
+    }
+
+    ## Extracting the right distribution
+
+    ## Checking if the parameter is at the first level (number of taxa)
+    param_level1 <- match(param, names(NTS))
+    if(!is.na(param_level1)) {
+        ## Extracting the distribution
+        return(as.vector(unlist(NTS[[param_level1]])))
+    } else {
+        ## Checking if the parameter is at the second level (number of characters)
+        param_level2 <- match(param, names(NTS[[1]]))
+        if(!is.na(param_level2)) {
+            ## Extracting the distribution
+            distribution <- sapply(NTS, function(x, param_level2) x[[param_level2]], param_level2, simplify = FALSE)
+            return(as.vector(unlist(distribution)))
+        } else {
+            ## Checking if the parameter is at the third level (method)
+            param_level3 <- match(param, names(NTS[[1]][[1]]))
+            if(!is.na(param_level3)) {
+                ## Extracting the distribution
+                distribution <- lapply(NTS, lapply, `[[`, param_level3)
+                return(as.vector(unlist(distribution)))
+            } else {
+                ## Checking if the parameter is at the fourth level (scenario)
[…]

### 2017-09-18 Functions/plot.results.R
edited las

@@ -71,7 +71,7 @@ plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, N
     
 
     if(what == "null") {
-        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), ...,
+        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), las = 1, ...,
             ## c100
             unlist(taxa_list$c100$bayesian$rand$maxi[,metric]), unlist(taxa_list$c100$parsimony$rand$maxi[,metric]),
             unlist(taxa_list$c100$bayesian$rand$mini[,metric]), unlist(taxa_list$c100$parsimony$rand$mini[,metric]),
@@ -103,7 +103,7 @@ plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, N
     }
 
     if(what == "best") {
-        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), ...,
+        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), las = 1, ...,
             ## c100
             unlist(taxa_list$c100$bayesian$norm$maxi[,metric]), unlist(taxa_list$c100$parsimony$norm$maxi[,metric]),
             unlist(taxa_list$c100$bayesian$norm$mini[,metric]), unlist(taxa_list$c100$parsimony$norm$mini[,metric]),
@@ -135,7 +135,7 @@ plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, N
     }
 
     if(what == "true") {
-        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols[1:4], 3), ...,
+        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols[1:4], 3), las = 1, ...,
             ## c100
             unlist(taxa_list$c100$bayesian$norm$true[,metric]), unlist(taxa_list$c100$parsimony$norm$true[,metric]),
             unlist(taxa_list$c100$bayesian$rand$true[,metric]), unlist(taxa_list$c100$parsimony$rand$true[,metric]),
[…]

### 2017-09-18 Analysis/EffectCorrelationStats.Rmd
stats pooled draft

@@
 =================
+
+Effect of character correlation per scenario
+-----------------
+
+```{r}
+## Pulling the data per scenario
+max_RF_norm <- pool.distribution(whole_data, param = "maxi", metric = "RF", best = "norm")
+min_RF_norm <- pool.distribution(whole_data, param = "mini", metric = "RF", best = "norm")
+rand_RF_norm <- pool.distribution(whole_data, param = "rand", metric = "RF", best = "norm")
+scenario_RF_norm <- matrix(data = c(max_RF_norm, min_RF_norm, rand_RF_norm),
+                           ncol = 3, dimnames = list(c(), c("max", "min", "rand")))
+
+max_RF_null <- pool.distribution(whole_data, param = "maxi", metric = "RF", best = "rand")
+min_RF_null <- pool.distribution(whole_data, param = "mini", metric = "RF", best = "rand")
+rand_RF_null <- pool.distribution(whole_data, param = "norm", metric = "RF", best = "rand")
+scenario_RF_null <- matrix(data = c(max_RF_null, min_RF_null, rand_RF_null),
+                           ncol = 3, dimnames = list(c(), c("max", "min", "norm")))
+
+max_Tr_norm <- pool.distribution(whole_data, param = "maxi", metric = "Tr", best = "norm")
+min_Tr_norm <- pool.distribution(whole_data, param = "mini", metric = "Tr", best = "norm")
+rand_Tr_norm <- pool.distribution(whole_data, param = "rand", metric = "Tr", best = "norm")
+scenario_Tr_norm <- matrix(data = c(max_Tr_norm, min_Tr_norm, rand_Tr_norm),
+                           ncol = 3, dimnames = list(c(), c("max", "min", "rand")))
+
+max_Tr_null <- pool.distribution(whole_data, param = "maxi", metric = "Tr", best = "rand")
+min_Tr_null <- pool.distribution(whole_data, param = "mini", metric = "Tr", best = "rand")
+rand_Tr_null <- pool.distribution(whole_data, param = "norm", metric = "Tr", best = "rand")
+scenario_Tr_null <- matrix(data = c(max_Tr_null, min_Tr_null, rand_Tr_null),
+                           ncol = 3, dimnames = list(c(), c("max", "min", "norm")))
+
+op <- par(mfrow = c(2,2))#, mar = c(2, 4, 1, 2))
+cols <- c("red", "orange", "green3", "lightgreen", "blue", "lightblue")
+
+boxplot(scenario_RF_norm, las = 1, horizontal = TRUE, xlab = "")
+
+plot.results.single(t25_list, metric = 2, what = "best",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    axislab = FALSE)
+plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
+    xlab = "", axislab = FALSE, NTS = TRUE)
+plot.results.single(t150_list, metric = 2, what = "best",
+    ylab = "RF for 150 taxa", NTS = TRUE)
+par(op)
+
+boxplot(scenario_RF_norm, col = )
+```
+
+Effect of character correlation per number of character
+-----------------
+
+Effect of character correlation per number of taxa
+-----------------
+
+Effect of character correlation per methods
+-----------------
 

### 2017-09-19 Analysis/Character_difference/Modifying_matrices.Rmd
Modifying matrices figure updated

@@
 
-<!-- ```{r, fig.width=12, fig.height=12}
+```{r, fig.width=12, fig.height=12, eval = FALSE, echo = FALSE}
+pdf("../../Writing/Figures/Modif_matrix.pdf", width = 7.2, height = 4.5)
 par(mfrow = c(2, 3), bty = "n", mar = c(1,1,1,1))
@@
 plot(char_diff_matrix_rand, type = "matrix", main = "E - Randomised matrix", legend = FALSE)
+dev.off()
 ```
- -->

### 2017-09-19 Functions/make.xtable.R
xtable wrapper function

@@ -1,0 +1,27 @@
+#' @title Making the xtable for publication
+#'
+#' @description Making the xtable for publication
+#'
+#' @param table the table to print
+#' @param digit the digit rounding
+#' @param caption the caption
+#' @param label the label (in LaTeX: \ref{label})
+#' @param longtable whether to use longtable
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE) {
+  
+  table <- xtable(table, digit = digit, caption = caption, label = label)
+
+  if(longtable == TRUE) {
+    print(table, tabular.environment = 'longtable', floating = FALSE, include.rownames = FALSE)
+  } else {
+    print(table, include.rownames = FALSE)
+  }
+}

### 2017-09-19 Functions/apply.test.R
pairwise tests wrapper function

@@ -162,3 +162,118 @@ apply.test <- function(data, metric, best, test, convert.row.names = TRUE) {
 
     return(test_results)
 }
+
+
+#' @title Applies a pairwise test to a (pooled) table or a list
+#'
+#' @description Applies a pairwise test to a (pooled) table or a list
+#'
+#' @param table a table or a list
+#' @param test the test to apply
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+apply.pair.test <- function(table, test) {
+    ## Check if table is a table or a list
+    if(class(table) == "matrix" || class(table) == "data.frame") {
+        data_list <- unlist(apply(table, 2, list), recursive = FALSE)
+    } else {
+        data_list <- table
+    }
+
+    ## Get the pairwise combinations
+    combinations <- combn(1:length(data_list), 2)
+    combo_names <- apply(combinations, 2, function(X, data_list) return(paste(names(data_list)[X[1]], names(data_list)[X[2]], sep = ":")), data_list)
+
+    ## Run the pairwise test
+    apply.test <- function(combn_col, data_list, test) {
+        return(test(data_list[[combn_col[1]]], data_list[[combn_col[2]]]))
+    }
+    results <- apply(combinations, 2, apply.test, data_list, test)
+
+    ## Output the results
+    if(class(results) == "numeric") {
+        names(results) <- combo_names
+        return(results)
+    } else {
+        remove.names <- function(X) {names(X) <- NULL; return(X)}
+        results <- lapply(results, lapply, remove.names)
+        names(results) <- combo_names
+        return(list("statistic" = unlist(lapply(results, `[[`, 1)), "p.value" = unlist(lapply(results, `[[`, 3))))
+    }
+}
+
+
+#' @title Applies a bhatt.coeff and a wilcox.test to a list of distributions and outputs a table
+#'
+#' @description Applies bhatt.coeff and a wilcox.test to a list of distributions and outputs a table
+#'
+#' @param distributions a list of distribution tables
+#' @param digit how many digits to display
+#' @param metric.label the metrics labels (can be missing)
+#' @param comp.label the comparisons labels (can be missing)
+#' @param correction which correction to apply
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+#' 
[…]

### 2017-09-19 Analysis/EffectCorrelationStats.Rmd
streamlined stats analysis

--- 2017-09-18 Analysis/EffectCorrelationStats.Rmd
@@
 ```{r, eval = TRUE, message = FALSE}
-library(knitr)
 library(dispRity)
@@
 
+Effect of character correlation per scenario
+-----------------
+
+```{r, warning = FALSE}
+## Pooling the data per scenarios
+scenario_RF_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
+                               metric = "RF", best = "norm")
+scenario_Tr_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
+                               metric = "Tr", best = "norm")
+scenario_RF_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
+                               metric = "RF", best = "rand")
+scenario_Tr_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
+                               metric = "Tr", best = "rand")
+
+## Preparing the table labels
+comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
+metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
+
+## Combining the data into a table
+scenarios <- pool.table(list(scenario_RF_norm, scenario_Tr_norm, scenario_RF_null, scenario_Tr_null),
+                    comp.label = comp.labels, metric.label = metric.labels, label.param = "scenario")
+```
+
+```{r, echo = FALSE}
+library(knitr)
+## Setting the table captions
+captions_pooled_scenarios <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled scenarios.")
+
+## Printing the tables
+kable(scenarios, row.names = FALSE, caption = captions_pooled_scenarios)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+library(xtable)
+make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Tab_pooledscenarios")
+```
+
+### Testing the effect of the scenario on the NTS
+
+```{r}
+## Applying an lm to all the data
+results <- test.effect.aov(whole_data, formula = NTS ~ scenario, fun = lm)
+```
+
+Checking the model assumptions
+
+```{r, echo = FALSE}
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```
+
+Applying a Bhattacharrya Coefficient and Wilcoxon test
+
+```{r, warning = FALSE}
+## Summarise into a table
+scenario_list <- list(scenario_RF_norm, scenario_Tr_norm, scenario_RF_null, scenario_Tr_null)
+comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
+metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
+test_table <- pair.test.table(scenario_list, comp.label = comp.labels, metric.label = metric.labels)
+```
+
+```{r, echo = FALSE}
+## Caption
+cap_pool_scen_test <- paste("Difference between the pooled scenarios.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+
+## Printing the tables
+kable(test_table, row.names = FALSE, caption = cap_pool_scen_test)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(test_table, caption = cap_pool_scen_test, label = "Tab_pooledscenarios_test")
+```
+
+Effect of character correlation per number of character
+-----------------
+
+Effect of character correlation per number of taxa
+-----------------
+
+Effect of character correlation per methods
+-----------------
+
 Individual results
@@
 ```{r}
-## Running the pairwise Wilcox tests:
+## Summarising the data
 sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL)
@@
 ```{r, echo = FALSE, eval = FALSE}
-library(xtable)
-## Wrapper function
-make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE) {
-
-  table <- xtable(table, digit = digit, caption = caption, label = label)
-
-  if(longtable == TRUE) {
-    print(table, tabular.environment = 'longtable', floating = FALSE, include.rownames = FALSE)
-  } else {
-    print(table, include.rownames = FALSE)
-  }
-}
-
 ## Labels
@@
 ```{r}
-## getting the aov data
-aov_data_RF_best <- get.aov.data(whole_data, "RF", "norm", combined = TRUE)
-aov_data_Tr_best <- get.aov.data(whole_data, "Triples", "norm", combined = TRUE)
-aov_data_RF_null <- get.aov.data(whole_data, "RF", "null", combined = TRUE)
-aov_data_Tr_null <- get.aov.data(whole_data, "Triples", "null", combined = TRUE)
-
-## Models with normalised tree score function of scenario (with nestedness)
-model_RF_best <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
-                      data = aov_data_RF_best)
-model_Tr_best <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
-                      data = aov_data_Tr_best)
-model_RF_null <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
-                      data = aov_data_RF_null)
-model_Tr_null <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
-                      data = aov_data_Tr_null)
-
-## Testing model assumptions
-## Normality of the residuals
-norm_RF_best <- shapiro.test(residuals(model_RF_best))
-norm_Tr_best <- shapiro.test(residuals(model_Tr_best))
-norm_RF_null <- shapiro.test(residuals(model_RF_null))
-norm_Tr_null <- shapiro.test(residuals(model_Tr_null))
-normality <- list(norm_RF_best, norm_Tr_best, norm_RF_null, norm_Tr_null)
-
-## Variance homoscedasticity
-homo_RF_best <- bartlett.test(NTS ~ scenario, data = aov_data_RF_best)
-homo_Tr_best <- bartlett.test(NTS ~ scenario, data = aov_data_Tr_best)
-homo_RF_null <- bartlett.test(NTS ~ scenario, data = aov_data_RF_null)
-homo_Tr_null <- bartlett.test(NTS ~ scenario, data = aov_data_Tr_null)
-homoscedasticity <- list(homo_RF_best, homo_Tr_best, homo_RF_null, homo_Tr_null)
-
-# ## Summarise the results
-results <- t(rbind(sapply(normality, function(x) x[c(1,2)]), 
-                   sapply(homoscedasticity, function(x) x[c(1,2,3)])))
-results <- matrix(as.numeric(results), ncol = 5)
-colnames(results) <- c("W", "p.value", "Bartlett's K^2", "df", "p.value")
-rownames(results) <- c("RF best", "Triples best", "RF null", "Triples null")
-results <- round(results, 3)
-
-# ## Check assumption's validities
-check.validity <- function(X) {ifelse(X[2] >= 0.05 && X[5] >= 0.05, TRUE, FALSE)}
-validity <- apply(results, 1, check.validity)
-
-## Report results
-results <- cbind(as.data.frame(results), validity)
+## Applying an lm to all the data with error
+results <- test.effect.aov(whole_data,
+  formula = NTS ~ scenario + (1|method) + (1|taxa) + (1|character), fun = lmer)
 ```
 
+Checking the model assumptions
+
 ```{r, echo = FALSE}
-kable(results, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
 ```
@@
 ```{r, eval = FALSE, echo = FALSE}
-xtable(results, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
 ```

### 2017-09-19 Analysis/EffectCorrelationStats.Rmd
Stats updated

@@
 
-Effect of character correlation
-===============
+# Effect of character correlation
 
-Loading the functions and the data
------------------------------------
+## Loading the functions and the data
 
@@
 
-Pooled results
-=================
+# Pooled results
 
-Effect of character correlation per scenario
------------------
+## Effect of character correlation per scenario
 
@@
 
-Effect of character correlation per number of character
------------------
+## Effect of character correlation per number of character
 
-Effect of character correlation per number of taxa
------------------
+```{r, warning = FALSE}
+## Pooling the data per character
+character_RF_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
+                               metric = "RF", best = "norm")
+character_Tr_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
+                               metric = "Tr", best = "norm")
+character_RF_null <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
+                               metric = "RF", best = "rand")
+character_Tr_null <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
+                               metric = "Tr", best = "rand")
 
-Effect of character correlation per methods
------------------
+## Preparing the table labels
+comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
+metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
 
-Individual results
-=================
+## Combining the data into a table
+characters <- pool.table(list(character_RF_norm, character_Tr_norm, character_RF_null, character_Tr_null),
+                    comp.label = comp.labels, metric.label = metric.labels, label.param = "character")
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_pooled_characters <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of characters.")
+
+## Printing the tables
+kable(characters, row.names = FALSE, caption = captions_pooled_characters)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(characters, caption = captions_pooled_characters, label = "Tab_pooledcharacters")
+```
+
+### Testing the effect of the number of characters on the NTS
+
+```{r}
+## Applying an lm to all the data
+results <- test.effect.aov(whole_data, formula = NTS ~ character, fun = lm)
+```
+
+Checking the model assumptions
+
+```{r, echo = FALSE}
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```
+
+Applying a Bhattacharrya Coefficient and Wilcoxon test
+
+```{r, warning = FALSE}
+## Summarise into a table
+character_list <- list(character_RF_norm, character_Tr_norm, character_RF_null, character_Tr_null)
+comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
+metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
+test_table <- pair.test.table(character_list, comp.label = comp.labels, metric.label = metric.labels)
+```
+
+```{r, echo = FALSE}
+## Caption
+cap_pool_char_test <- paste("Difference between the pooled number of characters.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+
+## Printing the tables
+kable(test_table, row.names = FALSE, caption = cap_pool_char_test)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(test_table, caption = cap_pool_char_test, label = "Tab_pooledscharacters_test")
+```
+
+## Effect of character correlation per number of taxa
+
+```{r, warning = FALSE}
+## Pooling the data per taxa
+taxa_RF_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
+                               metric = "RF", best = "norm")
+taxa_Tr_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
+                               metric = "Tr", best = "norm")
+taxa_RF_null <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
+                               metric = "RF", best = "rand")
+taxa_Tr_null <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
+                               metric = "Tr", best = "rand")
+
+## Preparing the table labels
+comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
+metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
+
+## Combining the data into a table
+taxa <- pool.table(list(taxa_RF_norm, taxa_Tr_norm, taxa_RF_null, taxa_Tr_null),
+                    comp.label = comp.labels, metric.label = metric.labels, label.param = "taxa")
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_pooled_taxa <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of taxa.")
+
+## Printing the tables
+kable(taxa, row.names = FALSE, caption = captions_pooled_taxa)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(taxas, caption = captions_pooled_taxas, label = "Tab_pooledtaxa")
+```
+
+### Testing the effect of the method on the NTS
+
+```{r}
+## Applying an lm to all the data
+results <- test.effect.aov(whole_data, formula = NTS ~ taxa, fun = lm)
+```
+
+Checking the model assumptions
+
+```{r, echo = FALSE}
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```
+
+Applying a Bhattacharrya Coefficient and Wilcoxon test
+
+```{r, warning = FALSE}
+## Summarise into a table
+taxa_list <- list(taxa_RF_norm, taxa_Tr_norm, taxa_RF_null, taxa_Tr_null)
+comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
+metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
+test_table <- pair.test.table(taxa_list, comp.label = comp.labels, metric.label = metric.labels)
+```
+
+```{r, echo = FALSE}
+## Caption
+cap_pool_tax_test <- paste("Difference between the pooled number of taxa.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+
+## Printing the tables
+kable(test_table, row.names = FALSE, caption = cap_pool_tax_test)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(test_table, caption = cap_pool_tax_test, label = "Tab_pooledstaxa_test")
+```
+
+## Effect of character correlation per methods
+
+```{r, warning = FALSE}
+## Pooling the data per taxa
+method_RF_norm <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
+                               metric = "RF", best = "norm")
+method_Tr_norm <- multi.pool(whole_data, param =c("bayesian", "parsimony"),
+                               metric = "Tr", best = "norm")
+method_RF_null <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
+                               metric = "RF", best = "rand")
+method_Tr_null <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
+                               metric = "Tr", best = "rand")
+
+## Preparing the table labels
+comp.labels <- c("Best", rep("", 3), "Null", rep("", 3))
+metric.labels <- rep(c("RF", "", "Tr", ""), 2)
+
+## Combining the data into a table
+methods <- pool.table(list(method_RF_norm, method_Tr_norm, method_RF_null, method_Tr_null),
+                    comp.label = comp.labels, metric.label = metric.labels, label.param = "method")
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_pooled_methods <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled methods.")
+
+## Printing the tables
+kable(methods, row.names = FALSE, caption = captions_pooled_methods)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(methods, caption = captions_pooled_methods, label = "Tab_pooledmethod")
+```
+
+### Testing the effect of the method on the NTS
+
+```{r}
+## Applying an lm to all the data
+results <- test.effect.aov(whole_data, formula = NTS ~ method, fun = lm)
+```
+
+Checking the model assumptions
+
+```{r, echo = FALSE}
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```
+
+Applying a Bhattacharrya Coefficient and Wilcoxon test
+
+```{r, warning = FALSE}
+## Summarise into a table
+methods_list <- list(method_RF_norm, method_Tr_norm, method_RF_null, method_Tr_null)
+comp.labels <- c("Best", rep("", 3), "Null", rep("", 3))
+metric.labels <- rep(c("RF", "", "Tr", ""), 2)
+test_table <- pair.test.table(methods_list, comp.label = comp.labels, metric.label = metric.labels)
+```
+
+```{r, echo = FALSE}
+## Caption
+cap_pool_met_test <- paste("Difference between the pooled methods.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+
+## Printing the tables
+kable(test_table, row.names = FALSE, caption = cap_pool_met_test)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(test_table, caption = cap_pool_met_test, label = "Tab_pooledsmethods_test")
+```
+
+# Individual results
 
@@
 
-Summary statistics
------------------
+## Summary statistics
 
@@
 
-Nested anova
------------------------------------
+## Nested anova
 
@@
 
-Distributions probability overlap (Bhattacharrya Coefficient)
------------------------------------
+## Distributions probability overlap (Bhattacharrya Coefficient)
 
@@
 
-Non-parametric pairwise comparisons
------------------------------------
+## Non-parametric pairwise comparisons
 

### 2017-09-19 Writing/draft.tex
re-edits to intro + methods (removed comparisons to null!)

@@
 The last six years have seen a fantastic expansion of the use of discrete morphological data in macroevolutionary studies.
-Traditionally confined to ``classic'' cladistic studies, these methods have recently been widely expanded and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep[e.g.][]{pyrondivergence2011,ronquista2012}.
-In parallel, much development has also been done on pre-phylogenetic analysis \citep[e.g.
+Traditionally confined to ``classic'' cladistic studies in palaeontology, these methods have recently been widely expanded and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep[e.g.][]{pyrondivergence2011,ronquista2012}.
+In parallel, much development has also been done on the pre-phylogenetic analysis \citep[e.g.
 data collection;][]{morphobank} as well as on post-phylogenetic ones \citep[e.g.
 morphological disparity analysis;][]{Close2015,Claddis}.
-Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011,simoes2017giant}; the inference method \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.
+Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011,simoes2017giant}; the topology inference method \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.
 
-One aspect of the data, however, is often neglected: the correlation between morphological characters.
-In fact, the assumption of character independence is at the base of phylogenetic inference and as been mentioned or implied throughout the history of phylogenetic methods \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-Nowadays, this assumption of character independence is nearly always accepted as a fact and rarely tested or questioned in empirical datasets.
-However - especially for discrete morphological data - this assumption of independence is probably violated more often than not due to the sheer nature of phylogenetic data.
-In practice, dependence between characters can be induced in at least three different ways:
+One of such caveats, however, is often neglected: the correlation between morphological characters.
+The assumption of character independence is at the base of phylogenetic inference and as been mentioned or implied throughout the history of phylogenetic methods \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
+Nowadays, this assumption of character independence is nearly always accepted as a fact and rarely tested or even questioned in empirical datasets.
+However - especially for discrete morphological data - this assumption of independence is probably violated more often than not due to the sheer nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
+In practice, characters dependence can come from at least three different ways:
 
@@
 
-    %TG: Change this induced thing
-
-    \item \textbf{Biological induce dependence between characters:} this is the result of a link between two characters through development and/or biological function.
-    These links are readily understood through developmental biology or understanding of shared function between features \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
+    \item \textbf{Biological dependence:} this is the result of a link between two characters through their development and/or biological function.
+    These links are readily studied through developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
     One trivial example would be the link between lower and upper molar characters in mammals: one character describing a feature of a lower molar will be expected to be correlated to another character describing another feature of the associated upper molar.
 
-    \item \textbf{Evolutionary induced dependence between characters:} this is the result of sets of characters co-evolving along branches in the phylogeny.
-    The character's correlation is thus due to shared evolutionary history of these characters.
+    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving along the same branch in a phylogeny: the character's correlation is thus due to their shared evolutionary history.
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
@@
 
-    %TG: This one can be reworded. Get a dichotomie a priori/a posteriori correlation. + add trade off in coding (more characters = more info but less independence)
-    \item \textbf{Coding induced dependence between characters:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
-    These links can be due to methodological errors: for example one could define a character describing a particular feature on a left molar as well as another character describing the same feature on a right molar.
-    It is worth noting, however, that these dependences could also be due to the nature of the available data, especially in palaeontology.
+    \item \textbf{Coding dependence:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
+    One such source of correlation could be due to methodological errors: for example by defining a character describing a particular feature on a left and right molar.
+    It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
     For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
@@
 \noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding!
+The three types sources are however dissociated by later (coding dependence) being detected prior to phylogenetic inference and the two other ones (biological and evolutionary) being only detected and/or tested \textit{a posteriori} based on a phylogenetic hypothesis.
 
-As mentioned above, these sources of dependence between characters are well studied in biology.
+These sources of dependence between characters are well studied in biology.
 Biological and evolutionary dependences are inherent parts to evo-devo and maroevolution and best practices to avoid coding induced dependences are commonly known and applied.
-However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inferences softwares that are blind to these distinctions.
+However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inferences software that are blind to these distinctions.
 This introduces a new, less studied, source of character dependence:
 
-\noindent \textbf{Software induced correlation between characters:} this is the result how software actually interprets the differences between characters.
-I.e.
-the vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different/similar tokens) leading to dependence between characters.
-For example, consider the following matrix containing four cetartiodactyles - say a pig (e.g.
-\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four characters - say (\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley:
+\noindent \textbf{Correlation between characters detected by the software}: this is the result how software actually interprets the differences between characters.
+The vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different/similar tokens).
+Therefore a great number of characters and - traditionally - a few number of tokens can easily lead to dependence between characters.
+For example, if we consider the following matrix containing four cetartiodactyles - say a pig (e.g.
+\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley:
 
@@
 
-If the three induced sources of correlation are fairly well studied, their actual interpretation in phylogenetic inference, has, to our knowledge, never been studied through a simulation protocol.
-In this study, we formally assess the effect of discrete character's correlation using simulated and empirical data.
-We propose a new metric to measure the difference between characters (as a proxy for these three sources of correlation) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences/similarities between characters.
-We found that @@@
-
-% However, only several attempts have been made on the actual effect of characters correlation on phylogenetic inference with two notable exceptions of \cite{Davalos01072014} on empirical data and \cite{ZouConvergence} on molecular data.
-%TG: put that somwhere
+If the three induced sources of correlation are fairly well studied, their actual interpretation by phylogenetic inference software, has, to our knowledge, never been studied through a simulation protocol (although they have been explored empirically for morphological data; \citealt{Davalos01072014}; or molecular data; \citealt{ZouConvergence}).
+In this study, we formally assess the effect of discrete character's correlation using simulated data.
+We propose a new metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences/similarities between characters.
+We found that overall, there is a detectable effect of character correlation on topology where an increase in character dependence results in a decrease in the ability to recover the correct topology.
+These results, however, vary in magnitude depending on the size of matrix and the inference method used.
 
@@
 \begin{enumerate}
-    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
-    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by duplicating characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
+    \item \textbf{Simulating matrices}: we simulated a discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
+This step resulted in 9 matrices.
+    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modified the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
     We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character difference to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
+This step resulted in 36 matrices.
     \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
+This step resulted in 72 topologies.
     \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on topology.
 \end{enumerate}
-Each steps are described below in more details, as well as our proposed definition of the Character Difference metric.
+Each step was replicated 40 times and are described below in more details, as well as our proposed definition for measuring the difference between characters.
 
@@
 \label{CD_description}
-We define characters as being correlated if they give the same phylogenetic information.
-In order to measure this similitude in phylogenetic information, we propose a new distance metric to measure the difference between two characters:
+To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information.
+In order to measure this, we propose a new distance metric to measure the difference between two characters:
 
@@
 \noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
-$CD$ is a continuous distance metric valid in a mathematical sense (see Supplementary Material 1).
+$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration Supplementary Material 1).
 Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we calculated the difference between character states in a qualitative way.
@@
 Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
-We standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
+We standardised each character by arbitrarily modifying their character state tokens by order of appearance.
 In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
@@
 
-When the character difference is equal to $0$, it means that characters convey the same phylogenetic signal.
-When the character difference is equal to $1$ it means it conveys the most different signal.
+When the character difference null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
+When the character difference maximal ($1$) it means it conveys the most different signal (i.e. characters are uncorrelated).
+It is important to stress that a character difference of $0$ (i.e. the same phylogenetic signal) does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal) .
 For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
+Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different phylogenetic interpretation.
 
@@
 To simulate the matrices we applied a protocol very similar to \citep{Guillerme2016146}.
-First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform ($0$,$1$) distribution maintaining $\lambda$ $>$ $\mu$ using the diversitree \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
+First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution maintaining $\lambda$ $>$ $\mu$ using the \texttt{diversitree} \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
 For each tree, we arbitrarily set the outgroup to be the first taxa (alphabetically) thus effectively rooting the trees on this taxa.
@@
     \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ and scaled to sum to $1$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
-    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][; see below]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
+    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
@@
 \subsection{Modifying the matrices}
-We calculated the pairwise Character Differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}. %TG: TO IMPLEMENT! 
+We calculated the pairwise character differences ($CD$) for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
 We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
 For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e.
-with an average character difference $<$$0.25$) and replaced them randomly by any of the remaining characters.
+with an average character difference $<0.25$) and replaced them randomly by any of the remaining characters.
 This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
 Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e.
-with an average character difference $<$$0.75$) and randomly replaced them with the remaining ones.
+with an average character difference $<0.75$) and randomly replaced them with the remaining ones.
 Finally, because this operation effectively changes the weight of characters (i.e.
-giving the characters $<$$0.25$ or $>$$0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification (i.e.
+giving the characters $<0.25$ or $>0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification (i.e.
 randomly weighting characters).
+Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
 This step resulted in a total of 1440 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
@@
 For the Bayesian consensus trees, the 25\% first trees of the posterior tree distribution were excluded as a burnin.
-The 2880 tree inferences took around 1 CPU century on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC}.
+The 2880 tree inferences took around 1 CPU century (@@@ current: 83 CPU years) on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC}.
 
@@
 
-The metrics scores where obtained using the TreeCmp java script \citep{Bogdanowicz2012}.
+The metrics scores where calculated using the \texttt{TreeCmp} javascript \citep{Bogdanowicz2012}.
 The measurements where then standardised using the Normalised Tree Similarity metric \citep[$NTS$; i.e.
@@
 
-Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at 1, the resulting residuals distribution of $NTS$ scores was not normal thus preventing the use of parametric nested anova for comparisons (see Supplementary @@@).
-Instead, we used the a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
-We considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
+Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at $1$, the resulting in the residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary @@@).
+%TG: TODO: residual normality + homscedasticity tests supp.
+Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in their p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
+when multiple trees are equal to the ``normal'' tree).
+Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
+Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
 Comparisons falling within these range can not be designated as strictly similar/different but can still be ranked (e.g.
@@
 
+%TG: This one can be reworded. Get a dichotomie a priori/a posteriori correlation. + add trade off in coding (more characters = more info but less independence)
+Furthermore it is important to distinguish how researchers can tackle these correlations: the coding dependence is by essence an \textit{a priori} source of correlation: the correlation is induced before the phylogenetic analysis.
+On the other hand Biological and Evolutionary dependences can only be determined \textit{a posteriori}: the correlation must be based on a phylogenetic hypothesis.
+Additionally there seem to be a trade in coding fewer characters (reducing \textit{a priori} dependence) but making the \textit{a posteriori} correlation more dependent on the coding.
+On the other hand, coding more characters increases \textit{a priori} dependence but will lead to the \textit{a posteriori} correlation being less dependent on it.
+
 \subsection{Conclusion}
@@
 
-\section{Data availability and repeatability}
+\section{Data availability, repeatability and reproducibility}
 The consensus trees are available on figshare at \url{@@@}.
-The code for reproducing the simulation is available and explained at \url{https://github.com/TGuillerme/CharactersCorrelation}.
-The vignette for reproducing the analysis and the figure is available at \url{http://htmlpreview.github.io/?https://github.com/TGuillerme/CharactersCorrelation/Analysis/EffectCorrelation.html}.
+The simulations are fully replicable following the explanations at \url{https://github.com/TGuillerme/CharactersCorrelation}.
+The post-simulation analysis, tables and figures (reported in this manuscript) are fully reproducible see (\url{https://github.com/TGuillerme/CharactersCorrelation}).
 

### 2017-09-19 Analysis/EffectCorrelationFigure.Rmd
Renamed figure file

--- 2017-09-18 Analysis/EffectCorrelation.Rmd
@@
 
-Results
------------------------------------
+# Results (plots)
 
@@
 
+# Results supplementary
+
+## Effect of character correlation compared to the null ("random") trees
+
 Comparison to 'null' tree (RF normalised distance)
@@
 
-Some stats
------------------------------------
+## Ability to recover the "true" tree
 
-Supp
------------------------------------
+How different are the "normal" and "random" trees from the true trees
 
-Plotting the results for 25 taxa
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the best tree (RF)
-plot.results.single(t25_list, metric = 2, what = "best",
-    main = "comparison to 'best' tree (25 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the best tree (Triplets)
-plot.results.single(t25_list, metric = 4, what = "best",
-    main = "comparison to 'best' tree (25 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the null tree (RF)
-plot.results.single(t25_list, metric = 2, what = "null",
-    main = "comparison to 'null' tree (25 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the null tree (Triplets)
-plot.results.single(t25_list, metric = 4, what = "null",
-    main = "comparison to 'null' tree (25 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-Plotting the results for 75 taxa
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the best tree (RF)
-plot.results.single(t75_list, metric = 2, what = "best",
-    main = "comparison to 'best' tree (75 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the best tree (Triplets)
-plot.results.single(t75_list, metric = 4, what = "best",
-    main = "comparison to 'best' tree (75 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the null tree (RF)
-plot.results.single(t75_list, metric = 2, what = "null",
-    main = "comparison to 'null' tree (75 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the null tree (Triplets)
-plot.results.single(t75_list, metric = 4, what = "null",
-    main = "comparison to 'null' tree (75 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-Plotting the results for 150 taxa
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the best tree (RF)
-plot.results.single(t150_list, metric = 2, what = "best",
-    main = "comparison to 'best' tree (150 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the best tree (Triplets)
-plot.results.single(t150_list, metric = 4, what = "best",
-    main = "comparison to 'best' tree (150 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the null tree (RF)
-plot.results.single(t150_list, metric = 2, what = "null",
-    main = "comparison to 'null' tree (150 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Comparisons to the null tree (Triplets)
-plot.results.single(t150_list, metric = 4, what = "null",
-    main = "comparison to 'null' tree (150 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-Checking the results for 25 taxa
+### 25 taxa
 
@@
 
-```{r, fig.width=12, fig.height=6}
-## Checking comparisons (RF)
-plot.results.single(t25_list, metric = 2, what = "check",
-    main = "Check comparison (25 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Checking comparisons (Triplets)
-plot.results.single(t25_list, metric = 4, what = "check",
-    main = "Check comparison (25 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-Checking the results for 75 taxa
+### 75 taxa
 
@@
 
-```{r, fig.width=12, fig.height=6}
-## Checking comparisons (RF)
-plot.results.single(t75_list, metric = 2, what = "check",
-    main = "Check comparison (75 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Checking comparisons (Triplets)
-plot.results.single(t75_list, metric = 4, what = "check",
-    main = "Check comparison (75 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-Checking the results for 150 taxa
+### 150 taxa
 
@@
 ```
-
-```{r, fig.width=12, fig.height=6}
-## Checking comparisons (RF)
-plot.results.single(t150_list, metric = 2, what = "check",
-    main = "Check comparison (150 taxa)", legend = TRUE, NTS = TRUE)
-```
-
-```{r, fig.width=12, fig.height=6}
-## Checking comparisons (Triplets)
-plot.results.single(t150_list, metric = 4, what = "check",
-    main = "Check comparison (150 taxa)", legend = TRUE, NTS = TRUE)
-```

### 2017-09-19 Analysis/EffectCorrelationStats.Rmd
Updated stats

@@
 whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)
+```
+
+```{r, echo = FALSE}
+library(knitr)
+library(xtable)
+```
+
+# Pooled results
+
+## Effect of character correlation per scenario
+
+```{r, warning = FALSE}
+## Pooling the data per scenarios
+scenario_RF_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
+                               metric = "RF", best = "norm")
+scenario_Tr_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
+                               metric = "Tr", best = "norm")
+
+## Preparing the table labels
+metric.labels <- c("RF", "", "", "Tr", "", "")
+
+## Combining the data into a table
+scenarios <- pool.table(list(scenario_RF_norm, scenario_Tr_norm),  metric.label = metric.labels,
+                        label.param = "scenario")
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_pooled_scenarios <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled scenarios.")
+
+## Printing the tables
+kable(scenarios, row.names = FALSE, caption = captions_pooled_scenarios)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Tab_pooledscenarios")
+```
+
+### Testing the effect of the scenario on the NTS
+
+```{r}
+## Applying an lm to all the data
+results <- test.effect.aov(whole_data, formula = NTS ~ scenario, fun = lm)
+## Removing the null tests
+results[[1]] <- results[[1]][-c(3,4),]
+```
+
+Checking the model assumptions
+
+```{r, echo = FALSE}
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```
+
+Applying a Bhattacharrya Coefficient and Wilcoxon test
+
+```{r, warning = FALSE}
+## Summarise into a table
+scenario_list <- list(scenario_RF_norm, scenario_Tr_norm)
+metric.labels <- c("RF", "", "", "Tr", "", "")
+test_table <- pair.test.table(scenario_list, metric.label = metric.labels)
+```
+
+```{r, echo = FALSE}
+## Caption
+cap_pool_scen_test <- paste("Difference between the pooled scenarios.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+
+## Printing the tables
+kable(test_table, row.names = FALSE, caption = cap_pool_scen_test)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(test_table, caption = cap_pool_scen_test, label = "Tab_pooledscenarios_test")
+```
+
+## Effect of character correlation per number of character
+
+```{r, warning = FALSE}
+## Pooling the data per character
+character_RF_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
+                               metric = "RF", best = "norm")
+character_Tr_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
+                               metric = "Tr", best = "norm")
+
+## Preparing the table labels
+metric.labels <- c("RF", "", "", "Tr", "", "")
+
+## Combining the data into a table
+characters <- pool.table(list(character_RF_norm, character_Tr_norm), metric.label = metric.labels,
+                         label.param = "character")
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_pooled_characters <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of characters.")
+
+## Printing the tables
+kable(characters, row.names = FALSE, caption = captions_pooled_characters)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(characters, caption = captions_pooled_characters, label = "Tab_pooledcharacters")
+```
+
+### Testing the effect of the number of characters on the NTS
+
+```{r}
+## Applying an lm to all the data
+results <- test.effect.aov(whole_data, formula = NTS ~ character, fun = lm)
+## Removing the null tests
+results[[1]] <- results[[1]][-c(3,4),]
+```
+
+Checking the model assumptions
+
+```{r, echo = FALSE}
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3, label = "Tab_signif")
+```
+
+Applying a Bhattacharrya Coefficient and Wilcoxon test
+
+```{r, warning = FALSE}
+## Summarise into a table
+character_list <- list(character_RF_norm, character_Tr_norm)
+metric.labels <- c("RF", "", "", "Tr", "", "")
+test_table <- pair.test.table(character_list, metric.label = metric.labels)
+```
+
+```{r, echo = FALSE}
+## Caption
+cap_pool_char_test <- paste("Difference between the pooled number of characters.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+
+## Printing the tables
+kable(test_table, row.names = FALSE, caption = cap_pool_char_test)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(test_table, caption = cap_pool_char_test, label = "Tab_pooledscharacters_test")
+```
+
+## Effect of character correlation per number of taxa
+
+```{r, warning = FALSE}
+## Pooling the data per taxa
+taxa_RF_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
+                               metric = "RF", best = "norm")
+taxa_Tr_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
+                               metric = "Tr", best = "norm")
+
+## Preparing the table labels
+metric.labels <- c("RF", "", "", "Tr", "", "")
+
+## Combining the data into a table
+taxa <- pool.table(list(taxa_RF_norm, taxa_Tr_norm), metric.label = metric.labels,
+                   label.param = "taxa")
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_pooled_taxa <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of taxa.")
+
+## Printing the tables
+kable(taxa, row.names = FALSE, caption = captions_pooled_taxa)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(taxa, caption = captions_pooled_taxa, label = "Tab_pooledtaxa")
+```
+
+### Testing the effect of the method on the NTS
+
+```{r}
+## Applying an lm to all the data
+results <- test.effect.aov(whole_data, formula = NTS ~ taxa, fun = lm)
+## Removing the null tests
+results[[1]] <- results[[1]][-c(3,4),]
+```
+
+Checking the model assumptions
+
+```{r, echo = FALSE}
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```
+
+Applying a Bhattacharrya Coefficient and Wilcoxon test
+
+```{r, warning = FALSE}
+## Summarise into a table
+taxa_list <- list(taxa_RF_norm, taxa_Tr_norm)
+metric.labels <- c("RF", "", "", "Tr", "", "")
+test_table <- pair.test.table(taxa_list, metric.label = metric.labels)
+```
+
+```{r, echo = FALSE}
+## Caption
+cap_pool_tax_test <- paste("Difference between the pooled number of taxa.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+
+## Printing the tables
+kable(test_table, row.names = FALSE, caption = cap_pool_tax_test)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(test_table, caption = cap_pool_tax_test, label = "Tab_pooledstaxa_test")
+```
+
+## Effect of character correlation per methods
+
+```{r, warning = FALSE}
+## Pooling the data per taxa
+method_RF_norm <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
+                               metric = "RF", best = "norm")
+method_Tr_norm <- multi.pool(whole_data, param =c("bayesian", "parsimony"),
+                               metric = "Tr", best = "norm")
+
+## Preparing the table labels
+metric.labels <- c("RF", "", "Tr", "")
+
+## Combining the data into a table
+methods <- pool.table(list(method_RF_norm, method_Tr_norm), metric.label = metric.labels,
+                      label.param = "method")
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_pooled_methods <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled methods.")
+
+## Printing the tables
+kable(methods, row.names = FALSE, caption = captions_pooled_methods)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(methods, caption = captions_pooled_methods, label = "Tab_pooledmethod")
+```
+
+### Testing the effect of the method on the NTS
+
+```{r}
+## Applying an lm to all the data
+results <- test.effect.aov(whole_data, formula = NTS ~ method, fun = lm)
+## Removing the null tests
+results[[1]] <- results[[1]][-c(3,4),]
+```
+
+Checking the model assumptions
+
+```{r, echo = FALSE}
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```
+
+Applying a Bhattacharrya Coefficient and Wilcoxon test
+
+```{r, warning = FALSE}
+## Summarise into a table
+methods_list <- list(method_RF_norm, method_Tr_norm)
+metric.labels <- c("RF", "", "Tr", "")
+test_table <- pair.test.table(methods_list, metric.label = metric.labels)
+```
+
+```{r, echo = FALSE}
+## Caption
+cap_pool_met_test <- paste("Difference between the pooled methods.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+
+## Printing the tables
+kable(test_table, row.names = FALSE, caption = cap_pool_met_test)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+make.xtable(test_table, caption = cap_pool_met_test, label = "Tab_pooledsmethods_test")
+```
+
+# Combined results
+
+We can also analyse the effect of each parameters combination separately:
+
+## Summary statistics
+
+We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
+Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
+The p-values are adjusted for multiple tests using Bonferroni-Holm correction.
+
+```{r}
+## Summarising the data
+sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL)
+sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL)
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_summary <- paste("Summary statistics of the normalised", rep(c("Robinson-Fould", "Triplets"), 2), "distance to the", c(rep("best", 2), rep("null", 2)), "tree.")
+
+## Printing the tables
+kable(sum_RF_best, row.names = FALSE, caption = captions_summary[1])
+kable(sum_Tr_best, row.names = FALSE, caption = captions_summary[2])
+```
+
+<!-- Table for publication -->
+
+```{r, echo = FALSE, eval = FALSE}
+## Labels
+labels_summary <- paste0("Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
+
+## Tables
+make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1])
+make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2])
+```
+
+## Nested anova
+
+Let's analysis the data using a nested anova.
+The models look at the effect of our scenarios ("maximised", "minimised", etc...) on the normalised tree score (NTS - using both the Robinson-Fould and Triplets scores) with a nested effect of the method (parsimony or bayesian), the number of taxa (25, 75, 150) and the number of characters (100, 150, 1000).
+
+```{r}
+## Applying an lm to all the data with error
+results <- test.effect.aov(whole_data, fun = lmer,
+                           formula = NTS ~ scenario + (1|method) + (1|taxa) + (1|character))
+## Removing the null tests
+results[[1]] <- results[[1]][-c(3,4),]
+```
+
+Checking the model assumptions
+
+```{r, echo = FALSE}
+kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, eval = FALSE, echo = FALSE}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```
+
+All our models violate a least one of the two models assumptions (normality of the residuals and homoscedasticity of variance) and thus cannot be used to compare the scenarios.
+This is due to the fact that the metric is bounded to 1.
+
+## Distributions probability overlap (Bhattacharrya Coefficient)
+
+We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
+A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don t overlap (significant effect of one scenario on topology).
+Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).
+
+```{r}
+## Running the pairwise Bhattacharrya Coefficient analysis:
+bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff)
+bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff)
+
+## Tables
+kable(bc_RF_best, row.names = FALSE, digit = 3)
+kable(bc_Tr_best, row.names = FALSE, digit = 3)
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_bc <- paste0("Bhattacharrya Coefficient (",rep(c("Robinson-Fould distance", "Triplets distance"), 1) ," from the ",c(rep("best", 2)), " tree)" , " between the different scenarios.")
+
+## Printing the tables
+kable(bc_RF_best, row.names = FALSE, caption = captions_bc[1], digit = 3)
+kable(bc_Tr_best, row.names = FALSE, caption = captions_bc[2], digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, echo = FALSE, eval = FALSE}
+## Labels
+labels_bc <- paste0("Tab_BC", rep(c("RF", "Tr")), c(rep("best", 2)))
+
+## Tables
+make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1])
+make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2])
+```
+
+## Non-parametric pairwise comparisons
+
+We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
+Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
+The p-values are adjusted for multiple tests using Bonferroni-Holm correction.
+
+```{r, warning = FALSE}
+## Running the pairwise Wilcox tests:
+wx_RF_best <- apply.test(whole_data, "RF", "norm", wilcox.test)
+wx_Tr_best <- apply.test(whole_data, "Triples", "norm", wilcox.test)
+
+## Applying Bonferroni correction
+for(column in c(5,7,9)) {
+  wx_RF_best[,column] <- p.adjust(wx_RF_best[,column], method = "bonferroni")
+  wx_Tr_best[,column] <- p.adjust(wx_Tr_best[,column], method = "bonferroni")
+}
+```
+
+```{r, echo = FALSE}
+## Setting the table captions
+captions_wx <- paste0("Non-parametric Wilcoxon test (",rep(c("Robinson-Fould distance", "Triplets distance")) ," from the ",c(rep("best", 2)), " tree)" , " between the different scenarios (p-values corrected using Bonferroni-Holm correction).")
+
+## Printing the tables
+kable(wx_RF_best, row.names = FALSE, caption = captions_wx[1], digit = 3)
+kable(wx_Tr_best, row.names = FALSE, caption = captions_wx[2], digit = 3)
+```
+
+<!-- Table for publication -->
+
+```{r, echo = FALSE, eval = FALSE}
+## Labels
+labels_wx <- paste0("Tab_BC", rep(c("RF", "Tr")), c(rep("best", 2)))
+
+## Tables
+make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1])
+make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2])
 ```

### 2017-09-19 Analysis/EffectCorrelationFullResults.Rmd
Full results

--- 2017-09-19 Analysis/EffectCorrelationStats.Rmd
@@
 whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)
+```
+
+```{r, echo = FALSE}
+library(knitr)
+library(xtable)
 ```
@@
 bc_Tr_null <- apply.test(whole_data, "Triples", "null", bhatt.coeff)
+```
 
+```{r, echo = FALSE}
 ## Tables

### 2017-09-26 Analysis/EffectCorrelationStats.Rmd
Updated stats

@@
 cap_pool_scen_test <- paste("Difference between the pooled scenarios.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction)")
 
@@
 methods_list <- list(method_RF_norm, method_Tr_norm)
-metric.labels <- c("RF", "", "Tr", "")
+metric.labels <- c("RF", "Tr")
 test_table <- pair.test.table(methods_list, metric.label = metric.labels)
@@
 bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff)
-
-## Tables
-kable(bc_RF_best, row.names = FALSE, digit = 3)
-kable(bc_Tr_best, row.names = FALSE, digit = 3)
 ```

### 2017-09-26 Writing/draft.tex
Results part of the draft

@@
 
-@@@ NOTE: THESE RESULTS ARE NOT FINAL - I thus didn't spend too much time describing them @@@
-
-\subsection{Effect of Character Differences on topology}
-The overall amount of pairwise $CD$ in a matrix has an effect of the ability to recover the correct topology.
-An overall decrease in $CD$ leads to an decrease in $NTS_{RF}$ (as a proxy for clade conservation - Fig.
-\ref{Fig:RF_results_best}) and in $NTS_{Tr}$ (as a proxy for taxa position - Fig.
-\ref{Fig:Tr_results_best}).
-Conversely, an increase in $CD$ leads to an increase in in $NTS_{RF}$ (Fig. \ref{Fig:RF_results_best}) and $NTS_{Tr}$ (Fig. \ref{Fig:Tr_results_best}).
-This beneficial effect can be contrasted however with a similar beneficial effect when simply randomising the character differences (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
-
-\subsubsection{Effect of the number of characters}
-This negative effect of decreased $CD$ (and positive of increased $CD$) on topology is increased with the number of characters.
-% Test pulled difference between character numbers.
-
-\subsubsection{Effect of the number of taxa}
-Conversely, the effect of variation in $CD$ doesn't seem to drastically increase with the number of taxa after 75 taxa in our simulations.
-% Test pulled difference between number of taxa.
-
-\subsection{Effect of Character Differences on the inference method}
-There seems to be a strong effect on the ability to recover the correct topology depending on the tree inference method.
-Parsimony trees seems to be more sensible than Bayesian ones.
-% Test pulled difference between methods.
+%TG: @@@ NOTE: THESE RESULTS ARE NOT FINAL - I thus didn't spend too much time describing them @@@
 
@@
 \centering
-   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
-\caption{\small{Effect of Character Difference on recovering the ``best'' topology.
+   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
+\caption{\small{Effect of Character Difference on recovering the ``normal'' topology.
 The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
-The x axis represents the different Character Difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
+The x axis represents the different character difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
 \label{Fig:RF_results_best}
@@
    \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
-\caption{Effect of Character Difference on recovering the ``best'' topology.
+\caption{Effect of Character Difference on recovering the ``normal'' topology.
 The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
@@
 
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Fri Jul 28 10:29:26 2017
+\subsection{Effect of Character Differences on topology}
+
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology with maximising the character difference leading to the less loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the character difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see supplementary material 3 for the full summary statistics).
+There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
+
 \begin{table}[ht]
 \centering
-\begin{tabular}{rlllrrrrrrrrr}
+\begin{tabular}{ll|r|rr}
   \hline
-Taxa & Char. & Method & \multicolumn{3}{c}{max \textit{vs}. min} & \multicolumn{3}{c}{max \textit{vs}. rand} & \multicolumn{3}{c}{min \textit{vs}. rand} \\
- & & & BC & W & p & BC & W & p & BC & W & p \\ 
+metric & test & bhatt.coeff & statistic & p.value \\ 
   \hline
-25 & 100 & Bay. & 0.67 & 1130.00 & \textbf{0.00} & 0.85 & 856.00 & \textbf{1.00} & 0.74 & 320.00 & \textbf{0.00} \\ 
-   & 350 &   & 0.71 & 1236.50 & \textbf{0.00} & \textbf{0.96} & 844.00 & \textbf{1.00} & 0.73 & 338.00 & \textbf{0.00} \\ 
-   & 1000 &   & 0.71 & 1321.50 & \textbf{0.00} & 0.89 & 802.00 & 1.00 & 0.67 & 226.00 & \textbf{0.00} \\ 
-75 & 100 &   & 0.48 & 1366.50 & \textbf{0.00} & 0.66 & 1134.00 & \textbf{0.01} & 0.69 & 264.00 & \textbf{0.00} \\ 
-   & 350 &   & 0.40 & 999.50 & \textbf{0.00} & 0.59 & 878.50 & \textbf{0.00} & 0.71 & 188.00 & \textbf{0.00} \\ 
-   & 1000 &   & 0.53 & 390.00 & \textbf{0.00} & 0.80 & 341.50 & 0.13 & 0.69 & 100.00 & 0.14 \\ 
-150 & 100 &   & \textbf{0.00} & 256.00 & \textbf{0.00} & \textbf{0.00} & 256.00 & \textbf{0.00} & 0.42 & 24.00 & \textbf{0.01} \\ 
-    & 350 &   & 0.07 & 361.00 & \textbf{0.00} & \textbf{0.05} & 361.00 & \textbf{0.00} & 0.39 & 56.50 & \textbf{0.02} \\ 
-    & 1000 &   & \textbf{0.00} & 121.00 & \textbf{0.00} & 0.13 & 119.00 & \textbf{0.01} & \textbf{0.00} & 0.00 & \textbf{0.00} \\ 
-25 & 100 & Par. & 0.63 & 357.00 & \textbf{0.00} & 0.91 & 234.00 & 1.00 & 0.71 & 85.00 & 0.10 \\ 
-   & 350 &   & 0.34 & 387.00 & \textbf{0.00} & 0.79 & 287.50 & 0.94 & 0.49 & 61.50 & \textbf{0.01} \\ 
-   & 1000 &  & 0.43 & 384.00 & \textbf{0.00} & 0.79 & 289.50 & 0.62 & 0.71 & 65.00 & \textbf{0.01} \\ 
-75 & 100 &   & 0.38 & 379.50 & \textbf{0.00} & 0.40 & 359.00 & \textbf{0.00} & 0.75 & 157.50 & 1.00 \\ 
-   & 350 &   & 0.25 & 393.00 & \textbf{0.00} & 0.76 & 311.00 & 0.13 & 0.65 & 55.50 & \textbf{0.01} \\ 
-   & 1000 &   & 0.15 & 393.00 & \textbf{0.00} & 0.46 & 362.00 & \textbf{0.00} & 0.46 & 58.00 & \textbf{0.01} \\ 
-150 & 100 &   & 0.34 & 357.50 & \textbf{0.00} & 0.49 & 356.50 & \textbf{0.00} & 0.73 & 197.00 & 1.00 \\ 
-    & 350 &   & 0.17 & 390.00 & \textbf{0.00} & 0.60 & 348.00 & \textbf{0.00} & 0.63 & 63.50 & \textbf{0.01} \\ 
-    & 1000 &   & \textbf{0.00} & 400.00 & \textbf{0.00} & 0.71 & 328.50 &\textbf{0.03} & 0.34 & 24.00 & \textbf{0.00} \\ 
+RF & maxi:mini & 0.588 & 439466.500 & \textbf{0.000} \\ 
+   & maxi:rand & 0.894 & 347624.500 & \textbf{0.000} \\ 
+   & mini:rand & 0.846 & 115008.500 & \textbf{0.000} \\ 
+Tr & maxi:mini & 0.628 & 443049.500 & \textbf{0.000} \\ 
+   & maxi:rand & 0.921 & 348074.000 & \textbf{0.000} \\ 
+   & mini:rand & 0.825 & 113171.500 & \textbf{0.000} \\ 
    \hline
 \end{tabular}
-\caption{Example of a result table: BC = Bhattacharyya Coefficient, W = wilcox test statistic and p = associated p-value} 
-\label{LABEL}
+\caption{Difference between the pooled scenarios.
+bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction), significant values are in bold.}
+\label{Tab_pooledscenarios_test}
 \end{table}
+
+\subsubsection{Effect of the number of characters}
+
+This effect of the character difference is not dependent on the number of characters for the clade conservation (i.e.
+$NTS_{RF}$) for a ``similar'' number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.76, 0.78, 0.8 respectively - see supplementary materials 3) with a significant effect only between 100 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+The number of characters affects the character difference more in terms of taxa placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+These difference have to be contrasted however by a really high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{ll|r|rr}
+  \hline
+metric & test & bhatt.coeff & statistic & p.value \\ 
+  \hline
+RF & c100:c350 & \textbf{0.989} & 228495.000 & 0.174 \\ 
+   & c100:c1000 & \textbf{0.983} & 218305.500 & \textbf{0.002} \\ 
+   & c350:c1000 & \textbf{0.988} & 232832.500 & 0.643 \\ 
+  Tr & c100:c350 & \textbf{0.965} & 203355.500 & \textbf{0.000} \\ 
+   & c100:c1000 & \textbf{0.959} & 189064.000 & \textbf{0.000} \\ 
+   & c350:c1000 & \textbf{0.983} & 226472.000 & 0.085 \\ 
+   \hline
+\end{tabular}
+\caption{Difference between the pooled number of characters. See Table \ref{Tab_pooledscenarios_test} for caption.} 
+\label{Tab_pooledscharacters_test}
+\end{table}
+
+\subsubsection{Effect of the number of taxa}
+
+Similarly than the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
+A low number of taxa (25) resulted in significant differences in both $NTS_{RF}$ and $NTS_{Tr}$ (medians for 25, 75 and 150 taxa equals 0.8, 0.76, 0.75 $NTS_{RF}$ and 0.75, 0.59 and 0.59 $NTS_{Tr}$respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
+Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{ll|r|rr}
+  \hline
+metric & test & bhatt.coeff & statistic & p.value \\ 
+  \hline
+RF & t25:t75 & \textbf{0.987} & 267274.500 & \textbf{0.003} \\ 
+   & t25:t150 &\textbf{0.979} & 272359.500 & \textbf{0.000} \\ 
+   & t75:t150 & \textbf{0.988} & 247619.500 & 1.000 \\ 
+Tr & t25:t75 & \textbf{0.975} & 282017.500 & \textbf{0.000} \\ 
+   & t25:t150 & \textbf{0.970} & 279730.000 & \textbf{0.000} \\ 
+   & t75:t150 & \textbf{0.989} & 240414.500 & 1.000 \\ 
+   \hline
+\end{tabular}
+\caption{Difference between the pooled number of taxa. See Table \ref{Tab_pooledscenarios_test} for caption.} 
+\label{Tab_pooledstaxa_test}
+\end{table}
+
+\subsection{Effect of Character Differences on the inference method}
+
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and Maximum Parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.82 and 0.68 respectively) but not in terms of individual taxa placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.72 and 0.61 respectively).
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{llrrr}
+  \hline
+metric & test & bhatt.coeff & statistic & p.value \\ 
+  \hline
+RF & bayesian:parsimony & 0.893 & 676940.500 & \textbf{0.000} \\ 
+Tr & bayesian:parsimony & \textbf{0.982} & 547140.000 & 0.147 \\ 
+   \hline
+\end{tabular}
+\caption{Difference between the pooled methods. See Table \ref{Tab_pooledscenarios_test} for caption.} 
+\label{Tab_pooledsmethods_test}
+\end{table}
+
+\subsection{Combined effects}
+
+When looking at the combined effect of each parameters, the maximised and minimised scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
+The same differences are observed when comparing the maximised scenario against the randomised one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the Maximum Parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
+Identically, there was always a significant difference between the minimised scenario and the randomised one was expect for the matrix of 150 taxa and 100 characters under Maximum Parsimony for $NTS_{RF}$ and the matrix of 150 and 1000 characters under Bayesian inference for $NTS_{Tr}$.
+The full list of comparisons and summary statistics are available in the supplementary materials 3.
 

### 2017-09-27 Writing/draft.tex
Abstract ideas

@@
 
-\noindent{\Large \bf Effect of violating character independence on tree topology}
+\noindent{\Large \bf Effect of discrete morphological character independence on tree topology}
 \bigskip
@@
 \begin{abstract}
-blablabla
+- When using discrete morphological characters for building phylogenies, the characters are assumed to be independent.
+
+- However, this assumption is nearly always wrong, simply because of the nature of morphological integration in organisms.
+
+- The dependence between characters can originate from biological features, shared phylogenetic history or errors from the researcher when coding characters.
+
+- Although the two first sources can be investigated by biologists post phylogenetic inference and the third one can be avoided pre inference with good practices, software do not distinguish between any of them.
+
+- In this study, we propose a new metric of raw character difference as a proxy for character (in)dependence.
+
+- Using thorough simulations and this metric, we test the effect of increasing or decreasing character difference on topology with a variable number of taxa and characters under both Bayesian and Maximum Parsimony inference.
+
+- Overall we found a positive effect of reducing character correlations on recovering the correct topology.
+
+- However, this effect is less important for matrices with a small number of taxa (~25) were reducing character correlation was not more effective than randomly drawing characters.
+
+- Furthermore, in bigger matrices (~350 taxa), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
+
+- These results suggest that ignoring the problem of character correlation or independence can easily impact topology in phylogenetic analysis.
+
+- However, encouragingly, they also suggest that, unless correlation is effectively maximised or minimised, probabilistic methods can easily accommodate for random correlation between characters.
+
 \end{abstract}
-
-The project implies inferring a total of 2880 phylogenetic trees (40 replicates of 9 data combination scenarios generating 4 trees each; both inferred in Bayesian and under Maximum Parsimony).
-The calculations for the project is ongoing since December 2016 a and is now 95.27\% completed (with a current cumulated computational time of around 80 CPU years that generated 3.2 TB of trees).
-The pipelines for analysing the data and producing the publishable figures and tables is already implemented and most of the manuscript is already written ()
-
-Overall we found a positive effect of reducing character correlations on recovering the correct topology.
-However, this effect is less important for matrices with a small number of taxa (~25) were reducing character correlation was not more effective than randomly drawing characters.
-Furthermore, in bigger matrices (~350 taxa), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
-
-These results suggest that ignoring the problem of character correlation or independence can easily impact topology in phylogenetic analysis.
-However, encouragingly, they also suggest that, unless correlation is effectively maximised or minimised, probabilistic methods can easily accommodate for random correlation between characters.
 

### 2017-09-27 Writing/draft.tex
Updated discussion

@@
 
-- Using thorough simulations and this metric, we test the effect of increasing or decreasing character difference on topology with a variable number of taxa and characters under both Bayesian and Maximum Parsimony inference.
+- Using thorough simulations and this metric, we test the effect of increasing or decreasing Character Difference on topology with a variable number of taxa and characters under both Bayesian and Maximum Parsimony inference.
 
@@
 
-\noindent (Keywords: )\\
+\noindent (Keywords: Character difference, correlation, topology, Bayesian, Maximum Parsimony)\\
 
@@
 
-The last six years have seen a fantastic expansion of the use of discrete morphological data in macroevolutionary studies.
+Since less than a decade, there has been a fantastic expansion of the use of discrete morphological data in macroevolutionary studies.
 Traditionally confined to ``classic'' cladistic studies in palaeontology, these methods have recently been widely expanded and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep[e.g.][]{pyrondivergence2011,ronquista2012}.
@@
 
-If the three induced sources of correlation are fairly well studied, their actual interpretation by phylogenetic inference software, has, to our knowledge, never been studied through a simulation protocol (although they have been explored empirically for morphological data; \citealt{Davalos01072014}; or molecular data; \citealt{ZouConvergence}).
+If the three sources of correlation are fairly well studied, their actual interpretation by phylogenetic inference software, has, to our knowledge, never been studied through a simulation protocol (although they have been explored empirically for morphological data; \citealt{Davalos01072014}; or molecular data; \citealt{ZouConvergence}).
 In this study, we formally assess the effect of discrete character's correlation using simulated data.
@@
     \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modified the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
-    We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character difference to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
+    We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising Character Difference to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
 This step resulted in 36 matrices.
@@
    \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
-\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using Maximum Parsimony and Bayesian inference (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).}
+\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using Maximum Parsimony and Bayesian inference (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
+$n$ and $m$ corresponds to the number of characters with a Character Difference $<0.25$ and $>0.75$ respectively.}
 \label{Fig:outline}
@@
 
-When the character difference null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
-When the character difference maximal ($1$) it means it conveys the most different signal (i.e. characters are uncorrelated).
-It is important to stress that a character difference of $0$ (i.e. the same phylogenetic signal) does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal) .
+When the Character Difference null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
+When the Character Difference maximal ($1$) it means it conveys the most different signal (i.e. characters are uncorrelated).
+It is important to stress that a Character Difference of $0$ (i.e. the same phylogenetic signal) does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal) .
 For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
@@
 \subsection{Modifying the matrices}
-We calculated the pairwise character differences ($CD$) for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
+We calculated the pairwise Character Differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
+We then modified the matrices to either maximise or minimise the pairwise Character Differences for each matrices using three different algorithms.
 For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e.
-with an average character difference $<0.25$) and replaced them randomly by any of the remaining characters.
-This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
-Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e.
-with an average character difference $<0.75$) and randomly replaced them with the remaining ones.
+with an average Character Difference $<0.25$) and replaced them randomly by any of the remaining characters.
+This operation increased the overall pairwise Character Difference in the matrix thus making the characters more dissimilar.
+Conversely, for minimising the pairwise Character Differences, we selected the most dissimilar characters (i.e.
+with an average Character Difference $<0.75$) and randomly replaced them with the remaining ones.
 Finally, because this operation effectively changes the weight of characters (i.e.
@@
 \caption{Example illustration of the protocol for modifying matrices.
-The matrices represent the pairwise character differences for 100 characters.
-Blue colours correspond to low character differences and orange colours correspond to high character differences.
-\textbf{A} - a random Birth-Death tree is simulated and used for generating the normal matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised character difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).}
+The matrices represent the pairwise Character Differences for 100 characters.
+Blue colours correspond to low Character Differences and orange colours correspond to high Character Differences.
+\textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised Character Difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).}
 \label{Fig:modif_matrix}
@@
 For the Bayesian consensus trees, the 25\% first trees of the posterior tree distribution were excluded as a burnin.
-The 2880 tree inferences took around 1 CPU century (@@@ current: 83 CPU years) on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC}.
+The 2880 tree inferences took around one CPU century on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC}.
 
@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of Character Difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Tables @@@).
+We explored the effect of Character Difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).
 
@@
 centering the metrics scores using the mean metric score for 1000 pairwise comparisons between random trees with $n$ taxa;][]{Bogdanowicz2012,Guillerme2016146}.
-When the normalised metric has a score of $1$ it means both trees are identical, when it has a score of $0$ it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
+When the normalised metric has a score of one it means both trees are identical, when it has a score of zero it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
 The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Fould ($NTS_{RF}$) Similarity reflects the conservation of clades (i.e.
@@
 
-Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at $1$, the resulting in the residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary @@@).
+Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one, the resulting in the residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary @@@).
 %TG: TODO: residual normality + homscedasticity tests supp.
@@
 The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
-The x axis represents the different character difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
+The x axis represents the different Character Difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
 \label{Fig:RF_results_best}
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology with maximising the character difference leading to the less loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the character difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see supplementary material 3 for the full summary statistics).
-There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
+The overall amount of Character Difference in a matrix has an effect of the ability to recover the correct topology with maximising the Character Difference leading to the less loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the Character Difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see supplementary material 3 for the full summary statistics).
+There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the Character Difference (Bhattacharrya coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the Character Difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
 
@@
 
-This effect of the character difference is not dependent on the number of characters for the clade conservation (i.e.
+This effect of the Character Difference is not dependent on the number of characters for the clade conservation (i.e.
 $NTS_{RF}$) for a ``similar'' number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.76, 0.78, 0.8 respectively - see supplementary materials 3) with a significant effect only between 100 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-The number of characters affects the character difference more in terms of taxa placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+The number of characters affects the Character Difference more in terms of taxa placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
 These difference have to be contrasted however by a really high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.
@@
 
-Similarly than the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
+Similarly than the effect of number of characters on Character Difference, the number of taxa seems to have only a marginal effect.
 A low number of taxa (25) resulted in significant differences in both $NTS_{RF}$ and $NTS_{Tr}$ (medians for 25, 75 and 150 taxa equals 0.8, 0.76, 0.75 $NTS_{RF}$ and 0.75, 0.59 and 0.59 $NTS_{Tr}$respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
@@
 
-When looking at the combined effect of each parameters, the maximised and minimised scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
-The same differences are observed when comparing the maximised scenario against the randomised one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the Maximum Parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
-Identically, there was always a significant difference between the minimised scenario and the randomised one was expect for the matrix of 150 taxa and 100 characters under Maximum Parsimony for $NTS_{RF}$ and the matrix of 150 and 1000 characters under Bayesian inference for $NTS_{Tr}$.
+When looking at the combined effect of each parameters, the ``maximised'' and ``minimised'' scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
+The same differences are observed when comparing the ``maximised'' scenario against the ``randomised'' one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the Maximum Parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
+Identically, there was always a significant difference between the ``minimised'' scenario and the ``randomised'' one was expect for the matrix of 150 taxa and 100 characters under Maximum Parsimony for $NTS_{RF}$ and the matrix of 150 and 1000 characters under Bayesian inference for $NTS_{Tr}$.
 The full list of comparisons and summary statistics are available in the supplementary materials 3.
@@
 
-% This goes with former results (see Parins-Fukuchi in review.)
+\subsection{Effect of Character Differences on topology}
+As expected, there is an significant effect of the characters difference in the ability to recover the correct topology.
+The Character Difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high Character Difference approximates a low level of character correlation and vice versa.
+When characters are correlated, one could expect the matrices to convey a really strong (but potentially misleading) phylogenetic signal since every characters agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasies.
+Intuitively, this would lead to the ``minimised'' Character Difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
+Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
 
-\subsection{Effect of Character Differences on topology}
-One way to look at the $CD$ metric is as a inverse of character correlation (see Methods \ref{CD_description}): a high $CD$ approximates a low level of character correlation and vice versa.
-When characters are correlated, one could expect the matrices to convey a really strong (but probably incorrect) phylogenetic signal (since every characters agrees with each other) and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasies.
-Intuitively this would lead to badly resolved trees in our scenarios where the $CD$ was high.
+Following these results, we encourage researchers to code characters that appear to be the most un-correlated \textit{a priori}.
+However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the Character Differences.
+Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e.
+as an equivalent of randomising the Character Differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).
 
-One first noteworthy point is that, on average, the matrices with the highest $CD$ where the ones that performed the best at recovering the ``normal'' topology which in turns confirms the intuition that discrete morphological characters should be independent in phylogenetics.
-This is also true when the characters are more similar to each other (low $CD$) resulting in worth trees due to the software treating all these characters as correlated.
-One second noteworthy point is that compared to simply randomising $CD$ in the matrices (effectively bootstrapping), maximising the $CD$ was, at least in small sized matrices, not more effective to recover the ``normal'' topology.
-Since many matrices are in the range size \citep{guillerme2016assessment} it is actually encouraging that actively coding characters to have high $CD$ values will not significantly improve the quality of the matrix compared to actually randomly coding characters.
-Coding characters with low $CD$ however will decrease the quality of the matrix.
+\subsubsection{Effect of the number of characters and taxa}
+Because of the nature of our simulation protocol, one could expect that the effect of character correlation would have increased with the number of characters (i.e.
+the more characters available, the more characters are modified in each scenario).
+Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance.)
 
-\subsubsection{Effect of the number of characters}
-As expected by the sheer design of this research project, the number of character has a nested effect on character correlation: increasing the number of characters in our simulation scenarios (i.e.
-``maximised'', ``minimised'' and ``randomised'') effectively increased the amount of high/low $CD$ values per matrix.
-This increase in character numbers thus increases the effect of $CD$ on tree topology with $CD$
-
-\subsubsection{Effect of the number of taxa}
-Same blabla (probably).
+Although we measured a significant difference between ``small'' a larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distribution that is always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
+This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
+For the number of characters, this suggests that the overall Character Difference metric is a good proxy for character correlation and is independent of the number of characters analysed.
+Similarly, using the a Normalised Tree Similarity metric ($NTS$) allows to account for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
+we here corrected for the expected difference when comparing two random trees with the same number of taxa).
 
 \subsection{Effect of Character Differences on the inference method}
-The rather strong nested effect of the method for recovering can be due to the sheer difference in both methods.
-These difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will more readily recover the ``right'' topology when analysing discrete morphological alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have been made in favour of parsimony: \citealt{spencerefficacy2013,goloboff2017weighted}).
+When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the Maximum Parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
+The difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will more readily recover the ``right'' topology when analysing discrete morphological alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have been made in favour of parsimony: \citealt{spencerefficacy2013,goloboff2017weighted}).
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e.
@@
 
-That being said, it is interesting to analysis this difference in tree inference method in the light of the effect of high/low $CD$.
+This being said, it is interesting to analysis this difference in tree inference method in the light of the effect of high/low Character Differences.
 In fact the differences observed here could be due to the inherent mechanisms of each method.
-When many characters are homoplasic (high $CD$) parsimony will cope with this information by generating longer trees (which will be more likely to be discarded in the tree search).
-Conversely Bayesian can cope with homoplasy more easily by increasing some branch length to accommodate more changes and will also incorporate sub-optimal trees in the final results.
+When many characters are homoplasic (high character difference) parsimony will cope with this information by generating longer trees (which will be more likely to be discarded in the tree search).
+Conversely, Bayesian analysis can (1) cope with homoplasy more easily by increasing branch length to accommodate more changes and (2) will also incorporate sub-optimal trees in the final results.
+These two points can explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.
 
 \subsection{Distinction between different character correlations}
-Although both biological and evolutionary induce correlation are what most evolutionary biologists are genuinely interested in and that coding induced correlation is readily avoided through best practices, software induced character correlation would be the main source of character dependence in systematic biology.
+The two types of correlations that most evolutionary biologists are genuinely interested in (biological and evolutionary) can only be studied \textit{a posteriori} using phylogenetic trees and should not be used as an \textit{a priori} criterion to validate/invalidate characters.
+The links between specific characters should studied explicitly through hypothesis testing based on an underlying phylogenetic framework - for example in evo-devo \citep[e.g.][]{goswami2006morphological} or in macroevolution \citep[e.g.][]{fitzjohn2014much} - and should thus be independent of the phylogenetic inference itself.
+In practice, therefore, only the correlation induced by data collection (i.e. coding correlation) can really affect the phylogeny \textit{a priori}.
 
-As we mentioned in the introduction of this study, we identified three different major sources of character correlation (biological, evolutionary and coding induced correlation).
-In practice, however, we argue that software induced character correlation would be the main source of character dependence in systematic biology (see Table \ref{Tab:example_matrix}).
-Therefore, in this study, we mainly focused on this source of correlation between characters.
-It is important to note, however, that because the characters were simulated using evolutionary models based on Birth-Death trees, evolutionary induced correlation is present in our simulation (though not analysed specifically).
-Conversely, biological induced correlation would be more complex to simulate and are out of the scope of this analysis.
+This dichotomy thus creates a trade of between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding; and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation being less dependent on the coding correlations.
 
-Additionally, the two types of correlations that most evolutionary biologists are genuinely interested in (biological and evolutionary) can only be studied \textit{a posteriori} using phylogenetic trees and should not be used as an \textit{a priori}.
-In fact, the links between specific characters should studied explicitly through hypothesis testing based on an underlying phylogenetic framework - for example in evo-devo \citep[e.g.][]{goswami2006morphological} or in macroevolution \citep[e.g.][]{fitzjohn2014much} - and should thus be independent of the phylogenetic inference itself.
-
-%TG: This one can be reworded. Get a dichotomie a priori/a posteriori correlation. + add trade off in coding (more characters = more info but less independence)
-Furthermore it is important to distinguish how researchers can tackle these correlations: the coding dependence is by essence an \textit{a priori} source of correlation: the correlation is induced before the phylogenetic analysis.
-On the other hand Biological and Evolutionary dependences can only be determined \textit{a posteriori}: the correlation must be based on a phylogenetic hypothesis.
-Additionally there seem to be a trade in coding fewer characters (reducing \textit{a priori} dependence) but making the \textit{a posteriori} correlation more dependent on the coding.
-On the other hand, coding more characters increases \textit{a priori} dependence but will lead to the \textit{a posteriori} correlation being less dependent on it.
+Effectively, this study focuses on the effect of coding induced correlation since our scenarios to maximise/minimise character difference are equivalent to actively choosing characters that are the most different/similar \textit{prior} to any phylogenetic analysis.
+The evolutionary correlation is also present in this simulations since the characters are modelled on Birth-Death trees.
+Similarly, biological induced correlation (i.e. modularity) can be present in the matrices for those characters randomly simulated with similar evolutionary simulation regimes.
+However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest.
 
 \subsection{Conclusion}
-As we presented in the introduction, dependence between characters can be induced through three main phenomenon: the biological, evolutionary or coding induced correlation between characters.
+Correlation between characters can be induced through three main phenomenon: biological features, evolutionary history or biases in coding the characters.
 Although one could argue (righteously!) that coding induced correlation is easily avoidable and that biological and evolutionary correlations are actually at the core of evolutionary biology, in the age of ever expanding morphological matrices \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the dependence between characters to increase stochastically.
@@
 
-We found that Character Differences as a proxy for character correlation is not a problem \textit{per se} as long as characters are not actively coded to be similar.
+We found that Character Differences as a proxy for character correlation have a strong effect on the recovering the ``normal'' topology
+
+ is not a problem \textit{per se} as long as characters are not actively coded to be similar.
 On the other hand, actively selecting characters based to favour independence does not highly increase the ability of recovering tree topology (compared to simply bootstrapping a matrix - at least for modest size matrices).

### 2017-09-27 Writing/draft.tex
Conclusion updated

@@
 
-We found that Character Differences as a proxy for character correlation have a strong effect on the recovering the ``normal'' topology
+We found that Character Differences as a proxy for character correlation have a strong effect on the recovering the ``normal'' topology: when character correlation was high (low Character Differences), the topology was always the furthest away from the ``normal'' topology.
+Conversely, when correlation between characters was low, the topology was always the closest to the ``normal'' topology.
+These results seems independent on the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
 
- is not a problem \textit{per se} as long as characters are not actively coded to be similar.
-On the other hand, actively selecting characters based to favour independence does not highly increase the ability of recovering tree topology (compared to simply bootstrapping a matrix - at least for modest size matrices).
+However, in modest size matrices (25 taxa 100 to 350 characters), the effect of actively choosing to minimise character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
 We hope that these results will encourage researchers working on discrete morphological characters to continue exploring morphology through discrete characters to better understand the Tree of Life.

### 2017-10-20 Writing/draft.tex
Removed capital letters for character difference

@@
 \begin{abstract}
-- When using discrete morphological characters for building phylogenies, the characters are assumed to be independent.
+Morphological characters for building phylogenies are assumed to be independent.
+However, this assumption is nearly always wrong simply because of the nature of morphological integration of organisms.
+Correlation between characters can originate from biological features, shared phylogenetic history or errors when coding or defining morphological characters.
+Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with good practices, phylogenetic software do not distinguish between any of them.
 
-- However, this assumption is nearly always wrong, simply because of the nature of morphological integration in organisms.
+In this study, we propose a new metric of raw character difference as a proxy for character correlation.
+Using thorough simulations, we test the effect of increasing or decreasing character differences on tree topology.
+Overall, we found an expected positive effect of reducing character correlations on recovering the correct topology.
+However, this effect is less important for matrices with a small number of taxa (~25) were reducing character correlation is not more effective than randomly drawing characters.
+Furthermore, in bigger matrices (~350 taxa), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
 
-- The dependence between characters can originate from biological features, shared phylogenetic history or errors from the researcher when coding characters.
-
-- Although the two first sources can be investigated by biologists post phylogenetic inference and the third one can be avoided pre inference with good practices, software do not distinguish between any of them.
-
-- In this study, we propose a new metric of raw character difference as a proxy for character (in)dependence.
-
-- Using thorough simulations and this metric, we test the effect of increasing or decreasing Character Difference on topology with a variable number of taxa and characters under both Bayesian and Maximum Parsimony inference.
-
-- Overall we found a positive effect of reducing character correlations on recovering the correct topology.
-
-- However, this effect is less important for matrices with a small number of taxa (~25) were reducing character correlation was not more effective than randomly drawing characters.
-
-- Furthermore, in bigger matrices (~350 taxa), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
-
-- These results suggest that ignoring the problem of character correlation or independence can easily impact topology in phylogenetic analysis.
-
-- However, encouragingly, they also suggest that, unless correlation is effectively maximised or minimised, probabilistic methods can easily accommodate for random correlation between characters.
+These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
+However, encouragingly, they also suggest that, unless correlation is actively maximised or minimised, probabilistic methods can easily accommodate for a random correlation between characters.
 
@@
 \noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding!
-The three types sources are however dissociated by later (coding dependence) being detected prior to phylogenetic inference and the two other ones (biological and evolutionary) being only detected and/or tested \textit{a posteriori} based on a phylogenetic hypothesis.
+The three types sources are however dissociated by the later (coding dependence) being detected prior to phylogenetic inference and the two other ones (biological and evolutionary) being only detected and/or tested \textit{a posteriori} based on a phylogenetic hypothesis.
 
@@
     \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modified the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
-    We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising Character Difference to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
+    We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character differences to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
 This step resulted in 36 matrices.
@@
 \end{enumerate}
-Each step was replicated 40 times and are described below in more details, as well as our proposed definition for measuring the difference between characters.
+Each step was replicated 40 times and are described below in more details, along with our proposed definition for measuring the difference between characters.
 
@@
 \caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using Maximum Parsimony and Bayesian inference (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
-$n$ and $m$ corresponds to the number of characters with a Character Difference $<0.25$ and $>0.75$ respectively.}
+$n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
 \label{Fig:outline}
@@
 \noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
-$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration Supplementary Material 1).
+$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
 Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we calculated the difference between character states in a qualitative way.
-Two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
+Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
 Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
@@
 
-When the Character Difference null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
-When the Character Difference maximal ($1$) it means it conveys the most different signal (i.e. characters are uncorrelated).
-It is important to stress that a Character Difference of $0$ (i.e. the same phylogenetic signal) does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal) .
+When the character difference null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
+When the character difference maximal ($1$) it means it conveys the most different signal (i.e. characters are uncorrelated).
+It is important to stress that a character difference of $0$ (i.e. the same phylogenetic signal) does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal) .
 For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
@@
 \subsection{Modifying the matrices}
-We calculated the pairwise Character Differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-We then modified the matrices to either maximise or minimise the pairwise Character Differences for each matrices using three different algorithms.
+We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
+We then modified the matrices to either maximise or minimise the pairwise character Differences for each matrices using three different algorithms.
 For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e.
-with an average Character Difference $<0.25$) and replaced them randomly by any of the remaining characters.
-This operation increased the overall pairwise Character Difference in the matrix thus making the characters more dissimilar.
-Conversely, for minimising the pairwise Character Differences, we selected the most dissimilar characters (i.e.
-with an average Character Difference $<0.75$) and randomly replaced them with the remaining ones.
+with an average character difference $<0.25$) and replaced them randomly by any of the remaining characters.
+This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
+Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e.
+with an average character difference $<0.75$) and randomly replaced them with the remaining ones.
 Finally, because this operation effectively changes the weight of characters (i.e.
@@
 \caption{Example illustration of the protocol for modifying matrices.
-The matrices represent the pairwise Character Differences for 100 characters.
-Blue colours correspond to low Character Differences and orange colours correspond to high Character Differences.
-\textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised Character Difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).}
+The matrices represent the pairwise character differences for 100 characters.
+Blue colours correspond to low character differences and orange colours correspond to high character differences.
+\textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised character difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).}
 \label{Fig:modif_matrix}
@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of Character Difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).
+We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).
 
@@
    \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
-\caption{\small{Effect of Character Difference on recovering the ``normal'' topology.
+\caption{\small{Effect of character difference on recovering the ``normal'' topology.
 The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
-The x axis represents the different Character Difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
+The x axis represents the different character difference scenarios and tree inference method with the maximised character difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised character difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised character difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
 \label{Fig:RF_results_best}
@@
    \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
-\caption{Effect of Character Difference on recovering the ``normal'' topology.
+\caption{Effect of character difference on recovering the ``normal'' topology.
 The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
@@
 
-\subsection{Effect of Character Differences on topology}
+\subsection{Effect of character differences on topology}
 
-The overall amount of Character Difference in a matrix has an effect of the ability to recover the correct topology with maximising the Character Difference leading to the less loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the Character Difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see supplementary material 3 for the full summary statistics).
-There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the Character Difference (Bhattacharrya coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the Character Difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology with maximising the character difference leading to the less loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the character difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see supplementary material 3 for the full summary statistics).
+There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
 
@@
 
-This effect of the Character Difference is not dependent on the number of characters for the clade conservation (i.e.
+This effect of the character difference is not dependent on the number of characters for the clade conservation (i.e.
 $NTS_{RF}$) for a ``similar'' number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.76, 0.78, 0.8 respectively - see supplementary materials 3) with a significant effect only between 100 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-The number of characters affects the Character Difference more in terms of taxa placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+The number of characters affects the character difference more in terms of taxa placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
 These difference have to be contrasted however by a really high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.
@@
 
-Similarly than the effect of number of characters on Character Difference, the number of taxa seems to have only a marginal effect.
+Similarly than the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
 A low number of taxa (25) resulted in significant differences in both $NTS_{RF}$ and $NTS_{Tr}$ (medians for 25, 75 and 150 taxa equals 0.8, 0.76, 0.75 $NTS_{RF}$ and 0.75, 0.59 and 0.59 $NTS_{Tr}$respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
@@
 
-\subsection{Effect of Character Differences on the inference method}
+\subsection{Effect of character differences on the inference method}
 
@@
 
-\subsection{Effect of Character Differences on topology}
+\subsection{Effect of character differences on topology}
 As expected, there is an significant effect of the characters difference in the ability to recover the correct topology.
-The Character Difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high Character Difference approximates a low level of character correlation and vice versa.
+The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a really strong (but potentially misleading) phylogenetic signal since every characters agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasies.
-Intuitively, this would lead to the ``minimised'' Character Difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
+Intuitively, this would lead to the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
 Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
@@
 Following these results, we encourage researchers to code characters that appear to be the most un-correlated \textit{a priori}.
-However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the Character Differences.
+However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
 Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e.
-as an equivalent of randomising the Character Differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).
+as an equivalent of randomising the character differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).
 
@@
 This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
-For the number of characters, this suggests that the overall Character Difference metric is a good proxy for character correlation and is independent of the number of characters analysed.
+For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation and is independent of the number of characters analysed.
 Similarly, using the a Normalised Tree Similarity metric ($NTS$) allows to account for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
@@
 
-\subsection{Effect of Character Differences on the inference method}
+\subsection{Effect of character differences on the inference method}
 When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the Maximum Parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
@@
 
-This being said, it is interesting to analysis this difference in tree inference method in the light of the effect of high/low Character Differences.
+This being said, it is interesting to analysis this difference in tree inference method in the light of the effect of high/low character differences.
 In fact the differences observed here could be due to the inherent mechanisms of each method.
@@
 
-We found that Character Differences as a proxy for character correlation have a strong effect on the recovering the ``normal'' topology: when character correlation was high (low Character Differences), the topology was always the furthest away from the ``normal'' topology.
+We found that character differences as a proxy for character correlation have a strong effect on the recovering the ``normal'' topology: when character correlation was high (low character differences), the topology was always the furthest away from the ``normal'' topology.
 Conversely, when correlation between characters was low, the topology was always the closest to the ``normal'' topology.

### 2017-10-23 Writing/draft.tex
Typos

@@
 \subsection{Simulating discrete morphological matrices}
-To simulate the matrices we applied a protocol very similar to \citep{Guillerme2016146}.
+To simulate the matrices we applied a protocol very similar to \cite{Guillerme2016146}.
 First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution maintaining $\lambda$ $>$ $\mu$ using the \texttt{diversitree} \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
@@
 We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-We then modified the matrices to either maximise or minimise the pairwise character Differences for each matrices using three different algorithms.
+We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
 For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e.
@@
 This step resulted in a total of 1440 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
-The algorithms for the three modifications are available on GitHub@@@. %(Modifying_matrices.html) %TG: "@@@" is my "to replace" sign.
+The algorithms for the three modifications are available on GitHub (\url{https://github.com/TGuillerme/CharactersCorrelation})
 
@@
 
-Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one, the resulting in the residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary @@@).
-%TG: TODO: residual normality + homscedasticity tests supp.
+Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one, the resulting in the residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary material 3).
 Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in their p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
@@
 Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
-Comparisons falling within these range can not be designated as strictly similar/different but can still be ranked (e.g.
+Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g.
 for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).
@@
 \section{Results}
-
-%TG: @@@ NOTE: THESE RESULTS ARE NOT FINAL - I thus didn't spend too much time describing them @@@
 
@@
 The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
-The x axis represents the different character difference scenarios and tree inference method with the maximised character difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised character difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised character difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
+The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under Maximum Parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
 \label{Fig:RF_results_best}
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology with maximising the character difference leading to the less loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the character difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see supplementary material 3 for the full summary statistics).
-There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology with maximising the character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the character difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see Supplementary material 3 for the full summary statistics).
+There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
 
@@
 \end{tabular}
-\caption{Difference between the pooled number of characters. See Table \ref{Tab_pooledscenarios_test} for caption.} 
+\caption{Difference between the pooled number of characters. See Table \ref{Tab_pooledscenarios_test} for full caption.} 
 \label{Tab_pooledscharacters_test}
@@
 \end{tabular}
-\caption{Difference between the pooled number of taxa. See Table \ref{Tab_pooledscenarios_test} for caption.} 
+\caption{Difference between the pooled number of taxa. See Table \ref{Tab_pooledscenarios_test} for full caption.} 
 \label{Tab_pooledstaxa_test}
@@
 \end{tabular}
-\caption{Difference between the pooled methods. See Table \ref{Tab_pooledscenarios_test} for caption.} 
+\caption{Difference between the pooled methods. See Table \ref{Tab_pooledscenarios_test} for full caption.} 
 \label{Tab_pooledsmethods_test}
@@
 When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the Maximum Parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
-The difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will more readily recover the ``right'' topology when analysing discrete morphological alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have been made in favour of parsimony: \citealt{spencerefficacy2013,goloboff2017weighted}).
+The difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will more readily recover the ``right'' topology when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have been made in favour of parsimony: \citealt{spencerefficacy2013,goloboff2017weighted}).
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e.
-those with in a $CI>0.26$ in a quick parsimony search;]{OReilly20160081} could also favour Parsimony analysis.
+those with in a $CI>0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour Parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
@@
 Conversely, Bayesian analysis can (1) cope with homoplasy more easily by increasing branch length to accommodate more changes and (2) will also incorporate sub-optimal trees in the final results.
-These two points can explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.
+These two points could explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.
 
@@
 The two types of correlations that most evolutionary biologists are genuinely interested in (biological and evolutionary) can only be studied \textit{a posteriori} using phylogenetic trees and should not be used as an \textit{a priori} criterion to validate/invalidate characters.
-The links between specific characters should studied explicitly through hypothesis testing based on an underlying phylogenetic framework - for example in evo-devo \citep[e.g.][]{goswami2006morphological} or in macroevolution \citep[e.g.][]{fitzjohn2014much} - and should thus be independent of the phylogenetic inference itself.
+The links between specific characters should be studied explicitly through hypothesis testing based on an underlying phylogenetic framework - for example in evo-devo \citep[e.g.][]{goswami2006morphological} or in macroevolution \citep[e.g.][]{fitzjohn2014much} - and should thus be independent of the phylogenetic inference itself.
 In practice, therefore, only the correlation induced by data collection (i.e. coding correlation) can really affect the phylogeny \textit{a priori}.

### 2017-10-30 Writing/draft.tex
added comment
adding cite

@@
 However - especially for discrete morphological data - this assumption of independence is probably violated more often than not due to the sheer nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
+% TG: add Brazeau et al. 201X somewhere?
 In practice, characters dependence can come from at least three different ways:
@@
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
-    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146}.
+    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
     This model assumes a equal transition rate between character states that might seem overly simplifying and excludes other observed transition patterns \citep[e.g.

### 2017-12-01 Writing/draft.tex
Updated affiliations

@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^*$, and Martin D. Brazeau$^1$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^1$$^2$$^*$, and Martin D. Brazeau$^2$$^3$}\\
 \noindent {\small \it 
-$^1$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\}
+$^1$School of Biological Sciences, University of Queensland, St. Lucia, Queensland, Australia.\\
+$^2$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\
+$^3$Department of Earth Sciences, Natural History Museum, Cromwell Road, London, SW75BD, United Kingdom.\\}
+
 \end{center}

### 2017-12-20 Writing/draft.tex
First intro edits/rewrites
Trying to focus the intro more directly on the problem and relying less on statements that resemble hearsay.

@@
 
-\noindent{\Large \bf Effect of discrete morphological character independence on tree topology}
+\noindent{\Large \bf Influence of different modes of morphological character correlation on tree inference}
 \bigskip
@@
 
-Since less than a decade, there has been a fantastic expansion of the use of discrete morphological data in macroevolutionary studies.
-Traditionally confined to ``classic'' cladistic studies in palaeontology, these methods have recently been widely expanded and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep[e.g.][]{pyrondivergence2011,ronquista2012}.
-In parallel, much development has also been done on the pre-phylogenetic analysis \citep[e.g.
+The last two decades have witnessed a resurgence of interest in the use of morphological character data in phylogenetic studies.
+This owes in large part to the use of fossils to untertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
+Morphological character data are often considered inferior to molecular sequence data, but are often the only source of data---especially for fossils.
+While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
+As morphological data are likely to continue to play an extensive role in phylogenetic analysis, it is essential to understand the circumstances under which morphological data might be expected to 'misbehave'.
+This opens up possibilities for predicting problematic datasets and possibly proposing new confidence measures in phylogenetic datasets.
+%The last decade has witnessed a resurgence of use of discrete morphological data in macroevolutionary studies.
+%Traditionally confined to ``classic'' cladistic studies in palaeontology, these methods have recently been widely expanded and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep[e.g.][]{pyrondivergence2011,ronquista2012}.
+%In parallel, much development has also been done on the pre-phylogenetic analysis \citep[e.g.
 data collection;][]{morphobank} as well as on post-phylogenetic ones \citep[e.g.
 morphological disparity analysis;][]{Close2015,Claddis}.
-Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011,simoes2017giant}; the topology inference method \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.
+%Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011,simoes2017giant}; the topology inference method \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.
 
-One of such caveats, however, is often neglected: the correlation between morphological characters.
-The assumption of character independence is at the base of phylogenetic inference and as been mentioned or implied throughout the history of phylogenetic methods \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-Nowadays, this assumption of character independence is nearly always accepted as a fact and rarely tested or even questioned in empirical datasets.
-However - especially for discrete morphological data - this assumption of independence is probably violated more often than not due to the sheer nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
+%One such caveat, however, is often neglected: the correlation between morphological characters.
+The non-independence of large numbers of morphological characters is often called upon in anticipation of problems with morphological data.
+The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony.
+\citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
+%Nowadays, this assumption of character independence is nearly always accepted as a fact and rarely tested or even questioned in empirical datasets.
+Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the sheer nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
 % TG: add Brazeau et al. 201X somewhere?
-In practice, characters dependence can come from at least three different ways:
+Before discussing character correlation further, it is important to understand that it may manifest itself in at least three distinct ways:
 

### 2017-12-21 Writing/draft.tex
Comments/changes on correl. types

@@
 %One such caveat, however, is often neglected: the correlation between morphological characters.
-The non-independence of large numbers of morphological characters is often called upon in anticipation of problems with morphological data.
+The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data.
 The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony.
@@
 
-    \item \textbf{Biological dependence:} this is the result of a link between two characters through their development and/or biological function.
-    These links are readily studied through developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
-    One trivial example would be the link between lower and upper molar characters in mammals: one character describing a feature of a lower molar will be expected to be correlated to another character describing another feature of the associated upper molar.
+    \item \textbf{Biological dependence:} this is the result of an intrinsic biological link between two characters through development, pleiotropy, and/or biological function.
+    %If it's a trivial example, then why use it?
+That said, it's unclear that this example would be genetic/developmental, and not fall under the second category.
+We should actually reconsider this example and how we express the difference between this and the next type of correlation.
+    %One trivial 
+    For example the lower and upper molar characters in mammals generally occlude one another. 
+    Therefore, one character describing a feature of a lower molar will be expected to be complemented by the surface of the occluding upper molar.
+    This will therefore provide a source of characters that would be expected to directly co-vary.
+     %They're not readily studied. They're actually quite difficult to study.
+    These links are usually the targets of comparative developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
 
-    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving along the same branch in a phylogeny: the character's correlation is thus due to their shared evolutionary history.
+    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving: the character's correlation is thus due to their shared evolutionary history.
+% MDB: Does the above really describe this correctly. This actually sounds a lot more like genuine synapomorphy than potentially problematic correlation.
+
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
-    For example, the apparition of the jaw as a character in gnatosthomes correlates with a whole sweep of characters present in modern gnatosthomes (e.g.
+    For example, the apparition of the jaw as a character in gnatosthomes correlates with a whole suite of characters present in modern gnatosthomes (e.g.
 beaks, tribosphenic molars, jaw protrusion, etc.).
 %TG: this links to inapplicable data!
+%MDB: The above is really coding dependence that you describe below, in a way.
+It's a character hierarchy.
+What we would like to discuss here is probably something more like limb reduction and axial elongation, for instance.
+These two things seem to co-vary extensively in tetrapods, with snake-like body forms evolving multiple times in numerous tetrapod lineages.
 

### 2017-12-24 Writing/draft.tex
minor changes following martin's comments

@@
 The last two decades have witnessed a resurgence of interest in the use of morphological character data in phylogenetic studies.
-This owes in large part to the use of fossils to untertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
-Morphological character data are often considered inferior to molecular sequence data, but are often the only source of data---especially for fossils.
+This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
+Morphological character data are often considered inferior to molecular sequence data (due to their inherent biases), but are often the only source of data---especially for fossils.
 While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
@@
 % MDB: Does the above really describe this correctly. This actually sounds a lot more like genuine synapomorphy than potentially problematic correlation.
-
+%TG: Yep, totally synapormophies.
+Although there is a subtle nuance in the way they are studied.
+In pure phylo approach (i.e.
+matrix -> tree), the synapomorphic characters are participating to the topology inference and are thus impossible to distinguish (i.e.
+the synapomorphies make the trees and the trees make the synapomorphies).
+In the papers cited just below they have a different approach.
+They assume a phylogeny and then "map" characters on to it.
+In this later case, the synapomorphies are effectively decoupled from the topology.
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
-    For example, the apparition of the jaw as a character in gnatosthomes correlates with a whole suite of characters present in modern gnatosthomes (e.g.
+    % For example, the apparition of the jaw as a character in gnatosthomes correlates with a whole suite of characters present in modern gnatosthomes (e.g.
 beaks, tribosphenic molars, jaw protrusion, etc.).
@@
 These two things seem to co-vary extensively in tetrapods, with snake-like body forms evolving multiple times in numerous tetrapod lineages.
+%TG: good call, I've changed the example.
+    For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
 
@@
 In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
-This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
+This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
@@
 \begin{itemize}
-    \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ and scaled to sum to $1$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
+    \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ and scaled to sum to $1$)
+    %TG: is that not a random Dirichlet? D(A,G,C,T)?
+     and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
     This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
@@
 
-This being said, it is interesting to analysis this difference in tree inference method in the light of the effect of high/low character differences.
+This being said, it is interesting to analyse this difference in tree inference method in the light of the effect of high/low character differences.
 In fact the differences observed here could be due to the inherent mechanisms of each method.

### 2018-01-05 Writing/draft.tex
comments on intro to work on

@@
 \begin{itemize}
-
-    \item \textbf{Biological dependence:} this is the result of an intrinsic biological link between two characters through development, pleiotropy, and/or biological function.
+%TODO: Change biological throughout
+    \item \textbf{Intra-organismal dependence:} this is the result of an intrinsic biological link between two characters through development, pleiotropy, and/or biological function.
     %If it's a trivial example, then why use it?
@@
      %They're not readily studied. They're actually quite difficult to study.
-    These links are usually the targets of comparative developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
+    These links can be the targets of comparative developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
+
+%TODO: add somewhere the problem that we use model organisms for comparative developmental bio which is shit. We may never be in a realistic position to actually know these interactions.
 
@@
     For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
+
+    %TODO: hierachical (logical dependency) problem (Wilkinson's papers 93-95 - Inapplicability).
+    %atomization: necessary evil; dataset size needed to overcome missing data
+    %biased density: e.g. skull: because they're complex, but localise information so high likelihood of type-1 correlation (cf a finite state space)
+    %                     things with high preservation potential (a finite state space, but also functional auto-correlation because the cause of high preservation potential can be e.g.
+biomechanical)
 \end{itemize}
 
-\noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding!
-The three types sources are however dissociated by the later (coding dependence) being detected prior to phylogenetic inference and the two other ones (biological and evolutionary) being only detected and/or tested \textit{a posteriori} based on a phylogenetic hypothesis.
+\noindent Of course, the three sources of dependence have an interaction: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding.
+Logical dependence, however, is easily distinguished prior to phylogenetic inference, while the two other ones (biological and evolutionary) are much harder.
+Intra-organismal dependence requires more detailed, often extremely time-consuming studies (and possibly beyond the limits of available technology).
+Even after all of the effort is expended, the results might then only be known for a single (model) species.
+Evolutionary dependence itself requires the resolution of a phylogenetic tree, and is best determined by independent character sets. 
+This is frequently accomplished by mapping morphological traits on molecular phylogenetic trees.
 
 These sources of dependence between characters are well studied in biology.
-Biological and evolutionary dependences are inherent parts to evo-devo and maroevolution and best practices to avoid coding induced dependences are commonly known and applied.
+Biological and evolutionary dependences are inherent parts to evo-devo and macroevolution and best practices to avoid coding-induced dependences are commonly known and applied.
 However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inferences software that are blind to these distinctions.
 This introduces a new, less studied, source of character dependence:
+
+%What the softwares are confronted with. Two dimensional matrix problem (renders the subtilties opac). The correlations are not distinguishable anymore.
 
@@
 Therefore a great number of characters and - traditionally - a few number of tokens can easily lead to dependence between characters.
-For example, if we consider the following matrix containing four cetartiodactyles - say a pig (e.g.
-\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley:
+For example, if we consider the following matrix containing four cetartiodactyls - say a pig (e.g.
+\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley:
 
@@
     \caption{Example of a matrix with software induced character correlation.
-\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley.}
+\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley.}
     \label{Tab:example_matrix}
@@
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e.
-those with in a $CI>0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour Parsimony analysis.
+those with in a $CI<0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour Parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.

### 2018-01-15 Writing/draft.tex
paper up

@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^2$$^*$, and Martin D. Brazeau$^2$$^3$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
 \noindent {\small \it 
@@
 However, this assumption is nearly always wrong simply because of the nature of morphological integration of organisms.
-Correlation between characters can originate from biological features, shared phylogenetic history or errors when coding or defining morphological characters.
+Correlation between characters can originate from intra-organismal features, shared phylogenetic history or errors when coding or defining morphological characters.
 Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with good practices, phylogenetic software do not distinguish between any of them.
@@
 
-%One such caveat, however, is often neglected: the correlation between morphological characters.
 The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data.
@@
 \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-%Nowadays, this assumption of character independence is nearly always accepted as a fact and rarely tested or even questioned in empirical datasets.
 Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the sheer nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
@@
 \begin{itemize}
-%TODO: Change biological throughout
     \item \textbf{Intra-organismal dependence:} this is the result of an intrinsic biological link between two characters through development, pleiotropy, and/or biological function.
-    %If it's a trivial example, then why use it?
-That said, it's unclear that this example would be genetic/developmental, and not fall under the second category.
-We should actually reconsider this example and how we express the difference between this and the next type of correlation.
-    %One trivial 
     For example the lower and upper molar characters in mammals generally occlude one another. 
@@
     This will therefore provide a source of characters that would be expected to directly co-vary.
-     %They're not readily studied. They're actually quite difficult to study.
     These links can be the targets of comparative developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
 
-%TODO: add somewhere the problem that we use model organisms for comparative developmental bio which is shit. We may never be in a realistic position to actually know these interactions.
-
     \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving: the character's correlation is thus due to their shared evolutionary history.
-% MDB: Does the above really describe this correctly. This actually sounds a lot more like genuine synapomorphy than potentially problematic correlation.
-%TG: Yep, totally synapormophies.
-Although there is a subtle nuance in the way they are studied.
-In pure phylo approach (i.e.
-matrix -> tree), the synapomorphic characters are participating to the topology inference and are thus impossible to distinguish (i.e.
-the synapomorphies make the trees and the trees make the synapomorphies).
-In the papers cited just below they have a different approach.
-They assume a phylogeny and then "map" characters on to it.
-In this later case, the synapomorphies are effectively decoupled from the topology.
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
-    % For example, the apparition of the jaw as a character in gnatosthomes correlates with a whole suite of characters present in modern gnatosthomes (e.g.
-beaks, tribosphenic molars, jaw protrusion, etc.).
-%TG: this links to inapplicable data!
-%MDB: The above is really coding dependence that you describe below, in a way.
-It's a character hierarchy.
-What we would like to discuss here is probably something more like limb reduction and axial elongation, for instance.
-These two things seem to co-vary extensively in tetrapods, with snake-like body forms evolving multiple times in numerous tetrapod lineages.
-%TG: good call, I've changed the example.
     For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
@@
     For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
-
-    %TODO: hierachical (logical dependency) problem (Wilkinson's papers 93-95 - Inapplicability).
-    %atomization: necessary evil; dataset size needed to overcome missing data
-    %biased density: e.g. skull: because they're complex, but localise information so high likelihood of type-1 correlation (cf a finite state space)
-    %                     things with high preservation potential (a finite state space, but also functional auto-correlation because the cause of high preservation potential can be e.g.
-biomechanical)
+    This coding dependency is linked to hierarchical dependency between characters [TODO: cite Wilkinson on inapplicability].
+    Finally this can also be due to a bias in the amount of characters available.
+    For example, in a skull, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity).
 \end{itemize}
@@
 \noindent Of course, the three sources of dependence have an interaction: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding.
-Logical dependence, however, is easily distinguished prior to phylogenetic inference, while the two other ones (biological and evolutionary) are much harder.
+Logical dependence, however, is easily distinguished prior to phylogenetic inference, while the two other ones (intra-organismal and evolutionary) are much harder.
 Intra-organismal dependence requires more detailed, often extremely time-consuming studies (and possibly beyond the limits of available technology).
@@
 However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inferences software that are blind to these distinctions.
+If fact, what the software are confronted with is a two dimensional matrix problem that renders the morphological subtleties described opaque.
 This introduces a new, less studied, source of character dependence:
-
-%What the softwares are confronted with. Two dimensional matrix problem (renders the subtilties opac). The correlations are not distinguishable anymore.
 
@@
 \subsection{Distinction between different character correlations}
-The two types of correlations that most evolutionary biologists are genuinely interested in (biological and evolutionary) can only be studied \textit{a posteriori} using phylogenetic trees and should not be used as an \textit{a priori} criterion to validate/invalidate characters.
+The two types of correlations that most evolutionary biologists are genuinely interested in (intra-organismal and evolutionary) can only be studied \textit{a posteriori} using phylogenetic trees and should not be used as an \textit{a priori} criterion to validate/invalidate characters.
 The links between specific characters should be studied explicitly through hypothesis testing based on an underlying phylogenetic framework - for example in evo-devo \citep[e.g.][]{goswami2006morphological} or in macroevolution \citep[e.g.][]{fitzjohn2014much} - and should thus be independent of the phylogenetic inference itself.
@@
 The evolutionary correlation is also present in this simulations since the characters are modelled on Birth-Death trees.
-Similarly, biological induced correlation (i.e. modularity) can be present in the matrices for those characters randomly simulated with similar evolutionary simulation regimes.
+Similarly, intra-organismal induced correlation (i.e. modularity) can be present in the matrices for those characters randomly simulated with similar evolutionary simulation regimes.
 However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest.
@@
 \subsection{Conclusion}
-Correlation between characters can be induced through three main phenomenon: biological features, evolutionary history or biases in coding the characters.
-Although one could argue (righteously!) that coding induced correlation is easily avoidable and that biological and evolutionary correlations are actually at the core of evolutionary biology, in the age of ever expanding morphological matrices \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the dependence between characters to increase stochastically.
+Correlation between characters can be induced through three main phenomenon: intra-organismal features, evolutionary history or biases in coding the characters.
+Although one could argue (righteously!) that coding induced correlation is easily avoidable and that intra-organismal and evolutionary correlations are actually at the core of evolutionary biology, in the age of ever expanding morphological matrices \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the dependence between characters to increase stochastically.
 Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.

### 2018-02-11 Writing/draft.tex
Small edits to intro; starting on methods

@@
     Therefore, one character describing a feature of a lower molar will be expected to be complemented by the surface of the occluding upper molar.
-    This will therefore provide a source of characters that would be expected to directly co-vary.
-    These links can be the targets of comparative developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
+	Characters of the occlusal surface of two opposing molars will be expected to directly covary.
+    These links can be the targets of comparative developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary} or functional investigations.
 
-    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving: the character's correlation is thus due to their shared evolutionary history.
+    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving due to selection: the character's correlation is thus due to their shared evolutionary history.
+	Unlike the case of intra-organismal dependence, there is no intrinsic constraint that causes these traits to covary.
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
     For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
+	Such covariances are generally studied in the context of a given phylogeny.
 
-    \item \textbf{Coding dependence:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
+    \item \textbf{Coding dependence:} this is the results of researcher methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
     One such source of correlation could be due to methodological errors: for example by defining a character describing a particular feature on a left and right molar.
@@
 
-We applied the following simulation protocol to assess the effect of characters correlation (Fig.\ref{Fig:outline}):
+We applied the following simulation protocol to assess the effect of character correlation on the accuracy of tree inference (Fig.\ref{Fig:outline}):
 \begin{enumerate}
-    \item \textbf{Simulating matrices}: we simulated a discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
+    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
 This step resulted in 9 matrices.

### 2018-02-11 Writing/draft.tex
comments/edits on beginning of methods

@@
 Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the different character states tokens will be used in order to infer a phylogenetic tree.
-It is thus still unclear what will be the effect of characters correlation on phylogenetic inference: how does these correlation really affect topology?
+% I'm unsure what point this introduction serves.
+It somehow doesn't quite connect with what you say next.
+I think we can simply focus on the fact that the phenomenon hasn't been empirically explored.
+% I think the point in this manuscript should be: OK, we believe two reasonable things: some characters will be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
+% The question then becomes: how bad does it have to be to be a real effect?
+% Along with this are subquestions:
+% If you've got intrinsic correlations, then those correlations could occur across anatomical systems. Would their effects cancel each other out?
+% Will "noisier" characters accurately cancel out the effects of correlation?
+
+It is thus still unclear what will be the effect of character correlation on phylogenetic inference: how does these correlation really affect topology?
 
@@
 
-We applied the following simulation protocol to assess the effect of character correlation on the accuracy of tree inference (Fig.\ref{Fig:outline}):
+To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different levels of correlation between some characters (Fig.\ref{Fig:outline}):
 \begin{enumerate}
@@
 This step resulted in 9 matrices.
-    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modified the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
+% HOW do you simulate them? What models do you use? What protocol? Are the simulated matrices expected to recover the correct topology in the absence of ourmanipulations? 
+    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
+% I don't understand this. Would a matrix with a "maximised" level of correlation simply be a uniform matrix?
+% What do you mean by "least" and "most different"? What is the measure of similarity? 
+% What do you mean "replacing them randomly by the remaining characters? Also: won't that simply create a correlation with another character in the dataset?
+
     We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character differences to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
 This step resulted in 36 matrices.
-    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
+
+\item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both maximum parsimony (MP) and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
 This step resulted in 72 topologies.
@@
 \end{enumerate}
-Each step was replicated 40 times and are described below in more details, along with our proposed definition for measuring the difference between characters.
+Each step was replicated 40 times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
 

### 2018-02-11 Writing/draft.tex
comments/edits on methods

@@
 
-\item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both maximum parsimony (MP) and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
-This step resulted in 72 topologies.
+\item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both maximum parsimony (MP) and Bayesian inference (BPP).
+Hereafter, the resultant topologies are called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees). This step resulted in 72 topologies.
     \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on topology.
@@
    \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
-\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using Maximum Parsimony and Bayesian inference (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
+\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using MP and BPP methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
 $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
@@
 In order to measure this, we propose a new distance metric to measure the difference between two characters:
+% Are you sure this is new? I can't read the formula because it's in LaTeX gobbledygook but there exist methods for measuring character similarity,
+% I can't be sure, but this could as well be Farris's (1970) advancement index as applied to characters rather than taxa?
 
-\subsubsection{Characters Difference ($CD$)}
+\subsubsection{Character Difference ($CD$)}
 \begin{equation}
@@
 \noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
+% What?
 $CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
-Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we calculated the difference between character states in a qualitative way.
+Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
 Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
@@
 This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
+% I don't get this part. What do you mean by standardized? And why are you doing this? Are you just editing the symbols?
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
@@
 First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution maintaining $\lambda$ $>$ $\mu$ using the \texttt{diversitree} \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
-For each tree, we arbitrarily set the outgroup to be the first taxa (alphabetically) thus effectively rooting the trees on this taxa.
+For each tree, we arbitrarily set the outgroup to be the first taxon (alphabetically) thus effectively rooting the trees on this taxon.
+% If we are using undirected characters, does it matter what the rooting is?
 These trees are hereafter called the ``true'' trees.
+% Yay! The Truth! Stop sciencing!
 We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
@@
      and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
-    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
+    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use an M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
     We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
-    This model assumes a equal transition rate between character states that might seem overly simplifying and excludes other observed transition patterns \citep[e.g.
+    This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g.
 Dollo characters;][]{Dollo,wright2015came}.
@@
 
-\noindent For each character, both models (morphological HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
-This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices.
-We re-simulated every invariant characters to obtain a matrix with no invariant characters.
-To ensure that our simulation where reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{OReilly20160081}.
+\noindent For each character, both models (morphological HKY-binary or M$k$) where chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
+These low evolutionary rate values allowed reduction in the number of homoplasic character changes, thus reinforcing the phylogenetic information in the matrices.
+We re-simulated every invariant characters to obtain a matrix with no invariant characters in order to better approximate real morphological data matrices.
+To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{OReilly20160081}.
 
-For each trees with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
-The matrices where generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
+For each tree with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
+The matrices were generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
 To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 40 times resulting in 360 ``normal'' morphological matrices.
@@
 \subsection{Modifying the matrices}
-We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
-For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e.
+We calculated the pairwise character differences for each generated matrix using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
+We then modified the matrices to either maximise or minimise the pairwise character differences for each matrix using three different algorithms.
+For maximising the pairwise differences between characters, we selected the characters that were the most similar to all the others (i.e.
 with an average character difference $<0.25$) and replaced them randomly by any of the remaining characters.
@@
 randomly weighting characters).
+% Clever.
 Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
@@
 \subsection{Inferring topologies}
-We inferred the topologies in both Bayesian and Maximum Parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
+We inferred the topologies with both BPP and MP using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
 For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
 The Maximum Parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
+% I'm not going to change all these, but "maximum parsimony" isn't a proper noun, so I wouldn't capitalise it.
 
-The Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$ - \texttt{lset nst=1 rates=gamma Ngammacat=4}) with an variable rate prior an exponential (0.5) shape (\texttt{prset ratepr=variable Shapepr=Exponential(0.5)}).
+Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$ - \texttt{lset nst=1 rates=gamma Ngammacat=4}) with an variable rate prior an exponential (0.5) shape (\texttt{prset ratepr=variable Shapepr=Exponential(0.5)}).
 We ran two runs of 6 chains each (2 hot, 4 cold) for a maximum of $1\times10^9$ generations with a sampling every 200 generations.

### 2018-03-15 Writing/CoverLetter.tex
Cover letter INIT

\documentclass[11pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
\usepackage[osf]{mathpazo}
\signature{Thomas Guillerme \\ Martin Brazeau}
\address{The University of Queensland \\School of Biological Sciences \\St Lucia QLD 4067, Australia \\guillert@tcd.ie}
\longindentation=0pt
\begin{document}

\begin{letter}{}
\opening{Dear Editors,}

Despite being at the base of almost every endeavour in modern Biology, the assumption made in tree inference that phylogenetic data is independent is probably mostly erroneous.
In fact, since the advent of Cladistics in the 1950s, evolutionary biologists have made the assumption that characters used to build a phylogeny are independent despite the fact that we are looking at their variance-covariance to build phylogenetic hypothesis.
This effect is somewhat negligible in molecular phylogenetics because the huge amount of available data is thought to temper down the correlation effect.
However, in smaller morphological datasets, this can lead to pretty big biases and can easily favour on topology from another simply with a ``judicious'' choice of characters.

Our research article, entitled ``Influence of different modes of morphological character correlation on tree inference'', is - to our knowledge - the first to thoroughly test the effect of character correlation on tree topology using a thorough simulation framework ($>$ 100 CPU years worth).
Using a novel metric to measure the amount of phylogenetic correlation in a matrix, we tested different scenario that either increase or decrease the amount of correlation of characters generated from a known ``true'' tree and compare how these levels of correlation affect our ability to recover this topology using both Maximum Parsimony and Bayesian inference.

Our results suggests that strong levels of correlation can negatively impact our ability to recover the correct tree topology and more so in Maximum Parsimony than in Baysian.
However, it is encouraging to see that low levels of correlation do not especially impact the topology, at least not to the extant where researchers should specifically collect only uncorrelated characters.
Finally, we show that, for smaller datasets, Maximum Parsimony and Bayesian inference are both pretty efficient to recover the correct topology, regardless of the amount of correlation

We look forward to hearing from you soon,

\closing{Yours sincerely,}

\end{letter}
\end{document}

%Editors:
%Sam Price, Daniele Silvestro

%Reviewers
%Graeme Lloyd, Dave Bapst, Emma Sherratt

### 2018-03-20 Writing/draft.tex
Comments edits on end of results and discussion

@@
 \noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
-% What?
+% What? Also, do you mean the taxon i? (not "taxa i").
 $CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
@@
 
-When the character difference null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
-When the character difference maximal ($1$) it means it conveys the most different signal (i.e. characters are uncorrelated).
+When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
+When the character difference is maximal ($1$) it means it conveys the greatest difference in phylogenetic signal (i.e. characters are uncorrelated).
 It is important to stress that a character difference of $0$ (i.e. the same phylogenetic signal) does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal) .
@@
 \subsection{Comparing topologies}
-We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
+We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
 We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).
 
-The metrics scores where calculated using the \texttt{TreeCmp} javascript \citep{Bogdanowicz2012}.
+The metric scores where calculated using the \texttt{TreeCmp} javascript \citep{Bogdanowicz2012}.
 The measurements where then standardised using the Normalised Tree Similarity metric \citep[$NTS$; i.e.
@@
 When the normalised metric has a score of one it means both trees are identical, when it has a score of zero it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
-The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Fould ($NTS_{RF}$) Similarity reflects the conservation of clades (i.e.
+The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Foulds ($NTS_{RF}$) Similarity reflects the conservation of clades (i.e.
 a score close to $1$ indicates that most clades are identical in both trees); and (2) the Normalised Triplets Similarity ($NTS_{Tr}$) reflects the position of taxa (i.e.
@@
 Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one, the resulting in the residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary material 3).
-Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in their p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
+%"the resulting in residuals..." I'm not sure what this sentence is supposed to say, so I can't suggest an edit.
+Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
 when multiple trees are equal to the ``normal'' tree).
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology with maximising the character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the character difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see Supplementary material 3 for the full summary statistics).
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology with maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the character difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see Supplementary material 3 for the full summary statistics).
 There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
@@
 \subsubsection{Effect of the number of characters}
+% Here and other similar subsection headings: the effect on what?
 
@@
 $NTS_{RF}$) for a ``similar'' number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.76, 0.78, 0.8 respectively - see supplementary materials 3) with a significant effect only between 100 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-The number of characters affects the character difference more in terms of taxa placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-These difference have to be contrasted however by a really high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.
+% The above sentence has too many "for"s. It's too long and needs to be split to be intelligible.
+The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.
 
@@
 
-Similarly than the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
+Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
 A low number of taxa (25) resulted in significant differences in both $NTS_{RF}$ and $NTS_{Tr}$ (medians for 25, 75 and 150 taxa equals 0.8, 0.76, 0.75 $NTS_{RF}$ and 0.75, 0.59 and 0.59 $NTS_{Tr}$respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
@@
 
-Regarding the inference method, there is a significant difference in clade conservation between Bayesian and Maximum Parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.82 and 0.68 respectively) but not in terms of individual taxa placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.72 and 0.61 respectively).
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and Maximum Parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.82 and 0.68 respectively) but not in terms of individual taxon placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.72 and 0.61 respectively).
 
@@
 
-When looking at the combined effect of each parameters, the ``maximised'' and ``minimised'' scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
+When looking at the combined effect of each parameter, the ``maximised'' and ``minimised'' scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
 The same differences are observed when comparing the ``maximised'' scenario against the ``randomised'' one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the Maximum Parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
@@
 \subsection{Effect of character differences on topology}
-As expected, there is an significant effect of the characters difference in the ability to recover the correct topology.
+As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
 The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
-When characters are correlated, one could expect the matrices to convey a really strong (but potentially misleading) phylogenetic signal since every characters agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasies.
-Intuitively, this would lead to the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
+When characters are correlated, one could expect the matrices to convey a really strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
+Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
 Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
@@
 Following these results, we encourage researchers to code characters that appear to be the most un-correlated \textit{a priori}.
+% Do we? I'm not sure that I would. In practice, this would probably force morphologists to remove most of their characters.
+% Making these kinds of recommendations leads to the risk that all you do is introduce biases into the dataset in another way.
 However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
+% How small? Be specific.
 Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e.
@@
 
-Although we measured a significant difference between ``small'' a larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distribution that is always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
+Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distribution that is always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
 This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
 For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation and is independent of the number of characters analysed.
-Similarly, using the a Normalised Tree Similarity metric ($NTS$) allows to account for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
-we here corrected for the expected difference when comparing two random trees with the same number of taxa).
+Similarly, using the a Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
+we corrected for the expected difference when comparing two random trees with the same number of taxa).
 
@@
 When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the Maximum Parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
-The difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will more readily recover the ``right'' topology when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have been made in favour of parsimony: \citealt{spencerefficacy2013,goloboff2017weighted}).
+The difference in the ability of each method to recover the ``correct'' topology has been heavily discussed in the last five years with some, indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have raised issues with these investigations \citealt{spencerefficacy2013,goloboff2017weighted}).
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e.
@@
 This being said, it is interesting to analyse this difference in tree inference method in the light of the effect of high/low character differences.
-In fact the differences observed here could be due to the inherent mechanisms of each method.
+% This is serious hedging ("we're not going to... but we're going to"). Could probably skip the above line.
+The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
 When many characters are homoplasic (high character difference) parsimony will cope with this information by generating longer trees (which will be more likely to be discarded in the tree search).
+% This doesn't make sense. Longer than one? How is generating longer trees coping? What do you mean by "coping"? 
 Conversely, Bayesian analysis can (1) cope with homoplasy more easily by increasing branch length to accommodate more changes and (2) will also incorporate sub-optimal trees in the final results.
 These two points could explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.
+% I'm not sure I'm convinced by this paragraph. Perhaps we should have a chat about it.
 
@@
 The two types of correlations that most evolutionary biologists are genuinely interested in (intra-organismal and evolutionary) can only be studied \textit{a posteriori} using phylogenetic trees and should not be used as an \textit{a priori} criterion to validate/invalidate characters.
+% Not sure I follow this line. Maybe there's another way to phrase it.
 The links between specific characters should be studied explicitly through hypothesis testing based on an underlying phylogenetic framework - for example in evo-devo \citep[e.g.][]{goswami2006morphological} or in macroevolution \citep[e.g.][]{fitzjohn2014much} - and should thus be independent of the phylogenetic inference itself.
 In practice, therefore, only the correlation induced by data collection (i.e. coding correlation) can really affect the phylogeny \textit{a priori}.
+% I don't understand this paragraph.
 
@@
 Effectively, this study focuses on the effect of coding induced correlation since our scenarios to maximise/minimise character difference are equivalent to actively choosing characters that are the most different/similar \textit{prior} to any phylogenetic analysis.
-The evolutionary correlation is also present in this simulations since the characters are modelled on Birth-Death trees.
+% This basically means choosing the dataset with the highest homoplasy... which will, in the long run, fuck up any method.
+% From a structural point of view, this line is out of place even.
+It makes a general point, but deep in the subsection.
+Such statements are better made earlier so a reader knows what the point of the section is.
+Evolutionary correlation is also present in this simulations since the characters are modelled on Birth-Death trees.
 Similarly, intra-organismal induced correlation (i.e. modularity) can be present in the matrices for those characters randomly simulated with similar evolutionary simulation regimes.

### 2018-03-20 Writing/draft.tex
Comments on conclusion

@@
 \subsection{Conclusion}
-Correlation between characters can be induced through three main phenomenon: intra-organismal features, evolutionary history or biases in coding the characters.
-Although one could argue (righteously!) that coding induced correlation is easily avoidable and that intra-organismal and evolutionary correlations are actually at the core of evolutionary biology, in the age of ever expanding morphological matrices \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the dependence between characters to increase stochastically.
+Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters.
+Although one could argue that coding-induced correlation is easily avoidable and that intra-organismal and evolutionary correlations are actually at the core of evolutionary biology, in the age of ever expanding morphological matrices \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the dependence between characters to increase stochastically.
+% I have no idea what thae above sentence means.
 Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
 
-We found that character differences as a proxy for character correlation have a strong effect on the recovering the ``normal'' topology: when character correlation was high (low character differences), the topology was always the furthest away from the ``normal'' topology.
+We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was high (low character differences), the topology was always the furthest away from the ``normal'' topology.
 Conversely, when correlation between characters was low, the topology was always the closest to the ``normal'' topology.
-These results seems independent on the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
+These results seem independent on the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
 
-However, in modest size matrices (25 taxa 100 to 350 characters), the effect of actively choosing to minimise character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
+However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
+% But at what point? How much correlation is needed to affect it?
 We hope that these results will encourage researchers working on discrete morphological characters to continue exploring morphology through discrete characters to better understand the Tree of Life.
+% This line sounds like something from an undergraduate term paper.
 

### 2018-03-21 Writing/draft.tex
Kind of tackled the comments (round 1)

@@
 The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data.
-The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony.
+The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony .
 \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
@@
 Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the different character states tokens will be used in order to infer a phylogenetic tree.
-% I'm unsure what point this introduction serves.
+%MB: I'm unsure what point this introduction serves.
 It somehow doesn't quite connect with what you say next.
 I think we can simply focus on the fact that the phenomenon hasn't been empirically explored.
-% I think the point in this manuscript should be: OK, we believe two reasonable things: some characters will be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
-% The question then becomes: how bad does it have to be to be a real effect?
-% Along with this are subquestions:
-% If you've got intrinsic correlations, then those correlations could occur across anatomical systems. Would their effects cancel each other out?
-% Will "noisier" characters accurately cancel out the effects of correlation?
+%MB: I think the point in this manuscript should be: OK, we believe two reasonable things: some characters will be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
+%MB: The question then becomes: how bad does it have to be to be a real effect?
+%MB: Along with this are subquestions:
+%MB: If you've got intrinsic correlations, then those correlations could occur across anatomical systems. Would their effects cancel each other out?
+%MB: Will "noisier" characters accurately cancel out the effects of correlation?
 
@@
 If the three sources of correlation are fairly well studied, their actual interpretation by phylogenetic inference software, has, to our knowledge, never been studied through a simulation protocol (although they have been explored empirically for morphological data; \citealt{Davalos01072014}; or molecular data; \citealt{ZouConvergence}).
-In this study, we formally assess the effect of discrete character's correlation using simulated data.
+Effectively, this study focuses on the effect of coding induced correlation since our scenarios to maximise/minimise character difference are equivalent to actively choosing characters that are the most different/similar \textit{prior} to any phylogenetic analysis.
+We formally assess the effect of discrete character's correlation using simulated data.
 We propose a new metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences/similarities between characters.
@@
 
-To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different levels of correlation between some characters (Fig.\ref{Fig:outline}):
+To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different levels of correlation between some characters (Fig.\ref{Fig:outline} - note that each step is described in more details below):
 \begin{enumerate}
@@
 This step resulted in 9 matrices.
-% HOW do you simulate them? What models do you use? What protocol? Are the simulated matrices expected to recover the correct topology in the absence of ourmanipulations? 
-    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
-% I don't understand this. Would a matrix with a "maximised" level of correlation simply be a uniform matrix?
-% What do you mean by "least" and "most different"? What is the measure of similarity? 
-% What do you mean "replacing them randomly by the remaining characters? Also: won't that simply create a correlation with another character in the dataset?
+%MB: HOW do you simulate them? What models do you use? What protocol? Are the simulated matrices expected to recover the correct topology in the absence of our manipulations? 
+%TG: I've specified above that this is just a summary.
+Details are below.
+Regarding the question, the recovery of the correct topology without manipulation is not perfect (of course) but ranges roughly between 0.2 and 1 (in both normalised triplets and RF).
+In a way though it's not that much of a big deal since the "True" tree is never known.
+Here we compare the modified matrices to the "Best" tree (the one we obtain by infering the tree from the simulated and unmodified "normal" matrices).
 
+    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least different or most different characters and replacing them randomly by the remaining characters.
+Our protocol for measuring character difference is detailed below.
+%MB: I don't understand this. Would a matrix with a "maximised" level of correlation simply be a uniform matrix?
+%MB: What do you mean by "least" and "most different"? What is the measure of similarity? 
+%MB: What do you mean "replacing them randomly by the remaining characters? Also: won't that simply create a correlation with another character in the dataset?
+%TG: Again, I've hihghlighted that precisions come below.
+To answer the questions, yes to both.
+However we do not remove/replace ALL the characters thus just changing the average correlation level.
+This is kind of illustrated in figure 2.
     We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character differences to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
@@
 In order to measure this, we propose a new distance metric to measure the difference between two characters:
-% Are you sure this is new? I can't read the formula because it's in LaTeX gobbledygook but there exist methods for measuring character similarity,
-% I can't be sure, but this could as well be Farris's (1970) advancement index as applied to characters rather than taxa?
+%MB: Are you sure this is new? I can't read the formula because it's in LaTeX gobbledygook but there exist methods for measuring character similarity,
+%MB: I can't be sure, but this could as well be Farris's (1970) advancement index as applied to characters rather than taxa?
+%TG: TODO: check Farris 1970 and whether this is new at all.
 
@@
 
-\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
-% What? Also, do you mean the taxon i? (not "taxa i").
+\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each characters states for the each $i^{th}$ taxon.
 $CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
@@
 Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
-We standardised each character by arbitrarily modifying their character state tokens by order of appearance.
+
+We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.
 In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
+%TG: Justification for the standardisation: basically, it makes comparisons faster and Fitch-like.
+This procedure was used to treat all the characters are unordered with no assumption on the meaning of the character state (e.g. in a binary character $0$ is not necessary ancestral to $1$).
+It also greatly improved the speed of our algorithm implementation to compare the characters.
 This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
-% I don't get this part. What do you mean by standardized? And why are you doing this? Are you just editing the symbols?
+%MB: I don't get this part. What do you mean by standardized? And why are you doing this? Are you just editing the symbols?
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
@@
 For each tree, we arbitrarily set the outgroup to be the first taxon (alphabetically) thus effectively rooting the trees on this taxon.
-% If we are using undirected characters, does it matter what the rooting is?
-These trees are hereafter called the ``true'' trees.
-% Yay! The Truth! Stop sciencing!
+%MB: If we are using undirected characters, does it matter what the rooting is?
+%TG: Not at all but it speeds up the MCMC
+These trees are hereafter called the ``true'' trees (see distinction below).
+%MB: Yay! The Truth! Stop sciencing!
 We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
@@
     \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ and scaled to sum to $1$)
-    %TG: is that not a random Dirichlet? D(A,G,C,T)?
+    %TG: is that not a random Dirichlet? D(A,G,C,T)? TODO: check
      and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
@@
 For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
-The Maximum Parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
-% I'm not going to change all these, but "maximum parsimony" isn't a proper noun, so I wouldn't capitalise it.
+The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
 
@@
 
-A strict majority rule tree was then calculated for both Bayesian an Maximum Parsimony trees.
+A strict majority rule tree was then calculated for both Bayesian an maximum parsimony trees.
 For the Bayesian consensus trees, the 25\% first trees of the posterior tree distribution were excluded as a burnin.
@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).
+We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).]
+%TG: True tree comparison stuff:
+Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
+First, in biology, this tree is by definition unknown.
+Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
+When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating birth-death trees simulating discrete morphological matrices from there.
 
@@
 
-Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one, the resulting in the residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary material 3).
-%"the resulting in residuals..." I'm not sure what this sentence is supposed to say, so I can't suggest an edit.
+Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one.
+The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary material 3).
+%MB: "the resulting in residuals..." I'm not sure what this sentence is supposed to say, so I can't suggest an edit.
+%TG: Probably a typo. Basically the residuals of lm or anovas cannot be normal because of this hard bound so we can't use parametric stats.
 Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
@@
 The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
-The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under Maximum Parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
+The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
 \label{Fig:RF_results_best}
@@
 
-\subsubsection{Effect of the number of characters}
-% Here and other similar subsection headings: the effect on what?
+\subsubsection{Number of characters}
+%MB: Here and other similar subsection headings: the effect on what?
+%TG: quick fix!
 
-This effect of the character difference is not dependent on the number of characters for the clade conservation (i.e.
-$NTS_{RF}$) for a ``similar'' number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.76, 0.78, 0.8 respectively - see supplementary materials 3) with a significant effect only between 100 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-% The above sentence has too many "for"s. It's too long and needs to be split to be intelligible.
+This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
+The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.76, 0.78, 0.8 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+%MB: The above sentence has too many "for"s. It's too long and needs to be split to be intelligible.
+%TG: fix attempt
 The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
@@
 
-\subsubsection{Effect of the number of taxa}
+\subsubsection{Number of taxa}
 
@@
 
-Regarding the inference method, there is a significant difference in clade conservation between Bayesian and Maximum Parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.82 and 0.68 respectively) but not in terms of individual taxon placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.72 and 0.61 respectively).
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.82 and 0.68 respectively) but not in terms of individual taxon placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.72 and 0.61 respectively).
 
@@
 
-\subsection{Combined effects}
+\subsection{Combined effects of taxa, characters and correlation on topology}
 
 When looking at the combined effect of each parameter, the ``maximised'' and ``minimised'' scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
-The same differences are observed when comparing the ``maximised'' scenario against the ``randomised'' one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the Maximum Parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
-Identically, there was always a significant difference between the ``minimised'' scenario and the ``randomised'' one was expect for the matrix of 150 taxa and 100 characters under Maximum Parsimony for $NTS_{RF}$ and the matrix of 150 and 1000 characters under Bayesian inference for $NTS_{Tr}$.
+The same differences are observed when comparing the ``maximised'' scenario against the ``randomised'' one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the maximum parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
+Identically, there was always a significant difference between the ``minimised'' scenario and the ``randomised'' one was expect for the matrix of 150 taxa and 100 characters under maximum parsimony for $NTS_{RF}$ and the matrix of 150 and 1000 characters under Bayesian inference for $NTS_{Tr}$.
 The full list of comparisons and summary statistics are available in the supplementary materials 3.
@@
 
-Following these results, we encourage researchers to code characters that appear to be the most un-correlated \textit{a priori}.
-% Do we? I'm not sure that I would. In practice, this would probably force morphologists to remove most of their characters.
-% Making these kinds of recommendations leads to the risk that all you do is introduce biases into the dataset in another way.
-However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
-% How small? Be specific.
-Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e.
+% Following these results, we encourage researchers to code characters that appear to be the most un-correlated \textit{a priori}.
+% %MB: Do we? I'm not sure that I would. In practice, this would probably force morphologists to remove most of their characters.
+% %MB: Making these kinds of recommendations leads to the risk that all you do is introduce biases into the dataset in another way.
+% %TG: Fair enough, I'm happy to drop this whole paragraph.
+% However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
+% %MB: How small? Be specific.
+% Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e.
 as an equivalent of randomising the character differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).
 
-\subsubsection{Effect of the number of characters and taxa}
+\subsubsection{Number of characters and taxa}
 Because of the nature of our simulation protocol, one could expect that the effect of character correlation would have increased with the number of characters (i.e.
@@
 \subsection{Effect of character differences on the inference method}
-When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the Maximum Parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
+When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
 The difference in the ability of each method to recover the ``correct'' topology has been heavily discussed in the last five years with some, indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have raised issues with these investigations \citealt{spencerefficacy2013,goloboff2017weighted}).
@@
 
-This being said, it is interesting to analyse this difference in tree inference method in the light of the effect of high/low character differences.
-% This is serious hedging ("we're not going to... but we're going to"). Could probably skip the above line.
+% This being said, it is interesting to analyse this difference in tree inference method in the light of the effect of high/low character differences.
+%MB: This is serious hedging ("we're not going to... but we're going to"). Could probably skip the above line.
+%TG: fair point
+% The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
+% When many characters are homoplasic (high character difference) parsimony will deal with this information by generating longer trees (i.e.
+) (which will be more likely to be discarded in the tree search).
+%MB: This doesn't make sense. Longer than one? How is generating longer trees coping? What do you mean by "coping"? 
+% Conversely, Bayesian analysis can (1) cope with homoplasy more easily by increasing branch length to accommodate more changes and (2) will also incorporate sub-optimal trees in the final results.
+% These two points could explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.
+%MB: I'm not sure I'm convinced by this paragraph. Perhaps we should have a chat about it.
+%TG: I've rephrased the parargaph below.
+This is the idea (a bit arms waving, I agree): for a given topology, when there is high homoplasy, parsimony will give the tree a high score (may changes per branches - even though branch length is not a parsimony parameter) and probably discard the topology.
+For the same topologyn and high homoplasy, Bayesian will stretch the branches (accounting for the many changes per branches) which will decrease the likelihood score but less fast than MP (because branch length is only *one* of the parameters accounting for the likelihood score).
+Also, in Bayesian, an decrease in likelihood score does not necessary mean that the tree will be discarded.
 The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
-When many characters are homoplasic (high character difference) parsimony will cope with this information by generating longer trees (which will be more likely to be discarded in the tree search).
-% This doesn't make sense. Longer than one? How is generating longer trees coping? What do you mean by "coping"? 
-Conversely, Bayesian analysis can (1) cope with homoplasy more easily by increasing branch length to accommodate more changes and (2) will also incorporate sub-optimal trees in the final results.
-These two points could explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.
-% I'm not sure I'm convinced by this paragraph. Perhaps we should have a chat about it.
+For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
+(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
+The tree search will thus likely result in a topology island that will not contain the given topology $T$.
+(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e.
+low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
+The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
+Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
+This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.
 
 \subsection{Distinction between different character correlations}
-The two types of correlations that most evolutionary biologists are genuinely interested in (intra-organismal and evolutionary) can only be studied \textit{a posteriori} using phylogenetic trees and should not be used as an \textit{a priori} criterion to validate/invalidate characters.
-% Not sure I follow this line. Maybe there's another way to phrase it.
-The links between specific characters should be studied explicitly through hypothesis testing based on an underlying phylogenetic framework - for example in evo-devo \citep[e.g.][]{goswami2006morphological} or in macroevolution \citep[e.g.][]{fitzjohn2014much} - and should thus be independent of the phylogenetic inference itself.
-In practice, therefore, only the correlation induced by data collection (i.e. coding correlation) can really affect the phylogeny \textit{a priori}.
-% I don't understand this paragraph.
+%TG: I've changed this paragraph - basically it's linking back to the intro saying the only correlation we should care about in phylogenetics is the one raising from data collection.
+Here we mention three different types of character correlations but evolutionary biologists are mainly interested in the intra-organismal and evolutionary correlations (e.g.
+in evo-devo \citealt{goswami2006morphological}; or in macroevolution \citealt{fitzjohn2014much}).
+These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not used \textit{a priori} as a criterion to select characters.
+In other words, intra-organisaml and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
+coding correlation) the only type of correlation that can affect the phylogeny \textit{a priori}.
+This dichotomy thus creates a trade of between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
+and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation being less dependent on the coding correlations.
 
-This dichotomy thus creates a trade of between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding; and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation being less dependent on the coding correlations.
-
-Effectively, this study focuses on the effect of coding induced correlation since our scenarios to maximise/minimise character difference are equivalent to actively choosing characters that are the most different/similar \textit{prior} to any phylogenetic analysis.
-% This basically means choosing the dataset with the highest homoplasy... which will, in the long run, fuck up any method.
-% From a structural point of view, this line is out of place even.
+% Effectively, this study focuses on the effect of coding induced correlation since our scenarios to maximise/minimise character difference are equivalent to actively choosing characters that are the most different/similar \textit{prior} to any phylogenetic analysis.
+TG: moved to intro.
+%MB: From a structural point of view, this line is out of place even.
 It makes a general point, but deep in the subsection.
 Such statements are better made earlier so a reader knows what the point of the section is.
-Evolutionary correlation is also present in this simulations since the characters are modelled on Birth-Death trees.
-Similarly, intra-organismal induced correlation (i.e. modularity) can be present in the matrices for those characters randomly simulated with similar evolutionary simulation regimes.
+It is important to note that the two other sources of character correlation could also be present in our simulations although they were not explicitly modelled:
+(1) evolutionary correlation is implied by simulating the characters based on Birth-Death trees;
+and (2) intra-organismal correlation could also be present in the matrices for those characters randomly simulated but sharing similar evolutionary simulation regimes (i.e.
+creating ``modules'' of characters).
 However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest.
@@
 \subsection{Conclusion}
-Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters.
-Although one could argue that coding-induced correlation is easily avoidable and that intra-organismal and evolutionary correlations are actually at the core of evolutionary biology, in the age of ever expanding morphological matrices \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the dependence between characters to increase stochastically.
-% I have no idea what thae above sentence means.
+Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon to investigate phylogenetic relationships.
+Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g.
+for characters hierarchy][]{deLaet2015,BrazeauNA} already exist.
+However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
+%MB: I have no idea what thae above sentence means.
+%TG: Modified. Is that clearer?
 Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
@@
 However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
-% But at what point? How much correlation is needed to affect it?
-We hope that these results will encourage researchers working on discrete morphological characters to continue exploring morphology through discrete characters to better understand the Tree of Life.
-% This line sounds like something from an undergraduate term paper.
+For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly decreased and increased correct topological recovery compared to randomly resample matrices.
+%MB: But at what point? How much correlation is needed to affect it? TG: added values.
+%TG: do we want to suggest something here to close the conclusion?
+I agree we don't wanna be presecriptive Bristolians but it'll be nice to finsh with something like: "we're not saying it's a problem per se but maybe check your data?"
 

### 2018-03-21 Writing/draft.tex
added caveats section

@@
 
+\subsection{Limitations}
+First, simulating evolutionary history is complex.
+Not only because the models we're using to infer phylogenies are ever improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible.
+This can lead to spuriously generated discrete morphological matrices
+However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree.
+This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.
+
+Second, measuring and modifying character correlation is difficult. 
+Increasing overall characters difference so that each pairwise differences is equal to 1 becomes more and more impossible with a large number of characters.
+In our algorithm, we maximised characters difference by replacing a characters $A$ highly correlated to another other character $B$ by a randomly chosen character $X$.
+This reduces the correlation between $A$ and $B$ but increases the correlation between $A$ and $X$!
+Additionally, when minimising $n$ character differences, the matrix will end containing only one character duplicated $n$ times (which will generate a unresolved tree by definition).
+Therefore we used an algorithm that only ``maximises'' or ``minimises'' character differences on average (see Fig. \ref{Fig:modif_matrix}).
+Although this was sufficient to generate detectable differences in topologies, it is not excluded that some randomly ``maximised'' or ``minimised'' matrices were actually not different than the ``randomised'' ones.
+
+Third, comparing phylogenetic inference methods is not trivial.
+As mentioned above, both maximum parsimony and Bayesian inference, although aiming (and often achieving) to infer evolutionary history only have similar outputs and vastly differ in how optimality is measured.
+But there are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
+However, we want to point out again that here we're no comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
+
 \subsection{Conclusion}

### 2018-03-21 Writing/draft.tex
Changed question paragraph in the intro

@@
 Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the different character states tokens will be used in order to infer a phylogenetic tree.
-%MB: I'm unsure what point this introduction serves.
-It somehow doesn't quite connect with what you say next.
-I think we can simply focus on the fact that the phenomenon hasn't been empirically explored.
-%MB: I think the point in this manuscript should be: OK, we believe two reasonable things: some characters will be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
-%MB: The question then becomes: how bad does it have to be to be a real effect?
-%MB: Along with this are subquestions:
+%MB: We believe two reasonable things: some characters will be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis. 
+Some character will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
+%MB: How bad does it have to be to be a real effect?
+Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence})
+How does these correlation really affect topology?
 %MB: If you've got intrinsic correlations, then those correlations could occur across anatomical systems. Would their effects cancel each other out?
+%TG: Actually not sure what you mean by this one and whether this is something we can test with this paper?
 %MB: Will "noisier" characters accurately cancel out the effects of correlation?
+Furthermore, we expect matrices with a high level of correlation to recover precise but inaccurate topologies but will matrices with low level of correlation (i.e.
+with high levels of homoplasy) actual cancel out the effects of correlation?
 
-It is thus still unclear what will be the effect of character correlation on phylogenetic inference: how does these correlation really affect topology?
-
-If the three sources of correlation are fairly well studied, their actual interpretation by phylogenetic inference software, has, to our knowledge, never been studied through a simulation protocol (although they have been explored empirically for morphological data; \citealt{Davalos01072014}; or molecular data; \citealt{ZouConvergence}).
-Effectively, this study focuses on the effect of coding induced correlation since our scenarios to maximise/minimise character difference are equivalent to actively choosing characters that are the most different/similar \textit{prior} to any phylogenetic analysis.
 We formally assess the effect of discrete character's correlation using simulated data.

### 2018-03-21 Writing/draft.tex
Comms + edits; going back to methods

@@
 \end{equation}
+%Why divide by 1/2 when you can just multiply by 2?
 
-\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each characters states for the each $i^{th}$ taxon.
+\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
 $CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
@@
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
+%MB: Splits rather than groups, really. Since, in a rooted tree, at most one symbol will need to define a clade. All others could be plesiomorphic states.
 
@@
 % However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
+%MB: Agreed that that is noteworthy.
+I suspect I know why this happens (and you might have explained it earlier).
+It might actually be a statistical problem with our approach.
+When character similaritiy is maximized, how well resolved are the trees?
 % %MB: How small? Be specific.
@@
 \subsection{Limitations}
+%MB: I think what this section _really_ needs is an assessment of the method used to simulate correlations.
+It was basically accomplished by duplicating characters.
+It wasn't accomplished by evolving some characters with a particular covariance model (which probably would've been better).
+We kind of faked-up that model.
 First, simulating evolutionary history is complex.
@@
 I agree we don't wanna be presecriptive Bristolians but it'll be nice to finsh with something like: "we're not saying it's a problem per se but maybe check your data?"
+%MB: Part of it is: we expect it to be a problem, but _when_ do we expect it to be a problem.
+There's the issue of "check your data", but it only raises the question: how?
+I think this is where we suggest avenues for future research (either here or move it up to the discussion).
+For instance, the simulations could be run in a way that more 'intelligently' places correlations in different parts of the tree.
+So, correlations are specifically modeled in an underlying stochastic process, for instance.
+We also don't really explore character congruence, that correlations can be ranked and spread over multiple internodes.
+We didn't look at ranking and we didn't look at competing correlations.
+There's a lot more that could be done here so, we might talk about that.
+Then we could talk about intelligent resampling measures, such as a targeted jackknifing that preferentially resamples characters that are highly similar to other characters etc.
 

### 2018-03-22 Writing/supplementary/Sup1-CharDiff.tex
equation fixes
changed CD equation

@@
 
+% \begin{equation}
+%     CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
+% \end{equation}
+
 \begin{equation}
-    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
+    CD_{(x,y)} = 1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}\right)
 \end{equation}
+
 \noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
@@
 
-For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ (because the difference between character states is considered as Fitch-like, i.e.
-the difference between two characters states \texttt{0} and \texttt{8} is equal to $1$ and the difference between two identical character states is equal to $0$):
+For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ because the difference between character states is considered as Fitch-like (unordered; i.e.
+the difference between two different characters states is equal to $1$ and the difference between two identical character states is equal to $0$):
 
@@
     0 \leq  \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} \leq \frac{1}{2}\\
-    0 \leq  \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} \leq 1\\
-    0 \leq  1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) \leq 1\\
+    0 \leq  2\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} \leq 1\\
+    0 \leq  1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}\right) \leq 1\\
     \end{gathered}
@@
     \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \frac{1}{2}\\
-    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= 1\\
-    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 0\\
+    2\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= 1\\
+    1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}\right) &= 0\\
     \end{aligned}
@@
     \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\\
-    \frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}} &= \frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\\
-    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right) &= 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)\\
+    2\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= 2\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\\
+    1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}\right) &= 1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\right)\\
     \end{aligned}
@@
 
-For any $n, j, b \in [1;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:
+For any $n, j, b \in [1;\infty)$ and $x_{n}$, $y_{j}$, $z_{b}$ $\in (0,1)$:
 
@@
     \text{if:}\\
-    1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
+     1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}\right)
      &\leq 
-     1 - \left(\frac{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}}}{\frac{1}{2}}\right) 
+     1 - 2\left(\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-z_{k}}}{j}-\frac{1}{2}}\right) 
      +
-     1 - \left(\frac{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}}}{\frac{1}{2}}\right) \\
+     1 - 2\left(\abs*{\frac{\sum_{a}^{b}\abs{z_{a}-y_{a}}}{b}-\frac{1}{2}}\right) \\
      \text{then, the opposite is:}\\
@@
      &\geq 
-     \underbrace{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-y_{k}}}{j}-\frac{1}{2}} }_{\in[0,1/2]}
+     \underbrace{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-z_{k}}}{j}-\frac{1}{2}} }_{\in[0,1/2]}
      +
-     \underbrace{\abs*{\frac{\sum_{a}^{b}\abs{x_{a}-y_{a}}}{b}-\frac{1}{2}} }_{\in[0,1/2]} - 2\\
+     \underbrace{\abs*{\frac{\sum_{a}^{b}\abs{z_{a}-y_{a}}}{b}-\frac{1}{2}} }_{\in[0,1/2]} - 2\\
      \text{thus:}\\
-     \in[0,\frac{1}{2}] &\geq \in[0,\frac{1}{2}] \in[0,\frac{1}{2}] (-2) \\
+     \in[0,\frac{1}{2}] &\geq \in[0,\frac{1}{2}] + \in[0,\frac{1}{2}] -2 \\
      \in[0,\frac{1}{2}] &\geq \in[-2,-1]\\

### 2018-03-22 Analysis/EffectCorrelationFigure.Rmd
added resolution plots

@@
 ```
+
+## Resolution
+
+How well resolved are the trees?
+These graphs represent the number of nodes (i.e. the tree resolution) in the consensus trees for 25, 75 or 50 taxa and 100, 350 or 1000 characters (c).
+The solid and dashed lines are respectively the 50% and 95% confidence intervals.
+The black dot is the median.
+
+```{r, fig.width=12, fig.height=6}
+## Nodes for the trees with 25 taxa
+plot.tree.resolution("../Data/Trees_out/Consensus_trees/", ntaxa = 25, col = c("darkgrey", "red", "green3", "blue", "grey", "orange", "lightgreen", "lightblue"))
+```
+
+```{r, fig.width=12, fig.height=6}
+## Nodes for the trees with 75 taxa
+plot.tree.resolution("../Data/Trees_out/Consensus_trees/", ntaxa = 75, col = c("darkgrey", "red", "green3", "blue", "grey", "orange", "lightgreen", "lightblue"))
+```
+
+```{r, fig.width=12, fig.height=6}
+## Nodes for the trees with 150 taxa
+plot.tree.resolution("../Data/Trees_out/Consensus_trees/", ntaxa = 150, col = c("darkgrey", "red", "green3", "blue", "grey", "orange", "lightgreen", "lightblue"))
+```

### 2018-03-22 Functions/plot.CI.R

@@ -1,0 +1,86 @@
+#' @title Plot confidence intervals
+#'
+#' @description Plot confidence intervals in a "boxplot" style as lines or polygons
+#'
+#' @param x a list of vectors
+#' @param type \code{"line"} or \code{"polygon"}
+#' @param CI the confidence intervals to plot (in percentages, \code{default = c(95, 50)}) 
+#' @param cent.tend the central tendency (\code{default = mean})
+#' @param col.x a vector of one or more colours to apply to x (default is \code{"black"} for lines and \code{"grey"} for polygons)
+#' @param ... Optional arguments to be handled by \code{\link[graphics]{polygon}, \code{\link[graphics]{lines} or \code{\link[graphics]{points} 
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+plot.CI <- function(x, type, CI = c(95, 50), cent.tend = mean, col.x, ...) {
+
+    ## Converts one or more CI into quantile probabilities
+    CI.converter <- function(CI) {
+        sort(c(50-CI/2, 50+CI/2)/100)
+    }
+
+    ## Convert the quantiles
+    quantiles <- CI.converter(CI)
+
+    ## Summarising the data
+    x_summary <- lapply(x, stats::quantile, probs = quantiles)
+
+    ## Number of plots
+    points_n <- length(x)
+    quantiles_n <- length(CI)
+
+    ## col.x
+    if(missing(col.x)) {
+        col.x <- ifelse(type == "line", "black", "grey")
+    }
+    if(length(col.x) != length(x)) {
+        col.x <- rep(col.x, length(x))
+    }
+
+    ## Options
+    dots <- list(...)
+
+    if(is.null(dots$border)) dots$border <- "black"
+    if(is.null(dots$point.col)) dots$point.col <- col.x
+    if(is.null(dots$width)) dots$width <- 0.5
+    if(is.null(dots$shift)) dots$shift <- 0
+    if(is.null(dots$shift)) dots$lwd <- 1.5
+
+    ## Plot polygons
+    if(type == "polygon") {
+        for (point in 1:points_n) {
+            for(cis in 1:quantiles_n) {
+                ## Setting X
[…]

### 2018-03-22 Functions/tree.resolution.R

@@ -1,0 +1,97 @@
+#' @title Tree resolution
+#'
+#' @description Get tree resolution
+#'
+#' @param 
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+# path = "../Data/Trees_out/Consensus_trees/Bayesian/Done/25t_100c/"
+# suffix = "con.tre"
+
+tree.resolution <- function(path, suffix) {
+
+    ## Read the tree files
+    trees <- list.files(path = path, pattern = suffix)
+
+    ## Get the tree format
+    type <- scan(paste0(path,trees[1]), what = "character", quiet = TRUE, n = 1)
+    if(type == "#NEXUS") {
+        read.tree.fun <- ape::read.nexus
+    } else {
+        read.tree.fun <- ape::read.tree
+    }
+
+    ## Categories
+    categories <- unique(unlist(lapply(lapply(strsplit(trees, split = "_"), function(X) return(X[4])), function(X) return(strsplit(X, split = paste0(".", suffix))[[1]]))))
+
+    ## Read the trees per categories
+    read.one.category <- function(category, path, suffix, read.tree.fun) {
+        return(unlist(lapply(as.list(list.files(path = path, pattern = paste0(category, ".", suffix))), function(X, read.tree.fun, path) return(ape::Nnode(read.tree.fun(paste0(path, X)))), read.tree.fun, path)))
+    }
+
+    ## Get the resolution per categories
+    resolutions <- lapply(as.list(categories), read.one.category, path, suffix, read.tree.fun)
+
+    ## Reorder the categories (norm, max, min, rand)
+    resolutions <- resolutions[c(3, 1, 2, 4)]
+    names(resolutions) <- categories[c(3, 1, 2, 4)]
+
+    return(resolutions)
+}
+
+# col <- c("darkgrey", "red", "green3", "blue", "grey", "orange", "lightgreen", "lightblue")
+
+# plot.tree.resolution("../Data/Trees_out/Consensus_trees/", ntaxa = 25, col = c("darkgrey", "red", "green3", "blue", "grey", "orange", "lightgreen", "lightblue"))
+
+plot.tree.resolution <- function(path, ntaxa, col, type = "line") {
+
+    ## placeholder lists
[…]

### 2018-03-23 Writing/draft.tex
changed CD equation
Added the fact that our simulation protocol is not biological

--- 2018-03-21 Writing/draft.tex
@@
 \begin{equation}
-    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
+    CD_{(x,y)} = 1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}\right)
 \end{equation}
 
-\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each characters states for the each $i^{th}$ taxon.
+%MB: Why divide by 1/2 when you can just multiply by 2?
+%TG: fair point. I've fixed it in the demonstration as well (supplementaries)
+
+\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
 $CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
@@
 This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
-%MB: I don't get this part. What do you mean by standardized? And why are you doing this? Are you just editing the symbols?
-Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
+Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).
 
@@
 
-% Following these results, we encourage researchers to code characters that appear to be the most un-correlated \textit{a priori}.
-% %MB: Do we? I'm not sure that I would. In practice, this would probably force morphologists to remove most of their characters.
-% %MB: Making these kinds of recommendations leads to the risk that all you do is introduce biases into the dataset in another way.
-% %TG: Fair enough, I'm happy to drop this whole paragraph.
 % However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
-% %MB: How small? Be specific.
 % Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e.
 as an equivalent of randomising the character differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).
+%MB: Agreed that that is noteworthy.
+I suspect I know why this happens (and you might have explained it earlier).
+It might actually be a statistical problem with our approach.
+When character similaritiy is maximized, how well resolved are the trees?
+%TG: I've moved that to a "potential application" paragraph in the discussion (please change the title of it, I had zero inspiration and find it ugly as feck!)
 
@@
 
-% This being said, it is interesting to analyse this difference in tree inference method in the light of the effect of high/low character differences.
-%MB: This is serious hedging ("we're not going to... but we're going to"). Could probably skip the above line.
-%TG: fair point
-% The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
-% When many characters are homoplasic (high character difference) parsimony will deal with this information by generating longer trees (i.e.
-) (which will be more likely to be discarded in the tree search).
-%MB: This doesn't make sense. Longer than one? How is generating longer trees coping? What do you mean by "coping"? 
-% Conversely, Bayesian analysis can (1) cope with homoplasy more easily by increasing branch length to accommodate more changes and (2) will also incorporate sub-optimal trees in the final results.
-% These two points could explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.
 %MB: I'm not sure I'm convinced by this paragraph. Perhaps we should have a chat about it.
@@
 
-% Effectively, this study focuses on the effect of coding induced correlation since our scenarios to maximise/minimise character difference are equivalent to actively choosing characters that are the most different/similar \textit{prior} to any phylogenetic analysis.
-TG: moved to intro.
-%MB: From a structural point of view, this line is out of place even.
-It makes a general point, but deep in the subsection.
-Such statements are better made earlier so a reader knows what the point of the section is.
 It is important to note that the two other sources of character correlation could also be present in our simulations although they were not explicitly modelled:
@@
 \subsection{Limitations}
+%MB: I think what this section _really_ needs is an assessment of the method used to simulate correlations.
+It was basically accomplished by duplicating characters.
+It wasn't accomplished by evolving some characters with a particular covariance model (which probably would've been better).
+We kind of faked-up that model.
+%TG: agree.
+
+TODO
+
 First, simulating evolutionary history is complex.
@@
 
+%TG: insisting on the shittiness of generatin correlation
 Second, measuring and modifying character correlation is difficult. 
-Increasing overall characters difference so that each pairwise differences is equal to 1 becomes more and more impossible with a large number of characters.
-In our algorithm, we maximised characters difference by replacing a characters $A$ highly correlated to another other character $B$ by a randomly chosen character $X$.
-This reduces the correlation between $A$ and $B$ but increases the correlation between $A$ and $X$!
-Additionally, when minimising $n$ character differences, the matrix will end containing only one character duplicated $n$ times (which will generate a unresolved tree by definition).
-Therefore we used an algorithm that only ``maximises'' or ``minimises'' character differences on average (see Fig. \ref{Fig:modif_matrix}).
-Although this was sufficient to generate detectable differences in topologies, it is not excluded that some randomly ``maximised'' or ``minimised'' matrices were actually not different than the ``randomised'' ones.
+In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation.
+In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
+This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
+%MB: We also don't really explore character congruence, that correlations can be ranked and spread over multiple internodes.
+%MB: We didn't look at ranking and we didn't look at competing correlations. There's a lot more that could be done here so, we might talk about that.
+%TG: not sure what you mean by the ranking stuff but it could definitely fit here.
+However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
+Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e.
+some sets of characters can be highly correlated but effectively describe independent information).
+Therefore, we made the choice to simplify our simulations by generating character correlation as a stochastic process rather than a biological one.
 
@@
 However, we want to point out again that here we're no comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
+
+\subsection{Potential applications}
+%TG: explaining that our "scenarios" are biased bootstraps
+Effectively, our simulation protocol bootstraps our data ``with bias''.
+In the ``randomised'' scenarios the data is simply randomly bootstrapped simply we randomly remove and resample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
+However, in the ``minimised'' and ``maximised'' scenario, the bootstrapping we remove the characters with the lowest/highest overall character difference.
+For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the left characters.
+
+%TG: explaining that we could use it for "clever" bootstrapping in big matrices
+It is noteworthy to point that in rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
+Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling (i.e.
+the equivalent of randomising the character differences in our analysis) will be sufficient to obtain a robust topology (\textit{cf.} actively collecting different characters).
+In matrices with more taxa, however, the ``maximised'' scenario resulted in better topological recovery than any other scenarios.
+Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.
 
@@
 %MB: But at what point? How much correlation is needed to affect it? TG: added values.
-%TG: do we want to suggest something here to close the conclusion?
-I agree we don't wanna be presecriptive Bristolians but it'll be nice to finsh with something like: "we're not saying it's a problem per se but maybe check your data?"
+
+%MB: Part of it is: we expect it to be a problem, but _when_ do we expect it to be a problem.
+There's the issue of "check your data", but it only raises the question: how?
+I think this is where we suggest avenues for future research (either here or move it up to the discussion).
 

### 2018-03-23 Writing/draft.tex
Slight title update
Chg first line abstract

@@
 
-\noindent{\Large \bf Influence of different modes of morphological character correlation on tree inference}
+\noindent{\Large \bf Influence of different modes of morphological character correlation on phylogenetic tree inference}
 \bigskip
@@
 \begin{abstract}
-Morphological characters for building phylogenies are assumed to be independent.
+Phylogenetic analysis algorithms require the assumption of character independence---a condition generally acknowledged to be violated by morphological data.
 However, this assumption is nearly always wrong simply because of the nature of morphological integration of organisms.

### 2018-03-26 Analysis/04-EffectCorrelation_lm.Rmd
lm analysis

---
title: "Effect of character correlation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Effect of character correlation

## Loading the functions and the data

Getting the functions and the data.

```{r, eval = TRUE, message = FALSE}
library(dispRity)
library(lme4)
library(lmerTest)
source("functions.R") ; load.functions(test = FALSE)
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")

## Getting the whole data
whole_data <- list("t25" = t25_list, "t75" = t75_list, "t150" = t150_list)
```

```{r, echo = FALSE}
library(knitr)
library(xtable)
```

## Linear models

### Running a single linear model

Here we measure the eventual correlation between a low/high character correlation and a low/high RF or Triplets score.

```{r}
## Loading the matrices CD
matrices_cd <- get.matrix.CD(path = "../Data/Simulations/Matrices", "25t_100c", verbose = TRUE, length.out = 35)

## Running the model
model_out <- run.lm(matrices_cd, whole_data, chain = "25t_100c", method = "Bayesian", metric = "RF")

## Plotting the output

## Colour differences
col <- c("blue", "orange", "green")
rounding <- 3

```

### 2018-03-26 Functions/correlation.lm.R
added pooled analysis

@@ -82,13 +82,42 @@ get.all.matrix.CD <- function(path, verbose = FALSE, length.out, cent.tend = NUL
     all_chains <- lapply(as.list(taxa), function(taxa, characters) return(as.list(paste(taxa, characters, sep = "_"))), characters)
 
     ## Apply get.matrix.CD on all_chains
-    all_char_diff <- lapply(all_chains, lapply, get.matrix.CD, path = path, verbose = verbose, length.out = length.out, cent.tend = cent.tend)
-    names(all_char_diff) <- taxa
-    for(tax in 1:length(all_char_diff)) {
+    all_char_diff <- list()
+    for(tax in 1:length(taxa)) {
+        all_char_diff[[tax]] <- list()
+        for(char in 1:length(characters)) {
+            all_char_diff[[tax]][[char]] <- get.matrix.CD(path = path, chain = all_chains[[tax]][[char]], verbose = verbose, length.out = length.out, cent.tend = cent.tend)
+        }
         names(all_char_diff[[tax]]) <- characters
     }
+    names(all_char_diff) <- taxa
 
     return(all_char_diff)
+}
+
+
+#' @title Pool results matrix CD
+#'
+#' @description Pools the results of the difference CD matrices
+#'
+#' @param list the list of matrices
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+pool.matrix.cd <- function(list) {
+
+    ## The results are a list from get.matrix.CDs
+    norm <- unlist(lapply(list, lapply, function(X) return(X$norm)))
+    mini <- unlist(lapply(list, lapply, function(X) return(X$mini)))
+    maxi <- unlist(lapply(list, lapply, function(X) return(X$maxi)))
+    rand <- unlist(lapply(list, lapply, function(X) return(X$rand)))
+
+    return(list("norm" = norm, "mini" = mini, "maxi" = maxi, "rand" = rand))
 }
 
 #' @title Runs one linear model
@@ -97,7 +126,7 @@ get.all.matrix.CD <- function(path, verbose = FALSE, length.out, cent.tend = NUL
 #'
 #' @param matrix_cd the list of matrices with CD values
 #' @param whole_data the list of all the data
-#' @param chain the chain name (corresponding to the one used to obtain the matrix_cd)
+#' @param chain the chain name (corresponding to the one used to obtain the matrix_cd). Use "pool" to pool the whole data.
[…]

### 2018-03-26 Functions/correlation.results.R

@@ -60,25 +60,36 @@ one.correlation.plot <- function(matrices_cd, whole_data, chain, metric, method,
         }
     }
 
+    ## Length of matrices_cd
+    length.out <- length(matrices_cd[[1]])
+
     ## Get the chain values
-    chain_values <- strsplit(chain, split = "_")[[1]]
-    taxa_number <- as.numeric(strsplit(chain_values[1], split = "t")[[1]])
-    if(!taxa_number %in% c(25, 75, 150)) {
-        stop(paste0("In ", chain, " taxa number (#t) must be 25, 75 or 150."))
-    }
-    if(taxa_number == 25) {
-        taxa_value <- 1
-    } else {
-        taxa_value <- ifelse(taxa_number == 75, 2, 3)
-    }
-    character_number <- as.numeric(strsplit(chain_values[2], split = "c")[[1]])
-    if(!character_number %in% c(100, 350, 1000)) {
-        stop(paste0("In ", chain, " character number (#c) must be 100, 350 or 1000."))
-    }
-    if(character_number == 100) {
-        character_value <- 1
-    } else {
-        character_value <- ifelse(taxa_number == 350, 2, 3)
+    if(chain != "pool") {
+        chain_values <- strsplit(chain, split = "_")[[1]]
+        taxa_number <- as.numeric(strsplit(chain_values[1], split = "t")[[1]])
+        if(!taxa_number %in% c(25, 75, 150)) {
+            stop(paste0("In ", chain, " taxa number (#t) must be 25, 75 or 150."))
+        }
+        if(taxa_number == 25) {
+            taxa_value <- 1
+        } else {
+            taxa_value <- ifelse(taxa_number == 75, 2, 3)
+        }
+        character_number <- as.numeric(strsplit(chain_values[2], split = "c")[[1]])
+        if(!character_number %in% c(100, 350, 1000)) {
+            stop(paste0("In ", chain, " character number (#c) must be 100, 350 or 1000."))
+        }
+        if(character_number == 100) {
+            character_value <- 1
+        } else {
+            character_value <- ifelse(taxa_number == 350, 2, 3)
+        }
+
+        ## Double check length.out
+        length_whole_data <- length(whole_data[[taxa_value]][[character_value]][[method_value]]$norm$mini[, metric_value])
+        if(length.out > length_whole_data) {
+            length.out <- length_whole_data
+        }
     }
 
[…]

### 2018-03-26 Writing/draft.tex
changed number of simulations from 40 to 35

@@
 \end{enumerate}
-Each step was replicated 40 times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
+Each step was replicated 35 times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
 
@@
 The matrices were generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 40 times resulting in 360 ``normal'' morphological matrices.
+To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 35 times resulting in 315 ``normal'' morphological matrices.
 
@@
 Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
-This step resulted in a total of 1440 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
+This step resulted in a total of 1260 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
 The algorithms for the three modifications are available on GitHub (\url{https://github.com/TGuillerme/CharactersCorrelation})

### 2018-03-26 Writing/draft.tex
Intro and abst edits

--- 2018-03-23 Writing/draft.tex
@@
 Phylogenetic analysis algorithms require the assumption of character independence---a condition generally acknowledged to be violated by morphological data.
-However, this assumption is nearly always wrong simply because of the nature of morphological integration of organisms.
-Correlation between characters can originate from intra-organismal features, shared phylogenetic history or errors when coding or defining morphological characters.
+%However, this assumption is nearly always wrong simply because of the nature of morphological integration of organisms.
+Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
 Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with good practices, phylogenetic software do not distinguish between any of them.
-
 In this study, we propose a new metric of raw character difference as a proxy for character correlation.
@@
 This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
-Morphological character data are often considered inferior to molecular sequence data (due to their inherent biases), but are often the only source of data---especially for fossils.
+Morphological character data are often considered inferior to molecular sequence data, but are often the only source of phylogenetic data for extinct species. 
 While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
@@
 \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the sheer nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
+Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the very nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
 % TG: add Brazeau et al. 201X somewhere?
@@
 	Characters of the occlusal surface of two opposing molars will be expected to directly covary.
-    These links can be the targets of comparative developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary} or functional investigations.
+	% We should probably cite some "Gunterology" here.
+	Pleiotropy also results in covariation between different aspects of phenotype. 
+	From a phylogenetic perspective, it can be especially pernicious because the relationship between the traits in question may have no obvious link from a morphological or functional comparison alone. 
+    Intra-organismal links can be the targets of comparative developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary} or functional investigations.
 
-    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving due to selection: the character's correlation is thus due to their shared evolutionary history.
-	Unlike the case of intra-organismal dependence, there is no intrinsic constraint that causes these traits to covary.
+    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving due to selection, likely related to functional links between two traits that help serve an overall lifestyle trait.
+	Unlike the case of intra-organismal dependence, there need not be an intrinsic constraint that causes these traits to covary.
+    For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
+	This is thought to correspond to adaptations for fossoriality or acquatic lifestyles.
+	Such covariances are generally studied in the context of a given phylogeny, often one derived from molecular data with morphological traits mapped on.
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
-    For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
-	Such covariances are generally studied in the context of a given phylogeny.
+	However, these methods do not give us a means to objectively control correlations that might adversely affect phylogenetic inference.
 
     \item \textbf{Coding dependence:} this is the results of researcher methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
-    One such source of correlation could be due to methodological errors: for example by defining a character describing a particular feature on a left and right molar.
+	Coding dependence manifests itself in several ways, particularly in coding redundant information.
+	For instance, coding for the same absence in different characters creates state transformations associated with the loss or gain of a particular character.
+	This occurs when a number of multistate characters include two variable feature states (e.g. large, small; red, blue etc.) in conjunction with absense.
     It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
@@
     Finally this can also be due to a bias in the amount of characters available.
-    For example, in a skull, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity).
+    For example, in skulls, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity).
 \end{itemize}
@@
 Logical dependence, however, is easily distinguished prior to phylogenetic inference, while the two other ones (intra-organismal and evolutionary) are much harder.
+However, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies.
+% Of course, at this point we could probably self-cite and cite De Laet.
 Intra-organismal dependence requires more detailed, often extremely time-consuming studies (and possibly beyond the limits of available technology).

### 2018-03-26 Analysis/01-EffectCorrelationStats.Rmd

--- 2017-09-26 Analysis/EffectCorrelationStats.Rmd
@@
 source("functions.R") ; load.functions(test = FALSE)
-dyn.load("../Functions/char.diff.so")
 load("../Data/NTS/matrix.nts.Rda")

### 2018-03-26 Analysis/04-EffectCorrelation_lm.Rmd

@@
 
-## Linear models
-
-### Running a single linear model
+# One single model example
 
@@
 
-```{r}
+```{r, fig.width=6, fig.height=6, message = FALSE}
 ## Loading the matrices CD
-matrices_cd <- get.matrix.CD(path = "../Data/Simulations/Matrices", "25t_100c", verbose = TRUE, length.out = 35)
+matrices_cd <- get.matrix.CD(path = "../Data/Simulations/Matrices", "25t_100c", verbose = TRUE, length.out = 35, cent.tend = mean)
 
 ## Running the model
-model_out <- run.lm(matrices_cd, whole_data, chain = "25t_100c", method = "Bayesian", metric = "RF")
+model_out <- run.lm(matrices_cd, whole_data, chain = "25t_100c", method = "Bayesian", metric = "RF", cent.tend = NULL)
 
 ## Plotting the output
+par(bty = "n")
+one.correlation.plot(matrices_cd, whole_data, chain = "25t_100c", method = "Bayesian", metric = "RF", model_out, cent.tend = NULL, legend = TRUE)
+```
 
-## Colour differences
-col <- c("blue", "orange", "green")
-rounding <- 3
+# Running all the models
+
+## Getting all matrices CD
+
+```{r, message = FALSE}
+## Loading and calculating all the matrices CDs
+all_matrices_cd <- get.all.matrix.CD(path = "../Data/Simulations/Matrices", verbose = TRUE, length.out = 35, cent.tend = mean)
+save(all_matrices_cd, file = "../Data/matrices_cd.Rda")
+# load("../Data/matrices_cd.Rda")
+```
+
+# Pooled models
+
+```{r}
+## Pool all the results into one list
+pooled_results <- pool.matrix.cd(all_matrices_cd)
+
+## Running the model
+model_out_bay_RF <- run.lm(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", cent.tend = NULL)
+model_out_bay_Tr <- run.lm(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "Triplets", cent.tend = NULL)
+model_out_par_RF <- run.lm(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "RF", cent.tend = NULL)
+model_out_par_Tr <- run.lm(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", cent.tend = NULL)
+
+## Plotting the output
+par(bty = "n", mfrow = c(2,2))
+one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", model_out_bay_RF, cent.tend = NULL, legend = TRUE, main = "Bayesian", xlab = "RF")
+one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", model_out_bay_Tr, cent.tend = NULL, legend = FALSE, main = "Bayesian", xlab = "RF")
+one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", model_out_par_RF, cent.tend = NULL, legend = FALSE, main = "Parsimony", xlab = "Triplets")
+one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", model_out_par_Tr, cent.tend = NULL, legend = FALSE, main = "Parsimony", xlab = "Triplets")
+```
+
+# Individual models
+
+## Bayesian
+
+### Robinson-Foulds
+
+```{r}
+## Running all the models
+all_models_out <- run.all.lm(all_matrices_cd, whole_data, method = "Bayesian", metric = "RF")
+
+## Getting the table
+kable(correlation.table(all_models_out), digits = 3, caption = "Linear model results for the Bayesian trees and RF metric.")
+```
+
+```{r, fig.width=12, fig.height=12}
+## Plotting the results
+all.correlation.plot(all_matrices_cd, whole_data, metric = "RF", method = "Bayesian", models = all_models_out, col = c("blue", "orange", "green"), rounding = 3, cent.tend = NULL)
+```
+
+### Triplets
+
+```{r}
+## Running all the models
+all_models_out <- run.all.lm(all_matrices_cd, whole_data, method = "Bayesian", metric = "Triplets")
+
+## Getting the table
+kable(correlation.table(all_models_out), digits = 3, caption = "Linear model results for the Bayesian trees and Triples metric.")
 
 ```
+
+```{r, fig.width=12, fig.height=12}
+## Plotting the results
+all.correlation.plot(all_matrices_cd, whole_data, metric = "Triplets", method = "Bayesian", models = all_models_out, col = c("blue", "orange", "green"), rounding = 3, cent.tend = NULL)
+```
+
+## Parsimony
+
+### Robinson-Foulds
+
+```{r}
+## Running all the models
+all_models_out <- run.all.lm(all_matrices_cd, whole_data, method = "Parsimony", metric = "RF")
+
+## Getting the table
+kable(correlation.table(all_models_out), digits = 3, caption = "Linear model results for the Parsimony trees and RF metric.")
+
+```
+
+```{r, fig.width=12, fig.height=12}
+## Plotting the results
+all.correlation.plot(all_matrices_cd, whole_data, metric = "RF", method = "Parsimony", models = all_models_out, col = c("blue", "orange", "green"), rounding = 3, cent.tend = NULL)
+```
+
+### Triplets
+
+```{r}
+## Running all the models
+all_models_out <- run.all.lm(all_matrices_cd, whole_data, method = "Parsimony", metric = "Triplets")
+
+## Getting the table
+kable(correlation.table(all_models_out), digits = 3, caption = "Linear model results for the Parsimony trees and Triples metric.")
+
+```
+
+```{r, fig.width=12, fig.height=12}
+## Plotting the results
+all.correlation.plot(all_matrices_cd, whole_data, metric = "Triplets", method = "Parsimony", models = all_models_out, col = c("blue", "orange", "green"), rounding = 3, cent.tend = NULL)
+```

### 2018-03-26 Functions/correlation.lm.R
typos

@@ -109,13 +109,13 @@ get.all.matrix.CD <- function(path, verbose = FALSE, length.out, cent.tend = NUL
 #' @author Thomas Guillerme
 #' @export
 
-pool.matrix.cd <- function(list) {
+pool.matrix.cd <- function(list, length.out) {
 
     ## The results are a list from get.matrix.CDs
-    norm <- unlist(lapply(list, lapply, function(X) return(X$norm)))
-    mini <- unlist(lapply(list, lapply, function(X) return(X$mini)))
-    maxi <- unlist(lapply(list, lapply, function(X) return(X$maxi)))
-    rand <- unlist(lapply(list, lapply, function(X) return(X$rand)))
+    norm <- unlist(lapply(list, lapply, function(X) return(X$norm[1:length.out])))
+    mini <- unlist(lapply(list, lapply, function(X) return(X$mini[1:length.out])))
+    maxi <- unlist(lapply(list, lapply, function(X) return(X$maxi[1:length.out])))
+    rand <- unlist(lapply(list, lapply, function(X) return(X$rand[1:length.out])))
 
     return(list("norm" = norm, "mini" = mini, "maxi" = maxi, "rand" = rand))
 }
@@ -164,6 +164,9 @@ run.lm <- function(matrices_cd, whole_data, chain, metric, method, cent.tend = N
 
     ## Length of matrices_cd
     length.out <- length(matrices_cd[[1]])
+    # if(chain == "pool") {
+    #     length.out <- length(matrices_cd[[1]])
+    # }
 
     ## Get the chain values
     if(chain != "pool") {
@@ -203,11 +206,13 @@ run.lm <- function(matrices_cd, whole_data, chain, metric, method, cent.tend = N
                         )
                     )
     } else {
+        warning("DEBUG: Change length.out value in run.lm and one.correlation.plot")
+        length.out <- 28
         data_metric <- data.frame(
             "Metric" = c(
-                unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$mini[, metric_value][1:length.out/9]), method_value, metric_value, length.out)),
-                unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$maxi[, metric_value][1:length.out/9]), method_value, metric_value, length.out)),
[…]

### 2018-03-26 Functions/correlation.results.R

@@ -28,9 +28,13 @@ one.correlation.plot <- function(matrices_cd, whole_data, chain, metric, method,
     ## Get the metric
     if(metric == "RF") {
         metric_value <- 2
+        ylim <- c(0,1)
+        y_legend <- c(0.2, 0.15, 0.1, 0.05)
     } else {
         if(metric == "Triplets") {
             metric_value <- 4
+            ylim <- c(-0.5,1)
+            y_legend <- c(-0.3, -0.35, -0.4, -0.45)
         } else {
             stop("metric must be \"RF\" or \"Triplets\".")
         }
@@ -39,11 +43,9 @@ one.correlation.plot <- function(matrices_cd, whole_data, chain, metric, method,
     ## Get the method
     if(method == "Bayesian") {
         method_value <- 1
-        ylim <- c(0,1)
     } else {
         if(method == "Parsimony") {
             method_value <- 2
-            ylim <- c(-0.5,1)
         } else {
             stop("metric must be \"Bayesian\" or \"Parsimony\".")
         }
@@ -117,10 +119,13 @@ one.correlation.plot <- function(matrices_cd, whole_data, chain, metric, method,
         y_maxi <- whole_data[[taxa_value]][[character_value]][[method_value]]$norm$maxi[, metric_value][1:length.out]
         y_rand <- whole_data[[taxa_value]][[character_value]][[method_value]]$norm$rand[, metric_value][1:length.out]
     } else {
-        y_mini <- unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$mini[, metric_value][1:length.out/9]), method_value, metric_value, length.out))
-        y_maxi <- unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$maxi[, metric_value][1:length.out/9]), method_value, metric_value, length.out))
-        y_rand <- unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$rand[, metric_value][1:length.out/9]), method_value, metric_value, length.out))
+        warning("DEBUG: Change length.out value in run.lm and one.correlation.plot")
+        length.out <- 28
+        y_mini <- unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$mini[1:length.out, metric_value]), method_value, metric_value, length.out))
[…]

### 2018-03-26 Analysis/04-EffectCorrelation_lm.Rmd
Added pooled analysis (dodgy)

@@
 ## Loading and calculating all the matrices CDs
-all_matrices_cd <- get.all.matrix.CD(path = "../Data/Simulations/Matrices", verbose = TRUE, length.out = 35, cent.tend = mean)
-save(all_matrices_cd, file = "../Data/matrices_cd.Rda")
-# load("../Data/matrices_cd.Rda")
+# all_matrices_cd <- get.all.matrix.CD(path = "../Data/Simulations/Matrices", verbose = TRUE, length.out = 35, cent.tend = mean)
+# save(all_matrices_cd, file = "../Data/matrices_cd.Rda")
+load("../Data/matrices_cd.Rda")
 ```
@@
 
-```{r}
+```{r, fig.width=12, fig.height=12, warnings = FALSE}
 ## Pool all the results into one list
-pooled_results <- pool.matrix.cd(all_matrices_cd)
+pooled_results <- pool.matrix.cd(all_matrices_cd, length.out = 28)
 
-## Running the model
+# ## Running the model
 model_out_bay_RF <- run.lm(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", cent.tend = NULL)

### 2018-03-27 Analysis/01-EffectCorrelationStats.Rmd
Automatic table build

@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Tab_pooledscenarios")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Tab_pooledscenarios", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
 xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_scen_test, label = "Tab_pooledscenarios_test")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(test_table, caption = cap_pool_scen_test, label = "Tab_pooledscenarios_test", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(characters, caption = captions_pooled_characters, label = "Tab_pooledcharacters")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(characters, caption = captions_pooled_characters, label = "Tab_pooledcharacters", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
 xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3, label = "Tab_signif")
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_char_test, label = "Tab_pooledscharacters_test")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(test_table, caption = cap_pool_char_test, label = "Tab_pooledscharacters_test", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(taxa, caption = captions_pooled_taxa, label = "Tab_pooledtaxa")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(taxa, caption = captions_pooled_taxa, label = "Tab_pooledtaxa", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_tax_test, label = "Tab_pooledstaxa_test")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(test_table, caption = cap_pool_tax_test, label = "Tab_pooledstaxa_test", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(methods, caption = captions_pooled_methods, label = "Tab_pooledmethod")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(methods, caption = captions_pooled_methods, label = "Tab_pooledmethod", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_met_test, label = "Tab_pooledsmethods_test")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(test_table, caption = cap_pool_met_test, label = "Tab_pooledsmethods_test", path = "../Writing/Tables/")
 ```
@@
 ## Tables
-make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1])
-make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2])
+make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1], path = "../Writing/Tables/")
+make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2], path = "../Writing/Tables/")
 ```
@@
 ## Tables
-make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1])
-make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2])
+make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1], path = "../Writing/Tables/")
+make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2], path = "../Writing/Tables/")
 ```
@@
 ## Tables
-make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1])
-make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2])
+make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1], path = "../Writing/Tables/")
+make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2], path = "../Writing/Tables/")
 ```
-
-# Pooled results
-
-## Effect of character correlation per scenario
-
-```{r, warning = FALSE}
-## Pooling the data per scenarios
-scenario_RF_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
-                               metric = "RF", best = "norm")
-scenario_Tr_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
-                               metric = "Tr", best = "norm")
-scenario_RF_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
-                               metric = "RF", best = "rand")
-scenario_Tr_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
-                               metric = "Tr", best = "rand")
-
-## Preparing the table labels
-comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
-metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
-
-## Combining the data into a table
-scenarios <- pool.table(list(scenario_RF_norm, scenario_Tr_norm, scenario_RF_null, scenario_Tr_null),
-                    comp.label = comp.labels, metric.label = metric.labels, label.param = "scenario")
-```
-
-```{r, echo = FALSE}
-library(knitr)
-## Setting the table captions
-captions_pooled_scenarios <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled scenarios.")
-
-## Printing the tables
-kable(scenarios, row.names = FALSE, caption = captions_pooled_scenarios)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-library(xtable)
-make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Tab_pooledscenarios")
-```
-
-### Testing the effect of the scenario on the NTS
-
-```{r}
-## Applying an lm to all the data
-results <- test.effect.aov(whole_data, formula = NTS ~ scenario, fun = lm)
-```
-
-Checking the model assumptions
-
-```{r, echo = FALSE}
-kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
-```
-
-Applying a Bhattacharrya Coefficient and Wilcoxon test
-
-```{r, warning = FALSE}
-## Summarise into a table
-scenario_list <- list(scenario_RF_norm, scenario_Tr_norm, scenario_RF_null, scenario_Tr_null)
-comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
-metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
-test_table <- pair.test.table(scenario_list, comp.label = comp.labels, metric.label = metric.labels)
-```
-
-```{r, echo = FALSE}
-## Caption
-cap_pool_scen_test <- paste("Difference between the pooled scenarios.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
-
-## Printing the tables
-kable(test_table, row.names = FALSE, caption = cap_pool_scen_test)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_scen_test, label = "Tab_pooledscenarios_test")
-```
-
-## Effect of character correlation per number of character
-
-```{r, warning = FALSE}
-## Pooling the data per character
-character_RF_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
-                               metric = "RF", best = "norm")
-character_Tr_norm <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
-                               metric = "Tr", best = "norm")
-character_RF_null <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
-                               metric = "RF", best = "rand")
-character_Tr_null <- multi.pool(whole_data, param = c("c100", "c350", "c1000"),
-                               metric = "Tr", best = "rand")
-
-## Preparing the table labels
-comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
-metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
-
-## Combining the data into a table
-characters <- pool.table(list(character_RF_norm, character_Tr_norm, character_RF_null, character_Tr_null),
-                    comp.label = comp.labels, metric.label = metric.labels, label.param = "character")
-```
-
-```{r, echo = FALSE}
-## Setting the table captions
-captions_pooled_characters <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of characters.")
-
-## Printing the tables
-kable(characters, row.names = FALSE, caption = captions_pooled_characters)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(characters, caption = captions_pooled_characters, label = "Tab_pooledcharacters")
-```
-
-### Testing the effect of the number of characters on the NTS
-
-```{r}
-## Applying an lm to all the data
-results <- test.effect.aov(whole_data, formula = NTS ~ character, fun = lm)
-```
-
-Checking the model assumptions
-
-```{r, echo = FALSE}
-kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
-```
-
-Applying a Bhattacharrya Coefficient and Wilcoxon test
-
-```{r, warning = FALSE}
-## Summarise into a table
-character_list <- list(character_RF_norm, character_Tr_norm, character_RF_null, character_Tr_null)
-comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
-metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
-test_table <- pair.test.table(character_list, comp.label = comp.labels, metric.label = metric.labels)
-```
-
-```{r, echo = FALSE}
-## Caption
-cap_pool_char_test <- paste("Difference between the pooled number of characters.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
-
-## Printing the tables
-kable(test_table, row.names = FALSE, caption = cap_pool_char_test)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_char_test, label = "Tab_pooledscharacters_test")
-```
-
-## Effect of character correlation per number of taxa
-
-```{r, warning = FALSE}
-## Pooling the data per taxa
-taxa_RF_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
-                               metric = "RF", best = "norm")
-taxa_Tr_norm <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
-                               metric = "Tr", best = "norm")
-taxa_RF_null <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
-                               metric = "RF", best = "rand")
-taxa_Tr_null <- multi.pool(whole_data, param = c("t25", "t75", "t150"),
-                               metric = "Tr", best = "rand")
-
-## Preparing the table labels
-comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
-metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
-
-## Combining the data into a table
-taxa <- pool.table(list(taxa_RF_norm, taxa_Tr_norm, taxa_RF_null, taxa_Tr_null),
-                    comp.label = comp.labels, metric.label = metric.labels, label.param = "taxa")
-```
-
-```{r, echo = FALSE}
-## Setting the table captions
-captions_pooled_taxa <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled number of taxa.")
-
-## Printing the tables
-kable(taxa, row.names = FALSE, caption = captions_pooled_taxa)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(taxas, caption = captions_pooled_taxas, label = "Tab_pooledtaxa")
-```
-
-### Testing the effect of the method on the NTS
-
-```{r}
-## Applying an lm to all the data
-results <- test.effect.aov(whole_data, formula = NTS ~ taxa, fun = lm)
-```
-
-Checking the model assumptions
-
-```{r, echo = FALSE}
-kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
-```
-
-Applying a Bhattacharrya Coefficient and Wilcoxon test
-
-```{r, warning = FALSE}
-## Summarise into a table
-taxa_list <- list(taxa_RF_norm, taxa_Tr_norm, taxa_RF_null, taxa_Tr_null)
-comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
-metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
-test_table <- pair.test.table(taxa_list, comp.label = comp.labels, metric.label = metric.labels)
-```
-
-```{r, echo = FALSE}
-## Caption
-cap_pool_tax_test <- paste("Difference between the pooled number of taxa.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
-
-## Printing the tables
-kable(test_table, row.names = FALSE, caption = cap_pool_tax_test)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_tax_test, label = "Tab_pooledstaxa_test")
-```
-
-## Effect of character correlation per methods
-
-```{r, warning = FALSE}
-## Pooling the data per taxa
-method_RF_norm <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
-                               metric = "RF", best = "norm")
-method_Tr_norm <- multi.pool(whole_data, param =c("bayesian", "parsimony"),
-                               metric = "Tr", best = "norm")
-method_RF_null <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
-                               metric = "RF", best = "rand")
-method_Tr_null <- multi.pool(whole_data, param = c("bayesian", "parsimony"),
-                               metric = "Tr", best = "rand")
-
-## Preparing the table labels
-comp.labels <- c("Best", rep("", 3), "Null", rep("", 3))
-metric.labels <- rep(c("RF", "", "Tr", ""), 2)
-
-## Combining the data into a table
-methods <- pool.table(list(method_RF_norm, method_Tr_norm, method_RF_null, method_Tr_null),
-                    comp.label = comp.labels, metric.label = metric.labels, label.param = "method")
-```
-
-```{r, echo = FALSE}
-## Setting the table captions
-captions_pooled_methods <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled methods.")
-
-## Printing the tables
-kable(methods, row.names = FALSE, caption = captions_pooled_methods)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(methods, caption = captions_pooled_methods, label = "Tab_pooledmethod")
-```
-
-### Testing the effect of the method on the NTS
-
-```{r}
-## Applying an lm to all the data
-results <- test.effect.aov(whole_data, formula = NTS ~ method, fun = lm)
-```
-
-Checking the model assumptions
-
-```{r, echo = FALSE}
-kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
-```
-
-Applying a Bhattacharrya Coefficient and Wilcoxon test
-
-```{r, warning = FALSE}
-## Summarise into a table
-methods_list <- list(method_RF_norm, method_Tr_norm, method_RF_null, method_Tr_null)
-comp.labels <- c("Best", rep("", 3), "Null", rep("", 3))
-metric.labels <- rep(c("RF", "", "Tr", ""), 2)
-test_table <- pair.test.table(methods_list, comp.label = comp.labels, metric.label = metric.labels)
-```
-
-```{r, echo = FALSE}
-## Caption
-cap_pool_met_test <- paste("Difference between the pooled methods.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
-
-## Printing the tables
-kable(test_table, row.names = FALSE, caption = cap_pool_met_test)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_met_test, label = "Tab_pooledsmethods_test")
-```
-
-# Individual results
-
-We can also analyse the effect of each parameters combination separately:
-
-## Summary statistics
-
-We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
-Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
-The p-values are adjusted for multiple tests using Bonferroni-Holm correction.
-
-```{r}
-## Summarising the data
-sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL)
-sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL)
-sum_RF_null <- apply.test(whole_data, "RF", "null", NULL)
-sum_Tr_null <- apply.test(whole_data, "Triples", "null", NULL)
-```
-
-```{r, echo = FALSE}
-## Setting the table captions
-captions_summary <- paste("Summary statistics of the normalised", rep(c("Robinson-Fould", "Triplets"), 2), "distance to the", c(rep("best", 2), rep("null", 2)), "tree.")
-
-## Printing the tables
-kable(sum_RF_best, row.names = FALSE, caption = captions_summary[1])
-kable(sum_Tr_best, row.names = FALSE, caption = captions_summary[2])
-kable(sum_RF_null, row.names = FALSE, caption = captions_summary[3])
-kable(sum_Tr_null, row.names = FALSE, caption = captions_summary[4])
-```
-
-<!-- Table for publication -->
-
-```{r, echo = FALSE, eval = FALSE}
-## Labels
-labels_summary <- paste0("Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
-
-## Tables
-make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1])
-make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2])
-make.xtable(sum_RF_null, caption = captions_summary[3], label = labels_summary[3])
-make.xtable(sum_Tr_null, caption = captions_summary[4], label = labels_summary[4])
-```
-
-## Nested anova
-
-Let's analysis the data using a nested anova.
-The models look at the effect of our scenarios ("maximised", "minimised", etc...) on the normalised tree score (NTS - using both the Robinson-Fould and Triplets scores) with a nested effect of the method (parsimony or bayesian), the number of taxa (25, 75, 150) and the number of characters (100, 150, 1000).
-
-```{r}
-## Applying an lm to all the data with error
-results <- test.effect.aov(whole_data,
-  formula = NTS ~ scenario + (1|method) + (1|taxa) + (1|character), fun = lmer)
-```
-
-Checking the model assumptions
-
-```{r, echo = FALSE}
-kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
-```
-
-<!-- Table for publication -->
-
-```{r, eval = FALSE, echo = FALSE}
-xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
-```
-
-All our models violate a least one of the two models assumptions (normality of the residuals and homoscedasticity of variance) and thus cannot be used to compare the scenarios.
-This is due to the fact that the metric is bounded to 1.
-
-## Distributions probability overlap (Bhattacharrya Coefficient)
-
-We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
-A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don t overlap (significant effect of one scenario on topology).
-Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).
-
-```{r}
-## Running the pairwise Bhattacharrya Coefficient analysis:
-bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff)
-bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff)
-bc_RF_null <- apply.test(whole_data, "RF", "null", bhatt.coeff)
-bc_Tr_null <- apply.test(whole_data, "Triples", "null", bhatt.coeff)
-
-## Tables
-kable(bc_RF_best, row.names = FALSE, digit = 3)
-kable(bc_RF_best, row.names = FALSE, digit = 3)
-kable(bc_RF_best, row.names = FALSE, digit = 3)
-kable(bc_RF_best, row.names = FALSE, digit = 3)
-```
-
-```{r, echo = FALSE}
-## Setting the table captions
-captions_bc <- paste0("Bhattacharrya Coefficient (",rep(c("Robinson-Fould distance", "Triplets distance"), 2) ," from the ",c(rep("best", 2), rep("null", 2)), " tree)" , " between the different scenarios.")
-
-## Printing the tables
-kable(bc_RF_best, row.names = FALSE, caption = captions_bc[1], digit = 3)
-kable(bc_Tr_best, row.names = FALSE, caption = captions_bc[2], digit = 3)
-kable(bc_RF_null, row.names = FALSE, caption = captions_bc[3], digit = 3)
-kable(bc_Tr_null, row.names = FALSE, caption = captions_bc[4], digit = 3)
-```
-
-<!-- Table for publication -->
-
-```{r, echo = FALSE, eval = FALSE}
-## Labels
-labels_bc <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
-
-## Tables
-make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1])
-make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2])
-make.xtable(bc_RF_null, caption = captions_bc[3], label = labels_bc[3])
-make.xtable(bc_Tr_null, caption = captions_bc[4], label = labels_bc[4])
-```
-
-## Non-parametric pairwise comparisons
-
-We can also apply a non-parametric pairwise comparison between scenarios for each parameters combinations.
-Note however that the resulting p-values are not always exact due to ties in the results (again, due to the metric being bounded to 1).
-The p-values are adjusted for multiple tests using Bonferroni-Holm correction.
-
-```{r, warning = FALSE}
-## Running the pairwise Wilcox tests:
-wx_RF_best <- apply.test(whole_data, "RF", "norm", wilcox.test)
-wx_Tr_best <- apply.test(whole_data, "Triples", "norm", wilcox.test)
-wx_RF_null <- apply.test(whole_data, "RF", "null", wilcox.test)
-wx_Tr_null <- apply.test(whole_data, "Triples", "null", wilcox.test)
-
-## Applying Bonferroni correction
-for(column in c(5,7,9)) {
-  wx_RF_best[,column] <- p.adjust(wx_RF_best[,column], method = "bonferroni")
-  wx_Tr_best[,column] <- p.adjust(wx_Tr_best[,column], method = "bonferroni")
-  wx_RF_null[,column] <- p.adjust(wx_RF_null[,column], method = "bonferroni")
-  wx_Tr_null[,column] <- p.adjust(wx_Tr_null[,column], method = "bonferroni")
-}
-```
-
-```{r, echo = FALSE}
-## Setting the table captions
-captions_wx <- paste0("Non-parametric Wilcoxon test (",rep(c("Robinson-Fould distance", "Triplets distance"), 2) ," from the ",c(rep("best", 2), rep("null", 2)), " tree)" , " between the different scenarios (p-values corrected using Bonferroni-Holm correction).")
-
-## Printing the tables
-kable(wx_RF_best, row.names = FALSE, caption = captions_wx[1], digit = 3)
-kable(wx_Tr_best, row.names = FALSE, caption = captions_wx[2], digit = 3)
-kable(wx_RF_null, row.names = FALSE, caption = captions_wx[3], digit = 3)
-kable(wx_Tr_null, row.names = FALSE, caption = captions_wx[4], digit = 3)
-```
-
-<!-- Table for publication -->
-
-```{r, echo = FALSE, eval = FALSE}
-## Labels
-labels_wx <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
-
-## Tables
-make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1])
-make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2])
-make.xtable(wx_RF_null, caption = captions_wx[3], label = labels_wx[3])
-make.xtable(wx_Tr_null, caption = captions_wx[4], label = labels_wx[4])
-```

### 2018-03-27 Analysis/02-EffectCorrelationFullResults.Rmd
Automatic table build

--- 2017-09-19 Analysis/EffectCorrelationFullResults.Rmd
@@
 
-```{r, eval = FALSE, echo = FALSE}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
 library(xtable)
-make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Tab_pooledscenarios")
+make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Full_Tab_pooledscenarios", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Full_Tab_signif", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_scen_test, label = "Tab_pooledscenarios_test")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(test_table, caption = cap_pool_scen_test, label = "Full_Tab_pooledscenarios_test", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(characters, caption = captions_pooled_characters, label = "Tab_pooledcharacters")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(characters, caption = captions_pooled_characters, label = "Full_Tab_pooledcharacters", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Full_Tab_signif")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_char_test, label = "Tab_pooledscharacters_test")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(test_table, caption = cap_pool_char_test, label = "Full_Tab_pooledscharacters_test", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(taxas, caption = captions_pooled_taxas, label = "Tab_pooledtaxa")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(taxa, caption = captions_pooled_taxa, label = "Full_Tab_pooledtaxa", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_tax_test, label = "Tab_pooledstaxa_test")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(test_table, caption = cap_pool_tax_test, label = "Full_Tab_pooledstaxa_test", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(methods, caption = captions_pooled_methods, label = "Tab_pooledmethod")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(methods, caption = captions_pooled_methods, label = "Full_Tab_pooledmethod", path = "../Writing/Tables/")
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
-make.xtable(test_table, caption = cap_pool_met_test, label = "Tab_pooledsmethods_test")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+make.xtable(test_table, caption = cap_pool_met_test, label = "Full_Tab_pooledsmethods_test", path = "../Writing/Tables/")
 ```
@@
 ## Labels
-labels_summary <- paste0("Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
+labels_summary <- paste0("Full_Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
 
 ## Tables
-make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1])
-make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2])
-make.xtable(sum_RF_null, caption = captions_summary[3], label = labels_summary[3])
-make.xtable(sum_Tr_null, caption = captions_summary[4], label = labels_summary[4])
+make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1], path = "../Writing/Tables/")
+make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2], path = "../Writing/Tables/")
+make.xtable(sum_RF_null, caption = captions_summary[3], label = labels_summary[3], path = "../Writing/Tables/")
+make.xtable(sum_Tr_null, caption = captions_summary[4], label = labels_summary[4], path = "../Writing/Tables/")
 ```
@@
 ## Labels
-labels_bc <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
+labels_bc <- paste0("Full_Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
 
 ## Tables
-make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1])
-make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2])
-make.xtable(bc_RF_null, caption = captions_bc[3], label = labels_bc[3])
-make.xtable(bc_Tr_null, caption = captions_bc[4], label = labels_bc[4])
+make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1], path = "../Writing/Tables/")
+make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2], path = "../Writing/Tables/")
+make.xtable(bc_RF_null, caption = captions_bc[3], label = labels_bc[3], path = "../Writing/Tables/")
+make.xtable(bc_Tr_null, caption = captions_bc[4], label = labels_bc[4], path = "../Writing/Tables/")
 ```
@@
 ## Labels
-labels_wx <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
+labels_wx <- paste0("Full_Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
 
 ## Tables
-make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1])
-make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2])
-make.xtable(wx_RF_null, caption = captions_wx[3], label = labels_wx[3])
-make.xtable(wx_Tr_null, caption = captions_wx[4], label = labels_wx[4])
+make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1], path = "../Writing/Tables/")
+make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2], path = "../Writing/Tables/")
+make.xtable(wx_RF_null, caption = captions_wx[3], label = labels_wx[3], path = "../Writing/Tables/")
+make.xtable(wx_Tr_null, caption = captions_wx[4], label = labels_wx[4], path = "../Writing/Tables/")
 ```

### 2018-03-27 Writing/draft.tex
Automatic table build

@@
 %Packages
-\usepackage{pdflscape}
 \usepackage{fixltx2e}
@@
 \usepackage{lineno}
-\usepackage{dcolumn}
-\usepackage{mathtools}
 \usepackage{pdflscape}
@@
 
-%MB: Why divide by 1/2 when you can just multiply by 2?
-%TG: fair point. I've fixed it in the demonstration as well (supplementaries)
-
 \noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{ll|r|rr}
-  \hline
-metric & test & bhatt.coeff & statistic & p.value \\ 
-  \hline
-RF & maxi:mini & 0.588 & 439466.500 & \textbf{0.000} \\ 
-   & maxi:rand & 0.894 & 347624.500 & \textbf{0.000} \\ 
-   & mini:rand & 0.846 & 115008.500 & \textbf{0.000} \\ 
-Tr & maxi:mini & 0.628 & 443049.500 & \textbf{0.000} \\ 
-   & maxi:rand & 0.921 & 348074.000 & \textbf{0.000} \\ 
-   & mini:rand & 0.825 & 113171.500 & \textbf{0.000} \\ 
-   \hline
-\end{tabular}
-\caption{Difference between the pooled scenarios.
+\input{Tables/Tab_pooledscenarios_test.tex}
+
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{ll|r|rr}
+%   \hline
+% metric & test & bhatt.coeff & statistic & p.value \\ 
+%   \hline
+% RF & maxi:mini & 0.588 & 439466.500 & \textbf{0.000} \\ 
+%    & maxi:rand & 0.894 & 347624.500 & \textbf{0.000} \\ 
+%    & mini:rand & 0.846 & 115008.500 & \textbf{0.000} \\ 
+% Tr & maxi:mini & 0.628 & 443049.500 & \textbf{0.000} \\ 
+%    & maxi:rand & 0.921 & 348074.000 & \textbf{0.000} \\ 
+%    & mini:rand & 0.825 & 113171.500 & \textbf{0.000} \\ 
+%    \hline
+% \end{tabular}
+% \caption{Difference between the pooled scenarios.
 bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction), significant values are in bold.}
-\label{Tab_pooledscenarios_test}
-\end{table}
+% \label{Tab_pooledscenarios_test}
+% \end{table}
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{ll|r|rr}
-  \hline
-metric & test & bhatt.coeff & statistic & p.value \\ 
-  \hline
-RF & c100:c350 & \textbf{0.989} & 228495.000 & 0.174 \\ 
-   & c100:c1000 & \textbf{0.983} & 218305.500 & \textbf{0.002} \\ 
-   & c350:c1000 & \textbf{0.988} & 232832.500 & 0.643 \\ 
-  Tr & c100:c350 & \textbf{0.965} & 203355.500 & \textbf{0.000} \\ 
-   & c100:c1000 & \textbf{0.959} & 189064.000 & \textbf{0.000} \\ 
-   & c350:c1000 & \textbf{0.983} & 226472.000 & 0.085 \\ 
-   \hline
-\end{tabular}
-\caption{Difference between the pooled number of characters. See Table \ref{Tab_pooledscenarios_test} for full caption.} 
-\label{Tab_pooledscharacters_test}
-\end{table}
+\input{Tables/Tab_pooledscharacters_test.tex}
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{ll|r|rr}
-  \hline
-metric & test & bhatt.coeff & statistic & p.value \\ 
-  \hline
-RF & t25:t75 & \textbf{0.987} & 267274.500 & \textbf{0.003} \\ 
-   & t25:t150 &\textbf{0.979} & 272359.500 & \textbf{0.000} \\ 
-   & t75:t150 & \textbf{0.988} & 247619.500 & 1.000 \\ 
-Tr & t25:t75 & \textbf{0.975} & 282017.500 & \textbf{0.000} \\ 
-   & t25:t150 & \textbf{0.970} & 279730.000 & \textbf{0.000} \\ 
-   & t75:t150 & \textbf{0.989} & 240414.500 & 1.000 \\ 
-   \hline
-\end{tabular}
-\caption{Difference between the pooled number of taxa. See Table \ref{Tab_pooledscenarios_test} for full caption.} 
-\label{Tab_pooledstaxa_test}
-\end{table}
+\input{Tables/Tab_pooledstaxa_test.tex}
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{llrrr}
-  \hline
-metric & test & bhatt.coeff & statistic & p.value \\ 
-  \hline
-RF & bayesian:parsimony & 0.893 & 676940.500 & \textbf{0.000} \\ 
-Tr & bayesian:parsimony & \textbf{0.982} & 547140.000 & 0.147 \\ 
-   \hline
-\end{tabular}
-\caption{Difference between the pooled methods. See Table \ref{Tab_pooledscenarios_test} for full caption.} 
-\label{Tab_pooledsmethods_test}
-\end{table}
+\input{Tables/Tab_pooledsmethods_test.tex}
 

### 2018-03-27 Functions/make.xtable.R

@@ -7,6 +7,7 @@
 #' @param caption the caption
 #' @param label the label (in LaTeX: \ref{label})
 #' @param longtable whether to use longtable
+#' @param path optional, a path for saving the table
 #' 
 #' @examples
 #'
@@ -15,13 +16,44 @@
 #' @author Thomas Guillerme
 #' @export
 
-make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE) {
-  
-  table <- xtable(table, digit = digit, caption = caption, label = label)
+make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE, path) {
 
-  if(longtable == TRUE) {
-    print(table, tabular.environment = 'longtable', floating = FALSE, include.rownames = FALSE)
-  } else {
-    print(table, include.rownames = FALSE)
-  }
+    ## Add significance values
+    if(all(names(table) == c("metric", "test", "bhatt.coeff", "statistic", "p.value"))){
+        for(row in 1:nrow(table)) {
+            if(as.numeric(table$bhatt.coeff[row]) >= 0.95 || as.numeric(table$bhatt.coeff[row]) <= 0.05) {
+                table$bhatt.coeff[row] <- paste0("BOLD",table$bhatt.coeff[row])
+            }
+            if(as.numeric(table$p.value[row]) <= 0.05) {
+                table$p.value[row] <- paste0("BOLD",table$p.value[row])
+            }
+        }
+    }
+
+    ## Bold cells function
+    bold.cells <- function(x) gsub('BOLD(.*)', paste0('\\\\textbf{\\1', '}'), x)
+
+    ## convert into xtable format
+    textable <- xtable(table, digit = digit, caption = caption, label = label)
+
+    ## Change attributes
+    if(all(names(table) == c("metric", "test", "bhatt.coeff", "statistic", "p.value"))){
+        ## Add a horizontal line
+        attr(textable, "align")[4] <- "r|"
+    }
+
+
+    if(!missing(path)) {
+        if(longtable == TRUE) {
+            cat(print(textable, tabular.environment = 'longtable', floating = FALSE, include.rownames = FALSE, sanitize.text.function = bold.cells), file = paste0(path, label, ".tex"))
+        } else {
+            cat(print(textable, include.rownames = FALSE, sanitize.text.function = bold.cells), file = paste0(path, label, ".tex"))
+        }
+    }
+
+    if(longtable == TRUE) {
[…]

### 2018-03-27 Functions/read.TreeCmp.R

@@ -5,6 +5,7 @@
 #' @param chain the general name of the chain (e.g. 25t_100c)
 #' @param path the path to where the files are stored
 #' @param type which type of tree (Bayesian or Parismony)
+#' @param length.in how many chains to read (e.g. 35)
 #'
 #' @return
 #' A list of metrics for the comparisons.
@@ -18,7 +19,7 @@
 
 #path_alt <- "/Users/TGuillerme/Projects/CharactersCorrelation/Data/Trees_out/Consensus_trees/Parsimony"
 
-read.TreeCmp <- function(chain, path = "/Users/TGuillerme/Projects/CharactersCorrelation/Data/Trees_out/Consensus_trees", type = "Bayesian") {
+read.TreeCmp <- function(chain, path = "/Users/TGuillerme/Projects/CharactersCorrelation/Data/Trees_out/Consensus_trees", type = "Bayesian", length.in) {
 
     path <- paste(path, type, sep = "/")
 
@@ -38,6 +39,10 @@ read.TreeCmp <- function(chain, path = "/Users/TGuillerme/Projects/CharactersCor
         stop('No files for rand has been found within the given chain name.')
     }
 
+    if(!missing(length.in)) {
+        files_norm <- files_norm[1:length.in]
+        files_rand <- files_rand[1:length.in]
+    }
 
     ## Load all the files
     matrices_norm <- list()

### 2018-03-27 Writing/Tables/Full_Tab_pooledsmethods_test.tex
Final results tables

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 13:20:38 2018
\begin{table}[ht]
\centering
\begin{tabular}{lllrrr}
  \hline
comp & metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
Best & RF & bayesian:parsimony & 0.891 & 579437.500 & 0.000 \\ 
   &  & bayesian:parsimony & 0.984 & 470621.500 & 0.168 \\ 
   & Tr & bayesian:parsimony & 0.847 & 657875.500 & 0.000 \\ 
   &  & bayesian:parsimony & 0.975 & 499754.000 & 0.000 \\ 
  Null & RF & bayesian:parsimony & 0.891 & 579437.500 & 0.000 \\ 
   &  & bayesian:parsimony & 0.984 & 470621.500 & 0.168 \\ 
   & Tr & bayesian:parsimony & 0.847 & 657875.500 & 0.000 \\ 
   &  & bayesian:parsimony & 0.975 & 499754.000 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Full_Tab_pooledsmethods_test}
\end{table}

### 2018-03-27 Writing/Tables/Tab_pooledmethod.tex
Final results tables

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:14 2018
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrrr}
  \hline
metric & method & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
RF & bayesian & 0.319 & 0.714 & 0.828 & 0.811 & 0.934 & 1.000 \\ 
   & parsimony & 0.076 & 0.496 & 0.679 & 0.684 & 0.911 & 1.000 \\ 
  Tr & bayesian & -0.473 & 0.354 & 0.738 & 0.618 & 0.947 & 1.000 \\ 
   & parsimony & -0.436 & 0.258 & 0.601 & 0.568 & 0.949 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Summary statistics of the normalised distances to the best or null tree for the pooled methods.} 
\label{Tab_pooledmethod}
\end{table}

### 2018-03-27 Writing/Tables/Tab_pooledscenarios_test.tex
Final results tables

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:13 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & maxi:mini & 0.573 & 356436.000 & \textbf{0} \\ 
   & maxi:rand & 0.873 & 287225.000 & \textbf{0} \\ 
   & mini:rand & 0.856 & 95841.500 & \textbf{0} \\ 
  Tr & maxi:mini & 0.614 & 358800.000 & \textbf{0} \\ 
   & maxi:rand & 0.908 & 288223.500 & \textbf{0} \\ 
   & mini:rand & 0.858 & 98507.500 & \textbf{0} \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled scenarios. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction)} 
\label{Tab_pooledscenarios_test}
\end{table}

### 2018-03-27 Writing/Tables/Tab_pooledscharacters_test.tex
Final results tables

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Mar 27 14:03:13 2018
\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
   \hline
\end{tabular}
\caption{Difference between the pooled number of characters. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)} 
\label{Tab_pooledscharacters_test}
\end{table}

### 2018-03-27 Analysis/03-EffectCorrelationFigure.Rmd
Updated figures pipeline

--- 2018-03-22 Analysis/EffectCorrelationFigure.Rmd
@@
 
-```{r, eval = FALSE, echo = FALSE}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
 pdf("../Writing/Figures/RF_results_best.pdf",width = 8,height = 8)
@@
 par(op)
-dev.off()
+# dev.off()
 ```
@@
 
-```{r, eval = FALSE, echo = FALSE}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
 pdf("../Writing/Figures/Tr_results_best.pdf", width = 8, height = 8)
@@
 
-```{r, eval = FALSE, echo = FALSE}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
 pdf("../Writing/Figures/RF_results_null.pdf", width = 8, height = 8)
@@
 
-```{r, eval = FALSE, echo = FALSE}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
 pdf("../Writing/Figures/Tr_results_null.pdf", width = 8, height = 8)

### 2018-03-27 Functions/tree.resolution.R

@@ -11,7 +11,7 @@
 #' @author Thomas Guillerme
 #' @export
 
-# path = "../Data/Trees_out/Consensus_trees/Bayesian/Done/25t_100c/"
+# path = "../Data/Trees_out/Consensus_trees/Bayesian/Trees/25t_100c/"
 # suffix = "con.tre"
 
 tree.resolution <- function(path, suffix) {
@@ -55,12 +55,12 @@ plot.tree.resolution <- function(path, ntaxa, col, type = "line") {
     bayesian_path <- parsimon_path <- list()
 
     ## Get the full path for both methods
-    bayesian_path[[1]] <- paste0(path, "Bayesian/Done/", ntaxa, "t_100c/")
-    parsimon_path[[1]] <- paste0(path, "Parsimony/Done/", ntaxa, "t_100c/")
-    bayesian_path[[2]] <- paste0(path, "Bayesian/Done/", ntaxa, "t_350c/")
-    parsimon_path[[2]] <- paste0(path, "Parsimony/Done/", ntaxa, "t_350c/")
-    bayesian_path[[3]] <- paste0(path, "Bayesian/Done/", ntaxa, "t_1000c/")
-    parsimon_path[[3]] <- paste0(path, "Parsimony/Done/", ntaxa, "t_1000c/")
+    bayesian_path[[1]] <- paste0(path, "Bayesian/Trees/", ntaxa, "t_100c/")
+    parsimon_path[[1]] <- paste0(path, "Parsimony/Trees/", ntaxa, "t_100c/")
+    bayesian_path[[2]] <- paste0(path, "Bayesian/Trees/", ntaxa, "t_350c/")
+    parsimon_path[[2]] <- paste0(path, "Parsimony/Trees/", ntaxa, "t_350c/")
+    bayesian_path[[3]] <- paste0(path, "Bayesian/Trees/", ntaxa, "t_1000c/")
+    parsimon_path[[3]] <- paste0(path, "Parsimony/Trees/", ntaxa, "t_1000c/")
 
 
     ## Get the trees

### 2018-03-27 Writing/draft.tex
updated figure 2 caption

@@
 Blue colours correspond to low character differences and orange colours correspond to high character differences.
-\textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised character difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).}
+\textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised character difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).
+The differences between the characters is low in \textbf{C} (minimised compared to \textbf{A}) implying a high correlation between the characters.
+Conversely, the character differences is high in \textbf{D} (maximised compared to \textbf{A}) implying a high correlation between the characters.}
 \label{Fig:modif_matrix}
@@
 
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{ll|r|rr}
-%   \hline
-% metric & test & bhatt.coeff & statistic & p.value \\ 
-%   \hline
-% RF & maxi:mini & 0.588 & 439466.500 & \textbf{0.000} \\ 
-%    & maxi:rand & 0.894 & 347624.500 & \textbf{0.000} \\ 
-%    & mini:rand & 0.846 & 115008.500 & \textbf{0.000} \\ 
-% Tr & maxi:mini & 0.628 & 443049.500 & \textbf{0.000} \\ 
-%    & maxi:rand & 0.921 & 348074.000 & \textbf{0.000} \\ 
-%    & mini:rand & 0.825 & 113171.500 & \textbf{0.000} \\ 
-%    \hline
-% \end{tabular}
-% \caption{Difference between the pooled scenarios.
-bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction), significant values are in bold.}
-% \label{Tab_pooledscenarios_test}
-% \end{table}
-
 \subsubsection{Number of characters}
-%MB: Here and other similar subsection headings: the effect on what?
-%TG: quick fix!
-
 This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
@@
 %TG: agree.
-
-TODO
-
 First, simulating evolutionary history is complex.
@@
 For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly decreased and increased correct topological recovery compared to randomly resample matrices.
+%TODO
 %MB: But at what point? How much correlation is needed to affect it? TG: added values.

### 2018-03-27 Analysis/02-EffectCorrelationFullResults.Rmd
added the looooong tables

@@
 
-```{r, echo = FALSE, eval = FALSE}
+```{r, echo = FALSE, eval = TRUE, print = FALSE, result = 'hide'}
 ## Labels
@@
 ## Tables
-make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1], path = "../Writing/Tables/")
-make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2], path = "../Writing/Tables/")
-make.xtable(sum_RF_null, caption = captions_summary[3], label = labels_summary[3], path = "../Writing/Tables/")
-make.xtable(sum_Tr_null, caption = captions_summary[4], label = labels_summary[4], path = "../Writing/Tables/")
+make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1], path = "../Writing/Tables/", longtable = TRUE)
+make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2], path = "../Writing/Tables/", longtable = TRUE)
+make.xtable(sum_RF_null, caption = captions_summary[3], label = labels_summary[3], path = "../Writing/Tables/", longtable = TRUE)
+make.xtable(sum_Tr_null, caption = captions_summary[4], label = labels_summary[4], path = "../Writing/Tables/", longtable = TRUE)
 ```

### 2018-03-27 Writing/supplementary/Sup1-CharDiff.tex
supplementary updated

@@
 
-\noindent{\Large \bf Effect of discrete character correlation on tree inference}
+\noindent{\Large \bf Influence of different modes of morphological character correlation on phylogenetic tree inference}
+\bigskip
 \noindent{\Large \bf Supplementary material 1 - Character difference}
 \bigskip
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
+\noindent {\small \it 
+$^1$School of Biological Sciences, University of Queensland, St. Lucia, Queensland, Australia.\\
+$^2$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\
+$^3$Department of Earth Sciences, Natural History Museum, Cromwell Road, London, SW75BD, United Kingdom.\\}
 
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^*$, and Martin D. Brazeau$^1$}\\
-\noindent {\small \it 
-$^1$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\}
 \end{center}
 \medskip
-\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\
+\noindent{*\bf Corresponding author.} \textit{guillert@tcd.ie}\\ 
+\vspace{1in}
 

### 2018-03-27 Writing/supplementary/Sup2-SimParam.tex
supplementary updated

@@
 
-\noindent{\Large \bf Effect of discrete character correlation on tree inference}
+\noindent{\Large \bf Influence of different modes of morphological character correlation on phylogenetic tree inference}
 \bigskip
 \noindent{\Large \bf Supplementary material 2 - Simulation parameters}
+\bigskip
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
+\noindent {\small \it 
+$^1$School of Biological Sciences, University of Queensland, St. Lucia, Queensland, Australia.\\
+$^2$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\
+$^3$Department of Earth Sciences, Natural History Museum, Cromwell Road, London, SW75BD, United Kingdom.\\}
 
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^*$, and Martin D. Brazeau$^1$}\\
-\noindent {\small \it 
-$^1$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\}
 \end{center}
 \medskip
-\noindent{*\bf Corresponding author.} \textit{t.guillerme@imperial.ac.uk}\\
+\noindent{*\bf Corresponding author.} \textit{guillert@tcd.ie}\\ 
+\vspace{1in}
 

### 2018-03-27 Writing/draft.tex
Updated final results

@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology with maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the character difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see Supplementary material 3 for the full summary statistics).
-There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see Supplementary material 3 for the full summary statistics).
+There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
 
@@
 This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
-The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.76, 0.78, 0.8 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 and 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
 %MB: The above sentence has too many "for"s. It's too long and needs to be split to be intelligible.
 %TG: fix attempt
-The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
 However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.
@@
 Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
-A low number of taxa (25) resulted in significant differences in both $NTS_{RF}$ and $NTS_{Tr}$ (medians for 25, 75 and 150 taxa equals 0.8, 0.76, 0.75 $NTS_{RF}$ and 0.75, 0.59 and 0.59 $NTS_{Tr}$respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
+A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
 Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
@@
 
-Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.82 and 0.68 respectively) but not in terms of individual taxon placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.72 and 0.61 respectively).
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively) but not in terms of individual taxon placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively).
 

### 2018-03-28 Analysis/04-EffectCorrelation_lm.Rmd
added beta regressions (?)

@@
 model_out <- run.lm(matrices_cd, whole_data, chain = "25t_100c", method = "Bayesian", metric = "RF", cent.tend = NULL)
+model_beta <- run.lm(matrices_cd, whole_data, chain = "25t_100c", method = "Bayesian", metric = "RF", cent.tend = NULL, type = "beta")
 
@@
 ```{r, fig.width=12, fig.height=12, warnings = FALSE}
-## Pool all the results into one list
-pooled_results <- pool.matrix.cd(all_matrices_cd, length.out = 28)
+# ## Pool all the results into one list
+# pooled_results <- pool.matrix.cd(all_matrices_cd, length.out = 28)
 
-# ## Running the model
-model_out_bay_RF <- run.lm(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", cent.tend = NULL)
-model_out_bay_Tr <- run.lm(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "Triplets", cent.tend = NULL)
-model_out_par_RF <- run.lm(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "RF", cent.tend = NULL)
-model_out_par_Tr <- run.lm(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", cent.tend = NULL)
+# # ## Running the model
+# model_out_bay_RF <- run.lm(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", cent.tend = NULL)
+# model_out_bay_Tr <- run.lm(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "Triplets", cent.tend = NULL)
+# model_out_par_RF <- run.lm(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "RF", cent.tend = NULL)
+# model_out_par_Tr <- run.lm(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", cent.tend = NULL)
 
-## Plotting the output
-par(bty = "n", mfrow = c(2,2))
-one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", model_out_bay_RF, cent.tend = NULL, legend = TRUE, main = "Bayesian", xlab = "RF")
-one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", model_out_bay_Tr, cent.tend = NULL, legend = FALSE, main = "Bayesian", xlab = "RF")
-one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", model_out_par_RF, cent.tend = NULL, legend = FALSE, main = "Parsimony", xlab = "Triplets")
-one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", model_out_par_Tr, cent.tend = NULL, legend = FALSE, main = "Parsimony", xlab = "Triplets")
+# ## Plotting the output
+# par(bty = "n", mfrow = c(2,2))
+# one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", model_out_bay_RF, cent.tend = NULL, legend = TRUE, main = "Bayesian", xlab = "RF")
+# one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Bayesian", metric = "RF", model_out_bay_Tr, cent.tend = NULL, legend = FALSE, main = "Bayesian", xlab = "RF")
+# one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", model_out_par_RF, cent.tend = NULL, legend = FALSE, main = "Parsimony", xlab = "Triplets")
+# one.correlation.plot(pooled_results, whole_data, chain = "pool", method = "Parsimony", metric = "Triplets", model_out_par_Tr, cent.tend = NULL, legend = FALSE, main = "Parsimony", xlab = "Triplets")
 ```
@@
 
+<!-- only 75t_100c correctly passes the lm assumptions -->
+
 ```{r, fig.width=12, fig.height=12}
@@
 
+<!-- only 75t_100c and 150t_100c correctly passes the lm assumptions -->
+
 ```{r, fig.width=12, fig.height=12}
@@
 
+<!-- only 150t_100c correctly passes the lm assumptions -->
+
 ```{r, fig.width=12, fig.height=12}

### 2018-03-28 Functions/correlation.lm.R

@@ -130,7 +130,8 @@ pool.matrix.cd <- function(list, length.out) {
 #' @param metric which metric to plot (\code{"RF"} or \code{"Triplets"})
 #' @param method which method to plot (\code{"Bayesian"} or \code{"Parsimony"}) 
 #' @param cent.tend the central tendency to plot for matrix_cd (\code{default = mean})
-#' 
+#' @param type which type of model to use: simple "linear" or "beta" regression?
+#'  
 #' @examples
 #'
 #' @seealso
@@ -138,7 +139,7 @@ pool.matrix.cd <- function(list, length.out) {
 #' @author Thomas Guillerme
 #' @export
 
-run.lm <- function(matrices_cd, whole_data, chain, metric, method, cent.tend = NULL) {
+run.lm <- function(matrices_cd, whole_data, chain, metric, method, cent.tend = NULL, type = "linear") {
 
     ## Get the metric
     if(metric == "RF") {
@@ -206,8 +207,8 @@ run.lm <- function(matrices_cd, whole_data, chain, metric, method, cent.tend = N
                         )
                     )
     } else {
-        warning("DEBUG: Change length.out value in run.lm and one.correlation.plot")
-        length.out <- 28
+        # TODO: change
+        length.out <- 35
         data_metric <- data.frame(
             "Metric" = c(
                 unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$mini[1:length.out, metric_value]), method_value, metric_value, length.out)),
@@ -236,7 +237,11 @@ run.lm <- function(matrices_cd, whole_data, chain, metric, method, cent.tend = N
 
 
     ## Running the model
-    return(lm(Metric ~ CD, data = data_table))
+    if(type != "beta") {
+        return(lm(Metric ~ CD, data = data_table))
+    } else {
+        return(betareg::betareg(Metric ~ CD, data = data_table))
+    }
 }
 
 

### 2018-03-28 Functions/correlation.results.R

@@ -119,8 +119,10 @@ one.correlation.plot <- function(matrices_cd, whole_data, chain, metric, method,
         y_maxi <- whole_data[[taxa_value]][[character_value]][[method_value]]$norm$maxi[, metric_value][1:length.out]
         y_rand <- whole_data[[taxa_value]][[character_value]][[method_value]]$norm$rand[, metric_value][1:length.out]
     } else {
-        warning("DEBUG: Change length.out value in run.lm and one.correlation.plot")
-        length.out <- 28
+        
+        ## TODO change
+        length.out <- 35
+
         y_mini <- unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$mini[1:length.out, metric_value]), method_value, metric_value, length.out))
         y_maxi <- unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$maxi[1:length.out, metric_value]), method_value, metric_value, length.out))
         y_rand <- unlist(lapply(whole_data, lapply, function(X, method_value, metric_value, length.out) return(X[[method_value]]$norm$rand[1:length.out, metric_value]), method_value, metric_value, length.out))

### 2018-04-03 Writing/draft.tex
Re-read through the methods up to methods - simulating data

@@
 \begin{abstract}
-Phylogenetic analysis algorithms require the assumption of character independence---a condition generally acknowledged to be violated by morphological data.
-%However, this assumption is nearly always wrong simply because of the nature of morphological integration of organisms.
+Phylogenetic analysis algorithms require the assumption of character independence - a condition generally acknowledged to be violated by morphological data.
 Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
@@
 Overall, we found an expected positive effect of reducing character correlations on recovering the correct topology.
-However, this effect is less important for matrices with a small number of taxa (~25) were reducing character correlation is not more effective than randomly drawing characters.
-Furthermore, in bigger matrices (~350 taxa), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony  trees.
+However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing character correlation is not more effective than randomly drawing characters.
+Furthermore, in bigger matrices (350 characters), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
 
@@
 This opens up possibilities for predicting problematic datasets and possibly proposing new confidence measures in phylogenetic datasets.
-%The last decade has witnessed a resurgence of use of discrete morphological data in macroevolutionary studies.
-%Traditionally confined to ``classic'' cladistic studies in palaeontology, these methods have recently been widely expanded and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep[e.g.][]{pyrondivergence2011,ronquista2012}.
-%In parallel, much development has also been done on the pre-phylogenetic analysis \citep[e.g.
-data collection;][]{morphobank} as well as on post-phylogenetic ones \citep[e.g.
-morphological disparity analysis;][]{Close2015,Claddis}.
-%Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011,simoes2017giant}; the topology inference method \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.
 
 The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data.
-The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony .
-\citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
+The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
 Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the very nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
-% TG: add Brazeau et al. 201X somewhere?
 Before discussing character correlation further, it is important to understand that it may manifest itself in at least three distinct ways:
@@
     Characters of the occlusal surface of two opposing molars will be expected to directly covary.
-	% We should probably cite some "Gunterology" here.
+    %MB: We should probably cite some "Gunterology" here.
     Pleiotropy also results in covariation between different aspects of phenotype. 
     From a phylogenetic perspective, it can be especially pernicious because the relationship between the traits in question may have no obvious link from a morphological or functional comparison alone.
-    Intra-organismal links can be the targets of comparative developmental biology \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary} or functional investigations.
+    Intra-organismal links can be the targets of comparative developmental biology \citep{goswami2010,kelly2010reduced,stoessel2013morphological,goswami2014macroevolutionary} or functional investigations.
 
@@
     For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
-    This coding dependency is linked to hierarchical dependency between characters [TODO: cite Wilkinson on inapplicability].
+    This coding dependency is linked to hierarchical dependency between characters \citep{wilkinson1995coping,BrazeauNA}.
     Finally this can also be due to a bias in the amount of characters available.
@@
 Logical dependence, however, is easily distinguished prior to phylogenetic inference, while the two other ones (intra-organismal and evolutionary) are much harder.
-However, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies.
-% Of course, at this point we could probably self-cite and cite De Laet.
+However, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{DeLaet2015,BrazeauNA}.
 Intra-organismal dependence requires more detailed, often extremely time-consuming studies (and possibly beyond the limits of available technology).
@@
 Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the different character states tokens will be used in order to infer a phylogenetic tree.
-%MB: We believe two reasonable things: some characters will be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis. 
 Some character will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
-%MB: How bad does it have to be to be a real effect?
-Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence})
+Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).
+
 How does these correlation really affect topology?
@@
 with high levels of homoplasy) actual cancel out the effects of correlation?
-
 We formally assess the effect of discrete character's correlation using simulated data.
@@
 This step resulted in 9 matrices.
-%MB: HOW do you simulate them? What models do you use? What protocol? Are the simulated matrices expected to recover the correct topology in the absence of our manipulations? 
-%TG: I've specified above that this is just a summary.
-Details are below.
-Regarding the question, the recovery of the correct topology without manipulation is not perfect (of course) but ranges roughly between 0.2 and 1 (in both normalised triplets and RF).
-In a way though it's not that much of a big deal since the "True" tree is never known.
-Here we compare the modified matrices to the "Best" tree (the one we obtain by infering the tree from the simulated and unmodified "normal" matrices).
 
@@
 Our protocol for measuring character difference is detailed below.
-%MB: I don't understand this. Would a matrix with a "maximised" level of correlation simply be a uniform matrix?
-%MB: What do you mean by "least" and "most different"? What is the measure of similarity? 
-%MB: What do you mean "replacing them randomly by the remaining characters? Also: won't that simply create a correlation with another character in the dataset?
-%TG: Again, I've hihghlighted that precisions come below.
-To answer the questions, yes to both.
-However we do not remove/replace ALL the characters thus just changing the average correlation level.
-This is kind of illustrated in figure 2.
+
     We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character differences to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
@@
 
-\item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both maximum parsimony  (MP) and Bayesian inference (BPP).
+\item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both maximum parsimony % (MP) %TG: not a big fan of the abbreviations - however if you think they're necessary, we can always add them.
+ and Bayesian inference.% (BPP). 
 Hereafter, the resultant topologies are called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees). This step resulted in 72 topologies.
@@
 %MB: I can't be sure, but this could as well be Farris's (1970) advancement index as applied to characters rather than taxa?
-%TG: TODO: check Farris 1970 and whether this is new at all.
+%TG: OK, I've checked Farris' stuff.
+Despite not beeing described (at least not the in 1970 Syst Biol paper), it seems that the advancement index is just the distance between two taxa.
+Appart from the taxa/character difference, it's kind of similar but (1) it's not ponderated (divided by n), centered (-1/2) and reversed (1-2(...)); so I'm not sure if we need to say it's similar.
+In the end it's a distance metric and it can be pretty similar to *any* distance metric.
+I think we can leave as it is (plus, I've reinvented the wheel by mathematicallty demonstrating it's a distance - my first only real math in a draft).
 
@@
 In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
-%TG: Justification for the standardisation: basically, it makes comparisons faster and Fitch-like.
 This procedure was used to treat all the characters are unordered with no assumption on the meaning of the character state (e.g. in a binary character $0$ is not necessary ancestral to $1$).
@@
 When the character difference is maximal ($1$) it means it conveys the greatest difference in phylogenetic signal (i.e. characters are uncorrelated).
-It is important to stress that a character difference of $0$ (i.e. the same phylogenetic signal) does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal) .
+It is important to stress that a character difference of $0$ (i.e.
+the same phylogenetic signal) does not mean the opposite of $1$ (i.e.
+\textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
 For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
-Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different phylogenetic interpretation.
+Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different set of splits (i.e. a different phylogenetic signal).
 
@@
 These trees are hereafter called the ``true'' trees (see distinction below).
-%MB: Yay! The Truth! Stop sciencing!
 We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
 \begin{itemize}
-    \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ and scaled to sum to $1$)
-    %TG: is that not a random Dirichlet? D(A,G,C,T)? TODO: check
-     and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
+    \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
     This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use an M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.

### 2018-04-04 Writing/draft.tex
Finished double checking the results

@@
 randomly weighting characters).
-% Clever.
 Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
@@
 Blue colours correspond to low character differences and orange colours correspond to high character differences.
-\textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised character difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).
+\textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised (\textbf{C}), minimised (\textbf{D}) or randomise character difference (\textbf{E}).
 The differences between the characters is low in \textbf{C} (minimised compared to \textbf{A}) implying a high correlation between the characters.
-Conversely, the character differences is high in \textbf{D} (maximised compared to \textbf{A}) implying a high correlation between the characters.}
+Conversely, the character differences is high in \textbf{D} (maximised compared to \textbf{A}) implying a low correlation between the characters.}
 \label{Fig:modif_matrix}
@@
 We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).]
-%TG: True tree comparison stuff:
 Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
-First, in biology, this tree is by definition unknown.
+First, in biology, this tree is always unknown.
 Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
-When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating birth-death trees simulating discrete morphological matrices from there.
+When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
 
@@
 The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary material 3).
-%MB: "the resulting in residuals..." I'm not sure what this sentence is supposed to say, so I can't suggest an edit.
-%TG: Probably a typo. Basically the residuals of lm or anovas cannot be normal because of this hard bound so we can't use parametric stats.
 Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.

### 2018-04-05 Analysis/Character_difference/01-Comparing_characters.Rmd
Rearranged analysis files

--- 2017-07-17 Analysis/Character_difference/Comparing_characters.Rmd
@@
 
-```{r}
+```{r, eval = FALSE}
 # Number of replicates:
@@
 
-```{r, fig.width = 12, fig.height = 12}
+```{r, fig.width = 12, fig.height = 12, eval = FALSE}
 op <- par(mfrow = c(2,2), bty = "n")
@@
 
-```{r, fig.width = 10, fig.height = 10}
+```{r, fig.width = 10, fig.height = 10, eval = FALSE}
 plot(first_character_differences[,1], type = "l", col = "lightblue", ylim = c(0,1), xlab = "",

### 2018-04-05 Writing/draft.tex
Updated full manuscript

@@
 Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
-
-% However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
-% Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e.
-as an equivalent of randomising the character differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).
-%MB: Agreed that that is noteworthy.
-I suspect I know why this happens (and you might have explained it earlier).
-It might actually be a statistical problem with our approach.
-When character similaritiy is maximized, how well resolved are the trees?
-%TG: I've moved that to a "potential application" paragraph in the discussion (please change the title of it, I had zero inspiration and find it ugly as feck!)
+Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary materials @@@).
 
@@
 the more characters available, the more characters are modified in each scenario).
-Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance.)
+Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance).
 
@@
 This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
-For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation and is independent of the number of characters analysed.
+For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation as it is independent of the number of characters analysed.
 Similarly, using the a Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
@@
 When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
-The difference in the ability of each method to recover the ``correct'' topology has been heavily discussed in the last five years with some, indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have raised issues with these investigations \citealt{spencerefficacy2013,goloboff2017weighted}).
+The difference in the ability of each method to recover the ``correct'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have raised issues with these investigations \citealt{spencerefficacy2013,goloboff2017weighted}).
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e.
@@
 \subsection{Distinction between different character correlations}
-%TG: I've changed this paragraph - basically it's linking back to the intro saying the only correlation we should care about in phylogenetics is the one raising from data collection.
 Here we mention three different types of character correlations but evolutionary biologists are mainly interested in the intra-organismal and evolutionary correlations (e.g.
@@
 \subsection{Limitations}
-%MB: I think what this section _really_ needs is an assessment of the method used to simulate correlations.
-It was basically accomplished by duplicating characters.
-It wasn't accomplished by evolving some characters with a particular covariance model (which probably would've been better).
-We kind of faked-up that model.
-%TG: agree.
 First, simulating evolutionary history is complex.
-Not only because the models we're using to infer phylogenies are ever improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible.
-This can lead to spuriously generated discrete morphological matrices
+Not only because the models we're using to infer phylogenies are ever improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible \citep[see constrasted discussions from][]{GoloboffEmpirical,OReillyEmpirical}.
 However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree.
@@
 For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly decreased and increased correct topological recovery compared to randomly resample matrices.
-%TODO
-%MB: But at what point? How much correlation is needed to affect it? TG: added values.
-
-%MB: Part of it is: we expect it to be a problem, but _when_ do we expect it to be a problem.
-There's the issue of "check your data", but it only raises the question: how?
-I think this is where we suggest avenues for future research (either here or move it up to the discussion).
 

### 2018-04-10 Writing/CoverLetter.tex
Rewrote the cover letter

@@
 Despite being at the base of almost every endeavour in modern Biology, the assumption made in tree inference that phylogenetic data is independent is probably mostly erroneous.
-In fact, since the advent of Cladistics in the 1950s, evolutionary biologists have made the assumption that characters used to build a phylogeny are independent despite the fact that we are looking at their variance-covariance to build phylogenetic hypothesis.
-This effect is somewhat negligible in molecular phylogenetics because the huge amount of available data is thought to temper down the correlation effect.
-However, in smaller morphological datasets, this can lead to pretty big biases and can easily favour on topology from another simply with a ``judicious'' choice of characters.
+This assumption was stated since the advent of Cladistics in the 1950s but is nowadays mainly ignored in phylogenetic endeavour.
+In fact, this effect can be neglected in molecular datasets because the huge amount of available data is thought to temper down the correlation effect.
+However, in morphological datasets, where the amount of characters is much smaller, character correlation can lead to significant biases especially by favouring one topology from another due to a specific choice of characters (whether this is done consciously by researchers or not).
+With the ``resurgence'' of morphological data phylogenetic analysis (namely with the Total Evidence tip-dating method, Ronquist \textit{et al.}, 2012, Sys.
+Biol), it is thus important to acknowledge limitations and caveats that come with the use of such data, namely how correlation patterns in morphological datasets can effect tree inference.
 
-Our research article, entitled ``Influence of different modes of morphological character correlation on tree inference'', is - to our knowledge - the first to thoroughly test the effect of character correlation on tree topology using a thorough simulation framework ($>$ 100 CPU years worth).
-Using a novel metric to measure the amount of phylogenetic correlation in a matrix, we tested different scenario that either increase or decrease the amount of correlation of characters generated from a known ``true'' tree and compare how these levels of correlation affect our ability to recover this topology using both Maximum Parsimony and Bayesian inference.
+In our research article, entitled ``Influence of different modes of morphological character correlation on tree inference'', we propose the first - to our knowledge - simulation based test of the effect of character correlation on tree topology using a thorough simulation framework ($>$ 100 CPU years worth).
+Using a novel metric to measure direct phylogenetic correlation between characters, we simulated discrete morphological datasets that we then modified to increase or decrease character correlation.
+We then measured the resulting topological differences (from both maximum parsimony and Bayesian trees) to measure the effect of character correlation in the light of our known simulated parameters.
 
-Our results suggests that strong levels of correlation can negatively impact our ability to recover the correct tree topology and more so in Maximum Parsimony than in Baysian.
-However, it is encouraging to see that low levels of correlation do not especially impact the topology, at least not to the extant where researchers should specifically collect only uncorrelated characters.
-Finally, we show that, for smaller datasets, Maximum Parsimony and Bayesian inference are both pretty efficient to recover the correct topology, regardless of the amount of correlation
+Our results suggest that datasets that have higher levels of internal correlation negatively impact our ability to recover the correct tree topology.
+This is too be contrasted with datasets that have low levels of correlation (i.e.
+more noisy) which can still recover the correct topology, depending on the amount of data and the inference method used.
+These results, although non-surprising to researchers in the field, are the first - to our knowledge - to actually quantify the effect of character correlation and provides practical data and reusable code allowing researcher to effectively measure the limitations of their datasets.
+Furthermore, we suggest that using our proposed new metric, it would be able to ``bias'' bootstrap pseudo-replications in phylogenetic analysis by favouring character with high or low correlation, depending on the purpose of the phylogenetic search.
 

### 2018-04-18 Writing/draft.tex
checked manuscript up until results

@@
 Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with good practices, phylogenetic software do not distinguish between any of them.
+
 In this study, we propose a new metric of raw character difference as a proxy for character correlation.
@@
 
-\noindent (Keywords: Character difference, correlation, topology, Bayesian, Maximum Parsimony)\\
+\noindent (Keywords: Character difference, correlation, topology, Bayesian, maximum parsimony)\\
 
@@
 
-The last two decades have witnessed a resurgence of interest in the use of morphological character data in phylogenetic studies.
+The last two decades have witnessed a ``resurgence'' of interest in the use of morphological character data in phylogenetic studies.
 This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
@@
 While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
-As morphological data are likely to continue to play an extensive role in phylogenetic analysis, it is essential to understand the circumstances under which morphological data might be expected to 'misbehave'.
+As morphological data are likely to continue to play an extensive role in phylogenetic analysis, it is essential to understand the circumstances under which morphological data might be expected to ``misbehave''.
 This opens up possibilities for predicting problematic datasets and possibly proposing new confidence measures in phylogenetic datasets.
@@
     Characters of the occlusal surface of two opposing molars will be expected to directly covary.
-    %MB: We should probably cite some "Gunterology" here.
     Pleiotropy also results in covariation between different aspects of phenotype. 
@@
     For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
-    This is thought to correspond to adaptations for fossoriality or acquatic lifestyles.
-    Such covariances are generally studied in the context of a given phylogeny, often one derived from molecular data with morphological traits mapped on.
+    This is thought to correspond to adaptations for fossoriality or aquatic lifestyles.
+    Such covariances are generally studied in the context of a given phylogeny, often one derived from molecular data with the morphological traits of interest mapped on it.
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
@@
     For instance, coding for the same absence in different characters creates state transformations associated with the loss or gain of a particular character.
-    This occurs when a number of multistate characters include two variable feature states (e.g. large, small; red, blue etc.) in conjunction with absense.
+    This occurs when a number of multistate characters include two variable feature states (e.g. large, small; red, blue etc.) in conjunction with absence.
     It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
@@
     Finally this can also be due to a bias in the amount of characters available.
-    For example, in skulls, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity).
+    For example, in skulls, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity to discrete characters).
 \end{itemize}
@@
 These sources of dependence between characters are well studied in biology.
-Biological and evolutionary dependences are inherent parts to evo-devo and macroevolution and best practices to avoid coding-induced dependences are commonly known and applied.
+Biological and evolutionary dependences are inherent parts to Evo-Devo and macroevolutionary studies and best practices to avoid coding-induced dependences are commonly known and applied.
 However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inferences software that are blind to these distinctions.
@@
 \noindent \textbf{Correlation between characters detected by the software}: this is the result how software actually interprets the differences between characters.
-The vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different/similar tokens).
+The vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different or similar tokens).
 Therefore a great number of characters and - traditionally - a few number of tokens can easily lead to dependence between characters.
@@
 In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
-Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the different character states tokens will be used in order to infer a phylogenetic tree.
-Some character will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
+Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states tokens will be used in order to infer the tree.
+Some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
 Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).
@@
 How does these correlation really affect topology?
-%MB: If you've got intrinsic correlations, then those correlations could occur across anatomical systems. Would their effects cancel each other out?
-%TG: Actually not sure what you mean by this one and whether this is something we can test with this paper?
-%MB: Will "noisier" characters accurately cancel out the effects of correlation?
-Furthermore, we expect matrices with a high level of correlation to recover precise but inaccurate topologies but will matrices with low level of correlation (i.e.
+We expect matrices with a high level of correlation to recover precise but inaccurate topologies but will matrices with low level of correlation (i.e.
 with high levels of homoplasy) actual cancel out the effects of correlation?
-We formally assess the effect of discrete character's correlation using simulated data.
-We propose a new metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences/similarities between characters.
+Here we formally assess the effect of discrete character's correlation using simulated data.
+We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences or similarities between characters.
 We found that overall, there is a detectable effect of character correlation on topology where an increase in character dependence results in a decrease in the ability to recover the correct topology.
-These results, however, vary in magnitude depending on the size of matrix and the inference method used.
+These results, however, vary greatly in magnitude depending on the size of matrix and the inference method used.
 
@@
 
-    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least different or most different characters and replacing them randomly by the remaining characters.
+    \item \textbf{Modifying matrices}: we changed the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least different or most different characters and replacing them randomly by the remaining characters.
 Our protocol for measuring character difference is detailed below.
@@
 
-\item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both maximum parsimony % (MP) %TG: not a big fan of the abbreviations - however if you think they're necessary, we can always add them.
- and Bayesian inference.% (BPP). 
-Hereafter, the resultant topologies are called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees). This step resulted in 72 topologies.
+\item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both maximum parsimony and Bayesian inference.
+Hereafter, the resultant topologies are called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
+This step resulted in 72 topologies.
     \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on topology.
@@
 In order to measure this, we propose a new distance metric to measure the difference between two characters:
-%MB: Are you sure this is new? I can't read the formula because it's in LaTeX gobbledygook but there exist methods for measuring character similarity,
-%MB: I can't be sure, but this could as well be Farris's (1970) advancement index as applied to characters rather than taxa?
 %TG: OK, I've checked Farris' stuff.
@@
 For each tree, we arbitrarily set the outgroup to be the first taxon (alphabetically) thus effectively rooting the trees on this taxon.
-%MB: If we are using undirected characters, does it matter what the rooting is?
-%TG: Not at all but it speeds up the MCMC
 These trees are hereafter called the ``true'' trees (see distinction below).
@@
 for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).
+
+The resulting full simulation was 4TB @@@ big so is not shared here (though the parameters are).
+However, the resulting consensus trees on which the topological differences are calculated are available at @@@.
 

### 2018-04-19 Functions/TreeCmp/src/treecmp/command/RunMCommand.java
zipped TreeCmp

@@ -1,124 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.command;
-
-import java.util.ArrayList;
-import treecmp.common.ProgressIndicator;
-import treecmp.common.StatCalculator;
-import treecmp.common.SummaryStatCalculator;
-import pal.tree.Tree;
-import treecmp.common.AlignWriter;
-import treecmp.common.ReportUtils;
-import treecmp.common.TreeCmpException;
-import treecmp.io.ResultWriter;
-import treecmp.io.TreeReader;
-import treecmp.config.ActiveMetricsSet;
-import treecmp.config.IOSettings;
-import treecmp.metric.Metric;
-
-public class RunMCommand extends Command {
-
-    public RunMCommand(int paramNumber, String name) {
-        super(paramNumber, name);
-    }
-
-    @Override
-    public void run() throws TreeCmpException{
-        super.run();       
-        out.init();
-        reader.open();
-        
-        matrixCompareExecute(reader, out);
-        
-        reader.close();
-        out.close();
-    }
-
-    private void matrixCompareEx(TreeReader reader, ResultWriter out, StatCalculator[] metrics) throws TreeCmpException {
-
-        pal.tree.Tree tree1,tree2 ;
-        ArrayList<Tree> tree_vec = new ArrayList<Tree>();
-        int k,counter,maxIt;
-        double val;
-        String row;
-
-        int mSize = metrics.length;
-
-        //initialize summary stat calculators
-        SummaryStatCalculator[] sStatCalc=new SummaryStatCalculator[mSize];
-        for(int i=0;i<mSize;i++){
-            sStatCalc[i]=new SummaryStatCalculator(metrics[i]);
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/command/RunSCommand.java

@@ -1,115 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.command;
-
-import treecmp.common.AlignWriter;
-import treecmp.common.ProgressIndicator;
-import treecmp.common.ReportUtils;
-import treecmp.common.StatCalculator;
-import treecmp.common.SummaryStatCalculator;
-import treecmp.common.TreeCmpException;
-import treecmp.io.ResultWriter;
-import treecmp.io.TreeReader;
-import treecmp.config.ActiveMetricsSet;
-import treecmp.metric.Metric;
-
-public class RunSCommand extends Command {
-
-    public RunSCommand(int paramNumber, String name) {
-        super(paramNumber, name);
-    }
-
-    @Override
-    public void run() throws TreeCmpException {
-        super.run();
-        out.init();
-        reader.open();
-        
-        pairCompareExecute(reader, out);
-        
-        reader.close();
-        out.close();
-    }
-
-    public void pairCompareExecute(TreeReader reader, ResultWriter out ) throws TreeCmpException{
-
-        Metric[] metrics = ActiveMetricsSet.getActiveMetricsSet().getActiveMetricsTable();
-        StatCalculator[] statsMetrics = new StatCalculator[metrics.length];
-
-        for(int i=0;i<metrics.length;i++){
-            statsMetrics[i]=new StatCalculator(metrics[i]);
-        }
-        pairCompareEx(reader, out, statsMetrics);
-    }
-
-private void pairCompareEx(TreeReader reader, ResultWriter out, StatCalculator[] metrics ) throws TreeCmpException {
-
-        pal.tree.Tree tree1 = reader.readNextTree();
-        pal.tree.Tree tree2;
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/commandline/CommandLineParser.java

@@ -1,372 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.commandline;
-
-import java.util.ArrayList;
-import java.util.Collections;
-import java.util.Comparator;
-import java.util.LinkedHashMap;
-import java.util.List;
-import java.util.ListIterator;
-import org.apache.commons.cli.CommandLine;
-import org.apache.commons.cli.GnuParser;
-import org.apache.commons.cli.HelpFormatter;
-import org.apache.commons.cli.Option;
-import org.apache.commons.cli.OptionGroup;
-import org.apache.commons.cli.Options;
-import org.apache.commons.cli.ParseException;
-import treecmp.config.*;
-import treecmp.metric.Metric;
-import treecmp.command.*;
-
-public class CommandLineParser {
-
-    private final static String S_DESC = "- Overlapping pair comparison mode. Every two neighboring trees are compared";
-    private final static String W_DESC = "- Window comparison mode. Every two trees within a window are compared.";
-    private final static String W_ARG = "size";
-    private final static String M_DESC = "- Matrix comparison mode. Every two trees in the input file are compared.";
-    private final static String R_DESC = "- Single tree to all tree mode. Each tree in the input file is compared to the single referenced tree.";
-    private final static String R_ARG = "refTreeFile";
-    private final static String I_DESC = "- Input file.";
-    private final static String I_ARG = "inputFile";
-    private final static String O_DESC = "- Output file.";
-    private final static String O_ARG = "outputFile";
-    private final static String P_DESC = "- Prune compared trees if needed (trees can have different leaf sets).";
-    private final static String SS_DESC = "- Report normalized distances.";
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/common/AlignInfo.java

@@ -1,69 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.common;
-
-import pal.tree.Tree;
-
-public class AlignInfo {
-
-    private IntNodePair[] aln;
-    private Tree t1;
-    private Tree t2;
-    private int totalCost;
-    private boolean useClusters;
-
-    public Tree getT1() {
-        return t1;
-    }
-
-    public void setT1(Tree t1) {
-        this.t1 = t1;
-    }
-
-    public Tree getT2() {
-        return t2;
-    }
-
-    public void setT2(Tree t2) {
-        this.t2 = t2;
-    }
-
-    public IntNodePair[] getAln() {
-        return aln;
-    }
-
-    public void setAln(IntNodePair[] aln) {
-        this.aln = aln;
-    }
-
-    public boolean isUseClusters() {
-        return useClusters;
-    }
-
-    public void setUseClusters(boolean useClusters) {
-        this.useClusters = useClusters;
-    }
-
-    public double getTotalCost() {
-        return totalCost;
-    }
-
-    public void setTotalCost(int totalCost) {
-        this.totalCost = totalCost;
-    }
-}

### 2018-04-19 Functions/TreeCmp/src/treecmp/common/AlignWriter.java

@@ -1,238 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.common;
-
-import java.text.MessageFormat;
-import java.util.Arrays;
-import java.util.Comparator;
-import java.util.HashSet;
-import java.util.Set;
-import pal.tree.Node;
-import pal.tree.NodeUtils;
-import pal.tree.Tree;
-import treecmp.config.IOSettings;
-import treecmp.io.ResultWriter;
-
-public class AlignWriter {
-
-    public final static String TREE = "Tree: ";
-    public final static String TREE1 = "Tree1: ";
-    public final static String TREE2 = "Tree2: ";
-    public final static String REF_TREE = "Referenced tree";
-    public final static String S_SPACE = " ";
-    public final static String NEW_LINE = System.getProperty("line.separator");
-    public final static String ALIGN = "Alignemnt: ";
-    public final static String ALIGN_COST = "Alignemnt cost (distance): {0}";
-    public final static String ALIGN_SEP = "-----------------";
-    public final static String EQUAL_CLUSTERS = "Equal clusters in both trees:";
-    public final static String ALIGN_CLUSTERS = "Aligned clusters:";
-    public final static String EQUAL_SPLITS = "Equal splits in both trees:";
-    public final static String ALIGN_SPLITS = "Aligned splits:";
-    public final static String PAIR_COST = "Cost: {0}";
-    public final static String MATCH_CLUST = "{0}. {1} <-> {2}";
-    public final static String MATCH_EMPTY = "EMPTY";
- 
-    private String outFileName;
-    private boolean isGenAlignments;
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/common/ClustIntersectInfoMatrix.java

@@ -1,226 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.common;
-
-import java.util.ArrayList;
-import java.util.BitSet;
-import java.util.List;
-import pal.misc.IdGroup;
-import pal.tree.Node;
-import pal.tree.Tree;
-import pal.tree.TreeUtils;
-
-
-public class ClustIntersectInfoMatrix {
-
-    protected int intT1Num;
-    protected int extT1Num;
-    protected int intT2Num;
-    protected int extT2Num;
-
-    protected Tree t1;
-    protected Tree t2;
-    protected IdGroup idGroup;
-
-    protected int[] alias1;
-    protected int[] alias2;
-
-    public short [][] intCladeSize;
-    public short [] cSize1;
-    public short [] cSize2;
-    public BitSet [] intT1toLeafT2;
-    public BitSet [] intT2toLeafT1;
-    public List<ClustPair> eqClustList;
-    public boolean [] eqClustT1;
-    public boolean [] eqClustT2;
-
-    public ClustIntersectInfoMatrix(Tree t1, Tree t2, IdGroup idGroup) {
-           this.t1 = t1;
-           this.t2 = t2;
-           this.idGroup = idGroup;
-    }
-    
-    public void init(){
-        intT1Num = t1.getInternalNodeCount();
-        extT1Num = t1.getExternalNodeCount();
-
-        intT2Num = t2.getInternalNodeCount();
-        extT2Num = t2.getExternalNodeCount();
-
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/common/LapSolver.java

@@ -1,853 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-
-/**
- *
- * The following code is based on Roy Jonker's implementation of LAP Solver,
- * see details below.
- */
-/************************************************************************
-*
-*  lap.cpp
-   version 1.0 - 4 September 1996
-   author: Roy Jonker @ MagicLogic Optimization Inc.
-   e-mail: roy_jonker@magiclogic.com
-
-   Code for Linear Assignment Problem, according to
-
-   "A Shortest Augmenting Path Algorithm for Dense and Sparse Linear
-    Assignment Problems," Computing 38, 325-340, 1987
-
-   by
-
-   R. Jonker and A. Volgenant, University of Amsterdam.
-*
-*************************************************************************/
-package treecmp.common;
-
-
-public class LapSolver {
-
-    static final int BIG=100000000;
-    static final long BIG_LONG=Long.MAX_VALUE;;
-    
-    public static int lap(int dim,
-        int assigncost[][],
-        int rowsol[],
-        int colsol[],
-        int u[],
-        int v[])
-
-// input:
-// dim        - problem size
-// assigncost - cost matrix
-
-// output:
-// rowsol     - column assigned to row in solution
-// colsol     - row assigned to column in solution
-// u          - dual variables, row reduction numbers
-// v          - dual variables, column reduction numbers
-
-{
-  boolean unassignedfound;
-  int  i, imin, numfree = 0, prvnumfree, f, i0, k, freerow;
-  int  j, j1, j2=0, endofpath=0, last=0, low, up;
-  int min=0, h, umin, usubmin, v2;
-
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/common/ProgressIndicator.java

@@ -1,138 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.common;
-
-import java.util.Locale;
-
-public class ProgressIndicator {
-
-    private final static int ROW_PRECISION=2;
-    private long startTime;
-    private int maxVal;
-
-    private long lastPrintTime;
-
-    /**
-     * max number of seconds without reporting of the progress
-     */
-    private int printInterval;
-    private long printIntervalMilis;
-    private int lastPrintVal;
-    private double printPercentInterval;
-
-    private String dataFormat;
-
-    public ProgressIndicator() {
-        maxVal=0;
-        lastPrintTime=0;
-        lastPrintVal=0;
-        printInterval=0;
-        printPercentInterval=0;
-        startTime=0;
-
-
-    }
-
-    public double getPrintPercentInterval() {
-        return printPercentInterval;
-    }
-
-    public void setPrintPercentInterval(double printPercentInterval) {
-        this.printPercentInterval = printPercentInterval;
-    }
-
-
-
-
-    public int getMaxVal() {
-        return maxVal;
-    }
-
-    public void setMaxVal(int maxVal) {
-        this.maxVal = maxVal;
-    }
-
-    public int getPrintInterval() {
-        return printInterval;
-    }
-
-    public void setPrintInterval(int printInterval) {
-        this.printInterval = printInterval;
-        this.printIntervalMilis=printInterval*1000;
-    }
-
-
-   public void init()
-    {
-       long now = System.currentTimeMillis();
-       this.dataFormat="%1$."+ROW_PRECISION+"f";
-       startTime=now;
-
-        if(lastPrintTime==0)
-        {
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/common/ReportUtils.java

@@ -1,165 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.common;
-
-import java.util.Locale;
-import treecmp.config.IOSettings;
-
-
-public class ReportUtils {
-    private final static int ROW_PRECISION = 4;
-    public final static String ROW_DATA_FORMAT="%1$."+ROW_PRECISION+"f";
-    private final static IOSettings ioSet = IOSettings.getIOSettings();
-    private final static String sep = ioSet.getSSep();
-    private final static boolean pruneTrees = ioSet.isPruneTrees();
-    private final static boolean randomComparison = ioSet.isRandomComparison();
-
-    public final static String NUM_COLUMN = "No";
-    public final static String T1_COLUMN = "Tree1";
-    public final static String T2_COLUMN = "Tree2";
-    public final static String T_COLUMN = "Tree";
-    public final static String T1_TAXA= "Tree1_taxa";
-    public final static String T2_TAXA= "Tree2_taxa";
-    public final static String T_TAXA= "Tree_taxa";
-    public final static String RT_TAXA= "RefTree_taxa";
-    public final static String COMMON_TAXA= "Common_taxa";
-    public final static String YULE_FRAC= "_toYuleAvg";
-    public final static String UNIF_FRAC= "_toUnifAvg";
-    public final static String NA_FRAC= "N/A";
-    //it t2 ==-1 do not print t2
-    public static String getResultRow(int rowNum, int t1, int t2, StatCalculator[] stats){
-
-        StringBuilder sb = new StringBuilder();
-        double dist, distToYuleAvg, distToUnifAvg;
-        String distStr;
-
-        //we do not need num in ref mode - tree number has the same valuse
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/common/SplitDist.java

@@ -1,218 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.common;
-
-import java.util.BitSet;
-import pal.misc.IdGroup;
-import pal.tree.Node;
-import pal.tree.NodeUtils;
-import pal.tree.SplitSystem;
-import pal.tree.Tree;
-
-public class SplitDist {
-
-    /** Creates a new instance of SplitDist */
-    public SplitDist() {
-    }
-
-    //dist([A1|A2],[B1|B2])=0.5*min((A1 xor B1)+(A2 xor B2),(A1 xor B2)+(A2 xor B1))
-    public static double getDist1(boolean[] split1, boolean[] split2) {
-        int n = split1.length;
-        int eq = 0, neq = 0, q;
-        double metric = 0.0;
-
-        for (int i = 0; i < n; i++) {
-            if (split1[i] == split2[i]) {
-                eq++;
-            } else {
-                neq++;
-            }
-        }
-
-        if (eq > neq) {
-            q = eq;
-        } else {
-            q = neq;
-        }
-
-        return (double) (n - q);
-    }
-
-    public static int getDist1Int(boolean[] split1, boolean[] split2) {
-        int n = split1.length;
-        int eq = 0, neq = 0, q;
-
-        for (int i = 0; i < n; i++) {
-            if (split1[i] == split2[i]) {
-                eq++;
-            } else {
-                neq++;
-            }
-        }
-
-        if (eq > neq) {
-            q = eq;
-        } else {
-            q = neq;
-        }
-
-        return (n - q);
-    }
-
-
-    public static double getDistToO_1(boolean[] split) {
-        int n = split.length;
-
-        return Math.ceil(Math.floor(n / 2.0) / 2.0);
-
-    }
-
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/common/SummaryStatCalculator.java

@@ -1,184 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.common;
-
-import treecmp.io.ResultWriter;
-import treecmp.config.IOSettings;
-import treecmp.metric.Metric;
-
-
-public class SummaryStatCalculator {
-
-
-    private Metric met;
-    private int count;
-    private double sum;
-    private double sq_sum;
-    private double min;
-    private double max;
-
-    public SummaryStatCalculator() {
-
-        this.count=0;
-        this.sum=0.0;
-        this.sq_sum=0.0;
-
-        this.min=Double.MAX_VALUE;
-        this.max=-Double.MAX_VALUE;
-    }
-
-
-    /**
-     *
-     * @param _met
-     */
-    public SummaryStatCalculator(Metric _met) {
-
-        //call non-parameter constructor
-        this();
-        this.met=_met;
-    }
-
-    public void addMetric(Metric _met)
-    {
-        this.met=_met;
-        this.clear();
-    }
-
-
-   public void clear()
-  {
-        this.count=0;
-        this.sum=0.0;
-        this.sq_sum=0.0;
-
-        this.min=Double.MAX_VALUE;
-        this.max=-Double.MAX_VALUE;
-
-    }
-
-
-  public double getMax()
-  {
-   return this.max;
-  }
-
-  public double getMin()
-  {
-    return this.min;
-  }
-
-  public double getAvg()
-  {
-      double avg=Double.POSITIVE_INFINITY;
-      if (count>0)
-          avg=this.sum/(double)count;
-
-      return avg;
-  }
-
-  public double getVariance()
-  {
-      double var=Double.POSITIVE_INFINITY;
-      double avg;
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/common/TreeCmpUtils.java

@@ -1,849 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.common;
-
-import java.util.ArrayList;
-import java.util.Comparator;
-import java.util.HashMap;
-import java.util.HashSet;
-import java.util.LinkedList;
-import java.util.List;
-import java.util.Map;
-import java.util.Set;
-import pal.misc.IdGroup;
-import pal.misc.Identifier;
-import pal.misc.SimpleIdGroup;
-import pal.tree.Node;
-import pal.tree.NodeUtils;
-import pal.tree.Tree;
-import pal.tree.TreeTool;
-import pal.tree.TreeUtils;
-
-public class TreeCmpUtils {
-
-  public static int[][] calcLcaMatrix(Tree tree, IdGroup idGroup) {
-
-        int leafNum = tree.getExternalNodeCount();
-        int intNum = tree.getInternalNodeCount();
-        if (idGroup == null)
-            idGroup = TreeUtils.getLeafIdGroup(tree);
-
-        int[] alias = TreeUtils.mapExternalIdentifiers(idGroup, tree);
-        int[][] lcaMatrix = new int[leafNum][leafNum];
-        for (int i=0;i<leafNum;i++)
-            lcaMatrix[i][i] = -1;
-        TCUtilsNode[] nodeInfoTab = new TCUtilsNode[intNum];
-        Node childNode, curNode = tree.getExternalNode(0);
-        int childCount, nodeIndex, childIndex;
-        boolean endLoop = true;
-
-        while (endLoop) {
-            if (curNode.isRoot()) {
-                endLoop = false;
-            }
-
-            if (!curNode.isLeaf()) {
-                nodeIndex = curNode.getNumber();
-                childCount = curNode.getChildCount();
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/config/ConfigSettings.java

@@ -1,163 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-package treecmp.config;
-
-import java.io.File;
-import java.io.IOException;
-import java.util.logging.Level;
-import java.util.logging.Logger;
-import javax.xml.parsers.DocumentBuilder;
-import javax.xml.parsers.DocumentBuilderFactory;
-import org.w3c.dom.*;
-import org.xml.sax.SAXException;
-import treecmp.metric.BaseMetric;
-
-
-public class ConfigSettings {
-
-    private static ConfigSettings config;
-    private String configFile;
-    private String dataDir;
-
-    protected ConfigSettings(String configFile, String dataDir) {
-        this.configFile = configFile;
-        this.dataDir = dataDir;
-    }
-
-    public static ConfigSettings getConfig() {        
-        return config;
-    }
-
-    public String getConfigFile() {
-        return configFile;
-    }
-
-    public String getDataDir() {
-        return dataDir;
-    }
-    
-     public static void initConfig(String configFile, String dataDir) {
-        if (config == null) {
-            config = new ConfigSettings(configFile, dataDir);
-            config.readConfigFromFile();
-        }
-    }
-
-
-    private void readConfigFromFile() {
-
-        try {
-
-            File xmlFile=new File(configFile);
-            DocumentBuilderFactory factory = DocumentBuilderFactory.newInstance();
-            // Use the factory to create a builder
-            DocumentBuilder builder = factory.newDocumentBuilder();
-            Document doc = builder.parse(xmlFile);
-            String className = "";
-            String metricName = "";
-            String commandLineName = "";
-            String metricDesc="";
-
-            String uniformFileName = "";
-            String yuleFileName = "";
-            String alnFileSuffix = "";
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/config/DefinedMetricsSet.java

@@ -1,64 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.config;
-
-import java.util.List;
-import java.util.ArrayList;
-import treecmp.metric.Metric;
-
-public class DefinedMetricsSet {
-
-    private static DefinedMetricsSet DMset;
-    private ArrayList<Metric> metricList;
-    
-    protected DefinedMetricsSet()
-    {
-        DMset=null;
-        metricList=new ArrayList<Metric>();
-        metricList.clear();
-
-    }
-    
-    public static DefinedMetricsSet getDefinedMetricsSet()
-    {
-        if(DMset==null)
-        {
-            DMset=new DefinedMetricsSet(); 
-        }
-        return DMset;
-    }
-         
-    public void addMetric(Metric m)
-    {
-
-        /**
-         *
-         * Here can be added a protection against adding the same metric more than onec
-         */
-
-        this.metricList.add(m);
-
-    }
-    public List<Metric> getDefinedMetrics()
-    {
-
-        return this.metricList;
-    }
-
-
-}

### 2018-04-19 Functions/TreeCmp/src/treecmp/config/IOSettings.java

@@ -1,150 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-package treecmp.config;
-
-
-public class IOSettings {
-
-    private static IOSettings IOConf;
-    private String inputFile;
-    private String outputFile;
-    private String sSep;
-    private boolean pruneTrees;
-    private boolean randomComparison;
-    private boolean optMsMcByRf;
-    private boolean genAlignments;
-    private boolean useMsMcFreeLeafSet;;
-    private int iStep;
-    //defalut false
-    private boolean calcCorrelation;
-    private boolean genSummary;
-
-    public boolean isUseMsMcFreeLeafSet() {
-        return useMsMcFreeLeafSet;
-    }
-
-    public void setUseMsMcFreeLeafSet(boolean useMsMcFreeLeafSet) {
-        this.useMsMcFreeLeafSet = useMsMcFreeLeafSet;
-    }
-
-    public boolean isGenSummary() {
-        return genSummary;
-    }
-
-    public void setGenSummary(boolean genSummary) {
-        this.genSummary = genSummary;
-    }
-
-    public boolean isRandomComparison() {
-        return randomComparison;
-    }
-
-    public void setRandomComparison(boolean randomComparison) {
-        this.randomComparison = randomComparison;
-    }
-
-    public boolean isCalcCorrelation() {
-        return calcCorrelation;
-    }
-
-    public void setCalcCorrelation(boolean calcCorrelation) {
-        this.calcCorrelation = calcCorrelation;
-    }
-
-    public String getSSep() {
-        return sSep;
-    }
-
-    public void setSSep(String sSep) {
-        this.sSep = sSep;
-    }
-
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/io/ResultWriter.java

@@ -1,1 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
    using the Matching Split distance and other metrics.
    Copyright (C) 2011,  Damian Bogdanowicz

    This program is free software: you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
    the Free Software Foundation, either version 3 of the License, or
    (at your option) any later version.

    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.

    You should have received a copy of the GNU General Public License
    along with this program.  If not, see <http://www.gnu.org/licenses/>. */

package treecmp.io;

import java.io.File;

import java.io.FileWriter;

import java.io.IOException;

public class ResultWriter {
    public final static String NEW_LINE = System.getProperty("line.separator");
    private String result;
    public boolean writeToFile;
    private String fileName;
    private File file;
    private FileWriter filewriter;

    /** Creates a new instance of ResultWriter */
    public ResultWriter() {

    }

    public ResultWriter(String txt, boolean writeToFile_, String fileName_) {
        this.result = txt;
        this.writeToFile = writeToFile_;
        this.fileName = fileName_;
    }

    public ResultWriter(String txt) {
        this.result = txt;
        this.writeToFile = false;
    }

    public void init() {

        if (this.writeToFile == true) {
            this.file = new File(this.fileName);

            try {
                this.filewriter = new FileWriter(this.file);
            } catch (IOException ex) {
                System.out.print("Error. There is a problem with an output file: "+this.fileName+"\n");
            }
        }
    }

    public void close() {

        if (this.filewriter != null) {
            try {
                this.filewriter.close();
            } catch (IOException ex) {
                ex.printStackTrace();
            }
        }
    }

    public void write() {

        if (this.writeToFile == false) {
            System.out.println(this.result);
        } else {
            try {
                this.filewriter.write(this.result + NEW_LINE);
            } catch (IOException ex) {
                ex.printStackTrace();
            }
        }

    }
    public void write_pure() {
        if (this.writeToFile == false) {
            System.out.println(this.result);
        } else {
            try {
                this.filewriter.write(this.result);
            } catch (IOException ex) {
                ex.printStackTrace();
            }
        }
    }
    public void setText(String text) {
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/io/TreeReader.java

@@ -1,1 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
    using the Matching Split distance and other metrics.
    Copyright (C) 2011,  Damian Bogdanowicz

    This program is free software: you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
    the Free Software Foundation, either version 3 of the License, or
    (at your option) any later version.

    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.

    You should have received a copy of the GNU General Public License
    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
package treecmp.io;

import java.io.*;
import pal.tree.Node;
import pal.tree.ReadTree;
import pal.tree.TreeParseException;
import pal.tree.TreeTool;


public class TreeReader {

    private FileInputStream FIStream;
    private BufferedInputStream BIStream;
    private String FileName;
    private int numberOfTrees;
    private int step;
    private boolean isFirstTree;

    public int getStep() {
        return step;
    }

    public void setStep(int step) {
        this.step = step;
    }

    public int getNumberOfTrees() {
        return numberOfTrees;
    }

    public int getEffectiveNumberOfTrees() {

        if (this.numberOfTrees == 0) {
            return 0;
        } else {
            return 1 + (this.numberOfTrees - 1) / this.step;
        }


    }

    /** Creates a new instance of TreeReader */
    public TreeReader(String filename) {

        this.step = 1;
        this.FileName = filename;

    }

    public int open() {

        isFirstTree = true;
        try {

            this.FIStream = new FileInputStream(this.FileName);
            this.BIStream = new BufferedInputStream(this.FIStream);

        } catch (FileNotFoundException e) {

            System.out.println("Error. Specified file: " + this.FileName + " not found.\n");
            return -1;
        }

        return 0;
    }

    public void close() {

        try {
            //this.FIStream.close();
            this.BIStream.close();
        } catch (IOException ex) {
            ex.printStackTrace();
        }

    }

    private String readUntil(char c) {

        int i;
        char ch;
        //String temp = new String("");
        StringBuilder s = new StringBuilder(8192);
        try {
            do {
                //i = this.FIStream.read();
                i = this.BIStream.read();
                if (i != -1 && i != (int) c) {
                    ch = (char) i;
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/metric/Metric.java

@@ -1,1 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
    using the Matching Split distance and other metrics.
    Copyright (C) 2011,  Damian Bogdanowicz

    This program is free software: you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
    the Free Software Foundation, either version 3 of the License, or
    (at your option) any later version.

    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.

    You should have received a copy of the GNU General Public License
    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
package treecmp.metric;

import pal.tree.*;
import treecmp.common.TreeCmpException;

public interface Metric extends Alignment {

    public double getDistance(Tree t1, Tree t2) throws TreeCmpException;
    
    public String getName();

    public String getCommandLineName();

    public void setCommandLineName(String commandLineName);

    public void setName(String name);

    public String getDescription();

    public void setDescription(String description);

    public void initData();

    public boolean isRooted();

    public boolean isDiffLeafSets();

}

### 2018-04-19 Functions/TreeCmp/src/treecmp/qt/BoolHolder.java

@@ -1,33 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-/** This code comes from the QuartetDist application, see
-    Chris Christiansen, Thomas Mailund, Christian NS Pedersen,
-    Martin Randers and Martin Stig Stissing, "Fast calculation of the quartet distance between trees of arbitrary
-    degrees", Algorithms for Molecular Biology, 1:16, 2006.
-*/
-
-package treecmp.qt;
-/**Container of a boolean*/
-public class BoolHolder {
-  public boolean bool;
-
-  public BoolHolder() {
-    bool = true;
-  }
-  
-}

### 2018-04-19 Functions/TreeCmp/src/treecmp/qt/DistResult.java

@@ -1,86 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-/** This code comes from the QuartetDist application, see
-    Chris Christiansen, Thomas Mailund, Christian NS Pedersen,
-    Martin Randers and Martin Stig Stissing, "Fast calculation of the quartet distance between trees of arbitrary
-    degrees", Algorithms for Molecular Biology, 1:16, 2006.
-*/
-package treecmp.qt;
-public class DistResult {
-  private long qdist, qsim, q1, q2, starbf, bfstar;
-
-  /**Construcs an object representing the results of a quartet distance
-   * calculation between two trees T1 and T2.
-   * @param qdist the quartet distance between prune(T1) and prune(T2),
-   * where prune removes all leaves T1 and T2 do not have in common.
-   * @param dsim the quartets the two trees have in common (topology must
-   * be the same).
-   * @param q1 is the number of quartets that exist in T1 but not in T2
-   * due to leaves present in T1 that are not in T2.
-   * @param q2 vice versa.
-   */
-  
-
-  /**Convenience constructor.*/
-  public DistResult(long qdist, long qsim, long q1, long q2) {
-    this.qdist = qdist;
-    this.qsim = qsim;
-    this.q1 = q1;
-    this.q2 = q2;
-  }
-
-  
-  public String toString() {
-    return ""+qdist;
-  }
-  public void print() {
-    String tmp = "";
-    tmp+="Quartet distance between trees T1 and T2:                               "+(qdist+q1+q2)+"\n";
-    tmp+="Number of quartets with the same topology in both input trees:          "+qsim+"\n";
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/qt/Distance.java

@@ -1,167 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-/** This code comes from the QuartetDist application, see
-    Chris Christiansen, Thomas Mailund, Christian NS Pedersen,
-    Martin Randers and Martin Stig Stissing, "Fast calculation of the quartet distance between trees of arbitrary
-    degrees", Algorithms for Molecular Biology, 1:16, 2006.
-*/
-
-package treecmp.qt;
-import java.util.*;
-/**An abstract class for algorithms calculating distances between
-   trees. Distance methods need only implement the calcDistance
-   method, and can assume that the trees contain the same leaves when
-   this method is called.
-*/
-public abstract class Distance {
-  
-  /**Finds the distance between the two given trees. The distance is
-     calculated on the subset of the trees that contain only the
-     intersection of the leaves. Basically just thins out the trees
-     ofg excess leaves, and then call the calcDistance, which must be
-     implemented for specific metrics.
-     @param t1 the first tree
-     @param t2 the second tree
-     @return the distance 
-  */
-  public long getDistance(Tree t1, Tree t2) {
-    if (!t1.isInitialized())
-      throw new RuntimeException("Tree t1 has not been initialized");
-    if (!t2.isInitialized())
-      throw new RuntimeException("Tree t2 has not been initialized");
-    try {
-      //Check if trees have same leaves;
-      LinkedList delete1 = new LinkedList();
-      LinkedList delete2 = new LinkedList();
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/qt/GeneralN2DQDistDoubleShort.java

@@ -1,206 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-/** This code comes from the QuartetDist application, see
-    Chris Christiansen, Thomas Mailund, Christian NS Pedersen,
-    Martin Randers and Martin Stig Stissing, "Fast calculation of the quartet distance between trees of arbitrary
-    degrees", Algorithms for Molecular Biology, 1:16, 2006.
-*/
-
-/** This is a modified and optimized version of the original code*/
-
-package treecmp.qt;
-import java.util.*;
-
-public class GeneralN2DQDistDoubleShort extends Distance{
-/**Calculates the quartet distance between two general unrooted
-     trees. The computation takes O(n^2d) time for trees that are
-     well resolved but up to quartic time depending on how unresulved the tree is.
-     @param t1 the first tree
-     @param t2 the second tree
-     @return the quartet distance
-  */
-  protected long calcDistance(Tree gt1, Tree gt2) {
-    if (!gt1.isInitialized() || !gt2.isInitialized())
-      throw new RuntimeException("Both trees must be initialized to use this algorithm");
-
-    short[][] intersect = gt1.calcInterSizesShort(gt2);
-
-    double shared = 0, diff = 0;
-    InnerNode[] inodes1 = gt1.getInnerNodes();
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/qt/GeneralN2DQDistLongShort.java

@@ -1,205 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-/** This code comes from the QuartetDist application, see
-    Chris Christiansen, Thomas Mailund, Christian NS Pedersen,
-    Martin Randers and Martin Stig Stissing, "Fast calculation of the quartet distance between trees of arbitrary
-    degrees", Algorithms for Molecular Biology, 1:16, 2006.
-*/
-
-/** This is a modified and optimized version of the original code*/
-
-package treecmp.qt;
-import java.util.*;
-
-public class GeneralN2DQDistLongShort extends Distance{
-/**Calculates the quartet distance between two general unrooted
-     trees. The computation takes O(n^2d) time for trees that are
-     well resolved but up to quartic time depending on how unresulved the tree is.
-     @param t1 the first tree
-     @param t2 the second tree
-     @return the quartet distance
-  */
-  protected long calcDistance(Tree gt1, Tree gt2) {
-    if (!gt1.isInitialized() || !gt2.isInitialized())
-      throw new RuntimeException("Both trees must be initialized to use this algorithm");
-
-    short[][] intersect = gt1.calcInterSizesShort(gt2);
-
-    long shared = 0, diff = 0;
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/qt/Node.java

@@ -1,115 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-/** This code comes from the QuartetDist application, see
-    Chris Christiansen, Thomas Mailund, Christian NS Pedersen,
-    Martin Randers and Martin Stig Stissing, "Fast calculation of the quartet distance between trees of arbitrary
-    degrees", Algorithms for Molecular Biology, 1:16, 2006.
-*/
-
-
-package treecmp.qt;
-import java.util.*;
-/**Class representing a general node*/
-public abstract class Node {
-  
-  /**Accumulates all the directed edges and leaves in the tree this
-     node belongs to. The edges, leaves and inner nodes are
-     accumulated in three given lists. The edge from the calling node is
-     also given, and should be null in the initial call. A Boolean is
-     threaded to carry information whether this tree is binary
-     (i.e. all internal nodes have degree three)
-     @param caller the calling node, null if it is the initial call
-     @param edgeacc the linked list to accumulate the edges in
-     @param leafacc the linked list to accumulate the leaves in
-     @param innernodeacc the linked list to accumulate the inner nodes in
-     @param binary holds whether the tree is binary
-  */
-  public abstract void accumulate(Edge caller, LinkedList edgeacc, LinkedList leafacc,
-				  LinkedList innernodeacc, BoolHolder binary);
-
-  /**Finds all leaves reachable from this node, except the ones that can
-    be reached by going opposite the caller edge
-    @param caller the caller edge
-    @param leafacc the linked list to save the leaves in.*/
-  public abstract void findLeaves(Edge caller, LinkedList leafacc);
-
-  
-  /**Finds a path from this node to the given leaf by searching in all directions except
-     the direction in which the given node lies
-     @param end the leaf we want to find a path to
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/qt/ParseException.java

@@ -1,29 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-/** This code comes from the QuartetDist application, see
-    Chris Christiansen, Thomas Mailund, Christian NS Pedersen,
-    Martin Randers and Martin Stig Stissing, "Fast calculation of the quartet distance between trees of arbitrary
-    degrees", Algorithms for Molecular Biology, 1:16, 2006.
-*/
-
-package treecmp.qt;
-public class ParseException extends Exception {
-  public ParseException(String msg) {
-    super(msg);
-  }
-}

### 2018-04-19 Functions/TreeCmp/src/treecmp/qt/SimpleParser.java

@@ -1,155 +1,0 @@
-/** This file is part of TreeCmp, a tool for comparing phylogenetic trees
-    using the Matching Split distance and other metrics.
-    Copyright (C) 2011,  Damian Bogdanowicz
-
-    This program is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    This program is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with this program.  If not, see <http://www.gnu.org/licenses/>. */
-
-/** This code comes from the QuartetDist application, see
-    Chris Christiansen, Thomas Mailund, Christian NS Pedersen,
-    Martin Randers and Martin Stig Stissing, "Fast calculation of the quartet distance between trees of arbitrary
-    degrees", Algorithms for Molecular Biology, 1:16, 2006.
-*/
-
-
-package treecmp.qt;
-/**A simple newick parser.*/
-public class SimpleParser {
-  private char[] newick;
-  private int position, startpos;
-
-  /**Parses the given string and creates a node representing the tree
-   *@param newickstr the string to parse
-   *@return the node representing the tree
-   *@exception ParseException if the string is not in newick format
-   */
-  public Node parse(String newickstr) throws ParseException {
-    newickstr = newickstr.trim();
-    if (!newickstr.endsWith(";"))
-      newickstr += ";"; //parsing is easier if the string always end with a ';'
-    newick = newickstr.toCharArray();
-    position = 0;
-    Node tree = parseNode();
-    checkEnd();
-    return tree;
-  }
-
-  /**Parses a node*/
-  private Node parseNode() throws ParseException {
-    if (newick[position] != '(')
-      throw new ParseException("Input not in newick format. Missing '(' at character "+position+
-			       "in the input string, counting from the first '('.");
-    Node current = new InnerNode();
-    position++;
-    
-    while (position < newick.length) {
-      if (newick[position] == ')') {
-	return current;
-      }
-      else if (newick[position] == '(') {
[…]

### 2018-04-19 Functions/TreeCmp/src/treecmp/statdata/IMetircDistrbHolder.java

@@ -1,16 +1,0 @@
-/*
- * To change this template, choose Tools | Templates
- * and open the template in the editor.
- */
-
-package treecmp.statdata;
-
-/**
- *
- * @author Damian
- */
-public interface IMetircDistrbHolder {
-           IMetricDistribution getDistribution(int n);
-           int getMaxLeafNum(int n);
-           int getMinLeafNum(int n);
-}

### 2018-04-19 Functions/TreeCmp/src/treecmp/statdata/IMetricDistribution.java

@@ -1,36 +1,0 @@
-/*
- * To change this template, choose Tools | Templates
- * and open the template in the editor.
- */
-
-package treecmp.statdata;
-
-/**
- *
- * @author Damian
- */
-public interface IMetricDistribution {
-
-     public int getLeafNum();
-
-     public double getAvg();
-     public double getStd();
-
-     public double getMin();
-     public double getMax();
-
-     public double get02Quantile();
-     public double get05Quantile();
-     public double get10Quantile();
-     public double get20Quantile();
-     public double get30Quantile();
-     public double get40Quantile();
-     public double get50Quantile();
-     public double get60Quantile();
-     public double get70Quantile();
-     public double get80Quantile();
-     public double get90Quantile();
-     public double get95Quantile();
-     public double get97Quantile();
-
-}

### 2018-04-19 Functions/TreeCmp/src/treecmp/statdata/MetircDistrbHolder.java

@@ -1,49 +1,0 @@
-/*
- * To change this template, choose Tools | Templates
- * and open the template in the editor.
- */
-
-package treecmp.statdata;
-
-import java.util.HashMap;
-import java.util.Map;
-
-/**
- *
- * @author Damian
- */
-public class MetircDistrbHolder implements IMetircDistrbHolder{
-
-
-    //protected IMetricDistribution distributions[];
-    private Map<Integer,IMetricDistribution> distMap;
-    private int minLeafNum;
-    private int maxLeafNum;
-
-    public MetircDistrbHolder(){
-        distMap  = new HashMap<Integer,IMetricDistribution>(1000);
-        minLeafNum = Integer.MAX_VALUE;
-        maxLeafNum = -1;
-    }
-
-    public IMetricDistribution getDistribution(int n) {
-        return distMap.get(n);
-    }
-
-    public int getMaxLeafNum(int n) {
-        return maxLeafNum;
-    }
-
-    public int getMinLeafNum(int n) {
-        return minLeafNum;
-    }
-
-    public void insertDistribution(IMetricDistribution distrb){
-        int leafNum = distrb.getLeafNum();
-        distMap.put(leafNum, distrb);
-        if(leafNum < minLeafNum)
-            minLeafNum = leafNum;
-        if(leafNum > maxLeafNum)
-            maxLeafNum = leafNum;
-    }
-}

### 2018-04-19 Functions/shells/CC_TreeCmp.sh
Added data simulation part to the README

@@ -1,7 +1,7 @@
 ##########################
 # Script for using TreeCmp between trees for Character Correlation
 ##########################
-#SYNTAX: TreeCmp -c <chain> -o <output> -f <treeformat> -d <draws> -r <rooted>
+#SYNTAX: sh CC_TreeCmp -c <chain> -r <ref> -o <output> -f <treeformat> -d <draws> -r <rooted>
 #with:
 #-c <chain> the name of the chain
 #-r <ref> the reference tree number (1 = normal, 2 = maximum, 3 = minimum, 4 = null)

### 2018-04-19 Analysis/Character_difference/02-Modifying_matrices.Rmd
Added missing links

--- 2017-09-19 Analysis/Character_difference/Modifying_matrices.Rmd
@@
 ## Simulate a matrix
-matrix_norm <- sim.morpho(tree, ncharacters, states = c(0.85,0.15),
+matrix_norm <- sim.morpho(tree, ncharacters, states = c(0.85, 0.15),
     model = "mixed", rates = c(rgamma, rate = 100, shape = 3),
@@
 while(check.morpho(matrix_norm)[2,] < 0.26) {
-    matrix_norm <- sim.morpho(tree, ncharacters, states = c(0.85,0.15),
+    matrix_norm <- sim.morpho(tree, ncharacters, states = c(0.85, 0.15),
         model = "mixed", rates = c(rgamma, rate = 100, shape = 3),

### 2018-04-19 Analysis/02-EffectCorrelationFullResults.Rmd
Uploaded analysis

@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, warning = FALSE}
 library(knitr)
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 library(xtable)
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, warning = FALSE}
 kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Full_Tab_signif", path = "../Writing/Tables/")
@@
 
-```{r, warning = FALSE}
+```{r, warning = FALSE, warning = FALSE}
 ## Summarise into a table
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, warning = FALSE}
 ## Caption
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 make.xtable(test_table, caption = cap_pool_scen_test, label = "Full_Tab_pooledscenarios_test", path = "../Writing/Tables/")
@@
 
-```{r, warning = FALSE}
+```{r, warning = FALSE, warning = FALSE}
 ## Pooling the data per character
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, warning = FALSE}
 ## Setting the table captions
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 make.xtable(characters, caption = captions_pooled_characters, label = "Full_Tab_pooledcharacters", path = "../Writing/Tables/")
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Full_Tab_signif")
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 make.xtable(test_table, caption = cap_pool_char_test, label = "Full_Tab_pooledscharacters_test", path = "../Writing/Tables/")
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 make.xtable(taxa, caption = captions_pooled_taxa, label = "Full_Tab_pooledtaxa", path = "../Writing/Tables/")
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 make.xtable(test_table, caption = cap_pool_tax_test, label = "Full_Tab_pooledstaxa_test", path = "../Writing/Tables/")
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 make.xtable(methods, caption = captions_pooled_methods, label = "Full_Tab_pooledmethod", path = "../Writing/Tables/")
@@
 
-```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide'}
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
 make.xtable(test_table, caption = cap_pool_met_test, label = "Full_Tab_pooledsmethods_test", path = "../Writing/Tables/")
@@
 
-```{r}
+```{r, warning = FALSE}
 ## Summarising the data
@@
 
-```{r, echo = FALSE, eval = TRUE, print = FALSE, result = 'hide'}
+```{r, echo = FALSE, eval = TRUE, print = FALSE, result = 'hide', warning = FALSE}
 ## Labels
@@
 
-```{r}
+```{r, warning = FALSE}
 ## Applying an lm to all the data with error
@@
 
-```{r}
+```{r, warning = FALSE}
 ## Running the pairwise Bhattacharrya Coefficient analysis:
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, warning = FALSE}
 ## Tables
@@
 
-```{r, warning = FALSE}
+```{r, warning = FALSE, warning = FALSE}
 ## Running the pairwise Wilcox tests:

### 2018-04-19 Writing/draft.tex
added reproducibility

@@
 
-The resulting full simulation was 4TB @@@ big so is not shared here (though the parameters are).
-However, the resulting consensus trees on which the topological differences are calculated are available at @@@.
+The resulting full simulation was 3.5TB big so is not shared here (though the parameters are).
+However, the resulting consensus trees on which the topological differences are calculated are available at \url{https://figshare.com/s/7a8fde8eaa39a3d3cf56}.
 
@@
 \centering
-   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
+   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf}
 \caption{Effect of character difference on recovering the ``normal'' topology.
@@
 The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 and 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-%MB: The above sentence has too many "for"s. It's too long and needs to be split to be intelligible.
-%TG: fix attempt
 The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
@@
 The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
-When characters are correlated, one could expect the matrices to convey a really strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
+When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
 Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
@@
 
-Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distribution that is always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
+Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distributions that are always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
 This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
@@
 The difference in the ability of each method to recover the ``correct'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have raised issues with these investigations \citealt{spencerefficacy2013,goloboff2017weighted}).
-In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e.
-those with in a $CI<0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour Parsimony analysis.
+In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[i.e.
+those with in a $CI<0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
 
-%MB: I'm not sure I'm convinced by this paragraph. Perhaps we should have a chat about it.
-%TG: I've rephrased the parargaph below.
-This is the idea (a bit arms waving, I agree): for a given topology, when there is high homoplasy, parsimony will give the tree a high score (may changes per branches - even though branch length is not a parsimony parameter) and probably discard the topology.
-For the same topologyn and high homoplasy, Bayesian will stretch the branches (accounting for the many changes per branches) which will decrease the likelihood score but less fast than MP (because branch length is only *one* of the parameters accounting for the likelihood score).
-Also, in Bayesian, an decrease in likelihood score does not necessary mean that the tree will be discarded.
 The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
@@
 \section{Data availability, repeatability and reproducibility}
-The consensus trees are available on figshare at \url{@@@}.
+The consensus trees are available on figshare at \url{https://figshare.com/s/7a8fde8eaa39a3d3cf56}.
 The simulations are fully replicable following the explanations at \url{https://github.com/TGuillerme/CharactersCorrelation}.

### 2018-06-04 Writing/SystematicBiology/Tab_pooledscharacters_test.tex
Cyst bill submission

--- 2018-03-27 Writing/Tables/Tab_pooledscharacters_test.tex
@@
 \end{tabular}
-\caption{Difference between the pooled number of characters.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)}
+\caption{Differences between the pooled number of characters.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
 \label{Tab_pooledscharacters_test}

### 2018-06-04 Writing/SystematicBiology/Tab_pooledsmethods_test.tex
Cyst bill submission

--- 2018-03-27 Writing/Tables/Tab_pooledscharacters_test.tex
@@
 % latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Tue Mar 27 14:03:13 2018
+% Tue Mar 27 14:03:14 2018
 \begin{table}[ht]
@@
   \hline
-RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
-   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
-   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
-  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
-   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
-   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
+RF & bayesian:parsimony & 0.891 & 579437.500 & \textbf{0} \\ 
+  Tr & bayesian:parsimony & \textbf{0.984} & 470621.500 & 0.084 \\ 
    \hline
 \end{tabular}
-\caption{Difference between the pooled number of characters.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)}
-\label{Tab_pooledscharacters_test}
+\caption{Differences between the pooled methods.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
+\label{Tab_pooledsmethods_test}
 \end{table}

### 2018-06-04 Writing/SystematicBiology/Tab_pooledstaxa_test.tex
Cyst bill submission

--- 2018-03-27 Writing/Tables/Tab_pooledscharacters_test.tex
@@
   \hline
-RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
-   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
-   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
-  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
-   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
-   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
+RF & t25:t75 & \textbf{0.976} & 218421.000 & \textbf{0.012} \\ 
+   & t25:t150 & \textbf{0.988} & 220529.000 & \textbf{0.004} \\ 
+   & t75:t150 & \textbf{0.99} & 201037.000 & 1 \\ 
+  Tr & t25:t75 & \textbf{0.976} & 233282.000 & \textbf{0} \\ 
+   & t25:t150 & \textbf{0.978} & 227288.000 & \textbf{0} \\ 
+   & t75:t150 & \textbf{0.992} & 194201.000 & 1 \\ 
    \hline
 \end{tabular}
-\caption{Difference between the pooled number of characters.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)}
-\label{Tab_pooledscharacters_test}
+\caption{Differences between the pooled number of taxa.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
+\label{Tab_pooledstaxa_test}
 \end{table}

### 2018-06-04 Writing/SystematicBiology/draft.tex
Cyst bill submission

--- 2018-04-19 Writing/draft.tex
@@
 
-\begin{figure}[!htbp]
+\begin{figure}
 \centering
-   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
-\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using MP and BPP methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
+   \includegraphics[width=0.9\textwidth]{outline.eps}
+\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using maximum parsimony and Bayesian inference methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
 $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
@@
 \noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
-$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
+$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
 Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
@@
 
-\begin{figure}[!htbp]
+\begin{figure}
 \centering
-   \includegraphics[width=1\textwidth]{Figures/Modif_matrix.pdf}
+   \includegraphics[width=0.9\textwidth]{Modif_matrix.eps}
 \caption{Example illustration of the protocol for modifying matrices.
@@
 \subsection{Inferring topologies}
-We inferred the topologies with both BPP and MP using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
+We inferred the topologies with both Bayesian inference and maximum parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
 For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).]
+We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 1 and 2).
 Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
@@
 Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one.
-The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary material 3).
+The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}).
 Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
@@
 
-\begin{figure}[!htbp]
+\begin{figure}
 \centering
-   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
+   \includegraphics[width=0.9\textwidth]{RF_results_best.eps}
 \caption{\small{Effect of character difference on recovering the ``normal'' topology.
@@
 
-\begin{figure}[!htbp]
+\begin{figure}
 \centering
-   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf}
+   \includegraphics[width=0.9\textwidth]{Tr_results_best.eps}
 \caption{Effect of character difference on recovering the ``normal'' topology.
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see Supplementary material 3 for the full summary statistics).
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see supplementary material 3 for the full summary statistics).
 There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
 
-\input{Tables/Tab_pooledscenarios_test.tex}
+\begin{table}[ht]
+\centering
+\begin{tabular}{llr|rl}
+  \hline
+metric & test & bhatt.coeff & statistic & p.value \\ 
+  \hline
+RF & maxi:mini & 0.573 & 356436.000 & \textbf{0} \\ 
+   & maxi:rand & 0.873 & 287225.000 & \textbf{0} \\ 
+   & mini:rand & 0.856 & 95841.500 & \textbf{0} \\ 
+  Tr & maxi:mini & 0.614 & 358800.000 & \textbf{0} \\ 
+   & maxi:rand & 0.908 & 288223.500 & \textbf{0} \\ 
+   & mini:rand & 0.858 & 98507.500 & \textbf{0} \\ 
+   \hline
+\end{tabular}
+\caption{Differences between the pooled scenarios.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
+\label{Tab_pooledscenarios_test}
+\end{table}
 
@@
 
-\input{Tables/Tab_pooledscharacters_test.tex}
+\begin{table}[ht]
+\centering
+\begin{tabular}{llr|rl}
+  \hline
+metric & test & bhatt.coeff & statistic & p.value \\ 
+  \hline
+RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
+   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
+   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
+  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
+   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
+   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
+   \hline
+\end{tabular}
+\caption{Differences between the pooled number of characters.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
+\label{Tab_pooledscharacters_test}
+\end{table}
 
@@
 
-\input{Tables/Tab_pooledstaxa_test.tex}
+\begin{table}[ht]
+\centering
+\begin{tabular}{llr|rl}
+  \hline
+metric & test & bhatt.coeff & statistic & p.value \\ 
+  \hline
+RF & t25:t75 & \textbf{0.976} & 218421.000 & \textbf{0.012} \\ 
+   & t25:t150 & \textbf{0.988} & 220529.000 & \textbf{0.004} \\ 
+   & t75:t150 & \textbf{0.99} & 201037.000 & 1 \\ 
+  Tr & t25:t75 & \textbf{0.976} & 233282.000 & \textbf{0} \\ 
+   & t25:t150 & \textbf{0.978} & 227288.000 & \textbf{0} \\ 
+   & t75:t150 & \textbf{0.992} & 194201.000 & 1 \\ 
+   \hline
+\end{tabular}
+\caption{Differences between the pooled number of taxa.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
+\label{Tab_pooledstaxa_test}
+\end{table}
 
@@
 
-\input{Tables/Tab_pooledsmethods_test.tex}
+\begin{table}[ht]
+\centering
+\begin{tabular}{llr|rl}
+  \hline
+metric & test & bhatt.coeff & statistic & p.value \\ 
+  \hline
+RF & bayesian:parsimony & 0.891 & 579437.500 & \textbf{0} \\ 
+  Tr & bayesian:parsimony & \textbf{0.984} & 470621.500 & 0.084 \\ 
+   \hline
+\end{tabular}
+\caption{Differences between the pooled methods.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
+\label{Tab_pooledsmethods_test}
+\end{table}
 
@@
 As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
-The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
+The character difference metric can be seen as the inverse of character correlation (see Methods): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
@@
 Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
-Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary materials @@@).
+Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary material 3, Figs 3, 4 and 5).
 
@@
 Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
-This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.
+This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.b
 
@@
 
-%TG: insisting on the shittiness of generatin correlation
 Second, measuring and modifying character correlation is difficult. 
@@
 This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
-%MB: We also don't really explore character congruence, that correlations can be ranked and spread over multiple internodes.
-%MB: We didn't look at ranking and we didn't look at competing correlations. There's a lot more that could be done here so, we might talk about that.
-%TG: not sure what you mean by the ranking stuff but it could definitely fit here.
 However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
@@
 \subsection{Potential applications}
-%TG: explaining that our "scenarios" are biased bootstraps
 Effectively, our simulation protocol bootstraps our data ``with bias''.
@@
 
-%TG: explaining that we could use it for "clever" bootstrapping in big matrices
 It is noteworthy to point that in rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
@@
 Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g.
-for characters hierarchy][]{deLaet2015,BrazeauNA} already exist.
+for characters hierarchy][]{DeLaet2015,BrazeauNA} already exist.
 However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
-%MB: I have no idea what thae above sentence means.
-%TG: Modified. Is that clearer?
 Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
@@
 
-% Suggested reviewers:
-%- Liliana Davalos for MPE / Joe O'Reilly, April Wright
-
 \bibliographystyle{sysbio}

### 2018-06-04 Writing/Tables/Tab_pooledscenarios_test.tex
compiled tables

@@
 \end{tabular}
-\caption{Difference between the pooled scenarios.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction)}
+\caption{Differences between the pooled scenarios.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
 \label{Tab_pooledscenarios_test}

### 2018-06-04 Writing/draft_v1.tex
draftv1 backup

--- 2018-06-04 Writing/SystematicBiology/draft.tex
@@
 
-\begin{figure}
+\begin{figure}[!htbp]
 \centering
-   \includegraphics[width=0.9\textwidth]{outline.eps}
-\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using maximum parsimony and Bayesian inference methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
+   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
+\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using MP and BPP methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
 $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
@@
 
-\begin{figure}
+\begin{figure}[!htbp]
 \centering
-   \includegraphics[width=0.9\textwidth]{Modif_matrix.eps}
+   \includegraphics[width=1\textwidth]{Figures/Modif_matrix.pdf}
 \caption{Example illustration of the protocol for modifying matrices.
@@
 \subsection{Inferring topologies}
-We inferred the topologies with both Bayesian inference and maximum parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
+We inferred the topologies with both BPP and MP using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
 For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
@@
 
-\begin{figure}
+\begin{figure}[!htbp]
 \centering
-   \includegraphics[width=0.9\textwidth]{RF_results_best.eps}
+   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
 \caption{\small{Effect of character difference on recovering the ``normal'' topology.
@@
 
-\begin{figure}
+\begin{figure}[!htbp]
 \centering
-   \includegraphics[width=0.9\textwidth]{Tr_results_best.eps}
+   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf}
 \caption{Effect of character difference on recovering the ``normal'' topology.
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{llr|rl}
-  \hline
-metric & test & bhatt.coeff & statistic & p.value \\ 
-  \hline
-RF & maxi:mini & 0.573 & 356436.000 & \textbf{0} \\ 
-   & maxi:rand & 0.873 & 287225.000 & \textbf{0} \\ 
-   & mini:rand & 0.856 & 95841.500 & \textbf{0} \\ 
-  Tr & maxi:mini & 0.614 & 358800.000 & \textbf{0} \\ 
-   & maxi:rand & 0.908 & 288223.500 & \textbf{0} \\ 
-   & mini:rand & 0.858 & 98507.500 & \textbf{0} \\ 
-   \hline
-\end{tabular}
-\caption{Differences between the pooled scenarios.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
-\label{Tab_pooledscenarios_test}
-\end{table}
+\input{Tables/Tab_pooledscenarios_test.tex}
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{llr|rl}
-  \hline
-metric & test & bhatt.coeff & statistic & p.value \\ 
-  \hline
-RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
-   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
-   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
-  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
-   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
-   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
-   \hline
-\end{tabular}
-\caption{Differences between the pooled number of characters.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
-\label{Tab_pooledscharacters_test}
-\end{table}
+\input{Tables/Tab_pooledscharacters_test.tex}
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{llr|rl}
-  \hline
-metric & test & bhatt.coeff & statistic & p.value \\ 
-  \hline
-RF & t25:t75 & \textbf{0.976} & 218421.000 & \textbf{0.012} \\ 
-   & t25:t150 & \textbf{0.988} & 220529.000 & \textbf{0.004} \\ 
-   & t75:t150 & \textbf{0.99} & 201037.000 & 1 \\ 
-  Tr & t25:t75 & \textbf{0.976} & 233282.000 & \textbf{0} \\ 
-   & t25:t150 & \textbf{0.978} & 227288.000 & \textbf{0} \\ 
-   & t75:t150 & \textbf{0.992} & 194201.000 & 1 \\ 
-   \hline
-\end{tabular}
-\caption{Differences between the pooled number of taxa.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
-\label{Tab_pooledstaxa_test}
-\end{table}
+\input{Tables/Tab_pooledstaxa_test.tex}
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{llr|rl}
-  \hline
-metric & test & bhatt.coeff & statistic & p.value \\ 
-  \hline
-RF & bayesian:parsimony & 0.891 & 579437.500 & \textbf{0} \\ 
-  Tr & bayesian:parsimony & \textbf{0.984} & 470621.500 & 0.084 \\ 
-   \hline
-\end{tabular}
-\caption{Differences between the pooled methods.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).}
-\label{Tab_pooledsmethods_test}
-\end{table}
+\input{Tables/Tab_pooledsmethods_test.tex}
 
@@
 As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
-The character difference metric can be seen as the inverse of character correlation (see Methods): a high character difference approximates a low level of character correlation and vice versa.
+The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
@@
 Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g.
-for characters hierarchy][]{DeLaet2015,BrazeauNA} already exist.
+for characters hierarchy][]{deLaet2015,BrazeauNA} already exist.
 However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.

### 2018-06-04 Writing/draft_highlighted.tex
new version draft

--- 2018-06-04 Writing/draft_v1.tex
@@
 \usepackage{pdflscape}
+
+\usepackage{color,soul}
 
@@
 Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
-Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with good practices, phylogenetic software do not distinguish between any of them.
+Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with \hl{careful character coding}, phylogenetic software do not distinguish between any of them.
 
@@
 
-    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving due to selection, likely related to functional links between two traits that help serve an overall lifestyle trait.
+    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving due to selection, likely related to functional links between two traits that help serve an overall lifestyle trait \citep{ClarkeMiddleton2008}.
     Unlike the case of intra-organismal dependence, there need not be an intrinsic constraint that causes these traits to covary.
@@
     This occurs when a number of multistate characters include two variable feature states (e.g. large, small; red, blue etc.) in conjunction with absence.
+    \hl{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.}
     It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
@@
 Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
+\hl{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.}
 
@@
 \begin{itemize}
-    \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
+    \item The \hl{``morphological HKY-binary'' model} \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
     This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use an M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
@@
 Dollo characters;][]{Dollo,wright2015came}.
-    Recently however, \cite{Wright01072016} have shown that an equal rate transition is still the most present in empirical data.
+    However, \cite{Wright01072016} \hl{have shown that equal transition rate is a reasonable assumption for many empirical datasets.}
 \end{itemize}
 
-\noindent For each character, both models (morphological HKY-binary or M$k$) where chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
+\noindent For each character, both models (``morphological HKY-binary'' or M$k$) where chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
 These low evolutionary rate values allowed reduction in the number of homoplasic character changes, thus reinforcing the phylogenetic information in the matrices.
@@
 To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{OReilly20160081}.
+\hl{Note that both models (``morphological HKY-binary' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
+However, the ``morphological HKY model'' is not the one \textit{directly}estimated in the phylogenetic software used in this protocol (MrBayes, see below).}
 
@@
 
-The resulting full simulation was 3.5TB big so is not shared here (though the parameters are).
+The resulting full simulation was 3.5TB \hl{in size} so is not shared here (though the parameters are).
 However, the resulting consensus trees on which the topological differences are calculated are available at \url{https://figshare.com/s/7a8fde8eaa39a3d3cf56}.

### 2018-06-04 Writing/response_to_reviewers.tex
Response to reviewers INIT

\documentclass[12pt,letterpaper]{article}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\pagenumbering{arabic}

%Pagination style and stuff
%\linespread{2}

\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\setlength\parindent{0pt}

\begin{document}

\textbf{RE: USYB-2018-099}\\
\bigskip
Dear Dr Jermiin,\\
\bigskip

We am very grateful to the three referees for their helpful and constructive comments, which we believe have helped me to significantly improve the manuscript.
We have taken all of their comments on board, and respond to their points in details below.

% For improving clarity in this document, reviewers' comments are displayed in blue, my responses in normal text and the changes to the manuscript in italic.
% Additionally, I've uploaded two versions of the revised manuscript: both have the exact same content but one has all the changes and additions to the main text highlighted in yellow to help with reviewing.

\section{Editors' comments}

\begin{enumerate}

\item{\textcolor{blue}{Along the lines set out by Reviewer 2, it would be wise to ensure that your metric is characterised as thoroughly as possible. In this regard, does it meet the following five conditions?}}

\begin{enumerate}
\item{\textcolor{blue}{Distinctiveness}}
\item{\textcolor{blue}{Non-negativity}}
\item{\textcolor{blue}{Symmetry}}
\item{\textcolor{blue}{Triangle inequality}}
\item{\textcolor{blue}{Four-point condition}}
\end{enumerate}

\item{\textcolor{blue}{If so, then it is actually an evolutionary distance, and it can be used as such; if not, then it is another type of distance. Establishing the type of distance may be overkill, but the characterization of a metric helps to determine what it can be used for, and what it cannot be used for.}}

\end{enumerate}

% ~~~~~~~~~~~~~~~~~~~~~~~~
%
% REVIEWER 1
%
% ~~~~~~~~~~~~~~~~~~~~~~~~

\section{Reviewer 1}

\subsection{Major suggestions}

\begin{enumerate}

\item{\textcolor{blue}{I have some concerns about the approach, however. I was somewhat confused by the CD metric. When quantifying n, how is it decided how many taxa have "comparable" characters? If you simulated the character for all taxa, do they not all have this character (i.e. n always equals the dataset size)? But more importantly, will two characters not be more similar if they share the same number of character states, independent
of if they are telling you the same evolutionary history? It seems to me that whether or not two characters share a state space is not the relevant axis along which to measure correlation.}}

\item{\textcolor{blue}{I also found the explanations confusing for what proportion of a matrix was repeat sampled. For example, in a matrix that you call maximised, n characters are replaced. But how many is that really? What is the range? I had a hard time putting a context to this and deciding how well I feel these simulations approximate reality.  Without knowing this, I was having difficulty understanding why the statement "when character correlation was high (low character differences), the topology was always the furthest away from the
``normal'' topology." should be true - does this simply mean that the same signal is being repeated over and over again, but that signal is low information relative to the topology? To my mind, highly correlated datasets would yield congruent trees, but that these trees may be discordant with the true tree, if the resampled characters were not representative of the true signal, or highly concordant with the true tree if they were. Perhaps this not being observed is a result of not performing comparisons to the true tree?}}

\item{\textcolor{blue}{I thought the reproducible science aspect was cool, and well done. Some aspects of the literature review and context for the problem need to be strengthened, particularly given that there is good work that has been done on this problem in a molecular model-based framework (see below). I've tried to note locations where I was confused by the prose, but this manuscript could certainly use a very close reading before resubmission.}}

\end{enumerate}

\subsection{Specific suggestions}

\begin{enumerate}

\item{\textcolor{blue}{"Phylogenetic analysis algorithms require the assumption of character independence - a condition generally acknowledged to be violated by morphological data."\\
I don't love this sentence because it emphasizes morphology as distinct, when we recognize non-independence as a problem in basically all data types (see Huelsenbeck and Nielsen, 1996, Effect of Nonindependent Substitution on Phylogenetic Accuracy and the references therein). I do think you should point to a citation here; Davalos et al. 2014 (Integrating Incomplete Fossils by Isolating Conflicting Signal in Saturated and Non-Independent Morphological Characters) is a good one.}}

\item{\textcolor{blue}{"good practices".\\
Change to "careful character coding"}}

We've changed the sentence following the reviewer's suggestions lines @@@.

\item{\textcolor{blue}{"The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data".\\
Evidence for this?}}

%TG: Find cite.

\item{\textcolor{blue}{"Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the very nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other."\\
I don't really understand what this sentence is adding. It feels repetitious in light of previous paragraphs, but also does little to establish why morphological data should be especially susceptible to this. I think there are parts of the genome where we would expect to find as much or more correlation than in morphological characters - the stem domains of rRNA genes, for example. I would expect the average character correlation to be weaker in DNA data, or at least more spread out in a way that makes it harder to detect.}}

%TG: Remove the sentence?

\item{\textcolor{blue}{"Evolutionary dependence: this is the result of sets of characters co-evolving due to selection, likely related to functional links between two traits that help serve an overall lifestyle trait."\\
Seems appropriate to cite Clarke and Middleton (2008) here.}}

We added the suggested citation (lines @@@).

\item{\textcolor{blue}{Coding dependence:\\
This section could use a figure illustrating these types of dependence, and what they mean. For example: "For instance, coding for the same absence in different characters creates state transformations associated with the loss or gain of a particular character." is a fairly confusing sentence. I took this to mean that if you are coding, for example, tail characters, you might atomize the tail into a suite of characters, all of which are absent in organisms with no tail. But I'm not sure I've understood your meaning correctly.}}

%TG: add a figure for the difference character dependencies?

We've clarified the sentence with an example as follows:
\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.} lines @@@

\item{\textcolor{blue}{"Correlation between characters detected by the software:"\\
I really struggled with this section. It seems like the problem here is still a coding problem - that the authors of these four characters have chosen four highly dependent characters to build the matrix. Even scoring by hand under parsimony would have problems here. I think what you need to decide here is what you want to communicate in this paragraph. Is it that software is not currently equipped to handle this issue?}}

%TG: probably insist on what the reviewer says: that the software cannot handle this.

\item{\textcolor{blue}{"To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different levels of correlation between some characters (Fig.1 - note that each step is described in more details below):"\\
Fig. 1 is the results, but this text implies it is a workflow diagram.}}

We doubled checked the figures links and made sure that this mention of Fig.1 now properly points to the workflow figure.

\item{\textcolor{blue}{"Additionally, we only consider differences for taxa with shared information (i.e. a Gower distance; Gower, 1971)."\\
I'm not sure what this means. Does this mean that you did not include taxa with missing information when calculating per-character distances?}}

We've clarified this sentence by adding the following example:
\textit{For example, for two characters X and Y for four taxa, if character X is {0,0,?,1} for the first, second, third and fourth taxa respectively and character Y is {0,0,1,1}, only the first, second and fourth taxa are compared for this character.}

\item{\textcolor{blue}{"This procedure was used to treat all the characters are unordered with no assumption on the meaning of the character state"\\
Why is this important? The text doesn't mention testing anything related to orderedness, or specifying anything related to orderedness in MrBayes.}}

%TG: answering that by saying this is a normal neutral procedure?

\item{\textcolor{blue}{"The morphological HKY model"\\
Is not real. This is just the HKY model. Goloboff (2017) would dispute that this model does not favor a Bayesian method, since it is still a Markovian model. That O'Reilly et al. used this model to generate binary data does not mean one has to - you could translate each of the four nucleotides to its own discrete character. This would probably get even further from the Mk model assumptions than collapsing the multistate character produced via the HKY model to a binary character.}}

We've enquoted the ``morphological HKY model'' and specified that it is still a Markovian model:
\textit{Note that both models (``morphological HKY'' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
However, the ``morphological HKY model'' is not the one} directly \textit{estimated in the phylogenetic software used in this protocol (MrBayes, see below).}(lines @@@).

%TG: maybe add a bit more?

\item{\textcolor{blue}{"Recently however, Wright et al. (2016) have shown that an equal rate transition is still the most present in empirical data."\\
Of the models tested - it's plausible there are other explainers of the data. I would change this to "have shown that equal transition rate is a reasonable assumption for many empirical datasets"}}

We've changed the sentence as suggest by the reviewer (lines @@@).

\item{\textcolor{blue}{"The resulting full simulation was 3.5TB big so is not shared here (though the 342 parameters are)."\\
3.5 TB in size.}}

We fixed this typo.

\item{\textcolor{blue}{Most of the time when RF is normalized, it's done so so that low scores are better - i.e. a low RF score indicates less distance to the true tree. I was very confused looking at the figures, until I realized this was not how normalizing had been performed.  "However, the effect of these sources of correlation was out of the scope of this study and would have required a posteriori changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest."\\
I'm not sure I follow this.}}

\item{\textcolor{blue}{"However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree. "\\
I understand why you did this, but I think you should also compare normal to true. That will give us some idea of the scope of how close we can get to true, under these conditions, and will help contextualize the results.}}

\end{enumerate}

% ~~~~~~~~~~~~~~~~~~~~~~~~
%
% REVIEWER 2
%
% ~~~~~~~~~~~~~~~~~~~~~~~~

\section{Reviewer 2}

\subsection{Major suggestions}

\begin{enumerate}

\item{\textcolor{blue}{1. To be a metric, there are several properties that need to be satisfied. The proposed function
is positively valued and symmetric (copied below from line 178): [...] But, the following metric property does not hold for all x; y:}}

\item{\textcolor{blue}{While some details of the simulation study are included, some important ones are missing. In particular, in the design of the study, how did you avoid or mitigate errors introduced by the programs simulating data? All of the programs use random-number generators and have their own biases. How have you accounted for this in the study of design to conclude that your results are due to signal, and not bias from the programs used?}}

\item{\textcolor{blue}{How did you ``randomly''randomly modify your matrices? Did you sample under some distribution? If so, which one?}}

\item{\textcolor{blue}{It is not obvious why generated character sequences are representative of biological sequences or sample the space overall. Were the parameters chosen for the HKY model to fit a particular evolutionary scenario? Are they meant to capture all of the space?}}

\item{\textcolor{blue}{Similarly, why did you choose 0.85 for the Mk model? Does it represent some important set of character sequences? How does it represent the space in general?}}

\item{\textcolor{blue}{There are statements that seem contradictory, like (line 335): ``Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability > 0:95 and different if they had an overlap probability > 0:05.''}}

\item{\textcolor{blue}{It is not clear how the conclusions were reached. For example, line 381-384: ``Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table 5 - median NTSRF of 0:828 and 0:679 respectively) but not in terms of individual taxon placements (Table 5 - median NTSTr of 0:738 and 0:601 respectively).'' Why is the 0:828 - 0:679 = 0:149 significant for this test and 0:738 - 0:601 = 0:137 is not?}}

\end{enumerate}

% ~~~~~~~~~~~~~~~~~~~~~~~~
%
% REVIEWER 3
%
% ~~~~~~~~~~~~~~~~~~~~~~~~

\section{Reviewer 3}

\subsection{Major suggestions}

\begin{enumerate}

\item{\textcolor{blue}{I feel that the manuscript is certainly worth publishing in Systematic Biology, although I worry that the simulation procedure used by the authors limits their ability to thoroughly test of the effect of between character correlation on the accuracy of estimating phylogenetic trees. The authors note this themselves in the discussion, but I feel more should be made of this point. If the simulations utilised an explicit distribution of character covariance then I think that interpretation of the results would have been more straightforward.}}

\item{\textcolor{blue}{Another worry I have is that the implied covariance at the tips does not necessarily reflect covariation over the full phylogeny. It is conceivable that two characters can appear to have near perfect correlation at the tips, but arrived at these states through completely independent pathways, particularly when simulating with a continuous time Markov chain where unobserved transitions to a new state and back to the original state are possible.}}

\item{\textcolor{blue}{It seems that the results reported by the authors are inherently driven by the phylogenetic signal present in their ``bootstrapped'' datasets. This effect would be present in any approach to the simulation of covarying characters, but as the authors do not explicitly simulate covariance but instead rely on a proxy they cannot be certain that their results are driven by true character covariation or purely a restructuring of phylogenetic signal by resampling. In a simulation framework in which covariance is actively enforced and controlled it would be possible to assess other aspects of character covariance that seem important for empirical datasets, such as correlation within morphological compartments. I think without tests of these sort we cannot be completely sure of the effect of character covariance on topology estimates. Despite this, the authors present an overview of the effect of character covariance at large, and I certainly think that this is a worthwhile contribution.}}

\item{\textcolor{blue}{A discussion of the ``effective'' number of characters might also be warranted. When maximising character covariance this likely requires the resampling of the same handful of characters multiple times, each of which provides the exact same phylogenetic signal. The redundancy of these characters seems likely to influence the accuracy of estimates. The choice of characters that are sampled by the bootstrap procedure will therefore constrain the accuracy of topology estimates. If the resampled matrix consists of a handful of characters with high consistency with the true tree, sampled many times over, then the tree estimate will be good (for at least part of the tree) and covariance will be high. Similarly, if the handful of regularly resampled characters are mainly inconsistent (with the true tree) then tree estimates will be poor but the implied covariance will also be high. I think this is evident from the reduced performance in larger matrices, where there will be a significant reduction in phylogenetic signal when attempting to maximise covariance.}}

\item{\textcolor{blue}{Overall, I think that the authors provide an overview of the effect of covariance on the estimation of phylogeny, but without an explicit simulation of patterns of covariation we do not yet have a definitive understanding of the effects of between character covariance.}}

\end{enumerate}

\subsection{Minor revisions:}

\begin{enumerate}

\item{\textcolor{blue}{2:2 – Phylogenetic algorithms don't necessarily require independence, they simply assume it.}}

\item{\textcolor{blue}{3:25 – I think that resurgence should be taken out of quotation marks, as no reference is supplied for the use of the phrase}}

\item{\textcolor{blue}{3:29 – In what ways are morphological data inferior? I think this needs to be qualified.}}

\item{\textcolor{blue}{4:45 – depend $->$ dependent}}

\item{\textcolor{blue}{7:108 – inferences $->$ inference}}

\item{\textcolor{blue}{8:134 – does $->$ do}}

\item{\textcolor{blue}{8:135 – Do we actually expect high levels of correlation to lead to inaccurate phylogenies? I think this depends on what is considered to be a high level of correlation and how that correlation is structured. Imagine 100 characters covarying strongly with each other in blocks of 10, but no constraint on the level of correlation across all 10 blocks. This would provide an ``effective'' character number of ~ 10, which could be enough to resolve a small tree.}}

\item{\textcolor{blue}{12:190 – are $->$ as}}

\item{\textcolor{blue}{15:240 – The reference for the 0.26 cut off used by these authors should also be cited}}

\item{\textcolor{blue}{18:286 – an $->$ and}}

\item{\textcolor{blue}{18:286 – converged $->$ converge}}

\item{\textcolor{blue}{18:298 – Surely strict isn't correct here as you are summarising a posterior sample? Is it not just a Majority rule tree?}}

\item{\textcolor{blue}{19:310 – where $->$ were}}

\item{\textcolor{blue}{19:311 – where $->$ were}}

\item{\textcolor{blue}{20:341 – big $->$ in size}}

\item{\textcolor{blue}{29:428 – really $->$ very}}

\item{\textcolor{blue}{29:434 – delete ‘a'}}

\item{\textcolor{blue}{30: 450 - < $->$ >}}

\item{\textcolor{blue}{30:457 – a score}}

\item{\textcolor{blue}{30:457 – on $->$ one}}

\item{\textcolor{blue}{30:469 – delete ‘b'}}

\item{\textcolor{blue}{31:479 – of $->$ off}}

\item{\textcolor{blue}{31:487 – and the use of models that utilise tree structure}}

\item{\textcolor{blue}{32:492 – worth $->$ worst}}

\item{\textcolor{blue}{32:511 – citation includes first name ``David''}}

\item{\textcolor{blue}{33:522 – I think the last sentence is redundant as the point has been made well before in the text.}}

\end{enumerate}

\end{document}

### 2018-06-04 Writing/response_to_reviewers.tex
Updated revisions

@@
 
-\item{\textcolor{blue}{Most of the time when RF is normalized, it's done so so that low scores are better - i.e.
+\item{\textcolor{blue}{Most of the time when RF is normalized, it's done so that low scores are better - i.e.
 a low RF score indicates less distance to the true tree.
 I was very confused looking at the figures, until I realized this was not how normalizing had been performed.
-"However, the effect of these sources of correlation was out of the scope of this study and would have required a posteriori changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest."\\
+}}
+
+%TG: not sure what to answer to this one. Yeah we've normalised using a different method that. Should I just insit on that in the text?
+
+\item{\textcolor{blue}{"However, the effect of these sources of correlation was out of the scope of this study and would have required a posteriori changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest."\\
 I'm not sure I follow this.}}
+
+We meant that modifying phylogenetic matrices after the phylogenetic inference is a way to test the effect of the sources of correlation but is not at all a good practice in systematics (it can be seen as dishonest to simply remove characters because they ``don't give the right signal'').
+We've removed the end of this sentence however to avoid any further confusion.
 
@@
 That will give us some idea of the scope of how close we can get to true, under these conditions, and will help contextualize the results.}}
+
+%TG: add this to the supplementatries
 
@@
 is positively valued and symmetric (copied below from line 178): [...] But, the following metric property does not hold for all x; y:}}
+
+We are really grateful that this reviewer picked up this flow in our mathematical demonstration of CD metric.
+This problem is due to a lack of indications from our side and we apologise for this.
+
+\subsubsection{Character states translation}
+In fact, before being plugged into the CD equation, characters needs to be modified following Felsenstein's \textit{xyz} notation.
+In other words, because the character states tokens are not representing any meaningful numerical value if the characters are unordered (e.g.
+there is no assumption that state 0 should come before state 1), they need to be reordered in a standard way so that the first token in the character is renamed to ``1'', the second to ``2'' etc.
+For example, in a 3 state character for four taxa: \texttt{X = {1, 0, 2, 1}}, the character tokens that are integers representing the states could equally have been \texttt{X = {E, A, I, E}} or \texttt{X = {blue, red, absent, blue}}.
+Therefore, to compare characters fairly, the tokens needs to be translated in a standard way such as the first token is replaced by the symbol ``1'', etc.
+In this example \texttt{X = {1, 0, 2, 1}} becomes \texttt{X' = {1, 2, 3, 1}}
+In the reviewer's demonstration, \texttt{x  = {1}} and \texttt{y  = {0}} thus become \texttt{x'  = {1}} and \texttt{y'  = {1}}.
+Hence $CD_{(x',y')} = 0 \Leftrightarrow x' = y'$.
+
+\subsubsection{CD as a metric to compare the splits between characters}
+Furthermore, the CD metric, used in this manuscript as a proxy for character correlation actually measures the difference between in characters in terms of phylogenetic signal.
+The character \texttt{X} described above implies 3 splits for four taxa (and of course, the character \texttt{X'} implies the exact same splits).
+Another character \texttt{Y = {1, 1, 1, 2}} will imply only two splits, non shared with \texttt{X} and a character \texttt{Z = {0, 2, 1, 0}} (\texttt{Z' = {1, 2, 3, 1}}) would imply the same splits as \texttt{X} (hence, $CD_{(X',Z')} = 0$ and $CD_{(X',Y')} = 0.5$).
+
+Following these two points, the reviewer counter example $CD_{(x,y)} = 0$ means that 1) $x'$ and $y'$ are identical and the characters $x$ and $y$ imply the same number of splits in a tree (here no splits since only one taxon is used - however, this would hold for any $n$ number of taxa).
+We have clarified these two points in the manuscript and in the mathematical demonstration.
+
+%TG: TODO: add the stuff to the manuscript
 
@@
 
+We've removed the quotes.
+
 \item{\textcolor{blue}{3:29 – In what ways are morphological data inferior? I think this needs to be qualified.}}
+
+We've added the following justification (and toned down the sentence):
+\textit{Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collected them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.}
 
@@
 
+We fixed this typo.
+
 \item{\textcolor{blue}{7:108 – inferences $->$ inference}}
 
+We fixed this typo.
+
 \item{\textcolor{blue}{8:134 – does $->$ do}}
+
+We fixed this typo.
 
@@
 
+%TG: need a way to phrase this but totally agree, in the reviewer's example, the matrix would look "honets" (lots of characters) but the results would be actually based on way fewer.
+
 \item{\textcolor{blue}{12:190 – are $->$ as}}
+
+We fixed this typo.
 
@@
 
+We've added the reference Sanderson \& Donoghue 1989.
+
 \item{\textcolor{blue}{18:286 – an $->$ and}}
+
+We fixed this typo.
 
@@
 
+We fixed this typo.
+
 \item{\textcolor{blue}{18:298 – Surely strict isn't correct here as you are summarising a posterior sample? Is it not just a Majority rule tree?}}
+
+We removed the term ``strict''.
 
@@
 
+We fixed this typo.
+
 \item{\textcolor{blue}{19:311 – where $->$ were}}
 
+We fixed this typo.
+
 \item{\textcolor{blue}{20:341 – big $->$ in size}}
+
+We fixed this typo.
 

### 2018-06-04 Writing/draft.tex
Updated revisions

@@
 
-The last two decades have witnessed a ``resurgence'' of interest in the use of morphological character data in phylogenetic studies.
+The last two decades have witnessed a resurgence of interest in the use of morphological character data in phylogenetic studies.
 This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
@@
 The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the very nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
+Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the very nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are \hl{dependent} on each other.
 Before discussing character correlation further, it is important to understand that it may manifest itself in at least three distinct ways:
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see Supplementary material 3 for the full summary statistics).
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see supplementary material 3 for the full summary statistics).
 There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
@@
 However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
-%MB: I have no idea what thae above sentence means.
-%TG: Modified. Is that clearer?
 Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.

### 2018-06-04 Writing/draft_highlighted.tex
Updated revisions in the supplementary

@@
 This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
-Morphological character data are often considered inferior to molecular sequence data, but are often the only source of phylogenetic data for extinct species. 
+\hl{Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collected them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.}
 While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
@@
 Biological and evolutionary dependences are inherent parts to Evo-Devo and macroevolutionary studies and best practices to avoid coding-induced dependences are commonly known and applied.
-However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inferences software that are blind to these distinctions.
+However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inference software that are blind to these distinctions.
 If fact, what the software are confronted with is a two dimensional matrix problem that renders the morphological subtleties described opaque.
@@
 
-How does these correlation really affect topology?
+How \hl{do} these correlation really affect topology?
 We expect matrices with a high level of correlation to recover precise but inaccurate topologies but will matrices with low level of correlation (i.e.
@@
 In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
-This procedure was used to treat all the characters are unordered with no assumption on the meaning of the character state (e.g. in a binary character $0$ is not necessary ancestral to $1$).
+This procedure was used to treat all the characters \hl{as} unordered with no assumption on the meaning of the character state (e.g. in a binary character $0$ is not necessary ancestral to $1$).
 It also greatly improved the speed of our algorithm implementation to compare the characters.
@@
 We re-simulated every invariant characters to obtain a matrix with no invariant characters in order to better approximate real morphological data matrices.
-To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{OReilly20160081}.
+To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{sanderson1989patterns,OReilly20160081}.
 \hl{Note that both models (``morphological HKY-binary' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
@@
 We automatically stopped the MCMC when the average standard deviation of split frequencies (ASDSF) between both runs fell below 0.01 (with a diagnosis every $1\times10^4$ generations - \texttt{mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES}).
-Due to cluster hardware requirements an to save some time, when chains didn't converged and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the unconverged chains.
+Due to cluster hardware requirements \hl{and} to save some time, when chains didn't \hl{converge} and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the unconverged chains.
 In practice, these few MCMC got stuck at an ASDSF around (but not below) 0.01.
 
-A strict majority rule tree was then calculated for both Bayesian an maximum parsimony trees.
+A majority rule tree was then calculated for both Bayesian an maximum parsimony trees.
 For the Bayesian consensus trees, the 25\% first trees of the posterior tree distribution were excluded as a burnin.
@@
 
-The metric scores where calculated using the \texttt{TreeCmp} javascript \citep{Bogdanowicz2012}.
-The measurements where then standardised using the Normalised Tree Similarity metric \citep[$NTS$; i.e.
+The metric scores were calculated using the \texttt{TreeCmp} javascript \citep{Bogdanowicz2012}.
+The measurements were then standardised using the Normalised Tree Similarity metric \citep[$NTS$; i.e.
 centering the metrics scores using the mean metric score for 1000 pairwise comparisons between random trees with $n$ taxa;][]{Bogdanowicz2012,Guillerme2016146}.
@@
 The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.
+However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.
 
@@
 creating ``modules'' of characters).
-However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest.
+However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices.% which are - when using empirical data - at best bad practice and at worth dishonest.
 

### 2018-06-04 Writing/supplementary/Sup1-CharDiff.tex
updated header
Updated supplementaries
Updated revisions in the supplementary

@@
 \noindent{\Large \bf Influence of different modes of morphological character correlation on phylogenetic tree inference}
+
 \bigskip
+
 \noindent{\Large \bf Supplementary material 1 - Character difference}
+
 \bigskip
+
 \noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
@@
 % \linenumbers
+\newpage
 
@@
 
-% \begin{equation}
-%     CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
-% \end{equation}
+$CD$, the Character Difference metric, is a distance that measures the difference between two characters in terms of shared phylogenetic splits.
+For two characters $x$ and $y$, $CD=0$ implies that the characters are identical, i.e.
+they convey the same phylogenetic signal in term of splits; $CD=1$ implies that the characters are the most different, i.e.
+they convey the most different phylogenetic signal in terms of splits.
 
@@
 
-\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
+\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each transformed characters states for the characters $x$ and $y$ and the taxa $i$.
+The characters are transformed following \cite{felsenstein2004inferring}'s \textit{xyz} notation where characters states are reorder so that:
 
-\newpage
+\begin{equation}
+x'_1 = 1 \text{ and } y'_1 = 1
+\end{equation}
+
+\noindent the following character states are then either translated using the same notation for each $i^{th}$ taxa where the state integer is incremented by $+1$ for each new encountered state:
+
+\begin{equation}
+\text{If } x_i = x_1 \text{ then } x'_i = 1 \text{; else } x'_i = x'_1 + 1
+\end{equation}
+
+\begin{equation}
+\text{If } y_i = y_1 \text{ then } y'_i = 1 \text{; else } y'_i = y'_1 + 1
+\end{equation}
+
+\noindent For example if:
+
+\texttt{x = {0,0,2,2,1,1}} and \texttt{y = {8,8,0,0,3,3}}
+
+\noindent then:
+
+\texttt{x' = {1,1,2,2,3,3}} and \texttt{y' = {1,2,2,2,3,3}}
+
+\indent The following proofs consider that the \textit{xyz} notation was applied.
+
 \section{Distance metric proof}
 
-The Character Difference ($CD$) metric is a distance $d: CD \times CD \rightarrow [0;\infty)$ with any character $x$, $y$ and $z$ $\in CD$ if an only if the followings are satisfied:
+The Character Difference ($CD$) metric is a distance $d: CD \times CD \rightarrow [0;\infty)$ with any character $x$, $y$ and $z$ (using the \textit{xyz} notation) $\in CD$ if an only if the followings are satisfied:
 
@@
 \end{equation}
+
+\newpage
 
@@
 Thus equation \ref{Positivity} is true.
+
+\newpage
 
@@
 
+\bibliographystyle{sysbio}
+\bibliography{References}
+
 \end{document}

### 2018-06-04 Writing/draft_highlighted.tex
Finished minor corrections

@@
 
-Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distributions that are always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
+Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distributions that are always \hl{very} high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
 This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
 For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation as it is independent of the number of characters analysed.
-Similarly, using the a Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
+Similarly, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
 we corrected for the expected difference when comparing two random trees with the same number of taxa).
@@
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[i.e.
-those with in a $CI<0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
+those with in a $CI>0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
@@
 For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
-(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
+(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
 The tree search will thus likely result in a topology island that will not contain the given topology $T$.
-(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e.
+(2) in Bayesian inference, the topology will also be given a bad score (i.e.
 low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
-The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
+The rate and the branch length being two parameters among others, the score (the likelihood) will change less drastically than for using parsimony.
 Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
-This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.b
+This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.
 
@@
 coding correlation) the only type of correlation that can affect the phylogeny \textit{a priori}.
-This dichotomy thus creates a trade of between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
+This dichotomy thus creates a trade \hl{off} between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
 and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation being less dependent on the coding correlations.
@@
 creating ``modules'' of characters).
-However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices.% which are - when using empirical data - at best bad practice and at worth dishonest.
+However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices.% which are - when using empirical data - at best bad practice and at worst dishonest.
 
@@
 But there are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
-However, we want to point out again that here we're no comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
+% However, we want to point out again that here we're no comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
 

### 2018-06-06 Writing/response_to_reviewers.tex
Finished minor corrections
Specified results locations

@@
 
-\item{\textcolor{blue}{I have some concerns about the approach, however.
-I was somewhat confused by the CD metric.
-When quantifying n, how is it decided how many taxa have "comparable" characters?
-If you simulated the character for all taxa, do they not all have this character (i.e.
-n always equals the dataset size)?
-But more importantly, will two characters not be more similar if they share the same number of character states, independent
-of if they are telling you the same evolutionary history? It seems to me that whether or not two characters share a state space is not the relevant axis along which to measure correlation.}}
+\item{\textcolor{blue}{I have some concerns about the approach, however. I was somewhat confused by the CD metric. When quantifying n, how is it decided how many taxa have "comparable" characters?
+If you simulated the character for all taxa, do they not all have this character (i.e. n always equals the dataset size)?
+But more importantly, will two characters not be more similar if they share the same number of character states, independent of if they are telling you the same evolutionary history?
+It seems to me that whether or not two characters share a state space is not the relevant axis along which to measure correlation.}}
+
+%TG: Specify the Gower distance (comparable)
+%TG: yes, n is fixed here (this is just for the CD definition)
+%TG: no, characters with missing data should be more dissimilar: since the similarity is measured in terms of shared splits, a missing data can always involve a new split (if the ? is a new state)
+%TG: and not sure about the directionality thing.
+
+%TG: TODO: clarify the CD metric definition
 
@@
 I had a hard time putting a context to this and deciding how well I feel these simulations approximate reality.
-Without knowing this, I was having difficulty understanding why the statement "when character correlation was high (low character differences), the topology was always the furthest away from the
-``normal'' topology." should be true - does this simply mean that the same signal is being repeated over and over again, but that signal is low information relative to the topology?
+Without knowing this, I was having difficulty understanding why the statement "when character correlation was high (low character differences), the topology was always the furthest away from the``normal'' topology."
+should be true - does this simply mean that the same signal is being repeated over and over again, but that signal is low information relative to the topology?
 To my mind, highly correlated datasets would yield congruent trees, but that these trees may be discordant with the true tree, if the resampled characters were not representative of the true signal, or highly concordant with the true tree if they were.
 Perhaps this not being observed is a result of not performing comparisons to the true tree?}}
+
+%TG: TODO: add the range of replaced character n and m
+%TG: TODO: add more explanantion on the replacements in the methods and the discussion (the reviewer's thoughts)
 
@@
 2014 (Integrating Incomplete Fossils by Isolating Conflicting Signal in Saturated and Non-Independent Morphological Characters) is a good one.}}
+
+%TG: TODO: add cites
 
@@
 I'm not sure I follow this.}}
+\label{dishonest}
 
@@
 
-\item{\textcolor{blue}{How did you ``randomly''randomly modify your matrices? Did you sample under some distribution? If so, which one?}}
+%TG: This is a tricky one.
+Simulations where generated using a random seed but each step of the simulation cascades down from it.
+In this context it is hard to control for how the random data was generated (seeds where used at the start of the matrix simulations and the start of the tree inferences to mitigate the effect).
+%TG: In the end this is always a problem with computers and I'm not sure how to solve them unless using true random generators (overkill?).
+
+\item{\textcolor{blue}{How did you ``randomly'' modify your matrices? Did you sample under some distribution? If so, which one?}}
+
+%TG: TODO: add to the methods the sampling distribution (runif(0,1)).
 
@@
 
+%TG: again, not sure what the reviewer means, the HKY picks the HKY scenario and the Mk the GTR one.
+%TG: and no, probably they don't capture all the space, morphological characters can be in any scenario (see Wright et al) but most of them seem to be GTR friendly.
+
 \item{\textcolor{blue}{Similarly, why did you choose 0.85 for the Mk model? Does it represent some important set of character sequences? How does it represent the space in general?}}
 
-\item{\textcolor{blue}{There are statements that seem contradictory, like (line 335): ``Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability > 0:95 and different if they had an overlap probability > 0:05.''}}
+%TG: the 0.85 is the proportion of binary characters. This was extracted from litterature.
+
+\item{\textcolor{blue}{There are statements that seem contradictory, like (line 335): ``Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability > 0.95 and different if they had an overlap probability > 0.05.''}}
+
+We fixed this type. We considered them different if they had an overlap probability $<$ 0.05.
 
@@
 For example, line 381-384: ``Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table 5 - median NTSRF of 0:828 and 0:679 respectively) but not in terms of individual taxon placements (Table 5 - median NTSTr of 0:738 and 0:601 respectively).'' Why is the 0:828 - 0:679 = 0:149 significant for this test and 0:738 - 0:601 = 0:137 is not?}}
+
+The significances was referring to the results in Table 5 (Wilcox test p value of 0 and 0.08 and Bhattacharrya Coefficient of 0.891 and 0.984 respectively) and the median values were extracted from the supplementary material 3 Table 4.
+We now detailed the values and where they are present for each of the tests to avoid confusion.
 
@@
 
+We fixed this typo.
+
 \item{\textcolor{blue}{29:434 – delete ‘a'}}
+
+We fixed this typo.
 
@@
 
+We fixed this typo.
+
 \item{\textcolor{blue}{30:457 – a score}}
+
+We fixed this typo throughout the paragraph.
 
@@
 
+We fixed this typo.
+
 \item{\textcolor{blue}{30:469 – delete ‘b'}}
+
+We fixed this typo.
 
@@
 
+We fixed this typo.
+
 \item{\textcolor{blue}{31:487 – and the use of models that utilise tree structure}}
+
+We fixed this typo.
 
@@
 
+We removed this sentence (see Reviewer 1 specific suggestion \ref{dishonest}).
+
 \item{\textcolor{blue}{32:511 – citation includes first name ``David''}}
 
+We fixed this reference.
+
 \item{\textcolor{blue}{33:522 – I think the last sentence is redundant as the point has been made well before in the text.}}
+
+We removed this sentence.
 

### 2018-06-06 Writing/draft_highlighted.tex
Specified results locations

@@
 Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
-Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
+Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $<0.05$.
 Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g.
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see supplementary material 3 for the full summary statistics).
-There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839\hl{; see supplementary materials 3, Table 1}) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628\hl{; see supplementary materials 3, Table 1}) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303\hl{; see supplementary materials 3, Table 1}).
+There is a significant difference between all scenarios (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test})
 
@@
 This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
-The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 and 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 2}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).
+The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3 \hl{Table 2}) with a significant difference between 100 and 350 or 1000 characters (\hl{\textit{p} = 0, \textit{p} = 0 and \textit{p} = 0.014 respectively}; Table \ref{Tab_pooledscharacters_test}).
 However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.
@@
 Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
-A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
+A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (\hl{\textit{p} = 0.012, \textit{p} = 0.004, \textit{p} = 0 and \textit{p} = 0 respectively}; Table \ref{Tab_pooledstaxa_test} and medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively; see supplementary materials 3 \hl{Table 3}).
 Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
@@
 
-Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively) but not in terms of individual taxon placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively).
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\hl{\textit{p} = 0} Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively; \hl{see supplementary materials 3, Table 4}) but not in terms of individual taxon placements (\hl{\textit{p} = 0.08}; Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively; \hl{see supplementary materials 3, Table 4}).
 

### 2018-06-07 Writing/response_to_reviewers.tex
Some equation proofs

@@
 Establishing the type of distance may be overkill, but the characterization of a metric helps to determine what it can be used for, and what it cannot be used for.}}
+
+In the supplementary materials 1, we mathematically define the characteristics of $CD$ and proved that it is a distance metric since it respects the four first points:
+
+\begin{enumerate}
+\item Distinctiveness: $CD_{(x,y)} = 0 \Leftrightarrow x = y$, proved in equation 10 (note that we corrected the error highlighted by reviewer 2 with a more restrictive definition of $x$ and $y$).
+\item Non-negativity: $CD_{(x,y)} \geq 0$, proved in equation 9.
+\item Symmetry: $CD_{x,y} = CD_{y,x}$, proved in equation 11.
+\item Triangle inequality: $CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}$, proved in equation 12.
+\end{enumerate}
+
+We did not found a proof for the Four-point condition that satisfied the equation:
+
+$CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )}$
+
+So we cannot confirm or reject that $CD$ is a evolutionary distance metric.
 
@@
 
+%TG: Not sure how to deal with this one.
+
 \item{\textcolor{blue}{Another worry I have is that the implied covariance at the tips does not necessarily reflect covariation over the full phylogeny.
 It is conceivable that two characters can appear to have near perfect correlation at the tips, but arrived at these states through completely independent pathways, particularly when simulating with a continuous time Markov chain where unobserved transitions to a new state and back to the original state are possible.}}
+
+%TG: Not sure about this one either: I guess its impossible to say both in simulations or in real life: in the end we only have the tips and their information.
 
@@
 Despite this, the authors present an overview of the effect of character covariance at large, and I certainly think that this is a worthwhile contribution.}}
+
+%TG: Maybe say something along the lines: "yeah, great idea we'll do it later"?
 
@@
 
+%TG: needs to emphasize this point: basically, smaller matrices matrices are more sensible to what the reviewer suggests (i.e.
+the chance of pick up the right or the wrong characters are higher).
+In bigger matrices, this variance is reduced because the number of good/bad characters is diffused.
+
 \item{\textcolor{blue}{Overall, I think that the authors provide an overview of the effect of covariance on the estimation of phylogeny, but without an explicit simulation of patterns of covariation we do not yet have a definitive understanding of the effects of between character covariance.}}
+
+We definitely agree with this statement and are working on a follow up project with more explicit patterns of covariation.
+However, one major bottleneck in such analysis is our ability to actually simulate discrete morphological characters (regardless of their pattern of covariance) that are meaningful in reflecting the actual patterns observed in empirical discrete morphological characters.
+Thankfully, the debate between Goloboff et al. and O'Reilly et al. will provide us with some more insightful and realistic methods to simulate discrete morphological characters.
+
+Secondly, as we highlighted throughout the manuscript (lines @@@), the main mechanisms that generate phylogenetic signal and variance/covariance between characters are the intra-organismal dependence and the evolutionary dependence described in the introduction.
+Both of which can unfortunately only be observed/analysed, \textit{a posteriori} based on a phylogenetic hypothesis.
 
@@
 \item{\textcolor{blue}{2:2 – Phylogenetic algorithms don't necessarily require independence, they simply assume it.}}
+
+We've rephrased this sentence to:
+\textit{Phylogenetic analysis algorithms assume character independence}. lines @@@
 

### 2018-06-07 Writing/draft_highlighted.tex
Some equation proofs

@@
 \begin{abstract}
-Phylogenetic analysis algorithms require the assumption of character independence - a condition generally acknowledged to be violated by morphological data.
+\hl{Phylogenetic analysis algorithms assume character independence} - a condition generally acknowledged to be violated by morphological data.
 Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 

### 2018-06-08 Writing/response_to_reviewers.tex
Updated figures labels

@@
 It seems to me that whether or not two characters share a state space is not the relevant axis along which to measure correlation.}}
+\label{gower}
 
-%TG: Specify the Gower distance (comparable)
-%TG: yes, n is fixed here (this is just for the CD definition)
-%TG: no, characters with missing data should be more dissimilar: since the similarity is measured in terms of shared splits, a missing data can always involve a new split (if the ? is a new state)
-%TG: and not sure about the directionality thing.
+Using a Gower based distance, we only compared the comparable taxa between characters.
+For example, for two characters \texttt{x = {0, 1, 2, 3}} and \texttt{y = {0, ?, ?, ?}}, the Character Difference metric will be effectively calculated using only the first taxon (\texttt{x = {0}} and \texttt{y = {0}}) and would result in a Character Difference of 0.
+As the reviewer points out, characters with less taxa in common will thus be less likely to be different.
+This is consistent with what the CD metric actually captures in terms of phylogenetic signal (the number of share splits - see Reviewer 2 comment \ref{proof}).
+In the example above, the both characters imply at least one split (between the first taxon and the others).
+In other words, in a phylogenetic software, both characters will support the same split.
 
-%TG: TODO: clarify the CD metric definition
+Besides, as the reviewer also correctly points out, we simulated our data with no missing data so this does not affect our results.
+The precision in the manuscript was just describing the properties of the new CD metric.
+
+We've clarified how missing data can be potentially treated in the methods section:
+\textit{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
+Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).} lines @@@
 
@@
 
-We've clarified this sentence by adding the following example:
-\textit{For example, for two characters X and Y for four taxa, if character X is {0,0,?,1} for the first, second, third and fourth taxa respectively and character Y is {0,0,1,1}, only the first, second and fourth taxa are compared for this character.}
+We've clarified this sentence (see Reviewer 1, Major suggestion \ref{gower}).
 
@@
 
-%TG: not sure what to answer to this one. Yeah we've normalised using a different method that. Should I just insit on that in the text?
+We've changed the labels of the figures to NTS(RF) and NTS(Tr) to avoid any confusions.
 
@@
 is positively valued and symmetric (copied below from line 178): [...] But, the following metric property does not hold for all x; y:}}
+\label{proof}
 

### 2018-06-08 Analysis/03-EffectCorrelationFigure.Rmd
Updated figures labels

@@
 plot.results.single(t25_list, metric = 2, what = "best",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "NTS(RF) for 25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
+plot.results.single(t75_list, metric = 2, what = "best", ylab = "NTS(RF) for 75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 2, what = "best",
-    ylab = "RF for 150 taxa", NTS = TRUE)
+    ylab = "NTS(RF) for 150 taxa", NTS = TRUE)
 par(op)
@@
 plot.results.single(t25_list, metric = 2, what = "best",
-    legend = FALSE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    legend = FALSE, NTS = TRUE, xlab = "", ylab = "NTS(RF) for 25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
+plot.results.single(t75_list, metric = 2, what = "best", ylab = "NTS(RF) for 75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 2, what = "best",
-    ylab = "RF for 150 taxa", NTS = TRUE)
+    ylab = "NTS(RF) for 150 taxa", NTS = TRUE)
 par(op)
@@
 plot.results.single(t25_list, metric = 4, what = "best",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "NTS(Tr) for 25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "best", ylab = "Tr for 75 taxa",
+plot.results.single(t75_list, metric = 4, what = "best", ylab = "NTS(Tr) for 75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 4, what = "best",
-    ylab = "Tr for 150 taxa", NTS = TRUE)
+    ylab = "NTS(Tr) for 150 taxa", NTS = TRUE)
 par(op)
@@
 plot.results.single(t25_list, metric = 4, what = "best",
-    legend = FALSE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    legend = FALSE, NTS = TRUE, xlab = "", ylab = "NTS(Tr) for 25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "best", ylab = "Tr for 75 taxa",
+plot.results.single(t75_list, metric = 4, what = "best", ylab = "NTS(Tr) for 75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 4, what = "best",
-    ylab = "Tr for 150 taxa", NTS = TRUE)
+    ylab = "NTS(Tr) for 150 taxa", NTS = TRUE)
 par(op)
@@
 plot.results.single(t25_list, metric = 2, what = "null",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "NTS(RF) for 25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 2, what = "null", ylab = "RF for 75 taxa",
+plot.results.single(t75_list, metric = 2, what = "null", ylab = "NTS(RF) for 75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 2, what = "null",
-    ylab = "RF for 150 taxa", NTS = TRUE)
+    ylab = "NTS(RF) for 150 taxa", NTS = TRUE)
 par(op)
@@
 plot.results.single(t25_list, metric = 4, what = "null",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "NTS(Tr) for 25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "null", ylab = "Tr for 75 taxa",
+plot.results.single(t75_list, metric = 4, what = "null", ylab = "NTS(Tr) for 75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 4, what = "null",
-    ylab = "Tr for 150 taxa", NTS = TRUE)
+    ylab = "NTS(Tr) for 150 taxa", NTS = TRUE)
 par(op)
@@
 plot.results.single(t25_list, metric = 4, what = "null",
-    legend = FALSE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
+    legend = FALSE, NTS = TRUE, xlab = "", ylab = "NTS(Tr) for 25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "null", ylab = "Tr for 75 taxa",
+plot.results.single(t75_list, metric = 4, what = "null", ylab = "NTS(Tr) for 75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 4, what = "null",
-    ylab = "Tr for 150 taxa", NTS = TRUE)
+    ylab = "NTS(Tr) for 150 taxa", NTS = TRUE)
 par(op)

### 2018-06-08 Writing/draft_highlighted.tex
Updated figures labels

@@
 Logical dependence, however, is easily distinguished prior to phylogenetic inference, while the two other ones (intra-organismal and evolutionary) are much harder.
-However, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{DeLaet2015,BrazeauNA}.
+However, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{de2015parsimony,BrazeauNA}.
 Intra-organismal dependence requires more detailed, often extremely time-consuming studies (and possibly beyond the limits of available technology).
@@
 \subsection{Measuring differences between characters}
-\label{CD_description}
+\label{CDdescription}
 To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information.
 In order to measure this, we propose a new distance metric to measure the difference between two characters:
-%TG: OK, I've checked Farris' stuff.
-Despite not beeing described (at least not the in 1970 Syst Biol paper), it seems that the advancement index is just the distance between two taxa.
-Appart from the taxa/character difference, it's kind of similar but (1) it's not ponderated (divided by n), centered (-1/2) and reversed (1-2(...)); so I'm not sure if we need to say it's similar.
-In the end it's a distance metric and it can be pretty similar to *any* distance metric.
-I think we can leave as it is (plus, I've reinvented the wheel by mathematicallty demonstrating it's a distance - my first only real math in a draft).
 
@@
 Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
-\hl{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.}
+\hl{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
+Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).}
 
@@
 As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
-The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
+The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
@@
 Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g.
-for characters hierarchy][]{deLaet2015,BrazeauNA} already exist.
+for characters hierarchy][]{de2015parsimony,BrazeauNA} already exist.
 However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
@@
 Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
-We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD metric proof.
+We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes and Abigail Pastore for helping with the CD metric proof and the \hl{three anonymous reviewers for there insightful comments}.
 

### 2018-06-08 Analysis/04-CheckMatrixCorrelation.Rmd
Character replacements

---
title: "Check matrix correlations"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Checking the number of character replaced in each matrix

```{r}
library(ape)
## Setting the path
PATH <- "../Data/Simulations/Matrices/"
```

The following function counts, _a posteriori_ the number of characters that where replaced in each matrix (by comparing them to the normal matrix)

```{r}
## Getting the proportion of replaced characters
get.proportion.replace <- function(matrix, path) {
    ## Getting one matrix in a comparable format
    get.one.matrix <- function(path, matrix, type) {
        return(unlist(apply(do.call(rbind, read.nexus.data(paste0(path, matrix, "_", type, ".nex"))), 2, list), recursive = FALSE))
    }

    ## Getting all the matrices
    norm <- get.one.matrix(path, matrix, "norm")
    maxi <- get.one.matrix(path, matrix, "maxi")
    mini <- get.one.matrix(path, matrix, "mini")
    rand <- get.one.matrix(path, matrix, "rand")

    ## Comparing all the matrices
    n <- length(norm)
    maxi <- length(which(!mapply(function(X, Y) all(X == Y), norm, maxi)))/n
    mini <- length(which(!mapply(function(X, Y) all(X == Y), norm, mini)))/n
    rand <- length(which(!mapply(function(X, Y) all(X == Y), norm, rand)))/n

    return(c("maxi" = maxi, "mini" = mini, "rand" = rand))
}
```

We can then apply it to the list of matrices

```{r}
## Creating the placeholder list for the results
replacements <- list(list(), list(), list())
## getting the values to loop through
taxa_val <- c(25, 75, 150)
character_val <- c(100, 350, 1000)
for(taxa in 1:3) {
    for(character in 1:3) {
        ## Get the taxa names
        names <- unique(gsub("_[a-z]*.nex", "",
          list.files(PATH, pattern = paste0(taxa_val[taxa], "t_", character_val[character], "c_"))))
        ## Get the proportions of replaced characters
        replacements[[taxa]][[character]] <- lapply(as.list(names), get.proportion.replace, PATH)
    }
}

## Combining the results
fun.maxi <- function(list) {
    return(unlist(lapply(list, function(X) return(X[[1]]))))
}
fun.mini <- function(list) {
    return(unlist(lapply(list, function(X) return(X[[2]]))))
}
fun.rand <- function(list) {
    return(unlist(lapply(list, function(X) return(X[[3]]))))
}
results_maxi <- lapply(replacements, lapply, fun.maxi)
results_mini <- lapply(replacements, lapply, fun.mini)
results_rand <- lapply(replacements, lapply, fun.rand)

## Naming the lists
c("25t", "75t", "100t") -> names(results_maxi) -> names(results_mini) -> names(results_rand)
c("100c", "350c", "1000c") -> names(results_maxi[[1]]) ->
                              names(results_maxi[[2]]) ->
                              names(results_maxi[[3]]) ->
                              names(results_mini[[1]]) ->
                              names(results_mini[[2]]) ->
                              names(results_mini[[3]]) ->
                              names(results_rand[[1]]) ->
                              names(results_rand[[2]]) ->
                              names(results_rand[[3]])
```

We can then represent the results using box plots:

```{r, fig.height = 6, fig.width = }

boxplot(results_rand[[3]], xaxis = "n")

plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, NTS = TRUE, ylab, xlab, axislab, ...) {

    ## Graphic parameters
    ## Colours
    cols <- c("red", "orange", "green3", "lightgreen", "blue", "lightblue")

    ## Y lab
    if(missing(ylab)) {
        if(metric == 1) ylab <- "Similarity (Matching Cluster)"
        if(metric == 2) ylab <- "Similarity (Robinson-Foulds)"
        if(metric == 3) ylab <- "Similarity (Nodal Split)"
        if(metric == 4) ylab <- "Similarity (Triplets)"
    }

    ## Y lim
    get.min.list <- function(X, metric) {
        return(min(unlist(X[,metric])))
    }

    min <- min(unlist(lapply(unlist(unlist(unlist(taxa_list, recursive = FALSE), recursive = FALSE), recursive = FALSE), get.min.list, metric)))
    if(min < 0) {
        ylim <- c(min,1)
    } else {
        ylim <- c(0,1)
    }

    ## X lab
    if(missing(xlab)) {
        xlab <- "Characters"
    }

    ## axis lab
    if(missing(axislab)) {
        axislab <- c("100", "350", "1000")
    }

    ## Horizontal line
    line <- 0

    if(what == "null") {
        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), las = 1, ...,
            ## c100
            unlist(taxa_list$c100$bayesian$rand$maxi[,metric]), unlist(taxa_list$c100$parsimony$rand$maxi[,metric]),
            unlist(taxa_list$c100$bayesian$rand$mini[,metric]), unlist(taxa_list$c100$parsimony$rand$mini[,metric]),
            unlist(taxa_list$c100$bayesian$rand$norm[,metric]), unlist(taxa_list$c100$parsimony$rand$norm[,metric]),
            ## c350
            unlist(taxa_list$c350$bayesian$rand$maxi[,metric]), unlist(taxa_list$c350$parsimony$rand$maxi[,metric]),
            unlist(taxa_list$c350$bayesian$rand$mini[,metric]), unlist(taxa_list$c350$parsimony$rand$mini[,metric]),
            unlist(taxa_list$c350$bayesian$rand$norm[,metric]), unlist(taxa_list$c350$parsimony$rand$norm[,metric]),
            ## c1000
            unlist(taxa_list$c1000$bayesian$rand$maxi[,metric]), unlist(taxa_list$c1000$parsimony$rand$maxi[,metric]),
            unlist(taxa_list$c1000$bayesian$rand$mini[,metric]), unlist(taxa_list$c1000$parsimony$rand$mini[,metric]),
            unlist(taxa_list$c1000$bayesian$rand$norm[,metric]), unlist(taxa_list$c1000$parsimony$rand$norm[,metric])
            )

        ## X axis
        axis(1, 1:18, labels = FALSE, tick = FALSE)
        axis(1, c(3.5, 9.5, 15.5), tick = FALSE, labels = axislab)
        ## Lines
        abline(v = 6.5) ; abline(v = 12.5)
        abline(h = line, col = "grey", lty = 2)
        ## Legend
        if(legend) {
            legend("bottomleft", col = cols, pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
                legend = c("Maximised (Bayesian)", "Maximised (Parsimony)",
                           "Minimised (Bayesian)", "Minimised (Parsimony)",
                           "Normal (Bayesian)", "Normal (Parsimony)")
                )
        }
    }

    if(what == "best") {
        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), las = 1, ...,
            ## c100
            unlist(taxa_list$c100$bayesian$norm$maxi[,metric]), unlist(taxa_list$c100$parsimony$norm$maxi[,metric]),
            unlist(taxa_list$c100$bayesian$norm$mini[,metric]), unlist(taxa_list$c100$parsimony$norm$mini[,metric]),
            unlist(taxa_list$c100$bayesian$norm$rand[,metric]), unlist(taxa_list$c100$parsimony$norm$rand[,metric]),
            ## c350
            unlist(taxa_list$c350$bayesian$norm$maxi[,metric]), unlist(taxa_list$c350$parsimony$norm$maxi[,metric]),
            unlist(taxa_list$c350$bayesian$norm$mini[,metric]), unlist(taxa_list$c350$parsimony$norm$mini[,metric]),
            unlist(taxa_list$c350$bayesian$norm$rand[,metric]), unlist(taxa_list$c350$parsimony$norm$rand[,metric]),
            ## c1000
            unlist(taxa_list$c1000$bayesian$norm$maxi[,metric]), unlist(taxa_list$c1000$parsimony$norm$maxi[,metric]),
            unlist(taxa_list$c1000$bayesian$norm$mini[,metric]), unlist(taxa_list$c1000$parsimony$norm$mini[,metric]),
            unlist(taxa_list$c1000$bayesian$norm$rand[,metric]), unlist(taxa_list$c1000$parsimony$norm$rand[,metric])
            )

        ## X axis
        axis(1, 1:18, labels = FALSE, tick = FALSE)
        axis(1, c(3.5, 9.5, 15.5), tick = FALSE, labels = axislab)
        ## Lines
        abline(v = 6.5) ; abline(v = 12.5)
        abline(h = line, col = "grey", lty = 2)
        ## Legend
        if(legend) {
            legend("bottomleft", col = cols, pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
                legend = c("Maximised (Bayesian)", "Maximised (Parsimony)",
                           "Minimised (Bayesian)", "Minimised (Parsimony)",
                           "Randomised (Bayesian)", "Randomised (Parsimony)")
                )
        }
    }

    if(what == "true") {
        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols[1:4], 3), las = 1, ...,
            ## c100
            unlist(taxa_list$c100$bayesian$norm$true[,metric]), unlist(taxa_list$c100$parsimony$norm$true[,metric]),
            unlist(taxa_list$c100$bayesian$rand$true[,metric]), unlist(taxa_list$c100$parsimony$rand$true[,metric]),
            ## c350
            unlist(taxa_list$c350$bayesian$norm$true[,metric]), unlist(taxa_list$c350$parsimony$norm$true[,metric]),
            unlist(taxa_list$c350$bayesian$rand$true[,metric]), unlist(taxa_list$c350$parsimony$rand$true[,metric]),
            ## c1000
            unlist(taxa_list$c1000$bayesian$norm$true[,metric]), unlist(taxa_list$c1000$parsimony$norm$true[,metric]),
            unlist(taxa_list$c1000$bayesian$rand$true[,metric]), unlist(taxa_list$c1000$parsimony$rand$true[,metric])
            )

        ## X axis
        axis(1, 1:12, labels = FALSE, tick = FALSE)
        axis(1, c(2.5, 6.5, 10.5), tick = FALSE, labels = axislab)
        ## Lines
        abline(v = 4.5) ; abline(v = 8.5)
        abline(h = line, col = "grey", lty = 2)
        ## Legend
        if(legend) {
            legend("bottomleft", col = cols[1:4], pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
                legend = c("Normal (Bayesian)", "Normal (Parsimony)",
                           "Randomised (Bayesian)", "Randomised (Parsimony)")
                )
        }
    }

    if(what == "check") {
        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols[1:4], 3), las = 1, ...,
            ## c100
            unlist(taxa_list$c100$bayesian$norm$rand[,metric]), unlist(taxa_list$c100$parsimony$norm$rand[,metric]),
            unlist(taxa_list$c100$bayesian$rand$norm[,metric]), unlist(taxa_list$c100$parsimony$rand$norm[,metric]),
            ## c350
            unlist(taxa_list$c350$bayesian$norm$rand[,metric]), unlist(taxa_list$c350$parsimony$norm$rand[,metric]),
            unlist(taxa_list$c350$bayesian$rand$norm[,metric]), unlist(taxa_list$c350$parsimony$rand$norm[,metric]),
            ## c1000
            unlist(taxa_list$c1000$bayesian$norm$rand[,metric]), unlist(taxa_list$c1000$parsimony$norm$rand[,metric]),
            unlist(taxa_list$c1000$bayesian$rand$norm[,metric]), unlist(taxa_list$c1000$parsimony$rand$norm[,metric])
            )

        ## X axis
        axis(1, 1:12, labels = FALSE, tick = FALSE)
        axis(1, c(2.5, 6.5, 10.5), tick = FALSE, labels = axislab)
        ## Lines
        abline(v = 4.5) ; abline(v = 8.5)
        abline(h = line, col = "grey", lty = 2)
        ## Legend
        if(legend) {
            legend("bottomleft", col = cols[1:4], pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
                legend = c("Normal (Bayesian)", "Normal (Parsimony)",
                           "Randomised (Bayesian)", "Randomised (Parsimony)")
                )
        }
    }
}

combine_results <- unlist(lapply(replacements[[1]][[1]], function(X) return(X[[1]])))

### 2018-06-08 Writing/response_to_reviewers.tex
Added four point proof

@@
 \begin{enumerate}
-\item Distinctiveness: $CD_{(x,y)} = 0 \Leftrightarrow x = y$, proved in equation 10 (note that we corrected the error highlighted by reviewer 2 with a more restrictive definition of $x$ and $y$).
-\item Non-negativity: $CD_{(x,y)} \geq 0$, proved in equation 9.
-\item Symmetry: $CD_{x,y} = CD_{y,x}$, proved in equation 11.
-\item Triangle inequality: $CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}$, proved in equation 12.
+\item Distinctiveness: $CD_{(x,y)} = 0 \Leftrightarrow x = y$, proved in equation 11 (note that we corrected the error highlighted by reviewer 2 with a more restrictive definition of $x$ and $y$).
+\item Non-negativity: $CD_{(x,y)} \geq 0$, proved in equation 10.
+\item Symmetry: $CD_{x,y} = CD_{y,x}$, proved in equation 12.
+\item Triangle inequality: $CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}$, proved in equation 13.
+\item Four point condition: $CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )}$, proved in equation 15.
 \end{enumerate}
-
-We did not found a proof for the Four-point condition that satisfied the equation:
-
-$CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )}$
-
-So we cannot confirm or reject that $CD$ is a evolutionary distance metric.
 

### 2018-06-08 Writing/supplementary/Sup1-CharDiff.tex
Added four point proof

@@
 
-\texttt{x' = {1,1,2,2,3,3}} and \texttt{y' = {1,2,2,2,3,3}}
+\texttt{x' = {1,1,2,2,3,3}} and \texttt{y' = {1,1,2,2,3,3}}
 
-\indent The following proofs consider that the \textit{xyz} notation was applied.
+\noindent The following proofs consider that the \textit{xyz} notation was applied.
 
@@
     CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}
+\end{equation}
+
+\noindent Furthermore, $CD$ is a evolutionary distance metric if it satisfies the four point condition:
+
+\begin{equation}
+    \label{fourpoint}
+    CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )}
 \end{equation}
@@
 
+\newpage
+
+\subsection{Four point condition}
+
+For any $n \in [1;\infty)$ and any four distinct characters $i$, $j$, $n$ and $m$, and if equations \ref{Positivity} and \ref{Subadditivity} are true, the pairwise distances $CD_{(i,j)}$, $CD_{(m,n)}$, $CD_{(i,m)}$, $CD_{(j,n)}$, $CD_{(j,m)}$ and $CD_{(i,n)}$ can be represented as a combination of any five values $a$, $b$, $c$, $d$ and $e$ $\geq 0$:
+
+\begin{figure}[!htbp]
+\centering
+   \includegraphics[width=0.5\textwidth]{Four_point.pdf}
+\end{figure}
+
+Then:
+
+\begin{equation}
+    \begin{aligned}
+    CD_{(i,j)} = a + b \\
+    CD_{(m,n)} = d + e \\
+    CD_{(i,m)} = a + c + d \\
+    CD_{(j,n)} = b + c + e \\
+    CD_{(j,m)} = b + c + d \\
+    CD_{(i,n)} = a + c + e \\
+    \end{aligned}
+\end{equation}
+
+Thus:
+
+\begin{equation}
+    \begin{gathered}
+    CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )} \\ 
+    a + b + d + e \leq \text{max} (a+c+d+b+c+e; b+c+d+a+c+e) \\
+    a + b + d + e \leq a+b+2c+e+d \\
+    2c \geq 0 \\
+    \end{gathered}
+\end{equation}
+
+Thus equation \ref{fourpoint} is true.
+
 \bibliographystyle{sysbio}

### 2018-06-11 Analysis/04-CheckMatrixCorrelation.Rmd
Proportion replacement

@@
     get.one.matrix <- function(path, matrix, type) {
-        return(unlist(apply(do.call(rbind, read.nexus.data(paste0(path, matrix, "_", type, ".nex"))), 2, list), recursive = FALSE))
+        return(unlist(apply(do.call(rbind,
+                                    read.nexus.data(paste0(path, matrix, "_", type, ".nex"))
+                            ), 2, list), recursive = FALSE))
     }
@@
 
-```{r, fig.height = 6, fig.width = }
-
-boxplot(results_rand[[3]], xaxis = "n")
-
-plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, NTS = TRUE, ylab, xlab, axislab, ...) {
-
-    ## Graphic parameters
-    ## Colours
-    cols <- c("red", "orange", "green3", "lightgreen", "blue", "lightblue")
-
-    ## Y lab
-    if(missing(ylab)) {
-        if(metric == 1) ylab <- "Similarity (Matching Cluster)"
-        if(metric == 2) ylab <- "Similarity (Robinson-Foulds)"
-        if(metric == 3) ylab <- "Similarity (Nodal Split)"
-        if(metric == 4) ylab <- "Similarity (Triplets)"
+```{r, fig.height = 18, fig.width = 6}
+## Plotting the different proportion results
+plot.proportions <- function(data, ...) {
+    ## Default colours
+    cols <- gray.colors(3)
+    ## Boxes
+    boxplot(xlab = "", ylab = "Proportion of replacement", xaxt = "n", ylim = c(0, 1), col = cols,
+            c(data[[1]], data[[2]], data[[3]]), ...)
+    legend("topleft", legend = names(data), pch = 19, col = cols, bty = "n")
+    ## X axis
+    axis(1, 1:9, labels = FALSE, tick = FALSE)
+    axis(1, c(2, 5, 8), tick = FALSE, labels = names(data))
+    ## Lines
+    abline(v = 3.5) ; abline(v = 6.5)
 }
 
-    ## Y lim
-    get.min.list <- function(X, metric) {
-        return(min(unlist(X[,metric])))
-    }
+par(mfrow = c(3,1))
+plot.proportions(results_maxi, main = "Maximised differences")
+plot.proportions(results_mini, main = "Minimised differences")
+plot.proportions(results_rand, main = "Randomised differences")
 
-    min <- min(unlist(lapply(unlist(unlist(unlist(taxa_list, recursive = FALSE), recursive = FALSE), recursive = FALSE), get.min.list, metric)))
-    if(min < 0) {
-        ylim <- c(min,1)
-    } else {
-        ylim <- c(0,1)
-    }
-
-    ## X lab
-    if(missing(xlab)) {
-        xlab <- "Characters"
-    }
-
-    ## axis lab
-    if(missing(axislab)) {
-        axislab <- c("100", "350", "1000")
-    }
-
-    ## Horizontal line
-    line <- 0
-
-    if(what == "null") {
-        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), las = 1, ...,
-            ## c100
-            unlist(taxa_list$c100$bayesian$rand$maxi[,metric]), unlist(taxa_list$c100$parsimony$rand$maxi[,metric]),
-            unlist(taxa_list$c100$bayesian$rand$mini[,metric]), unlist(taxa_list$c100$parsimony$rand$mini[,metric]),
-            unlist(taxa_list$c100$bayesian$rand$norm[,metric]), unlist(taxa_list$c100$parsimony$rand$norm[,metric]),
-            ## c350
-            unlist(taxa_list$c350$bayesian$rand$maxi[,metric]), unlist(taxa_list$c350$parsimony$rand$maxi[,metric]),
-            unlist(taxa_list$c350$bayesian$rand$mini[,metric]), unlist(taxa_list$c350$parsimony$rand$mini[,metric]),
-            unlist(taxa_list$c350$bayesian$rand$norm[,metric]), unlist(taxa_list$c350$parsimony$rand$norm[,metric]),
-            ## c1000
-            unlist(taxa_list$c1000$bayesian$rand$maxi[,metric]), unlist(taxa_list$c1000$parsimony$rand$maxi[,metric]),
-            unlist(taxa_list$c1000$bayesian$rand$mini[,metric]), unlist(taxa_list$c1000$parsimony$rand$mini[,metric]),
-            unlist(taxa_list$c1000$bayesian$rand$norm[,metric]), unlist(taxa_list$c1000$parsimony$rand$norm[,metric])
-            )
-
-        ## X axis
-        axis(1, 1:18, labels = FALSE, tick = FALSE)
-        axis(1, c(3.5, 9.5, 15.5), tick = FALSE, labels = axislab)
-        ## Lines
-        abline(v = 6.5) ; abline(v = 12.5)
-        abline(h = line, col = "grey", lty = 2)
-        ## Legend
-        if(legend) {
-            legend("bottomleft", col = cols, pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
-                legend = c("Maximised (Bayesian)", "Maximised (Parsimony)",
-                           "Minimised (Bayesian)", "Minimised (Parsimony)",
-                           "Normal (Bayesian)", "Normal (Parsimony)")
-                )
-        }
-    }
-
-    if(what == "best") {
-        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), las = 1, ...,
-            ## c100
-            unlist(taxa_list$c100$bayesian$norm$maxi[,metric]), unlist(taxa_list$c100$parsimony$norm$maxi[,metric]),
-            unlist(taxa_list$c100$bayesian$norm$mini[,metric]), unlist(taxa_list$c100$parsimony$norm$mini[,metric]),
-            unlist(taxa_list$c100$bayesian$norm$rand[,metric]), unlist(taxa_list$c100$parsimony$norm$rand[,metric]),
-            ## c350
-            unlist(taxa_list$c350$bayesian$norm$maxi[,metric]), unlist(taxa_list$c350$parsimony$norm$maxi[,metric]),
-            unlist(taxa_list$c350$bayesian$norm$mini[,metric]), unlist(taxa_list$c350$parsimony$norm$mini[,metric]),
-            unlist(taxa_list$c350$bayesian$norm$rand[,metric]), unlist(taxa_list$c350$parsimony$norm$rand[,metric]),
-            ## c1000
-            unlist(taxa_list$c1000$bayesian$norm$maxi[,metric]), unlist(taxa_list$c1000$parsimony$norm$maxi[,metric]),
-            unlist(taxa_list$c1000$bayesian$norm$mini[,metric]), unlist(taxa_list$c1000$parsimony$norm$mini[,metric]),
-            unlist(taxa_list$c1000$bayesian$norm$rand[,metric]), unlist(taxa_list$c1000$parsimony$norm$rand[,metric])
-            )
-
-        ## X axis
-        axis(1, 1:18, labels = FALSE, tick = FALSE)
-        axis(1, c(3.5, 9.5, 15.5), tick = FALSE, labels = axislab)
-        ## Lines
-        abline(v = 6.5) ; abline(v = 12.5)
-        abline(h = line, col = "grey", lty = 2)
-        ## Legend
-        if(legend) {
-            legend("bottomleft", col = cols, pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
-                legend = c("Maximised (Bayesian)", "Maximised (Parsimony)",
-                           "Minimised (Bayesian)", "Minimised (Parsimony)",
-                           "Randomised (Bayesian)", "Randomised (Parsimony)")
-                )
-        }
-    }
-
-    if(what == "true") {
-        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols[1:4], 3), las = 1, ...,
-            ## c100
-            unlist(taxa_list$c100$bayesian$norm$true[,metric]), unlist(taxa_list$c100$parsimony$norm$true[,metric]),
-            unlist(taxa_list$c100$bayesian$rand$true[,metric]), unlist(taxa_list$c100$parsimony$rand$true[,metric]),
-            ## c350
-            unlist(taxa_list$c350$bayesian$norm$true[,metric]), unlist(taxa_list$c350$parsimony$norm$true[,metric]),
-            unlist(taxa_list$c350$bayesian$rand$true[,metric]), unlist(taxa_list$c350$parsimony$rand$true[,metric]),
-            ## c1000
-            unlist(taxa_list$c1000$bayesian$norm$true[,metric]), unlist(taxa_list$c1000$parsimony$norm$true[,metric]),
-            unlist(taxa_list$c1000$bayesian$rand$true[,metric]), unlist(taxa_list$c1000$parsimony$rand$true[,metric])
-            )
-
-        ## X axis
-        axis(1, 1:12, labels = FALSE, tick = FALSE)
-        axis(1, c(2.5, 6.5, 10.5), tick = FALSE, labels = axislab)
-        ## Lines
-        abline(v = 4.5) ; abline(v = 8.5)
-        abline(h = line, col = "grey", lty = 2)
-        ## Legend
-        if(legend) {
-            legend("bottomleft", col = cols[1:4], pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
-                legend = c("Normal (Bayesian)", "Normal (Parsimony)",
-                           "Randomised (Bayesian)", "Randomised (Parsimony)")
-                )
-        }
-    }
-
-    if(what == "check") {
-        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols[1:4], 3), las = 1, ...,
-            ## c100
-            unlist(taxa_list$c100$bayesian$norm$rand[,metric]), unlist(taxa_list$c100$parsimony$norm$rand[,metric]),
-            unlist(taxa_list$c100$bayesian$rand$norm[,metric]), unlist(taxa_list$c100$parsimony$rand$norm[,metric]),
-            ## c350
-            unlist(taxa_list$c350$bayesian$norm$rand[,metric]), unlist(taxa_list$c350$parsimony$norm$rand[,metric]),
-            unlist(taxa_list$c350$bayesian$rand$norm[,metric]), unlist(taxa_list$c350$parsimony$rand$norm[,metric]),
-            ## c1000
-            unlist(taxa_list$c1000$bayesian$norm$rand[,metric]), unlist(taxa_list$c1000$parsimony$norm$rand[,metric]),
-            unlist(taxa_list$c1000$bayesian$rand$norm[,metric]), unlist(taxa_list$c1000$parsimony$rand$norm[,metric])
-            )
-
-        ## X axis
-        axis(1, 1:12, labels = FALSE, tick = FALSE)
-        axis(1, c(2.5, 6.5, 10.5), tick = FALSE, labels = axislab)
-        ## Lines
-        abline(v = 4.5) ; abline(v = 8.5)
-        abline(h = line, col = "grey", lty = 2)
-        ## Legend
-        if(legend) {
-            legend("bottomleft", col = cols[1:4], pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
-                legend = c("Normal (Bayesian)", "Normal (Parsimony)",
-                           "Randomised (Bayesian)", "Randomised (Parsimony)")
-                )
-        }
-    }
-}
-
-combine_results <- unlist(lapply(replacements[[1]][[1]], function(X) return(X[[1]])))
+```

### 2018-06-11 Writing/response_to_reviewers.tex
Added reference and molecular data part

@@
 
-\item{\textcolor{blue}{I thought the reproducible science aspect was cool, and well done.
-Some aspects of the literature review and context for the problem need to be strengthened, particularly given that there is good work that has been done on this problem in a molecular model-based framework (see below).
-I've tried to note locations where I was confused by the prose, but this manuscript could certainly use a very close reading before resubmission.}}
-
 \end{enumerate}
@@
 2014 (Integrating Incomplete Fossils by Isolating Conflicting Signal in Saturated and Non-Independent Morphological Characters) is a good one.}}
+\label{molecular}
 
-%TG: TODO: add cites
+We slightly rephrased the sentence in the abstract (See reviewer 3 minor revision \ref{abstract}) and expanded on the distinction between both character types this in the introduction following the reviewer pertinent biobliography suggestion:
+
+\textit{This independence is assumed for both molecular and morphological characters (Huelsenbeck and Nielsen, 1999; D\'{a}valos et al., 2014, Zou and Zhang, 2016).
+However, several distinctions can be made between the amount and the nature of the correlation.
+First, molecular and morphological data often differs in size and in the amount of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but sometimes up to more than ten; Guillerme and Cooper, 2016).
+Second, this discrepancy has been suggested to fewer stronger correlations in morphological data (i.e.
+``concentrated'' correlations) versus weaker and more ``spread'' correlations in molecular data (i.e.
+``diffused'' correlations - though some regions of the genome can be thought as more correlated than others; Huelsenbeck and Nielsen, 1999).
+Furthermore, because of this ``spread'' correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
+changes in morphological states can be more readily linked to a function cf.
+changes in nucleotide states ; see below).} lines @@@
 
@@
 
-%TG: Find cite.
+We now refer to  D\'{a}valos et al., 2014 and Zou and Zhang, 2016's work on the problem of morphological character correlation in contrast of molecular data.
 
@@
 
-%TG: Remove the sentence?
+We removed this sentence and added the reviewer's suggestion on DNA character susceptible to be more correlated in some regions (see Reviewer 1, specific suggestion \ref{molecular}).
 
@@
 \item{\textcolor{blue}{2:2 – Phylogenetic algorithms don't necessarily require independence, they simply assume it.}}
+\label{abstract}
 

### 2018-06-11 Writing/draft_highlighted.tex
Added reference and molecular data part

@@
 
-The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data.
+The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data \cite[e.g.][]{Davalos01072014, ZouConvergence}.
 The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the very nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are \hl{dependent} on each other.
+\hl{This independence is assumed for both molecular and morphological characters} \citep{huelsenbeck1999effect,Davalos01072014,ZouConvergence}.
+\hl{However, several distinctions can be made between the amount and the nature of the correlation.
+First, molecular and morphological data often differs in size and in the amount of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but sometimes up to more than ten;}
+\citealt{Guillerme2016146}
+\hl{);
+Second, this discrepancy has been suggested to fewer stronger correlations in morphological data (i.e.
+``concentrated'' correlations) \textit{versus} weaker and more ``spread'' correlations in molecular data}
+\citep[\hl{i.e. ``diffused'' correlations - though some regions of the genome can be thought as more correlated than others}][]{huelsenbeck1999effect}.
+\hl{Furthermore, because of this ``spread'' correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
+changes in morphological states can be more readily linked to a function \textit{cf.} changes in nucleotide states ; see below).}
 Before discussing character correlation further, it is important to understand that it may manifest itself in at least three distinct ways:
@@
 It also greatly improved the speed of our algorithm implementation to compare the characters.
-This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
+This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} (following the \textit{xyz} notation in \citealt{felsenstein2004inferring}; \hl{as used in} \citealt{Davalos01072014}).
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).

### 2018-06-11 Writing/response_to_reviewers.tex
Added comparisons to the true trees in the supplementaries

@@
 
-%TG: probably insist on what the reviewer says: that the software cannot handle this.
+We've added a paragraph highlithing how the problem indeed comes from the fact that current software are not equipped to handle this issue:
+
+\textit{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
+In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} more than the other characters since this would be \textit{a priori} the characters that are the most likely to reflect the group's evolutionary history in terms of split by separating the artiodactyles from the cetaceans.
+In other words, this is the correlation between characters that depicts the evolutionary history of this group.
+However, the other characters depict the same history (i.e.
+the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleens, the same split would still be estimated by the software).
+One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.} lines @@@
 
@@
 
-%TG: answering that by saying this is a normal neutral procedure?
+We removed the mention of orderdness of the characters since, indeed, it might be more confusing than helpful in this context (we didn't not simulate any ordered characters and the CD metric does not consider characters as ordered - see supplementary material 1).
+We've changed the sentence to the following:
+
+\textit{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in Felsenstein, 2004; as used in D\'{a}valos et al., 2014).} lines @@@
 
@@
 
-%TG: add this to the supplementatries
+We've added the comparisons between the ``normal'' and ``randomised'' trees to the ``true'' tree in the supplementary material 3, figures 6 to 8.
+We've added a reference to these figures in the main text:
+
+\textit{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 6, 7 and 8.} lines @@@
 

### 2018-06-11 Analysis/03-EffectCorrelationFigure.Rmd
Added comparisons to the true trees in the supplementaries

@@
 
+<!-- For generating the figure for the supplementary-->
+
+```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
+pdf("../Writing/Figures/25t_true_tree.pdf", width = 12, height = 8.25)
+op <- par(mfrow=c(2,1))
+plot.results.single(t25_list, metric = 2, what = "true",
+    main = "comparison to 'true' tree (25 taxa)", legend = TRUE, NTS = TRUE)
+plot.results.single(t25_list, metric = 4, what = "true",
+    main = "", legend = FALSE, NTS = TRUE)
+par(op)
+# dev.off()
+```
+
 ### 75 taxa
@@
 
+```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
+pdf("../Writing/Figures/75t_true_tree.pdf", width = 12, height = 8.25)
+op <- par(mfrow=c(2,1))
+plot.results.single(t75_list, metric = 2, what = "true",
+    main = "comparison to 'true' tree (75 taxa)", legend = TRUE, NTS = TRUE)
+plot.results.single(t75_list, metric = 4, what = "true",
+    main = "", legend = FALSE, NTS = TRUE)
+par(op)
+# dev.off()
+```
+
 ### 150 taxa
@@
     main = "comparison to 'true' tree (150 taxa)", legend = TRUE, NTS = TRUE)
+```
+
+```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
+pdf("../Writing/Figures/150t_true_tree.pdf", width = 12, height = 8.25)
+op <- par(mfrow=c(2,1))
+plot.results.single(t150_list, metric = 2, what = "true",
+    main = "comparison to 'true' tree (150 taxa)", legend = TRUE, NTS = TRUE)
+plot.results.single(t150_list, metric = 4, what = "true",
+    main = "", legend = FALSE, NTS = TRUE)
+par(op)
+# dev.off()
 ```

### 2018-06-11 Writing/draft_highlighted.tex
Added comparisons to the true trees in the supplementaries

@@
 Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states tokens will be used in order to infer the tree.
-Some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
+\hl{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
+In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} more than the other characters since this would be \textit{a priori} the characters that are the most likely to reflect the group's evolutionary history in terms of split by separating the artiodactyles from the cetaceans.
+In other words, this is the correlation between characters that depicts the evolutionary history of this group.
+However, the other characters depict the same history (i.e.
+the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleens, the same split would still be estimated by the software).
+One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.}
+
+This illustrates how some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
 Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).
@@
 In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
-This procedure was used to treat all the characters \hl{as} unordered with no assumption on the meaning of the character state (e.g. in a binary character $0$ is not necessary ancestral to $1$).
-It also greatly improved the speed of our algorithm implementation to compare the characters.
-This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} (following the \textit{xyz} notation in \citealt{felsenstein2004inferring}; \hl{as used in} \citealt{Davalos01072014}).
+\hl{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in }
+\citealt{felsenstein2004inferring}; \hl{as used in} \citealt{Davalos01072014}).
+This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).
@@
 When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
+\hl{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 6, 7 and 8.}
 

### 2018-06-11 Writing/response_to_reviewers.tex
Added precisions for the CD metric

@@
 
+%TG: TODO: introduce this
+
+\textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines @@@
+
+\textit{Similarly, an hypothetical comparison of two character for only one taxa \texttt{x = \{0\}} and \texttt{y = \{1\}} would result in comparing two identical characters \texttt{x' = \{1\}} and \texttt{y' = \{1\}}.} lines @@@
+
 %TG: TODO: add the stuff to the manuscript

### 2018-06-11 Writing/draft_highlighted.tex
Added precisions for the CD metric

@@
 \label{CDdescription}
-To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information.
+To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information \hl{(i.e.
+entirely correlated characters would define identical splits between a set of taxa)}.
 In order to measure this, we propose a new distance metric to measure the difference between two characters:
@@
 This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
+\hl{Similarly, an hypothetical comparison of two character for only one taxa \texttt{x = \{0\}} and \texttt{y = \{1\}} would result in comparing two identical characters \texttt{x' = \{1\}} and \texttt{y' = \{1\}}.}
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).

### 2018-06-11 Writing/supplementary/Sup1-CharDiff.tex
Added precisions for the CD metric

@@
 $CD$, the Character Difference metric, is a distance that measures the difference between two characters in terms of shared phylogenetic splits.
+
 For two characters $x$ and $y$, $CD=0$ implies that the characters are identical, i.e.
@@
 \noindent The following proofs consider that the \textit{xyz} notation was applied.
+
+\subsection{Justification for the character states translation}
+Because the character states tokens are not representing any meaningful numerical value if the characters are unordered (e.g.
+there is no assumption that state 0 should come before state 1), they need to be reordered in a standard way so that the first token in the character is renamed to ``1'', the second to ``2'' etc.
+For example, in a 3 state character for four taxa: \texttt{X = {1, 0, 2, 1}}, the character tokens that are integers representing the states could equally have been \texttt{X = {E, A, I, E}} or \texttt{X = {blue, red, absent, blue}}.
+Therefore, to compare characters fairly, the tokens needs to be translated in a standard way such as the first token is replaced by the symbol ``1'', etc.
+In this example \texttt{X = {1, 0, 2, 1}} becomes \texttt{X' = {1, 2, 3, 1}}
+
+\subsection{Interpretation}
+For a characters \texttt{X = {1, 0, 2, 1}} implying three splits between four taxa and two characters \texttt{Y = {1, 1, 1, 2}} and \texttt{Z = {0, 2, 1, 0}} implying respectively two and three splits between four taxa, $CD_{(X,Y)} = 0.5$ and $CD_{(X,Z)} = 0$.
+This can be interpreted as the probability of characters \texttt{X}, \texttt{Y} and \texttt{X}, \texttt{Z} giving a conflicting phylogenetic signal between the four taxa.
+Here \texttt{X}, \texttt{Y} will conflict for 50\% of the taxa and \texttt{X}, \texttt{X} for 0\% of them.
 

### 2018-06-13 Writing/response_to_reviewers.tex
Added comment on the randomness

@@
 Another character \texttt{Y = {1, 1, 1, 2}} will imply only two splits, non shared with \texttt{X} and a character \texttt{Z = {0, 2, 1, 0}} (\texttt{Z' = {1, 2, 3, 1}}) would imply the same splits as \texttt{X} (hence, $CD_{(X',Z')} = 0$ and $CD_{(X',Y')} = 0.5$).
+In other words the probability of X and Z generating conflicting split signal is null and this probability is of 0.5 for X and Y (i.e. half of the splits will be conflicting).
 
@@
 
-%TG: TODO: introduce this
+We've added more explanations in the supplementary materials 1 on (1) the character translation, (2) the four-point metric proof and (3) on how to interpret the $CD$ metric.
+Additionally, we added more precisions in the text:
 
@@
 \textit{Similarly, an hypothetical comparison of two character for only one taxa \texttt{x = \{0\}} and \texttt{y = \{1\}} would result in comparing two identical characters \texttt{x' = \{1\}} and \texttt{y' = \{1\}}.} lines @@@
-
-%TG: TODO: add the stuff to the manuscript
 
@@
 
-%TG: This is a tricky one.
-Simulations where generated using a random seed but each step of the simulation cascades down from it.
-In this context it is hard to control for how the random data was generated (seeds where used at the start of the matrix simulations and the start of the tree inferences to mitigate the effect).
-%TG: In the end this is always a problem with computers and I'm not sure how to solve them unless using true random generators (overkill?).
+%TG: really not sure about this comment...
+This is an epistemological point I think and I don't think we can solve this problem, even using true random-number generator.
+The error due to chaos is by definition impossible to take into account...
+
+We do agree that generating true randomness in current software is tricky but we are not sure how to answer this comment.
+In our simulations, random numbers where used at different levels:
+
+\begin{enumerate}
+    \item The generation of the ``true'' tree.
+    \item The generation of the discrete morphological characters (each 100, 350 and 1000).
+    \item The removal of the characters (for the ``randomised'' matrices).
+    \item The starting seeds for the tree inferences (in both MrBayes and PAUP*).
+\end{enumerate}
+
+Because each of this steps are dependent in a cascade (i.e.
+the random seed used to generated the ``true'' tree influence the generation of the discrete characters which influences the ``randomised'' matrix which influences the results from the random starting seed in the tree inference).
+However, we consider the influence of these cascading random numbers negligible in this analysis.
+Furthermore, we assumed that the multiplication of negligible effects will result in a near null error due to the pseudo random-number generators.
+In fine, our results are bound to be somehow biased by our prior believes and experience as well as by the tools (hardware and software) and the methods used in this manuscript but this is more of an epistemological point and is not specific to this study.
 

### 2018-06-13 Writing/response_to_reviewers.tex
advanced reviewer 2 comments

@@
 %Packages
+\usepackage{natbib}
 \usepackage{pdflscape}
@@
 
-%TG: TODO: add to the methods the sampling distribution (runif(0,1)).
+Every replaced characters where randomly selected from a uniform distribution (0,1). We have now specified this in the text:
+
+\textit{Each replaced characters (i.e. the characters with a weight of +$1$) where randomly selected from the remaining characters using a uniform distribution (0,1).} lines @@@
 
@@
 
-%TG: again, not sure what the reviewer means, the HKY picks the HKY scenario and the Mk the GTR one.
-%TG: and no, probably they don't capture all the space, morphological characters can be in any scenario (see Wright et al) but most of them seem to be GTR friendly.
+There is actually a vigorous debate surrounding how to generate biologically meaningful morphological characters \citep[][;confidential reviews and personal communications witht he authors]{OReilly20160081,goloboff2017weighted,puttick2017uncertain,GoloboffEmpirical,OReillyEmpirical}.
+This ongoing debate, however, focuses mainly on inference methods where data simulation is a tangential aspect of it.
+Therefore, we decided to base our simulation protocol on the empirical observations showing that half of 206 analysed datasets supported a M$k$-like model \citep[][Fig.
+3]{Wright01072016} and the other half being a combination of models not explicitly described here but that are believed to be covered with the HKY-binarised model \citep{puttick2017uncertain,OReillyEmpirical}.
+In our simulations, the idea was to cover a broad range of morphological characters (with variable state numbers, evolution and transition rates) in order to capture a realistic image of the ``discrete morphological character space''.
+
+%TG: add some text to the manuscript parphrasing this? It's kind of already written in there...
 
@@
 
-%TG: the 0.85 is the proportion of binary characters. This was extracted from litterature.
+This 0.85 proportion of binary characters and 0.15 of three state ones was extracted from the 100 empirical matrices analysed in \cite{Guillerme2016146} (Supplementary appendix 1.
+Appendix A: Tree inference) and \cite{ZouConvergence}.
+We've specified this is the text:
 
-\item{\textcolor{blue}{There are statements that seem contradictory, like (line 335): ``Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability > 0.95 and different if they had an overlap probability > 0.05.''}}
+\textit{We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters (from empirical observations in: Guillerme and Cooper, 2016b; Zou and Zhang, 2016).}
+
+\item{\textcolor{blue}{There are statements that seem contradictory, like (line 335): ``Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>$ 0.95 and different if they had an overlap probability $>$ 0.05.''}}
 

### 2018-06-13 Writing/draft_highlighted.tex
advanced reviewer 2 comments

@@
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
-    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
+    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep[\hl{from empirical observations in: }][]{Guillerme2016146,ZouConvergence}.
     This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g.
@@
 Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e.
-with an average character difference $<0.75$) and randomly replaced them with the remaining ones.
+with an average character difference $>0.75$) and randomly replaced them with the remaining ones.
 Finally, because this operation effectively changes the weight of characters (i.e.
@@
 randomly weighting characters).
+\hl{Each replaced characters (i.e. the characters with a weight of +$1$) where randomly selected from the remaining characters using a uniform distribution (0,1).}
 Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.

### 2018-06-19 Analysis/04-CheckMatrixCorrelation.Rmd
Added comparisons tests

@@
 ## Naming the lists
-c("25t", "75t", "100t") -> names(results_maxi) -> names(results_mini) -> names(results_rand)
+c("25t", "75t", "150t") -> names(results_maxi) -> names(results_mini) -> names(results_rand)
 c("100c", "350c", "1000c") -> names(results_maxi[[1]]) ->
@@
     axis(1, 1:9, labels = FALSE, tick = FALSE)
-    axis(1, c(2, 5, 8), tick = FALSE, labels = names(data))
+    axis(1, c(2, 5, 8), tick = FALSE, labels = names(data[[1]]))
     ## Lines
@@
 ```
+
+<!-- Figure version -->
+
+```{r, fig.height = 18, fig.width = 6, eval = TRUE, echo = FALSE}
+pdf("../Writing/Figures/Proportion_replaced_maxi.pdf", width = 6, height = 6)
+plot.proportions(results_maxi, main = "Maximised differences")
+pdf("../Writing/Figures/Proportion_replaced_mini.pdf", width = 6, height = 6)
+plot.proportions(results_mini, main = "Minimised differences")
+pdf("../Writing/Figures/Proportion_replaced_rand.pdf", width = 6, height = 6)
+plot.proportions(results_rand, main = "Randomised differences")
+```
+
+## Testing the differences in proportions
+
+We can then test the differences between categories.
+First we'll try an aov:
+
+```{r}
+## Data frame converter
+convert.df <- function(results) {
+    convert.sub.df <- function(list, names) {
+        df <- data.frame(unlist(list), as.factor(rep(names(list), each = length(list[[1]]))))
+        colnames(df) <- names
+        rownames(df) <- NULL
+        return(df)
+    }
+    df_list <- lapply(results, convert.sub.df, names = c("proportion", "characters"))
+    df_list <- cbind(do.call(rbind, df_list),
+                     "taxa" = as.factor(rep(names(results), each = nrow(df_list[[1]]))))
+    rownames(df_list) <- NULL
+    return(df_list)
+}
+
+## Generating the nested tables
+table_maxi <- convert.df(results_maxi)
+table_mini <- convert.df(results_mini)
+table_rand <- convert.df(results_rand)
+
+## Combining the tables
+all_results <- cbind(rbind(table_maxi, table_mini, table_rand),
+                    "scenario" = as.factor(rep(c("maxi", "mini", "rand"), each = nrow(table_maxi))))
+
+## Run the linear model for testing assumptions
+model <- lm(proportion ~ scenario, data = all_results)
+(residuals_normality <- shapiro.test(residuals(model)))
+(data_homostedasticity <- bartlett.test(proportion ~ scenario, data = all_results))
+
+## Normality assumption fails
+```
+
+However, the assumptions are violated so we need to go non-parametric (Wilcox test and Bhattacharyya Coefficient)
+
+```{r}
+## Pairwise tests wrapper
+pair.test <- function(data, factor, test) {
+    ## Separating the data according to factors
+    factor <- which(colnames(data) == factor)
+    splited_data <- split(data, data[, factor])
+
+    ## Combination of tests
+    combinations <- combn(1:length(splited_data), 2)
+
+    ## Applying the tests
+    tests <- apply(combinations, 2, function(X, test, data)
+                 return(test(data[[X[1]]][,1], data[[X[2]]][,1])), test = test, data = splited_data)
+    names_test <- apply(combinations, 2, function(X, data)
+                   return(paste0(names(data)[X[1]], " - ", names(data)[X[2]])), data = splited_data)
+
+    names(tests) <- names_test
+    return(tests)
+}
+
+## Running the wilcox tests for each pooled factor
+wilcox_scen <- pair.test(all_results, factor = "scenario", wilcox.test)
+wilcox_taxa <- pair.test(all_results, factor = "taxa", wilcox.test)
+wilcox_char <- pair.test(all_results, factor = "characters", wilcox.test)
+
+## Extracting the statistic and the p_value only
+wilcox_scen_w <- unlist(lapply(wilcox_scen, function(X) return(unname(X$statistic))))
+wilcox_taxa_w <- unlist(lapply(wilcox_taxa, function(X) return(unname(X$statistic))))
+wilcox_char_w <- unlist(lapply(wilcox_char, function(X) return(unname(X$statistic))))
+wilcox_scen_p <- unlist(lapply(wilcox_scen, function(X) return(X$p.value)))
+wilcox_taxa_p <- unlist(lapply(wilcox_taxa, function(X) return(X$p.value)))
+wilcox_char_p <- unlist(lapply(wilcox_char, function(X) return(X$p.value)))
+
+## Running the Bhattacharrya tests for each pooled factor
+bhattc_scen <- pair.test(all_results, factor = "scenario", bhatt.coeff)
+bhattc_taxa <- pair.test(all_results, factor = "taxa", bhatt.coeff)
+bhattc_char <- pair.test(all_results, factor = "characters", bhatt.coeff)
+
+```
+
+And here are the results:
+
+```{r, echo = FALSE}
+kable(data.frame("Statistic" = wilcox_scen_w, "p_value" = wilcox_scen_p), digits = 3, 
+      caption = "Pairwise Wilcoxon test for the different number of scenarios.")
+kable(data.frame("bhatt.coeff" = bhattc_scen), digits = 3, 
+      caption = "Pairwise Bhattacharrya Coefficient for the different number of scenarios.")
+
+kable(data.frame("Statistic" = wilcox_taxa_w, "p_value" = wilcox_taxa_p), digits = 3, 
+      caption = "Pairwise Wilcoxon test for the different number of taxa.")
+kable(data.frame("bhatt.coeff" = bhattc_taxa), digits = 3, 
+      caption = "Pairwise Bhattacharrya Coefficient for the different number of taxa.")
+
+kable(data.frame("Statistic" = wilcox_char_w, "p_value" = wilcox_char_p), digits = 3, 
+      caption = "Pairwise Wilcoxon test for the different number of characters.")
+kable(data.frame("bhatt.coeff" = bhattc_char), digits = 3, 
+      caption = "Pairwise Bhattacharrya Coefficient for the different number of characters.")
+```

### 2018-06-20 Analysis/04-CheckMatrixCorrelation.Rmd
added libraries
Added statistics table

@@
 ```{r}
+library(dispRity)
 library(ape)
+library(knitr)
 ## Setting the path
@@
 wilcox_char_w <- unlist(lapply(wilcox_char, function(X) return(unname(X$statistic))))
-wilcox_scen_p <- unlist(lapply(wilcox_scen, function(X) return(X$p.value)))
-wilcox_taxa_p <- unlist(lapply(wilcox_taxa, function(X) return(X$p.value)))
-wilcox_char_p <- unlist(lapply(wilcox_char, function(X) return(X$p.value)))
+wilcox_scen_p <- p.adjust(unlist(lapply(wilcox_scen, function(X) return(X$p.value))), "bonferroni")
+wilcox_taxa_p <- p.adjust(unlist(lapply(wilcox_taxa, function(X) return(X$p.value))), "bonferroni")
+wilcox_char_p <- p.adjust(unlist(lapply(wilcox_char, function(X) return(X$p.value))), "bonferroni")
 
@@
 
+## Combine the results
+combined_results <- as.data.frame(rbind(
+        cbind("statistic" = wilcox_scen_w, "p.value" = wilcox_scen_p, "bhatt.coeff" = bhattc_scen),
+        cbind("statistic" = wilcox_taxa_w, "p.value" = wilcox_taxa_p, "bhatt.coeff" = bhattc_taxa),
+        cbind("statistic" = wilcox_char_w, "p.value" = wilcox_char_p, "bhatt.coeff" = bhattc_char)))
 ```
@@
 ```{r, echo = FALSE}
-kable(data.frame("Statistic" = wilcox_scen_w, "p_value" = wilcox_scen_p), digits = 3, 
-      caption = "Pairwise Wilcoxon test for the different number of scenarios.")
-kable(data.frame("bhatt.coeff" = bhattc_scen), digits = 3, 
-      caption = "Pairwise Bhattacharrya Coefficient for the different number of scenarios.")
+caption <- paste0("Pairwise Wilcoxon test and Bhattacharrya Coefficients for the difference in",
+  " proportion of character replacement for the different scenarios, number of taxa and characters.")
+kable(combined_results, digits = 3, caption = caption)
+```
 
-kable(data.frame("Statistic" = wilcox_taxa_w, "p_value" = wilcox_taxa_p), digits = 3, 
-      caption = "Pairwise Wilcoxon test for the different number of taxa.")
-kable(data.frame("bhatt.coeff" = bhattc_taxa), digits = 3, 
-      caption = "Pairwise Bhattacharrya Coefficient for the different number of taxa.")
-
-kable(data.frame("Statistic" = wilcox_char_w, "p_value" = wilcox_char_p), digits = 3, 
-      caption = "Pairwise Wilcoxon test for the different number of characters.")
-kable(data.frame("bhatt.coeff" = bhattc_char), digits = 3, 
-      caption = "Pairwise Bhattacharrya Coefficient for the different number of characters.")
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
+library(xtable)
+source("../Functions/make.xtable.R")
+make.xtable(combined_results, caption = caption, label = "Tab_proportion_replacement",
+            path = "../Writing/Tables/", include.rownames = TRUE)
 ```

### 2018-06-20 Functions/make.xtable.R
Updated writing table to work for the proportion tests

@@ -8,6 +8,7 @@
 #' @param label the label (in LaTeX: \ref{label})
 #' @param longtable whether to use longtable
 #' @param path optional, a path for saving the table
+#' @param include.rownames logical, whether to include the rownames (default = FALSE)
 #' 
 #' @examples
 #'
@@ -16,10 +17,13 @@
 #' @author Thomas Guillerme
 #' @export
 
-make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE, path) {
+make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE, path, include.rownames = FALSE) {
+
+    ## Rounding
+    table <- round(table, digits = digit)
 
     ## Add significance values
-    if(all(names(table) == c("metric", "test", "bhatt.coeff", "statistic", "p.value"))){
+    if(all(!is.na(match(colnames(table), c("bhatt.coeff", "statistic", "p.value"))))){
         for(row in 1:nrow(table)) {
             if(as.numeric(table$bhatt.coeff[row]) >= 0.95 || as.numeric(table$bhatt.coeff[row]) <= 0.05) {
                 table$bhatt.coeff[row] <- paste0("BOLD",table$bhatt.coeff[row])
@@ -37,7 +41,7 @@ make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE, pat
     textable <- xtable(table, digit = digit, caption = caption, label = label)
 
     ## Change attributes
-    if(all(names(table) == c("metric", "test", "bhatt.coeff", "statistic", "p.value"))){
+    if(all(colnames(table) == c("metric", "test", "bhatt.coeff", "statistic", "p.value"))){
         ## Add a horizontal line
         attr(textable, "align")[4] <- "r|"
     }
@@ -45,15 +49,15 @@ make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE, pat
 
     if(!missing(path)) {
         if(longtable == TRUE) {
-            cat(print(textable, tabular.environment = 'longtable', floating = FALSE, include.rownames = FALSE, sanitize.text.function = bold.cells), file = paste0(path, label, ".tex"))
+            cat(print(textable, tabular.environment = 'longtable', floating = FALSE, include.rownames = include.rownames, sanitize.text.function = bold.cells), file = paste0(path, label, ".tex"))
         } else {
-            cat(print(textable, include.rownames = FALSE, sanitize.text.function = bold.cells), file = paste0(path, label, ".tex"))
[…]

### 2018-06-20 Writing/response_to_reviewers.tex
response up

@@
 
-% For improving clarity in this document, reviewers' comments are displayed in blue, my responses in normal text and the changes to the manuscript in italic.
-% Additionally, I've uploaded two versions of the revised manuscript: both have the exact same content but one has all the changes and additions to the main text highlighted in yellow to help with reviewing.
+For improving clarity in this document, reviewers' comments are displayed in blue, our responses in normal text and the changes to the manuscript in italic.
+Additionally, we've uploaded two versions of the revised manuscript: both have the exact same content but one has all the changes and additions to the main text highlighted in yellow to help with reviewing.
 
@@
 
+We've added the proportion of replaced characters in each matrices for each ``minimised'', ``maximised'' and ``randomised'' scenarios in the Supplementary material 3, figures @@@.
+It is worth noting that there is a significant difference in terms of number of number of replaced characters in each scenarios (the ``minimised'' matrices have, on average, more characters replaced than the ``randomised'' that has in turns more characters replaced than the ``maximised'') but non in terms of number of taxa or characters.
+
+This can lead the
 %TG: TODO: add the range of replaced character n and m
@@
 
-%TG: Not sure about this one either: I guess its impossible to say both in simulations or in real life: in the end we only have the tips and their information.
+We agree with this reviewer's comment which can be especially thought to be the case for discrete with a fast rate of evolution.
+Such characters will likely swap between states along the branches following a hidden Markov chain.
+Unfortunately, however, although the state of a character on a branch can be modelled, it is nearly always impossible to actually observe them.
+
+We've added this remark in the caveats section:
+\textit{It is also worth noting that character similarity at the tips does not necessarily reflects that the characters underwent the same evolutionary history.
+This can specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches (e.g.
+Revell, 2014).
+However, it is worth noting that in empirical studies, such state changes along the branches are nearly always unknown.}
 
@@
 
+1- Covariance is approximated by using the same tree for the simulation
+2- The Character difference is an effective metric for measuring covariance for discrete characters.
+This is better than a simple correlation measurement because it makes it a distance it has the properties of a distance metric (which is not the case for correlations with multiple states)
+
+For example, for three characters \texttt{A = \{1,2,0,0,0\}}, \texttt{B = \{2,3,4,4,4\}} and \texttt{C = \{1,2,1,0,0\}}, $CD_{(A, C)} \leq CD_{(A, B)} + CD_{(B, C)}$ ($0.4 \leq 0 + 0.4$) but this is not the case for a correlation measurement where $cor_{(A, C)} > cor_{(A, B)} + cor_{(B, C)}$ ($0.868 \leq -0.687 + -0.534$).
+Furthermore, using a correlation metric, two character with the same splits (e.g.
+A and B).
+%TG: DOUBLE CHECK THIS
 %TG: Maybe say something along the lines: "yeah, great idea we'll do it later"?
@@
 
-%TG: need a way to phrase this but totally agree, in the reviewer's example, the matrix would look "honets" (lots of characters) but the results would be actually based on way fewer.
+%TG: need a way to phrase this but totally agree, in the reviewer's example, the matrix would look "honest" (lots of characters) but the results would be actually based on way fewer.
 

### 2018-06-20 Writing/draft_highlighted.tex
response up

@@
 This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.
+\hl{It is also worth noting that character similarity at the tips does not necessarily reflects that the characters underwent the same evolutionary history.
+This can specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches}
+ \citep[e.g.][]{revell2014ancestral}.
+\hl{However, it is worth noting that in empirical studies, such state changes along the branches are nearly always unknown.}
 

### 2018-06-20 Writing/response_to_reviewers.tex
updated table and figures order

@@
 
-We've added the comparisons between the ``normal'' and ``randomised'' trees to the ``true'' tree in the supplementary material 3, figures 6 to 8.
+We've added the comparisons between the ``normal'' and ``randomised'' trees to the ``true'' tree in the supplementary material 3, figures 9 to 11.
 We've added a reference to these figures in the main text:
 
-\textit{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 6, 7 and 8.} lines @@@
+\textit{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 9, 10 and 11.} lines @@@
 
@@
 
-The significances was referring to the results in Table 5 (Wilcox test p value of 0 and 0.08 and Bhattacharrya Coefficient of 0.891 and 0.984 respectively) and the median values were extracted from the supplementary material 3 Table 4.
+The significances was referring to the results in Table 5 (Wilcox test p value of 0 and 0.08 and Bhattacharrya Coefficient of 0.891 and 0.984 respectively) and the median values were extracted from the supplementary material 3 Table 5.
 We now detailed the values and where they are present for each of the tests to avoid confusion.

### 2018-06-20 Writing/draft_highlighted.tex
updated table and figures order

@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 1 and 2).
+We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 4 and 5).
 Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
@@
 When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
-\hl{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 6, 7 and 8.}
+\hl{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 9, 10 and 11.}
 
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839\hl{; see supplementary materials 3, Table 1}) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628\hl{; see supplementary materials 3, Table 1}) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303\hl{; see supplementary materials 3, Table 1}).
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839\hl{; see supplementary materials 3, Table 2}) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628\hl{; see supplementary materials 3, Table 2}) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303\hl{; see supplementary materials 3, Table 2}).
 There is a significant difference between all scenarios (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test})
@@
 This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
-The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 2}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).
-The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3 \hl{Table 2}) with a significant difference between 100 and 350 or 1000 characters (\hl{\textit{p} = 0, \textit{p} = 0 and \textit{p} = 0.014 respectively}; Table \ref{Tab_pooledscharacters_test}).
+The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 3}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).
+The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3 \hl{Table 3}) with a significant difference between 100 and 350 or 1000 characters (\hl{\textit{p} = 0, \textit{p} = 0 and \textit{p} = 0.014 respectively}; Table \ref{Tab_pooledscharacters_test}).
 However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.
@@
 Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
-A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (\hl{\textit{p} = 0.012, \textit{p} = 0.004, \textit{p} = 0 and \textit{p} = 0 respectively}; Table \ref{Tab_pooledstaxa_test} and medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively; see supplementary materials 3 \hl{Table 3}).
+A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (\hl{\textit{p} = 0.012, \textit{p} = 0.004, \textit{p} = 0 and \textit{p} = 0 respectively}; Table \ref{Tab_pooledstaxa_test} and medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively; see supplementary materials 3 \hl{Table 4}).
 Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
@@
 
-Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\hl{\textit{p} = 0} Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively; \hl{see supplementary materials 3, Table 4}) but not in terms of individual taxon placements (\hl{\textit{p} = 0.08}; Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively; \hl{see supplementary materials 3, Table 4}).
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\hl{\textit{p} = 0} Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively; \hl{see supplementary materials 3, Table 5}) but not in terms of individual taxon placements (\hl{\textit{p} = 0.08}; Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively; \hl{see supplementary materials 3, Table 5}).
 
@@
 Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
-Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary material 3, Figs 3, 4 and 5).
+Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 6, 7 and 8}).
 

### 2018-06-20 Writing/response_to_reviewers.tex
Reviewer 1 nearly done?

@@
 
-We've added the proportion of replaced characters in each matrices for each ``minimised'', ``maximised'' and ``randomised'' scenarios in the Supplementary material 3, figures @@@.
-It is worth noting that there is a significant difference in terms of number of number of replaced characters in each scenarios (the ``minimised'' matrices have, on average, more characters replaced than the ``randomised'' that has in turns more characters replaced than the ``maximised'') but non in terms of number of taxa or characters.
+We've added the proportion of replaced characters in each matrices for each ``minimised'', ``maximised'' and ``randomised'' scenarios in the supplementary material 3, figures 1, 2 and 3.
+It is worth noting that there is a significant difference in terms of number of replaced characters in each scenarios between the ``maximised'' and the ``minimised'' and ``randomised'' matrices but non in terms of number of taxa or characters.
 
-This can lead the
-%TG: TODO: add the range of replaced character n and m
-%TG: TODO: add more explanantion on the replacements in the methods and the discussion (the reviewer's thoughts)
+This could potentially explain some of the differences between the scenarios (i.e.
+why it seems better to maximise character difference rather than minimising it) but does not explain the differences in smaller matrices nor between the ``minimised'' and ``randomised'' scenarios and neither between the two inference methods (since they both use the same matrices).
+Furthermore, the maximising/minimising character differences was done \textit{a priori} without any targeted number of characters to replace to make the analysis less idiosynchratic.
+In other words, we decided to modify the randomly simulated matrices to globally increase/decrease character differences which is expected to decrease anyway in bigger matrices.
+When using a small number of character states (i.e.
+two or three), the matrices with few taxa and many characters are expected to have an overall low character difference because of the smaller combinatoric possibilities for each characters.
+
+We've added mentions to the proportion of replaced characters analysis in the supplementary in the main text at lines @@@ and added a paragraph explaining this caveat in the discussion:
+
+%TG: TODO: added pragraph!
+\textit{.} lines @@@
 
@@
 
-%TG: add some text to the manuscript parphrasing this? It's kind of already written in there...
+%TG: add some text to the manuscript paraphrasing this? It's kind of already written in there...
 
@@
 If the simulations utilised an explicit distribution of character covariance then I think that interpretation of the results would have been more straightforward.}}
+
+% As noted in the introduction, the aim of this manuscript was mainly to see how do software handle correlation between discrete morphological characters.
+% In other words, how can we distinguish between the complex sources of correlation in morphology
 

### 2018-06-28 Writing/response_to_reviewers.tex
Response updates

@@
 
-% As noted in the introduction, the aim of this manuscript was mainly to see how do software handle correlation between discrete morphological characters.
-% In other words, how can we distinguish between the complex sources of correlation in morphology
+As noted in the introduction, the aim of this manuscript was mainly to see how do software handle correlation between discrete morphological characters.
+In other words, how can we distinguish between the complex sources of correlation in phenotype using only discrete morphological characters?
+We do agree with the reviewer that a thorough analysis of the variance/covariance structure of simulated ``discretised'' characters (see reviewer 3, major suggestion \ref{hiden}) would have broad more insight into the nature of character correlation on phylogenetics inference in general (regardless of the type of character and correlation) but we feel that this is out of the scope of this study.
+
+We have however made clearer in the introduction what were the objectives of this current manuscript:
 
@@
 It is conceivable that two characters can appear to have near perfect correlation at the tips, but arrived at these states through completely independent pathways, particularly when simulating with a continuous time Markov chain where unobserved transitions to a new state and back to the original state are possible.}}
+\label{hiden}
 

### 2018-06-28 Writing/draft_highlighted.tex
Response updates

@@
 One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.}
-
+% \hl{Furthermore, phenotypic variations can be continuous rather than discrete and ``discretising'' such characters can have some influence on tree inference} \citep{parins2017use}.
+% \hl{In this case, ``discretised'' }.
 This illustrates how some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
@@
 
-How \hl{do} these correlation really affect topology?
-We expect matrices with a high level of correlation to recover precise but inaccurate topologies but will matrices with low level of correlation (i.e.
-with high levels of homoplasy) actual cancel out the effects of correlation?
+How \hl{do} these \hl{specific} correlation really affect topology?
+% In the specific case of discrete morphological characters, how does the reduc phenotypic information and its different sources of correlation is 
+% More specifically, how can phylogenetic inference software distinguish between the complex sources of correlation tha
+% In other words, how can we distinguish between the complex sources of correlation in phenotype using only discrete morphological characters?
+% \cite{parins2017use}
 Here we formally assess the effect of discrete character's correlation using simulated data.

### 2018-11-27 Functions/shells/set.jobs.sh
added /30days/ save

@@ -23,7 +23,8 @@ chain1=$1
 method=$2
 CPU=$3
 job_time=$4
-chain2=$5
+username=$5
+chain2=$6
 
 ## Set up the methods
 if echo $method | grep 'both' > /dev/null
@@ -81,7 +82,7 @@ then
     echo "#PBS -A UQ-SCI-BiolSci" >> ${chain1}.template
     echo "#-#PBS -l nodes=1:ppn=${CPU},mem=32gb,vmem=32GB,walltime=${job_time}:00:00" >> ${chain1}.template
     echo "#-#PBS -m n" >> ${chain1}.template
-    echo "#PBS -N TEMPLATE" >> ${chain1}.template
+    echo "#PBS -N TEMPLATE.mbcmd" >> ${chain1}.template
     echo "#PBS -l select=1:ncpus=${CPU}:mpiprocs=${CPU}:mem=32gb:vmem=32GB,walltime=${job_time}:00:00" >> ${chain1}.template
 
     echo "## Load modules" >> ${chain1}.template
@@ -92,20 +93,25 @@ then
     echo "## Entry time" >> ${chain1}.template
     echo "echo \"Entry time\"" >> ${chain1}.template
     echo "date" >> ${chain1}.template
-    echo "" >> ${chain1}.template
 
     echo "## Run the script" >> ${chain1}.template
-    echo "mpiexec -np ${CPU} mb \$HOME/CharSim/TEMPLATE" >> ${chain1}.template
+    echo "mpiexec -np ${CPU} mb \$HOME/CharSim/TEMPLATE.mbcmd" >> ${chain1}.template
 
     echo "## Exit time" >> ${chain1}.template
     echo "echo \"Exit time\"" >> ${chain1}.template
     echo "date" >> ${chain1}.template
 
+    echo "## Transfer files" >> ${chain1}.template
+    echo "mv \$HOME/CharSim/TEMPLATE* /30days/${username}/" >> ${chain1}.template
+    echo "echo \"File transfer OK\"" >> ${chain1}.template
+
+
+
     ## Split the template into the multiple chain jobs
-    sed 's/TEMPLATE/'"${chain1}"'_norm.mbcmd/g' ${chain1}.template > ${chain1}_norm.mbjob
[…]

### 2018-11-27 Writing/draft_highlighted.tex
Updated formula + added note for similarity
awoonga

@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, Abigail I. Pastore$^{1}$, and Martin D. Brazeau$^{2,3}$}\\
 \noindent {\small \it 
@@
 \begin{equation}
-    CD_{(x,y)} = 1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}\right)
+    CD_{(x,y)} = \abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}}
 \end{equation}
+
+%CD is similar to https://onlinelibrary.wiley.com/doi/full/10.1111/pala.12384 but by character rather than by species
 
@@
 \section{Acknowledgements}
-Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
+Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232). @@@ plus awoonga
 We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes and Abigail Pastore for helping with the CD metric proof and the \hl{three anonymous reviewers for there insightful comments}.

### 2018-12-27 Writing/draft_highlighted.tex
Up acknow

@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, Abigail I. Pastore$^{1}$, and Martin D. Brazeau$^{2,3}$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
 \noindent {\small \it 
@@
 Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232). @@@ plus awoonga
-We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes and Abigail Pastore for helping with the CD metric proof and the \hl{three anonymous reviewers for there insightful comments}.
+We thank Alberto Pascual Garcia for input in the design of the simulations protocol, Guillermo Herraiz Yebes and Abigail Pastore for helping with the CD metric proof and Orlin Todorov for assistance with the computations.
+We also thank the \hl{three anonymous reviewers for there constructive and insightful comments}.
 

### 2019-03-01 Functions/make.xtable.R
Updated functions for new data

@@ -20,7 +20,7 @@
 make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE, path, include.rownames = FALSE) {
 
     ## Rounding
-    table <- round(table, digits = digit)
+    table[,which(unlist(lapply(table, class)) == "numeric")] <- round(table[,which(unlist(lapply(table, class)) == "numeric")], digits = digit)
 
     ## Add significance values
     if(all(!is.na(match(colnames(table), c("bhatt.coeff", "statistic", "p.value"))))){

### 2019-03-01 Functions/read.TreeCmp.R

@@ -4,8 +4,8 @@
 #'
 #' @param chain the general name of the chain (e.g. 25t_100c)
 #' @param path the path to where the files are stored
-#' @param type which type of tree (Bayesian or Parismony)
-#' @param length.in how many chains to read (e.g. 35)
+#' @param type which type of tree (Bayesian or Parsimony)
+#' @param length.in how many chains to read (e.g. 20) - can be left missing to read all the available files
 #'
 #' @return
 #' A list of metrics for the comparisons.

### 2019-03-01 Functions/shells/get.con.tree.sh

@@ -7,7 +7,7 @@ PATH_OUT="/Users/TGuillerme/Projects/CharactersCorrelation/Data/Trees_out/Consen
 for folder in ${chain_prefix}*
 do
     ## Get the chain name
-    chain_name=$(echo ${folder} | sed 's/\/Volumes\/CharacterCorrelation\/CharacterCorrelationData\///g')
+    chain_name=$(echo ${folder} | sed 's/\/Volumes\/CharacterCorrelation\/CharacterCorrelationNew\///g')
     
     list=$(ls ${PATH_OUT})
     if echo $list | grep -q ${chain_prefix}

### 2019-04-16 Analysis/Character_difference/02-Modifying_matrices.Rmd
Updated analysis

@@
 ## Seed for repeatability
-set.seed(0)
+set.seed(123)
 ```
@@
 ## Character difference
-char_diff_matrix_maxi <- char.diff(matrix_maxi)
+char_diff_matrix_maxi <- char.diff(matrix_maxi[[1]])
 par(mfrow = c(1, 2))
@@
 ## Character difference
-char_diff_matrix_mini <- char.diff(matrix_mini)
+char_diff_matrix_mini <- char.diff(matrix_mini[[1]])
 par(mfrow = c(1, 2))
@@
 ## Character difference
-char_diff_matrix_rand <- char.diff(matrix_rand)
+char_diff_matrix_rand <- char.diff(matrix_rand[[1]])
 par(mfrow = c(1, 2))
@@
 ```{r, fig.width=12, fig.height=12, eval = FALSE, echo = FALSE}
-pdf("../../Writing/Figures/Modif_matrix.pdf", width = 7.2, height = 4.5)
+# pdf("../../Writing/Figures/Modif_matrix.pdf", width = 7.2, height = 4.5)
 par(mfrow = c(2, 3), bty = "n", mar = c(1,1,1,1))

### 2019-04-16 Writing/Tables/Full_Tab_pooledsmethods_test.tex
added the looooong tables
Updated analysis
Updated results

@@
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Tue Mar 27 13:20:38 2018
+% latex table generated in R 3.5.3 by xtable 1.8-3 package
+% Tue Apr 16 16:10:50 2019
 \begin{table}[ht]
@@
   \hline
-Best & RF & bayesian:parsimony & 0.891 & 579437.500 & 0.000 \\ 
-   &  & bayesian:parsimony & 0.984 & 470621.500 & 0.168 \\ 
-   & Tr & bayesian:parsimony & 0.847 & 657875.500 & 0.000 \\ 
-   &  & bayesian:parsimony & 0.975 & 499754.000 & 0.000 \\ 
-  Null & RF & bayesian:parsimony & 0.891 & 579437.500 & 0.000 \\ 
-   &  & bayesian:parsimony & 0.984 & 470621.500 & 0.168 \\ 
-   & Tr & bayesian:parsimony & 0.847 & 657875.500 & 0.000 \\ 
-   &  & bayesian:parsimony & 0.975 & 499754.000 & 0.000 \\ 
+Best & RF & bayesian:parsimony & 0.939 & 192659.000 & 0.000 \\ 
+   &  & bayesian:parsimony & 0.986 & 157677.500 & 0.082 \\ 
+   & Tr & bayesian:parsimony & 0.944 & 193142.000 & 0.000 \\ 
+   &  & bayesian:parsimony & 0.975 & 161621.000 & 0.008 \\ 
+  Null & RF & bayesian:parsimony & 0.939 & 192659.000 & 0.000 \\ 
+   &  & bayesian:parsimony & 0.986 & 157677.500 & 0.082 \\ 
+   & Tr & bayesian:parsimony & 0.944 & 193142.000 & 0.000 \\ 
+   &  & bayesian:parsimony & 0.975 & 161621.000 & 0.008 \\ 
    \hline

### 2019-04-16 Writing/Tables/Tab_pooledmethod.tex
Updated results

@@
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Tue Mar 27 14:03:14 2018
+% latex table generated in R 3.5.3 by xtable 1.8-3 package
+% Tue Apr 16 16:08:54 2019
 \begin{table}[ht]
@@
   \hline
-RF & bayesian & 0.319 & 0.714 & 0.828 & 0.811 & 0.934 & 1.000 \\ 
-   & parsimony & 0.076 & 0.496 & 0.679 & 0.684 & 0.911 & 1.000 \\ 
-  Tr & bayesian & -0.473 & 0.354 & 0.738 & 0.618 & 0.947 & 1.000 \\ 
-   & parsimony & -0.436 & 0.258 & 0.601 & 0.568 & 0.949 & 1.000 \\ 
+RF & bayesian & 0.385 & 0.753 & 0.851 & 0.824 & 0.924 & 1.000 \\ 
+   & parsimony & 0.371 & 0.642 & 0.739 & 0.742 & 0.865 & 1.000 \\ 
+  Tr & bayesian & -0.196 & 0.560 & 0.802 & 0.717 & 0.945 & 1.000 \\ 
+   & parsimony & -0.155 & 0.525 & 0.750 & 0.693 & 0.900 & 1.000 \\ 
    \hline

### 2019-04-16 Writing/Tables/Tab_pooledscenarios_test.tex
Updated results

--- 2018-03-27 Writing/Tables/Tab_pooledscenarios_test.tex
@@
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Tue Mar 27 14:03:13 2018
+% latex table generated in R 3.5.3 by xtable 1.8-3 package
+% Tue Apr 16 16:07:23 2019
 \begin{table}[ht]
 \centering
-\begin{tabular}{llr|rl}
+\begin{tabular}{llr|rr}
   \hline
@@
   \hline
-RF & maxi:mini & 0.573 & 356436.000 & \textbf{0} \\ 
-   & maxi:rand & 0.873 & 287225.000 & \textbf{0} \\ 
-   & mini:rand & 0.856 & 95841.500 & \textbf{0} \\ 
-  Tr & maxi:mini & 0.614 & 358800.000 & \textbf{0} \\ 
-   & maxi:rand & 0.908 & 288223.500 & \textbf{0} \\ 
-   & mini:rand & 0.858 & 98507.500 & \textbf{0} \\ 
+RF & maxi:mini & 0.848 & 32012.500 & 0.000 \\ 
+   & maxi:rand & 0.941 & 44161.000 & 0.000 \\ 
+   & mini:rand & 0.954 & 80382.500 & 0.000 \\ 
+  Tr & maxi:mini & 0.992 & 60680.500 & 0.839 \\ 
+   & maxi:rand & 0.980 & 60447.000 & 0.713 \\ 
+   & mini:rand & 0.982 & 64222.500 & 1.000 \\ 
    \hline

### 2019-04-16 Writing/Tables/Tab_pooledscenarios_test.tex
bold main results

@@
   \hline
-RF & maxi:mini & 0.848 & 32012.500 & 0.000 \\ 
-   & maxi:rand & 0.941 & 44161.000 & 0.000 \\ 
-   & mini:rand & 0.954 & 80382.500 & 0.000 \\ 
-  Tr & maxi:mini & 0.992 & 60680.500 & 0.839 \\ 
-   & maxi:rand & 0.980 & 60447.000 & 0.713 \\ 
-   & mini:rand & 0.982 & 64222.500 & 1.000 \\ 
+RF & maxi:mini & 0.848 & 32012.500 & \textbf{0.000} \\ 
+   & maxi:rand & 0.941 & 44161.000 & \textbf{0.000} \\ 
+   & mini:rand & \textbf{0.954} & 80382.500 & \textbf{0.000} \\ 
+  Tr & maxi:mini & \textbf{0.992} & 60680.500 & 0.839 \\ 
+   & maxi:rand & \textbf{0.980} & 60447.000 & 0.713 \\ 
+   & mini:rand & \textbf{0.982} & 64222.500 & 1.000 \\ 
    \hline

### 2019-04-17 Analysis/04-CheckMatrixCorrelation.Rmd
New character correlation check

@@
 
-# Checking the number of character replaced in each matrix
+# Checking the number of character replaced in each matrix and the change in character correlation
 
@@
 ## Setting the path
-PATH <- "../Data/Simulations/Matrices/"
+PATH <- "../Data/Simulations/"
 ```
 
-The following function counts, _a posteriori_ the number of characters that where replaced in each matrix (by comparing them to the normal matrix)
+The following function counts the effective change in number of characters for each matrix: the average character correlation change, their distributions, the number of changed characters, and their character correlation change.
+
+## Load the data
+
+We can use the following functions to load the already estimated parameters (`read.matrix.param`) during the simulations or directly the matrices (`read.matrix`) to specifically double check the parameters (not in this document).
 
 ```{r}
-## Getting the proportion of replaced characters
-get.proportion.replace <- function(matrix, path) {
-    ## Getting one matrix in a comparable format
-    get.one.matrix <- function(path, matrix, type) {
-        return(unlist(apply(do.call(rbind,
-                                    read.nexus.data(paste0(path, matrix, "_", type, ".nex"))
-                            ), 2, list), recursive = FALSE))
+## Loading a matrix
+read.matrix <- function(chain_name, path) {
+    ## Read the four matrices
+    norm <- do.call(rbind, read.nexus.data(paste0(path, chain_name, "_norm.nex")))
+    maxi <- do.call(rbind, read.nexus.data(paste0(path, chain_name, "_maxi.nex")))
+    mini <- do.call(rbind, read.nexus.data(paste0(path, chain_name, "_mini.nex")))
+    rand <- do.call(rbind, read.nexus.data(paste0(path, chain_name, "_rand.nex")))
+    return(list(norm = norm, maxi = maxi, mini = mini, rand = rand))
 }
 
-    ## Getting all the matrices
-    norm <- get.one.matrix(path, matrix, "norm")
-    maxi <- get.one.matrix(path, matrix, "maxi")
-    mini <- get.one.matrix(path, matrix, "mini")
-    rand <- get.one.matrix(path, matrix, "rand")
+## Loading a matrix parameter
+read.matrix.param <- function(chain_name, path) {
+    ## Reading the parameters
+    maxi <- read.csv(paste0(path, chain_name, "_changes_maxi.csv"), row.names = 1)
+    mini <- read.csv(paste0(path, chain_name, "_changes_mini.csv"), row.names = 1)
+    rand <- read.csv(paste0(path, chain_name, "_changes_rand.csv"), row.names = 1)
 
-    ## Comparing all the matrices
-    n <- length(norm)
-    maxi <- length(which(!mapply(function(X, Y) all(X == Y), norm, maxi)))/n
-    mini <- length(which(!mapply(function(X, Y) all(X == Y), norm, mini)))/n
-    rand <- length(which(!mapply(function(X, Y) all(X == Y), norm, rand)))/n
+    ## Splitting the parameters
+    char_diff_change <- rbind(maxi = maxi[1,], mini = mini[1,], rand = rand[1,])
+    colnames(char_diff_change) <- c("original", "modified")
 
-    return(c("maxi" = maxi, "mini" = mini, "rand" = rand))
+    ## Getting the list of changes
+    characters_changes <- list(maxi = maxi[-1,], mini = mini[-1,], rand = rand[-1,])
+
+    ## Return both
+    return(list(differences = char_diff_change, changes = characters_changes))
 }
@@
 
-We can then apply it to the list of matrices
+We can load all the parameters arranged into a list of list of scenarios (e.g. 25 taxa and 100 characters, etc.):
 
-```{r}
+```{r, echo = FALSE}
 ## Creating the placeholder list for the results
-replacements <- list(list(), list(), list())
+parameters <- list(list(), list(), list())
 ## getting the values to loop through
@@
 character_val <- c(100, 350, 1000)
+## Set the path
+param_path <- paste0(PATH, "Parameters/")
+
+## Loading the parameters
 for(taxa in 1:3) {
     for(character in 1:3) {
-        ## Get the taxa names
-        names <- unique(gsub("_[a-z]*.nex", "",
-          list.files(PATH, pattern = paste0(taxa_val[taxa], "t_", character_val[character], "c_"))))
-        ## Get the proportions of replaced characters
-        replacements[[taxa]][[character]] <- lapply(as.list(names), get.proportion.replace, PATH)
+        ## Get the chain names
+        chain_names <- list.files(path = param_path, pattern = paste0(taxa_val[taxa], "t_", character_val[character], "c_"))
+        ## Remove the suffix
+        chain_names <- unique(gsub("_changes_\\S+.csv", "", chain_names, perl = TRUE))
+
+        ## Get the matrix change parameters
+        parameters[[taxa]][[character]] <- lapply(as.list(chain_names), read.matrix.param, path = param_path)
     }
+    names(parameters[[taxa]]) <- paste0(character_val, "c")
 }
-
-## Combining the results
-fun.maxi <- function(list) {
-    return(unlist(lapply(list, function(X) return(X[[1]]))))
-}
-fun.mini <- function(list) {
-    return(unlist(lapply(list, function(X) return(X[[2]]))))
-}
-fun.rand <- function(list) {
-    return(unlist(lapply(list, function(X) return(X[[3]]))))
-}
-results_maxi <- lapply(replacements, lapply, fun.maxi)
-results_mini <- lapply(replacements, lapply, fun.mini)
-results_rand <- lapply(replacements, lapply, fun.rand)
-
-## Naming the lists
-c("25t", "75t", "150t") -> names(results_maxi) -> names(results_mini) -> names(results_rand)
-c("100c", "350c", "1000c") -> names(results_maxi[[1]]) ->
-                              names(results_maxi[[2]]) ->
-                              names(results_maxi[[3]]) ->
-                              names(results_mini[[1]]) ->
-                              names(results_mini[[2]]) ->
-                              names(results_mini[[3]]) ->
-                              names(results_rand[[1]]) ->
-                              names(results_rand[[2]]) ->
-                              names(results_rand[[3]])
+names(parameters) <- paste0(taxa_val, "t")
 ```
 
-We can then represent the results using box plots:
+## Average character difference change
+
+Getting the difference between the character difference in the normal matrix and the one in the modified ones:
+
+```{r}
+## Get character difference change value
+get.char.diff.change <- function(chain) {
+    return(chain$differences)
+}
+parameters_difference <- lapply(parameters, lapply, lapply, get.char.diff.change)
+
+## Get the differences
+parameters_diff <- lapply(parameters_difference, lapply, lapply, function(X) apply(X, 1, diff))
+```
+
+Plotting the results
 
 ```{r, fig.height = 18, fig.width = 6}
+## Extract the results for each type of scenarios
+extract.results <- function(list, what) {
+    return(unname(unlist(lapply(list, function(X) return(X[what])))))
+}
+differences_maxi <- lapply(parameters_diff, lapply, extract.results, what = "maxi")
+differences_mini <- lapply(parameters_diff, lapply, extract.results, what = "mini")
+differences_rand <- lapply(parameters_diff, lapply, extract.results, what = "rand")
+
+## overal_range (for plotting)
+overal_range <- range(c(unlist(differences_maxi), unlist(differences_mini), unlist(differences_rand)))
+
 ## Plotting the different proportion results
-plot.proportions <- function(data, ...) {
+plot.proportions <- function(data, ylim = c(-1, 1), xlim = c(0.8, 9.2), leg.pos, ylab, ...) {
     ## Default colours
-    cols <- gray.colors(3)
+    cols <- gray.colors(length(data))
     ## Boxes
-    boxplot(xlab = "", ylab = "Proportion of replacement", xaxt = "n", ylim = c(0, 1), col = cols,
-            c(data[[1]], data[[2]], data[[3]]), ...)
-    legend("topleft", legend = names(data), pch = 19, col = cols, bty = "n")
+    boxplot(NULL, xlab = "", ylab = ylab, xaxt = "n", yaxt = "n",  ylim = ylim, xlim = xlim, ...)
+    legend(leg.pos, legend = names(data), pch = 19, col = cols, bty = "n")
     ## X axis
@@
     ## Lines
-    abline(v = 3.5) ; abline(v = 6.5)
+    abline(v = 3.5) ; abline(v = 6.5) ; abline(h = 0, lty = 2)
+
+    ## Data
+    boxplot(c(data[[1]], data[[2]], data[[3]]), col = cols, add = TRUE, xaxt = "n")
 }
-
 par(mfrow = c(3,1))
-plot.proportions(results_maxi, main = "Maximised differences")
-plot.proportions(results_mini, main = "Minimised differences")
-plot.proportions(results_rand, main = "Randomised differences")
-
+plot.proportions(differences_maxi, ylim = overal_range, main = "Maximised", leg.pos = "bottomleft",
+    ylab = "Character differences change")
+plot.proportions(differences_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
+    ylab = "Character differences change")
+plot.proportions(differences_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft",
+    ylab = "Character differences change")
 ```
@@
 ```{r, fig.height = 18, fig.width = 6, eval = TRUE, echo = FALSE}
-pdf("../Writing/Figures/Proportion_replaced_maxi.pdf", width = 6, height = 6)
-plot.proportions(results_maxi, main = "Maximised differences")
-pdf("../Writing/Figures/Proportion_replaced_mini.pdf", width = 6, height = 6)
-plot.proportions(results_mini, main = "Minimised differences")
-pdf("../Writing/Figures/Proportion_replaced_rand.pdf", width = 6, height = 6)
-plot.proportions(results_rand, main = "Randomised differences")
+pdf("../Writing/Figures/Difference_change_maxi.pdf", width = 6, height = 6)
+plot.proportions(differences_maxi, ylim = overal_range, main = "Maximised", leg.pos = "bottomleft",
+    ylab = "Character differences change")
+pdf("../Writing/Figures/Difference_change_mini.pdf", width = 6, height = 6)
+plot.proportions(differences_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
+    ylab = "Character differences change")
+pdf("../Writing/Figures/Difference_change_rand.pdf", width = 6, height = 6)
+plot.proportions(differences_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft",
+    ylab = "Character differences change")
 ```
 
-## Testing the differences in proportions
+### Testing the differences in character difference
 
-We can then test the differences between categories.
-First we'll try an aov:
+We can test the differences in character differences using the combined Wilcoxon test and Bhattacharyya coefficients.
+
+First we need to convert it into a usable table format:
 
@@
 ## Generating the nested tables
-table_maxi <- convert.df(results_maxi)
-table_mini <- convert.df(results_mini)
-table_rand <- convert.df(results_rand)
+table_maxi <- convert.df(differences_maxi)
+table_mini <- convert.df(differences_mini)
+table_rand <- convert.df(differences_rand)
 
@@
                     "scenario" = as.factor(rep(c("maxi", "mini", "rand"), each = nrow(table_maxi))))
-
-## Run the linear model for testing assumptions
-model <- lm(proportion ~ scenario, data = all_results)
-(residuals_normality <- shapiro.test(residuals(model)))
-(data_homostedasticity <- bartlett.test(proportion ~ scenario, data = all_results))
-
-## Normality assumption fails
 ```
-
-However, the assumptions are violated so we need to go non-parametric (Wilcox test and Bhattacharyya Coefficient)
 
@@
 source("../Functions/make.xtable.R")
-make.xtable(combined_results, caption = caption, label = "Tab_proportion_replacement",
+make.xtable(combined_results, caption = caption, label = "Tab_difference_change",
             path = "../Writing/Tables/", include.rownames = TRUE)
 ```
+
+## Number of characters duplicated
+
+We can also look at the number of characters that are duplicated in each
+
+```{r}
+## Get character difference change value
+get.char.replace <- function(chain) {
+    ## Get the duplicates
+    all_duplicates <- lapply(chain$changes, function(X) return(length(unique(X[,2]))))
+    ## Make them proportional
+    return(mapply(function(x,y) return(x/y), all_duplicates, as.list(c(100, 350, 1000)),
+                 SIMPLIFY = FALSE))
+
+}
+## Get the number of duplicated characters
+parameters_duplicates <- lapply(parameters, lapply, lapply, get.char.replace)
+```
+
+Plotting the results
+
+```{r, fig.height = 18, fig.width = 6}
+## Extract the results for each type of scenarios
+duplicates_maxi <- lapply(parameters_duplicates, lapply, extract.results, what = "maxi")
+duplciates_mini <- lapply(parameters_duplicates, lapply, extract.results, what = "mini")
+duplicates_rand <- lapply(parameters_duplicates, lapply, extract.results, what = "rand")
+
+## overal_range (for plotting)
+overal_range <- range(c(unlist(duplicates_maxi), unlist(duplciates_mini), unlist(duplicates_rand)))
+
+## Plotting the different proportion results
+par(mfrow = c(3,1))
+plot.proportions(duplicates_maxi, ylim = overal_range, main = "Maximised", leg.pos = "topleft",
+    ylab = "Proportion of duplicated characters")
+plot.proportions(duplciates_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
+    ylab = "Proportion of duplicated characters")
+plot.proportions(duplicates_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft",
+    ylab = "Proportion of duplicated characters")
+```
+
+<!-- Figure version -->
+
+```{r, fig.height = 18, fig.width = 6, eval = TRUE, echo = FALSE}
+pdf("../Writing/Figures/Difference_change_maxi.pdf", width = 6, height = 6)
+plot.proportions(differences_maxi, ylim = overal_range, main = "Maximised", leg.pos = "bottomleft")
+pdf("../Writing/Figures/Difference_change_mini.pdf", width = 6, height = 6)
+plot.proportions(differences_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft")
+pdf("../Writing/Figures/Difference_change_rand.pdf", width = 6, height = 6)
+plot.proportions(differences_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft"))
+```
+
+### Testing the differences in character difference
+
+We can test the differences in character differences using the combined Wilcoxon test and Bhattacharyya coefficients.
+
+First we need to convert it into a usable table format:
+
+```{r}
+## Data frame converter
+convert.df <- function(results) {
+    convert.sub.df <- function(list, names) {
+        df <- data.frame(unlist(list), as.factor(rep(names(list), each = length(list[[1]]))))
+        colnames(df) <- names
+        rownames(df) <- NULL
+        return(df)
+    }
+    df_list <- lapply(results, convert.sub.df, names = c("proportion", "characters"))
+    df_list <- cbind(do.call(rbind, df_list),
+                     "taxa" = as.factor(rep(names(results), each = nrow(df_list[[1]]))))
+    rownames(df_list) <- NULL
+    return(df_list)
+}
+
+## Generating the nested tables
+table_maxi <- convert.df(differences_maxi)
+table_mini <- convert.df(differences_mini)
+table_rand <- convert.df(differences_rand)
+
+## Combining the tables
+all_results <- cbind(rbind(table_maxi, table_mini, table_rand),
+                    "scenario" = as.factor(rep(c("maxi", "mini", "rand"), each = nrow(table_maxi))))
+```
+
+```{r}
+## Pairwise tests wrapper
+pair.test <- function(data, factor, test) {
+    ## Separating the data according to factors
+    factor <- which(colnames(data) == factor)
+    splited_data <- split(data, data[, factor])
+
+    ## Combination of tests
+    combinations <- combn(1:length(splited_data), 2)
+
+    ## Applying the tests
+    tests <- apply(combinations, 2, function(X, test, data)
+                 return(test(data[[X[1]]][,1], data[[X[2]]][,1])), test = test, data = splited_data)
+    names_test <- apply(combinations, 2, function(X, data)
+                   return(paste0(names(data)[X[1]], " - ", names(data)[X[2]])), data = splited_data)
+
+    names(tests) <- names_test
+    return(tests)
+}
+
+## Running the wilcox tests for each pooled factor
+wilcox_scen <- pair.test(all_results, factor = "scenario", wilcox.test)
+wilcox_taxa <- pair.test(all_results, factor = "taxa", wilcox.test)
+wilcox_char <- pair.test(all_results, factor = "characters", wilcox.test)
+
+## Extracting the statistic and the p_value only
+wilcox_scen_w <- unlist(lapply(wilcox_scen, function(X) return(unname(X$statistic))))
+wilcox_taxa_w <- unlist(lapply(wilcox_taxa, function(X) return(unname(X$statistic))))
+wilcox_char_w <- unlist(lapply(wilcox_char, function(X) return(unname(X$statistic))))
+wilcox_scen_p <- p.adjust(unlist(lapply(wilcox_scen, function(X) return(X$p.value))), "bonferroni")
+wilcox_taxa_p <- p.adjust(unlist(lapply(wilcox_taxa, function(X) return(X$p.value))), "bonferroni")
+wilcox_char_p <- p.adjust(unlist(lapply(wilcox_char, function(X) return(X$p.value))), "bonferroni")
+
+## Running the Bhattacharrya tests for each pooled factor
+bhattc_scen <- pair.test(all_results, factor = "scenario", bhatt.coeff)
+bhattc_taxa <- pair.test(all_results, factor = "taxa", bhatt.coeff)
+bhattc_char <- pair.test(all_results, factor = "characters", bhatt.coeff)
+
+## Combine the results
+combined_results <- as.data.frame(rbind(
+        cbind("statistic" = wilcox_scen_w, "p.value" = wilcox_scen_p, "bhatt.coeff" = bhattc_scen),
+        cbind("statistic" = wilcox_taxa_w, "p.value" = wilcox_taxa_p, "bhatt.coeff" = bhattc_taxa),
+        cbind("statistic" = wilcox_char_w, "p.value" = wilcox_char_p, "bhatt.coeff" = bhattc_char)))
+```
+
+And here are the results:
+
+```{r, echo = FALSE}
+caption <- paste0("Pairwise Wilcoxon test and Bhattacharrya Coefficients for the difference in",
+  " proportion of character replacement for the different scenarios, number of taxa and characters.")
+kable(combined_results, digits = 3, caption = caption)
+```
+
+```{r, eval = TRUE, echo = FALSE, print = FALSE, results = 'hide', warning = FALSE}
+library(xtable)
+source("../Functions/make.xtable.R")
+make.xtable(combined_results, caption = caption, label = "Tab_difference_change",
+            path = "../Writing/Tables/", include.rownames = TRUE)
+```

### 2019-04-17 Analysis/04-CheckMatrixCorrelation.Rmd
Updated character correlations

@@
 ```{r, echo = FALSE}
-caption <- paste0("Pairwise Wilcoxon test and Bhattacharrya Coefficients for the difference in",
-  " proportion of character replacement for the different scenarios, number of taxa and characters.")
+caption <- paste0("Pairwise Wilcoxon test and Bhattacharrya Coefficients for the differences in",
+  "character difference for the different scenarios, number of taxa and characters.")
 kable(combined_results, digits = 3, caption = caption)
@@
 get.char.replace <- function(chain) {
-    ## Get the duplicates
-    all_duplicates <- lapply(chain$changes, function(X) return(length(unique(X[,2]))))
-    ## Make them proportional
-    return(mapply(function(x,y) return(x/y), all_duplicates, as.list(c(100, 350, 1000)),
-                 SIMPLIFY = FALSE))
-
+    return(lapply(chain$changes, function(X) return(length(unique(X[,2])))))
 }
@@
 ```{r, fig.height = 18, fig.width = 6}
+## Make the characters proportional
+make.prop <- function(list) {
+    return(mapply(function(x,y) return(x/y), list, as.list(c(100, 350, 1000)),
+                 SIMPLIFY = FALSE))
+}
+
 ## Extract the results for each type of scenarios
-duplicates_maxi <- lapply(parameters_duplicates, lapply, extract.results, what = "maxi")
-duplciates_mini <- lapply(parameters_duplicates, lapply, extract.results, what = "mini")
-duplicates_rand <- lapply(parameters_duplicates, lapply, extract.results, what = "rand")
+duplicates_maxi <- lapply(lapply(parameters_duplicates, lapply, extract.results, what = "maxi"),
+                        make.prop)
+duplicates_mini <- lapply(lapply(parameters_duplicates, lapply, extract.results, what = "mini"),
+                        make.prop)
+duplicates_rand <- lapply(lapply(parameters_duplicates, lapply, extract.results, what = "rand"),
+                        make.prop)
 
 ## overal_range (for plotting)
-overal_range <- range(c(unlist(duplicates_maxi), unlist(duplciates_mini), unlist(duplicates_rand)))
+overal_range <- c(0,1)
 
@@
     ylab = "Proportion of duplicated characters")
-plot.proportions(duplciates_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
+plot.proportions(duplicates_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
     ylab = "Proportion of duplicated characters")
@@
 ```{r, fig.height = 18, fig.width = 6, eval = TRUE, echo = FALSE}
-pdf("../Writing/Figures/Difference_change_maxi.pdf", width = 6, height = 6)
-plot.proportions(differences_maxi, ylim = overal_range, main = "Maximised", leg.pos = "bottomleft")
-pdf("../Writing/Figures/Difference_change_mini.pdf", width = 6, height = 6)
-plot.proportions(differences_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft")
-pdf("../Writing/Figures/Difference_change_rand.pdf", width = 6, height = 6)
-plot.proportions(differences_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft"))
+pdf("../Writing/Figures/Proportion_duplicated_maxi.pdf", width = 6, height = 6)
+plot.proportions(duplicates_maxi, ylim = overal_range, main = "Maximised", leg.pos = "topleft",
+    ylab = "Proportion of duplicated characters")
+pdf("../Writing/Figures/Proportion_duplicated_mini.pdf", width = 6, height = 6)
+plot.proportions(duplicates_mini, ylim = overal_range, main = "Minimised", leg.pos = "topleft",
+    ylab = "Proportion of duplicated characters")
+pdf("../Writing/Figures/Proportion_duplicated_rand.pdf", width = 6, height = 6)
+plot.proportions(duplicates_rand, ylim = overal_range, main = "Randomised", leg.pos = "topleft",
+    ylab = "Proportion of duplicated characters")
 ```
@@
 ```{r}
-## Data frame converter
-convert.df <- function(results) {
-    convert.sub.df <- function(list, names) {
-        df <- data.frame(unlist(list), as.factor(rep(names(list), each = length(list[[1]]))))
-        colnames(df) <- names
-        rownames(df) <- NULL
-        return(df)
-    }
-    df_list <- lapply(results, convert.sub.df, names = c("proportion", "characters"))
-    df_list <- cbind(do.call(rbind, df_list),
-                     "taxa" = as.factor(rep(names(results), each = nrow(df_list[[1]]))))
-    rownames(df_list) <- NULL
-    return(df_list)
-}
-
 ## Generating the nested tables
-table_maxi <- convert.df(differences_maxi)
-table_mini <- convert.df(differences_mini)
-table_rand <- convert.df(differences_rand)
+table_maxi <- convert.df(duplicates_maxi)
+table_mini <- convert.df(duplicates_mini)
+table_rand <- convert.df(duplicates_rand)
 
@@
 ```{r}
-## Pairwise tests wrapper
-pair.test <- function(data, factor, test) {
-    ## Separating the data according to factors
-    factor <- which(colnames(data) == factor)
-    splited_data <- split(data, data[, factor])
-
-    ## Combination of tests
-    combinations <- combn(1:length(splited_data), 2)
-
-    ## Applying the tests
-    tests <- apply(combinations, 2, function(X, test, data)
-                 return(test(data[[X[1]]][,1], data[[X[2]]][,1])), test = test, data = splited_data)
-    names_test <- apply(combinations, 2, function(X, data)
-                   return(paste0(names(data)[X[1]], " - ", names(data)[X[2]])), data = splited_data)
-
-    names(tests) <- names_test
-    return(tests)
-}
-
 ## Running the wilcox tests for each pooled factor
@@
 caption <- paste0("Pairwise Wilcoxon test and Bhattacharrya Coefficients for the difference in",
-  " proportion of character replacement for the different scenarios, number of taxa and characters.")
+  " proportion of character duplicated for the different scenarios, number of taxa and characters.")
 kable(combined_results, digits = 3, caption = caption)
@@
 source("../Functions/make.xtable.R")
-make.xtable(combined_results, caption = caption, label = "Tab_difference_change",
+make.xtable(combined_results, caption = caption, label = "Tab_proportion_duplicated",
             path = "../Writing/Tables/", include.rownames = TRUE)

### 2019-04-18 Analysis/05-char.diffModification.Rmd
Modification of CD

---
title: "Modification of the unexpected behaviour of the Character Difference"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

The version of the Character Difference metric proposed in the manuscript version USYB-2018-099 had a flawed behaviour which did not capture correctly the difference between characters as defined by the difference in phylogenetic splits implied by the two characters.

This is the old version of the Character Difference metric (in manuscript USYB-2018-099 and `dispRity` package `v1.0.0`):

* Old version of the Character Difference metric:

\begin{equation}
    CD.old_{(x,y)} = 1 - 2\left(\frac{\sum_{i}^{n}|x_{i} - y_{i}|}{n}-\frac{1}{2}\right)
\end{equation}

This version was problematic as the differences where not linear creating the maximum difference when only half of the characters where different (see below).
It has now been corrected in this newer version (current manuscript and `dispRity` package >= `v1.1.0`) that is linear.

* New version of the Character Difference metric

\begin{equation}
    CD.new_{(x,y)} = \frac{\sum_{i}^{n}|x_{i} - y_{i}|}{n-1}
\end{equation}

## Behaviour of both metric versions

If we consider an hypothetical discrete character matrix of an unknown number of taxa and 100 characters, the difference between characters can range from 0 (all characters are identical) to 100 (every character are 
totally different).
Considering this, we can compare both metrics considering a distance vector (`dist`) ranging from 0 to 100 and plot the Character difference for each differences:

```{r, fig.width = 8, fig.height = 8}
## Older version
old.char.diff <- function(distance, count) {
    ## The distance
    difference <- distance/count
    ## The Character Difference
    return( 1 - ( abs(difference-0.5)/0.5 ))
}
## New version
new.char.diff <- function(distance, count) {
    ## The distance (note the (count-1))
    difference <- distance/(count-1)
    ## The Character Difference (equal to the distance / (count - 1))
    return(difference)
}

## The distances to consider
distance <- seq(from = 0:100)

## The change in Character difference metric
par(bty = "n")
plot(NULL, xlim = c(0, 100), ylim = c(0, 1), ylab = "Character Difference", xlab = "Differences")
lines(old.char.diff(distance, count = 100), col = "blue")
lines(new.char.diff(distance, count = 100), col = "orange")
legend("topleft", c("Old", "New"), col = c("blue", "orange"), lty = 1)
```

> Note that the number of taxa is unknown since in practice, a matrix that will satisfy this assumption above is highly hypothetical and is unlikely to be used in practice (i.e. that will require a matrix with more taxa than character which is rarely used in evolutionary biology).

### 2019-04-18 Writing/response_to_reviewers.tex
Updated response

@@
 
-We am very grateful to the three referees for their helpful and constructive comments, which we believe have helped me to significantly improve the manuscript.
+We are very grateful to the three referees for their helpful and constructive comments, which we believe have helped me to significantly improve the manuscript.
 We have taken all of their comments on board, and respond to their points in details below.
@@
 
+After working on the required proof of the metric, we figured out that the formerly proposed Character Difference had a non-expected behaviour of being non-linear.
+We have now modified the metric to be linear (see illustration here: \url{https://github.com/TGuillerme/CharactersCorrelation/blob/master/Analysis/05-char.diffModification.pdf}).
+We thus made some major changes to the manuscript that were not suggested by the reviewers below:
 \begin{enumerate}
-\item{\textcolor{blue}{Distinctiveness}}
-\item{\textcolor{blue}{Non-negativity}}
-\item{\textcolor{blue}{Symmetry}}
-\item{\textcolor{blue}{Triangle inequality}}
-\item{\textcolor{blue}{Four-point condition}}
+    \item We change the Character Difference metric to be a scaled Hamming Distance rather than the former proposed distance.
+    \item We thus removed the supplementary material 1 wit the proof for the distance metric (since Hamming distances are well established distance metrics).
+    \item We reran all the simulations from scratch with the inclusion of the updated Character Difference metric.
+    \item For time constraint reason we reran 20 replicates (rather than 35 in the previous manuscript) for a total of 40 CPU years.
+    \item We propose a new interpretation to the updated results.
 \end{enumerate}
 
-\item{\textcolor{blue}{If so, then it is actually an evolutionary distance, and it can be used as such; if not, then it is another type of distance.
-Establishing the type of distance may be overkill, but the characterization of a metric helps to determine what it can be used for, and what it cannot be used for.}}
+Of course, we also took all the reviewers comments on board (excluding the ones rendered obsolete by the modified metrics such as the proof for the four point condition).
 
-In the supplementary materials 1, we mathematically define the characteristics of $CD$ and proved that it is a distance metric since it respects the four first points:
-
-\begin{enumerate}
-\item Distinctiveness: $CD_{(x,y)} = 0 \Leftrightarrow x = y$, proved in equation 11 (note that we corrected the error highlighted by reviewer 2 with a more restrictive definition of $x$ and $y$).
-\item Non-negativity: $CD_{(x,y)} \geq 0$, proved in equation 10.
-\item Symmetry: $CD_{x,y} = CD_{y,x}$, proved in equation 12.
-\item Triangle inequality: $CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}$, proved in equation 13.
-\item Four point condition: $CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )}$, proved in equation 15.
-\end{enumerate}
+Note that, if you consider it necessary, we'll happily add the characterisation of the new metric to the supplementary although, again, since it is now simply a scaled Hamming distance, we believed it was unwarranted for this revised version of the manuscript.
 
@@
 As the reviewer points out, characters with less taxa in common will thus be less likely to be different.
-This is consistent with what the CD metric actually captures in terms of phylogenetic signal (the number of share splits - see Reviewer 2 comment \ref{proof}).
+This is consistent with what the CD metric actually captures in terms of phylogenetic signal (the number of share splits).
 In the example above, the both characters imply at least one split (between the first taxon and the others).
@@
 
-We've added the proportion of replaced characters in each matrices for each ``minimised'', ``maximised'' and ``randomised'' scenarios in the supplementary material 3, figures 1, 2 and 3.
-It is worth noting that there is a significant difference in terms of number of replaced characters in each scenarios between the ``maximised'' and the ``minimised'' and ``randomised'' matrices but non in terms of number of taxa or characters.
+We've added the proportion of replaced (i.e.
+duplicated) characters in each matrices for each ``minimised'', ``maximised'' and ``randomised'' scenarios along with the changes in overall Character Differences in the supplementary material 3, figures 1, 2 and 3.
+ @@@TO CHECK + change SM numbering
 
-This could potentially explain some of the differences between the scenarios (i.e.
-why it seems better to maximise character difference rather than minimising it) but does not explain the differences in smaller matrices nor between the ``minimised'' and ``randomised'' scenarios and neither between the two inference methods (since they both use the same matrices).
-Furthermore, the maximising/minimising character differences was done \textit{a priori} without any targeted number of characters to replace to make the analysis less idiosynchratic.
+The maximising/minimising character differences was done \textit{a priori} without any targeted number of characters to replace to make the analysis less idiosynchratic.
 In other words, we decided to modify the randomly simulated matrices to globally increase/decrease character differences which is expected to decrease anyway in bigger matrices.
@@
 
-%TG: TODO: added pragraph!
+%TG: TODO: add pragraph!
 \textit{.} lines @@@
@@
 
-%TG: add a figure for the difference character dependencies?
+%TG: add a figure for the difference character dependencies? @@@
 
@@
 
-We've added a paragraph highlithing how the problem indeed comes from the fact that current software are not equipped to handle this issue:
+We've added a paragraph highlighting how the problem indeed comes from the fact that current software are not equipped to handle this issue:
 

### 2019-04-18 Writing/draft_highlighted.tex
cite to add
Hamming distance
Updated response

@@
 % Palaeontology
+
+% CITE: https://academic.oup.com/sysbio/article/68/2/267/5145070#130574594
+
+% Mention Hamming distance rather than Gower?
 
@@
 \begin{equation}
-    CD_{(x,y)} = \abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}}
+    CD_{(x,y)} = \frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}
 \end{equation}
@@
 \noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
-$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
+%TG: Modifications in this sentence not in the letter yet
+$CD$ is a continuous \hl{scaled Hamming distance} metric bounded between $0$ and $1$.% (see the mathematical demonstration in the supplementary material 1).
 Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
@@
 \hl{Note that both models (``morphological HKY-binary' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
-However, the ``morphological HKY model'' is not the one \textit{directly}estimated in the phylogenetic software used in this protocol (MrBayes, see below).}
+However, the ``morphological HKY model'' is not the one \textit{directly} estimated in the phylogenetic software used in this protocol (MrBayes, see below).}
 
@@
 The matrices were generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 35 times resulting in 315 ``normal'' morphological matrices.
+To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 35@@@ times resulting in 315 ``normal'' morphological matrices.
 

### 2019-04-19 Writing/supplementary/Sup1-CharDiff.tex
Updated char diff demo

@@
 \begin{equation}
-    CD_{(x,y)} = 1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}\right)
+    CD_{(x,y)} = \frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}
 \end{equation}
@@
 For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ because the difference between character states is considered as Fitch-like (unordered; i.e.
-the difference between two different characters states is equal to $1$ and the difference between two identical character states is equal to $0$):
+the difference between two different characters states is equal to $1$ and the difference between two identical character states is equal to $0$).
+Furthermore, following the $xyz$ notation described above, for any $n > 1$, $x_{1}-y_{1}=0$.
+Therefore the sum of the differences $\sum_{i=1}^{n}\abs{x_{i}-y_{i}} \in [0;(n-1)]$.
 
@@
     0 \leq  \abs{x_{n}-y_{n}} \leq 1\\
-    0 \leq  \sum_{i=1}^{n}\abs{x_{i}-y_{i}} \leq n\\
-    0 \leq  \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} \leq 1\\
-    0 \leq  \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} \leq \frac{1}{2}\\
-    0 \leq  2\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} \leq 1\\
-    0 \leq  1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}\right) \leq 1\\
+    0 \leq  \sum_{i=1}^{n}\abs{x_{i}-y_{i}} \leq (n-1)\\
+    0 \leq  \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1} \leq 1\\
     \end{gathered}
@@
     \sum_{i}^{n}\abs{x_{i}-y_{i}} &= 0\\
-    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= 0\\
-    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \frac{1}{2}\\
-    2\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= 1\\
-    1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}\right) &= 0\\
+    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1} &= 0\\
     \end{aligned}
@@
     \sum_{i}^{n}\abs{x_{i}-y_{i}} &= \sum_{i}^{n}\abs{y_{i}-x_{i}}\\
-    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n} &= \frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}\\
-    \abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= \abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\\
-    2\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} &= 2\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\\
-    1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}\right) &= 1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n}-\frac{1}{2}}\right)\\
+    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1} &= \frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n-1}\\
     \end{aligned}
@@
 
-For any $n, j, b \in [1;\infty)$ and $x_{n}$, $y_{j}$, $z_{b}$ $\in (0,1)$:
+For any $n, \in [1;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:
 
 \begin{equation}
-    \label{subadditivity_demo}
+    \label{equality_demo}
     \begin{aligned}
-    CD_{x,y} &\leq CD_{x,z} + CD_{z,y}\\
-    \text{if:}\\
-     1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}}\right)
+    \text{since:}\\
+    \underbrace{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1}}_{\in[0,1]}
     &\leq
-     1 - 2\left(\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-z_{k}}}{j}-\frac{1}{2}}\right) 
+    \underbrace{\frac{\sum_{i}^{n}\abs{x_{i}-z_{i}}}{n-1}}_{\in[0,1]}
     +
-     1 - 2\left(\abs*{\frac{\sum_{a}^{b}\abs{z_{a}-y_{a}}}{b}-\frac{1}{2}}\right) \\
-     \text{then, the opposite is:}\\
-     \underbrace{\abs*{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n}-\frac{1}{2}} }_{\in[0,1/2]}
-     &\geq 
-     \underbrace{\abs*{\frac{\sum_{k}^{j}\abs{x_{k}-z_{k}}}{j}-\frac{1}{2}} }_{\in[0,1/2]}
-     +
-     \underbrace{\abs*{\frac{\sum_{a}^{b}\abs{z_{a}-y_{a}}}{b}-\frac{1}{2}} }_{\in[0,1/2]} - 2\\
+    \underbrace{\frac{\sum_{i}^{n}\abs{z_{i}-y_{i}}}{n-1}}_{\in[0,1]}\\
     \text{thus:}\\
-     \in[0,\frac{1}{2}] &\geq \in[0,\frac{1}{2}] + \in[0,\frac{1}{2}] -2 \\
-     \in[0,\frac{1}{2}] &\geq \in[-2,-1]\\
+    \in[0,1] &\leq \in[0,1] + \in[0,1] \\
+    \in[0,1] &\leq \in[0,2]\\
     \end{aligned}
 \end{equation}
-
 Thus equation \ref{Subadditivity} is true.

### 2019-04-19 Writing/response_to_reviewers.tex
changes to results part 1

@@
 \begin{enumerate}
+\label{metric_change}
 
@@
 
-Note that, if you consider it necessary, we'll happily add the characterisation of the new metric to the supplementary although, again, since it is now simply a scaled Hamming distance, we believed it was unwarranted for this revised version of the manuscript.
+We have also updated the Character Difference metric characterisation in the supplementary material 1.
+However, since it is now simply a scaled Hamming distance, these demonstrations might be unwarranted for this revised version of the manuscript.
 
@@
 
-We are really grateful that this reviewer picked up this flow in our mathematical demonstration of CD metric.
-This problem is due to a lack of indications from our side and we apologise for this.
+We have reviewed the formula of the CD metric following an unexpected non-linear behaviour (see comment \ref{metric_change}).
+Furtheremore, we'd like to highlight the following points:
+
+% We are really grateful that this reviewer picked up this flow in our mathematical demonstration of CD metric.
+% This problem is due to a lack of indications from our side and we apologise for this.
 
@@
 Following these two points, the reviewer counter example $CD_{(x,y)} = 0$ means that 1) $x'$ and $y'$ are identical and the characters $x$ and $y$ imply the same number of splits in a tree (here no splits since only one taxon is used - however, this would hold for any $n$ number of taxa).
-We have clarified these two points in the manuscript and in the mathematical demonstration.
+We have clarified these two points in the manuscript.
 

### 2019-04-19 Writing/draft_highlighted.tex
changes to results part 1

@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839\hl{; see supplementary materials 3, Table 2}) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628\hl{; see supplementary materials 3, Table 2}) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303\hl{; see supplementary materials 3, Table 2}).
-There is a significant difference between all scenarios (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test})
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.869 and median $NTS_{Tr}$ = 0.787\hl{; see supplementary materials 3, Table 3}) followed by randomising the characters (median $NTS_{RF}$ = 0.814 and median $NTS_{Tr}$ = 0.791\hl{; see supplementary materials 3, Table 2}) and maximising the character difference (median $NTS_{RF}$ = 0.698 and median $NTS_{Tr}$ = 0.754\hl{; see supplementary materials 3, Table 2}).
+There is a significant difference between all scenarios \hl{in terms of clade conservation ($NTS_{RF}$)} (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
+\hl{There is however no difference in terms of taxa displacement ($NTS_{Tr}$) between the three scenarios (with a BC > 0.98 for each comparison; Table \ref{Tab_pooledscenarios_test}).}
 
@@
 \subsubsection{Number of characters}
-This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
+\hl{This effect of the character difference affects clade conservation ($NTS_{RF}$) only by big differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ \ref{Tab_pooledscharacters_test}).}
+\hl{For closer differences of characters (i.e. 100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests where different ($p = 0$) but with high distribution overlap ($BC > 0.96$).}
+
+is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
+
 The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 3}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).

### 2019-04-22 Writing/draft_highlighted.tex
Single comparisons done

@@
 \subsubsection{Number of characters}
-\hl{This effect of the character difference affects clade conservation ($NTS_{RF}$) only by big differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ \ref{Tab_pooledscharacters_test}).}
-\hl{For closer differences of characters (i.e. 100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests where different ($p = 0$) but with high distribution overlap ($BC > 0.96$).}
-
-is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
-
-The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 3}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).
-The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3 \hl{Table 3}) with a significant difference between 100 and 350 or 1000 characters (\hl{\textit{p} = 0, \textit{p} = 0 and \textit{p} = 0.014 respectively}; Table \ref{Tab_pooledscharacters_test}).
-However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.
+\hl{This effect of the character difference affects clade conservation ($NTS_{RF}$) only by big differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ - table \ref{Tab_pooledscharacters_test}).}
+\hl{For closer differences between characters numbers (i.e.
+100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests showed clear differences ($p = 0$) but with high distribution overlap ($BC > 0.96$).}
+\hl{Regarding taxa displacements ($NTS_{Tr}$), there was an effect of the number of characters (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - table \ref{Tab_pooledscharacters_test}).}
 
@@
 Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
-A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (\hl{\textit{p} = 0.012, \textit{p} = 0.004, \textit{p} = 0 and \textit{p} = 0 respectively}; Table \ref{Tab_pooledstaxa_test} and medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively; see supplementary materials 3 \hl{Table 4}).
-Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
+\hl{A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ (median $NTS_{RF}$ of 0.846 and 0.787 for 25 and 75 taxa respectively with a $p = 0$ and a $BC$ marginally $< 0.95$ - table \ref{Tab_pooledstaxa_test} and supplementary materials 3, table 5) but no differences between 75 and 150 taxa and for $NTS_{Tr}$ ($BC > 0.95$ - table \ref{Tab_pooledstaxa_test}).}
+The significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
 
@@
 
-Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\hl{\textit{p} = 0} Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively; \hl{see supplementary materials 3, Table 5}) but not in terms of individual taxon placements (\hl{\textit{p} = 0.08}; Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively; \hl{see supplementary materials 3, Table 5}).
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\hl{\textit{p} = 0} Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of \hl{0.851} and \hl{0.739} respectively; \hl{see supplementary materials 3, Table 6}) but not in terms of individual taxon placements (\hl{\textit{p} = 0.041 but $BC > 0.95$}; Table \ref{Tab_pooledsmethods_test}).
 
@@
 
-\subsection{Combined effects of taxa, characters and correlation on topology}
+%TODO:
+% \subsection{Combined effects of taxa, characters and correlation on topology}
 
-When looking at the combined effect of each parameter, the ``maximised'' and ``minimised'' scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
-The same differences are observed when comparing the ``maximised'' scenario against the ``randomised'' one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the maximum parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
-Identically, there was always a significant difference between the ``minimised'' scenario and the ``randomised'' one was expect for the matrix of 150 taxa and 100 characters under maximum parsimony for $NTS_{RF}$ and the matrix of 150 and 1000 characters under Bayesian inference for $NTS_{Tr}$.
-The full list of comparisons and summary statistics are available in the supplementary materials 3.
+% When looking at the combined effect of each parameter, the ``maximised'' and ``minimised'' scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
+% The same differences are observed when comparing the ``maximised'' scenario against the ``randomised'' one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the maximum parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
+% Identically, there was always a significant difference between the ``minimised'' scenario and the ``randomised'' one was expect for the matrix of 150 taxa and 100 characters under maximum parsimony for $NTS_{RF}$ and the matrix of 150 and 1000 characters under Bayesian inference for $NTS_{Tr}$.
+% The full list of comparisons and summary statistics are available in the supplementary materials 3.
 

### 2019-04-23 Analysis/02-EffectCorrelationFullResults.Rmd
Updated analysis for methods

@@
 methods_list <- list(method_RF_norm, method_Tr_norm, method_RF_null, method_Tr_null)
-comp.labels <- c("Best", rep("", 3), "Null", rep("", 3))
-metric.labels <- rep(c("RF", "", "Tr", ""), 2)
+comp.labels <- c("Best", rep("", 1), "Null", rep("", 1))
+metric.labels <- rep(c("RF", "Tr"), 2)
 test_table <- pair.test.table(methods_list, comp.label = comp.labels, metric.label = metric.labels)

### 2019-04-23 Writing/Tables/Full_Tab_pooledsmethods_test.tex
Updated analysis for methods

@@
 % latex table generated in R 3.5.3 by xtable 1.8-3 package
-% Tue Apr 16 16:10:50 2019
+% Tue Apr 23 11:45:02 2019
 \begin{table}[ht]
@@
 Best & RF & bayesian:parsimony & 0.939 & 192659.000 & 0.000 \\ 
-   &  & bayesian:parsimony & 0.986 & 157677.500 & 0.082 \\ 
-   & Tr & bayesian:parsimony & 0.944 & 193142.000 & 0.000 \\ 
-   &  & bayesian:parsimony & 0.975 & 161621.000 & 0.008 \\ 
-  Null & RF & bayesian:parsimony & 0.939 & 192659.000 & 0.000 \\ 
-   &  & bayesian:parsimony & 0.986 & 157677.500 & 0.082 \\ 
-   & Tr & bayesian:parsimony & 0.944 & 193142.000 & 0.000 \\ 
-   &  & bayesian:parsimony & 0.975 & 161621.000 & 0.008 \\ 
+   & Tr & bayesian:parsimony & 0.986 & 157677.500 & 0.082 \\ 
+  Null & RF & bayesian:parsimony & 0.944 & 193142.000 & 0.000 \\ 
+   & Tr & bayesian:parsimony & 0.975 & 161621.000 & 0.008 \\ 
    \hline

### 2019-04-23 Analysis/02-EffectCorrelationFullResults.Rmd
added combined results test tables

@@
 ## Labels
-labels_wx <- paste0("Full_Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
+labels_wx <- paste0("Full_Tab_Bayesian_WX", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
 
 ## Tables
-make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1], path = "../Writing/Tables/")
-make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2], path = "../Writing/Tables/")
-make.xtable(wx_RF_null, caption = captions_wx[3], label = labels_wx[3], path = "../Writing/Tables/")
-make.xtable(wx_Tr_null, caption = captions_wx[4], label = labels_wx[4], path = "../Writing/Tables/")
+## Bayesians
+make.xtable(wx_RF_best[1:9,], caption = captions_wx[1], label = labels_wx[1], path = "../Writing/Tables/")
+make.xtable(wx_Tr_best[1:9,], caption = captions_wx[2], label = labels_wx[2], path = "../Writing/Tables/")
+make.xtable(wx_RF_null[1:9,], caption = captions_wx[3], label = labels_wx[3], path = "../Writing/Tables/")
+make.xtable(wx_Tr_null[1:9,], caption = captions_wx[4], label = labels_wx[4], path = "../Writing/Tables/")
+
+labels_wx <- paste0("Full_Tab_Parsimony_WX", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))
+
+## Parsimonies
+make.xtable(wx_RF_best[10:18,], caption = captions_wx[1], label = labels_wx[1], path = "../Writing/Tables/")
+make.xtable(wx_Tr_best[10:18,], caption = captions_wx[2], label = labels_wx[2], path = "../Writing/Tables/")
+make.xtable(wx_RF_null[10:18,], caption = captions_wx[3], label = labels_wx[3], path = "../Writing/Tables/")
+make.xtable(wx_Tr_null[10:18,], caption = captions_wx[4], label = labels_wx[4], path = "../Writing/Tables/")
 ```

### 2019-04-23 Writing/Tables/Full_Tab_Bayesian_WXRFbest.tex
added combined results test tables

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 23 14:45:48 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & max.min.W & max.min.p & max.rand.W & max.rand.p & min.rand.W & min.rand.p \\ 
  \hline
bayesian & 25 & 100 & 136.500 & 1.000 & 167.000 & 1.000 & 219.000 & 1.000 \\ 
   &  & 350 & 123.500 & 0.707 & 148.000 & 1.000 & 216.000 & 1.000 \\ 
   &  & 1000 & 51.500 & 0.001 & 81.000 & 0.024 & 218.000 & 1.000 \\ 
   & 75 & 100 & 5.500 & 0.000 & 51.500 & 0.001 & 381.000 & 0.000 \\ 
   &  & 350 & 20.000 & 0.000 & 52.500 & 0.001 & 301.500 & 0.112 \\ 
   &  & 1000 & 84.000 & 0.032 & 118.000 & 0.494 & 296.000 & 0.174 \\ 
   & 150 & 100 & 25.500 & 0.000 & 108.000 & 0.239 & 340.000 & 0.003 \\ 
   &  & 350 & 69.500 & 0.008 & 121.500 & 0.625 & 327.500 & 0.011 \\ 
   &  & 1000 & 105.500 & 0.198 & 122.500 & 0.669 & 252.500 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Robinson-Fould distance from the best tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Bayesian_WXRFbest}
\end{table}

### 2019-04-23 Writing/Tables/Full_Tab_Bayesian_WXRFnull.tex
added combined results test tables

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 23 14:45:48 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & max.min.W & max.min.p & max.rand.W & max.rand.p & min.rand.W & min.rand.p \\ 
  \hline
bayesian & 25 & 100 & 315.500 & 0.033 & 310.500 & 0.051 & 183.500 & 1.000 \\ 
   &  & 350 & 301.000 & 0.115 & 248.000 & 1.000 & 135.500 & 1.000 \\ 
   &  & 1000 & 345.000 & 0.002 & 228.000 & 1.000 & 65.500 & 0.005 \\ 
   & 75 & 100 & 353.000 & 0.001 & 252.500 & 1.000 & 59.500 & 0.003 \\ 
   &  & 350 & 355.000 & 0.001 & 258.000 & 1.000 & 64.500 & 0.005 \\ 
   &  & 1000 & 293.500 & 0.213 & 209.500 & 1.000 & 113.500 & 0.359 \\ 
   & 150 & 100 & 310.500 & 0.053 & 226.000 & 1.000 & 98.000 & 0.108 \\ 
   &  & 350 & 315.500 & 0.033 & 266.000 & 1.000 & 100.500 & 0.133 \\ 
   &  & 1000 & 308.000 & 0.065 & 260.000 & 1.000 & 128.000 & 0.954 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Robinson-Fould distance from the null tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Bayesian_WXRFnull}
\end{table}

### 2019-04-23 Writing/Tables/Full_Tab_Bayesian_WXTrbest.tex
added combined results test tables

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 23 14:45:48 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & max.min.W & max.min.p & max.rand.W & max.rand.p & min.rand.W & min.rand.p \\ 
  \hline
bayesian & 25 & 100 & 214.000 & 1.000 & 179.000 & 1.000 & 157.500 & 1.000 \\ 
   &  & 350 & 241.500 & 1.000 & 198.500 & 1.000 & 158.000 & 1.000 \\ 
   &  & 1000 & 158.500 & 1.000 & 96.000 & 0.092 & 165.000 & 1.000 \\ 
   & 75 & 100 & 101.000 & 0.121 & 143.000 & 1.000 & 235.000 & 1.000 \\ 
   &  & 350 & 143.000 & 1.000 & 137.000 & 1.000 & 180.000 & 1.000 \\ 
   &  & 1000 & 151.000 & 1.000 & 140.000 & 1.000 & 175.000 & 1.000 \\ 
   & 150 & 100 & 135.000 & 1.000 & 226.000 & 1.000 & 293.000 & 0.202 \\ 
   &  & 350 & 219.000 & 1.000 & 228.000 & 1.000 & 213.000 & 1.000 \\ 
   &  & 1000 & 178.000 & 1.000 & 170.000 & 1.000 & 196.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Triplets distance from the best tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Bayesian_WXTrbest}
\end{table}

### 2019-04-23 Writing/Tables/Full_Tab_Bayesian_WXTrnull.tex
added combined results test tables

% latex table generated in R 3.5.3 by xtable 1.8-3 package
% Tue Apr 23 14:45:48 2019
\begin{table}[ht]
\centering
\begin{tabular}{lllrrrrrr}
  \hline
method & taxa & characters & max.min.W & max.min.p & max.rand.W & max.rand.p & min.rand.W & min.rand.p \\ 
  \hline
bayesian & 25 & 100 & 244.000 & 1.000 & 300.500 & 0.123 & 246.500 & 1.000 \\ 
   &  & 350 & 257.500 & 1.000 & 254.000 & 1.000 & 205.500 & 1.000 \\ 
   &  & 1000 & 316.000 & 0.032 & 243.500 & 1.000 & 130.000 & 1.000 \\ 
   & 75 & 100 & 255.000 & 1.000 & 223.000 & 1.000 & 168.000 & 1.000 \\ 
   &  & 350 & 278.000 & 0.648 & 260.000 & 1.000 & 165.500 & 1.000 \\ 
   &  & 1000 & 238.000 & 1.000 & 258.000 & 1.000 & 212.000 & 1.000 \\ 
   & 150 & 100 & 240.000 & 1.000 & 188.000 & 1.000 & 147.000 & 1.000 \\ 
   &  & 350 & 229.000 & 1.000 & 229.000 & 1.000 & 198.000 & 1.000 \\ 
   &  & 1000 & 286.500 & 0.360 & 252.000 & 1.000 & 176.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Non-parametric Wilcoxon test (Triplets distance from the null tree) between the different scenarios (p-values corrected using Bonferroni-Holm correction).} 
\label{Full_Tab_Bayesian_WXTrnull}
\end{table}

### 2019-04-24 Writing/draft_highlighted.tex
Updated manuscript

@@
 Using thorough simulations, we test the effect of increasing or decreasing character differences on tree topology.
-Overall, we found an expected positive effect of reducing character correlations on recovering the correct topology.
-However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing character correlation is not more effective than randomly drawing characters.
+Overall, we found \hl{a negative} effect of reducing character correlations on recovering the correct topology.
+However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where \hl{reducing or increasing} character correlation is not more effective than randomly drawing characters.
 Furthermore, in bigger matrices (350 characters), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
@@
 \end{enumerate}
-Each step was replicated 35 times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
+Each step was replicated \hl{20} times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
 
@@
 There is a significant difference between all scenarios \hl{in terms of clade conservation ($NTS_{RF}$)} (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
-\hl{There is however no difference in terms of taxa displacement ($NTS_{Tr}$) between the three scenarios (with a BC > 0.98 for each comparison; Table \ref{Tab_pooledscenarios_test}).}
+\hl{There is however no difference in terms of taxa displacement ($NTS_{Tr}$) between the three scenarios (with a BC > 0.98 for each comparison - Table}
+\ref{Tab_pooledscenarios_test}).
 
@@
 \subsubsection{Number of characters}
-\hl{This effect of the character difference affects clade conservation ($NTS_{RF}$) only by big differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ - table \ref{Tab_pooledscharacters_test}).}
+\hl{This effect of the character difference affects clade conservation ($NTS_{RF}$) only by big differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ - Table}
+\ref{Tab_pooledscharacters_test}).
 \hl{For closer differences between characters numbers (i.e.
 100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests showed clear differences ($p = 0$) but with high distribution overlap ($BC > 0.96$).}
-\hl{Regarding taxa displacements ($NTS_{Tr}$), there was an effect of the number of characters (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - table \ref{Tab_pooledscharacters_test}).}
+\hl{Regarding taxa displacements ($NTS_{Tr}$), there was an effect of the number of characters (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - Table}
+\ref{Tab_pooledscharacters_test}).
 
@@
 Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
-\hl{A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ (median $NTS_{RF}$ of 0.846 and 0.787 for 25 and 75 taxa respectively with a $p = 0$ and a $BC$ marginally $< 0.95$ - table \ref{Tab_pooledstaxa_test} and supplementary materials 3, table 5) but no differences between 75 and 150 taxa and for $NTS_{Tr}$ ($BC > 0.95$ - table \ref{Tab_pooledstaxa_test}).}
+\hl{A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ (median $NTS_{RF}$ of 0.846 and 0.787 for 25 and 75 taxa respectively with a $p = 0$ and a $BC$ marginally $< 0.95$ - Table}
+\ref{Tab_pooledstaxa_test}
+\hl{and supplementary materials 3, table 5) but no differences between 75 and 150 taxa and for $NTS_{Tr}$ ($BC > 0.95$ - Table}
+\ref{Tab_pooledstaxa_test}).
 The significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
@@
 
-%TODO:
-% \subsection{Combined effects of taxa, characters and correlation on topology}
+\subsection{Combined effects of taxa, characters and correlation on topology}
 
-% When looking at the combined effect of each parameter, the ``maximised'' and ``minimised'' scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
-% The same differences are observed when comparing the ``maximised'' scenario against the ``randomised'' one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the maximum parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
-% Identically, there was always a significant difference between the ``minimised'' scenario and the ``randomised'' one was expect for the matrix of 150 taxa and 100 characters under maximum parsimony for $NTS_{RF}$ and the matrix of 150 and 1000 characters under Bayesian inference for $NTS_{Tr}$.
-% The full list of comparisons and summary statistics are available in the supplementary materials 3.
+\hl{When looking at the combined effect of each parameters, the ``maximised'', ``minimised'' and ``randomised'' scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).}
+\hl{When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences for the taxa displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).}
+\hl{In terms of clade conservation ($NTS_{RF}$), however, ``maximised'' and ``minimised'' scenarios produced nearly always consistently different topologies using Bayesian inference (apart for the matrices with 25 taxa and 100 and 350 characters and the matrices with 150 taxa and 350 characters) but not using parsimony inference (only the matrices with 75 taxa and 100 or 350 characters produced different $NTS_{RF}$ distributions - supplementary material 3, tables 19 and 20).}
+%TODO: [discussion] This can be explained by lower resolution for Bayesian
 
@@
 Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
-Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
-Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 6, 7 and 8}).
+\hl{Overall, throughout our simulation, the ``minimised'' character difference scenario (i.e.
+increased correlation) tend to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e.
+decreased correlation) led to slightly worst trees (Fig.}
+\ref{Fig:RF_results_best}.
+\hl{This did not however seem to affect the taxa placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical results (Fig.}
+\ref{Fig:Tr_results_best}).
+\hl{However this results needs to be contrasted by the fact that}
+the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
+Furthermore, the inference method, the number of taxa and the number of characters all had a specific effect on the ability to recover the ``correct'' topology.
 
@@
 
-Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distributions that are always \hl{very} high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
-This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
-For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation as it is independent of the number of characters analysed.
-Similarly, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
+Although we measured a significant difference between small and larger matrices (\hl{in terms of number of characters, between 100 and 1000 characters - $BC$ $NTS_{RF}=0.914$ and $BC$ $NTS_{Tr}=0.826$}; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}),
+these differences have to be contrasted with the probability of overlap between the results distributions that are always \hl{very} high between every matrices sizes (Bhattacharrya Coefficients $>0.949$ for both the different number of characters and taxa).
+This suggests that the overall character difference metric \hl{can be used as a proxy} for character correlation as it is \hl{not directly dependent} on the number of characters analysed.
+Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data.
+Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
 we corrected for the expected difference when comparing two random trees with the same number of taxa).
@@
 
+%TODO: Bayesian recovers the ``true'' tree topology slighltly but consistently better (supplementary materials 3, figures 12, 13, and 14).
+
 The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
@@
 Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
-This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.
+This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
+\hl{This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 12, 13, and 14).}
 
@@
 These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not used \textit{a priori} as a criterion to select characters.
-In other words, intra-organisaml and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
+In other words, intra-organismal and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
 coding correlation) the only type of correlation that can affect the phylogeny \textit{a priori}.
@@
 Second, measuring and modifying character correlation is difficult. 
-In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation.
+In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation
+\hl{(supplementary materials 3, figures 1, 2 and 3).}
 In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
@@
 But there are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
-% However, we want to point out again that here we're no comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
+However, we want to point out again that here we're no comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
 
@@
 Effectively, our simulation protocol bootstraps our data ``with bias''.
-In the ``randomised'' scenarios the data is simply randomly bootstrapped simply we randomly remove and resample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
+In the ``randomised'' scenarios the data is simply randomly bootstrapped simply we randomly remove and re-sample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
 However, in the ``minimised'' and ``maximised'' scenario, the bootstrapping we remove the characters with the lowest/highest overall character difference.
-For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the left characters.
+For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the characters left.
 
@@
 the equivalent of randomising the character differences in our analysis) will be sufficient to obtain a robust topology (\textit{cf.} actively collecting different characters).
-In matrices with more taxa, however, the ``maximised'' scenario resulted in better topological recovery than any other scenarios.
+In matrices with more taxa, however, the \hl{``minimised''} scenario resulted in better topological recovery than any other scenarios.
 Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.
@@
 
-We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was high (low character differences), the topology was always the furthest away from the ``normal'' topology.
-Conversely, when correlation between characters was low, the topology was always the closest to the ``normal'' topology.
+We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was low (high character differences), the topology was always the furthest away from the ``normal'' topology.
+Conversely, when correlation between characters was high, the topology was always the closest to the ``normal'' topology.
 These results seem independent on the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
 
-However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
-For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly decreased and increased correct topological recovery compared to randomly resample matrices.
+However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise \hl{or maximised} character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
+For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly increased and decreased correct topological recovery compared to randomly resample matrices.
 
@@
 \section{Acknowledgements}
-Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232). @@@ plus awoonga
+Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232) and the Australian QRIScloud Awoonga cluster services.
 We thank Alberto Pascual Garcia for input in the design of the simulations protocol, Guillermo Herraiz Yebes and Abigail Pastore for helping with the CD metric proof and Orlin Todorov for assistance with the computations.

### 2019-04-24 Writing/draft_highlighted.tex
Updated intro and methods

@@
 \citealt{Guillerme2016146}
-\hl{);
-Second, this discrepancy has been suggested to fewer stronger correlations in morphological data (i.e.
+\hl{).
+Second, this discrepancy has been suggested to be stronger correlations in morphological data (i.e.
 ``concentrated'' correlations) \textit{versus} weaker and more ``spread'' correlations in molecular data}
@@
 We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences or similarities between characters.
-We found that overall, there is a detectable effect of character correlation on topology where an increase in character dependence results in a decrease in the ability to recover the correct topology.
+We found that overall, there is a detectable effect of character correlation on topology where \hl{a decrease} in character dependence results in a decrease in the ability to recover the correct topology.
 These results, however, vary greatly in magnitude depending on the size of matrix and the inference method used.
@@
 %TG: Modifications in this sentence not in the letter yet
-$CD$ is a continuous \hl{scaled Hamming distance} metric bounded between $0$ and $1$.% (see the mathematical demonstration in the supplementary material 1).
+$CD$ is a continuous \hl{scaled Hamming distance} metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
 Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
@@
 The matrices were generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 35@@@ times resulting in 315 ``normal'' morphological matrices.
+To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step \hl{20} times resulting in \hl{180} ``normal'' morphological matrices.
 
@@
 Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
-This step resulted in a total of 1260 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
+This step resulted in a total of \hl{720} matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
 The algorithms for the three modifications are available on GitHub (\url{https://github.com/TGuillerme/CharactersCorrelation})
@@
 \subsection{Inferring topologies}
-We inferred the topologies with both BPP and MP using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
-For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
+We inferred the topologies with both Bayesian inference and maximum parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
+For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees (\hl{the first taxa in the taxa list}).
 The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
@@
 For the Bayesian consensus trees, the 25\% first trees of the posterior tree distribution were excluded as a burnin.
-The 2880 tree inferences took around one CPU century on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC}.
+The \hl{1440} tree inferences took around one CPU century on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC} \hl{and the Australian QRIScloud Awoonga clusters (3Ghz)}.
 
@@
 When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
-\hl{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 9, 10 and 11.}
+\hl{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 12, 13 and 14.}
 
@@
 
-Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one.
 The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}).
-Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
+Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its \textit{p}-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
 when multiple trees are equal to the ``normal'' tree).
-Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
+Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146}.
 Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $<0.05$.
@@
 
-The resulting full simulation was 3.5TB \hl{in size} so is not shared here (though the parameters are).
+The resulting full simulation was \hl{2TB in size} so is not shared here (though the parameters are).
 However, the resulting consensus trees on which the topological differences are calculated are available at \url{https://figshare.com/s/7a8fde8eaa39a3d3cf56}.

### 2019-04-24 Writing/response_to_reviewers.tex
Finished response

@@
 \textit{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
-Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).} lines @@@
+Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).} lines 210-215.
 
@@
 We've added the proportion of replaced (i.e.
-duplicated) characters in each matrices for each ``minimised'', ``maximised'' and ``randomised'' scenarios along with the changes in overall Character Differences in the supplementary material 3, figures 1, 2 and 3.
- @@@TO CHECK + change SM numbering
+duplicated) characters in each matrices for each ``minimised'', ``maximised'' and ``randomised'' scenarios along with the changes in overall Character Differences in the supplementary material 3, figures 4, 5 and 6.
 
@@
 
-We've added mentions to the proportion of replaced characters analysis in the supplementary in the main text at lines @@@ and added a paragraph explaining this caveat in the discussion:
+We've added mentions to the proportion of replaced characters analysis in the supplementary and added a paragraph explaining this caveat in the discussion:
 
 %TG: TODO: add pragraph!
-\textit{.} lines @@@
+\textit{Furthermore, it is important to note that on average, less characters where duplicated (replace) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6).
+However, this change in the amount of replaced characters did not changed the way average character differences changed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
+This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.} lines 452-461.
 
@@
 changes in morphological states can be more readily linked to a function cf.
-changes in nucleotide states ; see below).} lines @@@
+changes in nucleotide states ; see below).} lines 43-57.
 
@@
 
-We've changed the sentence following the reviewer's suggestions lines @@@.
+We've changed the sentence following the reviewer's suggestions lines 7.
 
@@
 
-We added the suggested citation (lines @@@).
+We added the suggested citation (lines 75).
 
@@
 
-%TG: add a figure for the difference character dependencies? @@@
+We do not feel that adding a figure for illustrating the matrix will improve clarity of it's already long and thorough description (accompanying a rather trivial example).
+However, we've trying to clarify the description with an example as follows:
 
-We've clarified the sentence with an example as follows:
-\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.} lines @@@
+\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.} lines 95-96.
 
@@
 the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleens, the same split would still be estimated by the software).
-One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.} lines @@@
+One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.} lines 143-154.
 
@@
 
-\textit{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in Felsenstein, 2004; as used in D\'{a}valos et al., 2014).} lines @@@
+\textit{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in Felsenstein, 2004; as used in D\'{a}valos et al., 2014).} lines 218-220.
 
@@
 \textit{Note that both models (``morphological HKY'' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
-However, the ``morphological HKY model'' is not the one} directly \textit{estimated in the phylogenetic software used in this protocol (MrBayes, see below).}(lines @@@).
+However, the ``morphological HKY model'' is not the one} directly \textit{estimated in the phylogenetic software used in this protocol (MrBayes, see below).} lines 270-274.
 
@@
 
-We've changed the sentence as suggest by the reviewer (lines @@@).
+We've changed the sentence as suggest by the reviewer (lines 260-261).
 
@@
 
-\textit{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 9, 10 and 11.} lines @@@
+\textit{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 12, 13 and 14.} lines 344-346
 
@@
 
+We are really grateful that this reviewer picked up this flow in our mathematical demonstration of CD metric.
+This problem is due to a lack of indications from our side and we apologise for this.
 We have reviewed the formula of the CD metric following an unexpected non-linear behaviour (see comment \ref{metric_change}).
 Furtheremore, we'd like to highlight the following points:
-
-% We are really grateful that this reviewer picked up this flow in our mathematical demonstration of CD metric.
-% This problem is due to a lack of indications from our side and we apologise for this.
 
@@
 
-\textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines @@@
+\textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines 199-200.
 
-\textit{Similarly, an hypothetical comparison of two character for only one taxa \texttt{x = \{0\}} and \texttt{y = \{1\}} would result in comparing two identical characters \texttt{x' = \{1\}} and \texttt{y' = \{1\}}.} lines @@@
+\textit{Similarly, an hypothetical comparison of two character for only one taxa \texttt{x = \{0\}} and \texttt{y = \{1\}} would result in comparing two identical characters \texttt{x' = \{1\}} and \texttt{y' = \{1\}}.} lines 221-223.
 
@@
 How have you accounted for this in the study of design to conclude that your results are due to signal, and not bias from the programs used?}}
-
-%TG: really not sure about this comment...
-This is an epistemological point I think and I don't think we can solve this problem, even using true random-number generator.
-The error due to chaos is by definition impossible to take into account...
 
@@
 
-\textit{Each replaced characters (i.e. the characters with a weight of +$1$) where randomly selected from the remaining characters using a uniform distribution (0,1).} lines @@@
+\textit{Each replaced characters (i.e. the characters with a weight of +$1$) where randomly selected from the remaining characters using a uniform distribution (0,1).} lines 296-297.
 
@@
 
-%TG: add some text to the manuscript paraphrasing this? It's kind of already written in there...
-
 \item{\textcolor{blue}{Similarly, why did you choose 0.85 for the Mk model? Does it represent some important set of character sequences? How does it represent the space in general?}}
@@
 
-\textit{We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters (from empirical observations in: Guillerme and Cooper, 2016b; Zou and Zhang, 2016).}
+\textit{We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters (from empirical observations in: Guillerme and Cooper, 2016b; Zou and Zhang, 2016).} lines 256-257.
 
@@
 If the simulations utilised an explicit distribution of character covariance then I think that interpretation of the results would have been more straightforward.}}
+\label{stramlined}
 
@@
 
-We have however made clearer in the introduction what were the objectives of this current manuscript:
-
-%TG: Not sure how to deal with this one.
+We have now clarified in the manuscript what our objectives where by streamlining the last paragraph of the introduction (lines 160-168).
+We have also explicitly added in the supplementary material 3, figs 1, 2 and 3 how the character difference changes between all matrices compared to the ``normal'' matrix.
 
@@
 Revell, 2014).
-However, it is worth noting that in empirical studies, such state changes along the branches are nearly always unknown.}
+However, it is worth noting that in empirical studies, such state changes along the branches are nearly always unknown.} lines 553-558.
 
@@
 
-1- Covariance is approximated by using the same tree for the simulation
-2- The Character difference is an effective metric for measuring covariance for discrete characters.
-This is better than a simple correlation measurement because it makes it a distance it has the properties of a distance metric (which is not the case for correlations with multiple states)
+We do agree with this reviewer's comment and believe that a study looking at difference in covariation levels between character in different morphological modules would be really useful (e.g.
+are cranial character more correlated than postcranial)?
+However we do feel that this is out of the scope of this study.
 
-For example, for three characters \texttt{A = \{1,2,0,0,0\}}, \texttt{B = \{2,3,4,4,4\}} and \texttt{C = \{1,2,1,0,0\}}, $CD_{(A, C)} \leq CD_{(A, B)} + CD_{(B, C)}$ ($0.4 \leq 0 + 0.4$) but this is not the case for a correlation measurement where $cor_{(A, C)} > cor_{(A, B)} + cor_{(B, C)}$ ($0.868 \leq -0.687 + -0.534$).
-Furthermore, using a correlation metric, two character with the same splits (e.g.
-A and B).
-%TG: DOUBLE CHECK THIS
-%TG: Maybe say something along the lines: "yeah, great idea we'll do it later"?
+We do not entirely agree with the reviewer however that our simulations did not ``explicitly simulate covariance'' between characters.
+In fact we do believe that we artificially increased this correlation by sampling all character rates from only on single distribution.
+Since characters where generated based on the same topology (and branch length) and the same distribution, it is highly likely that they would covary a lot.
+This is also reflected by the fact that, increase correlation (reducing character difference), rather few characters needed to be duplicated (see supplementary material 3, figures 1 to 6).
 
@@
 
-%TG: needs to emphasize this point: basically, smaller matrices matrices are more sensible to what the reviewer suggests (i.e.
-the chance of pick up the right or the wrong characters are higher).
-In bigger matrices, this variance is reduced because the number of good/bad characters is diffused.
+We've added the following part to the discussion:
+
+\textit{Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a bigger effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
+If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
+Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.} lines 467-473
 
@@
 
-Secondly, as we highlighted throughout the manuscript (lines @@@), the main mechanisms that generate phylogenetic signal and variance/covariance between characters are the intra-organismal dependence and the evolutionary dependence described in the introduction.
+Secondly, as we highlighted throughout the manuscript, the main mechanisms that generate phylogenetic signal and variance/covariance between characters are the intra-organismal dependence and the evolutionary dependence described in the introduction.
 Both of which can unfortunately only be observed/analysed, \textit{a posteriori} based on a phylogenetic hypothesis.
@@
 We've rephrased this sentence to:
-\textit{Phylogenetic analysis algorithms assume character independence}. lines @@@
+\textit{Phylogenetic analysis algorithms assume character independence} line 2.
 
@@
 We've added the following justification (and toned down the sentence):
-\textit{Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collected them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.}
+\textit{Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collected them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.} lines 29-31.
 
@@
 
-%TG: need a way to phrase this but totally agree, in the reviewer's example, the matrix would look "honest" (lots of characters) but the results would be actually based on way fewer.
+We have removed this section (see Reviewer 3, Major Suggestion \ref{stramlined})
 

### 2019-04-24 Writing/draft_highlighted.tex
Finished response

@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 4 and 5).
+We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 7 and 8).
 Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
@@
 the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
-Furthermore, the inference method, the number of taxa and the number of characters all had a specific effect on the ability to recover the ``correct'' topology.
+\hl{Furthermore, it is important to note that on average, less characters where duplicated (replace) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6).
+However, this change in the amount of replaced characters did not changed the way average character differences changed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
+This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.}
 
@@
 Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance).
+\hl{Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a bigger effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
+If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
+Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.}
 

### 2019-05-20 Writing/response_to_reviewers.tex
Updated Martin's comments

@@
     \item We change the Character Difference metric to be a scaled Hamming Distance rather than the former proposed distance.
-    \item We thus removed the supplementary material 1 wit the proof for the distance metric (since Hamming distances are well established distance metrics).
     \item We reran all the simulations from scratch with the inclusion of the updated Character Difference metric.
-    \item For time constraint reason we reran 20 replicates (rather than 35 in the previous manuscript) for a total of 40 CPU years.
+    \item For time constraint reasons we reran 20 replicates (rather than 35 in the previous manuscript) for a total of 40 CPU years.
     \item We propose a new interpretation to the updated results.
@@
 
-Besides, as the reviewer also correctly points out, we simulated our data with no missing data so this does not affect our results.
+As the reviewer also correctly points out, we simulated our data with no missing data so this does not affect our results.
 The precision in the manuscript was just describing the properties of the new CD metric.
@@
 
-%TG: TODO: add pragraph!
 \textit{Furthermore, it is important to note that on average, less characters where duplicated (replace) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6).
@@
 First, molecular and morphological data often differs in size and in the amount of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but sometimes up to more than ten; Guillerme and Cooper, 2016).
-Second, this discrepancy has been suggested to fewer stronger correlations in morphological data (i.e.
-``concentrated'' correlations) versus weaker and more ``spread'' correlations in molecular data (i.e.
+Second, this difference between both types of characters has been suggested to be stronger correlations in morphological data (i.e.
+``concentrated'' correlations) \textit{versus} weaker and more ``spread'' correlations in molecular data (i.e.
 ``diffused'' correlations - though some regions of the genome can be thought as more correlated than others; Huelsenbeck and Nielsen, 1999).
@@
 
-We do not feel that adding a figure for illustrating the matrix will improve clarity of it's already long and thorough description (accompanying a rather trivial example).
+We do not feel that adding a figure for illustrating the matrix will improve clarity of an already long and thorough description.
 However, we've trying to clarify the description with an example as follows:
 
-\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.} lines 95-96.
+\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.} lines 95-96 and \textit{In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.} lines 102-104.
 
@@
 \textit{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
-In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} more than the other characters since this would be \textit{a priori} the characters that are the most likely to reflect the group's evolutionary history in terms of split by separating the artiodactyles from the cetaceans.
+In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} that seem to be more likely to reflect the split by separating artiodactyles from the cetaceans.
 In other words, this is the correlation between characters that depicts the evolutionary history of this group.
 However, the other characters depict the same history (i.e.
-the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleens, the same split would still be estimated by the software).
-One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.} lines 143-154.
+the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleen, the same split would still be estimated by the software).
+One solution would be to a) weight some characters (e.g.
+giving \textbf{C3} and \textbf{C4} a lower weight) and b) make some characters dependent on each other (e.g.
+making \textbf{C3} and \textbf{C4} dependent on \textbf{C1}).
+However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.} lines 143-154.
 
@@
 
-We doubled checked the figures links and made sure that this mention of Fig.1 now properly points to the workflow figure.
+We double checked the figure links and made sure that this mention of Fig.1 now properly points to the workflow figure.
 
@@
 
-We meant that modifying phylogenetic matrices after the phylogenetic inference is a way to test the effect of the sources of correlation but is not at all a good practice in systematics (it can be seen as dishonest to simply remove characters because they ``don't give the right signal'').
+We meant that modifying phylogenetic matrices after the phylogenetic inference is a way to test the effect of the sources of correlation but is probably not a good practice in systematics.
+Although re-weighting characters based on consistency or retention indices can be used to objectively improve the topology, changing the matrix's characters \textit{a posteriori}, it can be seen as dishonest to remove characters because they ``don't give the right signal'').
 We've removed the end of this sentence however to avoid any further confusion.
@@
 
-\textit{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 12, 13 and 14.} lines 344-346
+\textit{However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.} lines 344-346
 
@@
 
-We are really grateful that this reviewer picked up this flow in our mathematical demonstration of CD metric.
-This problem is due to a lack of indications from our side and we apologise for this.
+We are really grateful that this reviewer picked up this flaw in our mathematical demonstration of the CD metric.
 We have reviewed the formula of the CD metric following an unexpected non-linear behaviour (see comment \ref{metric_change}).
-Furtheremore, we'd like to highlight the following points:
+Furtheremore, we'd like to highlight the following point:
 
-\subsubsection{Character states translation}
-In fact, before being plugged into the CD equation, characters needs to be modified following Felsenstein's \textit{xyz} notation.
-In other words, because the character states tokens are not representing any meaningful numerical value if the characters are unordered (e.g.
-there is no assumption that state 0 should come before state 1), they need to be reordered in a standard way so that the first token in the character is renamed to ``1'', the second to ``2'' etc.
+In fact, before being plugged into the CD equation, characters need to be modified following Felsenstein's \textit{xyz} notation.
+In other words, because the character state tokens are non-metric values if the characters are unordered they can be translated in a standard way so that the first token in the character is renamed to ``1'', the second to ``2'' etc.
 For example, in a 3 state character for four taxa: \texttt{X = {1, 0, 2, 1}}, the character tokens that are integers representing the states could equally have been \texttt{X = {E, A, I, E}} or \texttt{X = {blue, red, absent, blue}}.
-Therefore, to compare characters fairly, the tokens needs to be translated in a standard way such as the first token is replaced by the symbol ``1'', etc.
-In this example \texttt{X = {1, 0, 2, 1}} becomes \texttt{X' = {1, 2, 3, 1}}
+In order to numerically compare X to any other character it is thus more simple to translate both characters in a standard way such as the first token is replaced by the symbol ``1'', etc.
+
+For example if comparing two characters \texttt{X = {1, 0, 2, 1}} and \texttt{Y = {2, 1, 0, 2}}, the absolute difference would be \texttt{X - Y = {-1, -1, 2, -1}}, or when using a Hamming distance \texttt{X - Y = {1, 1, 1, 1}} showing maximum difference between both characters.
+However, because these tokens are non-numeric and effectively represent the same phylogenetic split, both characters are translated to \texttt{X' = {1, 2, 3, 1}} and \texttt{Y' = {1, 2, 3, 1}} leading to a difference of \texttt{X' - Y' = {0, 0, 0,0}}.
+
 In the reviewer's demonstration, \texttt{x  = {1}} and \texttt{y  = {0}} thus become \texttt{x' = {1}} and \texttt{y' = {1}}.
 Hence $CD_{(x',y')} = 0 \Leftrightarrow x' = y'$.
+This is also reflected in the new $CD$ formula that makes comparing two characters for a single taxon nonsensical: the denominator is equal to $0$ if $n = 1$.
 
-\subsubsection{CD as a metric to compare the splits between characters}
 Furthermore, the CD metric, used in this manuscript as a proxy for character correlation actually measures the difference between in characters in terms of phylogenetic signal.
-The character \texttt{X} described above implies 3 splits for four taxa (and of course, the character \texttt{X'} implies the exact same splits).
-Another character \texttt{Y = {1, 1, 1, 2}} will imply only two splits, non shared with \texttt{X} and a character \texttt{Z = {0, 2, 1, 0}} (\texttt{Z' = {1, 2, 3, 1}}) would imply the same splits as \texttt{X} (hence, $CD_{(X',Z')} = 0$ and $CD_{(X',Y')} = 0.5$).
-In other words the probability of X and Z generating conflicting split signal is null and this probability is of 0.5 for X and Y (i.e. half of the splits will be conflicting).
+The character \texttt{X = {1, 0, 2, 1, 0}} described above implies two splits (and of course same goes for the translated \texttt{X' = {1, 2, 3, 1, 2}}).
+Another character \texttt{Y = {1, 1, 1, 2, 2}} will imply only one split, not shared with \texttt{X} and a character \texttt{Z = {0, 2, 1, 0, 2}} (\texttt{Z' = {1, 2, 3, 1, 2}}) would imply the same split as \texttt{X} (hence, $CD_{(X',Z')} = 0$ and $CD_{(X',Y')} = 0.75$).
+In other words the probability of X and Z generating conflicting split signal is null and this probability for X and Y is of 0.75.
 

### 2019-05-20 Writing/draft_highlighted.tex
Updated Martin's comments

@@
 \hl{).
-Second, this discrepancy has been suggested to be stronger correlations in morphological data (i.e.
+Second, this difference between both types of characters has been suggested to be stronger correlations in morphological data (i.e.
 ``concentrated'' correlations) \textit{versus} weaker and more ``spread'' correlations in molecular data}
@@
     This coding dependency is linked to hierarchical dependency between characters \citep{wilkinson1995coping,BrazeauNA}.
+    \hl{In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.}
     Finally this can also be due to a bias in the amount of characters available.
@@
 \hl{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
-In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} more than the other characters since this would be \textit{a priori} the characters that are the most likely to reflect the group's evolutionary history in terms of split by separating the artiodactyles from the cetaceans.
+In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} that seem to be more likely to reflect the split by separating artiodactyles from the cetaceans.
 In other words, this is the correlation between characters that depicts the evolutionary history of this group.
 However, the other characters depict the same history (i.e.
-the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleens, the same split would still be estimated by the software).
-One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.}
-% \hl{Furthermore, phenotypic variations can be continuous rather than discrete and ``discretising'' such characters can have some influence on tree inference} \citep{parins2017use}.
-% \hl{In this case, ``discretised'' }.
+the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleen, the same split would still be estimated by the software).
+One solution would be to a) weight some characters (e.g.
+giving \textbf{C3} and \textbf{C4} a lower weight) and b) make some characters dependent on each other (e.g.
+making \textbf{C3} and \textbf{C4} dependent on \textbf{C1}).
+However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.}
 This illustrates how some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
@@
 When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
-\hl{However, we did report the comparison between the ``true'' trees and ``normal'' and ``randomised'' ones is available in the supplementary materials 3, Figs 12, 13 and 14.}
+\hl{However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.}
 

### 2019-05-20 Writing/response_to_reviewers.tex
reviewer 2 done with the random numbers detailed

@@
 
-We've added more explanations in the supplementary materials 1 on (1) the character translation, (2) the four-point metric proof and (3) on how to interpret the $CD$ metric.
+We've correct the supplementary materials 1 and added more explanations on (1) the character translation, (2) the four-point metric proof and (3) on how to interpret the $CD$ metric.
 Additionally, we added more precisions in the text:
@@
 \textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines 199-200.
-
-\textit{Similarly, an hypothetical comparison of two character for only one taxa \texttt{x = \{0\}} and \texttt{y = \{1\}} would result in comparing two identical characters \texttt{x' = \{1\}} and \texttt{y' = \{1\}}.} lines 221-223.
 
@@
 We do agree that generating true randomness in current software is tricky but we are not sure how to answer this comment.
-In our simulations, random numbers where used at different levels:
+The three first steps (generating the tree, the matrices and modifying the matrices) were using the default random number generator implemented in R, the Mersenne Twister algorithm that has a cycle of $2^{32}-1 = 4.2e^{9}$ unsigned random integers (see random numbers documentation in R: \texttt{?.Random.seed}).
+The last step (generating the trees) was using both the random number generated by default my MrBayes and PAUP* based on the seeds from the different HPC we used (the Imperial College London Cluster Services and the Australian QRIScloud Awoonga cluster).
 
 \begin{enumerate}
-    \item The generation of the ``true'' tree.
-    \item The generation of the discrete morphological characters (each 100, 350 and 1000).
-    \item The removal of the characters (for the ``randomised'' matrices).
-    \item The starting seeds for the tree inferences (in both MrBayes and PAUP*).
+    \item \textbf{The generation of the ``true'' tree.}
+    The generation of the true tree involves the drawing of two random parameters (birth and death parameters) and then builds the tree by drawing a probability of branching (birth) or extinction (death) at each time point until it reaches the desired number of taxa.
+    This means at least $2 + 2tn$ where $t$ are the evaluated time points and $n$ the number of required taxa.
+    For all 60 trees we thus have $120 + 500t$ draws.
+This means this whole step would cycle through all random numbers only if $t > 8.4e^{6}$.
+As a magnitude of comparisons when $t$ is one order of magnitude lower, $t = 1e^{5}$ it will typically lead to trees with $>> 1e{4}$ taxa (i.e.
+2 orders of magnitude bigger than our biggest trees).
+
+    \item \textbf{The generation of the discrete morphological characters (each 100, 350 and 1000).}
+    The number of seeds involved in the generation of each model depends on the two models with a total of $5 + (2n-2) = 2n + 3$ random numbers for the HkY-binarised model (4 states proportions + 1 rate + number of branches) and $2 + (2n-2) = 2n$ for the Mk model (1 states numbers + 1 rate + number of branches).
+    If if we simplify the seed calculations by saying only the HKY-model is chosen, we have at maximum $20(2n + 3)c$ random numbers; where $c$ is the number of characters.
+    This results in a maximum of $1.4e{7}$ random numbers ($<< 4.2e^{9}$ total random numbers).
+
+    \item \textbf{The removal of the characters (for the ``randomised'' matrices).}
+    This step is also variable but we can simplify it by assuming the extreme case where 1) all characters are selected to be replaced and 2) they are all replaced by all characters.
+This results in a maximum of $20(5c)$ random numbers per matrix (2 for both replacements and 3 for each ``minimised'', ``maximised'' and ``randomised'' matrices).
+    This results in a maximum of $1.4e{5}$ random numbers ($<< 4.2e^{9}$ total random numbers).
 \end{enumerate}
 
-Because each of this steps are dependent in a cascade (i.e.
-the random seed used to generated the ``true'' tree influence the generation of the discrete characters which influences the ``randomised'' matrix which influences the results from the random starting seed in the tree inference).
-However, we consider the influence of these cascading random numbers negligible in this analysis.
-Furthermore, we assumed that the multiplication of negligible effects will result in a near null error due to the pseudo random-number generators.
-In fine, our results are bound to be somehow biased by our prior believes and experience as well as by the tools (hardware and software) and the methods used in this manuscript but this is more of an epistemological point and is not specific to this study.
+Therefore, if cumulating all the seeds used in the generation of the matrices (trees + characters) and the modifications of the matrices is likely to be no more than $6.4e^{7}$.
+Therefore we do not believe that the random number generation was a proper since we sampled two orders of magnitude less random numbers than the total ones available.
 
@@
 
-Every replaced characters where randomly selected from a uniform distribution (0,1). We have now specified this in the text:
+Every replaced characters were randomly selected randomly from the total sample of characters (i.e.
+randomly sampling from the uniform integers distribution (1, $c$); where $c$ is the number of remaining characters).
+We have now specified this in the text:
 
-\textit{Each replaced characters (i.e. the characters with a weight of +$1$) where randomly selected from the remaining characters using a uniform distribution (0,1).} lines 296-297.
+\textit{Each replaced characters (i.e. the characters with a weight of +$1$) were randomly selected by sampling them randomly from the list of the remaining characters.} lines 296-297.
 
@@
 
-There is actually a vigorous debate surrounding how to generate biologically meaningful morphological characters \citep[][;confidential reviews and personal communications witht he authors]{OReilly20160081,goloboff2017weighted,puttick2017uncertain,GoloboffEmpirical,OReillyEmpirical}.
+There is actually a vigorous debate surrounding how to generate biologically meaningful morphological characters \citep[][]{OReilly20160081,goloboff2017weighted,puttick2017uncertain,GoloboffEmpirical,OReillyEmpirical}.
 This ongoing debate, however, focuses mainly on inference methods where data simulation is a tangential aspect of it.
@@
 Appendix A: Tree inference) and \cite{ZouConvergence}.
-We've specified this is the text:
+We've specified this in the text:
 
@@
 
-We fixed this type. We considered them different if they had an overlap probability $<$ 0.05.
+We fixed this typo. We considered them different if they had an overlap probability $<$ 0.05.
 
@@
 
-The significances was referring to the results in Table 5 (Wilcox test p value of 0 and 0.08 and Bhattacharrya Coefficient of 0.891 and 0.984 respectively) and the median values were extracted from the supplementary material 3 Table 5.
+The significances was referring to the results in Table 5 (Wilcoxon test $p$ value of 0 and 0.08 and Bhattacharrya Coefficient of 0.891 and 0.984 respectively) and the median values were extracted from the supplementary material 3 Table 5.
 We now detailed the values and where they are present for each of the tests to avoid confusion.

### 2019-05-20 Writing/draft_highlighted.tex
reviewer 2 done with the random numbers detailed

@@
 \noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
-%TG: Modifications in this sentence not in the letter yet
 $CD$ is a continuous \hl{scaled Hamming distance} metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
@@
 This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
-\hl{Similarly, an hypothetical comparison of two character for only one taxa \texttt{x = \{0\}} and \texttt{y = \{1\}} would result in comparing two identical characters \texttt{x' = \{1\}} and \texttt{y' = \{1\}}.}
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).
-
 When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
@@
 randomly weighting characters).
-\hl{Each replaced characters (i.e. the characters with a weight of +$1$) where randomly selected from the remaining characters using a uniform distribution (0,1).}
+\hl{Each replaced characters (i.e. the characters with a weight of +$1$) were randomly selected by sampling them randomly from the list of the remaining characters.}
 Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.

### 2019-05-20 Writing/supplementary/Sup1-CharDiff.tex
reviewer 2 done with the random numbers detailed

@@
 Therefore, to compare characters fairly, the tokens needs to be translated in a standard way such as the first token is replaced by the symbol ``1'', etc.
-In this example \texttt{X = {1, 0, 2, 1}} becomes \texttt{X' = {1, 2, 3, 1}}
+In this example \texttt{X = {1, 0, 2, 1}} becomes \texttt{X' = {1, 2, 3, 1}}.
+
+Note that the first two translated tokens are always equal (\texttt{x = {1, ...}} and \texttt{y = {1, ...}}) hence the denominator of $CD$ being $n - 1$ to reflect this equality.
+Because of this, $CD$ can only be applied for $n \in [2;\infty)$ (i.e. if $n = 1$, the denominator is 0).
 
 \subsection{Interpretation}
-For a characters \texttt{X = {1, 0, 2, 1}} implying three splits between four taxa and two characters \texttt{Y = {1, 1, 1, 2}} and \texttt{Z = {0, 2, 1, 0}} implying respectively two and three splits between four taxa, $CD_{(X,Y)} = 0.5$ and $CD_{(X,Z)} = 0$.
+For a character \texttt{X = {1, 0, 2, 1, 0}} implying two splits between five taxa and two characters \texttt{Y = {1, 1, 1, 2, 2}} and \texttt{Z = {0, 2, 1, 0, 2}} implying respectively one and two splits between five taxa, $CD_{(X,Y)} = 0.75$ and $CD_{(X,Z)} = 0$.
 This can be interpreted as the probability of characters \texttt{X}, \texttt{Y} and \texttt{X}, \texttt{Z} giving a conflicting phylogenetic signal between the four taxa.
-Here \texttt{X}, \texttt{Y} will conflict for 50\% of the taxa and \texttt{X}, \texttt{X} for 0\% of them.
+Here \texttt{X}, \texttt{Y} will conflict for 70\% of the taxa and \texttt{X}, \texttt{X} for 0\% of them.
 
@@
 
-For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ because the difference between character states is considered as Fitch-like (unordered; i.e.
+For any $n \in [2;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ because the difference between character states is considered as Fitch-like (unordered; i.e.
 the difference between two different characters states is equal to $1$ and the difference between two identical character states is equal to $0$).
@@
 
-For any $n \in [1;\infty)$ and $x_{n} = y_{n}$ then:
+For any $n \in [2;\infty)$ and $x_{n} = y_{n}$ then:
 %TG: Is that valid for a demonstration?
@@
 
-For any $n \in [1;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$, similarly to equation \ref{positive_demo}:
+For any $n \in [2;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$, similarly to equation \ref{positive_demo}:
 
@@
 
-For any $n, \in [1;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:
+For any $n, \in [2;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:
 
@@
 
-For any $n \in [1;\infty)$ and any four distinct characters $i$, $j$, $n$ and $m$, and if equations \ref{Positivity} and \ref{Subadditivity} are true, the pairwise distances $CD_{(i,j)}$, $CD_{(m,n)}$, $CD_{(i,m)}$, $CD_{(j,n)}$, $CD_{(j,m)}$ and $CD_{(i,n)}$ can be represented as a combination of any five values $a$, $b$, $c$, $d$ and $e$ $\geq 0$:
+For any $n \in [2;\infty)$ and any four distinct characters $i$, $j$, $n$ and $m$, and if equations \ref{Positivity} and \ref{Subadditivity} are true, the pairwise distances $CD_{(i,j)}$, $CD_{(m,n)}$, $CD_{(i,m)}$, $CD_{(j,n)}$, $CD_{(j,m)}$ and $CD_{(i,n)}$ can be represented as a combination of any five values $a$, $b$, $c$, $d$ and $e$ $\geq 0$:
 

### 2019-05-20 Writing/response_to_reviewers.tex
finished updated response

@@
 We have also updated the Character Difference metric characterisation in the supplementary material 1.
-However, since it is now simply a scaled Hamming distance, these demonstrations might be unwarranted for this revised version of the manuscript.
 
@@
 \textit{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
-Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).} lines 210-215.
+Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).} lines 213-218.
 
@@
 However, this change in the amount of replaced characters did not changed the way average character differences changed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
-This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.} lines 452-461.
+This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.} lines 453-462.
 
@@
 changes in morphological states can be more readily linked to a function cf.
-changes in nucleotide states ; see below).} lines 43-57.
+changes in nucleotide states ; see below).} lines 43-58.
 
@@
 making \textbf{C3} and \textbf{C4} dependent on \textbf{C1}).
-However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.} lines 143-154.
+However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.} lines 145-157.
 
@@
 
-\textit{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in Felsenstein, 2004; as used in D\'{a}valos et al., 2014).} lines 218-220.
+\textit{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in Felsenstein, 2004; as used in D\'{a}valos et al., 2014).} lines 221-223.
 
@@
 \textit{Note that both models (``morphological HKY'' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
-However, the ``morphological HKY model'' is not the one} directly \textit{estimated in the phylogenetic software used in this protocol (MrBayes, see below).} lines 270-274.
+However, the ``morphological HKY model'' is not the one} directly \textit{estimated in the phylogenetic software used in this protocol (MrBayes, see below).} lines 271-275.
 
@@
 
-\textit{However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.} lines 344-346
+\textit{However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.} lines 346-347
 
@@
 
-\textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines 199-200.
+\textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines 202-203.
 
@@
 We do agree that generating true randomness in current software is tricky but we are not sure how to answer this comment.
-The three first steps (generating the tree, the matrices and modifying the matrices) were using the default random number generator implemented in R, the Mersenne Twister algorithm that has a cycle of $2^{32}-1 = 4.2e^{9}$ unsigned random integers (see random numbers documentation in R: \texttt{?.Random.seed}).
+The three first steps (generating the tree, the matrices and modifying the matrices) were using the default random number generator implemented in R, the Mersenne Twister algorithm that has a cycle of $2^{32}-1 = 4.2e^{9}$ unsigned random integers (see random numbers documentation in R: \texttt{?.Random.seed} - we used the implementation from \texttt{R} v3.5 that differs in the starting seed from the new \texttt{R} v3.6).
 The last step (generating the trees) was using both the random number generated by default my MrBayes and PAUP* based on the seeds from the different HPC we used (the Imperial College London Cluster Services and the Australian QRIScloud Awoonga cluster).
@@
 
-\textit{Each replaced characters (i.e. the characters with a weight of +$1$) were randomly selected by sampling them randomly from the list of the remaining characters.} lines 296-297.
+\textit{Each replaced characters (i.e. the characters with a weight of +$1$) were randomly selected by sampling them randomly from the list of the remaining characters.} lines 297-299.
 
@@
 
-\textit{We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters (from empirical observations in: Guillerme and Cooper, 2016b; Zou and Zhang, 2016).} lines 256-257.
+\textit{We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters (from empirical observations in: Guillerme and Cooper, 2016b; Zou and Zhang, 2016).} lines 256-258.
 
@@
 
-As noted in the introduction, the aim of this manuscript was mainly to see how do software handle correlation between discrete morphological characters.
-In other words, how can we distinguish between the complex sources of correlation in phenotype using only discrete morphological characters?
-We do agree with the reviewer that a thorough analysis of the variance/covariance structure of simulated ``discretised'' characters (see reviewer 3, major suggestion \ref{hiden}) would have broad more insight into the nature of character correlation on phylogenetics inference in general (regardless of the type of character and correlation) but we feel that this is out of the scope of this study.
+As noted in the introduction, the aim of this manuscript was mainly to see does different levels of correlation between discrete morphological characters affect tree topology.
+We also review different sources of correlation for characters and discuss how the multidimensional complexity of a phenotype (and the correlations therein) is not reflected in two dimensional character matrices.
+We do agree with the reviewer that a thorough analysis of the variance/covariance structure of simulated ``discretised'' characters (see reviewer 3, major suggestion \ref{hiden}) would have broad more insight into the nature of character correlation on phylogenetics inference in general (regardless of the type of character and correlation) but due to the length required to run the current ``simplistic'' simulation, we feel that this would be out of the scope of this manuscript.
 
-We have now clarified in the manuscript what our objectives where by streamlining the last paragraph of the introduction (lines 160-168).
+We have now clarified in the manuscript what our objectives were by streamlining the last paragraph of the introduction (lines 160-168).
 We have also explicitly added in the supplementary material 3, figs 1, 2 and 3 how the character difference changes between all matrices compared to the ``normal'' matrix.
@@
 
-We agree with this reviewer's comment which can be especially thought to be the case for discrete with a fast rate of evolution.
+We agree with this reviewer's comment which can be especially thought to be the case for discrete characters with a fast rate of evolution.
 Such characters will likely swap between states along the branches following a hidden Markov chain.
-Unfortunately, however, although the state of a character on a branch can be modelled, it is nearly always impossible to actually observe them.
+
+However, we feel that modelling this process might not be useful in this specific study:
+First, in empirical data two characters can be strongly correlated for biological and evolutionary reasons (e.g.
+the absence of an astragalus and the presence of baleen in table 1) leading to a strong correlation of both characters at the tips (or even nodes) but does not mean that both character's hidden Markov chain are correlated.
+Second, by comparing characters at any time point rather than simply at the tips (which can be problematic since our trees were not all ultrametric) we would probably not be able to compare the effect of the ``true'' characters' correlation since this information will not be used by the phylogenetic software (MrBayes and PAUP*) that will only use tip values.
 
@@
 \textit{It is also worth noting that character similarity at the tips does not necessarily reflects that the characters underwent the same evolutionary history.
-This can specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches (e.g.
+This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches (e.g.
 Revell, 2014).
-However, it is worth noting that in empirical studies, such state changes along the branches are nearly always unknown.} lines 553-558.
+However, it is worth noting that in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).} lines 554-560.
 
@@
 are cranial character more correlated than postcranial)?
-However we do feel that this is out of the scope of this study.
+In our simulation protocol we effectively only simulated characters from a single ``module'' since they were generated based on the same topology (and branch length) and the same distribution (rates where sampled from a single distribution).
+Therefore, we believe that the correlation approximated by the $CD$ metric was actually explicitly simulated through our protocol.
+This is also reflected by the fact that, in practice, when increasing the correlation (reducing character difference), rather few characters needed to be duplicated (see supplementary material 3, figures 1 to 6).
 
-We do not entirely agree with the reviewer however that our simulations did not ``explicitly simulate covariance'' between characters.
-In fact we do believe that we artificially increased this correlation by sampling all character rates from only on single distribution.
-Since characters where generated based on the same topology (and branch length) and the same distribution, it is highly likely that they would covary a lot.
-This is also reflected by the fact that, increase correlation (reducing character difference), rather few characters needed to be duplicated (see supplementary material 3, figures 1 to 6).
+We hope to expand this study in the future by create different groups of characters sharing the same overall correlation by generating ``bunches'' of characters based on different rates (to simulate rate homogeneity) and different trees (branch length and topology - to simulate variation in ``gene'' trees).
 
@@
 If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
-Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.} lines 467-473
+Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.} lines 468-474
 

### 2019-05-20 Writing/draft_highlighted.tex
finished updated response

@@
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
-    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep[\hl{from empirical observations in: }][]{Guillerme2016146,ZouConvergence}.
+    \hl{We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters} \citep[\hl{from empirical observations in: }][]{Guillerme2016146,ZouConvergence}.
     This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g.
@@
 \hl{It is also worth noting that character similarity at the tips does not necessarily reflects that the characters underwent the same evolutionary history.
-This can specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches}
+This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches}
 \citep[e.g.][]{revell2014ancestral}.
-\hl{However, it is worth noting that in empirical studies, such state changes along the branches are nearly always unknown.}
+\hl{However, it is worth noting that in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).}
 

### 2019-05-30 Writing/draft_highlighted.tex

@@
 \begin{document}
-
-%Running head
-\begin{flushright}
-Version dated: \today
-\end{flushright}
+\[\[\[%Running head
+%\begin{flushright}
+%Version dated: \today
+%\end{flushright}
 \bigskip
 \noindent RH: Characters correlation
-
 \bigskip
@@
 \begin{center}
-
 \noindent{\Large \bf Influence of different modes of morphological character correlation on phylogenetic tree inference}
@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
+\noindent {\normalsize \sc Thomas Guillerme $^{1,2,*}$, and Martin D. Brazeau $^{2,3}$}\\
 \noindent {\small \it 
-$^1$School of Biological Sciences, University of Queensland, St. Lucia, Queensland, Australia.\\
+$^1$ School of Biological Sciences, University of Queensland, St. Lucia, Queensland, Australia.\\
 $^2$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\
@@
 Overall, we found \hl{a negative} effect of reducing character correlations on recovering the correct topology.
-However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where \hl{reducing or increasing} character correlation is not more effective than randomly drawing characters.
+However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where \hl{reducing or increasing} character correlation is not more effective EFFECTIVE FOR WHAT?
+than randomly drawing characters.
 Furthermore, in bigger matrices (350 characters), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
@@
 This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
-\hl{Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collected them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.}
+\hl{Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collecting them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.}
 While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
@@
 \hl{This independence is assumed for both molecular and morphological characters} \citep{huelsenbeck1999effect,Davalos01072014,ZouConvergence}.
-\hl{However, several distinctions can be made between the amount and the nature of the correlation.
-First, molecular and morphological data often differs in size and in the amount of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but sometimes up to more than ten;}
+\hl{However, threel distinctions can be made between the amount and the nature of the correlation.
+First, molecular and morphological data often differs in size and in the amount of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but sometimes more than ten;}
 \citealt{Guillerme2016146}
 \hl{).
-Second, this difference between both types of characters has been suggested to be stronger correlations in morphological data (i.e.
+Second, characters has been suggested to have stronger correlations in morphological data (i.e.
 ``concentrated'' correlations) \textit{versus} weaker and more ``spread'' correlations in molecular data}
 \citep[\hl{i.e. ``diffused'' correlations - though some regions of the genome can be thought as more correlated than others}][]{huelsenbeck1999effect}.
-\hl{Furthermore, because of this ``spread'' correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
+\hl{Furthermore, because of the ``spread'' correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
 changes in morphological states can be more readily linked to a function \textit{cf.} changes in nucleotide states ; see below).}
-Before discussing character correlation further, it is important to understand that it may manifest itself in at least three distinct ways:
+Before discussing character correlation further, it is important to understand that correlation may manifest itself in at least three distinct ways:
 
@@
 
-\noindent \textbf{Correlation between characters detected by the software}: this is the result how software actually interprets the differences between characters.
+\noindent \textbf{Correlation between characters detected by the software}: this is the result of how software actually interprets the differences between characters.
 The vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different or similar tokens).
@@
 For example, if we consider the following matrix containing four cetartiodactyls - say a pig (e.g.
-\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley:
+\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
 
@@
 In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
-Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states tokens will be used in order to infer the tree.
+Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states' tokens will be used in order to infer the tree.
 \hl{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
@@
 However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.}
-This illustrates how some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
+This illustrates how some characters will therefore be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
 Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).
@@
 Here we formally assess the effect of discrete character's correlation using simulated data.
-We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences or similarities between characters.
+We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease WHICH ONE?
+the overall differences or similarities between characters.
 We found that overall, there is a detectable effect of character correlation on topology where \hl{a decrease} in character dependence results in a decrease in the ability to recover the correct topology.
@@
 
+Well this is worrying-because it seems like authors could use these results to justify introducing or ignoring character correlations.
+It seems like it would be important to test the effect of correlations in uninformative vs informative characters.
+Reconstructors wouldn't know whether each character is informative or not, so they would more likely try to avoid it?
+But yeah, maybe that would take another year?
 \section{Methods}
@@
 This step resulted in 72 topologies.
-    \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on topology.
+    \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on the recovered topology.
 \end{enumerate}
-Each step was replicated \hl{20} times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
+All steps were replicated \hl{20} times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
 
@@
 \end{equation}
-
+%This is not making sense to me right now... I'll ask you about it when you get home. Nope I got it now.
 %CD is similar to https://onlinelibrary.wiley.com/doi/full/10.1111/pala.12384 but by character rather than by species
@@
 Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different set of splits (i.e. a different phylogenetic signal).
+%UG FRUSTRATING -A LOT OF MY EDITS HAVE GOTTEN DELETED :(   OH WELL I'LL JUST PUT BACK THE SUPER IMPORTANT ONES. x
+%soooo CD_AC = (0+0+1+1)/(4-1) =2/3  right now.
+C needs to change to C={0,0,2,3}  then you get (0+1+1+1)/3=1.
+I suggest putting the rough calculation in like I demonstrated-because it's a little bit confusing to figure out.
 
@@
 
+%I can't find Table 2 in this document...
+
+%I strongly suggest that you labelled these figures more on the fig-put less in the legend-give them titles of Robinson Ford or triplet whatever.
+Then in the first line of the fig text- you can say "Robinson Ford distance shows the degree of clade conservatism whatever that means...
+"  so the reader can immediately know the difference between fig 2 and 3.
+To label: (1) text on ylab "Normalized tree similarity"- then each row can just be "# taxa" (2) mtext on xlab "Number of characters"  (3)  legend c("more correlated", "less correlated", "randomised") but leave bayesian and parsimony for the fig text.
+The goal being to know what the figure says at a glance-the fig text can help clarify
 \begin{figure}[!htbp]
@@
 \hl{There is however no difference in terms of taxa displacement ($NTS_{Tr}$) between the three scenarios (with a BC > 0.98 for each comparison - Table}
-\ref{Tab_pooledscenarios_test}).
+\ref{Tab_pooledscenarios_test}).  HERE  CAN YOU GIVE A LITTLE MORE INTUITION ON WHAT THE RESULTS MEAN
+
+%maybe this is table 2?
+You say "differennce between the pooled scenarios"  I don't remember what you are talking about-maybe describe the pooled scenarios for the forgetful reader.
+I have no idea what table 2 is even saying-the table text could use more hand holding explanations.
+# same issues with table 3.
+I think I personally would've done these stats differently as a multilevel model so that the bonferroni corrections weren't necessary-but I suppose it doesnt fucking matter with p values like that.
+However, table 4 is perfect: don't change a thing.
 
@@
 
-\hl{When looking at the combined effect of each parameters, the ``maximised'', ``minimised'' and ``randomised'' scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).}
-\hl{When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences for the taxa displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).}
+\hl{When looking at the combined effect of each parameter, the ``maximised'', ``minimised'' and ``randomised'' scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).}
+\hl{When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences  between  tree similarity of differening numbers of taxa, characters, tree correlation, or tree construction method for the taxa displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).}
 \hl{In terms of clade conservation ($NTS_{RF}$), however, ``maximised'' and ``minimised'' scenarios produced nearly always consistently different topologies using Bayesian inference (apart for the matrices with 25 taxa and 100 and 350 characters and the matrices with 150 taxa and 350 characters) but not using parsimony inference (only the matrices with 75 taxa and 100 or 350 characters produced different $NTS_{RF}$ distributions - supplementary material 3, tables 19 and 20).}
@@
 \subsection{Effect of character differences on topology}
-As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
+As expected, there is a significant effect of the character difference in the ability to recover the correct topology.   This is only true for the RF method, no?
 The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
-Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
-\hl{Overall, throughout our simulation, the ``minimised'' character difference scenario (i.e.
-increased correlation) tend to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e.
+Intuitively, ONE WOULD HYPOTHESIZE THAT the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (very homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
+I DONT SEE ANY DATA THAT SUPPORTS THIS STATEMENT
+\hl{HOWEVER, throughout our simulation, the ``minimised'' character difference scenario (i.e.
+increased correlation) tendED to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e.
 decreased correlation) led to slightly worst trees (Fig.}
 \ref{Fig:RF_results_best}.
-\hl{This did not however seem to affect the taxa placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical results (Fig.}
+\hl{This (WHAT IS THIS?,  CHARACTER CORRELATION?) did not however APPEAR to affect the taxa placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical DIFFERENCES (Fig.}
 \ref{Fig:Tr_results_best}).
 \hl{However this results needs to be contrasted by the fact that}
-the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
-\hl{Furthermore, it is important to note that on average, less characters where duplicated (replace) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6).
-However, this change in the amount of replaced characters did not changed the way average character differences changed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
+the trees generated by the ``minimised'' CORRELATION scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
+\hl{Furthermore, it is important to note that on average, less characters where duplicated (replaceD) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices   WHY DID THIS HAPPEN?WHY IS IT IMPORTANT ?
+(supplementary material 3, Figs 4, 5 and 6).
+However, (THIS CLAUSE IS CONFUSING AND KIND OF SOUNDS LIKE BULLSHIT ) this change in the amount of replaced characters did not change the way average character differences VARYed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
 This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.}
+I SUDDENLY REALLY DONT LIKE THE TERM CHARACTER DIFFERENCES-ITS THE OPPOSITE OF CHARACTER CORRALTION?
+IT DOESNT SOUND LIKE IT INTUITIVELY.
+ITS KIND OF CONFUSING TO HAVE SO MANY TERMS MEANING THE SAME THING -MAXIMIZED V MINIMIZED, CHARACTER CORRALATION, CHARACTER DIFFERENCE.
+I'M TIRED AND GRUMPY.
 
@@
 
-Although we measured a significant difference between small and larger matrices (\hl{in terms of number of characters, between 100 and 1000 characters - $BC$ $NTS_{RF}=0.914$ and $BC$ $NTS_{Tr}=0.826$}; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}),
-these differences have to be contrasted with the probability of overlap between the results distributions that are always \hl{very} high between every matrices sizes (Bhattacharrya Coefficients $>0.949$ for both the different number of characters and taxa).
+Although we measured a significant difference between smaller and larger matrices (\hl{in terms of number of characters, between 100 and 1000 characters - $BC$ $NTS_{RF}=0.914$ and $BC$ $NTS_{Tr}=0.826$}; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}),
+these differences have to be contrasted with the probability of overlap between the results distributions that are always \hl{very} high between every matrix's size (Bhattacharrya Coefficients $>0.949$ for both the different number of characters and taxa).
 This suggests that the overall character difference metric \hl{can be used as a proxy} for character correlation as it is \hl{not directly dependent} on the number of characters analysed.
 Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data.
+DID YOU TALK ABOUT POOLING DATA IN THE METHOSA?
+I'M NOT SURE WHAT YOU ARE REFERING TO.
 Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
@@
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[i.e.
-those with in a $CI>0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
+those with in a $CI>0.25$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
@@
 
-The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
-For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
-(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
+The differences in performance of the BAYESIAN AND PARSIMONY APPROACHES observed here could be due to the inherent mechanisms of each method.
+For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scoreS differently:
+(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the (WHICH TREE?
+TRUE TREE?)tree.
 The tree search will thus likely result in a topology island that will not contain the given topology $T$.
@@
 in evo-devo \citealt{goswami2006morphological}; or in macroevolution \citealt{fitzjohn2014much}).
-These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not used \textit{a priori} as a criterion to select characters.
+These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not BE used \textit{a priori} as a criterion to select characters.
 In other words, intra-organismal and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
@@
 This dichotomy thus creates a trade \hl{off} between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
-and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation being less dependent on the coding correlations.
+and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation TO BE less dependent on the coding correlations.
 
@@
 However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices.% which are - when using empirical data - at best bad practice and at worst dishonest.
+I DONT  REALLY UNDERSTAND
 
@@
 Third, comparing phylogenetic inference methods is not trivial.
-As mentioned above, both maximum parsimony and Bayesian inference, although aiming (and often achieving) to infer evolutionary history only have similar outputs and vastly differ in how optimality is measured.
+As mentioned above, both maximum parsimony and Bayesian inference, although aiming (and often achieving) to infer evolutionary history, only have similar outputs and vastly differ in how optimality is measured.
 But there are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
-However, we want to point out again that here we're no comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
+However, we want to point out again that here we're noT comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
 
@@
 Effectively, our simulation protocol bootstraps our data ``with bias''.
-In the ``randomised'' scenarios the data is simply randomly bootstrapped simply we randomly remove and re-sample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
+In the ``randomised'' scenarios the data is simply randomly bootstrapped, simply we randomly remove and re-sample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
 However, in the ``minimised'' and ``maximised'' scenario, the bootstrapping we remove the characters with the lowest/highest overall character difference.
@@
 Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.
+HMM IS THIS CIRCULAR LOGIC?
+YOU KNOW THAT CHARACTERS ARE ON AVERAGE INFOMRATIVE-SO OF COURSE INCREASING THE REPRESENTATION OF INFROMATIVE CHARACTERS INCREASES YOUR ABILITY TO CORRELTY RSOLVE.
+IF CHRACTERS WERE UNIFORMATIVE AND CORRELATED THIS RESULT WOULD NOT HOLD
 
@@
 Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
+(DID YOU LOOK AT WHICH TOPOLOGIES WERE INDUCED??)
 
@@
 Conversely, when correlation between characters was high, the topology was always the closest to the ``normal'' topology.
-These results seem independent on the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
+These results APPEAR independent oF the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
 
-However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise \hl{or maximised} character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
+However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise \hl{or maximised} character correlation was not more significant than simply RANDOMINSING (I.E.
+bootstrapping) the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
 For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly increased and decreased correct topological recovery compared to randomly resample matrices.
@@
 
+\]
+\]
+\]
 \end{document}

### 2019-05-31 Writing/draft_highlighted.tex
Updated Abigail's comments

@@
 \begin{document}
-\[\[\[%Running head
-%\begin{flushright}
-%Version dated: \today
-%\end{flushright}
+%Running head
+\begin{flushright}
+Version dated: \today
+\end{flushright}
+
 \bigskip
@@
 
-\noindent {\normalsize \sc Thomas Guillerme $^{1,2,*}$, and Martin D. Brazeau $^{2,3}$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
+% Abigail I. Pastore$^{1}$
 \noindent {\small \it 
-$^1$ School of Biological Sciences, University of Queensland, St. Lucia, Queensland, Australia.\\
+$^1$School of Biological Sciences, University of Queensland, St. Lucia, Queensland, Australia.\\
 $^2$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\
@@
 Overall, we found \hl{a negative} effect of reducing character correlations on recovering the correct topology.
-However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where \hl{reducing or increasing} character correlation is not more effective EFFECTIVE FOR WHAT?
-than randomly drawing characters.
+However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where \hl{reducing or increasing} character correlation is not more effective \hl{in getting the expected topology} than randomly drawing characters.
 Furthermore, in bigger matrices (350 characters), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
@@
 \hl{This independence is assumed for both molecular and morphological characters} \citep{huelsenbeck1999effect,Davalos01072014,ZouConvergence}.
-\hl{However, threel distinctions can be made between the amount and the nature of the correlation.
+\hl{However, several distinctions can be made between the amount and the nature of the correlation.
 First, molecular and morphological data often differs in size and in the amount of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but sometimes more than ten;}
@@
 Here we formally assess the effect of discrete character's correlation using simulated data.
-We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease WHICH ONE?
-the overall differences or similarities between characters.
+We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
 We found that overall, there is a detectable effect of character correlation on topology where \hl{a decrease} in character dependence results in a decrease in the ability to recover the correct topology.
@@
 
-Well this is worrying-because it seems like authors could use these results to justify introducing or ignoring character correlations.
+%AP Well this is worrying-because it seems like authors could use these results to justify introducing or ignoring character correlations.
 It seems like it would be important to test the effect of correlations in uninformative vs informative characters.
@@
 \end{equation}
-%This is not making sense to me right now... I'll ask you about it when you get home. Nope I got it now.
 %CD is similar to https://onlinelibrary.wiley.com/doi/full/10.1111/pala.12384 but by character rather than by species
@@
 \textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
-For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
+For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,0,1,2\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$ \hl{$CD_{(A,B)} = (0+0+0+0)/(4-1) = 0$ and $CD_{(A,C)} = (0+1+1+1)/(4-1) = 1$}.
 Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different set of splits (i.e. a different phylogenetic signal).
-%UG FRUSTRATING -A LOT OF MY EDITS HAVE GOTTEN DELETED :(   OH WELL I'LL JUST PUT BACK THE SUPER IMPORTANT ONES. x
-%soooo CD_AC = (0+0+1+1)/(4-1) =2/3  right now.
-C needs to change to C={0,0,2,3}  then you get (0+1+1+1)/3=1.
-I suggest putting the rough calculation in like I demonstrated-because it's a little bit confusing to figure out.
 
@@
 
-%I can't find Table 2 in this document...
-
-%I strongly suggest that you labelled these figures more on the fig-put less in the legend-give them titles of Robinson Ford or triplet whatever.
+%TODO: I strongly suggest that you labelled these figures more on the fig-put less in the legend-give them titles of Robinson Ford or triplet whatever.
 Then in the first line of the fig text- you can say "Robinson Ford distance shows the degree of clade conservatism whatever that means...
@@
 The goal being to know what the figure says at a glance-the fig text can help clarify
+
 \begin{figure}[!htbp]
@@
 \hl{There is however no difference in terms of taxa displacement ($NTS_{Tr}$) between the three scenarios (with a BC > 0.98 for each comparison - Table}
-\ref{Tab_pooledscenarios_test}).  HERE  CAN YOU GIVE A LITTLE MORE INTUITION ON WHAT THE RESULTS MEAN
+\ref{Tab_pooledscenarios_test}).  %AB: TODO: HERE  CAN YOU GIVE A LITTLE MORE INTUITION ON WHAT THE RESULTS MEAN
 
-%maybe this is table 2?
+%TODO: maybe this is table 2?
 You say "differennce between the pooled scenarios"  I don't remember what you are talking about-maybe describe the pooled scenarios for the forgetful reader.
@@
 \hl{When looking at the combined effect of each parameter, the ``maximised'', ``minimised'' and ``randomised'' scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).}
-\hl{When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences  between  tree similarity of differening numbers of taxa, characters, tree correlation, or tree construction method for the taxa displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).}
+\hl{When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences  between  tree similarity of differening numbers of taxa, characters, tree correlation, or tree inference method for the taxa displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).}
 \hl{In terms of clade conservation ($NTS_{RF}$), however, ``maximised'' and ``minimised'' scenarios produced nearly always consistently different topologies using Bayesian inference (apart for the matrices with 25 taxa and 100 and 350 characters and the matrices with 150 taxa and 350 characters) but not using parsimony inference (only the matrices with 75 taxa and 100 or 350 characters produced different $NTS_{RF}$ distributions - supplementary material 3, tables 19 and 20).}
-%TODO: [discussion] This can be explained by lower resolution for Bayesian
 
@@
 \subsection{Effect of character differences on topology}
-As expected, there is a significant effect of the character difference in the ability to recover the correct topology.   This is only true for the RF method, no?
+As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
+%AB: This is only true for the RF method, no?
 The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
-Intuitively, ONE WOULD HYPOTHESIZE THAT the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (very homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
-I DONT SEE ANY DATA THAT SUPPORTS THIS STATEMENT
-\hl{HOWEVER, throughout our simulation, the ``minimised'' character difference scenario (i.e.
-increased correlation) tendED to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e.
+Intuitively, one would hypothesise that the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (very homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
+%AB: TODO: I DONT SEE ANY DATA THAT SUPPORTS THIS STATEMENT
+\hl{However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
+increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e.
 decreased correlation) led to slightly worst trees (Fig.}
 \ref{Fig:RF_results_best}.
-\hl{This (WHAT IS THIS?,  CHARACTER CORRELATION?) did not however APPEAR to affect the taxa placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical DIFFERENCES (Fig.}
+\hl{This
+%AB: TODO (WHAT IS THIS?,  CHARACTER CORRELATION?)
+did not however appear to affect the taxa placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.}
 \ref{Fig:Tr_results_best}).
 \hl{However this results needs to be contrasted by the fact that}
-the trees generated by the ``minimised'' CORRELATION scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
-\hl{Furthermore, it is important to note that on average, less characters where duplicated (replaceD) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices   WHY DID THIS HAPPEN?WHY IS IT IMPORTANT ?
-(supplementary material 3, Figs 4, 5 and 6).
-However, (THIS CLAUSE IS CONFUSING AND KIND OF SOUNDS LIKE BULLSHIT ) this change in the amount of replaced characters did not change the way average character differences VARYed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
+the trees generated by the ``minimised'' correlation scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
+\hl{Furthermore, it is important to note that on average, less characters where duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6).
+%AB: TODO:  WHY DID THIS HAPPEN?WHY IS IT IMPORTANT ? 
+However,
+%AB: TODO (THIS CLAUSE IS CONFUSING AND KIND OF SOUNDS LIKE BULLSHIT )
+this change in the amount of replaced characters did not change the way average character differences varyed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
 This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.}
-I SUDDENLY REALLY DONT LIKE THE TERM CHARACTER DIFFERENCES-ITS THE OPPOSITE OF CHARACTER CORRALTION?
+%AB: TODO: I SUDDENLY REALLY DONT LIKE THE TERM CHARACTER DIFFERENCES-ITS THE OPPOSITE OF CHARACTER CORRALTION?
 IT DOESNT SOUND LIKE IT INTUITIVELY.
@@
 Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data.
-DID YOU TALK ABOUT POOLING DATA IN THE METHOSA?
-I'M NOT SURE WHAT YOU ARE REFERING TO.
+%AB: TODO: DID YOU TALK ABOUT POOLING DATA IN THE METHOSA? I'M NOT SURE WHAT YOU ARE REFERING TO.
 Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
@@
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[i.e.
-those with in a $CI>0.25$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
+those with in a $CI>0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
@@
 
-The differences in performance of the BAYESIAN AND PARSIMONY APPROACHES observed here could be due to the inherent mechanisms of each method.
-For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scoreS differently:
-(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the (WHICH TREE?
-TRUE TREE?)tree.
+The differences in performance of the Bayesian and parsimony approaches observed here could be due to the inherent mechanisms of each method.
+For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scores differently:
+(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
+%AB: TODO: (WHICH TREE? TRUE TREE?)
 The tree search will thus likely result in a topology island that will not contain the given topology $T$.
@@
 in evo-devo \citealt{goswami2006morphological}; or in macroevolution \citealt{fitzjohn2014much}).
-These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not BE used \textit{a priori} as a criterion to select characters.
+These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not be used \textit{a priori} as a criterion to select characters.
 In other words, intra-organismal and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
@@
 This dichotomy thus creates a trade \hl{off} between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
-and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation TO BE less dependent on the coding correlations.
+and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation to be less dependent on the coding correlations.
 
@@
 However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices.% which are - when using empirical data - at best bad practice and at worst dishonest.
-I DONT  REALLY UNDERSTAND
+AB: TODO: I DONT  REALLY UNDERSTAND
 
@@
 But there are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
-However, we want to point out again that here we're noT comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
+However, we want to point out again that here we are not comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
 
@@
 Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.
-HMM IS THIS CIRCULAR LOGIC?
+%AB: TODO: HMM IS THIS CIRCULAR LOGIC?
 YOU KNOW THAT CHARACTERS ARE ON AVERAGE INFOMRATIVE-SO OF COURSE INCREASING THE REPRESENTATION OF INFROMATIVE CHARACTERS INCREASES YOUR ABILITY TO CORRELTY RSOLVE.
@@
 Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
-(DID YOU LOOK AT WHICH TOPOLOGIES WERE INDUCED??)
+%AB: TODO: (DID YOU LOOK AT WHICH TOPOLOGIES WERE INDUCED??)
 
@@
 Conversely, when correlation between characters was high, the topology was always the closest to the ``normal'' topology.
-These results APPEAR independent oF the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
+These results appear independent of the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
 
-However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise \hl{or maximised} character correlation was not more significant than simply RANDOMINSING (I.E.
+However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise \hl{or maximised} character correlation was not more significant than simply randomising (i.e.
 bootstrapping) the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
@@
 
-\]
-\]
-\]
 \end{document}

### 2019-05-31 Analysis/01-EffectCorrelationStats.Rmd
Updated figures and tables

@@
 test_table <- pair.test.table(scenario_list, metric.label = metric.labels)
+## Change the ":" into vs.
+test_table$test <- gsub(":", " vs. ", test_table$test)
 ```
@@
 ## Caption
-cap_pool_scen_test <- paste("Difference between the pooled scenarios.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction)")
+cap_pool_scen_test <- paste("Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for each minimised (mini), maximised (maxi) and randomised (rand) character differences.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).")
 
@@
 test_table <- pair.test.table(character_list, metric.label = metric.labels)
+## Change the ":" into vs.
+test_table$test <- gsub(":", " vs. ", test_table$test)
 ```
@@
 ## Caption
-cap_pool_char_test <- paste("Difference between the pooled number of characters.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+cap_pool_char_test <- paste("Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of characters: 100 (c100), 350 (c350) and 1000 (c1000).
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).")
 
@@
 test_table <- pair.test.table(taxa_list, metric.label = metric.labels)
+## Change the ":" into vs.
+test_table$test <- gsub(":", " vs. ", test_table$test)
 ```
@@
 ## Caption
-cap_pool_tax_test <- paste("Difference between the pooled number of taxa.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+cap_pool_tax_test <- paste("Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of taxa: 25 (t25), 75 (t75) and 150 (t150).
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).")
 
@@
 test_table <- pair.test.table(methods_list, metric.label = metric.labels)
+## Change the ":" into vs.
+test_table$test <- gsub(":", " vs. ", test_table$test)
 ```
@@
 ## Caption
-cap_pool_met_test <- paste("Difference between the pooled methods.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")
+cap_pool_met_test <- paste("Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the two different inference methods: bayesian or maximum parsimony.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).")
 

### 2019-05-31 Analysis/02-EffectCorrelationFullResults.Rmd
Updated figures and tables

@@
 test_table <- pair.test.table(scenario_list, comp.label = comp.labels, metric.label = metric.labels)
+## Change the ":" into vs.
+test_table$test <- gsub(":", " vs. ", test_table$test)
 ```
@@
 test_table <- pair.test.table(character_list, comp.label = comp.labels, metric.label = metric.labels)
+## Change the ":" into vs.
+test_table$test <- gsub(":", " vs. ", test_table$test)
 ```
@@
 test_table <- pair.test.table(taxa_list, comp.label = comp.labels, metric.label = metric.labels)
+## Change the ":" into vs.
+test_table$test <- gsub(":", " vs. ", test_table$test)
 ```
@@
 test_table <- pair.test.table(methods_list, comp.label = comp.labels, metric.label = metric.labels)
+## Change the ":" into vs.
+test_table$test <- gsub(":", " vs. ", test_table$test)
 ```

### 2019-05-31 Analysis/03-EffectCorrelationFigure.Rmd
Updated figures and tables

@@
 ```{r, fig.width=12, fig.height=12}
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+op <- par(mfrow=c(3,1), mar = c(2, 5, 1, 2))
 plot.results.single(t25_list, metric = 2, what = "best",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "NTS(RF) for 25 taxa",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 2, what = "best", ylab = "NTS(RF) for 75 taxa",
+plot.results.single(t75_list, metric = 2, what = "best", ylab = "Robinson-Foulds Normalised Tree Similarity\n75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 2, what = "best",
-    ylab = "NTS(RF) for 150 taxa", NTS = TRUE)
+    ylab = "150 taxa", NTS = TRUE)
 par(op)
@@
 pdf("../Writing/Figures/RF_results_best.pdf",width = 8,height = 8)
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+op <- par(mfrow=c(3,1), mar = c(2, 5, 1, 2))
 plot.results.single(t25_list, metric = 2, what = "best",
-    legend = FALSE, NTS = TRUE, xlab = "", ylab = "NTS(RF) for 25 taxa",
+    legend = FALSE, NTS = TRUE, xlab = "", ylab = "25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 2, what = "best", ylab = "NTS(RF) for 75 taxa",
+plot.results.single(t75_list, metric = 2, what = "best", ylab = "Robinson-Foulds Normalised Tree Similarity\n75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 2, what = "best",
-    ylab = "NTS(RF) for 150 taxa", NTS = TRUE)
+    ylab = "150 taxa", NTS = TRUE, legend = TRUE, leg.pos = "bottomright", leg.cex = 1.1)
 par(op)
@@
 ```{r, fig.width=12, fig.height=12}
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+op <- par(mfrow=c(3,1), mar = c(2, 5, 1, 2))
 plot.results.single(t25_list, metric = 4, what = "best",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "NTS(Tr) for 25 taxa",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "best", ylab = "NTS(Tr) for 75 taxa",
+plot.results.single(t75_list, metric = 4, what = "best", ylab = "Triplets Normalised Tree Similarity\n75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 4, what = "best",
-    ylab = "NTS(Tr) for 150 taxa", NTS = TRUE)
+    ylab = "150 taxa", NTS = TRUE, legend = TRUE, leg.pos = "bottomright", leg.cex = 1.1)
 par(op)
@@
 pdf("../Writing/Figures/Tr_results_best.pdf", width = 8, height = 8)
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+op <- par(mfrow=c(3,1), mar = c(2, 5, 1, 2))
 plot.results.single(t25_list, metric = 4, what = "best",
-    legend = FALSE, NTS = TRUE, xlab = "", ylab = "NTS(Tr) for 25 taxa",
+    legend = FALSE, NTS = TRUE, xlab = "", ylab = "25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "best", ylab = "NTS(Tr) for 75 taxa",
+plot.results.single(t75_list, metric = 4, what = "best", ylab = "Triplets Normalised Tree Similarity\n75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 4, what = "best",
-    ylab = "NTS(Tr) for 150 taxa", NTS = TRUE)
+    ylab = "150 taxa", NTS = TRUE, legend = TRUE, leg.pos = "bottomright", leg.cex = 1.1)
 par(op)
@@
 ```{r, fig.width=12, fig.height=12}
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+op <- par(mfrow=c(3,1), mar = c(2, 5, 1, 3))
 plot.results.single(t25_list, metric = 2, what = "null",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "NTS(RF) for 25 taxa",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 2, what = "null", ylab = "NTS(RF) for 75 taxa",
+plot.results.single(t75_list, metric = 2, what = "null", ylab = "Robinson-Foulds Normalised Tree Similarity\n75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 2, what = "null",
-    ylab = "NTS(RF) for 150 taxa", NTS = TRUE)
+    ylab = "150 taxa", NTS = TRUE)
 par(op)
@@
 pdf("../Writing/Figures/RF_results_null.pdf", width = 8, height = 8)
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+op <- par(mfrow=c(3,1), mar = c(2, 5, 1, 2))
 plot.results.single(t25_list, metric = 2, what = "null",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "NTS(RF) for 25 taxa",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 2, what = "null", ylab = "NTS(RF) for 75 taxa",
+plot.results.single(t75_list, metric = 2, what = "null", ylab = "Robinson-Foulds Normalised Tree Similarity\n75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 2, what = "null",
-    ylab = "NTS(RF) for 150 taxa", NTS = TRUE)
+    ylab = "150 taxa", NTS = TRUE)
 par(op)
@@
 ```{r, fig.width=12, fig.height=12}
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+op <- par(mfrow=c(3,1), mar = c(2, 5, 1, 2))
 plot.results.single(t25_list, metric = 4, what = "null",
-    legend = TRUE, NTS = TRUE, xlab = "", ylab = "NTS(Tr) for 25 taxa",
+    legend = TRUE, NTS = TRUE, xlab = "", ylab = "25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "null", ylab = "NTS(Tr) for 75 taxa",
+plot.results.single(t75_list, metric = 4, what = "null", ylab = "Triplets Normalised Tree Similarity\n75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 4, what = "null",
-    ylab = "NTS(Tr) for 150 taxa", NTS = TRUE)
+    ylab = "150 taxa", NTS = TRUE)
 par(op)
@@
 pdf("../Writing/Figures/Tr_results_null.pdf", width = 8, height = 8)
-op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
+op <- par(mfrow=c(3,1), mar = c(2, 5, 1, 2))
 plot.results.single(t25_list, metric = 4, what = "null",
-    legend = FALSE, NTS = TRUE, xlab = "", ylab = "NTS(Tr) for 25 taxa",
+    legend = FALSE, NTS = TRUE, xlab = "", ylab = "25 taxa",
     axislab = FALSE)
-plot.results.single(t75_list, metric = 4, what = "null", ylab = "NTS(Tr) for 75 taxa",
+plot.results.single(t75_list, metric = 4, what = "null", ylab = "Triplets Normalised Tree Similarity\n75 taxa",
     xlab = "", axislab = FALSE, NTS = TRUE)
 plot.results.single(t150_list, metric = 4, what = "null",
-    ylab = "NTS(Tr) for 150 taxa", NTS = TRUE)
+    ylab = "150 taxa", NTS = TRUE)
 par(op)

### 2019-05-31 Functions/plot.results.R
Added legend controls

@@ -11,6 +11,8 @@
 #' @param ylab a \code{character} string to be passed to \code{boxplot(..., ylab)}
 #' @param xlab a \code{character} string to be passed to \code{boxplot(..., xlab)}
 #' @param axislab a \code{character} string to be passed to \code{axis(..., labels)}
+#' @param leg.pos the legend position (default is \code{"bottomleft"})
+#' @param leg.cex the legend text size (default is \code{1})
 #' @param ... any additional argument to be passed to \code{boxplot}
 #'
 #' @details
@@ -30,7 +32,7 @@
 #' 
 
 
-plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, NTS = TRUE, ylab, xlab, axislab, ...) {
+plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, NTS = TRUE, ylab, xlab, axislab, leg.pos = "bottomleft", leg.cex = 0.8, ...) {
     
     ## Graphic parameters
     ## Colours
@@ -63,7 +65,7 @@ plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, N
 
     ## axis lab
     if(missing(axislab)) {
-        axislab <- c("100", "350", "1000")
+        axislab <- c("100", "350\ncharacters", "1000")
     }
 
     ## Horizontal line
@@ -94,7 +96,7 @@ plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, N
         abline(h = line, col = "grey", lty = 2)
         ## Legend
         if(legend) {
-            legend("bottomleft", col = cols, pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
+            legend(leg.pos, cex = leg.cex, col = cols, pch = 19, bty = "n", title = "Character differences:", bg = "white", border = "white",
                 legend = c("Maximised (Bayesian)", "Maximised (Parsimony)",
                            "Minimised (Bayesian)", "Minimised (Parsimony)",
                            "Normal (Bayesian)", "Normal (Parsimony)")
@@ -126,7 +128,7 @@ plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, N
         abline(h = line, col = "grey", lty = 2)
         ## Legend
         if(legend) {
-            legend("bottomleft", col = cols, pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
[…]

### 2019-05-31 Writing/draft_highlighted.tex
Updated manuscript following Abigail's comments

@@
 
-%AP Well this is worrying-because it seems like authors could use these results to justify introducing or ignoring character correlations.
-It seems like it would be important to test the effect of correlations in uninformative vs informative characters.
-Reconstructors wouldn't know whether each character is informative or not, so they would more likely try to avoid it?
-But yeah, maybe that would take another year?
 \section{Methods}
@@
 \textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
-For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,0,1,2\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$ \hl{$CD_{(A,B)} = (0+0+0+0)/(4-1) = 0$ and $CD_{(A,C)} = (0+1+1+1)/(4-1) = 1$}.
+For example with three characters \texttt{A = \{0,1,2,1\}}, \texttt{B = \{2,0,1,0\}} and \texttt{C = \{0,0,1,2\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$ \hl{(translated into \texttt{A' = \{1,2,3,2\}} \texttt{B' = \{1,2,3,2\}} and \texttt{C' = \{1,1,2,3\}}; $CD_{(A,B)} = (0+0+0+0)/(4-1) = 0$ and $CD_{(A,C)} = (0+1+1+1)/(4-1) = 1$)}.
 Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different set of splits (i.e. a different phylogenetic signal).
@@
 
-%TODO: I strongly suggest that you labelled these figures more on the fig-put less in the legend-give them titles of Robinson Ford or triplet whatever.
-Then in the first line of the fig text- you can say "Robinson Ford distance shows the degree of clade conservatism whatever that means...
-"  so the reader can immediately know the difference between fig 2 and 3.
-To label: (1) text on ylab "Normalized tree similarity"- then each row can just be "# taxa" (2) mtext on xlab "Number of characters"  (3)  legend c("more correlated", "less correlated", "randomised") but leave bayesian and parsimony for the fig text.
-The goal being to know what the figure says at a glance-the fig text can help clarify
-
 \begin{figure}[!htbp]
@@
    \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
-\caption{\small{Effect of character difference on recovering the ``normal'' topology.
+\caption{\small{Effect of character difference on recovering the ``normal'' topology in terms of clade conservation ($NTS_{RF}$).
 The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
@@
    \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf}
-\caption{Effect of character difference on recovering the ``normal'' topology.
+\caption{Effect of character difference on recovering the ``normal'' topology in terms of Triplets ($NTS_{Tr}$).
 The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
@@
 The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.869 and median $NTS_{Tr}$ = 0.787\hl{; see supplementary materials 3, Table 3}) followed by randomising the characters (median $NTS_{RF}$ = 0.814 and median $NTS_{Tr}$ = 0.791\hl{; see supplementary materials 3, Table 2}) and maximising the character difference (median $NTS_{RF}$ = 0.698 and median $NTS_{Tr}$ = 0.754\hl{; see supplementary materials 3, Table 2}).
-There is a significant difference between all scenarios \hl{in terms of clade conservation ($NTS_{RF}$)} (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
+There is a significant difference between all scenarios \hl{in terms of clade conservation ($NTS_{RF}$)} (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability of overlap between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
 \hl{There is however no difference in terms of taxa displacement ($NTS_{Tr}$) between the three scenarios (with a BC > 0.98 for each comparison - Table}
-\ref{Tab_pooledscenarios_test}).  %AB: TODO: HERE  CAN YOU GIVE A LITTLE MORE INTUITION ON WHAT THE RESULTS MEAN
-
-%TODO: maybe this is table 2?
-You say "differennce between the pooled scenarios"  I don't remember what you are talking about-maybe describe the pooled scenarios for the forgetful reader.
-I have no idea what table 2 is even saying-the table text could use more hand holding explanations.
-# same issues with table 3.
-I think I personally would've done these stats differently as a multilevel model so that the bonferroni corrections weren't necessary-but I suppose it doesnt fucking matter with p values like that.
-However, table 4 is perfect: don't change a thing.
+\ref{Tab_pooledscenarios_test}).
 
@@
 \subsection{Effect of character differences on topology}
-As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
-%AB: This is only true for the RF method, no?
+As expected, there is a significant effect of the character difference in the ability to recover the correct topology \hl{(in terms of clades)}.
 The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
-Intuitively, one would hypothesise that the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (very homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
-%AB: TODO: I DONT SEE ANY DATA THAT SUPPORTS THIS STATEMENT
+Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (\hl{i.e.
+because the characters are not different, they all support the same topology that might be incorrect}), the ``maximised'' scenario to lead to poorly resolved trees (\hl{i.e.
+because characters are different, the matrix has a lot of homoplasy leading to badly resolved topologies}) and the ``randomised'' scenario to perform the best at recovering the correct topology (\hl{i.e.
+a simple bootstrap procedure}).
 \hl{However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
@@
 \ref{Fig:RF_results_best}.
-\hl{This
-%AB: TODO (WHAT IS THIS?,  CHARACTER CORRELATION?)
-did not however appear to affect the taxa placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.}
+\hl{This effect of increasing or reducing character difference did not however appear to affect the taxa placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.}
 \ref{Fig:Tr_results_best}).
 \hl{However this results needs to be contrasted by the fact that}
-the trees generated by the ``minimised'' correlation scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
-\hl{Furthermore, it is important to note that on average, less characters where duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6).
-%AB: TODO:  WHY DID THIS HAPPEN?WHY IS IT IMPORTANT ? 
-However,
-%AB: TODO (THIS CLAUSE IS CONFUSING AND KIND OF SOUNDS LIKE BULLSHIT )
-this change in the amount of replaced characters did not change the way average character differences varyed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
+the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
+\hl{Furthermore, it is important to note that on average, less characters where duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
 This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.}
-%AB: TODO: I SUDDENLY REALLY DONT LIKE THE TERM CHARACTER DIFFERENCES-ITS THE OPPOSITE OF CHARACTER CORRALTION?
-IT DOESNT SOUND LIKE IT INTUITIVELY.
-ITS KIND OF CONFUSING TO HAVE SO MANY TERMS MEANING THE SAME THING -MAXIMIZED V MINIMIZED, CHARACTER CORRALATION, CHARACTER DIFFERENCE.
-I'M TIRED AND GRUMPY.
 
@@
 This suggests that the overall character difference metric \hl{can be used as a proxy} for character correlation as it is \hl{not directly dependent} on the number of characters analysed.
-Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data.
-%AB: TODO: DID YOU TALK ABOUT POOLING DATA IN THE METHOSA? I'M NOT SURE WHAT YOU ARE REFERING TO.
+Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data (\hl{i.e.
+when combining the simulations together}).
 Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
@@
 
-%TODO: Bayesian recovers the ``true'' tree topology slighltly but consistently better (supplementary materials 3, figures 12, 13, and 14).
-
 The differences in performance of the Bayesian and parsimony approaches observed here could be due to the inherent mechanisms of each method.
 For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scores differently:
-(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
-%AB: TODO: (WHICH TREE? TRUE TREE?)
+(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the focal topology.
 The tree search will thus likely result in a topology island that will not contain the given topology $T$.
@@
 creating ``modules'' of characters).
-However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices.% which are - when using empirical data - at best bad practice and at worst dishonest.
-AB: TODO: I DONT  REALLY UNDERSTAND
+However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices \hl{(i.e.
+inferring the topology first and looking at character's correlation based on the observed topology}\citealt{Lande1983,Maddison1990,Pagel1994}).
 
@@
 In matrices with more taxa, however, the \hl{``minimised''} scenario resulted in better topological recovery than any other scenarios.
-Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.
-%AB: TODO: HMM IS THIS CIRCULAR LOGIC?
-YOU KNOW THAT CHARACTERS ARE ON AVERAGE INFOMRATIVE-SO OF COURSE INCREASING THE REPRESENTATION OF INFROMATIVE CHARACTERS INCREASES YOUR ABILITY TO CORRELTY RSOLVE.
-IF CHRACTERS WERE UNIFORMATIVE AND CORRELATED THIS RESULT WOULD NOT HOLD
+Applying this kind of bootstraps that minimised character difference by biasing the random sampling could thus \hl{help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.}
 
@@
 Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
-%AB: TODO: (DID YOU LOOK AT WHICH TOPOLOGIES WERE INDUCED??)
 

### 2019-05-31 Writing/Tables/Tab_pooledscenarios_test.tex
Updated tables

--- 2019-04-16 Writing/Tables/Tab_pooledscenarios_test.tex
@@
-% latex table generated in R 3.5.3 by xtable 1.8-3 package
-% Tue Apr 16 16:07:23 2019
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Fri May 31 14:42:24 2019
 \begin{table}[ht]
@@
   \hline
-RF & maxi:mini & 0.848 & 32012.500 & 0.000 \\ 
-   & maxi:rand & 0.941 & 44161.000 & 0.000 \\ 
-   & mini:rand & 0.954 & 80382.500 & 0.000 \\ 
-  Tr & maxi:mini & 0.992 & 60680.500 & 0.839 \\ 
-   & maxi:rand & 0.980 & 60447.000 & 0.713 \\ 
-   & mini:rand & 0.982 & 64222.500 & 1.000 \\ 
+RF & maxi vs. mini & 0.848 & 32012.500 & \textbf{0.000} \\ 
+   & maxi vs. rand & 0.941 & 44161.000 & \textbf{0.000} \\ 
+   & mini vs. rand & \textbf{0.954} & 80382.500 & \textbf{0.000} \\ 
+  Tr & maxi vs. mini & \textbf{0.992} & 60680.500 & 0.839 \\ 
+   & maxi vs. rand & \textbf{0.980} & 60447.000 & 0.713 \\ 
+   & mini vs. rand & \textbf{0.982} & 64222.500 & 1.000 \\ 
    \hline
 \end{tabular}
-\caption{Difference between the pooled scenarios.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction)}
+\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for each minimised (mini), maximised (maxi) and randomised (rand) character differences.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).}
 \label{Tab_pooledscenarios_test}

### 2019-06-21 Writing/draft_highlighted.tex
highlighted draft done!

@@
 %Running head
-\begin{flushright}
+\begin{flushright}s
 Version dated: \today
@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, Abigail I. Pastore$^{1}$, and Martin D. Brazeau$^{2,3}$}\\
 % Abigail I. Pastore$^{1}$
@@
 Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with \hl{careful character coding}, phylogenetic software do not distinguish between any of them.
-
 In this study, we propose a new metric of raw character difference as a proxy for character correlation.
@@
 Furthermore, in bigger matrices (350 characters), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
-
 These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
@@
 \hl{However, several distinctions can be made between the amount and the nature of the correlation.
-First, molecular and morphological data often differs in size and in the amount of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but sometimes more than ten;}
+First, molecular and morphological data often differs in size and in the number of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but may have an arbitrary number of states (up to that supported by phylogenetic software;}
 \citealt{Guillerme2016146}
 \hl{).
-Second, characters has been suggested to have stronger correlations in morphological data (i.e.
-``concentrated'' correlations) \textit{versus} weaker and more ``spread'' correlations in molecular data}
-\citep[\hl{i.e. ``diffused'' correlations - though some regions of the genome can be thought as more correlated than others}][]{huelsenbeck1999effect}.
-\hl{Furthermore, because of the ``spread'' correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
+Second, correlations have been proposed to be stronger in morphological data (i.e. ``concentrated'' correlations) \textit{versus} weaker and more ``spread-out'' correlations in molecular data}
+\citep[\hl{i.e. ``diffused'' correlations - though some regions of the genome may be thought as more correlated than others}][]{huelsenbeck1999effect}.
+\hl{Furthermore, because of the ``spread'' of correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
 changes in morphological states can be more readily linked to a function \textit{cf.} changes in nucleotide states ; see below).}
-Before discussing character correlation further, it is important to understand that correlation may manifest itself in at least three distinct ways:
+Before discussing character correlation further,
+\hl{we distinguish three different modes of forms of character correlation:}
 
@@
     Characters of the occlusal surface of two opposing molars will be expected to directly covary.
-    Pleiotropy also results in covariation between different aspects of phenotype. 
+    Pleiotropy also results in covariation between different aspects of phenotype \hl{and may thus induce correlations}. 
     From a phylogenetic perspective, it can be especially pernicious because the relationship between the traits in question may have no obvious link from a morphological or functional comparison alone.
@@
 Even after all of the effort is expended, the results might then only be known for a single (model) species.
-Evolutionary dependence itself requires the resolution of a phylogenetic tree, and is best determined by independent character sets. 
+Evolutionary dependence itself requires the \hl{accurate} resolution of a phylogenetic tree, and is best determined by independent character sets. 
 This is frequently accomplished by mapping morphological traits on molecular phylogenetic trees.
@@
 
-\noindent \textbf{Correlation between characters detected by the software}: this is the result of how software actually interprets the differences between characters.
-The vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different or similar tokens).
-Therefore a great number of characters and - traditionally - a few number of tokens can easily lead to dependence between characters.
+\noindent \textbf{Characters compatibility}: this is the result of how software actually interprets the differences between characters.
+\hl{The majority of phylogenetic algorithms cannot really interpret character descriptions and state meanings (i.e.
+character ontologies), and instead must interpret discrete tokens as signifiers of shared similarity.}
+Therefore, \hl{phylogenetic character matrices that contain a great number of characters but} a few number of tokens can easily lead to dependence between characters.
 For example, if we consider the following matrix containing four cetartiodactyls - say a pig (e.g.
@@
 In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
-Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states' tokens will be used in order to infer the tree.
-\hl{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
+Yet a phylogenetic software will treat all these four characters in exactly the same way.
+\hl{In fact, for practical reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
 In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} that seem to be more likely to reflect the split by separating artiodactyles from the cetaceans.
@@
 However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.}
+\hl{It is worth noting that partitioning strategies have been used to reduce this problem}
+\citep[see][and references within]{Rosa2019}
+\hl{.}
 This illustrates how some characters will therefore be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
-Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).
 
 How \hl{do} these \hl{specific} correlation really affect topology?
+Although it has been tackled empirically (for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}), the question has never been explored through a thorough simulation framework.
 % In the specific case of discrete morphological characters, how does the reduc phenotypic information and its different sources of correlation is 
@@
 
-    \item \textbf{Modifying matrices}: we changed the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least different or most different characters and replacing them randomly by the remaining characters.
+    \item \textbf{Modifying matrices}: we changed the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least different or most different characters and replacing them randomly by \hl{characters from the remaining set}.
 Our protocol for measuring character difference is detailed below.
@@
 In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
-\hl{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in }
+\hl{This procedure allows comparison of characters regardless of the significance given to their tokens (following the \textit{xyz} notation in }
 \citealt{felsenstein2004inferring}; \hl{as used in} \citealt{Davalos01072014}).
@@
 
-\noindent For each character, both models (``morphological HKY-binary'' or M$k$) where chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
+\noindent For each character, both models (``morphological HKY-binary'' or M$k$) were chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
 These low evolutionary rate values allowed reduction in the number of homoplasic character changes, thus reinforcing the phylogenetic information in the matrices.
@@
 We inferred the topologies with both Bayesian inference and maximum parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
-For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees (\hl{the first taxa in the taxa list}).
+For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees (\hl{the first taxon in the taxa list}).
 The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
@@
 \caption{\small{Effect of character difference on recovering the ``normal'' topology in terms of clade conservation ($NTS_{RF}$).
-The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
+The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 f from top to bottom respectively.
 The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
@@
 There is a significant difference between all scenarios \hl{in terms of clade conservation ($NTS_{RF}$)} (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability of overlap between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
-\hl{There is however no difference in terms of taxa displacement ($NTS_{Tr}$) between the three scenarios (with a BC > 0.98 for each comparison - Table}
+\hl{There is however no difference in terms of taxon displacement ($NTS_{Tr}$) between the three scenarios (with a BC $>$ 0.98 for each comparison - Table}
 \ref{Tab_pooledscenarios_test}).
@@
 \subsubsection{Number of characters}
-\hl{This effect of the character difference affects clade conservation ($NTS_{RF}$) only by big differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ - Table}
+\hl{This effect of the character difference affects clade conservation ($NTS_{RF}$) only over large differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ - Table}
 \ref{Tab_pooledscharacters_test}).
@@
 100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests showed clear differences ($p = 0$) but with high distribution overlap ($BC > 0.96$).}
-\hl{Regarding taxa displacements ($NTS_{Tr}$), there was an effect of the number of characters (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - Table}
+\hl{Regarding taxon displacements ($NTS_{Tr}$), there was an effect of the number of characters (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - Table}
 \ref{Tab_pooledscharacters_test}).
@@
 \hl{When looking at the combined effect of each parameter, the ``maximised'', ``minimised'' and ``randomised'' scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).}
-\hl{When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences  between  tree similarity of differening numbers of taxa, characters, tree correlation, or tree inference method for the taxa displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).}
-\hl{In terms of clade conservation ($NTS_{RF}$), however, ``maximised'' and ``minimised'' scenarios produced nearly always consistently different topologies using Bayesian inference (apart for the matrices with 25 taxa and 100 and 350 characters and the matrices with 150 taxa and 350 characters) but not using parsimony inference (only the matrices with 75 taxa and 100 or 350 characters produced different $NTS_{RF}$ distributions - supplementary material 3, tables 19 and 20).}
+\hl{When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences  between  tree similarity of differening numbers of taxa, characters, tree correlation, or tree inference method for the taxon displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).}
+\hl{In terms of clade conservation ($NTS_{RF}$), however, ``maximised'' and ``minimised'' scenarios produced nearly always consistently different topologies using Bayesian inference (apart for the matrices with 25 taxa and 100 and 350 characters and the matrices with 150 taxa and 350 characters) but not using parsimony (only the matrices with 75 taxa and 100 or 350 characters produced different $NTS_{RF}$ distributions - supplementary material 3, tables 19 and 20).}
 
@@
 \subsection{Effect of character differences on topology}
-As expected, there is a significant effect of the character difference in the ability to recover the correct topology \hl{(in terms of clades)}.
+As expected, there is a significant effect of character difference in the ability to recover the correct topology \hl{(in terms of clades)}.
 The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
@@
 \ref{Fig:RF_results_best}.
-\hl{This effect of increasing or reducing character difference did not however appear to affect the taxa placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.}
+\hl{This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.}
 \ref{Fig:Tr_results_best}).
-\hl{However this results needs to be contrasted by the fact that}
+\hl{However this results needs to be contrasted with the fact that}
 the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
-\hl{Furthermore, it is important to note that on average, less characters where duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+\hl{Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
 However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
@@
 Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance).
-\hl{Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a bigger effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
+\hl{Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
 If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
@@
 coding correlation) the only type of correlation that can affect the phylogeny \textit{a priori}.
-This dichotomy thus creates a trade \hl{off} between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
+This dichotomy thus creates a trade-off between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
 and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation to be less dependent on the coding correlations.
@@
 This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.
-\hl{It is also worth noting that character similarity at the tips does not necessarily reflects that the characters underwent the same evolutionary history.
+\hl{It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
 This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches}
 \citep[e.g.][]{revell2014ancestral}.
-\hl{However, it is worth noting that in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).}
+\hl{However, in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).}
 
@@
 Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232) and the Australian QRIScloud Awoonga cluster services.
-We thank Alberto Pascual Garcia for input in the design of the simulations protocol, Guillermo Herraiz Yebes and Abigail Pastore for helping with the CD metric proof and Orlin Todorov for assistance with the computations.
+We thank Alberto Pascual Garcia for input in the design of the simulations protocol, Guillermo Herraiz Yebes for helping with the CD metric proof and Orlin Todorov for assistance with the computations.
 We also thank the \hl{three anonymous reviewers for there constructive and insightful comments}.

### 2019-06-21 Writing/response_to_reviewers.tex
Updated line numbering in response

@@
 \textit{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
-Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).} lines 213-218.
+Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).} lines 212-2178.
 
@@
 
-\textit{Furthermore, it is important to note that on average, less characters where duplicated (replace) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6).
+\textit{Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6).
 However, this change in the amount of replaced characters did not changed the way average character differences changed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
-This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.} lines 453-462.
+This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.} lines 459-469.
 
@@
 However, several distinctions can be made between the amount and the nature of the correlation.
-First, molecular and morphological data often differs in size and in the amount of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but sometimes up to more than ten; Guillerme and Cooper, 2016).
-Second, this difference between both types of characters has been suggested to be stronger correlations in morphological data (i.e.
-``concentrated'' correlations) \textit{versus} weaker and more ``spread'' correlations in molecular data (i.e.
-``diffused'' correlations - though some regions of the genome can be thought as more correlated than others; Huelsenbeck and Nielsen, 1999).
-Furthermore, because of this ``spread'' correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
+First, molecular and morphological data often differs in size and in the number of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but may have an arbitrary number of states (up to that supported by phylogenetic software; Guillerme and Cooper, 2016).
+Second, correlations have been proposed to be stronger in morphological data (i.e.
+``concentrated'' correlations) \textit{versus} weaker and more ``spread-out'' correlations in molecular data (i.e.
+``diffused'' correlations - though some regions of the genome may be thought as more correlated than others; Huelsenbeck and Nielsen, 1999).
+Furthermore, because of this ``spread'' of correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
 changes in morphological states can be more readily linked to a function cf.
-changes in nucleotide states ; see below).} lines 43-58.
+changes in nucleotide states ; see below).} lines 42-57.
 
@@
 
-We added the suggested citation (lines 75).
+We added the suggested citation (line 74).
 
@@
 
-\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.} lines 95-96 and \textit{In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.} lines 102-104.
+\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.} lines 94-95 and \textit{In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.} lines 101-103.
 
@@
 
-\textit{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
+\textit{In fact, for practical reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
 In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} that seem to be more likely to reflect the split by separating artiodactyles from the cetaceans.
@@
 making \textbf{C3} and \textbf{C4} dependent on \textbf{C1}).
-However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.} lines 145-157.
+However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.} lines 143-156.
 
@@
 
-\textit{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in Felsenstein, 2004; as used in D\'{a}valos et al., 2014).} lines 221-223.
+\textit{This procedure allows comparison of characters regardless of the significance given to their tokens (following the \textit{xyz} notation in Felsenstein, 2004; as used in D\'{a}valos et al., 2014).} lines 220-222.
 
@@
 \textit{Note that both models (``morphological HKY'' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
-However, the ``morphological HKY model'' is not the one} directly \textit{estimated in the phylogenetic software used in this protocol (MrBayes, see below).} lines 271-275.
+However, the ``morphological HKY model'' is not the one} directly \textit{estimated in the phylogenetic software used in this protocol (MrBayes, see below).} lines 272-276.
 
@@
 
-We've changed the sentence as suggest by the reviewer (lines 260-261).
+We've changed the sentence as suggest by the reviewer (lines 262-263).
 
@@
 
-\textit{However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.} lines 346-347
+\textit{However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.} lines 347-3478.
 
@@
 
-\textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines 202-203.
+\textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines 201-202.
 
@@
 
-\textit{Each replaced characters (i.e. the characters with a weight of +$1$) were randomly selected by sampling them randomly from the list of the remaining characters.} lines 297-299.
+\textit{Each replaced characters (i.e. the characters with a weight of +$1$) were randomly selected by sampling them randomly from the list of the remaining characters.} lines 298-300.
 
@@
 
-\textit{We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters (from empirical observations in: Guillerme and Cooper, 2016b; Zou and Zhang, 2016).} lines 256-258.
+\textit{We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters (from empirical observations in: Guillerme and Cooper, 2016b; Zou and Zhang, 2016).} lines 257-259.
 
@@
 
-We have now clarified in the manuscript what our objectives were by streamlining the last paragraph of the introduction (lines 160-168).
+We have now clarified in the manuscript what our objectives were by streamlining the last paragraph of the introduction (lines 159-170).
 We have also explicitly added in the supplementary material 3, figs 1, 2 and 3 how the character difference changes between all matrices compared to the ``normal'' matrix.
@@
 We've added this remark in the caveats section:
-\textit{It is also worth noting that character similarity at the tips does not necessarily reflects that the characters underwent the same evolutionary history.
+\textit{It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
 This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches (e.g.
 Revell, 2014).
-However, it is worth noting that in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).} lines 554-560.
+However, it is worth noting that in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).} lines 563-568.
 
@@
 
-\textit{Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a bigger effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
+\textit{Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
 If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
-Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.} lines 468-474
+Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.} lines 475-481.
 
@@
 We've added the following justification (and toned down the sentence):
-\textit{Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collected them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.} lines 29-31.
+\textit{Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collected them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.} lines 28-30.
 

### 2019-06-21 Writing/draft.tex
Finalised version done up to the discussion

@@
 \begin{document}
-
 %Running head
-\begin{flushright}
+\begin{flushright}s
 Version dated: \today
 \end{flushright}
+
 \bigskip
 \noindent RH: Characters correlation
-
 \bigskip
@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, Abigail I. Pastore$^{1}$, and Martin D. Brazeau$^{2,3}$}\\
+% Abigail I. Pastore$^{1}$
 \noindent {\small \it 
@@
 
+% CITE: https://academic.oup.com/sysbio/article/68/2/267/5145070#130574594
+
+% Mention Hamming distance rather than Gower?
+
 \newpage
 \begin{abstract}
-Phylogenetic analysis algorithms require the assumption of character independence - a condition generally acknowledged to be violated by morphological data.
+Phylogenetic analysis algorithms assume character independence - a condition generally acknowledged to be violated by morphological data.
 Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
-Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with good practices, phylogenetic software do not distinguish between any of them.
-
+Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
 In this study, we propose a new metric of raw character difference as a proxy for character correlation.
 Using thorough simulations, we test the effect of increasing or decreasing character differences on tree topology.
-Overall, we found an expected positive effect of reducing character correlations on recovering the correct topology.
-However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing character correlation is not more effective than randomly drawing characters.
+Overall, we found a negative effect of reducing character correlations on recovering the correct topology.
+However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing or increasing character correlation is not more effective in getting the expected topology than randomly drawing characters.
 Furthermore, in bigger matrices (350 characters), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
-
 These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
@@
 This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
-Morphological character data are often considered inferior to molecular sequence data, but are often the only source of phylogenetic data for extinct species. 
+Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collecting them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.
 While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
@@
 
-The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data.
+The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data \cite[e.g.][]{Davalos01072014, ZouConvergence}.
 The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the very nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are \hl{dependent} on each other.
-Before discussing character correlation further, it is important to understand that it may manifest itself in at least three distinct ways:
+This independence is assumed for both molecular and morphological characters \citep{huelsenbeck1999effect,Davalos01072014,ZouConvergence}.
+However, several distinctions can be made between the amount and the nature of the correlation.
+First, molecular and morphological data often differs in size and in the number of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but may have an arbitrary number of states (up to that supported by phylogenetic software; \citealt{Guillerme2016146}).
+Second, correlations have been proposed to be stronger in morphological data (i.e.
+``concentrated'' correlations) \textit{versus} weaker and more ``spread-out'' correlations in molecular data \citep[i.e.
+``diffused'' correlations - though some regions of the genome may be thought as more correlated than others][]{huelsenbeck1999effect}.
+Furthermore, because of the ``spread'' of correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
+changes in morphological states can be more readily linked to a function \textit{cf.} changes in nucleotide states ; see below).
+Before discussing character correlation further, we distinguish three different modes of forms of character correlation:
 
@@
     Characters of the occlusal surface of two opposing molars will be expected to directly covary.
-    Pleiotropy also results in covariation between different aspects of phenotype. 
+    Pleiotropy also results in covariation between different aspects of phenotype and may thus induce correlations. 
     From a phylogenetic perspective, it can be especially pernicious because the relationship between the traits in question may have no obvious link from a morphological or functional comparison alone.
@@
     This occurs when a number of multistate characters include two variable feature states (e.g. large, small; red, blue etc.) in conjunction with absence.
+    For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.
     It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
@@
     This coding dependency is linked to hierarchical dependency between characters \citep{wilkinson1995coping,BrazeauNA}.
+    In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.
     Finally this can also be due to a bias in the amount of characters available.
@@
 Logical dependence, however, is easily distinguished prior to phylogenetic inference, while the two other ones (intra-organismal and evolutionary) are much harder.
-However, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{DeLaet2015,BrazeauNA}.
+However, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{de2015parsimony,BrazeauNA}.
 Intra-organismal dependence requires more detailed, often extremely time-consuming studies (and possibly beyond the limits of available technology).
 Even after all of the effort is expended, the results might then only be known for a single (model) species.
-Evolutionary dependence itself requires the resolution of a phylogenetic tree, and is best determined by independent character sets. 
+Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree, and is best determined by independent character sets. 
 This is frequently accomplished by mapping morphological traits on molecular phylogenetic trees.
@@
 
-\noindent \textbf{Correlation between characters detected by the software}: this is the result how software actually interprets the differences between characters.
-The vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different or similar tokens).
-Therefore a great number of characters and - traditionally - a few number of tokens can easily lead to dependence between characters.
+\noindent \textbf{Characters compatibility}: this is the result of how software actually interprets the differences between characters.
+The majority of phylogenetic algorithms cannot really interpret character descriptions and state meanings (i.e.
+character ontologies), and instead must interpret discrete tokens as signifiers of shared similarity.
+Therefore, phylogenetic character matrices that contain a great number of characters but a few number of tokens can easily lead to dependence between characters.
 For example, if we consider the following matrix containing four cetartiodactyls - say a pig (e.g.
-\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley:
+\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
 
@@
 In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
-Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states tokens will be used in order to infer the tree.
-Some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
-Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).
+Yet a phylogenetic software will treat all these four characters in exactly the same way.
+In fact, for practical reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
+In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} that seem to be more likely to reflect the split by separating artiodactyles from the cetaceans.
+In other words, this is the correlation between characters that depicts the evolutionary history of this group.
+However, the other characters depict the same history (i.e.
+the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleen, the same split would still be estimated by the software).
+One solution would be to a) weight some characters (e.g.
+giving \textbf{C3} and \textbf{C4} a lower weight) and b) make some characters dependent on each other (e.g.
+making \textbf{C3} and \textbf{C4} dependent on \textbf{C1}).
+However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.
+Alternatively, partitioning strategies have been used to reduce this problem \citep[see][and references within]{Rosa2019}.
+This illustrates how some characters will therefore be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
 
-How does these correlation really affect topology?
-We expect matrices with a high level of correlation to recover precise but inaccurate topologies but will matrices with low level of correlation (i.e.
-with high levels of homoplasy) actual cancel out the effects of correlation?
+How do these specific correlation really affect topology?
+Although it has been tackled empirically (for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}), the question has never been explored through a thorough simulation framework.
+% In the specific case of discrete morphological characters, how does the reduc phenotypic information and its different sources of correlation is 
+% More specifically, how can phylogenetic inference software distinguish between the complex sources of correlation tha
+% In other words, how can we distinguish between the complex sources of correlation in phenotype using only discrete morphological characters?
+% \cite{parins2017use}
 Here we formally assess the effect of discrete character's correlation using simulated data.
-We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences or similarities between characters.
-We found that overall, there is a detectable effect of character correlation on topology where an increase in character dependence results in a decrease in the ability to recover the correct topology.
+We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
+We found that overall, there is a detectable effect of character correlation on topology where a decrease in character dependence results in a decrease in the ability to recover the correct topology.
 These results, however, vary greatly in magnitude depending on the size of matrix and the inference method used.
@@
 
-    \item \textbf{Modifying matrices}: we changed the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least different or most different characters and replacing them randomly by the remaining characters.
+    \item \textbf{Modifying matrices}: we changed the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least different or most different characters and replacing them randomly by characters from the remaining set.
 Our protocol for measuring character difference is detailed below.
@@
 This step resulted in 72 topologies.
-    \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on topology.
+    \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on the recovered topology.
 \end{enumerate}
-Each step was replicated 35 times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
+All steps were replicated 20 times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
 
@@
 \subsection{Measuring differences between characters}
-\label{CD_description}
-To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information.
+\label{CDdescription}
+To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information (i.e.
+entirely correlated characters would define identical splits between a set of taxa).
 In order to measure this, we propose a new distance metric to measure the difference between two characters:
-%TG: OK, I've checked Farris' stuff.
-Despite not beeing described (at least not the in 1970 Syst Biol paper), it seems that the advancement index is just the distance between two taxa.
-Appart from the taxa/character difference, it's kind of similar but (1) it's not ponderated (divided by n), centered (-1/2) and reversed (1-2(...)); so I'm not sure if we need to say it's similar.
-In the end it's a distance metric and it can be pretty similar to *any* distance metric.
-I think we can leave as it is (plus, I've reinvented the wheel by mathematicallty demonstrating it's a distance - my first only real math in a draft).
 
@@
 \begin{equation}
-    CD_{(x,y)} = 1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}\right)
+    CD_{(x,y)} = \frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}
 \end{equation}
+%CD is similar to https://onlinelibrary.wiley.com/doi/full/10.1111/pala.12384 but by character rather than by species
 
 \noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
-$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
+$CD$ is a continuous scaled Hamming distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
 Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
@@
 Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
+For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
+Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).
 
@@
 In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
-This procedure was used to treat all the characters are unordered with no assumption on the meaning of the character state (e.g. in a binary character $0$ is not necessary ancestral to $1$).
-It also greatly improved the speed of our algorithm implementation to compare the characters.
-This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
+This procedure allows comparison of characters regardless of the significance given to their tokens (following the \textit{xyz} notation in \citealt{felsenstein2004inferring}; as used in \citealt{Davalos01072014}).
+This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
 Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).
-
 When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
@@
 \textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
-For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
+For example with three characters \texttt{A = \{0,1,2,1\}}, \texttt{B = \{2,0,1,0\}} and \texttt{C = \{0,0,1,2\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$ (translated into \texttt{A' = \{1,2,3,2\}} \texttt{B' = \{1,2,3,2\}} and \texttt{C' = \{1,1,2,3\}}; $CD_{(A,B)} = (0+0+0+0)/(4-1) = 0$ and $CD_{(A,C)} = (0+1+1+1)/(4-1) = 1$).
 Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different set of splits (i.e. a different phylogenetic signal).
@@
 \begin{itemize}
-    \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
+    \item The ``morphological HKY-binary'' model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
     This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use an M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
-    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
+    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep[from empirical observations in: ][]{Guillerme2016146,ZouConvergence}.
     This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g.
 Dollo characters;][]{Dollo,wright2015came}.
-    Recently however, \cite{Wright01072016} have shown that an equal rate transition is still the most present in empirical data.
+    However, \cite{Wright01072016} have shown that equal transition rate is a reasonable assumption for many empirical datasets.
 \end{itemize}
 
-\noindent For each character, both models (morphological HKY-binary or M$k$) where chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
+\noindent For each character, both models (``morphological HKY-binary'' or M$k$) were chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
 These low evolutionary rate values allowed reduction in the number of homoplasic character changes, thus reinforcing the phylogenetic information in the matrices.
 We re-simulated every invariant characters to obtain a matrix with no invariant characters in order to better approximate real morphological data matrices.
-To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{OReilly20160081}.
+To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{sanderson1989patterns,OReilly20160081}.
+Note that both models (``morphological HKY-binary' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
+However, the ``morphological HKY model'' is not the one \textit{directly} estimated in the phylogenetic software used in this protocol (MrBayes, see below).
 
@@
 The matrices were generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 35 times resulting in 315 ``normal'' morphological matrices.
+To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 20 times resulting in 180 ``normal'' morphological matrices.
 
@@
 randomly weighting characters).
+Each replaced characters (i.e. the characters with a weight of +$1$) were randomly selected by sampling them randomly from the list of the remaining characters.
 Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
-This step resulted in a total of 1260 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
+This step resulted in a total of 720 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
 The algorithms for the three modifications are available on GitHub (\url{https://github.com/TGuillerme/CharactersCorrelation})
@@
 \subsection{Inferring topologies}
-We inferred the topologies with both BPP and MP using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
-For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
+We inferred the topologies with both Bayesian inference and maximum parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
+For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees (the first taxon in the taxa list).
 The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
@@
 We automatically stopped the MCMC when the average standard deviation of split frequencies (ASDSF) between both runs fell below 0.01 (with a diagnosis every $1\times10^4$ generations - \texttt{mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES}).
-Due to cluster hardware requirements an to save some time, when chains didn't converged and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the unconverged chains.
+Due to cluster hardware requirements and to save some time, when chains didn't converge and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the unconverged chains.
 In practice, these few MCMC got stuck at an ASDSF around (but not below) 0.01.
 
-A strict majority rule tree was then calculated for both Bayesian an maximum parsimony trees.
+A majority rule tree was then calculated for both Bayesian an maximum parsimony trees.
 For the Bayesian consensus trees, the 25\% first trees of the posterior tree distribution were excluded as a burnin.
-The 2880 tree inferences took around one CPU century on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC}.
+The 1440 tree inferences took around one CPU century on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC} and the Australian QRIScloud Awoonga clusters (3Ghz).
 
@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 1 and 2).
+We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 7 and 8).
 Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
@@
 When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
+However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.
 
-The metric scores where calculated using the \texttt{TreeCmp} javascript \citep{Bogdanowicz2012}.
-The measurements where then standardised using the Normalised Tree Similarity metric \citep[$NTS$; i.e.
+The metric scores were calculated using the \texttt{TreeCmp} javascript \citep{Bogdanowicz2012}.
+The measurements were then standardised using the Normalised Tree Similarity metric \citep[$NTS$; i.e.
 centering the metrics scores using the mean metric score for 1000 pairwise comparisons between random trees with $n$ taxa;][]{Bogdanowicz2012,Guillerme2016146}.
@@
 
-Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one.
 The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}).
-Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
+Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its \textit{p}-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
 when multiple trees are equal to the ``normal'' tree).
-Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
-Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
+Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146}.
+Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $<0.05$.
 Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g.
@@
 
-The resulting full simulation was 3.5TB big so is not shared here (though the parameters are).
+The resulting full simulation was 2TB in size so is not shared here (though the parameters are).
 However, the resulting consensus trees on which the topological differences are calculated are available at \url{https://figshare.com/s/7a8fde8eaa39a3d3cf56}.
@@
    \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
-\caption{\small{Effect of character difference on recovering the ``normal'' topology.
-The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
+\caption{\small{Effect of character difference on recovering the ``normal'' topology in terms of clade conservation ($NTS_{RF}$).
+The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 f from top to bottom respectively.
 The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see supplementary material 3 for the full summary statistics).
-There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.869 and median $NTS_{Tr}$ = 0.787; see supplementary materials 3, Table 3) followed by randomising the characters (median $NTS_{RF}$ = 0.814 and median $NTS_{Tr}$ = 0.791; see supplementary materials 3, Table 2) and maximising the character difference (median $NTS_{RF}$ = 0.698 and median $NTS_{Tr}$ = 0.754; see supplementary materials 3, Table 2).
+There is a significant difference between all scenarios in terms of clade conservation ($NTS_{RF}$) (maximising, minimising and randomising; \textit{p} = 0; Table \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability of overlap between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
+There is however no difference in terms of taxon displacement ($NTS_{Tr}$) between the three scenarios (with a BC $>$ 0.98 for each comparison - Table \ref{Tab_pooledscenarios_test}).
 
@@
 \subsubsection{Number of characters}
-This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
-The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 and 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.
+This effect of the character difference affects clade conservation ($NTS_{RF}$) only over large differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ - Table \ref{Tab_pooledscharacters_test}).
+For closer differences between characters numbers (i.e.
+100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests showed clear differences ($p = 0$) but with high distribution overlap ($BC > 0.96$).
+Regarding taxon displacements ($NTS_{Tr}$), there was an effect of the number of characters (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - Table \ref{Tab_pooledscharacters_test}).
 
@@
 Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
-A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
-Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
+A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ (median $NTS_{RF}$ of 0.846 and 0.787 for 25 and 75 taxa respectively with a $p = 0$ and a $BC$ marginally $< 0.95$ - Table \ref{Tab_pooledstaxa_test} and supplementary materials 3, table 5) but no differences between 75 and 150 taxa and for $NTS_{Tr}$ ($BC > 0.95$ - Table \ref{Tab_pooledstaxa_test}).
+The significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
 
@@
 
-Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively) but not in terms of individual taxon placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively).
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\textit{p} = 0 Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.851 and 0.739 respectively; see supplementary materials 3, Table 6) but not in terms of individual taxon placements (\textit{p} = 0.041 but $BC > 0.95$; Table \ref{Tab_pooledsmethods_test}).
 
@@
 
-When looking at the combined effect of each parameter, the ``maximised'' and ``minimised'' scenarios are always significantly different with no high probability of overlap for both $NTS_{RF}$ and $NTS_{Tr}$ (Wilcoxon rank test p.value $<0.05$ and Bhattacharrya Coefficient $<0.95$ - see supplementary material 3).
-The same differences are observed when comparing the ``maximised'' scenario against the ``randomised'' one expect for: (1) the Bayesian inference with 25 taxa (with 100, 350 and 1000 characters) and with 75 taxa with 1000 characters for both $NTS_{RF}$ and $NTS_{Tr}$; and (2) the maximum parsimony for 25 taxa (with 350 and 1000) characters for both $NTS_{RF}$ and $NTS_{Tr}$ and 75 taxa with 100 characters for $NTS_{Tr}$.
-Identically, there was always a significant difference between the ``minimised'' scenario and the ``randomised'' one was expect for the matrix of 150 taxa and 100 characters under maximum parsimony for $NTS_{RF}$ and the matrix of 150 and 1000 characters under Bayesian inference for $NTS_{Tr}$.
-The full list of comparisons and summary statistics are available in the supplementary materials 3.
+When looking at the combined effect of each parameter, the ``maximised'', ``minimised'' and ``randomised'' scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).
+When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences  between  tree similarity of differening numbers of taxa, characters, tree correlation, or tree inference method for the taxon displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).
+In terms of clade conservation ($NTS_{RF}$), however, ``maximised'' and ``minimised'' scenarios produced nearly always consistently different topologies using Bayesian inference (apart for the matrices with 25 taxa and 100 and 350 characters and the matrices with 150 taxa and 350 characters) but not using parsimony (only the matrices with 75 taxa and 100 or 350 characters produced different $NTS_{RF}$ distributions - supplementary material 3, tables 19 and 20).
 
@@
 \subsection{Effect of character differences on topology}
-As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
-The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
+As expected, there is a significant effect of character difference in the ability to recover the correct topology \hl{(in terms of clades)}.
+The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
-Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
-Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
-Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary material 3, Figs 3, 4 and 5).
+Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (\hl{i.e.
+because the characters are not different, they all support the same topology that might be incorrect}), the ``maximised'' scenario to lead to poorly resolved trees (\hl{i.e.
+because characters are different, the matrix has a lot of homoplasy leading to badly resolved topologies}) and the ``randomised'' scenario to perform the best at recovering the correct topology (\hl{i.e.
+a simple bootstrap procedure}).
+\hl{However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
+increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e.
+decreased correlation) led to slightly worst trees (Fig.}
+\ref{Fig:RF_results_best}.
+\hl{This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.}
+\ref{Fig:Tr_results_best}).
+\hl{However this results needs to be contrasted with the fact that}
+the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
+\hl{Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
+This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.}
 
@@
 Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance).
+\hl{Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
+If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
+Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.}
 
-Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distributions that are always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
-This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
-For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation as it is independent of the number of characters analysed.
-Similarly, using the a Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
+Although we measured a significant difference between smaller and larger matrices (\hl{in terms of number of characters, between 100 and 1000 characters - $BC$ $NTS_{RF}=0.914$ and $BC$ $NTS_{Tr}=0.826$}; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}),
+these differences have to be contrasted with the probability of overlap between the results distributions that are always \hl{very} high between every matrix's size (Bhattacharrya Coefficients $>0.949$ for both the different number of characters and taxa).
+This suggests that the overall character difference metric \hl{can be used as a proxy} for character correlation as it is \hl{not directly dependent} on the number of characters analysed.
+Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data (\hl{i.e.
+when combining the simulations together}).
+Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
 we corrected for the expected difference when comparing two random trees with the same number of taxa).
@@
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[i.e.
-those with in a $CI<0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
+those with in a $CI>0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
 
-The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
-For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
-(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
+The differences in performance of the Bayesian and parsimony approaches observed here could be due to the inherent mechanisms of each method.
+For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scores differently:
+(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the focal topology.
 The tree search will thus likely result in a topology island that will not contain the given topology $T$.
-(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e.
+(2) in Bayesian inference, the topology will also be given a bad score (i.e.
 low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
-The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
+The rate and the branch length being two parameters among others, the score (the likelihood) will change less drastically than for using parsimony.
 Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
-This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.b
+This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
+\hl{This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 12, 13, and 14).}
 
@@
 in evo-devo \citealt{goswami2006morphological}; or in macroevolution \citealt{fitzjohn2014much}).
-These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not used \textit{a priori} as a criterion to select characters.
-In other words, intra-organisaml and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
+These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not be used \textit{a priori} as a criterion to select characters.
+In other words, intra-organismal and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
 coding correlation) the only type of correlation that can affect the phylogeny \textit{a priori}.
-This dichotomy thus creates a trade of between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
-and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation being less dependent on the coding correlations.
+This dichotomy thus creates a trade-off between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
+and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation to be less dependent on the coding correlations.
 
@@
 creating ``modules'' of characters).
-However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest.
+However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices \hl{(i.e.
+inferring the topology first and looking at character's correlation based on the observed topology}\citealt{Lande1983,Maddison1990,Pagel1994}).
 
@@
 This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.
+\hl{It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
+This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches}
+\citep[e.g.][]{revell2014ancestral}.
+\hl{However, in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).}
 
 Second, measuring and modifying character correlation is difficult. 
-In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation.
+In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation
+\hl{(supplementary materials 3, figures 1, 2 and 3).}
 In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
@@
 Third, comparing phylogenetic inference methods is not trivial.
-As mentioned above, both maximum parsimony and Bayesian inference, although aiming (and often achieving) to infer evolutionary history only have similar outputs and vastly differ in how optimality is measured.
+As mentioned above, both maximum parsimony and Bayesian inference, although aiming (and often achieving) to infer evolutionary history, only have similar outputs and vastly differ in how optimality is measured.
 But there are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
-However, we want to point out again that here we're no comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
+However, we want to point out again that here we are not comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
 
@@
 Effectively, our simulation protocol bootstraps our data ``with bias''.
-In the ``randomised'' scenarios the data is simply randomly bootstrapped simply we randomly remove and resample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
+In the ``randomised'' scenarios the data is simply randomly bootstrapped, simply we randomly remove and re-sample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
 However, in the ``minimised'' and ``maximised'' scenario, the bootstrapping we remove the characters with the lowest/highest overall character difference.
-For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the left characters.
+For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the characters left.
 
@@
 the equivalent of randomising the character differences in our analysis) will be sufficient to obtain a robust topology (\textit{cf.} actively collecting different characters).
-In matrices with more taxa, however, the ``maximised'' scenario resulted in better topological recovery than any other scenarios.
-Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.
+In matrices with more taxa, however, the \hl{``minimised''} scenario resulted in better topological recovery than any other scenarios.
+Applying this kind of bootstraps that minimised character difference by biasing the random sampling could thus \hl{help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.}
 
@@
 
-We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was high (low character differences), the topology was always the furthest away from the ``normal'' topology.
-Conversely, when correlation between characters was low, the topology was always the closest to the ``normal'' topology.
-These results seem independent on the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
+We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was low (high character differences), the topology was always the furthest away from the ``normal'' topology.
+Conversely, when correlation between characters was high, the topology was always the closest to the ``normal'' topology.
+These results appear independent of the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
 
-However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
-For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly decreased and increased correct topological recovery compared to randomly resample matrices.
+However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise \hl{or maximised} character correlation was not more significant than simply randomising (i.e.
+bootstrapping) the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
+For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly increased and decreased correct topological recovery compared to randomly resample matrices.
 
@@
 \section{Acknowledgements}
-Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
-We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD metric proof.
+Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232) and the Australian QRIScloud Awoonga cluster services.
+We thank Alberto Pascual Garcia for input in the design of the simulations protocol, Guillermo Herraiz Yebes for helping with the CD metric proof and Orlin Todorov for assistance with the computations.
+We also thank the \hl{three anonymous reviewers for there constructive and insightful comments}.
 

### 2019-06-21 Writing/draft.tex
Finalised version (to check)

@@
 %Running head
-\begin{flushright}s
+\begin{flushright}
 Version dated: \today
@@
 \subsection{Effect of character differences on topology}
-As expected, there is a significant effect of character difference in the ability to recover the correct topology \hl{(in terms of clades)}.
+As expected, there is a significant effect of character difference in the ability to recover the correct topology (in terms of clades).
 The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
-Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (\hl{i.e.
-because the characters are not different, they all support the same topology that might be incorrect}), the ``maximised'' scenario to lead to poorly resolved trees (\hl{i.e.
-because characters are different, the matrix has a lot of homoplasy leading to badly resolved topologies}) and the ``randomised'' scenario to perform the best at recovering the correct topology (\hl{i.e.
-a simple bootstrap procedure}).
-\hl{However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
+Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (i.e.
+because the characters are not different, they all support the same topology that might be incorrect), the ``maximised'' scenario to lead to poorly resolved trees (i.e.
+because characters are different, the matrix has a lot of homoplasy leading to badly resolved topologies) and the ``randomised'' scenario to perform the best at recovering the correct topology (i.e.
+a simple bootstrap procedure).
+However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
 increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e.
-decreased correlation) led to slightly worst trees (Fig.}
-\ref{Fig:RF_results_best}.
-\hl{This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.}
-\ref{Fig:Tr_results_best}).
-\hl{However this results needs to be contrasted with the fact that}
-the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, \hl{Figs 9, 10 and 11}).
-\hl{Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+decreased correlation) led to slightly worst trees (Fig.
+\ref{Fig:RF_results_best}).
+This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
+However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 9, 10 and 11).
+Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
 However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
-This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.}
+This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.
 
@@
 Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance).
-\hl{Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
+Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
 If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
-Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.}
+Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.
 
-Although we measured a significant difference between smaller and larger matrices (\hl{in terms of number of characters, between 100 and 1000 characters - $BC$ $NTS_{RF}=0.914$ and $BC$ $NTS_{Tr}=0.826$}; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}),
-these differences have to be contrasted with the probability of overlap between the results distributions that are always \hl{very} high between every matrix's size (Bhattacharrya Coefficients $>0.949$ for both the different number of characters and taxa).
-This suggests that the overall character difference metric \hl{can be used as a proxy} for character correlation as it is \hl{not directly dependent} on the number of characters analysed.
-Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data (\hl{i.e.
-when combining the simulations together}).
+Although we measured a significant difference between smaller and larger matrices (in terms of number of characters, between 100 and 1000 characters - $BC$ $NTS_{RF}=0.914$ and $BC$ $NTS_{Tr}=0.826$; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}),
+these differences have to be contrasted with the probability of overlap between the results distributions that are always very high between every matrix's size (Bhattacharrya Coefficients $>0.949$ for both the different number of characters and taxa).
+This suggests that the overall character difference metric can be used as a proxy for character correlation as it is not directly dependent on the number of characters analysed.
+Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data (i.e.
+when combining the simulations together).
 Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
@@
 This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
-\hl{This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 12, 13, and 14).}
+This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 12, 13, and 14).
 
@@
 creating ``modules'' of characters).
-However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices \hl{(i.e.
-inferring the topology first and looking at character's correlation based on the observed topology}\citealt{Lande1983,Maddison1990,Pagel1994}).
+However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices (i.e.
+inferring the topology first and looking at character's correlation based on the observed topology \citealt{Lande1983,Maddison1990,Pagel1994}).
 
@@
 This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.
-\hl{It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
-This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches}
-\citep[e.g.][]{revell2014ancestral}.
-\hl{However, in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).}
+It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
+This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches \citep[e.g.][]{revell2014ancestral}.
+However, in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).
 
 Second, measuring and modifying character correlation is difficult. 
-In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation
-\hl{(supplementary materials 3, figures 1, 2 and 3).}
+In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 1, 2 and 3).
 In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
@@
 the equivalent of randomising the character differences in our analysis) will be sufficient to obtain a robust topology (\textit{cf.} actively collecting different characters).
-In matrices with more taxa, however, the \hl{``minimised''} scenario resulted in better topological recovery than any other scenarios.
-Applying this kind of bootstraps that minimised character difference by biasing the random sampling could thus \hl{help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.}
+In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery than any other scenarios.
+Applying this kind of bootstraps that minimised character difference by biasing the random sampling could thus help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.
 
@@
 
-However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise \hl{or maximised} character correlation was not more significant than simply randomising (i.e.
+However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise or maximised character correlation was not more significant than simply randomising (i.e.
 bootstrapping) the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
@@
 We thank Alberto Pascual Garcia for input in the design of the simulations protocol, Guillermo Herraiz Yebes for helping with the CD metric proof and Orlin Todorov for assistance with the computations.
-We also thank the \hl{three anonymous reviewers for there constructive and insightful comments}.
+We also thank the three anonymous reviewers for there constructive and insightful comments.
 

### 2019-06-21 Writing/SystematicBiology/draft.tex
Resubmission check

--- 2018-04-19 Writing/draft.tex
@@
 \begin{document}
-
 %Running head
@@
 \end{flushright}
+
 \bigskip
 \noindent RH: Characters correlation
-
 \bigskip
@@
 
-The last two decades have witnessed a ``resurgence'' of interest in the use of morphological character data in phylogenetic studies.
+The last two decades have witnessed a resurgence of interest in the use of morphological character data in phylogenetic studies.
 This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
-Morphological character data are often considered inferior to molecular sequence data, but are often the only source of phylogenetic data for extinct species. 
+Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collecting them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.
 While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
@@
 
-The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data.
+The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data \cite[e.g.][]{Davalos01072014, ZouConvergence}.
 The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the very nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other.
-Before discussing character correlation further, it is important to understand that it may manifest itself in at least three distinct ways:
+This independence is assumed for both molecular and morphological characters \citep{huelsenbeck1999effect,Davalos01072014,ZouConvergence}.
+However, several distinctions can be made between the amount and the nature of the correlation.
+First, molecular and morphological data often differs in size and in the number of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but may have an arbitrary number of states (up to that supported by phylogenetic software; \citealt{Guillerme2016146}).
+Second, correlations have been proposed to be stronger in morphological data (i.e.
+``concentrated'' correlations) \textit{versus} weaker and more ``spread-out'' correlations in molecular data \citep[i.e.
+``diffused'' correlations - though some regions of the genome may be thought as more correlated than others][]{huelsenbeck1999effect}.
+Furthermore, because of the ``spread'' of correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
+changes in morphological states can be more readily linked to a function \textit{cf.} changes in nucleotide states ; see below).
+Before discussing character correlation further, we distinguish three different modes of forms of character correlation:
 
@@
 \centering
-   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
+   \includegraphics[width=0.9\textwidth]{outline.eps}
 \caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using MP and BPP methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
@@
 \begin{equation}
-    CD_{(x,y)} = 1 - 2\left(\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}\right)
+    CD_{(x,y)} = \frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}
 \end{equation}
+%CD is similar to https://onlinelibrary.wiley.com/doi/full/10.1111/pala.12384 but by character rather than by species
 
 \noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
-$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
+$CD$ is a continuous scaled Hamming distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
 Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
@@
 \centering
-   \includegraphics[width=1\textwidth]{Figures/Modif_matrix.pdf}
+   \includegraphics[width=1\textwidth]{Modif_matrix.eps}
 \caption{Example illustration of the protocol for modifying matrices.
@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).]
+We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 7 and 8).
 Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
@@
 
-Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one.
-The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see Supplementary material 3).
-Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
+The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}).
+Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its \textit{p}-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
 when multiple trees are equal to the ``normal'' tree).
-Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
-Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
+Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146}.
+Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $<0.05$.
 Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g.
@@
 \centering
-   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
-\caption{\small{Effect of character difference on recovering the ``normal'' topology.
-The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
+   \includegraphics[width=1\textwidth]{RF_results_best.eps}
+\caption{\small{Effect of character difference on recovering the ``normal'' topology in terms of clade conservation ($NTS_{RF}$).
+The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 f from top to bottom respectively.
 The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
@@
 \centering
-   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf}
-\caption{Effect of character difference on recovering the ``normal'' topology.
+   \includegraphics[width=1\textwidth]{Tr_results_best.eps}
+\caption{Effect of character difference on recovering the ``normal'' topology in terms of Triplets ($NTS_{Tr}$).
 The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see Supplementary material 3 for the full summary statistics).
-There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.869 and median $NTS_{Tr}$ = 0.787; see supplementary materials 3, Table 3) followed by randomising the characters (median $NTS_{RF}$ = 0.814 and median $NTS_{Tr}$ = 0.791; see supplementary materials 3, Table 2) and maximising the character difference (median $NTS_{RF}$ = 0.698 and median $NTS_{Tr}$ = 0.754; see supplementary materials 3, Table 2).
+There is a significant difference between all scenarios in terms of clade conservation ($NTS_{RF}$) (maximising, minimising and randomising; \textit{p} = 0; Table \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability of overlap between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
+There is however no difference in terms of taxon displacement ($NTS_{Tr}$) between the three scenarios (with a BC $>$ 0.98 for each comparison - Table \ref{Tab_pooledscenarios_test}).
 
-\input{Tables/Tab_pooledscenarios_test.tex}
+\input{Tab_pooledscenarios_test.tex}
 
 \subsubsection{Number of characters}
-This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
-The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 and 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.
+This effect of the character difference affects clade conservation ($NTS_{RF}$) only over large differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ - Table \ref{Tab_pooledscharacters_test}).
+For closer differences between characters numbers (i.e.
+100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests showed clear differences ($p = 0$) but with high distribution overlap ($BC > 0.96$).
+Regarding taxon displacements ($NTS_{Tr}$), there was an effect of the number of characters (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - Table \ref{Tab_pooledscharacters_test}).
 
-\input{Tables/Tab_pooledscharacters_test.tex}
+\input{Tab_pooledscharacters_test.tex}
 
@@
 Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
-A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
-Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
+A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ (median $NTS_{RF}$ of 0.846 and 0.787 for 25 and 75 taxa respectively with a $p = 0$ and a $BC$ marginally $< 0.95$ - Table \ref{Tab_pooledstaxa_test} and supplementary materials 3, table 5) but no differences between 75 and 150 taxa and for $NTS_{Tr}$ ($BC > 0.95$ - Table \ref{Tab_pooledstaxa_test}).
+The significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
 
-\input{Tables/Tab_pooledstaxa_test.tex}
+\input{Tab_pooledstaxa_test.tex}
 
@@
 
-Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively) but not in terms of individual taxon placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively).
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\textit{p} = 0 Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.851 and 0.739 respectively; see supplementary materials 3, Table 6) but not in terms of individual taxon placements (\textit{p} = 0.041 but $BC > 0.95$; Table \ref{Tab_pooledsmethods_test}).
 
-\input{Tables/Tab_pooledsmethods_test.tex}
+\input{Tab_pooledsmethods_test.tex}
 
@@
 \subsection{Effect of character differences on topology}
-As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
-The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
+As expected, there is a significant effect of character difference in the ability to recover the correct topology (in terms of clades).
+The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
-Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
-Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
-Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary materials @@@).
+Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (i.e.
+because the characters are not different, they all support the same topology that might be incorrect), the ``maximised'' scenario to lead to poorly resolved trees (i.e.
+because characters are different, the matrix has a lot of homoplasy leading to badly resolved topologies) and the ``randomised'' scenario to perform the best at recovering the correct topology (i.e.
+a simple bootstrap procedure).
+However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
+increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e.
+decreased correlation) led to slightly worst trees (Fig.
+\ref{Fig:RF_results_best}).
+This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
+However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 9, 10 and 11).
+Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
+This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.
 
@@
 Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance).
+Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
+If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
+Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.
 
-Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distributions that are always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
-This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
-For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation as it is independent of the number of characters analysed.
-Similarly, using the a Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
+Although we measured a significant difference between smaller and larger matrices (in terms of number of characters, between 100 and 1000 characters - $BC$ $NTS_{RF}=0.914$ and $BC$ $NTS_{Tr}=0.826$; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}),
+these differences have to be contrasted with the probability of overlap between the results distributions that are always very high between every matrix's size (Bhattacharrya Coefficients $>0.949$ for both the different number of characters and taxa).
+This suggests that the overall character difference metric can be used as a proxy for character correlation as it is not directly dependent on the number of characters analysed.
+Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data (i.e.
+when combining the simulations together).
+Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
 we corrected for the expected difference when comparing two random trees with the same number of taxa).
@@
 In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[i.e.
-those with in a $CI<0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
+those with in a $CI>0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
 
-The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
-For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
-(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
+The differences in performance of the Bayesian and parsimony approaches observed here could be due to the inherent mechanisms of each method.
+For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scores differently:
+(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the focal topology.
 The tree search will thus likely result in a topology island that will not contain the given topology $T$.
-(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e.
+(2) in Bayesian inference, the topology will also be given a bad score (i.e.
 low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
-The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
+The rate and the branch length being two parameters among others, the score (the likelihood) will change less drastically than for using parsimony.
 Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
-This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.
+This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
+This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 12, 13, and 14).
 
@@
 creating ``modules'' of characters).
-However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest.
+However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices (i.e.
+inferring the topology first and looking at character's correlation based on the observed topology \citealt{Lande1983,Maddison1990,Pagel1994}).
 
@@
 This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.
+It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
+This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches \citep[e.g.][]{revell2014ancestral}.
+However, in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).
 
-%TG: insisting on the shittiness of generatin correlation
 Second, measuring and modifying character correlation is difficult. 
-In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation.
+In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 1, 2 and 3).
 In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
 This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
-%MB: We also don't really explore character congruence, that correlations can be ranked and spread over multiple internodes.
-%MB: We didn't look at ranking and we didn't look at competing correlations. There's a lot more that could be done here so, we might talk about that.
-%TG: not sure what you mean by the ranking stuff but it could definitely fit here.
 However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
@@
 Third, comparing phylogenetic inference methods is not trivial.
-As mentioned above, both maximum parsimony and Bayesian inference, although aiming (and often achieving) to infer evolutionary history only have similar outputs and vastly differ in how optimality is measured.
+As mentioned above, both maximum parsimony and Bayesian inference, although aiming (and often achieving) to infer evolutionary history, only have similar outputs and vastly differ in how optimality is measured.
 But there are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
-However, we want to point out again that here we're no comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
+However, we want to point out again that here we are not comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
 
 \subsection{Potential applications}
-%TG: explaining that our "scenarios" are biased bootstraps
 Effectively, our simulation protocol bootstraps our data ``with bias''.
-In the ``randomised'' scenarios the data is simply randomly bootstrapped simply we randomly remove and resample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
+In the ``randomised'' scenarios the data is simply randomly bootstrapped, simply we randomly remove and re-sample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
 However, in the ``minimised'' and ``maximised'' scenario, the bootstrapping we remove the characters with the lowest/highest overall character difference.
-For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the left characters.
+For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the characters left.
 
-%TG: explaining that we could use it for "clever" bootstrapping in big matrices
 It is noteworthy to point that in rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
@@
 the equivalent of randomising the character differences in our analysis) will be sufficient to obtain a robust topology (\textit{cf.} actively collecting different characters).
-In matrices with more taxa, however, the ``maximised'' scenario resulted in better topological recovery than any other scenarios.
-Applying this kind of bootstraps that maximises character difference by biasing the random sampling could thus results in better resolved trees.
+In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery than any other scenarios.
+Applying this kind of bootstraps that minimised character difference by biasing the random sampling could thus help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.
 
@@
 Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g.
-for characters hierarchy][]{deLaet2015,BrazeauNA} already exist.
+for characters hierarchy][]{de2015parsimony,BrazeauNA} already exist.
 However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
-%MB: I have no idea what thae above sentence means.
-%TG: Modified. Is that clearer?
 Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
 
-We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was high (low character differences), the topology was always the furthest away from the ``normal'' topology.
-Conversely, when correlation between characters was low, the topology was always the closest to the ``normal'' topology.
-These results seem independent on the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
+We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was low (high character differences), the topology was always the furthest away from the ``normal'' topology.
+Conversely, when correlation between characters was high, the topology was always the closest to the ``normal'' topology.
+These results appear independent of the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
 
-However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise character correlation was not more significant than simply bootstrapping the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
-For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly decreased and increased correct topological recovery compared to randomly resample matrices.
+However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise or maximised character correlation was not more significant than simply randomising (i.e.
+bootstrapping) the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
+For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly increased and decreased correct topological recovery compared to randomly resample matrices.
 
@@
 \section{Acknowledgements}
-Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
-We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD metric proof.
-
-% Suggested reviewers:
-%- Liliana Davalos for MPE / Joe O'Reilly, April Wright
+Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232) and the Australian QRIScloud Awoonga cluster services.
+We thank Alberto Pascual Garcia for input in the design of the simulations protocol, Guillermo Herraiz Yebes for helping with the CD metric proof and Orlin Todorov for assistance with the computations.
+We also thank the three anonymous reviewers for there constructive and insightful comments.
 

### 2019-10-15 Writing/Palaeontology/Sup1-CharDiff.tex
Prepared Palaeontology submission

--- 2019-05-20 Writing/supplementary/Sup1-CharDiff.tex
@@
 \usepackage{lineno}
-\usepackage{dcolumn}
-\usepackage{mathtools}
+\usepackage{pdflscape}
+
+\usepackage{color,soul}
 
@@
 \end{center}}
-\renewcommand{\subsection}[1]{%
+\renewcommand{\subsubsection}[1]{%
 \vspace{2ex}
@@
 \begin{center}
-
 \noindent{\Large \bf Influence of different modes of morphological character correlation on phylogenetic tree inference}
-
 \bigskip
@@
 
-\bigskip
-
-\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, and Martin D. Brazeau$^{2,3}$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, Abigail I. Pastore$^{1}$, and Martin D. Brazeau$^{2,3}$}\\
+% Abigail I. Pastore$^{1}$
 \noindent {\small \it 
@@
 This can be interpreted as the probability of characters \texttt{X}, \texttt{Y} and \texttt{X}, \texttt{Z} giving a conflicting phylogenetic signal between the four taxa.
-Here \texttt{X}, \texttt{Y} will conflict for 70\% of the taxa and \texttt{X}, \texttt{X} for 0\% of them.
-
-\section{Distance metric proof}
-
-The Character Difference ($CD$) metric is a distance $d: CD \times CD \rightarrow [0;\infty)$ with any character $x$, $y$ and $z$ (using the \textit{xyz} notation) $\in CD$ if an only if the followings are satisfied:
-
-\begin{equation}
-    \label{Positivity}
-    CD_{(x,y)} \geq 0
-\end{equation}
-\begin{equation}
-    \label{Identity}
-    CD_{(x,y)} = 0 \Leftrightarrow x = y
-\end{equation}
-\begin{equation}
-    \label{Equality}
-    CD_{x,y} = CD_{y,x}
-\end{equation}
-\begin{equation}
-    \label{Subadditivity}
-    CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}
-\end{equation}
-
-\noindent Furthermore, $CD$ is a evolutionary distance metric if it satisfies the four point condition:
-
-\begin{equation}
-    \label{fourpoint}
-    CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )}
-\end{equation}
-
-\newpage
-
-\subsection{Character difference positivity}
-
-For any $n \in [2;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ because the difference between character states is considered as Fitch-like (unordered; i.e.
-the difference between two different characters states is equal to $1$ and the difference between two identical character states is equal to $0$).
-Furthermore, following the $xyz$ notation described above, for any $n > 1$, $x_{1}-y_{1}=0$.
-Therefore the sum of the differences $\sum_{i=1}^{n}\abs{x_{i}-y_{i}} \in [0;(n-1)]$.
-
-\begin{equation}
-    \label{positive_demo}
-    \begin{gathered}
-    0 \leq  \abs{x_{n}-y_{n}} \leq 1\\
-    0 \leq  \sum_{i=1}^{n}\abs{x_{i}-y_{i}} \leq (n-1)\\
-    0 \leq  \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1} \leq 1\\
-    \end{gathered}
-\end{equation}
-
-Thus equation \ref{Positivity} is true.
-
-\newpage
-
-\subsection{Character difference identity}
-
-For any $n \in [2;\infty)$ and $x_{n} = y_{n}$ then:
-%TG: Is that valid for a demonstration?
-\begin{equation}
-    \label{identitydemo}
-    \begin{aligned}
-    \abs{x_{n}-y_{n}} &= 0\\
-    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= 0\\
-    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1} &= 0\\
-    \end{aligned}
-\end{equation}
-
-Thus equation \ref{Identity} is true.
-
-\subsection{Character difference equality}
-
-For any $n \in [2;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$, similarly to equation \ref{positive_demo}:
-
-\begin{equation}
-    \label{equality_demo}
-    \begin{aligned}
-    \abs{x_{n}-y_{n}} &= \abs{y_{n}-x_{n}}\\
-    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= \sum_{i}^{n}\abs{y_{i}-x_{i}}\\
-    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1} &= \frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n-1}\\
-    \end{aligned}
-\end{equation}
-
-Thus equation \ref{Equality} is true.
-
-\subsection{Character difference subadditivity}
-
-For any $n, \in [2;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:
-
-\begin{equation}
-    \label{equality_demo}
-    \begin{aligned}
-    \text{since:}\\
-    \underbrace{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1}}_{\in[0,1]}
-    &\leq
-    \underbrace{\frac{\sum_{i}^{n}\abs{x_{i}-z_{i}}}{n-1}}_{\in[0,1]}
-    +
-    \underbrace{\frac{\sum_{i}^{n}\abs{z_{i}-y_{i}}}{n-1}}_{\in[0,1]}\\
-    \text{thus:}\\
-    \in[0,1] &\leq \in[0,1] + \in[0,1] \\
-    \in[0,1] &\leq \in[0,2]\\
-    \end{aligned}
-\end{equation}
-Thus equation \ref{Subadditivity} is true.
-
-\newpage
-
-\subsection{Four point condition}
-
-For any $n \in [2;\infty)$ and any four distinct characters $i$, $j$, $n$ and $m$, and if equations \ref{Positivity} and \ref{Subadditivity} are true, the pairwise distances $CD_{(i,j)}$, $CD_{(m,n)}$, $CD_{(i,m)}$, $CD_{(j,n)}$, $CD_{(j,m)}$ and $CD_{(i,n)}$ can be represented as a combination of any five values $a$, $b$, $c$, $d$ and $e$ $\geq 0$:
-
-\begin{figure}[!htbp]
-\centering
-   \includegraphics[width=0.5\textwidth]{Four_point.pdf}
-\end{figure}
-
-Then:
-
-\begin{equation}
-    \begin{aligned}
-    CD_{(i,j)} = a + b \\
-    CD_{(m,n)} = d + e \\
-    CD_{(i,m)} = a + c + d \\
-    CD_{(j,n)} = b + c + e \\
-    CD_{(j,m)} = b + c + d \\
-    CD_{(i,n)} = a + c + e \\
-    \end{aligned}
-\end{equation}
-
-Thus:
-
-\begin{equation}
-    \begin{gathered}
-    CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )} \\ 
-    a + b + d + e \leq \text{max} (a+c+d+b+c+e; b+c+d+a+c+e) \\
-    a + b + d + e \leq a+b+2c+e+d \\
-    2c \geq 0 \\
-    \end{gathered}
-\end{equation}
-
-Thus equation \ref{fourpoint} is true.
+Here \texttt{X}, \texttt{Y} will conflict for 75\% of the taxa and \texttt{X}, \texttt{X} for 0\% of them.
 

### 2019-10-15 Writing/Palaeontology/Tab_pooledscharacters_test.tex
Prepared Palaeontology submission

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Fri May 31 14:42:25 2019
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & c100 vs. c350 & \textbf{0.965} & 48263.500 & \textbf{0.000} \\ 
   & c100 vs. c1000 & 0.914 & 37714.500 & \textbf{0.000} \\ 
   & c350 vs. c1000 & \textbf{0.976} & 53138.000 & \textbf{0.000} \\ 
  Tr & c100 vs. c350 & 0.929 & 40253.000 & \textbf{0.000} \\ 
   & c100 vs. c1000 & 0.826 & 27923.000 & \textbf{0.000} \\ 
   & c350 vs. c1000 & \textbf{0.957} & 48164.500 & \textbf{0.000} \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of characters: 100 (c100), 350 (c350) and 1000 (c1000). Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscharacters_test}
\end{table}

### 2019-10-15 Writing/Palaeontology/Tab_pooledsmethods_test.tex
Prepared Palaeontology submission

--- 2019-10-15 Writing/Palaeontology/Tab_pooledscharacters_test.tex
@@
   \hline
-RF & c100 vs. c350 & \textbf{0.965} & 48263.500 & \textbf{0.000} \\ 
-   & c100 vs. c1000 & 0.914 & 37714.500 & \textbf{0.000} \\ 
-   & c350 vs. c1000 & \textbf{0.976} & 53138.000 & \textbf{0.000} \\ 
-  Tr & c100 vs. c350 & 0.929 & 40253.000 & \textbf{0.000} \\ 
-   & c100 vs. c1000 & 0.826 & 27923.000 & \textbf{0.000} \\ 
-   & c350 vs. c1000 & \textbf{0.957} & 48164.500 & \textbf{0.000} \\ 
+RF & bayesian vs. parsimony & 0.939 & 192659.000 & \textbf{0.000} \\ 
+  Tr & bayesian vs. parsimony & \textbf{0.986} & 157677.500 & \textbf{0.041} \\ 
    \hline
 \end{tabular}
-\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of characters: 100 (c100), 350 (c350) and 1000 (c1000).
+\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the two different inference methods: bayesian or maximum parsimony.
 Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).}
-\label{Tab_pooledscharacters_test}
+\label{Tab_pooledsmethods_test}
 \end{table}

### 2019-10-15 Writing/Palaeontology/Tab_pooledstaxa_test.tex
Prepared Palaeontology submission

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Fri May 31 14:42:25 2019
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & t25 vs. t75 & 0.949 & 79945.000 & \textbf{0.000} \\ 
   & t25 vs. t150 & \textbf{0.954} & 75906.000 & \textbf{0.000} \\ 
   & t75 vs. t150 & \textbf{0.984} & 60328.000 & 0.654 \\ 
  Tr & t25 vs. t75 & \textbf{0.960} & 72814.000 & \textbf{0.024} \\ 
   & t25 vs. t150 & \textbf{0.961} & 72716.000 & \textbf{0.027} \\ 
   & t75 vs. t150 & \textbf{0.980} & 64637.000 & 1.000 \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of taxa: 25 (t25), 75 (t75) and 150 (t150). Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledstaxa_test}
\end{table}

### 2019-10-15 Writing/Palaeontology/draft.tex
Prepared Palaeontology submission

--- 2019-06-21 Writing/SystematicBiology/draft.tex
@@
+% https://www.palass.org/sites/default/files/media/publications/for_authors/ITA_2016_v1.pdf
+% Abstract < 300w ---> OK
+% Keywirds < 6 ---> OK
 \documentclass[12pt,letterpaper]{article}
@@
    \includegraphics[width=0.9\textwidth]{outline.eps}
-\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using MP and BPP methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
+\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using maximum parsimony and Bayesian inference methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines).
 $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}

### 2019-10-21 Writing/Palaeontology/draft.tex
requirements
added R cite

@@
 % Abstract < 300w ---> OK
-% Keywirds < 6 ---> OK
+% Keywords < 6 ---> OK
+% Length ~25 pages ---> NOT OK
+
 \documentclass[12pt,letterpaper]{article}
@@
 \subsection{Simulating discrete morphological matrices}
-To simulate the matrices we applied a protocol very similar to \cite{Guillerme2016146}.
+To simulate the matrices we applied a protocol very similar to \cite{Guillerme2016146} using \texttt{R} \citep{R}.
 First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution maintaining $\lambda$ $>$ $\mu$ using the \texttt{diversitree} \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.

### 2019-10-22 Writing/Palaeontology/draft.tex
Updated figures numbers for SOM3

@@
 We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 7 and 8).
+We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 10 and 11).
 Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
@@
 When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
-However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.
+However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 15, 16 and 17.
 
@@
 This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
-However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 9, 10 and 11).
-Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
-However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
+However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
+Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 4, 5 and 6).
 This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.
@@
 This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
-This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 12, 13, and 14).
+This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 15, 16, and 17).
 
@@
 Second, measuring and modifying character correlation is difficult. 
-In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 1, 2 and 3).
+In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 4, 5 and 6).
 In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.

### 2019-10-22 Functions/get.brlen.distribution.R
Added proportion of 0 branch lengths

@@ -1,0 +1,70 @@
+#' @title 
+#'
+#' @description 
+#'
+#' @param path the path where to find the true trees
+#' @param ntaxa the number of taxa per trees (25, 75 or 150)
+#' @param internals logical, whether to get all the branches distributions (including internal branches, TRUE) or just the total branch length to the tips (FALSE, default)
+#' @param zeros a list of values under which branch lengths are considered as 0.
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+get.brlen.distribution <- function(path, ntaxa, internals = FALSE, zeros = c(0)) {
+
+    ## Get the tree paths
+    trees_in <- list.files(path)
+    trees_in <- trees_in[grep(paste0(ntaxa, "t_"), trees_in)]
+    trees_in <- paste0(path, trees_in)
+
+    ## Read all the trees
+    tree_list <- lapply(as.list(trees_in), function(x) read.nexus(x))
+
+    ## Get the branch lengths
+    if(internals == TRUE) {
+        ## Get all the branch lengths
+        brlengths_list <- lapply(tree_list, function(x) return(x$edge.length[-1]))
+    } else {
+        ## Get the total branch lengths for each tip
+        brlengths_list <- lapply(tree_list, function(x, ntaxa) node.depth.edgelength(x)[1:ntaxa], ntaxa = ntaxa)
+    }
+
+    ## Get the density lines
+    brlengths_densities <- lapply(brlengths_list, function(x) density(x))
+    ## Get the proportion of 0 branch lengths
+    zero_list <- list()
+    for(zero in 1:length(zeros)) {
+        zero_list[[zero]] <- lapply(brlengths_list, function(x) ifelse(any(x < zeros[zero]), which(x < zeros[zero]), 0))
+    }
+    names(zero_list) <- as.character(zeros)
+
+    return(list(densities = brlengths_densities, zeros = zero_list))
+}
+
+plot.brlen.distribution <- function(path, ntaxa, zeros = zeros, cols, ...) {
+    ## Get the distributions
+    brlengths_int <- get.brlen.distribution(path, ntaxa, internals = TRUE, zeros = zeros)
+
+    ## Get the xlimits
+    xlim <- c(0, max(unlist(lapply(brlengths_int$densities, function(x) max(x$x)))))
[…]

### 2019-10-22 Writing/Palaeontology/draft.tex
Cut down results

@@
 The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-This independence is assumed for both molecular and morphological characters \citep{huelsenbeck1999effect,Davalos01072014,ZouConvergence}.
+% This independence is assumed for both molecular and morphological characters \citep{huelsenbeck1999effect,Davalos01072014,ZouConvergence}.
 However, several distinctions can be made between the amount and the nature of the correlation.
-First, molecular and morphological data often differs in size and in the number of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but may have an arbitrary number of states (up to that supported by phylogenetic software; \citealt{Guillerme2016146}).
-Second, correlations have been proposed to be stronger in morphological data (i.e.
+First, molecular and morphological data often differs in size and in the number of states analysed.
+ % (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but may have an arbitrary number of states (up to that supported by phylogenetic software; \citealt{Guillerme2016146}).
+Second, correlations have been proposed to be stronger in morphological data
+(i.e.
 ``concentrated'' correlations) \textit{versus} weaker and more ``spread-out'' correlations in molecular data \citep[i.e.
 ``diffused'' correlations - though some regions of the genome may be thought as more correlated than others][]{huelsenbeck1999effect}.
-Furthermore, because of the ``spread'' of correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
+% Furthermore, because of the ``spread'' of correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
 changes in morphological states can be more readily linked to a function \textit{cf.} changes in nucleotide states ; see below).
-Before discussing character correlation further, we distinguish three different modes of forms of character correlation:
+% Before discussing character correlation further,
+Here we distinguish three different modes of forms of character correlation:
 
@@
     Characters of the occlusal surface of two opposing molars will be expected to directly covary.
-    Pleiotropy also results in covariation between different aspects of phenotype and may thus induce correlations. 
-    From a phylogenetic perspective, it can be especially pernicious because the relationship between the traits in question may have no obvious link from a morphological or functional comparison alone.
+    % Pleiotropy also results in covariation between different aspects of phenotype and may thus induce correlations. 
+    % From a phylogenetic perspective, it can be especially pernicious because the relationship between the traits in question may have no obvious link from a morphological or functional comparison alone.
     Intra-organismal links can be the targets of comparative developmental biology \citep{goswami2010,kelly2010reduced,stoessel2013morphological,goswami2014macroevolutionary} or functional investigations.
@@
     For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
-    This is thought to correspond to adaptations for fossoriality or aquatic lifestyles.
-    Such covariances are generally studied in the context of a given phylogeny, often one derived from molecular data with the morphological traits of interest mapped on it.
+    % This is thought to correspond to adaptations for fossoriality or aquatic lifestyles.
+    % Such covariances are generally studied in the context of a given phylogeny, often one derived from molecular data with the morphological traits of interest mapped on it.
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
@@
     For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.
-    It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
-    For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
+    % It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
+    % For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
     This coding dependency is linked to hierarchical dependency between characters \citep{wilkinson1995coping,BrazeauNA}.
-    In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.
-    Finally this can also be due to a bias in the amount of characters available.
-    For example, in skulls, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity to discrete characters).
+    % In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.
+    % Finally this can also be due to a bias in the amount of characters available.
+    % For example, in skulls, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity to discrete characters).
 \end{itemize}
@@
 Intra-organismal dependence requires more detailed, often extremely time-consuming studies (and possibly beyond the limits of available technology).
-Even after all of the effort is expended, the results might then only be known for a single (model) species.
+% Even after all of the effort is expended, the results might then only be known for a single (model) species.
 Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree, and is best determined by independent character sets. 
-This is frequently accomplished by mapping morphological traits on molecular phylogenetic trees.
+% This is frequently accomplished by mapping morphological traits on molecular phylogenetic trees.
 
-These sources of dependence between characters are well studied in biology.
+% These sources of dependence between characters are well studied in biology.
 Biological and evolutionary dependences are inherent parts to Evo-Devo and macroevolutionary studies and best practices to avoid coding-induced dependences are commonly known and applied.
@@
 character ontologies), and instead must interpret discrete tokens as signifiers of shared similarity.
-Therefore, phylogenetic character matrices that contain a great number of characters but a few number of tokens can easily lead to dependence between characters.
-For example, if we consider the following matrix containing four cetartiodactyls - say a pig (e.g.
-\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
+Therefore, phylogenetic character matrices that contain a great number of characters but a few number of tokens can easily lead to correlation between characters.
+For example, if we consider the following matrix containing four cetartiodactyls - say a pig (\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
 
@@
 Yet a phylogenetic software will treat all these four characters in exactly the same way.
-In fact, for practical reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
+% In fact, for practical reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
 In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} that seem to be more likely to reflect the split by separating artiodactyles from the cetaceans.
-In other words, this is the correlation between characters that depicts the evolutionary history of this group.
-However, the other characters depict the same history (i.e.
-the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleen, the same split would still be estimated by the software).
+% In other words, this is the correlation between characters that depicts the evolutionary history of this group.
+However, the other characters depict the same history %(i.e. the same split)
+ but for the wrong reasons. % (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleen, the same split would still be estimated by the software).
 One solution would be to a) weight some characters (e.g.
@@
 Here we formally assess the effect of discrete character's correlation using simulated data.
-We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
+We propose % a new
+distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
 We found that overall, there is a detectable effect of character correlation on topology where a decrease in character dependence results in a decrease in the ability to recover the correct topology.
@@
     \item \textbf{Modifying matrices}: we changed the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least different or most different characters and replacing them randomly by characters from the remaining set.
-Our protocol for measuring character difference is detailed below.
-
-    We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character differences to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
-This step resulted in 36 matrices.
+    We also randomised the ``normal'' matrices without bias towards maximising or minimising character differences (hereafter called the ``randomised'' matrices).
 
@@
     Hereafter, the resultant topologies are called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
-This step resulted in 72 topologies.
+
     \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on the recovered topology.
+
 \end{enumerate}
-All steps were replicated 20 times and are described below in more detail, along with our proposed definition for measuring the difference between characters.
+All steps were replicated 20 times and are described below in more detail.
 
@@
 entirely correlated characters would define identical splits between a set of taxa).
-In order to measure this, we propose a new distance metric to measure the difference between two characters:
+In order to measure this, we used the following scaled Hamming distance between two characters:
 
@@
 \noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
-$CD$ is a continuous scaled Hamming distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
-Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
-Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
-Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
-For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
-Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).
+$CD$ is a continuous scaled Hamming distance metric bounded between $0$ and $1$.
+Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way %.
+% Two same character states tokens have a difference of zero and two different ones have a difference of one
+(e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
+% Additionally, we only consider differences for taxa with shared information \citep{GowerDist}.
+% For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
+% Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).
 
@@
 This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
-Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).
+% Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).
 When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
@@
 \textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
-For example with three characters \texttt{A = \{0,1,2,1\}}, \texttt{B = \{2,0,1,0\}} and \texttt{C = \{0,0,1,2\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$ (translated into \texttt{A' = \{1,2,3,2\}} \texttt{B' = \{1,2,3,2\}} and \texttt{C' = \{1,1,2,3\}}; $CD_{(A,B)} = (0+0+0+0)/(4-1) = 0$ and $CD_{(A,C)} = (0+1+1+1)/(4-1) = 1$).
-Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different set of splits (i.e. a different phylogenetic signal).
+% For example with three characters \texttt{A = \{0,1,2,1\}}, \texttt{B = \{2,0,1,0\}} and \texttt{C = \{0,0,1,2\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$ (translated into \texttt{A' = \{1,2,3,2\}} \texttt{B' = \{1,2,3,2\}} and \texttt{C' = \{1,1,2,3\}}; $CD_{(A,B)} = (0+0+0+0)/(4-1) = 0$ and $CD_{(A,C)} = (0+1+1+1)/(4-1) = 1$).
+% Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different set of splits (i.e.
+a different phylogenetic signal).
 
@@
 To simulate the matrices we applied a protocol very similar to \cite{Guillerme2016146} using \texttt{R} \citep{R}.
-First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution maintaining $\lambda$ $>$ $\mu$ using the \texttt{diversitree} \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
+First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution with $\lambda$ $>$ $\mu$\citep[\texttt{diversitree} v0.9-8;][]{fitzjohndiversitree2012} and saving the trees after reaching either 25, 75 or 150 taxa (see supplementary material 3 Figs 1, 2 and 3).
 For each tree, we arbitrarily set the outgroup to be the first taxon (alphabetically) thus effectively rooting the trees on this taxon.
 These trees are hereafter called the ``true'' trees (see distinction below).
+
 We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
@@
     \item The ``morphological HKY-binary'' model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
-    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use an M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
+    % This model has the advantage of not favouring Bayesian inference \citep[; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
-    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep[from empirical observations in: ][]{Guillerme2016146,ZouConvergence}.
-    This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g.
+    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
+    % This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g.
 Dollo characters;][]{Dollo,wright2015came}.
-    However, \cite{Wright01072016} have shown that equal transition rate is a reasonable assumption for many empirical datasets.
+    % However, \cite{Wright01072016} have shown that equal transition rate is a reasonable assumption for many empirical datasets.
 \end{itemize}
@@
 To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{sanderson1989patterns,OReilly20160081}.
-Note that both models (``morphological HKY-binary' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
-However, the ``morphological HKY model'' is not the one \textit{directly} estimated in the phylogenetic software used in this protocol (MrBayes, see below).
+% Note that both models (``morphological HKY-binary' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
+% However, the ``morphological HKY model'' is not the one \textit{directly} estimated in the phylogenetic software used in this protocol (MrBayes, see below).
 
@@
 The matrices were generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 20 times resulting in 180 ``normal'' morphological matrices.
+% To estimate the variance of our simulations and assess the effect of our random parameters,
+We repeated this step 20 times resulting in 180 ``normal'' morphological matrices.
 
@@
 randomly weighting characters).
-Each replaced characters (i.e. the characters with a weight of +$1$) were randomly selected by sampling them randomly from the list of the remaining characters.
-Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
+Each replaced characters were randomly selected by sampling them randomly from the list of the remaining characters.
+% Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
 This step resulted in a total of 720 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
@@
 For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees (the first taxon in the taxa list).
-The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
+The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates.
+ % (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
 
-Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$ - \texttt{lset nst=1 rates=gamma Ngammacat=4}) with an variable rate prior an exponential (0.5) shape (\texttt{prset ratepr=variable Shapepr=Exponential(0.5)}).
-We ran two runs of 6 chains each (2 hot, 4 cold) for a maximum of $1\times10^9$ generations with a sampling every 200 generations.
-We automatically stopped the MCMC when the average standard deviation of split frequencies (ASDSF) between both runs fell below 0.01 (with a diagnosis every $1\times10^4$ generations - \texttt{mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES}).
+Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$)
+ % - \texttt{lset nst=1 rates=gamma Ngammacat=4})
+with an variable rate prior an exponential (0.5) shape.
+ % (\texttt{prset ratepr=variable Shapepr=Exponential(0.5)}).
+The MCMC was ran over two runs of 6 chains each for a maximum of $1\times10^9$ generations with a sampling every 200 generations with an automatic stop if the average standard deviation of split frequency (ASDSF) fell below 0.01.
+% \texttt{mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES}).
 Due to cluster hardware requirements and to save some time, when chains didn't converge and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the unconverged chains.
@@
 
-A majority rule tree was then calculated for both Bayesian an maximum parsimony trees.
-For the Bayesian consensus trees, the 25\% first trees of the posterior tree distribution were excluded as a burnin.
+A majority rule tree was then calculated for both maximum parsimony and Bayesian trees (discarding the 25\% first trees).
 The 1440 tree inferences took around one CPU century on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC} and the Australian QRIScloud Awoonga clusters (3Ghz).
@@
 Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
-When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
-However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 15, 16 and 17.
+% When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
+% However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 15, 16 and 17.
 
@@
 centering the metrics scores using the mean metric score for 1000 pairwise comparisons between random trees with $n$ taxa;][]{Bogdanowicz2012,Guillerme2016146}.
-When the normalised metric has a score of one it means both trees are identical, when it has a score of zero it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
-The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Foulds ($NTS_{RF}$) Similarity reflects the conservation of clades (i.e.
-a score close to $1$ indicates that most clades are identical in both trees); and (2) the Normalised Triplets Similarity ($NTS_{Tr}$) reflects the position of taxa (i.e.
-a score close to $1$ indicates that most taxa have the same neighbours in both trees).
+% When the normalised metric has a score of one it means both trees are identical, when it has a score of zero it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
+The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Foulds Similarity ($NTS_{RF}$)  reflects the conservation of clades
+ % (i.e. a score close to $1$ indicates that most clades are identical in both trees)
+; and (2) the Normalised Triplets Similarity ($NTS_{Tr}$) reflects the position of taxa.
+% (i.e. a score close to $1$ indicates that most taxa have the same neighbours in both trees).
 
-The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}).
-Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its \textit{p}-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
+% The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}).
+% Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its \textit{p}-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
 when multiple trees are equal to the ``normal'' tree).
-Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146}.
-Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $<0.05$.
-Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g.
+To measure the effect of character correlation, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep{holm1979simple} and a the probability of overlap between two distributions \citep[the Bhattacharyya Coefficient: $BC$;][]{Bhattacharyya}.
+% Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $<0.05$.
+% Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g.
 for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).
@@
 
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.869 and median $NTS_{Tr}$ = 0.787; see supplementary materials 3, Table 3) followed by randomising the characters (median $NTS_{RF}$ = 0.814 and median $NTS_{Tr}$ = 0.791; see supplementary materials 3, Table 2) and maximising the character difference (median $NTS_{RF}$ = 0.698 and median $NTS_{Tr}$ = 0.754; see supplementary materials 3, Table 2).
-There is a significant difference between all scenarios in terms of clade conservation ($NTS_{RF}$) (maximising, minimising and randomising; \textit{p} = 0; Table \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability of overlap between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
-There is however no difference in terms of taxon displacement ($NTS_{Tr}$) between the three scenarios (with a BC $>$ 0.98 for each comparison - Table \ref{Tab_pooledscenarios_test}).
+The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information followed by randomising and maximising the character difference (see supplementary materials 3, Tables 3, 4 and 5).
+% (median $NTS_{RF}$ = 0.869 and median $NTS_{Tr}$ = 0.787; see supplementary materials 3, Table 3) followed by randomising the characters (median $NTS_{RF}$ = 0.814 and median $NTS_{Tr}$ = 0.791; see supplementary materials 3, Table 2) and maximising the character difference (median $NTS_{RF}$ = 0.698 and median $NTS_{Tr}$ = 0.754; see supplementary materials 3, Table 2).
+There is a significant difference between all scenarios in terms of clade conservation ($NTS_{RF}$ - Table \ref{Tab_pooledscenarios_test}).
+ % (maximising, minimising and randomising; \textit{p} = 0; Table \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (BC = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability of overlap between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
+However, there is no difference in terms of taxon displacement between the three scenarios ($NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}).
 
@@
 \subsubsection{Number of characters}
-This effect of the character difference affects clade conservation ($NTS_{RF}$) only over large differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ - Table \ref{Tab_pooledscharacters_test}).
-For closer differences between characters numbers (i.e.
-100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests showed clear differences ($p = 0$) but with high distribution overlap ($BC > 0.96$).
-Regarding taxon displacements ($NTS_{Tr}$), there was an effect of the number of characters (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - Table \ref{Tab_pooledscharacters_test}).
+This effect of the character difference affects clade conservation ($NTS_{RF}$) only over large differences in number of characters (100 \textit{vs.} 1000 - Table \ref{Tab_pooledscharacters_test}).
+% , median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ ).
+For closer differences between characters numbers (100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests showed clear differences ($p = 0$) but with high distribution overlap ($BC > 0.96$).
+Regarding taxon displacements ($NTS_{Tr}$), there was an effect of the number of characters  although the with a high distribution overlap between 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
+% (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - Table \ref{Tab_pooledscharacters_test}).
 
@@
 Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
-A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ (median $NTS_{RF}$ of 0.846 and 0.787 for 25 and 75 taxa respectively with a $p = 0$ and a $BC$ marginally $< 0.95$ - Table \ref{Tab_pooledstaxa_test} and supplementary materials 3, table 5) but no differences between 75 and 150 taxa and for $NTS_{Tr}$ ($BC > 0.95$ - Table \ref{Tab_pooledstaxa_test}).
+A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in $NTS_{RF}$ (Table \ref{Tab_pooledstaxa_test}) but no differences between 75 and 150 taxa for $NTS_{Tr}$ (Table \ref{Tab_pooledstaxa_test}).
+ % (median $NTS_{RF}$ of 0.846 and 0.787 for 25 and 75 taxa respectively with a $p = 0$ and a $BC$ marginally $< 0.95$ - Table \ref{Tab_pooledstaxa_test} and supplementary materials 3, table 5) but no differences between 75 and 150 taxa and for $NTS_{Tr}$ ($BC > 0.95$ - Table \ref{Tab_pooledstaxa_test}).
 The significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
@@
 
-Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\textit{p} = 0 Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.851 and 0.739 respectively; see supplementary materials 3, Table 6) but not in terms of individual taxon placements (\textit{p} = 0.041 but $BC > 0.95$; Table \ref{Tab_pooledsmethods_test}).
+Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\textit{p} = 0 Table \ref{Tab_pooledsmethods_test}) but not in terms of individual taxon placements (\textit{p} = 0.041 but $BC > 0.95$; Table \ref{Tab_pooledsmethods_test}).
 
@@
 When looking at the combined effect of each parameter, the ``maximised'', ``minimised'' and ``randomised'' scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).
-When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences  between  tree similarity of differening numbers of taxa, characters, tree correlation, or tree inference method for the taxon displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).
+When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences between tree similarity of differencing numbers of taxa, characters, tree correlation, or tree inference method for the taxon displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).
 In terms of clade conservation ($NTS_{RF}$), however, ``maximised'' and ``minimised'' scenarios produced nearly always consistently different topologies using Bayesian inference (apart for the matrices with 25 taxa and 100 and 350 characters and the matrices with 150 taxa and 350 characters) but not using parsimony (only the matrices with 75 taxa and 100 or 350 characters produced different $NTS_{RF}$ distributions - supplementary material 3, tables 19 and 20).
@@
 Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+% Add mention to this difference in character duplicates likely to affect the results
 However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 4, 5 and 6).

### 2019-10-23 Writing/Palaeontology/draft.tex
Shortened version (-10 pages!)

@@
 \section{Introduction}
-
 The last two decades have witnessed a resurgence of interest in the use of morphological character data in phylogenetic studies.
@@
 Second, correlations have been proposed to be stronger in morphological data
-(i.e.
-``concentrated'' correlations) \textit{versus} weaker and more ``spread-out'' correlations in molecular data \citep[i.e.
-``diffused'' correlations - though some regions of the genome may be thought as more correlated than others][]{huelsenbeck1999effect}.
+%(i.e. ``concentrated'' correlations)
+but weaker and more ``spread-out'' correlations in molecular data \citep[][]{huelsenbeck1999effect}.
+%i.e.
+``diffused'' correlations - though some regions of the genome may be thought as more correlated than others
 % Furthermore, because of the ``spread'' of correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
@@
 \section{Methods}
-
 To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different levels of correlation between some characters (Fig.\ref{Fig:outline} - note that each step is described in more details below):
@@
 \label{CDdescription}
-To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information (i.e.
-entirely correlated characters would define identical splits between a set of taxa).
+To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information.
+%(i.e. entirely correlated characters would define identical splits between a set of taxa).
 In order to measure this, we used the following scaled Hamming distance between two characters:
@@
 We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.
-In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
+We replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
 This procedure allows comparison of characters regardless of the significance given to their tokens (following the \textit{xyz} notation in \citealt{felsenstein2004inferring}; as used in \citealt{Davalos01072014}).
@@
 When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
-When the character difference is maximal ($1$) it means it conveys the greatest difference in phylogenetic signal (i.e. characters are uncorrelated).
-It is important to stress that a character difference of $0$ (i.e.
-the same phylogenetic signal) does not mean the opposite of $1$ (i.e.
-\textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
+When the character difference is maximal ($1$) it means it conveys the greatest difference in phylogenetic signal. % (i.e. characters are uncorrelated).
+However, a character difference of $0$ %(i.e. the same phylogenetic signal)
+does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
 % For example with three characters \texttt{A = \{0,1,2,1\}}, \texttt{B = \{2,0,1,0\}} and \texttt{C = \{0,0,1,2\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$ (translated into \texttt{A' = \{1,2,3,2\}} \texttt{B' = \{1,2,3,2\}} and \texttt{C' = \{1,1,2,3\}}; $CD_{(A,B)} = (0+0+0+0)/(4-1) = 0$ and $CD_{(A,C)} = (0+1+1+1)/(4-1) = 1$).
@@
 These trees are hereafter called the ``true'' trees (see distinction below).
-
 We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
@@
 We then modified the matrices to either maximise or minimise the pairwise character differences for each matrix using three different algorithms.
-For maximising the pairwise differences between characters, we selected the characters that were the most similar to all the others (i.e.
-with an average character difference $<0.25$) and replaced them randomly by any of the remaining characters.
+For maximising the pairwise differences between characters, we selected the characters that were the most similar to all the others (mean character difference $<0.25$) and replaced them randomly by any of the remaining characters.
 This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
-Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e.
-with an average character difference $>0.75$) and randomly replaced them with the remaining ones.
+Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (mean character difference $>0.75$) and randomly replaced them with the remaining ones.
 Finally, because this operation effectively changes the weight of characters (i.e.
-giving the characters $<0.25$ or $>0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification (i.e.
+giving the characters $<0.25$ or $>0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification.% (i.e.
 randomly weighting characters).
@@
 The metric scores were calculated using the \texttt{TreeCmp} javascript \citep{Bogdanowicz2012}.
-The measurements were then standardised using the Normalised Tree Similarity metric \citep[$NTS$; i.e.
-centering the metrics scores using the mean metric score for 1000 pairwise comparisons between random trees with $n$ taxa;][]{Bogdanowicz2012,Guillerme2016146}.
+The measurements were then standardised using the Normalised Tree Similarity metric \citep[$NTS$; centering and scaling the metric score to the number of taxa;][]{Bogdanowicz2012,Guillerme2016146}.
 % When the normalised metric has a score of one it means both trees are identical, when it has a score of zero it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
@@
 for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).
-
 The resulting full simulation was 2TB in size so is not shared here (though the parameters are).
@@
 \subsection{Effect of character differences on topology}
-
 The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information followed by randomising and maximising the character difference (see supplementary materials 3, Tables 3, 4 and 5).
@@
 \subsubsection{Number of taxa}
-
 Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
@@
 \subsection{Effect of character differences on the inference method}
-
 Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\textit{p} = 0 Table \ref{Tab_pooledsmethods_test}) but not in terms of individual taxon placements (\textit{p} = 0.041 but $BC > 0.95$; Table \ref{Tab_pooledsmethods_test}).
@@
 \subsection{Combined effects of taxa, characters and correlation on topology}
-
 When looking at the combined effect of each parameter, the ``maximised'', ``minimised'' and ``randomised'' scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).
@@
 As expected, there is a significant effect of character difference in the ability to recover the correct topology (in terms of clades).
-The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
+% The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
 When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
-Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (i.e.
+% Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (i.e.
 because the characters are not different, they all support the same topology that might be incorrect), the ``maximised'' scenario to lead to poorly resolved trees (i.e.
@@
 However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
-increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e.
-decreased correlation) led to slightly worst trees (Fig.
+increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario led to slightly worst trees (Fig.
 \ref{Fig:RF_results_best}).
@@
 However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
-Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
-% Add mention to this difference in character duplicates likely to affect the results
-However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 4, 5 and 6).
+Furthermore, fewer characters were replaced, on average, in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6)% - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+% However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 4, 5 and 6).
 This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.
@@
 \subsubsection{Number of characters and taxa}
-Because of the nature of our simulation protocol, one could expect that the effect of character correlation would have increased with the number of characters (i.e.
+% Because of the nature of our simulation protocol, one could expect that the effect of character correlation would have increased with the number of characters (i.e.
 the more characters available, the more characters are modified in each scenario).
-Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance).
-Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
-If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
+% Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance).
+Overall, it is expected that minimising character difference %(increasing character correlation)
+ is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
+If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good.
 Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.
 
-Although we measured a significant difference between smaller and larger matrices (in terms of number of characters, between 100 and 1000 characters - $BC$ $NTS_{RF}=0.914$ and $BC$ $NTS_{Tr}=0.826$; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}),
-these differences have to be contrasted with the probability of overlap between the results distributions that are always very high between every matrix's size (Bhattacharrya Coefficients $>0.949$ for both the different number of characters and taxa).
-This suggests that the overall character difference metric can be used as a proxy for character correlation as it is not directly dependent on the number of characters analysed.
-Regarding the number of taxa, our results suggest that the effect of character correlation on recovering the right topology is independent on the size of the matrix when pooling the data (i.e.
-when combining the simulations together).
-Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e.
+Although we measured a significant difference between smaller and larger matrices (Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), %in terms of number of characters, between 100 and 1000 characters;
+these differences have to be contrasted with the Bhattacharrya Coefficients that are always very high between every matrix's size.
+% This suggests that the overall character difference metric can be used as a proxy for character correlation as it is not directly dependent on the number of characters analysed.
+Regarding the number of taxa, our results suggest the effect of character correlation on topology is independent on the size of the matrix when pooling the data.
+Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered.% (i.e.
 we corrected for the expected difference when comparing two random trees with the same number of taxa).
 
-\subsection{Effect of character differences on the inference method}
-When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
-The difference in the ability of each method to recover the ``correct'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have raised issues with these investigations \citealt{spencerefficacy2013,goloboff2017weighted}).
-In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[i.e.
-those with in a $CI>0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
+\subsubsection{Effect of character differences on the inference method}
+When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (Table \ref{Tab_pooledsmethods_test}).
+% both using a Wilcoxon test and the Bhattacharrya Coefficient; 
+The difference in the ability of each method to recover the ``correct'' topology has been recently discussed with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; but see \citealt{spencerefficacy2013,goloboff2017weighted}).
+In this study, it is possible that our simulation protocol for generating the characters (using some M$k$-based characters) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[using matrices with a $CI>0.26$;][]{OReilly20160081} could also favour maximum parsimony analysis.
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
 
-The differences in performance of the Bayesian and parsimony approaches observed here could be due to the inherent mechanisms of each method.
-For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scores differently:
-(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the focal topology.
-The tree search will thus likely result in a topology island that will not contain the given topology $T$.
-(2) in Bayesian inference, the topology will also be given a bad score (i.e.
+% The differences in performance of the Bayesian and parsimony approaches observed here could be due to the inherent mechanisms of each method.
+% For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scores differently:
+% (1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the focal topology.
+% The tree search will thus likely result in a topology island that will not contain the given topology $T$.
+% (2) in Bayesian inference, the topology will also be given a bad score (i.e.
 low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
-The rate and the branch length being two parameters among others, the score (the likelihood) will change less drastically than for using parsimony.
-Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
-This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
-This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 15, 16, and 17).
+% The rate and the branch length being two parameters among others, the score (the likelihood) will change less drastically than for using parsimony.
+% Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
+% This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
+% This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 15, 16, and 17).
 
@@
 These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not be used \textit{a priori} as a criterion to select characters.
-In other words, intra-organismal and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
+% In other words, intra-organismal and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
 coding correlation) the only type of correlation that can affect the phylogeny \textit{a priori}.
-This dichotomy thus creates a trade-off between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
+This creates a trade-off between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
 and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation to be less dependent on the coding correlations.
 
-It is important to note that the two other sources of character correlation could also be present in our simulations although they were not explicitly modelled:
-(1) evolutionary correlation is implied by simulating the characters based on Birth-Death trees;
-and (2) intra-organismal correlation could also be present in the matrices for those characters randomly simulated but sharing similar evolutionary simulation regimes (i.e.
-creating ``modules'' of characters).
-However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices (i.e.
-inferring the topology first and looking at character's correlation based on the observed topology \citealt{Lande1983,Maddison1990,Pagel1994}).
+The two other sources of character correlation are present in our simulations although they were not explicitly modelled:
+(1) evolutionary correlation is implied by simulating the characters using Birth-Death trees;
+and (2) intra-organismal correlation is present in the matrices for the characters randomly simulated but sharing similar evolutionary simulation regimes.% (i.e. creating ``modules'' of characters).
+However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices \citep{Lande1983,Maddison1990,Pagel1994}.
+%(i.e. inferring the topology first and looking at character's correlation based on the observed topology \citealt{Lande1983,Maddison1990,Pagel1994}).
 
@@
 This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.
-It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
-This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches \citep[e.g.][]{revell2014ancestral}.
-However, in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).
+% It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
+% This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches \citep[e.g.][]{revell2014ancestral}.
+% However, in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).
 
@@
 However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
-Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e.
+% Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e.
 some sets of characters can be highly correlated but effectively describe independent information).
@@
 Third, comparing phylogenetic inference methods is not trivial.
-As mentioned above, both maximum parsimony and Bayesian inference, although aiming (and often achieving) to infer evolutionary history, only have similar outputs and vastly differ in how optimality is measured.
-But there are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
+As mentioned above, both maximum parsimony and Bayesian inference
+% although aiming (and often achieving) to infer evolutionary history,
+have similar outputs nut vastly differ in how optimality is measured.
+There are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
 However, we want to point out again that here we are not comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
@@
 Effectively, our simulation protocol bootstraps our data ``with bias''.
-In the ``randomised'' scenarios the data is simply randomly bootstrapped, simply we randomly remove and re-sample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
-However, in the ``minimised'' and ``maximised'' scenario, the bootstrapping we remove the characters with the lowest/highest overall character difference.
+In the ``randomised'' scenarios the data is simply randomly bootstrapped. %, simply we randomly remove and re-sample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
+However, in the ``minimised'' and ``maximised'' scenario, we remove the characters with the lowest/highest overall character difference.
 For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the characters left.
 
-It is noteworthy to point that in rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
-Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling (i.e.
-the equivalent of randomising the character differences in our analysis) will be sufficient to obtain a robust topology (\textit{cf.} actively collecting different characters).
-In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery than any other scenarios.
+In rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
+Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling
+%(i.e. the equivalent of randomising the character differences in our analysis)
+will be sufficient to obtain a robust topology.% (\textit{cf.} actively collecting different characters).
+In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery.
 Applying this kind of bootstraps that minimised character difference by biasing the random sampling could thus help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.
@@
 \subsection{Conclusion}
-Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon to investigate phylogenetic relationships.
+Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon.% to investigate phylogenetic relationships.
 Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g.
-for characters hierarchy][]{de2015parsimony,BrazeauNA} already exist.
-However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
-Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
+][]{de2015parsimony,BrazeauNA} already exist.
+However, with
+%the regain of popularity in discrete morphological data and
+the expansion of dataset size \citep[e.g.][> 1000 characters]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
+Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate between these correlation types, it is important to understand to what extant topologies can be induced by such bias.
 
@@
 These results appear independent of the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
-
 However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise or maximised character correlation was not more significant than simply randomising (i.e.
 bootstrapping) the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
-For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly increased and decreased correct topological recovery compared to randomly resample matrices.
+% For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly increased and decreased correct topological recovery compared to randomly resample matrices.
 

### 2019-10-26 Writing/Palaeontology/draft.tex
Update draft.tex
AP comments on Intro and Discussion
Nice intro, Discussion is not too far off.

@@
 % Keywords < 6 ---> OK
-% Length ~25 pages ---> NOT OK
 
@@
 Overall, we found a negative effect of reducing character correlations on recovering the correct topology.
+%ap- will you confront the fact that this says that having character correlation is an advantage?
 However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing or increasing character correlation is not more effective in getting the expected topology than randomly drawing characters.
 Furthermore, in bigger matrices (350 characters), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
+%ap but if cc increase recovering correct topology, then Bayesian is bad and we should use parsimony.
+Needs clarifying.
 These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
@@
 %(i.e. ``concentrated'' correlations)
-but weaker and more ``spread-out'' correlations in molecular data \citep[][]{huelsenbeck1999effect}.
+but weaker and more ``spread-out'' in molecular data \citep[][]{huelsenbeck1999effect}.
 %i.e.
@@
 % Before discussing character correlation further,
-Here we distinguish three different modes of forms of character correlation:
+Here we distinguish three different modes of character correlation:
 
@@
     Intra-organismal links can be the targets of comparative developmental biology \citep{goswami2010,kelly2010reduced,stoessel2013morphological,goswami2014macroevolutionary} or functional investigations.
+%AP - I would drop this sentence and say whether this type of correlation is useful or not in less uncertain terms.
 
@@
     Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
-    However, these methods do not give us a means to objectively control correlations that might adversely affect phylogenetic inference.
+    However, these methods do not give us a means to objectively control correlations that might adversely affect phylogenetic inference. %AP this is good.
 
@@
     % For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
-    This coding dependency is linked to hierarchical dependency between characters \citep{wilkinson1995coping,BrazeauNA}.
+    This coding dependency is linked to hierarchical dependency between characters \citep{wilkinson1995coping,BrazeauNA}. %AP-Can cut this sentence
     % In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.
@@
 
-\noindent Of course, the three sources of dependence have an interaction: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding.
-Logical dependence, however, is easily distinguished prior to phylogenetic inference, while the two other ones (intra-organismal and evolutionary) are much harder.
-However, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{de2015parsimony,BrazeauNA}.
-Intra-organismal dependence requires more detailed, often extremely time-consuming studies (and possibly beyond the limits of available technology).
+\noindent Of course, the three sources of dependence can have an interaction: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding.
+Logical dependence, however, is easily distinguished prior to phylogenetic inference, while intra-organismal and evolutionary dependence are much harder to parse.
+%AP-what is 'logical dependece'?
+its not defined...
+I guess you meaning "coding'...
+why is it easy to distinguish?
+Meanwhile, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{de2015parsimony,BrazeauNA}.
+Intra-organismal dependence requires more detailed, often extremely time-consuming studies (possibly beyond the limits of available technology).
 % Even after all of the effort is expended, the results might then only be known for a single (model) species.
@@
 However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inference software that are blind to these distinctions.
-If fact, what the software are confronted with is a two dimensional matrix problem that renders the morphological subtleties described opaque.
+If fact, what the software are confronted with is a two dimensional matrix problem that renders all morphological dependencies invisible.
 This introduces a new, less studied, source of character dependence:
 
-\noindent \textbf{Characters compatibility}: this is the result of how software actually interprets the differences between characters.
-The majority of phylogenetic algorithms cannot really interpret character descriptions and state meanings (i.e.
+\noindent \textbf{Character compatibility}: this is the result of how software actually interprets the differences between characters.
+The majority of phylogenetic algorithms cannot interpret character descriptions and state meanings (i.e.
 character ontologies), and instead must interpret discrete tokens as signifiers of shared similarity.
 Therefore, phylogenetic character matrices that contain a great number of characters but a few number of tokens can easily lead to correlation between characters.
-For example, if we consider the following matrix containing four cetartiodactyls - say a pig (\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
+%AP I do not understand this definition.
+The example did not clear it up.
+What are tokens?
+For example, if we consider the following matrix containing four cetartiodactyls - say a pig (\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus (ankle bone); \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
 
@@
     \caption{Example of a matrix with software induced character correlation.
-\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley.}
+\textbf{C1}: presence (1) or absence (0) of an astragalus (ankle bone); \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley.}
     \label{Tab:example_matrix}
@@
 Alternatively, partitioning strategies have been used to reduce this problem \citep[see][and references within]{Rosa2019}.
-This illustrates how some characters will therefore be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
+This example illustrates how some characters will therefore be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
+%AP-what is "covary in a non-phlyogenetic way'??
+this section-I don't get it.
+
+%AP it also kind of bothers me that you said there were 3 ways to have CC- but wait no, actually 4. But then below is 3 again... confusing
 
 How do these specific correlation really affect topology?
-Although it has been tackled empirically (for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}), the question has never been explored through a thorough simulation framework.
+Although it has been tackled empirically (for morphological data \citealt{Davalos01072014}, and molecular data \citealt{ZouConvergence}), the question has never been explored through a thorough simulation framework.
+%AP I hope we come back to these studies in the discussion.
 % In the specific case of discrete morphological characters, how does the reduc phenotypic information and its different sources of correlation is 
@@
 % \cite{parins2017use}
-Here we formally assess the effect of discrete character's correlation using simulated data.
+Here we formally assess the effect of discrete characters' correlation using simulated data.
 We propose % a new
-distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
+using the Hamming distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
 We found that overall, there is a detectable effect of character correlation on topology where a decrease in character dependence results in a decrease in the ability to recover the correct topology.
@@
 \section{Methods}
-To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different levels of correlation between some characters (Fig.\ref{Fig:outline} - note that each step is described in more details below):
+To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different levels of correlation between some characters (Fig.\ref{Fig:outline} - note that each step is described in more details below): %AP different average correlation between characters?
 \begin{enumerate}
-    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
+    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 or 150 taxa, for 100, 350 or 1000 characters, hereafter called the ``normal'' matrices.
 This step resulted in 9 matrices.
@@
 $CD$ is a continuous scaled Hamming distance metric bounded between $0$ and $1$.
-Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way %.
+Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states as binary %.
 % Two same character states tokens have a difference of zero and two different ones have a difference of one
-(e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
+(e.g. $2 - 2 = 0$ or $1 - 8 = 1$).
 % Additionally, we only consider differences for taxa with shared information \citep{GowerDist}.
@@
 When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated). 
-When the character difference is maximal ($1$) it means it conveys the greatest difference in phylogenetic signal. % (i.e. characters are uncorrelated).
-However, a character difference of $0$ %(i.e. the same phylogenetic signal)
-does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
+When the character difference is maximal ($1$) it means the characters convey the greatest difference in phylogenetic signal. % (i.e. characters are uncorrelated).
+%However, a character difference of $0$ %(i.e. the same phylogenetic signal)
+%does not mean the opposite of $1$ (i.e.
+\textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
+%AP this sentence was awkward or ungrammatical or incorrect, not sure which.
+Don't talk about "opposites" it doesnt mean anything.
 % For example with three characters \texttt{A = \{0,1,2,1\}}, \texttt{B = \{2,0,1,0\}} and \texttt{C = \{0,0,1,2\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$ (translated into \texttt{A' = \{1,2,3,2\}} \texttt{B' = \{1,2,3,2\}} and \texttt{C' = \{1,1,2,3\}}; $CD_{(A,B)} = (0+0+0+0)/(4-1) = 0$ and $CD_{(A,C)} = (0+1+1+1)/(4-1) = 1$).
@@
 To simulate the matrices we applied a protocol very similar to \cite{Guillerme2016146} using \texttt{R} \citep{R}.
-First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution with $\lambda$ $>$ $\mu$\citep[\texttt{diversitree} v0.9-8;][]{fitzjohndiversitree2012} and saving the trees after reaching either 25, 75 or 150 taxa (see supplementary material 3 Figs 1, 2 and 3).
+First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution with $\lambda$ $>$ $\mu$\citep[\texttt{diversitree} v0.9-8;][]{fitzjohndiversitree2012} and saving the trees after reaching either 25, 75 or 150 taxa (see supplementary material 3, Figs 1, 2 and 3).
 For each tree, we arbitrarily set the outgroup to be the first taxon (alphabetically) thus effectively rooting the trees on this taxon.
@@
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
-    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
+    We drew the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
     % This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g.
@@
 \section{Discussion}
+%AP- ok so im understanding that its important to measure CD because it tells you wether you are inflating your odds of getting a well resolved tree.
+If you find low CD it means you should check your characters to see if you are committing coding dependency or if your traits are evolutionarily dependent then its probably an honest phylogenetic signal.
+I'm not sure how Intra-organismal fits in though.
+So basically the message of this paper should be "Always check your characters for correllation using this sweet metric-and if its high -check to see what kind of correlation you have"(This should go in the abstract as well) It'd then be nice to 'prescribe' methods for checking which kind of correlation you have.
+Ok, Now I'm going to read the discussion and see if you did this.
 
-\subsection{Effect of character differences on topology}
-As expected, there is a significant effect of character difference in the ability to recover the correct topology (in terms of clades).
+\subsection{Effect of character differences on topology}%AP-the first sentence of the discussion should refer back to the title in my opinion.
+Ie.
+We find that different modes of character correlation have a strong effect on the ability to recover the correct phylogenetic tree.
+As expected, there is a significant effect of character difference (the inverse of character correlation) on the ability to recover the correct topology (in terms of clades).
 % The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
-When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
+When characters are correlated, matrices convey a strong (but potentially misleading) phylogenetic signal since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
+%AP do you have to use 'homoplasy' here?
+I had to look it up (for the billionth time) understand that it means independent trait origins- but dont understand how homoplasy would be related to weaker signal.
+Nor why a maximized tree would have higher homoplasy.
 % Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (i.e.
@@
 This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.
+%AP this paragraph is so dry and boring, refer to the supplements??-get out of the weeds!  this is just rehashing results with no interpretation.  I suggest rewriting completely.
 
@@
 Overall, it is expected that minimising character difference %(increasing character correlation)
- is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
+ is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal. %ap- I would change "correct" to "more resolved"
 If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good.
-Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.
-
-Although we measured a significant difference between smaller and larger matrices (Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), %in terms of number of characters, between 100 and 1000 characters;
+Similarly, if the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree, then the estimates will be poor.
+%AP-The above paragraph isnt about characters or taxa, however it would be good for the first paragraph of the discussion.
+While matrices with more characters or taxa tended to be better resolved than matrices with fewer characters or taxa, the effect of character correlation on topology is independent of the size of the matrix.
+(Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), %in terms of number of characters, between 100 and 1000 characters;
 these differences have to be contrasted with the Bhattacharrya Coefficients that are always very high between every matrix's size.
+%AP- the Bhattacharrya stuff doesnt make sense, do you mean BC is always high between matrices of different sizes? Why does it matter
 % This suggests that the overall character difference metric can be used as a proxy for character correlation as it is not directly dependent on the number of characters analysed.
-Regarding the number of taxa, our results suggest the effect of character correlation on topology is independent on the size of the matrix when pooling the data.
+Regarding the number of taxa, our results suggest the effect of character correlation on topology is independent on the size of the matrix when pooling the data. %ap Cut-as I summarized above
 Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered.% (i.e.
 we corrected for the expected difference when comparing two random trees with the same number of taxa).
+%ap-This seems like weeds
 
@@
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
+%AP-Solid paragraph
 
@@
 These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not be used \textit{a priori} as a criterion to select characters.
+%AP-It'd be great to talk here about the problem with a priori selection creating circular and biased reasoning in a bit of depth.
 % In other words, intra-organismal and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
@@
 However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices \citep{Lande1983,Maddison1990,Pagel1994}.
+%AP We should probably change the title of the paper then...
 %(i.e. inferring the topology first and looking at character's correlation based on the observed topology \citealt{Lande1983,Maddison1990,Pagel1994}).
@@
 However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree.
-This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to each other (the outputs) rather than to the simulation inputs.
+This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to the other outputs rather than to the simulation inputs.
 % It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
@@
 
-Second, measuring and modifying character correlation is difficult. 
-In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 4, 5 and 6).
+Second, modifying character correlation is difficult.
+%AP- the whole point is to start using CD to measure it...
+you could maybe drop this paragraph and subsume the literature discussion into the discussion I suggested above about the 3 types of dependency and the potential for circular logic.
+But maybe add a few lines on what you believe is state of the art for dealing with it, and then tie it up with a CHARGE to find betters ways to deal with it like in your monotreme study
+In our simulation protocol we chose to create simulations by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 4, 5 and 6).
 In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
 This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
-However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
+However, because of the number of characters, it is currently too complex to attribute correlation to each kind of biological  and coding dependence and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
 % Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e.
@@
 
-\subsection{Potential applications}
+\subsection{Potential applications} %AP -yup  nice- maybe move above of caveats
 Effectively, our simulation protocol bootstraps our data ``with bias''.
@@
 \subsection{Conclusion}
-Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon.% to investigate phylogenetic relationships.
+Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon.
+% to investigate phylogenetic relationships.
 Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g.
@@
 the expansion of dataset size \citep[e.g.][> 1000 characters]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
-Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate between these correlation types, it is important to understand to what extant topologies can be induced by such bias.
+Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate between these correlation types, it is important to understand to what extent topologies can be induced by such bias.
 
@@
 However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise or maximised character correlation was not more significant than simply randomising (i.e.
-bootstrapping) the matrix, suggesting that character correlation is more a problem in large discrete morphological matrices.
+bootstrapping) the matrix, suggesting that character correlation is more of a problem in large discrete morphological matrices.
+We recommend that all morphological phylogenetic analyses should report the level of character difference so that we can begin to detect and address the problem of coding dependence on inflating our confidence in our phylogenetic reconstructions.
+%Or some similar charge
 % For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly increased and decreased correct topological recovery compared to randomly resample matrices.

### 2019-11-01 Writing/Palaeontology/draft.tex
Martin comments

--- 2019-10-23 Writing/Palaeontology/draft.tex
@@
 Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
-Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
-In this study, we propose a new metric of raw character difference as a proxy for character correlation.
+Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be diminished \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
+In this study, use a metric of raw character difference as a proxy for character correlation.
 Using thorough simulations, we test the effect of increasing or decreasing character differences on tree topology.
-Overall, we found a negative effect of reducing character correlations on recovering the correct topology.
+Overall, we found a negative effect of reducing character correlations on recovering the correct topology
+This means that dataset with low correlation between characters will make it more complicated to estimate a correct topology.
 However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing or increasing character correlation is not more effective in getting the expected topology than randomly drawing characters.
-Furthermore, in bigger matrices (350 characters), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
+Furthermore, in bigger matrices (350 characters), the inference method has a strong effect with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
 These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
@@
 This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
-Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collecting them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.
+Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collecting them and their sometimes subjectivity, they are usually the only source of phylogenetic data for extinct species.
 While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
@@
 This opens up possibilities for predicting problematic datasets and possibly proposing new confidence measures in phylogenetic datasets.
+%MB: I think a light rewriting  of this opening paragraph would be good, focusing it more on a palaeontological audience (given the new target journal).
+We can cut the apologetics for morphology and focus it on palaeontology-relevant stuff  a bit more.
 
@@
 However, several distinctions can be made between the amount and the nature of the correlation.
-First, molecular and morphological data often differs in size and in the number of states analysed.
+First, molecular and morphological data often differ in the size and number of states analysed.
  % (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but may have an arbitrary number of states (up to that supported by phylogenetic software; \citealt{Guillerme2016146}).
@@
 Here we distinguish three different modes of forms of character correlation:
+% MB: These considerations are interesting, and "modes" are also referred to in the title.
+The discussion does come back to these, but not sure it adds much.
+It could be an opportunity to remove almost a page and keep the overall flow more focused.
+Indeed, this is almost all said before (a bit more succinctly) by Mark Wilkinson.
 
@@
 \noindent \textbf{Characters compatibility}: this is the result of how software actually interprets the differences between characters.
+% MB: Well, software can maximise it, but compatibility itself is just a property of the data.
 The majority of phylogenetic algorithms cannot really interpret character descriptions and state meanings (i.e.
@@
 Therefore, phylogenetic character matrices that contain a great number of characters but a few number of tokens can easily lead to correlation between characters.
+% MB: I am having trouble interpreting this.
 For example, if we consider the following matrix containing four cetartiodactyls - say a pig (\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
@@
 This illustrates how some characters will therefore be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
+% MB: Is all of this necessary?
 
@@
 Although it has been tackled empirically (for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}), the question has never been explored through a thorough simulation framework.
+% MB: You may wish to add one line to say what the fundamental importance of doing so is. What will this paper add that those did not?
 % In the specific case of discrete morphological characters, how does the reduc phenotypic information and its different sources of correlation is 
@@
 However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
-increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario led to slightly worst trees (Fig.
+increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario led to slightly worse trees (Fig.
 \ref{Fig:RF_results_best}).
-This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
+This effect of increasing or reducing character difference did not appear to affect taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
 However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
-Furthermore, fewer characters were replaced, on average, in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6)% - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+Furthermore, fewer characters were replaced, on average, in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6).% - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
 % However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 4, 5 and 6).
-This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.
+This suggest that character difference can be decreased with a smaller number of characters replacements than increasing it.
 
@@
 
-The two other sources of character correlation are present in our simulations although they were not explicitly modelled:
+The two other sources 
+%MB: These are the same types of correlation you mention in the preceding paragraph.
+ of character correlation are present in our simulations although they were not explicitly modelled:
 (1) evolutionary correlation is implied by simulating the characters using Birth-Death trees;
@@
 First, simulating evolutionary history is complex.
-Not only because the models we're using to infer phylogenies are ever improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible \citep[see constrasted discussions from][]{GoloboffEmpirical,OReillyEmpirical}.
+Not only because the models we're using to infer phylogenies are ever-improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible \citep[see constrasted discussions from][]{GoloboffEmpirical,OReillyEmpirical}.
 However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree.
@@
 However, in the ``minimised'' and ``maximised'' scenario, we remove the characters with the lowest/highest overall character difference.
-For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other and randomly resample from the characters left.
+For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other
+% MB: I think a word is missing.
+ and randomly resample from the characters left.
 
@@
 Applying this kind of bootstraps that minimised character difference by biasing the random sampling could thus help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.
+%MB: Rewrite.
+Avoid informal constructions like "kind of...".
+As a side note: "kind of..." is a harmless thing to say, but after some things it can mean everything, like after "you're going to live" or "It's a boy!".
 
@@
 the expansion of dataset size \citep[e.g.][> 1000 characters]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
-Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate between these correlation types, it is important to understand to what extant topologies can be induced by such bias.
+Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate between these correlation types, it is important to understand to what extent topologies can be induced %MB: Not sure this is word we want.
+ by such bias.
 
 We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was low (high character differences), the topology was always the furthest away from the ``normal'' topology.
-Conversely, when correlation between characters was high, the topology was always the closest to the ``normal'' topology.
+Conversely, when correlation between characters was high, the topology was always the closest %MB: closer than what?
+ to the ``normal'' topology.
 These results appear independent of the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
@@
 % For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly increased and decreased correct topological recovery compared to randomly resample matrices.
+% MB: I think the conclusion could be briefer: focus more on the broad-level implications and take it back to what this means for palaeontologists.
 

### 2019-11-01 Writing/Palaeontology/draft.tex

--- 2019-10-26 Writing/Palaeontology/draft.tex
@@
 Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
-Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
-In this study, we propose a new metric of raw character difference as a proxy for character correlation.
+Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be diminished \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
+In this study, use a metric of raw character difference as a proxy for character correlation.
 Using thorough simulations, we test the effect of increasing or decreasing character differences on tree topology.
@@
 %ap- will you confront the fact that this says that having character correlation is an advantage?
+This means that dataset with low correlation between characters will make it more complicated to estimate a correct topology.
 However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing or increasing character correlation is not more effective in getting the expected topology than randomly drawing characters.
-Furthermore, in bigger matrices (350 characters), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
-%ap but if cc increase recovering correct topology, then Bayesian is bad and we should use parsimony.
-Needs clarifying.
+Furthermore, in bigger matrices (350 characters), the inference method has a strong effect with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
+%ap but if cc increase recovering correct topology, then Bayesian is bad and we should use parsimony. Needs clarifying.
 These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
@@
 % Before discussing character correlation further,
-Here we distinguish three different modes of character correlation:
+Here we distinguish three different modes of forms of character correlation:
+% MB: These considerations are interesting, and "modes" are also referred to in the title.
+The discussion does come back to these, but not sure it adds much.
+It could be an opportunity to remove almost a page and keep the overall flow more focused.
+Indeed, this is almost all said before (a bit more succinctly) by Mark Wilkinson.
 
@@
 
-\noindent \textbf{Character compatibility}: this is the result of how software actually interprets the differences between characters.
-The majority of phylogenetic algorithms cannot interpret character descriptions and state meanings (i.e.
+\noindent \textbf{Characters compatibility}: this is the result of how software actually interprets the differences between characters.
+% MB: Well, software can maximise it, but compatibility itself is just a property of the data.
+The majority of phylogenetic algorithms cannot really interpret character descriptions and state meanings (i.e.
 character ontologies), and instead must interpret discrete tokens as signifiers of shared similarity.
@@
 What are tokens?
-For example, if we consider the following matrix containing four cetartiodactyls - say a pig (\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus (ankle bone); \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
+% MB: I am having trouble interpreting this.
+For example, if we consider the following matrix containing four cetartiodactyls - say a pig (\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
 
@@
 this section-I don't get it.
-
-%AP it also kind of bothers me that you said there were 3 ways to have CC- but wait no, actually 4. But then below is 3 again... confusing
+% MB: Is all of this necessary?
 
@@
 %AP I hope we come back to these studies in the discussion.
+% MB: You may wish to add one line to say what the fundamental importance of doing so is. What will this paper add that those did not?
 % In the specific case of discrete morphological characters, how does the reduc phenotypic information and its different sources of correlation is 
@@
 However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
-increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario led to slightly worst trees (Fig.
+increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario led to slightly worse trees (Fig.
 \ref{Fig:RF_results_best}).
-This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
+This effect of increasing or reducing character difference did not appear to affect taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
 However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
-Furthermore, fewer characters were replaced, on average, in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6)% - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
+Furthermore, fewer characters were replaced, on average, in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6).% - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
 % However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 4, 5 and 6).
@@
 the expansion of dataset size \citep[e.g.][> 1000 characters]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
-Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate between these correlation types, it is important to understand to what extent topologies can be induced by such bias.
+Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate between these correlation types, it is important to understand to what extent topologies can be induced %MB: Not sure this is word we want.
+ by such bias.
 
 We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was low (high character differences), the topology was always the furthest away from the ``normal'' topology.
-Conversely, when correlation between characters was high, the topology was always the closest to the ``normal'' topology.
+Conversely, when correlation between characters was high, the topology was always the closest %MB: closer than what?
+ to the ``normal'' topology.
 These results appear independent of the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.

### 2019-11-01 Writing/Palaeontology/draft.tex
Updated draft

@@
 \begin{center}
-\noindent{\Large \bf Influence of different modes of morphological character correlation on phylogenetic tree inference}
+\noindent{\Large \bf Influence of morphological character correlation on phylogenetic tree inference}
 \bigskip
@@
 Overall, we found a negative effect of reducing character correlations on recovering the correct topology.
-%ap- will you confront the fact that this says that having character correlation is an advantage?
 This means that dataset with low correlation between characters will make it more complicated to estimate a correct topology.
@@
 Furthermore, in bigger matrices (350 characters), the inference method has a strong effect with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
-%ap but if cc increase recovering correct topology, then Bayesian is bad and we should use parsimony. Needs clarifying.
 These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
@@
 The last two decades have witnessed a resurgence of interest in the use of morphological character data in phylogenetic studies.
-This owes in large part to the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
-Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collecting them and their sometimes subjectivity, they are usually the only source of phylogenetic data for extinct species.
-While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any empirical investigations into their statistical properties.
-As morphological data are likely to continue to play an extensive role in phylogenetic analysis, it is essential to understand the circumstances under which morphological data might be expected to ``misbehave''.
-This opens up possibilities for predicting problematic datasets and possibly proposing new confidence measures in phylogenetic datasets.
-%MB: I think a light rewriting  of this opening paragraph would be good, focusing it more on a palaeontological audience (given the new target journal).
-We can cut the apologetics for morphology and focus it on palaeontology-relevant stuff  a bit more.
+This owes in large part to the advent of new models, methods and software \citep{lewisa2001,ronquista2012,Ronquist2012mrbayes,heath2014fossilized} and the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
+While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any investigations into their statistical properties.
+As morphological data are likely to continue to play an extensive role in phylogenetic analysis (namely for neontological analysis), it is essential to understand the circumstances under which morphological data might be expected to lead to erroneous phylogenetic inferences.
+This opens up possibilities for analysing problematic datasets and possibly proposing new confidence measures in phylogenetic datasets.
 
@@
 The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-% This independence is assumed for both molecular and morphological characters \citep{huelsenbeck1999effect,Davalos01072014,ZouConvergence}.
 However, several distinctions can be made between the amount and the nature of the correlation.
 First, molecular and morphological data often differ in the size and number of states analysed.
- % (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but may have an arbitrary number of states (up to that supported by phylogenetic software; \citealt{Guillerme2016146}).
-Second, correlations have been proposed to be stronger in morphological data
-%(i.e. ``concentrated'' correlations)
-but weaker and more ``spread-out'' in molecular data \citep[][]{huelsenbeck1999effect}.
-%i.e.
-``diffused'' correlations - though some regions of the genome may be thought as more correlated than others
-% Furthermore, because of the ``spread'' of correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
-changes in morphological states can be more readily linked to a function \textit{cf.} changes in nucleotide states ; see below).
-% Before discussing character correlation further,
-Here we distinguish three different modes of forms of character correlation:
-% MB: These considerations are interesting, and "modes" are also referred to in the title.
-The discussion does come back to these, but not sure it adds much.
-It could be an opportunity to remove almost a page and keep the overall flow more focused.
-Indeed, this is almost all said before (a bit more succinctly) by Mark Wilkinson.
+Second, correlations have been proposed to be stronger in morphological data but weaker and more ``spread-out'' in molecular data \citep[][]{huelsenbeck1999effect}.
+Here we distinguish three different modes of forms of character correlation (described in more details in \citealt{wilkison1992ordered,wilkinson1995character,wilkinson1995coping}):
 
 \begin{itemize}
-    \item \textbf{Intra-organismal dependence:} this is the result of an intrinsic biological link between two characters through development, pleiotropy, and/or biological function.
-    For example the lower and upper molar characters in mammals generally occlude one another. 
-    Therefore, one character describing a feature of a lower molar will be expected to be complemented by the surface of the occluding upper molar.
-    Characters of the occlusal surface of two opposing molars will be expected to directly covary.
-    % Pleiotropy also results in covariation between different aspects of phenotype and may thus induce correlations. 
-    % From a phylogenetic perspective, it can be especially pernicious because the relationship between the traits in question may have no obvious link from a morphological or functional comparison alone.
-    Intra-organismal links can be the targets of comparative developmental biology \citep{goswami2010,kelly2010reduced,stoessel2013morphological,goswami2014macroevolutionary} or functional investigations.
-%AP - I would drop this sentence and say whether this type of correlation is useful or not in less uncertain terms.
+    \item \textbf{Intra-organismal dependence:} this is the result of an intrinsic biological link between two characters.
+    For example the occlusal surface of two opposing molars will be expected to directly covary.
 
-    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving due to selection, likely related to functional links between two traits that help serve an overall lifestyle trait \citep{ClarkeMiddleton2008}.
-    Unlike the case of intra-organismal dependence, there need not be an intrinsic constraint that causes these traits to covary.
+    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving. 
     For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
-    % This is thought to correspond to adaptations for fossoriality or aquatic lifestyles.
-    % Such covariances are generally studied in the context of a given phylogeny, often one derived from molecular data with the morphological traits of interest mapped on it.
-    Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
-    However, these methods do not give us a means to objectively control correlations that might adversely affect phylogenetic inference. %AP this is good.
 
     \item \textbf{Coding dependence:} this is the results of researcher methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
-    Coding dependence manifests itself in several ways, particularly in coding redundant information.
-    For instance, coding for the same absence in different characters creates state transformations associated with the loss or gain of a particular character.
-    This occurs when a number of multistate characters include two variable feature states (e.g. large, small; red, blue etc.) in conjunction with absence.
-    For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.
-    % It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
-    % For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
-    This coding dependency is linked to hierarchical dependency between characters \citep{wilkinson1995coping,BrazeauNA}. %AP-Can cut this sentence
-    % In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.
-    % Finally this can also be due to a bias in the amount of characters available.
-    % For example, in skulls, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity to discrete characters).
+    For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail \citep{wilkinson1995coping,BrazeauNA}.
 \end{itemize}
@@
 \noindent Of course, the three sources of dependence can have an interaction: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding.
-Logical dependence, however, is easily distinguished prior to phylogenetic inference, while intra-organismal and evolutionary dependence are much harder to parse.
-%AP-what is 'logical dependece'?
-its not defined...
-I guess you meaning "coding'...
-why is it easy to distinguish?
+Coding dependence, however, is easily distinguished prior to phylogenetic inference, while intra-organismal and evolutionary dependence are much harder to parse.
 Meanwhile, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{de2015parsimony,BrazeauNA}.
 Intra-organismal dependence requires more detailed, often extremely time-consuming studies (possibly beyond the limits of available technology).
-% Even after all of the effort is expended, the results might then only be known for a single (model) species.
 Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree, and is best determined by independent character sets.
-% This is frequently accomplished by mapping morphological traits on molecular phylogenetic trees.
 
-% These sources of dependence between characters are well studied in biology.
 Biological and evolutionary dependences are inherent parts to Evo-Devo and macroevolutionary studies and best practices to avoid coding-induced dependences are commonly known and applied.
@@
 
-\noindent \textbf{Characters compatibility}: this is the result of how software actually interprets the differences between characters.
-% MB: Well, software can maximise it, but compatibility itself is just a property of the data.
-The majority of phylogenetic algorithms cannot really interpret character descriptions and state meanings (i.e.
-character ontologies), and instead must interpret discrete tokens as signifiers of shared similarity.
-Therefore, phylogenetic character matrices that contain a great number of characters but a few number of tokens can easily lead to correlation between characters.
-%AP I do not understand this definition.
-The example did not clear it up.
-What are tokens?
-% MB: I am having trouble interpreting this.
-For example, if we consider the following matrix containing four cetartiodactyls - say a pig (\textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four binary characters - say (\textbf{C1}: presence (1) or absence (0) of an astragalus; \textbf{C2}: presence (0) or absence (1) of baleen; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragalus with a double pulley.
+\begin{itemize}
+    \item \textbf{Characters compatibility}: this is the result of how software actually interprets the differences between characters.
+    The majority of phylogenetic algorithms cannot interpret character descriptions and state meanings, and instead must interpret discrete states as signifiers of shared similarity: for each characters, species are grouped if they have identical characters.
+    Therefore, character with the same number of states will have a chance of being correlated regardless of mode of form of character correlation.
+\end{itemize}
 
-\begin{table}
-\center
-    \begin{tabular}{r|cccc}
-            & C1 & C2 & C3 & C4\\
-        \hline
-        \textit{Sus} & 1 & 1 & 1 & 1\\
-        \textit{Cervus} & 1 & 1 & 1 & 1\\
-        \textit{Hippopotamus} & 1 & 1 & 1 & 1\\
-        \textit{Balaenoptera} & 0 & 0 & 0 & 0\\
-    \end{tabular}
-    \caption{Example of a matrix with software induced character correlation.
-\textbf{C1}: presence (1) or absence (0) of an astragalus (ankle bone); \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragalus with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley.}
-    \label{Tab:example_matrix}
-\end{table}
-
-In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
-Yet a phylogenetic software will treat all these four characters in exactly the same way.
-% In fact, for practical reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
-In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} that seem to be more likely to reflect the split by separating artiodactyles from the cetaceans.
-% In other words, this is the correlation between characters that depicts the evolutionary history of this group.
-However, the other characters depict the same history %(i.e. the same split)
- but for the wrong reasons. % (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleen, the same split would still be estimated by the software).
-One solution would be to a) weight some characters (e.g.
-giving \textbf{C3} and \textbf{C4} a lower weight) and b) make some characters dependent on each other (e.g.
-making \textbf{C3} and \textbf{C4} dependent on \textbf{C1}).
-However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.
-Alternatively, partitioning strategies have been used to reduce this problem \citep[see][and references within]{Rosa2019}.
-This example illustrates how some characters will therefore be expected to covary in a non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
-%AP-what is "covary in a non-phlyogenetic way'??
-this section-I don't get it.
-% MB: Is all of this necessary?
-
-How do these specific correlation really affect topology?
+\noindent How does character correlation really affect topology?
 Although it has been tackled empirically (for morphological data \citealt{Davalos01072014}, and molecular data \citealt{ZouConvergence}), the question has never been explored through a thorough simulation framework.
-%AP I hope we come back to these studies in the discussion.
-% MB: You may wish to add one line to say what the fundamental importance of doing so is. What will this paper add that those did not?
-% In the specific case of discrete morphological characters, how does the reduc phenotypic information and its different sources of correlation is 
-% More specifically, how can phylogenetic inference software distinguish between the complex sources of correlation tha
-% In other words, how can we distinguish between the complex sources of correlation in phenotype using only discrete morphological characters?
-% \cite{parins2017use}
-Here we formally assess the effect of discrete characters' correlation using simulated data.
-We propose % a new
-using the Hamming distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
-We found that overall, there is a detectable effect of character correlation on topology where a decrease in character dependence results in a decrease in the ability to recover the correct topology.
+Simulations can allow the design of dataset with controlled levels of correlation and for which the correct expected answer is known by design.
+Furthermore, using simulations will allow to draw generalities on the effect of character correlation on phylogenetic inference conversely to potential idiosyncrasies of empirical datasets.
+Here we use a Hamming distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
+We found that overall, there is a clear effect of character correlation on topology where a decrease in character correlation results in a decrease in the ability to recover the correct topology.
 These results, however, vary greatly in magnitude depending on the size of matrix and the inference method used.
@@
 \section{Methods}
-To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different levels of correlation between some characters (Fig.\ref{Fig:outline} - note that each step is described in more details below): %AP different average correlation between characters?
+To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different average correlation between characters (Fig.\ref{Fig:outline} - note that each step is described in more details below):
 \begin{enumerate}
@@
 To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information.
-%(i.e. entirely correlated characters would define identical splits between a set of taxa).
 In order to measure this, we used the following scaled Hamming distance between two characters:
@@
 \end{equation}
-%CD is similar to https://onlinelibrary.wiley.com/doi/full/10.1111/pala.12384 but by character rather than by species
 
@@
 $CD$ is a continuous scaled Hamming distance metric bounded between $0$ and $1$.
-Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states as binary %.
-% Two same character states tokens have a difference of zero and two different ones have a difference of one
-(e.g. $2 - 2 = 0$ or $1 - 8 = 1$).
-% Additionally, we only consider differences for taxa with shared information \citep{GowerDist}.
-% For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
-% Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).
+Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states as binary (e.g. $2 - 2 = 0$ or $1 - 8 = 1$).
 
@@
 This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
-% Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).
 When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated). 
-When the character difference is maximal ($1$) it means the characters convey the greatest difference in phylogenetic signal. % (i.e. characters are uncorrelated).
-%However, a character difference of $0$ %(i.e. the same phylogenetic signal)
-%does not mean the opposite of $1$ (i.e.
-\textit{not} the opposite phylogenetic signal but the most different number of implied splits) .
-%AP this sentence was awkward or ungrammatical or incorrect, not sure which.
-Don't talk about "opposites" it doesnt mean anything.
-% For example with three characters \texttt{A = \{0,1,2,1\}}, \texttt{B = \{2,0,1,0\}} and \texttt{C = \{0,0,1,2\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$ (translated into \texttt{A' = \{1,2,3,2\}} \texttt{B' = \{1,2,3,2\}} and \texttt{C' = \{1,1,2,3\}}; $CD_{(A,B)} = (0+0+0+0)/(4-1) = 0$ and $CD_{(A,C)} = (0+1+1+1)/(4-1) = 1$).
-% Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different set of splits (i.e.
-a different phylogenetic signal).
+When the character difference is maximal ($1$) it means the characters convey the greatest difference in phylogenetic signal
 
@@
     \item The ``morphological HKY-binary'' model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
-    % This model has the advantage of not favouring Bayesian inference \citep[; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
     We drew the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
-    % This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g.
-Dollo characters;][]{Dollo,wright2015came}.
-    % However, \cite{Wright01072016} have shown that equal transition rate is a reasonable assumption for many empirical datasets.
 \end{itemize}
@@
 To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{sanderson1989patterns,OReilly20160081}.
-% Note that both models (``morphological HKY-binary' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
-% However, the ``morphological HKY model'' is not the one \textit{directly} estimated in the phylogenetic software used in this protocol (MrBayes, see below).
 
@@
 The matrices were generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-% To estimate the variance of our simulations and assess the effect of our random parameters,
 We repeated this step 20 times resulting in 180 ``normal'' morphological matrices.
@@
 Finally, because this operation effectively changes the weight of characters (i.e.
-giving the characters $<0.25$ or $>0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification.% (i.e.
-randomly weighting characters).
+giving the characters $<0.25$ or $>0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification.
 Each replaced characters were randomly selected by sampling them randomly from the list of the remaining characters.
-% Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
 This step resulted in a total of 720 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
@@
 The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates.
- % (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).
 
-Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$)
- % - \texttt{lset nst=1 rates=gamma Ngammacat=4})
-with an variable rate prior an exponential (0.5) shape.
- % (\texttt{prset ratepr=variable Shapepr=Exponential(0.5)}).
+Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$) with an variable rate prior an exponential (0.5) shape.
 The MCMC was ran over two runs of 6 chains each for a maximum of $1\times10^9$ generations with a sampling every 200 generations with an automatic stop if the average standard deviation of split frequency (ASDSF) fell below 0.01.
-% \texttt{mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES}).
 Due to cluster hardware requirements and to save some time, when chains didn't converge and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the unconverged chains.
@@
 Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
-% When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
-% However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 15, 16 and 17.
 
@@
 The measurements were then standardised using the Normalised Tree Similarity metric \citep[$NTS$; centering and scaling the metric score to the number of taxa;][]{Bogdanowicz2012,Guillerme2016146}.
-% When the normalised metric has a score of one it means both trees are identical, when it has a score of zero it means the trees are no more different than expected by chance and when it has a score $<0$ the trees are more different than expected by chance.
-The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Foulds Similarity ($NTS_{RF}$)  reflects the conservation of clades
- % (i.e. a score close to $1$ indicates that most clades are identical in both trees)
-; and (2) the Normalised Triplets Similarity ($NTS_{Tr}$) reflects the position of taxa.
-% (i.e. a score close to $1$ indicates that most taxa have the same neighbours in both trees).
+The normalised score for both metrics thus reflects two distinct aspects of tree topology: (1) the Normalised Robinson-Foulds Similarity ($NTS_{RF}$)  reflects the conservation of clades; and (2) the Normalised Triplets Similarity ($NTS_{Tr}$) reflects the position of taxa.
 
-% The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}).
-% Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its \textit{p}-value calculation due to the presence of equal values in the $NTS$ distributions (e.g.
-when multiple trees are equal to the ``normal'' tree).
 To measure the effect of character correlation, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep{holm1979simple} and a the probability of overlap between two distributions \citep[the Bhattacharyya Coefficient: $BC$;][]{Bhattacharyya}.
-% Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $<0.05$.
-% Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g.
-for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).
 The resulting full simulation was 2TB in size so is not shared here (though the parameters are).
@@
 The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information followed by randomising and maximising the character difference (see supplementary materials 3, Tables 3, 4 and 5).
-% (median $NTS_{RF}$ = 0.869 and median $NTS_{Tr}$ = 0.787; see supplementary materials 3, Table 3) followed by randomising the characters (median $NTS_{RF}$ = 0.814 and median $NTS_{Tr}$ = 0.791; see supplementary materials 3, Table 2) and maximising the character difference (median $NTS_{RF}$ = 0.698 and median $NTS_{Tr}$ = 0.754; see supplementary materials 3, Table 2).
 There is a significant difference between all scenarios in terms of clade conservation ($NTS_{RF}$ - Table \ref{Tab_pooledscenarios_test}).
- % (maximising, minimising and randomising; \textit{p} = 0; Table \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (BC = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability of overlap between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
 However, there is no difference in terms of taxon displacement between the three scenarios ($NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}).
@@
 This effect of the character difference affects clade conservation ($NTS_{RF}$) only over large differences in number of characters (100 \textit{vs.} 1000 - Table \ref{Tab_pooledscharacters_test}).
-% , median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ ).
 For closer differences between characters numbers (100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests showed clear differences ($p = 0$) but with high distribution overlap ($BC > 0.96$).
 Regarding taxon displacements ($NTS_{Tr}$), there was an effect of the number of characters  although the with a high distribution overlap between 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
-% (median $NTS_{Tr}$ of 0.614, 0.793 and 0.896 for respectively 100, 350 and 1000 characters - supplementary materials 3, table 4) although the with a high distribution overlap between 350 and 1000 characters ($BC > 0.95$ - Table \ref{Tab_pooledscharacters_test}).
 
@@
 A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in $NTS_{RF}$ (Table \ref{Tab_pooledstaxa_test}) but no differences between 75 and 150 taxa for $NTS_{Tr}$ (Table \ref{Tab_pooledstaxa_test}).
- % (median $NTS_{RF}$ of 0.846 and 0.787 for 25 and 75 taxa respectively with a $p = 0$ and a $BC$ marginally $< 0.95$ - Table \ref{Tab_pooledstaxa_test} and supplementary materials 3, table 5) but no differences between 75 and 150 taxa and for $NTS_{Tr}$ ($BC > 0.95$ - Table \ref{Tab_pooledstaxa_test}).
 The significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).
@@
 \section{Discussion}
-%AP- ok so im understanding that its important to measure CD because it tells you wether you are inflating your odds of getting a well resolved tree.
+% @@@AP- ok so im understanding that its important to measure CD because it tells you wether you are inflating your odds of getting a well resolved tree.
 If you find low CD it means you should check your characters to see if you are committing coding dependency or if your traits are evolutionarily dependent then its probably an honest phylogenetic signal.
@@
 
-\subsection{Effect of character differences on topology}%AP-the first sentence of the discussion should refer back to the title in my opinion.
-Ie.
-We find that different modes of character correlation have a strong effect on the ability to recover the correct phylogenetic tree.
-As expected, there is a significant effect of character difference (the inverse of character correlation) on the ability to recover the correct topology (in terms of clades).
-% The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
-When characters are correlated, matrices convey a strong (but potentially misleading) phylogenetic signal since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
-%AP do you have to use 'homoplasy' here?
-I had to look it up (for the billionth time) understand that it means independent trait origins- but dont understand how homoplasy would be related to weaker signal.
-Nor why a maximized tree would have higher homoplasy.
-% Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (i.e.
-because the characters are not different, they all support the same topology that might be incorrect), the ``maximised'' scenario to lead to poorly resolved trees (i.e.
-because characters are different, the matrix has a lot of homoplasy leading to badly resolved topologies) and the ``randomised'' scenario to perform the best at recovering the correct topology (i.e.
-a simple bootstrap procedure).
-However, throughout our simulation, the ``minimised'' character difference scenario (i.e.
+\subsection{Effect of character differences on topology}
+We find that there is a significant influence of morphological character correlation on inferring tree topology.
+When characters are correlated, matrices convey a strong (but potentially misleading) phylogenetic signal since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of characters giving a conflicting signal (homoplasy).
+Throughout our simulation, the ``minimised'' character difference scenario (i.e.
 increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario led to slightly worse trees (Fig.
 \ref{Fig:RF_results_best}).
-This effect of increasing or reducing character difference did not appear to affect taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
-However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
-Furthermore, fewer characters were replaced, on average, in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6).% - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
-% However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 4, 5 and 6).
+These results suggest that, in matrices with high character differences, bootstrapping or other data randomisation methods can effectively improve the ability to recover the correct topology.
+
+However, this effect of increasing or reducing character difference did not appear to affect taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
+Furthermore, these results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
+Furthermore, fewer characters were replaced, on average, in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6).
 This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.
-%AP this paragraph is so dry and boring, refer to the supplements??-get out of the weeds!  this is just rehashing results with no interpretation.  I suggest rewriting completely.
 
 \subsubsection{Number of characters and taxa}
-% Because of the nature of our simulation protocol, one could expect that the effect of character correlation would have increased with the number of characters (i.e.
-the more characters available, the more characters are modified in each scenario).
-% Similarly, one could expect the number of taxa to have an effect of the raw ability to recover the ``normal'' topology (i.e. the more taxa, the more likely taxa are misplaced by chance).
-Overall, it is expected that minimising character difference %(increasing character correlation)
- is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal. %ap- I would change "correct" to "more resolved"
-If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good.
-Similarly, if the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree, then the estimates will be poor.
-%AP-The above paragraph isnt about characters or taxa, however it would be good for the first paragraph of the discussion.
-While matrices with more characters or taxa tended to be better resolved than matrices with fewer characters or taxa, the effect of character correlation on topology is independent of the size of the matrix.
-(Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), %in terms of number of characters, between 100 and 1000 characters;
-these differences have to be contrasted with the Bhattacharrya Coefficients that are always very high between every matrix's size.
-%AP- the Bhattacharrya stuff doesnt make sense, do you mean BC is always high between matrices of different sizes? Why does it matter
-% This suggests that the overall character difference metric can be used as a proxy for character correlation as it is not directly dependent on the number of characters analysed.
-Regarding the number of taxa, our results suggest the effect of character correlation on topology is independent on the size of the matrix when pooling the data. %ap Cut-as I summarized above
-Furthermore, using the Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered.% (i.e.
-we corrected for the expected difference when comparing two random trees with the same number of taxa).
-%ap-This seems like weeds
+We found an effect of the number of characters on the ability to recover the ``best'' topology with different levels of character correlations only between the two extreme number of characters (100 and 100; Tables \ref{Tab_pooledscharacters_test}).
+This is expected since a higher number of characters can results in better topologies with a higher probability of ``noisy'' signal to cancel out (\textit{c.f.} in smaller matrices).
+While matrices with more characters or taxa tended to be better resolved than matrices with fewer characters or taxa, the effect of character correlation on topology is independent of the size of the matrix (Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}).
+This can be explain partly because of our method to compare trees using the Normalised Tree Similarity \citep[$NTS$;][]{Bogdanowicz2012} which corrects for the fact that topological differences are proportional to the number of taxa considered.
 
@@
 When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (Table \ref{Tab_pooledsmethods_test}).
-% both using a Wilcoxon test and the Bhattacharrya Coefficient; 
 The difference in the ability of each method to recover the ``correct'' topology has been recently discussed with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; but see \citealt{spencerefficacy2013,goloboff2017weighted}).
@@
 It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.
-%AP-Solid paragraph
-
-% The differences in performance of the Bayesian and parsimony approaches observed here could be due to the inherent mechanisms of each method.
-% For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scores differently:
-% (1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the focal topology.
-% The tree search will thus likely result in a topology island that will not contain the given topology $T$.
-% (2) in Bayesian inference, the topology will also be given a bad score (i.e.
-low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
-% The rate and the branch length being two parameters among others, the score (the likelihood) will change less drastically than for using parsimony.
-% Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
-% This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
-% This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 15, 16, and 17).
 
@@
 in evo-devo \citealt{goswami2006morphological}; or in macroevolution \citealt{fitzjohn2014much}).
-These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not be used \textit{a priori} as a criterion to select characters.
-%AP-It'd be great to talk here about the problem with a priori selection creating circular and biased reasoning in a bit of depth.
-% In other words, intra-organismal and evolutionary correlation should be studied based on an underlying phylogenetic framework making the correlation induced by data collection (i.e.
-coding correlation) the only type of correlation that can affect the phylogeny \textit{a priori}.
+These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not be used \textit{a priori} as a criterion to select characters (to avoid creating circular and biased reasoning).
 This creates a trade-off between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
@@
 
-The two other sources 
-%MB: These are the same types of correlation you mention in the preceding paragraph.
- of character correlation are present in our simulations although they were not explicitly modelled:
+Intra-organismal and evolutionary correlations are present in our simulations although they were not explicitly modelled:
 (1) evolutionary correlation is implied by simulating the characters using Birth-Death trees;
-and (2) intra-organismal correlation is present in the matrices for the characters randomly simulated but sharing similar evolutionary simulation regimes.% (i.e. creating ``modules'' of characters).
+and (2) intra-organismal correlation is present in the matrices for the characters randomly simulated but sharing similar evolutionary simulation regimes.
 However, the effect of these sources of correlation was out of the scope of this study and would have required \textit{a posteriori} changes to the matrices \citep{Lande1983,Maddison1990,Pagel1994}.
-%AP We should probably change the title of the paper then...
-%(i.e. inferring the topology first and looking at character's correlation based on the observed topology \citealt{Lande1983,Maddison1990,Pagel1994}).
+
+\subsection{Potential applications}
+Effectively, our simulation protocol bootstraps our data ``with bias''.
+In the ``randomised'' scenarios the data is simply randomly bootstrapped.
+However, in the ``minimised'' and ``maximised'' scenario, we remove the characters with the lowest/highest overall character difference.
+For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with each other and randomly resample from the characters left.
+
+In rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
+Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling
+will be sufficient to obtain a robust topology.
+In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery.
+Applying this type of bootstraps that minimised character difference by biasing the random sampling could thus help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.
+Overall, we suggest measuring (and reporting) the levels of character difference in a matrix to help understand potential caveats in the distribution of optimal trees (e.g.
+a low average character difference will expect to generate more robust trees in bigger matrices).
 
@@
 This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to the other outputs rather than to the simulation inputs.
-% It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
-% This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches \citep[e.g.][]{revell2014ancestral}.
-% However, in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).
 
-Second, modifying character correlation is difficult.
-%AP- the whole point is to start using CD to measure it...
+% Second, modifying character correlation is difficult.
+% @@@AP- the whole point is to start using CD to measure it...
 you could maybe drop this paragraph and subsume the literature discussion into the discussion I suggested above about the 3 types of dependency and the potential for circular logic.
 But maybe add a few lines on what you believe is state of the art for dealing with it, and then tie it up with a CHARGE to find betters ways to deal with it like in your monotreme study
-In our simulation protocol we chose to create simulations by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 4, 5 and 6).
-In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
-This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
-However, because of the number of characters, it is currently too complex to attribute correlation to each kind of biological  and coding dependence and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
-% Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e.
-some sets of characters can be highly correlated but effectively describe independent information).
-Therefore, we made the choice to simplify our simulations by generating character correlation as a stochastic process rather than a biological one.
+% In our simulation protocol we chose to create simulations by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 4, 5 and 6).
+% In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
+% This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
+% However, because of the number of characters, it is currently too complex to attribute correlation to each kind of biological  and coding dependence and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
+% Therefore, we made the choice to simplify our simulations by generating character correlation as a stochastic process rather than a biological one.
 
-Third, comparing phylogenetic inference methods is not trivial.
-As mentioned above, both maximum parsimony and Bayesian inference
-% although aiming (and often achieving) to infer evolutionary history,
-have similar outputs nut vastly differ in how optimality is measured.
+Second, comparing phylogenetic inference methods is not trivial.
+As mentioned above, both maximum parsimony and Bayesian inference have similar outputs nut vastly differ in how optimality is measured.
 There are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
@@
 
-\subsection{Potential applications} %AP -yup  nice- maybe move above of caveats
-Effectively, our simulation protocol bootstraps our data ``with bias''.
-In the ``randomised'' scenarios the data is simply randomly bootstrapped. %, simply we randomly remove and re-sample characters (i.e. giving the weight of $0$ to some and $>1$ to other).
-However, in the ``minimised'' and ``maximised'' scenario, we remove the characters with the lowest/highest overall character difference.
-For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with other
-% MB: I think a word is missing.
- and randomly resample from the characters left.
-
-In rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
-Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling
-%(i.e. the equivalent of randomising the character differences in our analysis)
-will be sufficient to obtain a robust topology.% (\textit{cf.} actively collecting different characters).
-In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery.
-Applying this kind of bootstraps that minimised character difference by biasing the random sampling could thus help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.
-%MB: Rewrite.
-Avoid informal constructions like "kind of...".
-As a side note: "kind of..." is a harmless thing to say, but after some things it can mean everything, like after "you're going to live" or "It's a boy!".
-
 \subsection{Conclusion}
 Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon.
-% to investigate phylogenetic relationships.
 Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g.
 ][]{de2015parsimony,BrazeauNA} already exist.
-However, with
-%the regain of popularity in discrete morphological data and
-the expansion of dataset size \citep[e.g.][> 1000 characters]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
-Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate between these correlation types, it is important to understand to what extent topologies can be induced %MB: Not sure this is word we want.
- by such bias.
+However, with the expansion of dataset size \citep[e.g.][> 1000 characters]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
+Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate between these correlation types, it is important to understand to what extent topologies can be influenced by such bias.
 
-We found that character differences as a proxy for character correlation have a strong effect on recovering the ``normal'' topology: when character correlation was low (high character differences), the topology was always the furthest away from the ``normal'' topology.
-Conversely, when correlation between characters was high, the topology was always the closest %MB: closer than what?
- to the ``normal'' topology.
-These results appear independent of the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
-However, in modest size matrices (25 taxa; 100 to 350 characters), the effect of actively choosing to minimise or maximised character correlation was not more significant than simply randomising (i.e.
-bootstrapping) the matrix, suggesting that character correlation is more of a problem in large discrete morphological matrices.
+We found that character correlation have a strong effect of recovering the correct topology with trees with high character differences always performing more poorly at recovering the correct topology than trees with low or randomised character differences.
+This is especially true in large matrices where we advice to reduce character difference by using bootstrapping with ``bias'' whereas this may not be major problem in modest size matrices (25 taxa; 100 to 350 characters).
 We recommend that all morphological phylogenetic analyses should report the level of character difference so that we can begin to detect and address the problem of coding dependence on inflating our confidence in our phylogenetic reconstructions.
-%Or some similar charge
-% For such matrices, minimising the character correlation (resampling characters $<25$\% different) or maximising it ($>75$\%) respectively significantly increased and decreased correct topological recovery compared to randomly resample matrices.
-% MB: I think the conclusion could be briefer: focus more on the broad-level implications and take it back to what this means for palaeontologists.
 

### 2019-11-04 Writing/Palaeontology/draft.tex
added Martin's two comments

@@
 Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be diminished \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
-In this study, use a metric of raw character difference as a proxy for character correlation.
+In this study, we use a metric of raw character difference as a proxy for character correlation.
 Using thorough simulations, we test the effect of increasing or decreasing character differences on tree topology.
 Overall, we found a negative effect of reducing character correlations on recovering the correct topology.
-This means that dataset with low correlation between characters will make it more complicated to estimate a correct topology.
+This means that datasets with low correlation between characters will make it more complicated to estimate a correct topology.
 However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing or increasing character correlation is not more effective in getting the expected topology than randomly drawing characters.
@@
 Although it has been tackled empirically (for morphological data \citealt{Davalos01072014}, and molecular data \citealt{ZouConvergence}), the question has never been explored through a thorough simulation framework.
-Simulations can allow the design of dataset with controlled levels of correlation and for which the correct expected answer is known by design.
-Furthermore, using simulations will allow to draw generalities on the effect of character correlation on phylogenetic inference conversely to potential idiosyncrasies of empirical datasets.
+sSimulations can allow the design of datasets with controlled levels of correlation and for which the correct expected answer is known by design.
+Furthermore, using simulations will allow drawing generalities on the effect of character correlation on phylogenetic inference and conversely to potential idiosyncrasies of empirical datasets.
 Here we use a Hamming distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
@@
 In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery.
-Applying this type of bootstraps that minimised character difference by biasing the random sampling could thus help understanding which characters generate more homoplasy and which nodes in the tree are conflictual.
+Applying this type of bootstraps that minimised character difference by biasing the random sampling could thus help understanding which characters generate more homoplasy, which nodes in the tree are conflictual or which clades might be most dependent on highly correlated characters.
 Overall, we suggest measuring (and reporting) the levels of character difference in a matrix to help understand potential caveats in the distribution of optimal trees (e.g.
 a low average character difference will expect to generate more robust trees in bigger matrices).
+%@@@ MB: I think we need to be careful here.
+Trees with a lot of high, consistent signal might appear to have low CD.
+What might be a better metric to investigate is the density of characters that have very high (like, perfectly correlated) CDs a hint that a lot of the data might reflect coding the same thing.
 
@@
 
-We found that character correlation have a strong effect of recovering the correct topology with trees with high character differences always performing more poorly at recovering the correct topology than trees with low or randomised character differences.
+We found that character correlation has a strong effect on recovering the correct topology with trees having high character differences always performing more poorly at recovering the correct topology than trees with low or randomised character differences.
+%@@@ MB: Might be worth pointing to this and adding this as a reassurance: correlations not driven by selection  (e.g.
+pleitropy) could cancel each other out if they are randomly distributed  with respect to traits.
 This is especially true in large matrices where we advice to reduce character difference by using bootstrapping with ``bias'' whereas this may not be major problem in modest size matrices (25 taxa; 100 to 350 characters).

### 2019-11-04 Writing/Palaeontology/draft.tex
AP comments

@@
 Overall, we found a negative effect of reducing character correlations on recovering the correct topology.
+%@@@ TG: AP -> accuracy precision trade off.
 This means that datasets with low correlation between characters will make it more complicated to estimate a correct topology.
@@
 Furthermore, in bigger matrices (350 characters), the inference method has a strong effect with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
-These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
+These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis and that analysing the correlations between characters could help understanding problematic phylogenetic matrices.
 However, encouragingly, they also suggest that, unless correlation is actively maximised or minimised, probabilistic methods can easily accommodate for a random correlation between characters.
@@
 
-Biological and evolutionary dependences are inherent parts to Evo-Devo and macroevolutionary studies and best practices to avoid coding-induced dependences are commonly known and applied.
+Intra-organismal and evolutionary dependences are inherent parts to Evo-Devo and macroevolutionary studies and best practices to avoid coding-induced dependences are commonly known and applied.
 However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inference software that are blind to these distinctions.
@@
 \begin{itemize}
-    \item \textbf{Characters compatibility}: this is the result of how software actually interprets the differences between characters.
+    \item \textbf{Character compatibility}: this is the result of how software actually interprets the differences between characters.
     The majority of phylogenetic algorithms cannot interpret character descriptions and state meanings, and instead must interpret discrete states as signifiers of shared similarity: for each characters, species are grouped if they have identical characters.
@@
 sSimulations can allow the design of datasets with controlled levels of correlation and for which the correct expected answer is known by design.
-Furthermore, using simulations will allow drawing generalities on the effect of character correlation on phylogenetic inference and conversely to potential idiosyncrasies of empirical datasets.
+Furthermore, using simulations will allow drawing generalities on the effect of character correlation on phylogenetic inference in contrast to potential idiosyncrasies of empirical datasets.
 Here we use a Hamming distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase or decrease the overall differences or similarities between characters.
@@
 These results suggest that, in matrices with high character differences, bootstrapping or other data randomisation methods can effectively improve the ability to recover the correct topology.
+%@@@ AP: does it???
+I dont think I agree!
+basically this just over inflates the weight of the characters you picked at random.
+So if its a bad character-you will recover an incorrect tree with greater confidence.
+No?
 
@@
 \subsubsection{Number of characters and taxa}
-We found an effect of the number of characters on the ability to recover the ``best'' topology with different levels of character correlations only between the two extreme number of characters (100 and 100; Tables \ref{Tab_pooledscharacters_test}).
-This is expected since a higher number of characters can results in better topologies with a higher probability of ``noisy'' signal to cancel out (\textit{c.f.} in smaller matrices).
+We found an effect of the number of characters on the ability to recover the ``best'' topology with different levels of character correlations only between the two extreme number of characters (100 and 1000; Tables \ref{Tab_pooledscharacters_test}).
+This is expected since a higher number of characters can result in better topologies with a higher probability of ``noisy'' signal to cancel (\textit{c.f.} in smaller matrices).
 While matrices with more characters or taxa tended to be better resolved than matrices with fewer characters or taxa, the effect of character correlation on topology is independent of the size of the matrix (Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}).
-This can be explain partly because of our method to compare trees using the Normalised Tree Similarity \citep[$NTS$;][]{Bogdanowicz2012} which corrects for the fact that topological differences are proportional to the number of taxa considered.
+This can be explained partly by our method of comparing trees using the Normalised Tree Similarity \citep[$NTS$;][]{Bogdanowicz2012} which corrects for the fact that topological differences are proportional to the number of taxa considered.
 
@@
 \subsection{Distinction between different character correlations}
-Here we mention three different types of character correlations but evolutionary biologists are mainly interested in the intra-organismal and evolutionary correlations (e.g.
+Here we mention three different types of character correlations but evolutionary biologists are mainly interested in intra-organismal and evolutionary correlations (e.g.
 in evo-devo \citealt{goswami2006morphological}; or in macroevolution \citealt{fitzjohn2014much}).
@@
 In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery.
-Applying this type of bootstraps that minimised character difference by biasing the random sampling could thus help understanding which characters generate more homoplasy, which nodes in the tree are conflictual or which clades might be most dependent on highly correlated characters.
+Applying this type of bootstraps that minimised character difference by biasing the random sampling could thus help in understanding which characters generate more homoplasy, which nodes in the tree are conflictual or which clades might be most dependent on highly correlated characters.
 Overall, we suggest measuring (and reporting) the levels of character difference in a matrix to help understand potential caveats in the distribution of optimal trees (e.g.
@@
 pleitropy) could cancel each other out if they are randomly distributed  with respect to traits.
-This is especially true in large matrices where we advice to reduce character difference by using bootstrapping with ``bias'' whereas this may not be major problem in modest size matrices (25 taxa; 100 to 350 characters).
+%@@@ AP: Rephrase above.
+This is especially true in large matrices where we advise to reduce character difference by using bootstrapping with ``bias'' whereas this may not be major problem in modest size matrices (25 taxa; 100 to 350 characters).
 We recommend that all morphological phylogenetic analyses should report the level of character difference so that we can begin to detect and address the problem of coding dependence on inflating our confidence in our phylogenetic reconstructions.

### 2019-11-04 Writing/Palaeontology/draft.tex
Updated draft

@@
 Overall, we found a negative effect of reducing character correlations on recovering the correct topology.
-%@@@ TG: AP -> accuracy precision trade off.
 This means that datasets with low correlation between characters will make it more complicated to estimate a correct topology.
@@
 Furthermore, in bigger matrices (350 characters), the inference method has a strong effect with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
-These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis and that analysing the correlations between characters could help understanding problematic phylogenetic matrices.
+These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
 However, encouragingly, they also suggest that, unless correlation is actively maximised or minimised, probabilistic methods can easily accommodate for a random correlation between characters.
@@
 If fact, what the software are confronted with is a two dimensional matrix problem that renders all morphological dependencies invisible.
-This introduces a new, less studied, source of character dependence:
-
-\begin{itemize}
-    \item \textbf{Character compatibility}: this is the result of how software actually interprets the differences between characters.
 The majority of phylogenetic algorithms cannot interpret character descriptions and state meanings, and instead must interpret discrete states as signifiers of shared similarity: for each characters, species are grouped if they have identical characters.
 Therefore, character with the same number of states will have a chance of being correlated regardless of mode of form of character correlation.
-\end{itemize}
 
@@
 \section{Discussion}
-% @@@AP- ok so im understanding that its important to measure CD because it tells you wether you are inflating your odds of getting a well resolved tree.
-If you find low CD it means you should check your characters to see if you are committing coding dependency or if your traits are evolutionarily dependent then its probably an honest phylogenetic signal.
-I'm not sure how Intra-organismal fits in though.
-So basically the message of this paper should be "Always check your characters for correllation using this sweet metric-and if its high -check to see what kind of correlation you have"(This should go in the abstract as well) It'd then be nice to 'prescribe' methods for checking which kind of correlation you have.
-Ok, Now I'm going to read the discussion and see if you did this.
 
@@
 These results suggest that, in matrices with high character differences, bootstrapping or other data randomisation methods can effectively improve the ability to recover the correct topology.
-%@@@ AP: does it???
-I dont think I agree!
-basically this just over inflates the weight of the characters you picked at random.
-So if its a bad character-you will recover an incorrect tree with greater confidence.
-No?
 
@@
 Applying this type of bootstraps that minimised character difference by biasing the random sampling could thus help in understanding which characters generate more homoplasy, which nodes in the tree are conflictual or which clades might be most dependent on highly correlated characters.
-Overall, we suggest measuring (and reporting) the levels of character difference in a matrix to help understand potential caveats in the distribution of optimal trees (e.g.
-a low average character difference will expect to generate more robust trees in bigger matrices).
-%@@@ MB: I think we need to be careful here.
-Trees with a lot of high, consistent signal might appear to have low CD.
-What might be a better metric to investigate is the density of characters that have very high (like, perfectly correlated) CDs a hint that a lot of the data might reflect coding the same thing.
+Overall, we suggest measuring (and reporting) the levels of character difference in a matrix to help understand potential caveats in the data.
+In fact, regions of the matrix where characters have a consistent low difference can reflect that they are coding for the same information.
 
@@
 This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to the other outputs rather than to the simulation inputs.
-
-% Second, modifying character correlation is difficult.
-% @@@AP- the whole point is to start using CD to measure it...
-you could maybe drop this paragraph and subsume the literature discussion into the discussion I suggested above about the 3 types of dependency and the potential for circular logic.
-But maybe add a few lines on what you believe is state of the art for dealing with it, and then tie it up with a CHARGE to find betters ways to deal with it like in your monotreme study
-% In our simulation protocol we chose to create simulations by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 4, 5 and 6).
-% In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
-% This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
-% However, because of the number of characters, it is currently too complex to attribute correlation to each kind of biological  and coding dependence and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
-% Therefore, we made the choice to simplify our simulations by generating character correlation as a stochastic process rather than a biological one.
 
@@
 
-We found that character correlation has a strong effect on recovering the correct topology with trees having high character differences always performing more poorly at recovering the correct topology than trees with low or randomised character differences.
-%@@@ MB: Might be worth pointing to this and adding this as a reassurance: correlations not driven by selection  (e.g.
-pleitropy) could cancel each other out if they are randomly distributed  with respect to traits.
-%@@@ AP: Rephrase above.
+We found that character correlation has a strong effect on recovering the correct topology: matrices with high character differences always performed more poorly at recovering the correct topology than trees with low or randomised character differences.
 This is especially true in large matrices where we advise to reduce character difference by using bootstrapping with ``bias'' whereas this may not be major problem in modest size matrices (25 taxa; 100 to 350 characters).
+However, in empirical datasets, if character difference is not driven by selection (e.g.
+pleiotropy) correlation is likely be cancelled out if the correlated characters are randomly distributed with respect to traits.
 We recommend that all morphological phylogenetic analyses should report the level of character difference so that we can begin to detect and address the problem of coding dependence on inflating our confidence in our phylogenetic reconstructions.

### 2020-01-21 Writing/Palaeontology/Reviews/response_to_reviewers.tex
Review INIT

--- 2019-06-21 Writing/response_to_reviewers.tex
@@
 
-\textbf{RE: USYB-2018-099}\\
+\textbf{RE: PALA-11-19-4635-OA}\\
 \bigskip
-Dear Dr Jermiin,\\
+Dear Dr Benson,\\
 \bigskip
@@
 \section{Editors' comments}
+
+\textcolor{blue}{Both referees had much to say.
+My summary is that the manuscript is highly technical and difficult to follow as currently written.
+But I don't think it needs to be.
+This is important for publication in Palaeontology, and indeed in other journals.
+It is in all of our interests that the work can influence a wide audience.
+With this in mind, the choices of language and structure would benefit from some attention, specifically with a view to making it more comprehensible, and cleaning up some of the concepts so they are framed less ambiguously.
+I agree especially with the point made by referee 2 about the difference between (1) confounding covariation among character state distributions, and (2) informative covariation between character states that is induced by shared evolutionary history (the signal we're trying to infer during phylogenetic inference).
+This is not incidental.
+To many people this is the point of such studies.}
+
+\textcolor{blue}{I have some illustrative comments of my own that help point to the type so issues I'm referring to.
+But note that these are not exhaustive.
+I'd like to see a deep and through consideration of these elements of communication, and those raised by the referees.
+It is likely that this constitutes fairly large modifications to the manuscript, but less so to the analyses and images.}
+
+\textcolor{blue}{My example points:}
+
+\begin{itemize}
+
+\item{\textcolor{blue}{The abstract lists "shared phylogenetic history" as a factor that causes character correlation in a list of three factors.
+Approximately the same list of factors is presented in the introduction (p.5, bulletted), where "evolutionary dependence" is stated in place of "shared phylogenetic history".
+This seems confusing as the two lists seems to serve the same purpose but have non-overalpping content.
+Shared evolutionary history is a positive thing (we're trying to infer evolutionary history) whereas "evolutionary dependence" of traits seems like it could be negative in some circumstances (for the body elongation example, it may cause it to infer a clade of long-bodied taxa).}}
+
+\item{\textcolor{blue}{Matrices that were modified to minimise character differences are referred to as 'minimised'.
+However, in the context of the main questions, 'minimisation' of character differences corresponds to maximisation of character state correlations.
+Since these two concepts have opposite polarities, it would be easier for readers to follow the work if it consistently referred to one or the other - and I'd propose that referring to the amount of character correlation would be best and more straightforward for readers.}}
+
+\item{\textcolor{blue}{'Normal' has multiple meanings and mathematically could mean 'orthogonal'.
+In the way the authors use it, it comes across as colloquial, and it doesn't contain much information (there is no objective meaning of 'normal' in this context).
+A more intuitive term might be 'unmodified' or 'simnulated', or something else, there are probably better options.}}
+
+\item{\textcolor{blue}{'In a few places 'phylogenetic signal' is used to refer to a situation where a topology is well-supported.
+But I had the impression this might not strictly correspond to support of the "true" phylogeny (correct me if I'm wrong however).
+Phylogenetic signal is a widely-used phrase with several different potential meanings.
+So to avoid confusion it would be useful each time if the text referred explictily to support for the "true" phylogeny underlying the simulations, or to support for nodes in the inferred phylogeny, whether they are present in the true phylogeny or not.
+Or tended to produce strong support for incorrect topologies.}}
+
+\end{itemize}
+
+\section{Referee 1}
+
+Comments to the Author
+Please see technical comments in the attached file: Guillerme_4635_TE
+
+\section{Referee 2}
+
+Comments to the Author
+See attached file...
+
+\section{Referee 3}
+
+\textcolor{blue}{Sadly however, I am skeptical that this paper really addresses the central issue with character correlation and inferring phylogenies.
+Particularly, their main conclusion is that character correlation is a 'good' thing, i.e.
+having increased character correlation actually results in more accurate topologies.}
+
+\textcolor{blue}{One of the central conclusions (from the abstract) is 'This means that datasets with low correlation between characters will make it more complicated to estimate a correct topology'.
+This is not a novel conclusion, nor does it really have anything to do with the issue that the authors claim to address.
+All this is really confirming is that inferring phylogenies from matrices that contain limited phylogenetic signal is difficult; it is not related to how correlated characters that show homoplasy impact our ability to infer accurate phylogenetic hypotheses.
+If character correlation really were a positive influence on topological accuracy then character matrices should be coded using a/p coding (rather than using hierarchical characters with some taxa being inapplicable for certain characters) as this coding method would maximize character correlation (e.g.
+Pleijel 1995, Cladistics).}
+
+\textcolor{blue}{Character correlation should only be a problem for inferring an accurate topology if the correlated characters contain misleading phylogenetic signal and change multiple times in concert on the tree, i.e.
+show coordinated homoplasy.
+This is effectively the 'Intra-organismal dependence' type of character correlation that the authors identify in the introduction, the type of character correlation that is probably of most concern regarding inferring phylogenetic trees.
+This type of dependent character state change is not only an issue for morphological phylogenetics, but is likely also a problem for molecular data.
+For example the presence of particular amino acids at particular sites in proteins is fundamental for determining 3D protein structure, limiting the substitutions that can occur at particular sites if proteins are to maintain this 3D structure, which is similar to the tooth occlusion example that is cited in the text.
+This a particularly relevant and interesting example, especially given that an empirical analysis investigating the performance of dental data from mammals found that they recover phylogenies that are more in conflict with those inferred from molecular data (Sansom et al.
+2017) citing character non-independence as a possible factor driving this incongruence.
+This paper is absent from the bibliography and should certainly be discussed, especially as the conclusions of this paper differ so radically.}
+
+\textcolor{blue}{In the discussion (P6.
+Line 339) the authors state '(1) evolutionary correlation is implied by simulating the characters using Birth-Death trees; and (2) intra-organismal correlation is present in the matrices for the characters randomly simulated but sharing similar evolutionary simulation regimes.
+However, the effect of these sources of correlation was out of the scope of this study and would have required a posteriori changes to the matrices'.
+This means that characters that are correlated due to shared evolutionary history and characters that show coordinated homoplasy are not distinguished by downstream analyses, limiting the applicability of the results for understanding issues raised by real world examples (such as the dental data analysis by Sansom et al.
+2017, Sys Bio).}
+
+\textcolor{blue}{In the conclusions the authors state: 'However, in empirical datasets, if character difference is not driven by selection (e.g.
+pleiotropy) correlation is likely be cancelled out if the correlated characters are randomly distributed with respect to traits.' I am not entirely sure what the authors mean by this sentence.
+I think that what they are suggesting is that character correlations that arise due to selection will have different impacts on accuracy of phylogenetic reconstructions vs those that arise from drift, which really is what this paper should be trying to address.
+Instead the analyses pool these types of character correlation rather than trying to disentangle them.}
+
+\subsection{Simulation protocol and methodology}
+
+\textcolor{blue}{I am not sure that I understand the logic of only comparing the results of the phylogenetic analyses to what the authors call the 'normal' tree and not the tree that generated the data.
+Surely the whole point of this exercise is to understand if character correlation actually results in misleading phylogenetic inferences.
+It strikes me that the way in which they treat their morphological matrices just results in the 'minimised' matrix having the phylogenetic signal that supports the 'normal' tree reinforced as the correlated characters will essentially just contain the same phylogenetic signal that supported the 'normal' tree, though not necessarily the true tree.
+For many of the simulated matrices, the normal tree could share very few clades with the tree that was used to generate the data itself, which is clearly demonstrated in O'Reilly et al 2016 (where the same binary HKY simulation method was also used).
+In this paper, the accuracy of the trees from downstream phylogenetic analyses are often very different to the generating tree.
+In any revision, I would want to see the impact of comparing the simulated matrices to the generating tree.
+}
+
+\subsection{Minor/stylistic comments}
+
+\textcolor{blue}{The merged PDF for review has identical font size for figure/table captions and the main text, making it difficult to read fluently.}
+
+\textcolor{blue}{The text has many typos, mostly pluralization (e.g. character vs characters). Randomised is misspelled in Fig 1. Please check the text and figures thoroughly.}
 
@@
 
-\item{\textcolor{blue}{29:434 – delete ‘a'}}
+\item{\textcolor{blue}{29:434 – delete `a'}}
 
@@
 
-\item{\textcolor{blue}{30: 450 - < $->$ >}}
+\item{\textcolor{blue}{30: 450 - $<$ $->$ $>$}}
 
@@
 
-\item{\textcolor{blue}{30:469 – delete ‘b'}}
+\item{\textcolor{blue}{30:469 – delete 'b'}}
 

### 2020-01-21 Writing/Palaeontology/Reviews/response_to_reviewers.tex
added reviewer 1 and 2 comments

@@
 
-Comments to the Author
-Please see technical comments in the attached file: Guillerme_4635_TE
+\subsection{General notes}
+\begin{itemize}
+\item{\textcolor{blue}{Please supply your main text in Word format rather than PDF; please supply all figures as separate files and include a list of figure captions at the end of the manuscript rather than embedded in the text.
+Please check journal submission guidelines:}}
+
+\item{\textcolor{blue}{Please respond directly to all referee comments, including the technical comments below.
+It is particularly important that you explain your reasoning if you have not followed any of the suggestions made in the reports.}}
+
+\item{\textcolor{blue}{Please upload your response to the reviewers as a separate document designated as a 'Supplementary File' with the other submission files.
+This will pull it into the automatically generated proof that is available to all reviewers.}}
+
+\item{\textcolor{blue}{Do any of the authors have an ORCID iD?
+This is a digital identifier that allows all publications to be permanently linked to the individual authors, even if they change institution or name.
+It is free to register, and we strongly recommend that at least the corresponding author should do so.
+There is more information available here: http://orcid.org/.
+If each author can link their ScholarOne account directly to ORCiD (which requires logging into ORCiD to authorise) they will automatically be added on this and any future paper you might submit, and metadata for a published paper will automatically be sent to their ORCiD account.
+I have attached some guidelines on how to do this.
+You can also simply add IDs to the manuscript (with the affiliation information) but our publisher cannot push publication data to ORCiD without consent.}}
+
+\end{itemize}
+
+\subsection{Data}
+
+\begin{itemize}
+\item{\textcolor{blue}{Thank you for uploading your data to GitHub.
+Please replace the link in the text with a citation to a data reference: Guillereme, T., Pastore, A.
+and Brazeau, M.
+20xx.
+Influence of different modes of morphological character correlation on phylogenetic tree inference.
+GitHub.
+https://github.com/TGuillerme/CharactersCorrelation}}
+
+\item{\textcolor{blue}{We no longer upload any supplementary information on our publisher's website.
+This information, including figures, needs to be properly curated.
+Can your supplementary figures be uploaded to GitHub or Figshare?
+If not, I suggest Dryad (the Palaeontological Association will pay the fee for a Dryad record for any data package associated with a paper in one of its journals).
+Please contact me directly for an upload link if you wish to use Dryad (other repositories such as MorphoBank are also acceptable).
+Any text should be supplied in a non-proprietary format (txt files can be supplied alongside a styled pdf version if you prefer).
+Dryad encourages uploading data in multiple formats if styling is important.}}
+
+\item{\textcolor{blue}{As above, any data citations in the text should be linked to a data reference including the doi for the record.}}
+\end{itemize}
+
+\subsection{References}
+\begin{itemize}
+\item{\textcolor{blue}{Where you have two references with identical citations (e.g.
+Guillerme \& Cooper 2016) please list these in the order in which they are cited in the text, irrespective of publication order, or alphabetical order of second author.
+Therefore Guillerme \& Cooper 2016a should be cited first, whichever reference this is.
+Having said that, I only found references of Guillerme \& Cooper 2016b so please check this.}}
+\end{itemize}
+
+\subsection{Figures}
+\begin{itemize}
+\item{\textcolor{blue}{Please note that although figures may be reproduced in colour in the online edition, there is a fee to reproduce them in print.
+If a figure appears in colour online and greyscale in print, the same file will be used for both, so please carefully consider your use of colours to make sure that it works in shades of grey.}}
+\item{\textcolor{blue}{Please supply all of your figures at a resolution of 600 dpi; preferably in tif format using LZW compression.
+Embedded photographs are fine at 300 dpi, but if any labelling is included the overall figure will require 600 dpi for printing.
+Please do not use jpg compression at any stage.
+Final widths should be either single column (80 mm), 2/3 page width (110 mm) or double column (166 mm).
+Please see the attached figure
+guidelines.}}
+\item{\textcolor{blue}{Please confirm the final intended width of each figure (166 mm = full page; 110 mm = 2/3 page; or 80 mm = single column) by adding this to the file name (add 166, 110 or 80).}}
+\item{\textcolor{blue}{Fig.
+1: Please note that although the journal uses British English, it uses suffix variants - ize, -iza and -izi where appropriate (rather than -ise etc.) This would be changed in the main text but it would be helpful if you could correct any instances on figures (e.g.
+Randomized, Maximized, Minimized).
+Similarly, journal style uses single quotes throughout (e.g.
+'True' tree)}}
+\item{\textcolor{blue}{Fig.
+2: Note spelling as above.
+Please do not refer to colours in the caption if this is intended for printing in colour.
+Generally, a visual legend is much better on figures anyway (there is room for this in the top right hand corner; it does not require a title).}}
+\item{\textcolor{blue}{Figs 3, 4: These figures would require colour both online and in print to be clear.
+If you do not have funding available to pay the colour printing fee (£150 for one figure, plus £50 for each additional one) please state this clearly in your cover letter as the Palaeontological Association is willing to cover the cost of essential colour where alternative funding is not available (at the Editor-in-Chief's discretion).
+Please view these two figures at full page width (166 mm wide) and check that all text labels are clearly visible.
+The axis labels in particular are currently too small; generally, text sizes should be in a range equivalent to 6–10 pt Arial (viewed at 100\%).}}
+\end{itemize}
+
+\subsection{Tables}
+\begin{itemize}
+\item{\textcolor{blue}{Please supply tables either separately, or at the end of the manuscript.}}
+\item{\textcolor{blue}{Please note that captions would be altered to include a single, short descriptive phrase, with all additional information/abbreviations moved to a footnote.}}
+\end{itemize}
 

### 2020-01-21 Writing/Palaeontology/Reviews/response_to_reviewers.tex
Reviewer 2

@@
 
-Comments to the Author
-See attached file...
+\begin{itemize}
+
+\item{\textcolor{blue}{Lines 22-23: 'probabilistic methods can easily accommodate for a random correlation between characters.'
+\\
+'random correlation' sounds like an oxymoron!
+I think that what the authors really mean is moderate amounts of correlation that basically mimic accelerated local rates of change on a small proportion of branches.
+From what the authors state later, it’s not the randomness that is important, but that it occurs infrequently relative to total taxon sampling.}}
+
+\item{\textcolor{blue}{Lines 50-55: Types of dependence.
+\\
+I found the definitions here adequate, but I think that they should be tied more closely to aspects of correlated character change as discussed elsewhere in systematics literature and even general evolutionary biology.
+I, myself, tend to think of correlated characters as falling into two general categories: functional suites and integrated modules.
+(Note that these need not be exclusive; modules can link functionally linked characters.)
+\\
+The reason why I prefer this is that we can model both functional suites and integration with extended Mk models.
+Suppose we have two binary characters.
+We will use matrix exponentiation to describe the probabilities of transitions between character combinations.
+Under independent change, our model would look like:
+\\
+Q=...
+\\
+We’ll set a=1 and let’s say we have a rate of change of π=0.05. Our transition matrix for time t=1 is now mexp(πQt) =
+\\
+Note that if we calculated the probability of going from 0 to 1 for either character, then we’ve have almost the same value: the difference is that there is some small chance that, over the same amount of time, we also went from 0 to 1 for the other character, and thus from 00 to 11.
+The probability of stasis is the probability of stasis for one character squared.
+\\
+Change here is independent; the condition of Character 1 has no effect how Character 2 evolves, and the two characters do not change together.
+(This is assuming continuous time change.) We can introduce a new Q matrix that allows Characters 1 and 2 to affect each other, and in which there are joint changes altering both characters:
+\\
+Q=...
+\\
+The two d values reflect what I think that authors mean by 'evolutionary dependence': 11 is an 'optimal' condition, and once you get to 10 or 01, then the probability that the other character changes increases (d1).
+Conversely, once you get to 11, then reversing back to 01 or 10 is less probable (d2).
+This is the sort of change we would expect in mosaic functional evolution, in which new changes were 'improving' on the overall morphotype.
+\\
+The off-diagonal reflects (j) joint change: a change 'deeper' in the organisms that affects both characters.
+Integration predicts this: we alter a module of linked characters, and we change states on multiple characters.
+We can get some independent change because different parts of the module can (sometimes!) change on their own; however, there are many other ways to change the module that will affect both (and possibly other) characters.
+Heterochronic shifts are one particular version of this model.
+\\
+This is where I was a little uncertain: does integration fit in with intra-organismal dependence?
+If so, then the authors should state that.
+At any rate, given the ubiquity of papers describing integration and modularity papers, I think that it would help the readers to understand the correlation models if the discussion was presented in terms of integration/modules and functional suites.}}
+
+\item{\textcolor{blue}{Lines 58-59: 'For example, two characters 'tail colour' and 'tail length' could be coded two times as an absence for a taxon with no tail (Wilkinson, 1995b; Brazeau et al., 2018).'
+\\
+Coding this as inapplicable makes this much less of a problem for likelihood, where it is treated as an unknown.
+In likelihood, unlike parsimony, the ancestral condition is not fixed: so, if tails are found in two subclades, then the marginals for the 'tailless' paraclade will have low values for tail being present: but there is some small probability that the tail was there, and then lost in parallel.
+If that tail was there, then it had States A, B, C, etc., for length, color, etc.
+Because we are multiplying this, it’s really P[tail present] X P([tail state A]+tail state B)) etc.}}
+
+\item{\textcolor{blue}{Lines 69-71: 'Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree, and is best determined by independent character sets.'
+\\
+I would add three points here.
+One, with fossil data (such as readers of Palaeontology use), we have very few options for independent character sets.
+Two, as the authors cite above, we have reason to think that correlated change happens with molecular evolution.
+The really important point is that what we want are character data that we know had independent character change: but no biological data are above suspicion here.
+Three, we do not need accurate resolution of trees if we have sets of characters that we suspect are correlated.
+Given such a set, we can run two sets of Markov Chain Monte Carlo analyses: one in which characters in the set evolve independently, and one in which we use more complex models in which the states are not the states of one character, but the combinations of states from 2+ characters (Billet \& Bardin 2018 Syst.
+Biol.
+68:267, Pagel 1994 Proc.
+R.
+Soc.
+B 255:37).
+We then can assess whether the correlated change model produces appreciably better posterior probabilities in the same way that we can contrast any other pair of different models.
+\\
+What Guillerme et al.
+are assessing is a little bit different from what Pagel and later workers considered: what happens if there is correlated character evolution happening among unknown sets of characters?
+Billet \& Bardin’s study focused on mammal teeth: but even non-biologists would expect mammal teeth to show correlated character change.
+This raises the issue with which readers of Palaeontology will be familiar: how do I have reason to suspect that there is correlated change in some group of characters, and what happens if I fail to guess?
+What this simulation study offers is an idea of what happens when there is some amount of correlated change, but we just do not know how much.}}
+
+\item{\textcolor{blue}{Lines 123 – 124: In order to measure this, we used the following scaled Hamming distance between two characters:
+\\
+I think that it would be worth comparing this to pairwise dissimilarity among taxa.
+After all, there have been numerous papers published in Palaeontology that use pairwise dissimilarity (also called phenetic dissimilarity or phenetic distance) to measure disparity.
+The difference is basically R-mode vs.
+Q-mode: for phenetic dissimilarity, we count the number of differences between two taxa and divide by the total number of comparable characters to get the proportion of different characters.
+Here, we count the number of differences between two characters and divide by n-1 to get (almost) the proportion of differing taxa.
+It basically is like generating scatterplots for every pair of characters, and then counting the number of taxa that fall above or below the line bisecting the axes (0,0 : m,m).}}
+
+\item{\textcolor{blue}{Lines 131-132: 'We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.'
+\\
+I think that the authors should state 'by order of derivation' or 'order of evolution.' I realize that their simulations evolve up to n taxa and then attempt to reconstruct phylogeny.
+However, many simulations in palaeobiological studies sample simulated taxa over time.
+In such cases, the order in which states evolve and the order in which they are sampled might not be identical, particularly for characters that evolve rapidly relative to sampling rates.}}
+
+\item{\textcolor{blue}{Lines 150 – 152: 'The 'morphological HKY-binary' model (O’Reilly et al.,
+2016) which is an HKY model (Hasegawa et al., 1985) with a random states frequency (sampled from a Dirichlet distribution Dir(1, 1, 1, 1))...'
+\\
+The authors should state that this allows variation in gains/losses of present absent characters to vary without specifying which characters are prone to gain/loss.
+The authors also should note that because it’s (1,1,1,1), the expectation is that gains and losses will be equiprobable.
+This does raise the question: why didn’t they use a beta distribution (1,1) given that they were dealing with 2-state characters?
+That’s just the special case of the Dirichlet.
+Using a 4 state character like this seems a little weird because net stasis will include both net stasis and net transition, and thus is P[net stasis] + P[net transition]; net change now is twice net transversion and thus 2*P[net transition].
+So, this will not quite be the same distribution as beta(1,1); the expectation will still be 50:50, but (I think!) the variance on the state-acquisition bias will be a little greater under the 'lumped' 4-state Dirichlet than under the Beta.}}
+
+\item{\textcolor{blue}{Lines 155-156: 'To generate more than binary states characters, we used the Mk model (Lewis, 2001).'
+\\
+How did the Mk model generate the 3-state characters?
+Shouldn’t the authors have used a Dirichlet (1,1,1) (or a Dirichlet (1,1,1,1,1,1) with lumping of 1\&2, 3\&4 and 5\&6) to actually generate the characters with individual rate bias?}}
+
+\item{\textcolor{blue}{Lines 159-161: 'For each character, both models ('morphological HKYbinary' or Mk) were chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution (β = 100 and α = 5).'
+\\
+The authors should make a couple of points clear here.
+The first is that they should make clear what this section is doing vs.
+the prior section.
+I would state plainly that the prior section was about generating variation in state gain/loss biases.
+This part is about generating variation in character change; basically, you are allowing each character to change at its own rate and (when it does change) to have its own bias in gains/losses of states; however, in all cases, every character’s particular rate and gain/loss bias is drawn from the same distribution.
+\\
+The second is that this is different from the way systematists usually use gamma distributions for rate variation, in which they have a separate rate parameter π (I will use π rather than α as Lewis did, simply to avoid confusion between rate parameter and the Gamma shape parameter α).
+They then set β=α so that the mean of the Gamma, α/(β=α), is 1.0 and the variance, α/(β=α)2, increases as α decreases.
+The expected rate of change then is π and the variance is π/α.
+\\
+By using separate shape (α) and rate (β) parameters without any independent rate parameter, these authors are making a mean rate π = α/β = 0.05 with variance of the rate being 0.0005.
+I think that the authors should just state this.
+(Also, I’m not sure that everyone knows reflexively what the mean and variance are given α and β, so it’s probably best to remind them.)}}
+
+\item{\textcolor{blue}{Lines 161 – 163: 'low evolutionary rate values allowed reduction in the number of homoplasic character changes, thus reinforcing the phylogenetic information in the matrices'
+\\
+Again, you need to make sure that the readers know how to convert Gamma parameters to rates!
+Also, is a rate of 0.05 really all that low?
+That’s a pretty typical rate per million years for fossil invertebrates.}}
+
+\item{\textcolor{blue}{Lines 163-164: 'We re-simulated every invariant characters to obtain a matrix with no invariant characters in order to better approximate real morphological data matrices.'
+\\
+Ah, but that also means that the true rate is >.05!
+Programs like RevBayes take into account ascertainment bias (unscored invariant characters), and individual runs could be limited to examine only variant characters; wouldn’t doing the latter and then accommodating ascertainment bias in the analyses be a little more realistic?
+(This is probably not a major issue, so I would not worry over much about it; however, it is worth mentioning ascertainment bias at this point and why that makes what you are doing more realistic than keeping invariant characters in the simulations.)}}
+
+\item{\textcolor{blue}{Lines 165-167: 'we only selected matrices with Consistency Indices (CI) superior to 0.26 (Sanderson and Donoghue, 1989; O’Reilly et al., 2016).'
+\\
+I have discovered that younger workers often do not know what the Consistency Index is!
+So, you might explain that this means about 3.85 (1/0.26) changes per derived character; if we have 50 binary characters, then it would be 192 changes (50/192 = 0.26).}}
+
+\item{\textcolor{blue}{Lines 200 – 202: 'Bayesian inference was run using an Mk model with ascertainment bias and four discrete gamma rate categories (Mkv 4G) with an variable rate prior an exponential (0.5) shape'
+\\
+I think that much more explanation of what these parameters and terms mean is in order!
+I note above that the authors should discuss ascertainment bias and why their simulations made sure that all characters had at least one derived taxa.
+At the very least, ascertainment bias needs to be defined here: I have found that a lot of morphological systematists do not know what it is.
+\\
+The authors use four Gamma rate categories: but how are the characters categorized?
+Again, the authors should remind the readers that this means that we are assuming that characters have a 1 in 4 chance (or however you divide the Gamma) of having the 1st, 2nd, 3rd or 4th quartile rate without assuming that any one character is fast or slow.
+\\
+Also, it should be explained that the Gamma distribution itself is varied over the runs, so that the tree will be considering new topologies, new average rates and new rate distributions.
+What this means is that the shift in the α=β parameter is drawn from an exponential distribution with a mean of 0.5.
+However, I think that it is also worth reminding the reader that every generation in the MCMC analysis will consider altered topologies, average rates and Gamma shape parameters; improvements are kept and (sometimes!) worse propositions are accepted, too, in order to better search hypothesis space.}}
+
+\item{\textcolor{blue}{Lines 202-204: The MCMC was ran over two runs of 6 chains each for a maximum of 1 x 109 generations with a sampling every 200 generations with an automatic stop if the average standard deviation of split frequency (ASDSF) fell below 0.01.
+\\
+The authors need to state that getting the ASDSF down to <0.01 is taken to mean 'convergence': that is, the individual Markov Chain has not been finding any major improvements in a long time, and that it is has been sampling the same (and 'good') part of hypothesis space for a while.
+\\
+To this end, the authors also have to define what convergence is!
+The next paragraph discusses what they do when they fail to reach convergence, but readers who do now know what convergence is (in this context) are going to be baffled.}}
+
+\item{\textcolor{blue}{Lines 215-218: 'We measured both the Robinson-Foulds distance (Robinson and Foulds, 1981) and the triplets distance (Dobson, 1975) between the trees inferred from the 'maximised', 'minimised' and 'randomised' matrices and the tree inferred from the 'normal' matrix.
+\\
+I think that a brief explanation of the metrics is in order here. If nothing else, then move lines 230 – 231 up to this paragraph so that readers will know right away why you are using both metrics.}}
+\end{itemize}
 
@@
 
-\begin{enumerate}
-\label{metric_change}
-
-\item{\textcolor{blue}{Along the lines set out by Reviewer 2, it would be wise to ensure that your metric is characterised as thoroughly as possible.
-In this regard, does it meet the following five conditions?}}
-
-After working on the required proof of the metric, we figured out that the formerly proposed Character Difference had a non-expected behaviour of being non-linear.
-We have now modified the metric to be linear (see illustration here: \url{https://github.com/TGuillerme/CharactersCorrelation/blob/master/Analysis/05-char.diffModification.pdf}).
-We thus made some major changes to the manuscript that were not suggested by the reviewers below:
-\begin{enumerate}
-    \item We change the Character Difference metric to be a scaled Hamming Distance rather than the former proposed distance.
-    \item We reran all the simulations from scratch with the inclusion of the updated Character Difference metric.
-    \item For time constraint reasons we reran 20 replicates (rather than 35 in the previous manuscript) for a total of 40 CPU years.
-    \item We propose a new interpretation to the updated results.
-\end{enumerate}
-
-Of course, we also took all the reviewers comments on board (excluding the ones rendered obsolete by the modified metrics such as the proof for the four point condition).
-
-We have also updated the Character Difference metric characterisation in the supplementary material 1.
-
-\end{enumerate}
-
-% ~~~~~~~~~~~~~~~~~~~~~~~~
-%
-% REVIEWER 1
-%
-% ~~~~~~~~~~~~~~~~~~~~~~~~
-
-\section{Reviewer 1}
-
-\subsection{Major suggestions}
-
-\begin{enumerate}
-
-\item{\textcolor{blue}{I have some concerns about the approach, however. I was somewhat confused by the CD metric. When quantifying n, how is it decided how many taxa have "comparable" characters?
-If you simulated the character for all taxa, do they not all have this character (i.e. n always equals the dataset size)?
-But more importantly, will two characters not be more similar if they share the same number of character states, independent of if they are telling you the same evolutionary history?
-It seems to me that whether or not two characters share a state space is not the relevant axis along which to measure correlation.}}
-\label{gower}
-
-Using a Gower based distance, we only compared the comparable taxa between characters.
-For example, for two characters \texttt{x = {0, 1, 2, 3}} and \texttt{y = {0, ?, ?, ?}}, the Character Difference metric will be effectively calculated using only the first taxon (\texttt{x = {0}} and \texttt{y = {0}}) and would result in a Character Difference of 0.
-As the reviewer points out, characters with less taxa in common will thus be less likely to be different.
-This is consistent with what the CD metric actually captures in terms of phylogenetic signal (the number of share splits).
-In the example above, the both characters imply at least one split (between the first taxon and the others).
-In other words, in a phylogenetic software, both characters will support the same split.
-
-As the reviewer also correctly points out, we simulated our data with no missing data so this does not affect our results.
-The precision in the manuscript was just describing the properties of the new CD metric.
-
-We've clarified how missing data can be potentially treated in the methods section:
-\textit{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
-Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).} lines 212-2178.
-
-\item{\textcolor{blue}{I also found the explanations confusing for what proportion of a matrix was repeat sampled.
-For example, in a matrix that you call maximised, n characters are replaced.
-But how many is that really?
-What is the range?
-I had a hard time putting a context to this and deciding how well I feel these simulations approximate reality.
-Without knowing this, I was having difficulty understanding why the statement "when character correlation was high (low character differences), the topology was always the furthest away from the``normal'' topology."
-should be true - does this simply mean that the same signal is being repeated over and over again, but that signal is low information relative to the topology?
-To my mind, highly correlated datasets would yield congruent trees, but that these trees may be discordant with the true tree, if the resampled characters were not representative of the true signal, or highly concordant with the true tree if they were.
-Perhaps this not being observed is a result of not performing comparisons to the true tree?}}
-
-We've added the proportion of replaced (i.e.
-duplicated) characters in each matrices for each ``minimised'', ``maximised'' and ``randomised'' scenarios along with the changes in overall Character Differences in the supplementary material 3, figures 4, 5 and 6.
-
-The maximising/minimising character differences was done \textit{a priori} without any targeted number of characters to replace to make the analysis less idiosynchratic.
-In other words, we decided to modify the randomly simulated matrices to globally increase/decrease character differences which is expected to decrease anyway in bigger matrices.
-When using a small number of character states (i.e.
-two or three), the matrices with few taxa and many characters are expected to have an overall low character difference because of the smaller combinatoric possibilities for each characters.
-
-We've added mentions to the proportion of replaced characters analysis in the supplementary and added a paragraph explaining this caveat in the discussion:
-
-\textit{Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6).
-However, this change in the amount of replaced characters did not changed the way average character differences changed across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
-This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.} lines 459-469.
-
-\end{enumerate}
-
-\subsection{Specific suggestions}
-
-\begin{enumerate}
-
-\item{\textcolor{blue}{"Phylogenetic analysis algorithms require the assumption of character independence - a condition generally acknowledged to be violated by morphological data."\\
-I don't love this sentence because it emphasizes morphology as distinct, when we recognize non-independence as a problem in basically all data types (see Huelsenbeck and Nielsen, 1996, Effect of Nonindependent Substitution on Phylogenetic Accuracy and the references therein).
-I do think you should point to a citation here; Davalos et al.
-2014 (Integrating Incomplete Fossils by Isolating Conflicting Signal in Saturated and Non-Independent Morphological Characters) is a good one.}}
-\label{molecular}
-
-We slightly rephrased the sentence in the abstract (See reviewer 3 minor revision \ref{abstract}) and expanded on the distinction between both character types this in the introduction following the reviewer pertinent biobliography suggestion:
-
-\textit{This independence is assumed for both molecular and morphological characters (Huelsenbeck and Nielsen, 1999; D\'{a}valos et al., 2014, Zou and Zhang, 2016).
-However, several distinctions can be made between the amount and the nature of the correlation.
-First, molecular and morphological data often differs in size and in the number of states analysed (molecular matrices are generally larger with mainly four states - not including missing data or indels - and morphological matrices are smaller with generally two states but may have an arbitrary number of states (up to that supported by phylogenetic software; Guillerme and Cooper, 2016).
-Second, correlations have been proposed to be stronger in morphological data (i.e.
-``concentrated'' correlations) \textit{versus} weaker and more ``spread-out'' correlations in molecular data (i.e.
-``diffused'' correlations - though some regions of the genome may be thought as more correlated than others; Huelsenbeck and Nielsen, 1999).
-Furthermore, because of this ``spread'' of correlations in molecular data, it can be harder to link the character correlation to a specific biological phenomenon (i.e.
-changes in morphological states can be more readily linked to a function cf.
-changes in nucleotide states ; see below).} lines 42-57.
-
-\item{\textcolor{blue}{"good practices".\\
-Change to "careful character coding"}}
-
-We've changed the sentence following the reviewer's suggestions lines 7.
-
-\item{\textcolor{blue}{"The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data".\\
-Evidence for this?}}
-
-We now refer to D\'{a}valos et al., 2014 and Zou and Zhang, 2016's work on the problem of morphological character correlation in contrast of molecular data.
-
-\item{\textcolor{blue}{"Especially for discrete morphological data, this assumption of independence is probably violated frequently due to the very nature of phylogenetic data: correlations are expected to occur (to some degree) when characters are depending on each other."\\
-I don't really understand what this sentence is adding.
-It feels repetitious in light of previous paragraphs, but also does little to establish why morphological data should be especially susceptible to this.
-I think there are parts of the genome where we would expect to find as much or more correlation than in morphological characters - the stem domains of rRNA genes, for example.
-I would expect the average character correlation to be weaker in DNA data, or at least more spread out in a way that makes it harder to detect.}}
-
-We removed this sentence and added the reviewer's suggestion on DNA character susceptible to be more correlated in some regions (see Reviewer 1, specific suggestion \ref{molecular}).
-
-\item{\textcolor{blue}{"Evolutionary dependence: this is the result of sets of characters co-evolving due to selection, likely related to functional links between two traits that help serve an overall lifestyle trait."\\
-Seems appropriate to cite Clarke and Middleton (2008) here.}}
-
-We added the suggested citation (line 74).
-
-\item{\textcolor{blue}{Coding dependence:\\
-This section could use a figure illustrating these types of dependence, and what they mean.
-For example: "For instance, coding for the same absence in different characters creates state transformations associated with the loss or gain of a particular character." is a fairly confusing sentence.
-I took this to mean that if you are coding, for example, tail characters, you might atomize the tail into a suite of characters, all of which are absent in organisms with no tail.
-But I'm not sure I've understood your meaning correctly.}}
-
-We do not feel that adding a figure for illustrating the matrix will improve clarity of an already long and thorough description.
-However, we've trying to clarify the description with an example as follows:
-
-\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.} lines 94-95 and \textit{In the example above, attributes of the tail (colour or length) are hierarchically dependent on the presence of the tail in the first place.} lines 101-103.
-
-\item{\textcolor{blue}{"Correlation between characters detected by the software:"\\
-I really struggled with this section.
-It seems like the problem here is still a coding problem - that the authors of these four characters have chosen four highly dependent characters to build the matrix.
-Even scoring by hand under parsimony would have problems here.
-I think what you need to decide here is what you want to communicate in this paragraph.
-Is it that software is not currently equipped to handle this issue?}}
-
-We've added a paragraph highlighting how the problem indeed comes from the fact that current software are not equipped to handle this issue:
-
-\textit{In fact, for practical reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
-In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} that seem to be more likely to reflect the split by separating artiodactyles from the cetaceans.
-In other words, this is the correlation between characters that depicts the evolutionary history of this group.
-However, the other characters depict the same history (i.e.
-the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleen, the same split would still be estimated by the software).
-One solution would be to a) weight some characters (e.g.
-giving \textbf{C3} and \textbf{C4} a lower weight) and b) make some characters dependent on each other (e.g.
-making \textbf{C3} and \textbf{C4} dependent on \textbf{C1}).
-However, this can also be a bad practice since it can easily lead to the resulting tree reflecting the researcher's prior expectations rather than the true evolutionary history.} lines 143-156.
-
-\item{\textcolor{blue}{"To assess the effects of character correlation on the accuracy of phylogenetic inference we generated a series of matrices exhibiting different levels of correlation between some characters (Fig.1 - note that each step is described in more details below):"\\
-Fig. 1 is the results, but this text implies it is a workflow diagram.}}
-
-We double checked the figure links and made sure that this mention of Fig.1 now properly points to the workflow figure.
-
-\item{\textcolor{blue}{"Additionally, we only consider differences for taxa with shared information (i.e. a Gower distance; Gower, 1971)."\\
-I'm not sure what this means. Does this mean that you did not include taxa with missing information when calculating per-character distances?}}
-
-We've clarified this sentence (see Reviewer 1, Major suggestion \ref{gower}).
-
-\item{\textcolor{blue}{"This procedure was used to treat all the characters are unordered with no assumption on the meaning of the character state"\\
-Why is this important? The text doesn't mention testing anything related to orderedness, or specifying anything related to orderedness in MrBayes.}}
-
-We removed the mention of orderdness of the characters since, indeed, it might be more confusing than helpful in this context (we didn't not simulate any ordered characters and the CD metric does not consider characters as ordered - see supplementary material 1).
-We've changed the sentence to the following:
-
-\textit{This procedure allows comparison of characters regardless of the significance given to their tokens (following the \textit{xyz} notation in Felsenstein, 2004; as used in D\'{a}valos et al., 2014).} lines 220-222.
-
-\item{\textcolor{blue}{"The morphological HKY model"\\
-Is not real.
-This is just the HKY model.
-Goloboff (2017) would dispute that this model does not favor a Bayesian method, since it is still a Markovian model.
-That O'Reilly et al.
-used this model to generate binary data does not mean one has to - you could translate each of the four nucleotides to its own discrete character.
-This would probably get even further from the Mk model assumptions than collapsing the multistate character produced via the HKY model to a binary character.}}
-
-We've enquoted the ``morphological HKY model'' and specified that it is still a Markovian model:
-\textit{Note that both models (``morphological HKY'' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
-However, the ``morphological HKY model'' is not the one} directly \textit{estimated in the phylogenetic software used in this protocol (MrBayes, see below).} lines 272-276.
-
-%TG: maybe add a bit more?
-
-\item{\textcolor{blue}{"Recently however, Wright et al. (2016) have shown that an equal rate transition is still the most present in empirical data."\\
-Of the models tested - it's plausible there are other explainers of the data. I would change this to "have shown that equal transition rate is a reasonable assumption for many empirical datasets"}}
-
-We've changed the sentence as suggest by the reviewer (lines 262-263).
-
-\item{\textcolor{blue}{"The resulting full simulation was 3.5TB big so is not shared here (though the 342 parameters are)."\\
-3.5 TB in size.}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{Most of the time when RF is normalized, it's done so that low scores are better - i.e.
-a low RF score indicates less distance to the true tree.
-I was very confused looking at the figures, until I realized this was not how normalizing had been performed.
-}}
-
-We've changed the labels of the figures to NTS(RF) and NTS(Tr) to avoid any confusions.
-
-\item{\textcolor{blue}{"However, the effect of these sources of correlation was out of the scope of this study and would have required a posteriori changes to the matrices which are - when using empirical data - at best bad practice and at worth dishonest."\\
-I'm not sure I follow this.}}
-\label{dishonest}
-
-We meant that modifying phylogenetic matrices after the phylogenetic inference is a way to test the effect of the sources of correlation but is probably not a good practice in systematics.
-Although re-weighting characters based on consistency or retention indices can be used to objectively improve the topology, changing the matrix's characters \textit{a posteriori}, it can be seen as dishonest to remove characters because they ``don't give the right signal'').
-We've removed the end of this sentence however to avoid any further confusion.
-
-\item{\textcolor{blue}{"However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree. "\\
-I understand why you did this, but I think you should also compare normal to true.
-That will give us some idea of the scope of how close we can get to true, under these conditions, and will help contextualize the results.}}
-
-We've added the comparisons between the ``normal'' and ``randomised'' trees to the ``true'' tree in the supplementary material 3, figures 9 to 11.
-We've added a reference to these figures in the main text:
-
-\textit{However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.} lines 347-3478.
-
-\end{enumerate}
-
-% ~~~~~~~~~~~~~~~~~~~~~~~~
-%
-% REVIEWER 2
-%
-% ~~~~~~~~~~~~~~~~~~~~~~~~
-
-\section{Reviewer 2}
-
-\subsection{Major suggestions}
-
-\begin{enumerate}
-
-\item{\textcolor{blue}{1. To be a metric, there are several properties that need to be satisfied. The proposed function
-is positively valued and symmetric (copied below from line 178): [...] But, the following metric property does not hold for all x; y:}}
-\label{proof}
-
-We are really grateful that this reviewer picked up this flaw in our mathematical demonstration of the CD metric.
-We have reviewed the formula of the CD metric following an unexpected non-linear behaviour (see comment \ref{metric_change}).
-Furtheremore, we'd like to highlight the following point:
-
-In fact, before being plugged into the CD equation, characters need to be modified following Felsenstein's \textit{xyz} notation.
-In other words, because the character state tokens are non-metric values if the characters are unordered they can be translated in a standard way so that the first token in the character is renamed to ``1'', the second to ``2'' etc.
-For example, in a 3 state character for four taxa: \texttt{X = {1, 0, 2, 1}}, the character tokens that are integers representing the states could equally have been \texttt{X = {E, A, I, E}} or \texttt{X = {blue, red, absent, blue}}.
-In order to numerically compare X to any other character it is thus more simple to translate both characters in a standard way such as the first token is replaced by the symbol ``1'', etc.
-
-For example if comparing two characters \texttt{X = {1, 0, 2, 1}} and \texttt{Y = {2, 1, 0, 2}}, the absolute difference would be \texttt{X - Y = {-1, -1, 2, -1}}, or when using a Hamming distance \texttt{X - Y = {1, 1, 1, 1}} showing maximum difference between both characters.
-However, because these tokens are non-numeric and effectively represent the same phylogenetic split, both characters are translated to \texttt{X' = {1, 2, 3, 1}} and \texttt{Y' = {1, 2, 3, 1}} leading to a difference of \texttt{X' - Y' = {0, 0, 0,0}}.
-
-In the reviewer's demonstration, \texttt{x  = {1}} and \texttt{y  = {0}} thus become \texttt{x' = {1}} and \texttt{y' = {1}}.
-Hence $CD_{(x',y')} = 0 \Leftrightarrow x' = y'$.
-This is also reflected in the new $CD$ formula that makes comparing two characters for a single taxon nonsensical: the denominator is equal to $0$ if $n = 1$.
-
-Furthermore, the CD metric, used in this manuscript as a proxy for character correlation actually measures the difference between in characters in terms of phylogenetic signal.
-The character \texttt{X = {1, 0, 2, 1, 0}} described above implies two splits (and of course same goes for the translated \texttt{X' = {1, 2, 3, 1, 2}}).
-Another character \texttt{Y = {1, 1, 1, 2, 2}} will imply only one split, not shared with \texttt{X} and a character \texttt{Z = {0, 2, 1, 0, 2}} (\texttt{Z' = {1, 2, 3, 1, 2}}) would imply the same split as \texttt{X} (hence, $CD_{(X',Z')} = 0$ and $CD_{(X',Y')} = 0.75$).
-In other words the probability of X and Z generating conflicting split signal is null and this probability for X and Y is of 0.75.
-
-Following these two points, the reviewer counter example $CD_{(x,y)} = 0$ means that 1) $x'$ and $y'$ are identical and the characters $x$ and $y$ imply the same number of splits in a tree (here no splits since only one taxon is used - however, this would hold for any $n$ number of taxa).
-We have clarified these two points in the manuscript.
-
-We've correct the supplementary materials 1 and added more explanations on (1) the character translation, (2) the four-point metric proof and (3) on how to interpret the $CD$ metric.
-Additionally, we added more precisions in the text:
-
-\textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines 201-202.
-
-\item{\textcolor{blue}{While some details of the simulation study are included, some important ones are missing.
-In particular, in the design of the study, how did you avoid or mitigate errors introduced by the programs simulating data?
-All of the programs use random-number generators and have their own biases.
-How have you accounted for this in the study of design to conclude that your results are due to signal, and not bias from the programs used?}}
-
-We do agree that generating true randomness in current software is tricky but we are not sure how to answer this comment.
-The three first steps (generating the tree, the matrices and modifying the matrices) were using the default random number generator implemented in R, the Mersenne Twister algorithm that has a cycle of $2^{32}-1 = 4.2e^{9}$ unsigned random integers (see random numbers documentation in R: \texttt{?.Random.seed} - we used the implementation from \texttt{R} v3.5 that differs in the starting seed from the new \texttt{R} v3.6).
-The last step (generating the trees) was using both the random number generated by default my MrBayes and PAUP* based on the seeds from the different HPC we used (the Imperial College London Cluster Services and the Australian QRIScloud Awoonga cluster).
-
-\begin{enumerate}
-    \item \textbf{The generation of the ``true'' tree.}
-    The generation of the true tree involves the drawing of two random parameters (birth and death parameters) and then builds the tree by drawing a probability of branching (birth) or extinction (death) at each time point until it reaches the desired number of taxa.
-    This means at least $2 + 2tn$ where $t$ are the evaluated time points and $n$ the number of required taxa.
-    For all 60 trees we thus have $120 + 500t$ draws.
-This means this whole step would cycle through all random numbers only if $t > 8.4e^{6}$.
-As a magnitude of comparisons when $t$ is one order of magnitude lower, $t = 1e^{5}$ it will typically lead to trees with $>> 1e{4}$ taxa (i.e.
-2 orders of magnitude bigger than our biggest trees).
-
-    \item \textbf{The generation of the discrete morphological characters (each 100, 350 and 1000).}
-    The number of seeds involved in the generation of each model depends on the two models with a total of $5 + (2n-2) = 2n + 3$ random numbers for the HkY-binarised model (4 states proportions + 1 rate + number of branches) and $2 + (2n-2) = 2n$ for the Mk model (1 states numbers + 1 rate + number of branches).
-    If if we simplify the seed calculations by saying only the HKY-model is chosen, we have at maximum $20(2n + 3)c$ random numbers; where $c$ is the number of characters.
-    This results in a maximum of $1.4e{7}$ random numbers ($<< 4.2e^{9}$ total random numbers).
-
-    \item \textbf{The removal of the characters (for the ``randomised'' matrices).}
-    This step is also variable but we can simplify it by assuming the extreme case where 1) all characters are selected to be replaced and 2) they are all replaced by all characters.
-This results in a maximum of $20(5c)$ random numbers per matrix (2 for both replacements and 3 for each ``minimised'', ``maximised'' and ``randomised'' matrices).
-    This results in a maximum of $1.4e{5}$ random numbers ($<< 4.2e^{9}$ total random numbers).
-\end{enumerate}
-
-Therefore, if cumulating all the seeds used in the generation of the matrices (trees + characters) and the modifications of the matrices is likely to be no more than $6.4e^{7}$.
-Therefore we do not believe that the random number generation was a proper since we sampled two orders of magnitude less random numbers than the total ones available.
-
-\item{\textcolor{blue}{How did you ``randomly'' modify your matrices? Did you sample under some distribution? If so, which one?}}
-
-Every replaced characters were randomly selected randomly from the total sample of characters (i.e.
-randomly sampling from the uniform integers distribution (1, $c$); where $c$ is the number of remaining characters).
-We have now specified this in the text:
-
-\textit{Each replaced characters (i.e. the characters with a weight of +$1$) were randomly selected by sampling them randomly from the list of the remaining characters.} lines 298-300.
-
-\item{\textcolor{blue}{It is not obvious why generated character sequences are representative of biological sequences or sample the space overall.
-Were the parameters chosen for the HKY model to fit a particular evolutionary scenario?
-Are they meant to capture all of the space?}}
-
-There is actually a vigorous debate surrounding how to generate biologically meaningful morphological characters \citep[][]{OReilly20160081,goloboff2017weighted,puttick2017uncertain,GoloboffEmpirical,OReillyEmpirical}.
-This ongoing debate, however, focuses mainly on inference methods where data simulation is a tangential aspect of it.
-Therefore, we decided to base our simulation protocol on the empirical observations showing that half of 206 analysed datasets supported a M$k$-like model \citep[][Fig.
-3]{Wright01072016} and the other half being a combination of models not explicitly described here but that are believed to be covered with the HKY-binarised model \citep{puttick2017uncertain,OReillyEmpirical}.
-In our simulations, the idea was to cover a broad range of morphological characters (with variable state numbers, evolution and transition rates) in order to capture a realistic image of the ``discrete morphological character space''.
-
-\item{\textcolor{blue}{Similarly, why did you choose 0.85 for the Mk model? Does it represent some important set of character sequences? How does it represent the space in general?}}
-
-This 0.85 proportion of binary characters and 0.15 of three state ones was extracted from the 100 empirical matrices analysed in \cite{Guillerme2016146} (Supplementary appendix 1.
-Appendix A: Tree inference) and \cite{ZouConvergence}.
-We've specified this in the text:
-
-\textit{We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters (from empirical observations in: Guillerme and Cooper, 2016b; Zou and Zhang, 2016).} lines 257-259.
-
-\item{\textcolor{blue}{There are statements that seem contradictory, like (line 335): ``Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>$ 0.95 and different if they had an overlap probability $>$ 0.05.''}}
-
-We fixed this typo. We considered them different if they had an overlap probability $<$ 0.05.
-
-\item{\textcolor{blue}{It is not clear how the conclusions were reached.
-For example, line 381-384: ``Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table 5 - median NTSRF of 0:828 and 0:679 respectively) but not in terms of individual taxon placements (Table 5 - median NTSTr of 0:738 and 0:601 respectively).'' Why is the 0:828 - 0:679 = 0:149 significant for this test and 0:738 - 0:601 = 0:137 is not?}}
-
-The significances was referring to the results in Table 5 (Wilcoxon test $p$ value of 0 and 0.08 and Bhattacharrya Coefficient of 0.891 and 0.984 respectively) and the median values were extracted from the supplementary material 3 Table 5.
-We now detailed the values and where they are present for each of the tests to avoid confusion.
-
-\end{enumerate}
-
-% ~~~~~~~~~~~~~~~~~~~~~~~~
-%
-% REVIEWER 3
-%
-% ~~~~~~~~~~~~~~~~~~~~~~~~
-
-\section{Reviewer 3}
-
-\subsection{Major suggestions}
-
-\begin{enumerate}
-
-\item{\textcolor{blue}{I feel that the manuscript is certainly worth publishing in Systematic Biology, although I worry that the simulation procedure used by the authors limits their ability to thoroughly test of the effect of between character correlation on the accuracy of estimating phylogenetic trees.
-The authors note this themselves in the discussion, but I feel more should be made of this point.
-If the simulations utilised an explicit distribution of character covariance then I think that interpretation of the results would have been more straightforward.}}
-\label{stramlined}
-
-As noted in the introduction, the aim of this manuscript was mainly to see does different levels of correlation between discrete morphological characters affect tree topology.
-We also review different sources of correlation for characters and discuss how the multidimensional complexity of a phenotype (and the correlations therein) is not reflected in two dimensional character matrices.
-We do agree with the reviewer that a thorough analysis of the variance/covariance structure of simulated ``discretised'' characters (see reviewer 3, major suggestion \ref{hiden}) would have broad more insight into the nature of character correlation on phylogenetics inference in general (regardless of the type of character and correlation) but due to the length required to run the current ``simplistic'' simulation, we feel that this would be out of the scope of this manuscript.
-
-We have now clarified in the manuscript what our objectives were by streamlining the last paragraph of the introduction (lines 159-170).
-We have also explicitly added in the supplementary material 3, figs 1, 2 and 3 how the character difference changes between all matrices compared to the ``normal'' matrix.
-
-\item{\textcolor{blue}{Another worry I have is that the implied covariance at the tips does not necessarily reflect covariation over the full phylogeny.
-It is conceivable that two characters can appear to have near perfect correlation at the tips, but arrived at these states through completely independent pathways, particularly when simulating with a continuous time Markov chain where unobserved transitions to a new state and back to the original state are possible.}}
-\label{hiden}
-
-We agree with this reviewer's comment which can be especially thought to be the case for discrete characters with a fast rate of evolution.
-Such characters will likely swap between states along the branches following a hidden Markov chain.
-
-However, we feel that modelling this process might not be useful in this specific study:
-First, in empirical data two characters can be strongly correlated for biological and evolutionary reasons (e.g.
-the absence of an astragalus and the presence of baleen in table 1) leading to a strong correlation of both characters at the tips (or even nodes) but does not mean that both character's hidden Markov chain are correlated.
-Second, by comparing characters at any time point rather than simply at the tips (which can be problematic since our trees were not all ultrametric) we would probably not be able to compare the effect of the ``true'' characters' correlation since this information will not be used by the phylogenetic software (MrBayes and PAUP*) that will only use tip values.
-
-We've added this remark in the caveats section:
-\textit{It is also worth noting that character similarity at the tips does not necessarily reflects identical evolutionary history.
-This may be specifically true when simulating characters based on the M$k$ model where the resulting tip states does not necessarily reflects variation of the state along the branches (e.g.
-Revell, 2014).
-However, it is worth noting that in empirical studies, such state changes along the branches are nearly always empirically unknown (even though they can be estimated by Maximum Likelihood).} lines 563-568.
-
-\item{\textcolor{blue}{It seems that the results reported by the authors are inherently driven by the phylogenetic signal present in their ``bootstrapped'' datasets.
-This effect would be present in any approach to the simulation of covarying characters, but as the authors do not explicitly simulate covariance but instead rely on a proxy they cannot be certain that their results are driven by true character covariation or purely a restructuring of phylogenetic signal by resampling.
-In a simulation framework in which covariance is actively enforced and controlled it would be possible to assess other aspects of character covariance that seem important for empirical datasets, such as correlation within morphological compartments.
-I think without tests of these sort we cannot be completely sure of the effect of character covariance on topology estimates.
-Despite this, the authors present an overview of the effect of character covariance at large, and I certainly think that this is a worthwhile contribution.}}
-
-We do agree with this reviewer's comment and believe that a study looking at difference in covariation levels between character in different morphological modules would be really useful (e.g.
-are cranial character more correlated than postcranial)?
-In our simulation protocol we effectively only simulated characters from a single ``module'' since they were generated based on the same topology (and branch length) and the same distribution (rates where sampled from a single distribution).
-Therefore, we believe that the correlation approximated by the $CD$ metric was actually explicitly simulated through our protocol.
-This is also reflected by the fact that, in practice, when increasing the correlation (reducing character difference), rather few characters needed to be duplicated (see supplementary material 3, figures 1 to 6).
-
-We hope to expand this study in the future by create different groups of characters sharing the same overall correlation by generating ``bunches'' of characters based on different rates (to simulate rate homogeneity) and different trees (branch length and topology - to simulate variation in ``gene'' trees).
-
-\item{\textcolor{blue}{A discussion of the ``effective'' number of characters might also be warranted.
-When maximising character covariance this likely requires the resampling of the same handful of characters multiple times, each of which provides the exact same phylogenetic signal.
-The redundancy of these characters seems likely to influence the accuracy of estimates.
-The choice of characters that are sampled by the bootstrap procedure will therefore constrain the accuracy of topology estimates.
-If the resampled matrix consists of a handful of characters with high consistency with the true tree, sampled many times over, then the tree estimate will be good (for at least part of the tree) and covariance will be high.
-Similarly, if the handful of regularly resampled characters are mainly inconsistent (with the true tree) then tree estimates will be poor but the implied covariance will also be high.
-I think this is evident from the reduced performance in larger matrices, where there will be a significant reduction in phylogenetic signal when attempting to maximise covariance.}}
-
-We've added the following part to the discussion:
-
-\textit{Overall, it is expected that minimising character difference (increasing character correlation) is likely to have a greater effect on recovering the ``correct'' topology since most characters will provide the same phylogenetic signal.
-If the ``minimised'' matrix consists of a handful of characters with high consistency with the ``best'' tree, then the tree estimate will be good (for at least part of the tree).
-Similarly, the characters in the ``minimised'' matrix are mainly inconsistent with the ``best'' tree then the estimates will be poor.} lines 475-481.
-
-\item{\textcolor{blue}{Overall, I think that the authors provide an overview of the effect of covariance on the estimation of phylogeny, but without an explicit simulation of patterns of covariation we do not yet have a definitive understanding of the effects of between character covariance.}}
-
-We definitely agree with this statement and are working on a follow up project with more explicit patterns of covariation.
-However, one major bottleneck in such analysis is our ability to actually simulate discrete morphological characters (regardless of their pattern of covariance) that are meaningful in reflecting the actual patterns observed in empirical discrete morphological characters.
-Thankfully, the debate between Goloboff et al. and O'Reilly et al. will provide us with some more insightful and realistic methods to simulate discrete morphological characters.
-
-Secondly, as we highlighted throughout the manuscript, the main mechanisms that generate phylogenetic signal and variance/covariance between characters are the intra-organismal dependence and the evolutionary dependence described in the introduction.
-Both of which can unfortunately only be observed/analysed, \textit{a posteriori} based on a phylogenetic hypothesis.
-
-\end{enumerate}
-
-\subsection{Minor revisions:}
-
-\begin{enumerate}
-
-\item{\textcolor{blue}{2:2 – Phylogenetic algorithms don't necessarily require independence, they simply assume it.}}
-\label{abstract}
-
-We've rephrased this sentence to:
-\textit{Phylogenetic analysis algorithms assume character independence} line 2.
-
-\item{\textcolor{blue}{3:25 – I think that resurgence should be taken out of quotation marks, as no reference is supplied for the use of the phrase}}
-
-We've removed the quotes.
-
-\item{\textcolor{blue}{3:29 – In what ways are morphological data inferior? I think this needs to be qualified.}}
-
-We've added the following justification (and toned down the sentence):
-\textit{Although morphological character data are sometimes considered inferior to molecular sequence data due to the difficulties of collected them and their sometimes subjectivity, they are often the only source of phylogenetic data for extinct species.} lines 28-30.
-
-\item{\textcolor{blue}{4:45 – depend $->$ dependent}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{7:108 – inferences $->$ inference}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{8:134 – does $->$ do}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{8:135 – Do we actually expect high levels of correlation to lead to inaccurate phylogenies?
-I think this depends on what is considered to be a high level of correlation and how that correlation is structured.
-Imagine 100 characters covarying strongly with each other in blocks of 10, but no constraint on the level of correlation across all 10 blocks.
-This would provide an ``effective'' character number of ~ 10, which could be enough to resolve a small tree.}}
-
-We have removed this section (see Reviewer 3, Major Suggestion \ref{stramlined})
-
-\item{\textcolor{blue}{12:190 – are $->$ as}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{15:240 – The reference for the 0.26 cut off used by these authors should also be cited}}
-
-We've added the reference Sanderson \& Donoghue 1989.
-
-\item{\textcolor{blue}{18:286 – an $->$ and}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{18:286 – converged $->$ converge}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{18:298 – Surely strict isn't correct here as you are summarising a posterior sample? Is it not just a Majority rule tree?}}
-
-We removed the term ``strict''.
-
-\item{\textcolor{blue}{19:310 – where $->$ were}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{19:311 – where $->$ were}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{20:341 – big $->$ in size}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{29:428 – really $->$ very}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{29:434 – delete `a'}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{30: 450 - $<$ $->$ $>$}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{30:457 – a score}}
-
-We fixed this typo throughout the paragraph.
-
-\item{\textcolor{blue}{30:457 – on $->$ one}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{30:469 – delete 'b'}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{31:479 – of $->$ off}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{31:487 – and the use of models that utilise tree structure}}
-
-We fixed this typo.
-
-\item{\textcolor{blue}{32:492 – worth $->$ worst}}
-
-We removed this sentence (see Reviewer 1 specific suggestion \ref{dishonest}).
-
-\item{\textcolor{blue}{32:511 – citation includes first name ``David''}}
-
-We fixed this reference.
-
-\item{\textcolor{blue}{33:522 – I think the last sentence is redundant as the point has been made well before in the text.}}
-
-We removed this sentence.
-
-\end{enumerate}
-
 \bibliographystyle{sysbio}

### 2020-04-21 Functions/compatibility.R
compatibility init

@@ -1,0 +1,86 @@
+#' @title Character compatibility functions
+#'
+#' @description 
+#'
+#' @param 
+#' 
+#' @examples
+#'
+#' @seealso
+#' 
+#' @author Thomas Guillerme
+#' @export
+
+# This code is modified from https://datadryad.org/stash/dataset/doi:10.5061/dryad.sh8b4 (functions pairCompat and compCount)
+
+
+
+
+set.seed(4)
+## A random tree with 15 tips
+tree <- rcoal(15)
+## Setting up the parameters
+my_rates = c(rgamma, rate = 10, shape = 5)
+## Mk matrix (15*50) (for Mkv models)
+matrix <- sim.morpho(tree, characters = 50, model = "ER", rates = my_rates,invariant = FALSE) 
+
+
+pair.compat <- function(i, j)    {
+
+    ## Get the tokens
+    tokens <- unique(c(i, j))
+
+    ## The number of possible pairwise combinations
+    n_tokens <- length(tokens)
+    expect_combn <- ncol(combn(1:n_tokens, 2))*2 + n_tokens
+
+    ## Intersection for any token
+    get.tokens <- function(token, character) return(which(character == token))
+    tokens_i <- unlist(sapply(tokens, get.tokens, character = i, simplify = FALSE))
+    tokens_j <- unlist(sapply(tokens, get.tokens, character = j, simplify = FALSE))
+    keeps <- intersect(tokens_i, tokens_j)
+
+    ## Get the observed combinations:
+    obs_combn <- unique(paste0(i[keeps], j[keeps]))
+
+    ## Measure compatibility
+    ## If the number of observed combinations is lower than the expected combinations, the characters are compatible
+    return(ifelse(length(obs_combn) < expect_combn, 0, 1))
+}
+comp.count <- function(matrix, pairwise)   {
+    compatibilities <- apply(pairwise, 2, function(x) pair.compat(matrix[,x[1]], matrix[,x[2]]))
+    return(sum(compatibilities))
+}
+pair.comp <- function(matrix, type = "compatible") {
+    pairwise_comparisons <- combn(1:ncol(matrix), 2)
+    incompatibilities <- comp.count(matrix, pairwise_comparisons)
+
+    if(type == "incompatible"){
+        return(incompatibilities)
+    } else {
+        return(ncol(pairwise_comparisons - incompatibilities))
+    }
+}
+
+
+pair.comp(matrix, "incompatible")
+
+
[…]

### 2020-04-21 Functions/compatibility.R
pseudo code version

@@ -77,10 +77,10 @@ pair.comp(matrix, "incompatible")
 intersect(c(which(i == "0"), which(i=="1")), c(which(j == "0"), which(j=="1")))
 paste(i)
 
-pairCompat.orig <- function(i, j, count) {
+pairCompat <- function(i, j, maximum.pairs = 2) {
     Keeps <- intersect(c(which(i == "0"), which(i=="1")), c(which(j == "0"), which(j=="1"))) #TG: This line basically removes NAs (making sure not comparing NAs, won't intersect)
     Combns <- paste(i[Keeps], j[Keeps]) #TG: This line combines both characters
     Pairs <- unique(Combns) #TG: This line counts the unique number of comparison pairs
-    Incompatibility <- ifelse(length(Pairs) < count, 0, 1) #TG: If there are less pairs than the maximum (count), character is compatible
+    Incompatibility <- ifelse(length(Pairs) < maximum.pairs, 0, 1) #TG: If there are less pairs than the maximum (count), character is compatible
     return(Incompatibility)
 }

### 2022-03-24 Writing/Palaeontology/draft.tex
Update draft.tex

@@
 \usepackage{pdflscape}
-
+\usepackage{xcolor}
 \usepackage{color,soul}
@@
 
+% TODO: change the affiliations
+
 \end{center}
@@
 Phylogenetic analysis algorithms assume character independence - a condition generally acknowledged to be violated by morphological data.
-Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
+Correlation between characters can originate from intra-organismal features, \color{blue}{evolutionary dependence} or forced by particular character-state coding schemes.
 Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be diminished \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
@@
 \subsubsection{Character Difference ($CD$)}
-\begin{equation}
-    CD_{(x,y)} = \frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}
-\end{equation}
+% \begin{equation}
+%     CD_{(x,y)} = \frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}
+% \end{equation}
 

### 2022-03-24 Writing/Palaeontology/draft.tex
Intro updated

@@
 While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any investigations into their statistical properties.
-As morphological data are likely to continue to play an extensive role in phylogenetic analysis (namely for neontological analysis), it is essential to understand the circumstances under which morphological data might be expected to lead to erroneous phylogenetic inferences.
+%MB: I think there are a few classic papers we could cite here.
+The dismissal tends to arise with the same 'workflow': calculate a molecular tree; calculate a morphological tree; see that they are different; shit on the morphological tree because of this that and the other, but never bother to check if the data actually suggest these phenomena.
+TG: which ones?
+As morphological data are likely to continue to play an extensive role in phylogenetic analysis,  %MB: cite Nature paper here. TG: which one?
+it is essential to understand the circumstances under which morphological data might be expected to lead to erroneous phylogenetic inferences.
 This opens up possibilities for analysing problematic datasets and possibly proposing new confidence measures in phylogenetic datasets.
@@
 The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-However, several distinctions can be made between the amount and the nature of the correlation.
-First, molecular and morphological data often differ in the size and number of states analysed.
-Second, correlations have been proposed to be stronger in morphological data but weaker and more ``spread-out'' in molecular data \citep[][]{huelsenbeck1999effect}.
+Non-independence of characters implies that some suites of characters are expected to transform in highly correlated ways.
+Not only does this violate the assumptions of phylogenetic analysis, but implies that if convergence has occurred in correlated sets of characters, then this would lead to spurious cladistic groupings.
 Here we distinguish three different modes of forms of character correlation (described in more details in \citealt{wilkison1992ordered,wilkinson1995character,wilkinson1995coping}):
@@
 
-    \item \textbf{Coding dependence:} this is the results of researcher methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
+    \item \textbf{\color{blue}{Logical} dependence:} this is the results of researcher methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
     For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail \citep{wilkinson1995coping,BrazeauNA}.
+    \color{blue}{Note that there are a lot of alternative methodological choices for this example \citep[e.g.
+using inapplicable tokens; ][]{Brazeau2011} and that these choices affect results more or less depending on the estimation method (parsimony methods contrast with probabilistic methods were ancestral conditions are respectively fixed or not).}
 \end{itemize}
@@
 \noindent Of course, the three sources of dependence can have an interaction: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding.
-Coding dependence, however, is easily distinguished prior to phylogenetic inference, while intra-organismal and evolutionary dependence are much harder to parse.
+\color{blue}{Logical dependence}, however, is easily distinguished prior to phylogenetic inference, while intra-organismal and evolutionary dependence are much harder to parse.
 Meanwhile, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{de2015parsimony,BrazeauNA}.
 Intra-organismal dependence requires more detailed, often extremely time-consuming studies (possibly beyond the limits of available technology).
-Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree, and is best determined by independent character sets.
+Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree \color{blue}{(but see useful methods from \citealt{Pagel1994} and \citealt{billet2019serial}), and is best determined by independent character sets.
+The latter is commonly hard to get because of the nature of the data: some common types of data (e.g.
+fossil data) offer very few options of independent character sets and no biological datasets can really be independent.
+It is also worth noting that the evolutionary and intra-organismal dependences induced correlations are by no means restricted to the use of discrete morphological data.
+In fact, by essence, these correlations appear to any type of data whether continuous (e.g.
+linear measurements or morphometric geometric measurements) or non-morphological (e.g.
+behaviour, or - much more commonly used - molecular data).}
 
@@
 Although it has been tackled empirically (for morphological data \citealt{Davalos01072014}, and molecular data \citealt{ZouConvergence}), the question has never been explored through a thorough simulation framework.
-sSimulations can allow the design of datasets with controlled levels of correlation and for which the correct expected answer is known by design.
+Simulations can allow the design of datasets with controlled levels of correlation and for which the correct expected answer is known by design.
 Furthermore, using simulations will allow drawing generalities on the effect of character correlation on phylogenetic inference in contrast to potential idiosyncrasies of empirical datasets.
@@
 \begin{enumerate}
-    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 or 150 taxa, for 100, 350 or 1000 characters, hereafter called the ``normal'' matrices.
+    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 or 150 taxa, for 100, 350 or 1000 characters, hereafter called the \color{blue}{\textit{unperturbed matrices}}.
 This step resulted in 9 matrices.

### 2022-03-24 Writing/Palaeontology/draft.tex
generating matrices bit

@@
 Phylogenetic analysis algorithms assume character independence - a condition generally acknowledged to be violated by morphological data.
-Correlation between characters can originate from intra-organismal features, \color{blue}{evolutionary dependence} or forced by particular character-state coding schemes.
+Correlation between characters can originate from intra-organismal features, \textcolor{blue}{evolutionary dependence} or forced by particular character-state coding schemes.
 Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be diminished \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
@@
 
-    \item \textbf{\color{blue}{Logical} dependence:} this is the results of researcher methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
+    \item \textbf{Logical dependence:} this is the results of researcher methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
     For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail \citep{wilkinson1995coping,BrazeauNA}.
-    \color{blue}{Note that there are a lot of alternative methodological choices for this example \citep[e.g.
+    \textcolor{blue}{Note that there are a lot of alternative methodological choices for this example \citep[e.g.
 using inapplicable tokens; ][]{Brazeau2011} and that these choices affect results more or less depending on the estimation method (parsimony methods contrast with probabilistic methods were ancestral conditions are respectively fixed or not).}
@@
 \noindent Of course, the three sources of dependence can have an interaction: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding.
-\color{blue}{Logical dependence}, however, is easily distinguished prior to phylogenetic inference, while intra-organismal and evolutionary dependence are much harder to parse.
+Logical dependence, however, is easily distinguished prior to phylogenetic inference, while intra-organismal and evolutionary dependence are much harder to parse.
 Meanwhile, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{de2015parsimony,BrazeauNA}.
 Intra-organismal dependence requires more detailed, often extremely time-consuming studies (possibly beyond the limits of available technology).
-Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree \color{blue}{(but see useful methods from \citealt{Pagel1994} and \citealt{billet2019serial}), and is best determined by independent character sets.
-The latter is commonly hard to get because of the nature of the data: some common types of data (e.g.
+Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree (but see useful methods from \citealt{Pagel1994} and \citealt{billet2019serial}), and is best determined by independent character sets.
+\textcolor{blue}{The latter is commonly hard to get because of the nature of the data: some common types of data (e.g.
 fossil data) offer very few options of independent character sets and no biological datasets can really be independent.
@@
 
-Intra-organismal and evolutionary dependences are inherent parts to Evo-Devo and macroevolutionary studies and best practices to avoid coding-induced dependences are commonly known and applied.
+\color{black}{Intra-organismal} and evolutionary dependences are inherent parts to Evo-Devo and macroevolutionary studies and best practices to avoid coding-induced dependences are commonly known and applied.
 However, eventually, all these characters, whether they are independent or not are analysed through phylogenetic inference software that are blind to these distinctions.
@@
 \begin{enumerate}
-    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 or 150 taxa, for 100, 350 or 1000 characters, hereafter called the \color{blue}{\textit{unperturbed matrices}}.
+    \item \textbf{Simulating matrices}: \textcolor{blue}{using a randomly generated topology (the \textit{starting tree})} we simulated discrete morphological matrices with 25, 75 or 150 taxa, for 100, 350 or 1000 characters, hereafter called the \textcolor{blue}{\textit{unperturbed matrices}}.
     This step resulted in 9 matrices.
 
-    \item \textbf{Modifying matrices}: we changed the ``normal'' matrices by modifying the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least different or most different characters and replacing them randomly by characters from the remaining set.
-    We also randomised the ``normal'' matrices without bias towards maximising or minimising character differences (hereafter called the ``randomised'' matrices).
+    \item \textcolor{blue}{\textbf{Inducing correlation}: To induce different levels of unpredicted correlation, we used the unperturbed matrices as the basis of a series of resampling manipulations that either enriched or depleted the matrix with correlated characters.
+    We modified the matrices by running three types of bootstrapping methods (i.e.
+removing a set of characters by a set of remaining ones): 1) a classical bootstrap (randomly removing/replacing characters) as our null change in character correlation (hereafter, the \textit{randomised matrices}); 2) a biassed bootstrap towards correlation (removing/replacing the most different characters) to generate matrices with maximised correlation (the \textit{maximised correlation matrices}) and 3) a biassed bootstrap away from correlation (removing/replacing the least different characters) to generate the \textit{minimised correlation matrices} (i.e.
+with a minimised amount of correlation).}
 
-    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both maximum parsimony and Bayesian inference.
-    Hereafter, the resultant topologies are called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
+    \item \textcolor{blue}{\textbf{Inferring topologies}: we inferred the topologies from the unperturbed, randomised, maximised correlation and minimised correlation matrices using both maximum parsimony and Bayesian inference.
+Hereafter, the resultant topologies are called the unperturbed, randomised, maximised correlation and minimised correlation trees.}
 
-    \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on the recovered topology.
+    \item \textcolor{blue}{\textbf{Comparing topologies}: finally, we compared the unperturbed tree to the trees recovered from analysis of the three different dataset types (randomised, maximised and minimised correlation) to measure the effect of character correlation on the recovered topology.}
 
@@
 All steps were replicated 20 times and are described below in more detail.
+\textcolor{blue}{Note that we compare to the topologies from the randomised, maximised and minimised correlation trees to the unperturbed tree rather than to the starting tree because differences with the starting tree can be due to other confounding factors which are out of the scope of this study (e.g.
+the algorithms used to simulate characters).}
 
@@
 
+\subsection{Simulating discrete morphological matrices}
+
+To simulate the matrices we applied a protocol very similar to \cite{Guillerme2016146} using \texttt{R} \citep{R}.
+First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution with $\lambda$ $>$ $\mu$ \citep[\texttt{diversitree} v0.9-8;][]{fitzjohndiversitree2012} and saving the trees after reaching either 25, 75 or 150 taxa (see supplementary material 3, Figs 1, 2 and 3).
+For each tree, we arbitrarily set the outgroup to be the first taxon (alphabetically) thus effectively rooting the trees on this taxon.
+These trees are hereafter called the \textcolor{blue}{\textit{starting trees}}.
+We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
+\begin{itemize}
+    \item The ``morphological HKY-binary'' model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
+    \textcolor{blue}{This allows the changes in character states to vary without specifying which change is more likely, hence expecting an equiprobable change in character state in all directions.}
+    \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
+    We drew the number of character states with a probability of $0.85$ for binary characters \textcolor{blue}{(using a $2\times2$ Q matrix)} and $0.15$ for three state characters (\textcolor{blue}{using a $3\times3$ Q matrix);} \citealt{Guillerme2016146,ZouConvergence}).
+%TG: this Q matrix bit is very jargony and unsincere...
+\end{itemize}
+
+\noindent For each character, both models (``morphological HKY-binary'' or M$k$) were chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
+\textcolor{blue}{This simulated, for each character, a variation in the change of character state at different rates while ensuring that these rates were drawn from the same distribution, resulting in an overall mean character rate of 0.05 with a variance of 0.0005.}
+These low evolutionary rate values allowed reduction in the number of homoplasic character changes, thus reinforcing the phylogenetic information in the matrices.
+\textcolor{blue}{We re-simulated every invariant character to obtain a matrix with no invariant characters (making the effective mean rate to $>0.05$).
+Although invariant characters are important for effectively measuring the true evolutionary rates, many empirical matrices do not contain invariant characters which is in turn taken into account by software that accommodate for this ascertainment bias.}
+We re-simulated every invariant characters to obtain a matrix with no invariant characters in order to better approximate real morphological data matrices.
+To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ (\textcolor{blue}{expecting at least 3.85 (1/0.26) changes per derived character}; \citealt{sanderson1989patterns,OReilly20160081}).
+
+For each tree with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
+The matrices were generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
+We repeated this step 20 times resulting in 180 ``normal'' morphological matrices.
+
 \subsection{Measuring differences between characters}
@@
 \subsubsection{Character Difference ($CD$)}
-% \begin{equation}
-%     CD_{(x,y)} = \frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}
-% \end{equation}
+\begin{equation}
+    CD_{(x,y)} = \frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n-1}
+\end{equation}
 
@@
 When the character difference is maximal ($1$) it means the characters convey the greatest difference in phylogenetic signal
-
-\subsection{Simulating discrete morphological matrices}
-To simulate the matrices we applied a protocol very similar to \cite{Guillerme2016146} using \texttt{R} \citep{R}.
-First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution with $\lambda$ $>$ $\mu$\citep[\texttt{diversitree} v0.9-8;][]{fitzjohndiversitree2012} and saving the trees after reaching either 25, 75 or 150 taxa (see supplementary material 3, Figs 1, 2 and 3).
-For each tree, we arbitrarily set the outgroup to be the first taxon (alphabetically) thus effectively rooting the trees on this taxon.
-These trees are hereafter called the ``true'' trees (see distinction below).
-We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
-\begin{itemize}
-    \item The ``morphological HKY-binary'' model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
-    \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
-    We drew the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
-\end{itemize}
-
-\noindent For each character, both models (``morphological HKY-binary'' or M$k$) were chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
-These low evolutionary rate values allowed reduction in the number of homoplasic character changes, thus reinforcing the phylogenetic information in the matrices.
-We re-simulated every invariant characters to obtain a matrix with no invariant characters in order to better approximate real morphological data matrices.
-To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{sanderson1989patterns,OReilly20160081}.
-
-For each tree with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
-The matrices were generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-We repeated this step 20 times resulting in 180 ``normal'' morphological matrices.
 

### 2022-03-24 Writing/Palaeontology/draft.tex
Updated till comparing trees bit

@@
 \subsection{Measuring differences between characters}
+
 \label{CDdescription}
-To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information.
-In order to measure this, we used the following scaled Hamming distance between two characters:
+
+\textcolor{blue}{In order to introduce more or remove correlation in the unperturbed matrices, we use a character difference metric that is the scaled Hamming distance between two characters.
+Using this metric, two characters can have a character difference ranging from 0 (no difference, the characters are perfectly correlated and will provide the same phylogenetic information) to 1 (the characters are completely different and uncorrelated providing a different phylogenetic information):}
 
@@
 
-We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.
-We replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
+We standardised each character by arbitrarily modifying their character state tokens (or symbols) by \textcolor{blue}{order of appearance in the matrix - for example for the first character (1st column) in the matrix, if the state coded for the first OTU (1st row) was \texttt{3}, it was translated into \texttt{1} for the whole character (across the 1st column); if the state coded for the second OTU (2nd row) was \texttt{0}, it was translated into \texttt{2} for the whole character, etc.}
+We replaced all the occurrences of the first token to be \texttt{1}, the second to be \texttt{2}, etc.
 This procedure allows comparison of characters regardless of the significance given to their tokens (following the \textit{xyz} notation in \citealt{felsenstein2004inferring}; as used in \citealt{Davalos01072014}).
-This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
+This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \textcolor{blue}{(see supplementary materials 1)}.
 When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated). 
@@
 We calculated the pairwise character differences for each generated matrix using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
-We then modified the matrices to either maximise or minimise the pairwise character differences for each matrix using three different algorithms.
-For maximising the pairwise differences between characters, we selected the characters that were the most similar to all the others (mean character difference $<0.25$) and replaced them randomly by any of the remaining characters.
-This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
-Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (mean character difference $>0.75$) and randomly replaced them with the remaining ones.
+We then modified the matrices to \textcolor{blue}{either decrease or increase character difference, respectively maximising or minimising correlation}.
+For \textcolor{blue}{decreasing the pairwise differences between characters, we selected the characters that were the most dissimilar to all the others (mean character difference $>0.75$) and replaced them randomly by any of the remaining characters.
+This operation made the characters in the matrix more similar to each other, thus maximising character correlation.}
+Conversely, for \textcolor{blue}{increasing} the pairwise character differences, we selected the most similar characters (mean character difference $<0.25$) and randomly replaced them with the remaining ones, \textcolor{blue}{thus minimising character correlation.}
 Finally, because this operation effectively changes the weight of characters (i.e.
-giving the characters $<0.25$ or $>0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification.
+giving the characters $>0.75$ or $<0.25$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification.
 Each replaced characters were randomly selected by sampling them randomly from the list of the remaining characters.
-This step resulted in a total of 720 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
+
+This step resulted in a total of 720 matrices (\textcolor{blue}{the unperturbed, randomised, maximised correlation and minimised correlation matrices} - see Fig.
+\ref{Fig:modif_matrix} for an illustration).
 The algorithms for the three modifications are available on GitHub (\url{https://github.com/TGuillerme/CharactersCorrelation})
@@
 Blue colours correspond to low character differences and orange colours correspond to high character differences.
-\textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised (\textbf{C}), minimised (\textbf{D}) or randomise character difference (\textbf{E}).
-The differences between the characters is low in \textbf{C} (minimised compared to \textbf{A}) implying a high correlation between the characters.
-Conversely, the character differences is high in \textbf{D} (maximised compared to \textbf{A}) implying a low correlation between the characters.}
+\textcolor{blue}{\textbf{A} - a random Birth-Death tree is simulated and used for generating the unperturbed matrix (\textbf{B}), characters in this matrix are then removed or duplicated to maximise (\textbf{C}) or minimise (\textbf{D}) character correlation; or to randomise characters (\textbf{E}).
+The differences between the characters is low in \textbf{C} (decreased compared to \textbf{A}) implying a maximised correlation between the characters.
+Conversely, the character differences is high in \textbf{D} (increased compared to \textbf{A}) implying a minimised correlation between the characters.}}
 \label{Fig:modif_matrix}
@@
 Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$) with an variable rate prior an exponential (0.5) shape.
+\textcolor{blue}{We’ve used the Mk model with ascertainment bias (Mkv) to correct for the fact that no invariant characters were present in the matrix (as is often the case with morphological data) allowing the model to estimate the correct evolutionary rates (see above).
+We used four discrete character rates categories making the assumption that each character has an independent $\frac{1}{4}$ chance to be in the $1^{st}$, $2^{nd}$, $3^{rd}$ or $4^{th}$ quartile of the rate distribution.
+Finally we allowed that the average rate of this distribution ($\alpha$) is drawn from an exponential distribution (with a mean of 0.5) in every iteration of the MCMC allowing an efficient exploration of the parameter space.}
 The MCMC was ran over two runs of 6 chains each for a maximum of $1\times10^9$ generations with a sampling every 200 generations with an automatic stop if the average standard deviation of split frequency (ASDSF) fell below 0.01.
+\textcolor{blue}{We used this method to diagnose when the 12 chains reached a similar region in the parameter space.
+In other words, we considered that when the differences in solutions between the chains reaches a stabilised solution independently (ASDSF $<0.01$), then the MCMC converged on the ``correct'' inference.}
 Due to cluster hardware requirements and to save some time, when chains didn't converge and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the unconverged chains.

### 2022-03-24 Functions/compatibility.R

@@ -1,86 +1,84 @@
-#' @title Character compatibility functions
-#'
-#' @description 
-#'
-#' @param 
-#' 
-#' @examples
-#'
-#' @seealso
-#' 
-#' @author Thomas Guillerme
-#' @export
+# #' @title Character compatibility functions
+# #'
+# #' @description 
+# #'
+# #' @param 
+# #' 
+# #' @examples
+# #'
+# #' @seealso
+# #' 
+# #' @author Thomas Guillerme
+# #' @export
 
-# This code is modified from https://datadryad.org/stash/dataset/doi:10.5061/dryad.sh8b4 (functions pairCompat and compCount)
-
-
+# # This code is modified from https://datadryad.org/stash/dataset/doi:10.5061/dryad.sh8b4 (functions pairCompat and compCount)
 
 
-set.seed(4)
-## A random tree with 15 tips
-tree <- rcoal(15)
-## Setting up the parameters
-my_rates = c(rgamma, rate = 10, shape = 5)
-## Mk matrix (15*50) (for Mkv models)
-matrix <- sim.morpho(tree, characters = 50, model = "ER", rates = my_rates,invariant = FALSE) 
+# # set.seed(4)
+# # ## A random tree with 15 tips
+# # tree <- rcoal(15)
+# # ## Setting up the parameters
+# # my_rates = c(rgamma, rate = 10, shape = 5)
+# # ## Mk matrix (15*50) (for Mkv models)
+# # matrix <- sim.morpho(tree, characters = 50, model = "ER", rates = my_rates,invariant = FALSE) 
 
 
-pair.compat <- function(i, j)    {
+# pair.compat <- function(i, j)    {
 
-    ## Get the tokens
-    tokens <- unique(c(i, j))
+#     ## Get the tokens
+#     tokens <- unique(c(i, j))
 
-    ## The number of possible pairwise combinations
-    n_tokens <- length(tokens)
-    expect_combn <- ncol(combn(1:n_tokens, 2))*2 + n_tokens
+#     ## The number of possible pairwise combinations
+#     n_tokens <- length(tokens)
+#     expect_combn <- ncol(combn(1:n_tokens, 2))*2 + n_tokens
 
-    ## Intersection for any token
-    get.tokens <- function(token, character) return(which(character == token))
-    tokens_i <- unlist(sapply(tokens, get.tokens, character = i, simplify = FALSE))
[…]

### 2022-03-24 Analysis/01-EffectCorrelationStats.Rmd
Updated functions and results for new nomenclature

@@
 ## Pooling the data per scenarios
-scenario_RF_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
+scenario_RF_norm <- multi.pool(whole_data, param = c("mini", "maxi", "rand"),
                                metric = "RF", best = "norm")
-scenario_Tr_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
+scenario_Tr_norm <- multi.pool(whole_data, param = c("mini", "maxi", "rand"),
                                metric = "Tr", best = "norm")

### 2022-03-24 Analysis/02-EffectCorrelationFullResults.Rmd
Updated functions and results for new nomenclature

@@
 ## Pooling the data per scenarios
-scenario_RF_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
+scenario_RF_norm <- multi.pool(whole_data, param = c("mini", "maxi", "rand"),
                                metric = "RF", best = "norm")
-scenario_Tr_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
+scenario_Tr_norm <- multi.pool(whole_data, param = c("mini", "maxi", "rand"),
                                metric = "Tr", best = "norm")
-scenario_RF_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
+scenario_RF_null <- multi.pool(whole_data, param = c("mini", "maxi", "norm"),
                                metric = "RF", best = "rand")
-scenario_Tr_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
+scenario_Tr_null <- multi.pool(whole_data, param = c("mini", "maxi", "norm"),
                                metric = "Tr", best = "rand")

### 2022-03-24 Writing/Palaeontology/draft.tex
Updated functions and results for new nomenclature

@@
 \end{figure}
-
+''
 \subsection{Simulating discrete morphological matrices}
@@
 \subsection{Comparing topologies}
-We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
-We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 10 and 11).
-Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
-First, in biology, this tree is always unknown.
-Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
+We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance (\textcolor{blue}{the differences in clade compositions}, \citealt{RF1981}) and the triplets distance (\textcolor{blue}{the differences in tips positions}; \citealt{dobson1975triplets}) between the trees inferred from the maximised correlation, minimised correlation and randomised matrices and the tree inferred from the unperturbed matrix (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 10 and 11).
+Note that we are not comparing the trees to the starting tree used to simulate the matrices.
+This is because, first, in biology, this tree is always unknown; and second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the unperturbed,  maximised correlation and minimised correlation trees.
 
@@
    \includegraphics[width=1\textwidth]{RF_results_best.eps}
-\caption{\small{Effect of character difference on recovering the ``normal'' topology in terms of clade conservation ($NTS_{RF}$).
-The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 f from top to bottom respectively.
+\caption{\small{Effect of character difference on recovering the topology from the unperturbed matrix in terms of clade conservation ($NTS_{RF}$).
+The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
 The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}

### 2022-03-24 Writing/Tables/Tab_pooledscharacters_test.tex
Updated results
bold main results
Updated functions and results for new nomenclature

@@
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Tue Mar 27 14:03:13 2018
+% latex table generated in R 4.1.3 by xtable 1.8-4 package
+% Thu Mar 24 14:24:05 2022
 \begin{table}[ht]
 \centering
-\begin{tabular}{llr|rl}
+\begin{tabular}{llr|rr}
   \hline
@@
   \hline
-RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
-   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
-   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
-  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
-   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
-   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
+RF & c100 vs. c350 & 0.965 & 48263.500 & 0.000 \\ 
+   & c100 vs. c1000 & 0.914 & 37714.500 & 0.000 \\ 
+   & c350 vs. c1000 & 0.976 & 53138.000 & 0.000 \\ 
+  Tr & c100 vs. c350 & 0.929 & 40253.000 & 0.000 \\ 
+   & c100 vs. c1000 & 0.826 & 27923.000 & 0.000 \\ 
+   & c350 vs. c1000 & 0.957 & 48164.500 & 0.000 \\ 
    \hline
 \end{tabular}
-\caption{Difference between the pooled number of characters.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)}
+\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of characters: 100 (c100), 350 (c350) and 1000 (c1000).
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).}
 \label{Tab_pooledscharacters_test}

### 2022-03-24 Analysis/characterCompat 2.R

@@ -1,0 +1,118 @@
+
+### dropNames and dropBlanks are utility functions for read.wilkdat
+dropNames <- function(x)		{
+	out <- x[-1]
+	return(out)
+}
+dropBlanks <- function(x)	{
+	Keeps <- which(x != "")
+	return(x[Keeps])
+}
+
+### read.wilkdat reads in a file in the format used by MATRIX.EXE and converts it into an actual matrix in R (i.e., it removes all the fluff and reformats the character strings)
+read.wilkdat <- function(xstring)	{
+	x <- read.table(xstring, sep="^", stringsAsFactors=FALSE)
+	Stop <- which(x=="codes") - 1
+	dataLines <- 2:Stop
+	x <- x[dataLines,]
+	Pieces <- sapply(x, function(i) strsplit(i, " ")[[1]])
+	Names <- sapply(Pieces, function(i) i[1])
+	NoNames <- lapply(Pieces, dropNames)
+	NoBlanks <- lapply(NoNames, dropBlanks)
+	Collapsed <- lapply(NoBlanks, paste, collapse="")
+	CharSplit <- lapply(Collapsed, function(j) unlist(strsplit(j, split="")))
+	Matrix <- do.call(rbind, CharSplit)
+	rownames(Matrix) <- Names
+	return(Matrix)
+}
+
+### Core function for character compatibility. Given a data matrix and a pair of characters (i & j), this computes whether they conflict (1) or are compatible (0)
+pairCompat <- function(i, j, count, type="conflict")	{
+	Keeps <- intersect(c(which(i == "0"), which(i=="1")), c(which(j == "0"), which(j=="1")))
+	Combns <- paste(i[Keeps], j[Keeps])
+	Pairs <- unique(Combns)
+	if (type=="conflict")	{
+		Incompatibility <- ifelse(length(Pairs) < count, 0, 1)
+		return(Incompatibility)
+	}
+	if (type=="nesting")		{
+		count <- 3
+		iDerived <- which(i[Keeps]=="1")
+		jDerived <- which(j[Keeps]=="1")
+		derLen <- c(length(iDerived), length(jDerived))
+		Nested <- length(intersect(iDerived, jDerived))
+		isNested <- ifelse(min(derLen)==Nested, 1, 0)
+		return(isNested)
+	}
+}
+
[…]

### 2022-03-24 Functions/apply.test.R

@@ -7,6 +7,7 @@
 #' @param best which best tree (rand or norm)
 #' @param test which test (if NULL, simply summarises the distribution)
 #' @param convert.row.names whether to convert the row names into columns data (TRUE) or not (FALSE)
+#' @param translate whether to translate the param names to the latest publication (maxi -> minimised, mini -> maximised, norm -> unperturbed, rand -> randomised)
 #' 
 #' @examples
 #'
@@ -15,7 +16,7 @@
 #' @author Thomas Guillerme
 #' @export
 
-apply.test <- function(data, metric, best, test, convert.row.names = TRUE) {
+apply.test <- function(data, metric, best, test, convert.row.names = TRUE, translate = TRUE) {
 
     ## Combined
     combined <- TRUE
@@ -45,9 +46,9 @@ apply.test <- function(data, metric, best, test, convert.row.names = TRUE) {
     list_names <- names(NTS_list)
     parameter_names <- c("t25.", "t75.", "t150.", "c100.", "c350.", "c1000.", "bayesian", "parsimony")
     if(best == "norm") {
-        parameter_names <- c(parameter_names, c("maxi", "mini", "rand"))
+        parameter_names <- c(parameter_names, c("minimised", "maximised", "randomised"))
     } else {
-        parameter_names <- c(parameter_names, c("norm", "maxi", "mini"))
+        parameter_names <- c(parameter_names, c("unperturbed", "minimised", "maximised"))
     }
     parameters <- lapply(as.list(parameter_names), grep, list_names, fixed = TRUE)
     names(parameters) <- parameter_names
@@ -87,7 +88,7 @@ apply.test <- function(data, metric, best, test, convert.row.names = TRUE) {
             results_out[-c(1,10,19,28,37,46), 2] <- "" #taxa
             results_out[-seq(from = 1, to = 52, by = 3), 3] <- "" #characters
             ## Adding the column names
-            colnames(results_out)[1:4] <- c("method", "taxa", "characters", "scenario")
+            colnames(results_out)[1:4] <- c("method", "taxa", "characters", "correlation")
 
             return(results_out)
         } else {
@@ -108,7 +109,7 @@ apply.test <- function(data, metric, best, test, convert.row.names = TRUE) {
         colnames(test_results) <- apply(combo_matrix, 2, function(X) paste(parameter_names[X], collapse = ""))
     
         ## Get the row names (min:max, etc..)
-        rownames(test_results) <- c("max-min", "max-rand", "min-rand")
[…]

### 2022-03-24 Functions/plot.results.R

@@ -34,7 +34,7 @@
 
 plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, NTS = TRUE, ylab, xlab, axislab, leg.pos = "bottomleft", leg.cex = 0.8, ...) {
     
-    ## Graphic parameters
+    ## Graphic parameters
     ## Colours
     cols <- c("red", "orange", "green3", "lightgreen", "blue", "lightblue")
     
@@ -75,20 +75,20 @@ plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, N
     if(what == "null") {
         boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), las = 1, ...,
             ## c100
-            unlist(taxa_list$c100$bayesian$rand$maxi[,metric]), unlist(taxa_list$c100$parsimony$rand$maxi[,metric]),
             unlist(taxa_list$c100$bayesian$rand$mini[,metric]), unlist(taxa_list$c100$parsimony$rand$mini[,metric]),
+            unlist(taxa_list$c100$bayesian$rand$maxi[,metric]), unlist(taxa_list$c100$parsimony$rand$maxi[,metric]),
             unlist(taxa_list$c100$bayesian$rand$norm[,metric]), unlist(taxa_list$c100$parsimony$rand$norm[,metric]),
             ## c350
-            unlist(taxa_list$c350$bayesian$rand$maxi[,metric]), unlist(taxa_list$c350$parsimony$rand$maxi[,metric]),
             unlist(taxa_list$c350$bayesian$rand$mini[,metric]), unlist(taxa_list$c350$parsimony$rand$mini[,metric]),
+            unlist(taxa_list$c350$bayesian$rand$maxi[,metric]), unlist(taxa_list$c350$parsimony$rand$maxi[,metric]),
             unlist(taxa_list$c350$bayesian$rand$norm[,metric]), unlist(taxa_list$c350$parsimony$rand$norm[,metric]),
             ## c1000
-            unlist(taxa_list$c1000$bayesian$rand$maxi[,metric]), unlist(taxa_list$c1000$parsimony$rand$maxi[,metric]),
             unlist(taxa_list$c1000$bayesian$rand$mini[,metric]), unlist(taxa_list$c1000$parsimony$rand$mini[,metric]),
+            unlist(taxa_list$c1000$bayesian$rand$maxi[,metric]), unlist(taxa_list$c1000$parsimony$rand$maxi[,metric]),
[…]

### 2022-03-24 Functions/pool.distribution.R

@@ -38,38 +38,41 @@ pool.distribution <- function(data, param, metric, best) {
 
     ## Extracting the right distribution
 
-    ## Checking if the parameter is at the first level (number of taxa)
+    ## Checking if the parameter is at the first level (number of taxa)
     param_level1 <- match(param, names(NTS))
     if(!is.na(param_level1)) {
-        ## Extracting the distribution
-        return(as.vector(unlist(NTS[[param_level1]])))
+        ## Extracting the distribution
+        output <- as.vector(unlist(NTS[[param_level1]]))
     } else {
-        ## Checking if the parameter is at the second level (number of characters)
+        ## Checking if the parameter is at the second level (number of characters)
         param_level2 <- match(param, names(NTS[[1]]))
         if(!is.na(param_level2)) {
             ## Extracting the distribution
             distribution <- sapply(NTS, function(x, param_level2) x[[param_level2]], param_level2, simplify = FALSE)
-            return(as.vector(unlist(distribution)))
+            output <- as.vector(unlist(distribution))
         } else {
             ## Checking if the parameter is at the third level (method)
             param_level3 <- match(param, names(NTS[[1]][[1]]))
             if(!is.na(param_level3)) {
                 ## Extracting the distribution
                 distribution <- lapply(NTS, lapply, `[[`, param_level3)
-                return(as.vector(unlist(distribution)))
+                output <- as.vector(unlist(distribution))
             } else {
                 ## Checking if the parameter is at the fourth level (scenario)
                 param_level4 <- match(param, names(NTS[[1]][[1]][[1]]))
                 if(!is.na(param_level4)) {
                     ## Extracting the distribution
                     distribution <- lapply(NTS, lapply, lapply, `[[`, param_level4)
-                    return(as.vector(unlist(distribution)))
+                    output <- as.vector(unlist(distribution))
                 } else {
                     stop("Parameter not found.")
                 }
             }
         }
     }
+
+    ## Translate or not
+    return(output)
 }
 
 #' @title Pool multiple distributions for multiple parameters
@@ -80,7 +83,8 @@ pool.distribution <- function(data, param, metric, best) {
 #' @param param which parameters
 #' @param metric which metrics
 #' @param best which comparison
-#' 
[…]

### 2022-03-24 Functions/read.TreeCmp.R

@@ -19,7 +19,7 @@
 
 #path_alt <- "/Users/TGuillerme/Projects/CharactersCorrelation/Data/Trees_out/Consensus_trees/Parsimony"
 
-read.TreeCmp <- function(chain, path = "/Users/TGuillerme/Projects/CharactersCorrelation/Data/Trees_out/Consensus_trees", type = "Bayesian", length.in) {
+read.TreeCmp <- function(chain, path = "~/Projects/CharactersCorrelation/Data/Trees_out/Consensus_trees", type = "Bayesian", length.in) {
 
     path <- paste(path, type, sep = "/")
 

### 2022-03-24 Analysis/01-EffectCorrelationStats.Rmd
Updated results

@@
 ## Caption
-cap_pool_scen_test <- paste("Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for each minimised (mini), maximised (maxi) and randomised (rand) character differences.
+cap_pool_scen_test <- paste("Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for each maximised, minimised and randomised correlation.
 Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).")

### 2022-03-24 Functions/apply.test.R

@@ -7,7 +7,6 @@
 #' @param best which best tree (rand or norm)
 #' @param test which test (if NULL, simply summarises the distribution)
 #' @param convert.row.names whether to convert the row names into columns data (TRUE) or not (FALSE)
-#' @param translate whether to translate the param names to the latest publication (maxi -> minimised, mini -> maximised, norm -> unperturbed, rand -> randomised)
 #' 
 #' @examples
 #'
@@ -16,7 +15,7 @@
 #' @author Thomas Guillerme
 #' @export
 
-apply.test <- function(data, metric, best, test, convert.row.names = TRUE, translate = TRUE) {
+apply.test <- function(data, metric, best, test, convert.row.names = TRUE) {
 
     ## Combined
     combined <- TRUE
@@ -219,6 +218,7 @@ apply.pair.test <- function(table, test) {
 #' @param metric.label the metrics labels (can be missing)
 #' @param comp.label the comparisons labels (can be missing)
 #' @param correction which correction to apply
+#' @param translate whether to translate the param names to the latest publication (maxi -> minimised, mini -> maximised, norm -> unperturbed, rand -> randomised)
 #' 
 #' @examples
 #'
@@ -227,7 +227,7 @@ apply.pair.test <- function(table, test) {
 #' @author Thomas Guillerme
 #' @export
 #' 
-pair.test.table <- function(distributions, digit = 3, metric.label, comp.label, correction = "bonferroni") {
+pair.test.table <- function(distributions, digit = 3, metric.label, comp.label, correction = "bonferroni", translate = TRUE) {
 
     tests <- c(bhatt.coeff, wilcox.test)
 
@@ -251,21 +251,21 @@ pair.test.table <- function(distributions, digit = 3, metric.label, comp.label, 
     ## Rounding
     table <- round(table, digit = digit)
 
-    ## Adding the metrics and comp (if not missing)
+    ## Adding the metrics and comp (if not missing)
     if(!missing(metric.label) && !missing(comp.label)) {
         labels <- data.frame("comp" = comp.label, "metric" = metric.label, rownames(table))
         colnames(labels)[length(labels)] <- "test"
         table_out <- cbind(labels, as.data.frame(table))
     }
 
-    ## Adding the metrics (if not missing)
+    ## Adding the metrics (if not missing)
     if(!missing(metric.label) && missing(comp.label)) {
         labels <- data.frame("metric" = metric.label, rownames(table))
         colnames(labels)[length(labels)] <- "test"
[…]

### 2022-03-24 Writing/Palaeontology/draft.tex
Updated results

@@
 
+%---------------------------------------------
+%
+%       METHODS
+%
+%---------------------------------------------
 \section{Methods}
@@
 
+%---------------------------------------------
+%
+%       RESULTS
+%
+%---------------------------------------------
 \section{Results}
@@
    \includegraphics[width=1\textwidth]{RF_results_best.eps}
-\caption{\small{Effect of character difference on recovering the topology from the unperturbed matrix in terms of clade conservation ($NTS_{RF}$).
+\caption{\small{Effect of character correlation on recovering the topology from the unperturbed matrix in terms of clade conservation ($NTS_{RF}$).
 The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively.
-The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
+The x axis represents the different correlation scenarios and tree inference method with the maximised correlation in Bayesian (red) and under maximum parsimony (orange), the minimised correlation in Bayesian (dark green) and under maximum parsimony (light green) and the randomised correlation in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
 \label{Fig:RF_results_best}
@@
    \includegraphics[width=1\textwidth]{Tr_results_best.eps}
-\caption{Effect of character difference on recovering the ``normal'' topology in terms of Triplets ($NTS_{Tr}$).
+\caption{Effect of character correlation on recovering the topology from the unperturbed matrix in terms of Triplets ($NTS_{Tr}$).
 The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
@@
 \subsection{Effect of character differences on topology}
-The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information followed by randomising and maximising the character difference (see supplementary materials 3, Tables 3, 4 and 5).
+\textcolor{blue}{The overall amount of character difference in a matrix has an effect of the ability to recover the topology from the unperturbed matrix when maximised character correlation leading to the smallest loss in phylogenetic information followed by randomising and minimising the character correlation} (see supplementary materials 3, Tables 3, 4 and 5).
 There is a significant difference between all scenarios in terms of clade conservation ($NTS_{RF}$ - Table \ref{Tab_pooledscenarios_test}).
@@
 \subsubsection{Number of characters}
-This effect of the character difference affects clade conservation ($NTS_{RF}$) only over large differences in number of characters (100 \textit{vs.} 1000 - Table \ref{Tab_pooledscharacters_test}).
+The effect of the character difference affects clade conservation ($NTS_{RF}$) only over large differences in number of characters (100 \textit{vs.} 1000 - Table \ref{Tab_pooledscharacters_test}).
 For closer differences between characters numbers (100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests showed clear differences ($p = 0$) but with high distribution overlap ($BC > 0.96$).
@@
 \subsection{Combined effects of taxa, characters and correlation on topology}
-When looking at the combined effect of each parameter, the ``maximised'', ``minimised'' and ``randomised'' scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).
+When looking at the combined effect of each parameter, the maximised, minimised or randomised correlation scenarios $NTS_{RF}$ and $NTS_{Tr}$ never overlap ($BC < 0.95$ - supplementary material 3, tables 16 and 17).
 When comparing the distribution using the Wilcoxon ranked tests, there was no clear differences between tree similarity of differencing numbers of taxa, characters, tree correlation, or tree inference method for the taxon displacement metric ($NTS_{Tr}$ - supplementary material 3, tables 21 and 22).
-In terms of clade conservation ($NTS_{RF}$), however, ``maximised'' and ``minimised'' scenarios produced nearly always consistently different topologies using Bayesian inference (apart for the matrices with 25 taxa and 100 and 350 characters and the matrices with 150 taxa and 350 characters) but not using parsimony (only the matrices with 75 taxa and 100 or 350 characters produced different $NTS_{RF}$ distributions - supplementary material 3, tables 19 and 20).
+In terms of clade conservation ($NTS_{RF}$), however, maximised and minimised correlation scenarios produced nearly always consistently different topologies using Bayesian inference (apart for the matrices with 25 taxa and 100 and 350 characters and the matrices with 150 taxa and 350 characters) but not using parsimony (only the matrices with 75 taxa and 100 or 350 characters produced different $NTS_{RF}$ distributions - supplementary material 3, tables 19 and 20).
 
@@
 Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232) and the Australian QRIScloud Awoonga cluster services.
-We thank Alberto Pascual Garcia for input in the design of the simulations protocol, Guillermo Herraiz Yebes for helping with the CD metric proof and Orlin Todorov for assistance with the computations.
+We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Orlin Todorov for assistance with the computations.
 We also thank the three anonymous reviewers for there constructive and insightful comments.

### 2022-03-24 Writing/Palaeontology/draft.tex
Updated discussion

@@
 
+%---------------------------------------------
+%
+%       DISCUSSION
+%
+%---------------------------------------------
 \section{Discussion}
@@
 When characters are correlated, matrices convey a strong (but potentially misleading) phylogenetic signal since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of characters giving a conflicting signal (homoplasy).
-Throughout our simulation, the ``minimised'' character difference scenario (i.e.
-increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario led to slightly worse trees (Fig.
+Throughout our simulation, the maximised character correlation scenario (i.e.
+decreased character difference) tended to result in trees slightly closer to the tree inferred from the unperturbed matrix and the minimised character correlation led to slightly worse trees (Fig.
 \ref{Fig:RF_results_best}).
-These results suggest that, in matrices with high character differences, bootstrapping or other data randomisation methods can effectively improve the ability to recover the correct topology.
+These results suggest that, in matrices with low character correlations, bootstrapping or other data randomisation methods can effectively improve the ability to recover the correct topology.
 
-However, this effect of increasing or reducing character difference did not appear to affect taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
-Furthermore, these results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
-Furthermore, fewer characters were replaced, on average, in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6).
-This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.
+However, this effect of increasing or reducing character correlation did not appear to affect taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped to much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
+Furthermore, these results needs to be contrasted with the fact that the trees inferred from the maximised character correlation matrices do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
+Furthermore, fewer characters were replaced, on average, in the maximised correlation matrices compared to both the minimised correlation and the randomised matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6).
 
 \subsubsection{Number of characters and taxa}
-We found an effect of the number of characters on the ability to recover the ``best'' topology with different levels of character correlations only between the two extreme number of characters (100 and 1000; Tables \ref{Tab_pooledscharacters_test}).
+We found an effect of the number of characters on the ability to recover the topology from the unperturbed matrices with different levels of character correlations only between the two extreme number of characters (100 and 1000; Tables \ref{Tab_pooledscharacters_test}).
 This is expected since a higher number of characters can result in better topologies with a higher probability of ``noisy'' signal to cancel (\textit{c.f.} in smaller matrices).
@@
 When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (Table \ref{Tab_pooledsmethods_test}).
-The difference in the ability of each method to recover the ``correct'' topology has been recently discussed with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; but see \citealt{spencerefficacy2013,goloboff2017weighted}).
+The difference in the ability of each method to recover the topology inferred from the unperturbed matrices has been discussed with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; but see \citealt{spencerefficacy2013,goloboff2017weighted}).
 In this study, it is possible that our simulation protocol for generating the characters (using some M$k$-based characters) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[using matrices with a $CI>0.26$;][]{OReilly20160081} could also favour maximum parsimony analysis.
@@
 \subsection{Distinction between different character correlations}
-Here we mention three different types of character correlations but evolutionary biologists are mainly interested in intra-organismal and evolutionary correlations (e.g.
+Here we mention three different types of character correlations but \textcolor{blue}{only look at raw (logical) correlation and do not distinguish between the different character dependencies mentioned in the introduction}.
+However, it is likely that evolutionary biologists are mainly interested in intra-organismal and evolutionary correlations (e.g.
 in evo-devo \citealt{goswami2006morphological}; or in macroevolution \citealt{fitzjohn2014much}).
@@
 \subsection{Potential applications}
-Effectively, our simulation protocol bootstraps our data ``with bias''.
-In the ``randomised'' scenarios the data is simply randomly bootstrapped.
-However, in the ``minimised'' and ``maximised'' scenario, we remove the characters with the lowest/highest overall character difference.
-For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with each other and randomly resample from the characters left.
+Effectively, our simulation protocol bootstraps data ``with bias''.
+In the randomised matrices the data is simply randomly bootstrapped.
+However, in the matrices with maximised and minimised correlation, we remove the characters with the highest/lowest overall character difference.
+For example, in the matrices with minimised correlation, we randomly remove some characters that are strongly correlated with each other and randomly resample from the characters left.
 
-In rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
-Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling
-will be sufficient to obtain a robust topology.
-In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery.
-Applying this type of bootstraps that minimised character difference by biasing the random sampling could thus help in understanding which characters generate more homoplasy, which nodes in the tree are conflictual or which clades might be most dependent on highly correlated characters.
+In rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when minimising or randomising the character correlation.
+Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling will be sufficient to obtain a robust topology.
+In matrices with more taxa, however, maximising the character correlation resulted in better topological recovery.
+Applying this type of bootstraps that decreased character difference by biasing the random sampling could thus help in understanding which characters generate more homoplasy, which nodes in the tree are conflictual or which clades might be most dependent on highly correlated characters.
 Overall, we suggest measuring (and reporting) the levels of character difference in a matrix to help understand potential caveats in the data.
@@
 Not only because the models we're using to infer phylogenies are ever-improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible \citep[see constrasted discussions from][]{GoloboffEmpirical,OReillyEmpirical}.
-However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree.
+However, we do not compare the trees resulting from the \textcolor{blue}{maximised, minimised or randomised correlation matrices to the starting tree but rather to the tree inferred from the unperturbed matrix.}
 This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to the other outputs rather than to the simulation inputs.
@@
 Second, comparing phylogenetic inference methods is not trivial.
-As mentioned above, both maximum parsimony and Bayesian inference have similar outputs nut vastly differ in how optimality is measured.
-There are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
+As mentioned above, both maximum parsimony and Bayesian inference have similar outputs but vastly differ in how optimality is measured.
+There are also difficulties in summarising both methods with consensus trees \citep{oReilly2017efficacy}.
 However, we want to point out again that here we are not comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.
+
+\textcolor{blue}{Third, by not focusing on modeling the sources of character correlation specifically (see section ``Distinction between different character correlations'' above), we do not allow to distinguish the specific effect of each correlation for downstream analyses (see \citealt{sansom2017dental}f or such an application on dental morphological data).
+However, these sources are often not available to authors, either because of the data availability or the resources available to the workers (i.e.
+time and money) and thus we here propose an approach showing the effect of character correlation when the sources of correlation are effectively unknown.}
 
@@
 
-We found that character correlation has a strong effect on recovering the correct topology: matrices with high character differences always performed more poorly at recovering the correct topology than trees with low or randomised character differences.
+We found that character correlation has an effect on recovering the correct topology: matrices with low character correlation always performed more poorly at recovering the correct topology than matrices with high or randomised character correlation.
 This is especially true in large matrices where we advise to reduce character difference by using bootstrapping with ``bias'' whereas this may not be major problem in modest size matrices (25 taxa; 100 to 350 characters).
@@
 pleiotropy) correlation is likely be cancelled out if the correlated characters are randomly distributed with respect to traits.
-We recommend that all morphological phylogenetic analyses should report the level of character difference so that we can begin to detect and address the problem of coding dependence on inflating our confidence in our phylogenetic reconstructions.
+We recommend that all morphological phylogenetic analyses should report the level of character correlation or differences so that we can begin to detect and address the problem of coding dependence on inflating our confidence in our phylogenetic reconstructions.
 

### 2022-03-24 Writing/Palaeontology/draft.tex
Updated affiliations

@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, Abigail I. Pastore$^{1}$, and Martin D. Brazeau$^{2,3}$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^{1,2,*}$, Abigail I. Pastore$^{2}$, and Martin D. Brazeau$^{3,4}$}\\
 % Abigail I. Pastore$^{1}$
 \noindent {\small \it 
-$^1$School of Biological Sciences, University of Queensland, St. Lucia, Queensland, Australia.\\
-$^2$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\
-$^3$Department of Earth Sciences, Natural History Museum, Cromwell Road, London, SW75BD, United Kingdom.\\}
-
-% TODO: change the affiliations
+$^1$School of Biosciences, The University of Sheffield, Sheffield, United Kingdom.\\
+$^2$School of Biological Sciences, University of Queensland, St. Lucia, Queensland, Australia.\\
+$^3$Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.\\
+$^4$Department of Earth Sciences, Natural History Museum, Cromwell Road, London, SW75BD, United Kingdom.\\}
 

### 2022-03-29 Writing/Palaeontology/Reviews/response_to_reviewers.tex
Editors comments sorted (missing 1)

@@
 We are very grateful to the three referees for their helpful and constructive comments, which we believe have helped me to significantly improve the manuscript.
-We have taken all of their comments on board, and respond to their points in details below.
+Unfortunately, due to circumstances out of our control (pandemic, moving continent and new job), it took us way longer to reply to the reviewers requests as we could have done.
+Fortunately, however, we have taken all of their comments on board, and respond to their points in details below.
 
@@
 
+%@@@TODO-1: add a line in the intro explaining this? 
+%TG: something like that "Note that investigating the shared evolutionary history is often part of the goal of phylogenetic analyses."
+
 \item{\textcolor{blue}{Matrices that were modified to minimise character differences are referred to as 'minimised'.
@@
 
+We have changed the naming of the matrices and the trees throughout the whole manuscript to now only focus on the character correlation aspect (\textit{c.f.} the character difference).
+The \textit{Normal} matrix/tree is now the \textit{Unperturbed} matrix/tree the \textit{Maximised (difference)} matrix/tree is now the \textit{Minimised (correlation)} matrix/tree and the \textit{Minimised (difference)} matrix/tree is now the \textit{Maximised (correlation)} matrix.
+The \textit{Randomised} matrix/tree stayed unchanged.
+
 \item{\textcolor{blue}{'Normal' has multiple meanings and mathematically could mean 'orthogonal'.
@@
 A more intuitive term might be 'unmodified' or 'simnulated', or something else, there are probably better options.}}
+
+We have now changed mentions to the \textit{Normal} tree as "the topology inferred from the unperturbed matrix" to avoid any confusion throughout the manuscript.
 
@@
 Or tended to produce strong support for incorrect topologies.}}
+
+We have now replaced mentions of "phylogenetic signal" to "\textit{phylogenetic information}" throughout the whole manuscript to avoid any confusion with other uses of the term "phylogenetic signal".
+Here we mean information (code) used to infer phylogenies.
 
@@
 From what the authors state later, it’s not the randomness that is important, but that it occurs infrequently relative to total taxon sampling.}}
+
+We have now changed "random correlation" to "\textit{randomised} correlation" throughout the manuscript to better reflect this comment.
 

### 2022-03-29 Writing/Palaeontology/draft.tex
Editors comments sorted (missing 1)

@@
 These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
-However, encouragingly, they also suggest that, unless correlation is actively maximised or minimised, probabilistic methods can easily accommodate for a random correlation between characters.
+However, encouragingly, they also suggest that, unless correlation is actively maximised or minimised, probabilistic methods can easily accommodate for a \textcolor{blue}{randomised} correlation between characters.
 
@@
 This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \textcolor{blue}{(see supplementary materials 1)}.
-When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated). 
-When the character difference is maximal ($1$) it means the characters convey the greatest difference in phylogenetic signal
+When the character difference is null ($0$) it means that characters convey the same \textcolor{blue}{phylogenetic information} (i.e. characters are entirely correlated). 
+When the character difference is maximal ($1$) it means the characters convey the greatest difference in \textcolor{blue}{phylogenetic information}.
 
@@
 We find that there is a significant influence of morphological character correlation on inferring tree topology.
-When characters are correlated, matrices convey a strong (but potentially misleading) phylogenetic signal since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of characters giving a conflicting signal (homoplasy).
+When characters are correlated, matrices convey a strong (but potentially misleading) \textcolor{blue}{phylogenetic information} since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker \textcolor{blue}{phylogenetic information}with a high amount of characters giving a conflicting signal (homoplasy).
 Throughout our simulation, the maximised character correlation scenario (i.e.

### 2022-03-29 Writing/Palaeontology/Reviews/response_to_reviewers.tex
Added Reviewer 2 responses

@@
 I think that what the authors really mean is moderate amounts of correlation that basically mimic accelerated local rates of change on a small proportion of branches.
-From what the authors state later, it’s not the randomness that is important, but that it occurs infrequently relative to total taxon sampling.}}
+From what the authors state later, it's not the randomness that is important, but that it occurs infrequently relative to total taxon sampling.}}
 
@@
 \\
-We’ll set a=1 and let’s say we have a rate of change of π=0.05. Our transition matrix for time t=1 is now mexp(πQt) =
+We'll set a=1 and let's say we have a rate of change of $\pi$=0.05. Our transition matrix for time t=1 is now mexp($\pi$Qt) =
 \\
-Note that if we calculated the probability of going from 0 to 1 for either character, then we’ve have almost the same value: the difference is that there is some small chance that, over the same amount of time, we also went from 0 to 1 for the other character, and thus from 00 to 11.
+Note that if we calculated the probability of going from 0 to 1 for either character, then we've have almost the same value: the difference is that there is some small chance that, over the same amount of time, we also went from 0 to 1 for the other character, and thus from 00 to 11.
 The probability of stasis is the probability of stasis for one character squared.
@@
 
+% @@@TODO-6: Check if that works with comments: https://docs.google.com/document/d/1ED7b-uxtEzJ5fUUXugBYMIq88zx2ySY9fHcJDcUekAs/edit?disco=AAAAMGcMrqU
+%TG: still to do. But I think it's kind of in there...
+
 \item{\textcolor{blue}{Lines 58-59: 'For example, two characters 'tail colour' and 'tail length' could be coded two times as an absence for a taxon with no tail (Wilkinson, 1995b; Brazeau et al., 2018).'
@@
 If that tail was there, then it had States A, B, C, etc., for length, color, etc.
-Because we are multiplying this, it’s really P[tail present] X P([tail state A]+tail state B)) etc.}}
+Because we are multiplying this, it's really P[tail present] X P([tail state A]+tail state B)) etc.}}
+
+We have added this precision in the text:
+
+\textit{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail (Wilkinson, 1995b; Brazeau et al., 2018).
+Note that there are a lot of alternative methodological choices for this example (e.g.
+using inapplicable tokens; Brazeau 2011) and that these choices affect results more or less depending on the estimation method (parsimony methods contrast with probabilistic methods were ancestral conditions are respectively fixed or not).
+Line @@@.}
 
@@
 are assessing is a little bit different from what Pagel and later workers considered: what happens if there is correlated character evolution happening among unknown sets of characters?
-Billet \& Bardin’s study focused on mammal teeth: but even non-biologists would expect mammal teeth to show correlated character change.
+Billet \& Bardin's study focused on mammal teeth: but even non-biologists would expect mammal teeth to show correlated character change.
 This raises the issue with which readers of Palaeontology will be familiar: how do I have reason to suspect that there is correlated change in some group of characters, and what happens if I fail to guess?
 What this simulation study offers is an idea of what happens when there is some amount of correlated change, but we just do not know how much.}}
+
+We are grateful for this insight and have added it to the manuscript:
+
+\textit{Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree (but see useful methods from Pagel 1994 and Billet and Bardin 2019), and is best determined by independent character sets.
+The later being commonly hard to get because of the nature of the data: some common type of data (e.g.
+fossil data) offer very few options of independent character sets and no biological datasets can really be independent.
+Line @@@.}
 
@@
 
+%@@@ TODO:
+
 \item{\textcolor{blue}{Lines 131-132: 'We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.'
@@
 
-\item{\textcolor{blue}{Lines 150 – 152: 'The 'morphological HKY-binary' model (O’Reilly et al.,
+We apologise for the confusion here: by ``order of appearance'' we meant:
+\textit{order of appearance in the matrix - for example for the first character (1st column) in the matrix, if the state coded for the first OTU (1st row) was \texttt{3}, it was translated into \texttt{1} for the whole character (across the 1st column); if the state coded for the second OTU (2nd row) was \texttt{0}, it was translated into \texttt{2} for the whole character, etc...} and have now specified that in the manuscript on line @@@.
+
+\item{\textcolor{blue}{Lines 150 – 152: 'The 'morphological HKY-binary' model (O'Reilly et al.,
 2016) which is an HKY model (Hasegawa et al., 1985) with a random states frequency (sampled from a Dirichlet distribution Dir(1, 1, 1, 1))...'
@@
 The authors should state that this allows variation in gains/losses of present absent characters to vary without specifying which characters are prone to gain/loss.
-The authors also should note that because it’s (1,1,1,1), the expectation is that gains and losses will be equiprobable.
-This does raise the question: why didn’t they use a beta distribution (1,1) given that they were dealing with 2-state characters?
-That’s just the special case of the Dirichlet.
+The authors also should note that because it's (1,1,1,1), the expectation is that gains and losses will be equiprobable.
+This does raise the question: why didn't they use a beta distribution (1,1) given that they were dealing with 2-state characters?
+That's just the special case of the Dirichlet.
 Using a 4 state character like this seems a little weird because net stasis will include both net stasis and net transition, and thus is P[net stasis] + P[net transition]; net change now is twice net transversion and thus 2*P[net transition].
 So, this will not quite be the same distribution as beta(1,1); the expectation will still be 50:50, but (I think!) the variance on the state-acquisition bias will be a little greater under the 'lumped' 4-state Dirichlet than under the Beta.}}
+
+We used the Dir(1,1,1,1) distribution to be consistent with previously published simulations (e.g. O'Reilly et al. 2016).
+We have however specified the following:
+
+\textit{This allows the changes in character states to vary without specifying which change is more likely, hence expecting an equiprobable change in character state in all directions.} Line @@@.
 
@@
 How did the Mk model generate the 3-state characters?
-Shouldn’t the authors have used a Dirichlet (1,1,1) (or a Dirichlet (1,1,1,1,1,1) with lumping of 1\&2, 3\&4 and 5\&6) to actually generate the characters with individual rate bias?}}
+Shouldn't the authors have used a Dirichlet (1,1,1) (or a Dirichlet (1,1,1,1,1,1) with lumping of 1\&2, 3\&4 and 5\&6) to actually generate the characters with individual rate bias?}}
 
-\item{\textcolor{blue}{Lines 159-161: 'For each character, both models ('morphological HKYbinary' or Mk) were chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution (β = 100 and α = 5).'
+We generated multi ($>2$) state characters with the Mk model using the \texttt{ape::rTraitDisc} function (through \texttt{dispRity::sim.morpho}) that calculates the transition probabilities for each branch using any sized Q matrices.
+We specified this in the text:
+
+\textit{we used the Mk model (Lewis 2001).
+We drew the number of character states with a probability of 0.85 for binary characters (using a 2x2 transition matrix Q) and 0.15 for three state characters (using a 3x3 Q matrix)}.
+Lines @@@
+
+\item{\textcolor{blue}{Lines 159-161: 'For each character, both models ('morphological HKYbinary' or Mk) were chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = 100 and $\alpha$ = 5).'
 \\
@@
 I would state plainly that the prior section was about generating variation in state gain/loss biases.
-This part is about generating variation in character change; basically, you are allowing each character to change at its own rate and (when it does change) to have its own bias in gains/losses of states; however, in all cases, every character’s particular rate and gain/loss bias is drawn from the same distribution.
+This part is about generating variation in character change; basically, you are allowing each character to change at its own rate and (when it does change) to have its own bias in gains/losses of states; however, in all cases, every character's particular rate and gain/loss bias is drawn from the same distribution.
 \\
-The second is that this is different from the way systematists usually use gamma distributions for rate variation, in which they have a separate rate parameter π (I will use π rather than α as Lewis did, simply to avoid confusion between rate parameter and the Gamma shape parameter α).
-They then set β=α so that the mean of the Gamma, α/(β=α), is 1.0 and the variance, α/(β=α)2, increases as α decreases.
-The expected rate of change then is π and the variance is π/α.
+The second is that this is different from the way systematists usually use gamma distributions for rate variation, in which they have a separate rate parameter $\pi$ (I will use $\pi$ rather than $\alpha$ as Lewis did, simply to avoid confusion between rate parameter and the Gamma shape parameter $\alpha$).
+They then set $\beta$=$\alpha$ so that the mean of the Gamma, $\alpha$/($\beta$=$\alpha$), is 1.0 and the variance, $\alpha$/($\beta$=$\alpha$)2, increases as $\alpha$ decreases.
+The expected rate of change then is $\pi$ and the variance is $\pi$/$\alpha$.
 \\
-By using separate shape (α) and rate (β) parameters without any independent rate parameter, these authors are making a mean rate π = α/β = 0.05 with variance of the rate being 0.0005.
+By using separate shape ($\alpha$) and rate ($\beta$) parameters without any independent rate parameter, these authors are making a mean rate $\pi$ = $\alpha$/$\beta$ = 0.05 with variance of the rate being 0.0005.
 I think that the authors should just state this.
-(Also, I’m not sure that everyone knows reflexively what the mean and variance are given α and β, so it’s probably best to remind them.)}}
+(Also, I'm not sure that everyone knows reflexively what the mean and variance are given $\alpha$ and $\beta$, so it's probably best to remind them.)}}
+
+We thank the reviewer for this clear explanation and agree that it is always best to be extra clear and remind the reader what these values actually mean. We've added the following precisions:
+
+\textit{This simulated, for each character, a variation in the change of character state at different rates while ensuring that these rates were drawn from the same distribution, resulting in an overall mean character rate of 0.05 with a variance of 0.0005.} Line @@@
 
@@
 Also, is a rate of 0.05 really all that low?
-That’s a pretty typical rate per million years for fossil invertebrates.}}
+That's a pretty typical rate per million years for fossil invertebrates.}}
+
+We explicitly specified the overall mean and variance of the rates distributions (see above) and removed the term ``low'' to reflect this comment.
 
@@
 Ah, but that also means that the true rate is >.05!
-Programs like RevBayes take into account ascertainment bias (unscored invariant characters), and individual runs could be limited to examine only variant characters; wouldn’t doing the latter and then accommodating ascertainment bias in the analyses be a little more realistic?
+Programs like RevBayes take into account ascertainment bias (unscored invariant characters), and individual runs could be limited to examine only variant characters; wouldn't doing the latter and then accommodating ascertainment bias in the analyses be a little more realistic?
 (This is probably not a major issue, so I would not worry over much about it; however, it is worth mentioning ascertainment bias at this point and why that makes what you are doing more realistic than keeping invariant characters in the simulations.)}}
 
-\item{\textcolor{blue}{Lines 165-167: 'we only selected matrices with Consistency Indices (CI) superior to 0.26 (Sanderson and Donoghue, 1989; O’Reilly et al., 2016).'
+We now specify all this as follows:
+\textit{We re-simulated every invariant character to obtain a matrix with no invariant characters (making the effective mean rate to > 0.05).
+Although invariant characters are important for effectively measuring the true evolutionary rates, many empirical matrices do not contain invariant characters which is in turn taken into account by software that accommodate for this ascertainment bias.} Line @@@
+
+\item{\textcolor{blue}{Lines 165-167: 'we only selected matrices with Consistency Indices (CI) superior to 0.26 (Sanderson and Donoghue, 1989; O'Reilly et al., 2016).'
 \\
@@
 So, you might explain that this means about 3.85 (1/0.26) changes per derived character; if we have 50 binary characters, then it would be 192 changes (50/192 = 0.26).}}
+
+We've now added the following precision:
+
+\textit{To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to 0.26 (expecting at least 3.85 (1/0.26) changes per derived character; Sanderson \& Donoghue 1989; O'Reilly et al.
+2016).} Line @@@
 
@@
 Also, it should be explained that the Gamma distribution itself is varied over the runs, so that the tree will be considering new topologies, new average rates and new rate distributions.
-What this means is that the shift in the α=β parameter is drawn from an exponential distribution with a mean of 0.5.
+What this means is that the shift in the $\alpha$=$\beta$ parameter is drawn from an exponential distribution with a mean of 0.5.
 However, I think that it is also worth reminding the reader that every generation in the MCMC analysis will consider altered topologies, average rates and Gamma shape parameters; improvements are kept and (sometimes!) worse propositions are accepted, too, in order to better search hypothesis space.}}
+
+We thank the reviewer for these precisions and again agree that it is important to make sure the reader understands the intrications of these parameters:
+
+\textit{Bayesian inference was run using an Mk model with ascertainment bias and four discrete gamma rate categories (Mkv 4$\Gamma$) with an variable rate prior an exponential ($\alpha$=0.5) shape.
+We've used the Mk model with ascertainment bias (Mkv) to correct for the fact that no invariant characters were present in the matrix (as is often the case with morphological data) allowing the model to estimate the correct evolutionary rates (see above).
+We used four discrete character rates categories making the assumption that each character has an independent 1/4 chance to be in the 1st, 2nd, 3rd or 4th quartile of the rate distribution.
+Finally we allowed that the average rate of this distribution (alpha) is drawn from an exponential distribution (with a mean of 0.5) in every iteration of the MCMC allowing an efficient exploration of the parameter space.} Lines @@@
 
@@
 
+We've added the following to explain the convergence and the ASDSF more clearly:
+
+\textit{We used this method to diagnose when the 12 chains reached a similar region in the parameter space.
+In other words, we considered that when the differences in solutions between the chains reaches a stabilised solution independently (ASDSF < 0.01), then the MCMC converged on the “correct” inference} Lines @@@
+
 \item{\textcolor{blue}{Lines 215-218: 'We measured both the Robinson-Foulds distance (Robinson and Foulds, 1981) and the triplets distance (Dobson, 1975) between the trees inferred from the 'maximised', 'minimised' and 'randomised' matrices and the tree inferred from the 'normal' matrix.
@@
 I think that a brief explanation of the metrics is in order here. If nothing else, then move lines 230 – 231 up to this paragraph so that readers will know right away why you are using both metrics.}}
+
+We now briefly define the two metrics when mentioning them:
+
+\textit{we measured both the Robinson-Foulds distance (the differences in clade compositions; Robinson \& Foulds 1981) and the triplets distance (the differences in tips positions; Dobson 1975) between the trees inferred}.
+Lines @@@
+
 \end{itemize}

### 2022-03-29 Writing/Palaeontology/draft.tex
Added Reviewer 2 responses

@@
 Meanwhile, the development of algorithms and software has not yet caught up with the need to deal with these interdependencies \citep{de2015parsimony,BrazeauNA}.
-Intra-organismal dependence requires more detailed, often extremely time-consuming studies (possibly beyond the limits of available technology).
-Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree (but see useful methods from \citealt{Pagel1994} and \citealt{billet2019serial}), and is best determined by independent character sets.
+\textcolor{blue}{Intra-organismal dependence requires more detailed, often extremely time-consuming studies (possibly beyond the limits of available technology).
+Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree (but see useful methods from \citealt{Pagel1994} and \citealt{billet2019serial}), and is best determined by independent character sets.}
 \textcolor{blue}{The latter is commonly hard to get because of the nature of the data: some common types of data (e.g.

### 2022-03-29 Writing/Palaeontology/Reviews/response_to_reviewers.tex
added reviewer 3

@@
 
-\textcolor{blue}{Sadly however, I am skeptical that this paper really addresses the central issue with character correlation and inferring phylogenies.
-Particularly, their main conclusion is that character correlation is a 'good' thing, i.e.
-having increased character correlation actually results in more accurate topologies.}
+\textcolor{blue}{Sadly however, I am skeptical that this paper really addresses the central issue with character correlation and inferring phylogenies.}
+
+We are also sad that we did not address the central issue(s) with character correlation although that was never our intention.
+We have now specified more clearly that what this paper ``offers is an idea of what happens when there is some amount of correlated change, but we just do not know how much.'' (as correctly worded by referee 2).
+We believe that the issue(s!) with character correlation and inferring phylogenies have to be addressed in an incremental way and hope that this manuscript will be a step forward in that direction.
+We have now clarified that throughout the manuscript:
+
+% @@@TODO-7.
+
+\textcolor{blue}{Particularly, their main conclusion is that character correlation is a 'good' thing, i.e. having increased character correlation actually results in more accurate topologies.}
+
+We do not claim this in the manuscript but rather that character correlation can, for specific datasets and methods, more or less bias topological inference. We have now clarified this
+
+% @@@TODO-8.
 
@@
 Pleijel 1995, Cladistics).}
+
+We do agree that these are valid limitations from this study and have acknowledged them more clearly:
+Third, by not focusing on modeling the sources of character correlation specifically (see section “Distinction between different character correlations” above), we do not allow to distinguish the specific effect of each correlation for downstream analyses (see Sansom et al.
+2017 for such an application on dental morphological data).
+However, these sources are often not available to authors, either because of the data availability or the resources available to the workers (i.e.
+time and money) and thus we here propose an approach showing the effect of character correlation when the sources of correlation are effectively unknown.
+Lines @@@
+% @@@TODO-9
 
@@
 
+We thank the reviewer for drawing our attention to this paper (Sansom et al. 2017) and the DNA example and have now added that to our manuscript:
+
+\textit{It is also worth noting that the evolutionary and intra-organismal dependences induced correlations are by no means restricted to the use of discrete morphological data.
+In fact, by essence, these correlations appear to any type of data whether continuous (e.g.
+linear measurements or morphometric geometric measurements) or non-morphological (e.g.
+behaviour, or - much more commonly used - molecular data).} Lines @@@
+% @@@TODO-10: add ref to Sansom somewhere in the discussion.
+
 \textcolor{blue}{In the discussion (P6.
@@
 
+We do agree that these are valid limitations from this study and have acknowledged them more clearly:
+
+\textit{Third, by not focusing on modeling the sources of character correlation specifically (see section “Distinction between different character correlations” above), we do not allow to distinguish the specific effect of each correlation for downstream analyses (see Sansom et al.
+2017 for such an application on dental morphological data).
+However, these sources are often not available to authors, either because of the data availability or the resources available to the workers (i.e.
+time and money) and thus we here propose an approach showing the effect of character correlation when the sources of correlation are effectively unknown.} Lines @@@
+
 \textcolor{blue}{In the conclusions the authors state: 'However, in empirical datasets, if character difference is not driven by selection (e.g.
@@
 
+% @@@TODO-11: specify this in the conclusion
+
 \subsection{Simulation protocol and methodology}
@@
 \textcolor{blue}{I am not sure that I understand the logic of only comparing the results of the phylogenetic analyses to what the authors call the 'normal' tree and not the tree that generated the data.
-Surely the whole point of this exercise is to understand if character correlation actually results in misleading phylogenetic inferences.
-It strikes me that the way in which they treat their morphological matrices just results in the 'minimised' matrix having the phylogenetic signal that supports the 'normal' tree reinforced as the correlated characters will essentially just contain the same phylogenetic signal that supported the 'normal' tree, though not necessarily the true tree.
+Surely the whole point of this exercise is to understand if character correlation actually results in misleading phylogenetic inferences.}
+
+The purpose of the “normal tree” (now described as the ``topology inferred from the unperturbed matrix'') is to isolate the variables being studied.
+Indeed, if the whole point of the exercise is to understand if character correlation actually results in misleading signal, one must isolate that variable from all others.
+This difference is crucial because our methods did not directly simulate covariation.
+We simulated datasets and then perturbed them.
+Thus the experimental ‘true tree’ (now described as "the starting tree") is the one recovered by the analyses in the absence of any of our experimental perturbations.
+Comparison with the simulation (i.e.
+“starting tree”) would introduce additional stochastic variables that arose during the unperturbed simulation of the datasets.
+Comparisons with that tree would be unable to isolate the effects of our experimental perturbations.
+
+\textcolor{blue}{It strikes me that the way in which they treat their morphological matrices just results in the 'minimised' matrix having the phylogenetic signal that supports the 'normal' tree reinforced as the correlated characters will essentially just contain the same phylogenetic signal that supported the 'normal' tree, though not necessarily the true tree.
 For many of the simulated matrices, the normal tree could share very few clades with the tree that was used to generate the data itself, which is clearly demonstrated in O'Reilly et al 2016 (where the same binary HKY simulation method was also used).
 In this paper, the accuracy of the trees from downstream phylogenetic analyses are often very different to the generating tree.
-In any revision, I would want to see the impact of comparing the simulated matrices to the generating tree.
-}
+In any revision, I would want to see the impact of comparing the simulated matrices to the generating tree.}
+
+We’ve added the comparisons between the trees from the modified matrices and the "starting tree" in the supplementary materials. %TG: is that done?
+However, as stated above, what these results actually show is the effect of A: changing the character correlation + B: the effect of phylogenetic inference + C: the effect of simulating discrete morphological characters.
+This last component C is not the focus of our study (as specified on lines @@@) and thus is removed from the main manuscript.
+In fact, there is a really rich amount of papers trying to understand the effect of B and C with no consensus so far (i.e.
+simulating realistic evolutionary data is complicated AND inferring a realistic evolutionary history from real data is complicated).
+Therefore, we chose to only focus on A and B and correct for C by effectively comparing to the topology inferred from the unperturbed matrix rather than the starting tree (i.e.
+making the results relative to the ``best'' inference possible with the simulated data).
 
@@
 
+% @@@TODO-13 we have changed the font size for the figures/tables captions and the main text.
+
 \textcolor{blue}{The text has many typos, mostly pluralization (e.g. character vs characters). Randomised is misspelled in Fig 1. Please check the text and figures thoroughly.}
+
+We have fixed the typo in figure 1 and checked the rest of the manuscript more thoroughly.
+%TG: TODO
 

### 2022-05-16 Analysis/03-EffectCorrelationFigure.Rmd
sync

@@
 ```
+
+## Branch length distribution
+
+How are the branch lengths distributed in the initial trees?
+
+```{r, fig.width=8, fig.height=8}
+## Branch lengths for the trees with 25 taxa
+plot.brlen.distribution("../Data/Simulations/Trees/", ntaxa = 25, cols = "grey", main = "Branch length distributions for \"true\" trees with 25 taxa", zeros = c(1e-16, 1e-10, 1e-05, 1e-01))
+```
+
+```{r, fig.width=8, fig.height=8}
+## Branch lengths for the trees with 25 taxa
+plot.brlen.distribution("../Data/Simulations/Trees/", ntaxa = 75, cols = "grey", main = "Branch length distributions for \"true\" trees with 75 taxa", zeros = c(1e-16, 1e-10, 1e-05, 1e-01))
+```
+
+```{r, fig.width=8, fig.height=8}
+## Branch lengths for the trees with 25 taxa
+plot.brlen.distribution("../Data/Simulations/Trees/", ntaxa = 150, cols = "grey", main = "Branch length distributions for \"true\" trees with 150 taxa", zeros = c(1e-16, 1e-10, 1e-05, 1e-01))
+```
+
+<!-- Figures for SOM: -->
+
+```{r, fig.width=8, fig.height=8, eval = TRUE, echo = FALSE}
+pdf("../Writing/Figures/Zero_brlen_25t.pdf", width = 8, height = 8)
+## Branch lengths for the trees with 25 taxa
+plot.brlen.distribution("../Data/Simulations/Trees/", ntaxa = 25, cols = "grey", main = "Branch length distributions for \"true\" trees with 25 taxa", zeros = c(1e-16, 1e-10, 1e-05, 1e-01))
+```
+
+```{r, fig.width=8, fig.height=8, eval = TRUE, echo = FALSE}
+pdf("../Writing/Figures/Zero_brlen_75t.pdf", width = 8, height = 8)
+## Branch lengths for the trees with 75 taxa
+plot.brlen.distribution("../Data/Simulations/Trees/", ntaxa = 75, cols = "grey", main = "Branch length distributions for \"true\" trees with 75 taxa", zeros = c(1e-16, 1e-10, 1e-05, 1e-01))
+```
+
+```{r, fig.width=8, fig.height=8, eval = TRUE, echo = FALSE}
+pdf("../Writing/Figures/Zero_brlen_150t.pdf", width = 8, height = 8)
+## Branch lengths for the trees with 150 taxa
+plot.brlen.distribution("../Data/Simulations/Trees/", ntaxa = 150, cols = "grey", main = "Branch length distributions for \"true\" trees with 150 taxa", zeros = c(1e-16, 1e-10, 1e-05, 1e-01))
+```

### 2022-05-16 Analysis/04-EffectCorrelation_lm.Rmd
sync

@@
 ## Loading the matrices CD
-matrices_cd <- get.matrix.CD(path = "../Data/Simulations/Matrices", "25t_100c", verbose = TRUE, length.out = 35, cent.tend = mean)
+matrices_cd <- get.matrix.CD(path = "../Data/Simulations/Matrices", chain = "25t_100c", verbose = TRUE, length.out = 35, cent.tend = mean)
 

### 2022-05-16 Analysis/Character_difference/01-Comparing_characters.Rmd
sync

--- 2017-07-17 Analysis/Character_difference/Comparing_characters.Rmd
@@
 
-```{r}
+```{r, eval = TRUE}
 # Number of replicates:
@@
 
-```{r, fig.width = 12, fig.height = 12}
+```{r, fig.width = 12, fig.height = 12, eval = TRUE}
 op <- par(mfrow = c(2,2), bty = "n")
@@
 
-```{r, fig.width = 10, fig.height = 10}
+```{r, fig.width = 10, fig.height = 10, eval = TRUE}
 plot(first_character_differences[,1], type = "l", col = "lightblue", ylim = c(0,1), xlab = "",

### 2022-05-16 Analysis/Character_difference/02-Modifying_matrices.Rmd
sync

@@
 ## The cran version of diversitree
-if(!require(diversitree)) install.packages("diversitree")
+if(!require(diversitree)) install_github("richfitz/diversitree")
 ```
@@
 
-```{r, fig.width=12, fig.height=12, eval = FALSE, echo = FALSE}
-# pdf("../../Writing/Figures/Modif_matrix.pdf", width = 7.2, height = 4.5)
+```{r, fig.width=12, fig.height=12, eval = TRUE, echo = FALSE}
+pdf("../../Writing/Figures/Modif_matrix2.pdf", width = 7.2, height = 4.5)
 par(mfrow = c(2, 3), bty = "n", mar = c(1,1,1,1))
 ## Tree
-plot(tree, cex = 0.8, main = "A - Simulated tree")
+plot(tree, cex = 0.8, main = "A - Starting tree")
 ## Normal matrix
-plot(char_diff_matrix_norm, type = "matrix", main = "B - Normal matrix", legend = FALSE)
+plot(char_diff_matrix_norm, type = "matrix", main = "B - Unperturbed matrix", legend = FALSE)
 ## Empty plot
@@
 ## Maximised matrix
-plot(char_diff_matrix_maxi, type = "matrix", main = "C - Maximised matrix", legend = FALSE)
+plot(char_diff_matrix_maxi, type = "matrix", main = "C - Minimised correlation matrix", legend = FALSE)
 ## Minimised matrix
-plot(char_diff_matrix_mini, type = "matrix", main = "D - Minimised matrix", legend = FALSE)
+plot(char_diff_matrix_mini, type = "matrix", main = "D - Maximised correlation matrix", legend = FALSE)
 ## Randomised matrix

### 2022-05-16 Analysis/Cluster/ClusterJobs.Rmd
sync

---
title: "Cluster Jobs"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Cluster Jobs

A quick 'n' dirty tutorial for (re)running jobs on the cluster.
The idea here is to get the missing consensus trees to finish the goddamn paper.
Here's a tutorial for doing so.
The data needed for that and the list of trees to rerun will be sent separately.

## What's in there?

### `/home/`

Once you log in, you automatically get redirected to `/home/`.
If I remember well, the absolute path is something like `/home/tguiller/`.
Once in there, you can navigate using the usual `ls`  and `cd`.
The folder of interest here should be somewhere like:

```
cd Charsim/Bayesian/
```

Not entirely sure about the spelling here since I'm doing this from the to of my head (though you'll figure it out easily, it's not like there's a shit ton of stuff on there).
Anyways, in there, you should find a tonne of files with ugly names.
The way the names work is pretty easy first set of digits with `t` is the number of taxa, second with `c` is the number of character, third is the chain number (starting with `0`) and then it's either the matrix, the job, the output, etc...
For example:

```
150t_1000c_0440.mbjob-bis.e268333
```

is the error output (`.e...`) of the 440st matrix with 150 taxa and 1000 characters for the MrBayes job (bis) (`.mbjob-bis`) that got assigned the cluster ID `268333`.

This folder will be for launching jobs and retrieving error messages.
If they don't fuck around with the cluster, you can ignore the error logs (it nearly always generate some anyways, sometimes the error being: "Error: you're using the cluster").

### `/work/`

This is where the stuff gets real.
The absolute path should be something like `/work/tguiller/` and then `Charsim/Bayesian/` (the same as for `/home/`).
This is where you can retrieve the jobs which is way more important/useful: it contains folders on the same chain name model as explained above (e.g. `25t_100c_0001/`).

These folders contain the MrBayes files produced by the job and the "scenario" considered.
I.e. how is the correlation modified:

 * `_norm` for the normal (base) scenarios
 * `_maxi` for the maximised differences scenarios
 * `_mini` for the minimised differences scenarios
 * `_rand` for the randomised differences scenarios

Usually the randomised and minimised are a pain to run (i.e. don't converge easily) but the other ones should be fine.
The file suffixes are the following

* `.ckp` or `.ckp~`: the checkpoint files for restarting the MCMC
* `.con.tre`: pretty critical, the consensus tree!
* `.log`: the log file (probably won't need to dwelve in there unless some huge crap happens)
* `.lstat`, `.pstat`, `vstat`, `tstat` and `.parts`: likelihood, posterior probabilities and partition files (ignore - but download!)
* `.run1.p` and `.run2.p` the posterior probabilities parameters (pretty critical!)
* `.run1.t` and `.run2.t` the posterior tree distribution (pretty critical!)

The `~` is for "previous" file.
For example if you have a `.ckp` file, it will rename it `.ckp~` before creating the new `.ckp` file.
This way if everything goes wrong, you can simply remove the corrupted file (e.g. the new `.ckp`) and rename the previous one by removing the `~` (e.g. `mv my_file.ckp~ my_file.ckp`).
Good guy Fred Ronquist to implement this!

> Note that the `~` will appear for unfinished chains (e.g. `my_file.run1.p~` and `my_file.run1.t~`); more about that later.

## Downloading stuff

For downloading the jobs and stuff you can either use the `ssh` with your normal `cp` and `mv` but I'll advice you to use a drag and drop client.
It's actually way more reliable looking at the structure of their fucking cluster (i.e. if you loose connection, it keeps track of what happened and what needs to be done).
Personally I use the [CyberDuck](https://cyberduck.io/) GUI which is safe, free, small and open.

## The drill

Here's what needs to be done:

  * checking the jobs status
  * checking finished jobs
  * Downloading the finished jobs
  * reruning eventual jobs

### Checking the job status

Easy one, once you're logged in, you can go with

```
qstat -a
```

That should give you info on the jobs running or pending (it can be completely empty of course).
Not much to do here but it's important to make sure that you don't download/modify folders of jobs that are running.

### Checking the finished jobs

For the list of jobs to check (see in the attached email) you have to check whether:

 * The job has produced a `.con.tre` file;   
 * Or, if not, whether the two associated `.run1.t` and `.run2.t` are past 5GB each

If it's any of these two conditions, we consider the job as finished and the files needs to be downloaded (see below).
Else, the jobs need to be rerun (see below).
In either case, you need to manual remove the temporary `~` files if they have been created for the five `.ckp`, `.run1.p`, `.run2.p`, `.run1.t` and `.run2.t` files.

### Downloading the finished jobs

Once the jobs are either done (`.con.tre` computed) or past 5GB per posterior tree file (`.run1.t`/`.run2.t`), download all the files for the concerned chain (e.g. `chainname_norm.*`).
For the jobs that reached >5GB per chain but without a consensus tree, you will need to compute the consensus tree via MrBayes (no rush on this step but I write the details down here anyways).

#### Computing the consensus trees

**This step is not priority**

If you have two posterior tree distributions that are > 5GB and still didn't converge, you'll need to manually compute the consensus tree.
To do so, you need to have MrBayes installed at the user level.

> To check if you have MrBayes at the user level, simply type `mb` in the terminal. If it runs MrBayes, simply quit (`q`), you're good. Else, [download/install MrBayes](https://sourceforge.net/projects/mrbayes/files/mrbayes/3.2.6/MrBayes-3.2.6_MACx64.zip/download?use_mirror=nchc) and move the executable from where you downloaded it into your user folder: `sudo mv MrBayes/mb /usr/local/bin/` (and `sudo mv MrBayes/mb-mpi /usr/local/bin/` if you want the mpi version there as well). Now you can call `mb` from wherever (check if it worked).

From the folder where you downloaded your chain results (e.g. say you've put that in `~/Downloads/Thomasshit/` where you'd have `~/Downloads/Thomasshit/150t_1000c_0440/` with all the downloaded files related to this chain within), you can copy/paste the [`make.con.tre.sh` script](https://github.com/TGuillerme/CharactersCorrelation/blob/master/Functions/shells/make.con.tre.sh) at the folder level (e.g.`~/Downloads/Thomasshit/`) and run it as follow:

```
sh make.con.tre.sh <chain> <tree>
```

With:

 * `<chain>` being the chain name (e.g. `150t_1000c_0440`)
 * `<tree>` being the tree "name" (either `norm`, `maxi`, `mini` or `rand` - without the `_`).

For example, for computing the `150t_1000c_0440_rand.con.tre` tree, you'd go:

```
sh make.con.tre.sh 150t_1000c_0440 rand
```

This takes couple of hours for a reasonable sized chain (e.g. ~ 5GB) so be sure to do it when you're not gonna move your machin.

### Reruning eventual jobs

For the jobs that are not finished, you need to re-generate a `.mbjob-bis` file and then submitting it.

#### Regenerating the `.mjob-bis` file

For that, navigate from the terminal into the `Jobs_in/` folder on your machine (from the zip I've sent you by email) and you can simply re-generate the `.mbjob` files by using:

```
sh rerun.missing.sh <chain> <broken> <missing>
```

With:

 * `<chain>` being the chain name (e.g. `150t_1000c_0440`)
 * `<broken>` being the number of the aborted (unfinished) chains (1 = norm, 2 = maxi, 3 = mini, 4 = rand)
 * `<missing>` being the chains that still needs to be ran (1 = missing rand, 2 = missing rand+mini, 3 = missing rand+mini+maxi, 4 = missing all)

For example, for the chain `150t_1000c_0440`, if you have the `_norm` tree computed (`150t_1000c_0440_norm.con.tre` present), the `150t_1000c_0440_maxi` and `150t_1000c_0440_mini` aborted (i.e. the `run1.p`, etc. files are present but not the consensus tree - and they didn't exceed 5GB), you'll need to rerun the `_maxi` and `_mini` chains and run the `_rand` one:

```
sh rerun.missing.sh 150t_1000c_0440 2,3 3
```

This will create the files for reruning `_maxi` and `_mini` and for running `_rand` (all in one job).
Concretely, this should generate three files:

 * `150t_1000c_0440_maxi.append` that's the matrix for appending the `_maxi` tree
 * `150t_1000c_0440_mini.append` same for the `_mini` tree
 * `150t_1000c_0440.mbjob-bis` the job file

Ignore the content of the `.append` files since it's generated automatically and should not be changed.
For the `150t_1000c_0440.mbjob-bis` file, the content should be pretty straightforwardly commented.
If you need to modify it for some reason (e.g. the missing chains are only `_maxi` and `_mini`), you can activate/deactivate some chains by commenting out the 4 associated lines (**DON'T DELETE ANY LINES**, the line numbering matters).
For example, for commenting out the `_rand` chain you can change the lines

```
pbsexec mpiexec mb $HOME/CharSim/Bayesian/150t_1000c_0440_rand.mbcmd
if [ -f "150t_1000c_0440_rand.con.tre" ] ; then echo "rand time out" ; else echo "rand aborted" ; fi
date
pbsdsh2 "cp $TMPDIR/* $WORK/CharSim/Bayesian/150t_1000c_0440/"
```

to

```
# pbsexec mpiexec mb $HOME/CharSim/Bayesian/150t_1000c_0440_rand.mbcmd
# if [ -f "150t_1000c_0440_rand.con.tre" ] ; then echo "rand time out" ; else echo "rand aborted" ; fi
# date
# pbsdsh2 "cp $TMPDIR/* $WORK/CharSim/Bayesian/150t_1000c_0440/"
```

Pretty straightforward.

> For jobs that don't need reruning (e.g. where the full MCMC is needed), you can still apply the same procedure but replacing `<broken>` by `0`. E.g. if you need to run the `_rand` tree, (not rerun) just go `sh rerun.missing.sh 150t_1000c_0440 1`

#### Resubmitting the job.

Once that's done, you'll need to transfer **all** the newly generated files (here the 3 ones listed above) on the server in `/home/tguiller/Charsim/Bayesian` (again not entirely sure about the path but you should be able to figure it out).
Overwrite any eventual file (that's the point of updating these fellows)!

After transferring the files, you can finally resubmit the jobs using the following (make sure you're in the right repo - `/home/tguiller/Charsim/Bayesian` - otherwise the cluster will do stupid things):

```
qsub <chain>.mbjob-bis
```

From the example above that'll be:

```
qsub 150t_1000c_0440.mbjob-bis
```

The cluster should then assign you a job ID number that you should keep track of it (an excel file would do the job) so you know where you job is.

And that's pretty much it, the jobs should be resubmitted like that until we get all consensus trees.
The jobs are set to run for 72 hours and can take some time in the queue so in an ideal world, you'll submit them on Monday and on Thursdays.

## Important stuff/tips

The cluster is an asshole so here are a couple of tips to double check or common mistakes:

 * Make sure to do the things in the right order (i.e. 1- retrieve/delete jobs; 2- write/upload new jobs; 3- submit new jobs). I often forgot to re-upload the jobs before submitting, best case it reran the same job up to the same point (loosing 72h) and worst case it decide to delete my stuff (again, the cluster is a douche!).
 * Make sure you delete the temporary jobs (`~`): MrBayes can deal with them no problems but they take space on the cluster and there's a space limit. Once you pass the limit, jobs results stay in the void and won't end up in your `work/` directory.

### 2022-05-16 Writing/CoverLetter.tex
added repeatability statement
sync

--- 2018-03-15 Writing/CoverLetter.tex
@@
 \usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
+\usepackage{hyperref}
 \usepackage[osf]{mathpazo}
-\signature{Thomas Guillerme \\ Martin Brazeau}
+\signature{Thomas Guillerme \\ Abigail Pastore \\ Martin Brazeau}
 \address{The University of Queensland \\School of Biological Sciences \\St Lucia QLD 4067, Australia \\guillert@tcd.ie}
@@
 
-Despite being at the base of almost every endeavour in modern Biology, the assumption made in tree inference that phylogenetic data is independent is probably mostly erroneous.
-In fact, since the advent of Cladistics in the 1950s, evolutionary biologists have made the assumption that characters used to build a phylogeny are independent despite the fact that we are looking at their variance-covariance to build phylogenetic hypothesis.
-This effect is somewhat negligible in molecular phylogenetics because the huge amount of available data is thought to temper down the correlation effect.
-However, in smaller morphological datasets, this can lead to pretty big biases and can easily favour on topology from another simply with a ``judicious'' choice of characters.
+The assumption made in tree inference that phylogenetic data is independent is probably mostly erroneous.
+This assumption about data correlation was stated since the advent of Cladistics in the 1950s but is nowadays mainly ignored in phylogenetic endeavour when using molecular data (where it can be neglected due to the huge amount of available data).
+In morphological datasets, however, where the amount of characters are much smaller, correlation can lead to significant biases especially by favouring some topologies due to a specific choice of characters (whether this is done consciously by researchers or not).
+With the ``resurgence'' of the use of morphological data in phylogenetic analysis (namely with the Total Evidence tip-dating method, Ronquist \textit{et al.}, 2012, Sys.
+Biol), it is thus important to acknowledge limitations and caveats that come with the use of such data, namely how correlation patterns in morphological datasets can effect tree inference.
 
-Our research article, entitled ``Influence of different modes of morphological character correlation on tree inference'', is - to our knowledge - the first to thoroughly test the effect of character correlation on tree topology using a thorough simulation framework ($>$ 100 CPU years worth).
-Using a novel metric to measure the amount of phylogenetic correlation in a matrix, we tested different scenario that either increase or decrease the amount of correlation of characters generated from a known ``true'' tree and compare how these levels of correlation affect our ability to recover this topology using both Maximum Parsimony and Bayesian inference.
+In our research article, entitled ``Influence of different modes of morphological characters correlation on tree inference'', we propose the first - to our knowledge - testing of the effect of characters correlation on tree topology based on a thorough simulation framework ($>$ 100 CPU years worth).
+Using a novel distance metric to measure direct phylogenetic correlation between characters, we simulated discrete morphological datasets that we then modified to increase or decrease the overall characters correlation within the datasets.
+We then measured the resulting topological differences (from both maximum parsimony and Bayesian trees) to assess the effect of characters correlation in the light of our known simulated parameters.
 
-Our results suggests that strong levels of correlation can negatively impact our ability to recover the correct tree topology and more so in Maximum Parsimony than in Baysian.
-However, it is encouraging to see that low levels of correlation do not especially impact the topology, at least not to the extant where researchers should specifically collect only uncorrelated characters.
-Finally, we show that, for smaller datasets, Maximum Parsimony and Bayesian inference are both pretty efficient to recover the correct topology, regardless of the amount of correlation
+Our results suggest that datasets that have higher levels of correlation between characters negatively impact our ability to recover the correct tree topology.
+This is to be contrasted with datasets that have low levels of correlation (i.e.
+that are more homoplasic) which can still recover the correct topology, depending on the amount of data and the inference method used.
+These results, although non-surprising to researchers in the field, are the first - to our knowledge - to actually quantify the effect of characters correlation and provides practical data and reusable code allowing researcher to effectively measure the limitations of their empirical datasets.
+Furthermore, we suggest that using our proposed new metric, one could improve bootstrap procedures in phylogenetic analysis by ``biasing'' the bootstrap to favour characters with high or low correlation, depending on the purpose of the phylogenetic search.
+Finally, the code and data for this whole analysis is entirely repeatable through \url{https://github.com/TGuillerme/CharactersCorrelation/}.
 

### 2022-05-16 Writing/Palaeontology/Tab_pooledscenarios_test.tex
sync

% latex table generated in R 4.1.3 by xtable 1.8-4 package
% Thu Mar 24 14:24:05 2022
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & maximised vs. minimised & 0.848 & 97587.500 & \textbf{0.000} \\ 
   & maximised vs. randomised & \textbf{0.954} & 80382.500 & \textbf{0.000} \\ 
   & minimised vs. randomised & 0.941 & 44161.000 & \textbf{0.000} \\ 
  Tr & maximised vs. minimised & \textbf{0.992} & 68919.500 & 0.839 \\ 
   & maximised vs. randomised & \textbf{0.982} & 64222.500 & 1.000 \\ 
   & minimised vs. randomised & \textbf{0.980} & 60447.000 & 0.713 \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for each maximised, minimised and randomised character correlation. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscenarios_test}
\end{table}

### 2022-05-16 Writing/Tables/Tab_pooledsmethods_test.tex
Final results tables
Updated results
bold main results
sync

--- 2018-03-27 Writing/Tables/Tab_pooledscharacters_test.tex
@@
   \hline
-RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
-   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
-   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
-  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
-   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
-   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
+RF & bayesian vs. parsimony & 0.939 & 192659.000 & 0.000 \\ 
+  Tr & bayesian vs. parsimony & 0.986 & 157677.500 & 0.041 \\ 
    \hline
 \end{tabular}
-\caption{Difference between the pooled number of characters.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)}
-\label{Tab_pooledscharacters_test}
+\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the two different inference methods: bayesian or maximum parsimony.
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).}
+\label{Tab_pooledsmethods_test}
 \end{table}

### 2022-05-16 Writing/Tables/Tab_pooledstaxa_test.tex
Final results tables
Updated results
Single comparisons done
sync

--- 2018-03-27 Writing/Tables/Tab_pooledscharacters_test.tex
@@
   \hline
-RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
-   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
-   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
-  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
-   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
-   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
+RF & t25 vs. t75 & 0.949 & 79945.000 & 0.000 \\ 
+   & t25 vs. t150 & 0.954 & 75906.000 & 0.000 \\ 
+   & t75 vs. t150 & 0.984 & 60328.000 & 0.654 \\ 
+  Tr & t25 vs. t75 & 0.960 & 72814.000 & 0.024 \\ 
+   & t25 vs. t150 & 0.961 & 72716.000 & 0.027 \\ 
+   & t75 vs. t150 & 0.980 & 64637.000 & 1.000 \\ 
    \hline
 \end{tabular}
-\caption{Difference between the pooled number of characters.
-Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)}
-\label{Tab_pooledscharacters_test}
+\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of taxa: 25 (t25), 75 (t75) and 150 (t150).
+Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).}
+\label{Tab_pooledstaxa_test}
 \end{table}

### 2022-05-16 Functions/shells/con.tree.sh

@@ -1,52 +1,0 @@
-#!/bin/sh
-
-##########################
-#Shell script for building up the MrBayes consensus trees
-##########################
-#SYNTAX:
-#sh con.tree.sh <chain_name>
-#with:
-#<chain_name> the name of the chain containing the norm/mini/maxi trees
-##########################
-#----
-#guillert(at)tcd.ie - 2016/11/18
-##########################
-
-## Step 1 - INPUT
-
-## Input values
-chain_name=$1
-
-## Renaming the matrices
-cp ${chain_name}_norm.nex ${chain_name}_norm
-cp ${chain_name}_maxi.nex ${chain_name}_maxi
-cp ${chain_name}_mini.nex ${chain_name}_mini
-
-## Generating the template file for calculating the consensus tree in MrBayes
-echo "set autoclose=yes nowarn=yes;
-execute <CHAIN>;
-sumt Relburnin=YES Burninfrac=0.25;" > contree.template
-
-## Copying the commands to the matrices
-cat contree.template | sed 's/<CHAIN>/'"${chain_name}"'_norm/g' > ${chain_name}_norm_contree.cmd
-cat contree.template | sed 's/<CHAIN>/'"${chain_name}"'_maxi/g' > ${chain_name}_maxi_contree.cmd
-cat contree.template | sed 's/<CHAIN>/'"${chain_name}"'_mini/g' > ${chain_name}_mini_contree.cmd
-
-## Running the trees
-echo "Generating the consensus tree:"
-mb ${chain_name}_norm_contree.cmd > /dev/null
-echo "."
-mb ${chain_name}_maxi_contree.cmd > /dev/null
-echo "."
-mb ${chain_name}_mini_contree.cmd > /dev/null
-echo "."
-echo "Done"
-
-## Cleaning the files
-rm contree.template
-rm ${chain_name}_norm
-rm ${chain_name}_maxi
-rm ${chain_name}_mini
-rm ${chain_name}_norm_contree.cmd
-rm ${chain_name}_maxi_contree.cmd
-rm ${chain_name}_mini_contree.cmd

### 2023-03-30 Writing/Palaeontology/draft.tex
Updated manuscript

@@
 \section{Introduction}
-The last two decades have witnessed a resurgence of interest in the use of morphological character data in phylogenetic studies.
-This owes in large part to the advent of new models, methods and software \citep{lewisa2001,ronquista2012,Ronquist2012mrbayes,heath2014fossilized} and the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
-While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any investigations into their statistical properties.
+%The last two decades have witnessed a resurgence of interest in the use of morphological character data in phylogenetic studies.
+%This owes in large part to the advent of new models, methods and software \citep{lewisa2001,ronquista2012,Ronquist2012mrbayes,heath2014fossilized} and the use of fossils to undertake at least partial reconstructions of phylogenetic trees, especially where ancestral states reconstructions or absolute calibrations of divergence times are necessary.
+%While there is a general appreciation of the limits of morphological data, they are frequently dismissed without any investigations into their statistical properties.
 %MB: I think there are a few classic papers we could cite here.
@@
 TG: which ones?
-As morphological data are likely to continue to play an extensive role in phylogenetic analysis,  %MB: cite Nature paper here. TG: which one?
-it is essential to understand the circumstances under which morphological data might be expected to lead to erroneous phylogenetic inferences.
-This opens up possibilities for analysing problematic datasets and possibly proposing new confidence measures in phylogenetic datasets.
+%As morphological data are likely to continue to play an extensive role in phylogenetic analysis,  %MB: cite Nature paper here. TG: which one?
+%it is essential to understand the circumstances under which morphological data might be expected to lead to erroneous phylogenetic inferences.
+%This opens up possibilities for analysing problematic datasets and possibly proposing new confidence measures in phylogenetic datasets.
 
-The non-independence of large numbers of morphological characters is often cited in anticipation of problems with morphological data \cite[e.g.][]{Davalos01072014, ZouConvergence}.
-The assumption of character independence is central to phylogenetic inference methods such as maximum likelihood and maximum parsimony \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
-Non-independence of characters implies that some suites of characters are expected to transform in highly correlated ways.
-Not only does this violate the assumptions of phylogenetic analysis, but implies that if convergence has occurred in correlated sets of characters, then this would lead to spurious cladistic groupings.
-Here we distinguish three different modes of forms of character correlation (described in more details in \citealt{wilkison1992ordered,wilkinson1995character,wilkinson1995coping}):
+%% MDB: I think we can just start here. The paragraph above is completely extraneous for readers of Palaeontology.
+The assumption of character independence is central to phylogenetic inference \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
+Contemporary methods like parsimony and maximum likelihood evaluate phylogenetic trees under the assumption that the input characters represent independent and identically distributed data points.
+However, there are many cases in which morphological data can be reasonably suspected to have evolved in a non-independent way \cite[e.g.][]{Davalos01072014, ZouConvergence}.
+Morphological character inter-dependencies can occur as a result of factors such as functional integration, underlying genetic modularity, or a  hierarchical relationship where one character depends directly on the state of another.
+These relationships can create suites of similar character distributions that may bias phylogenetic analysis and deliver spurious results whose problems cannot be easily detected. 
+At the same time, it is usually impossible to know whether similarities are homologous or the result of convergent evolution without prior knowledge of phylogeny itself.
+This creates an impasse for morphologists, especially for the phylogenies of extinct groups where there is usually no likelihood of recovering molecular sequence data.
+
+In this paper, we explore the circumstances under which character correlation might be expected to be a problem.
+Despite the realistic \textcolor{blue}{hurdles} posed by character correlations, it remains unclear that can be expected to be a pervasive problem.
+It is therefore worthwhile to characterise different types of character correlations that can impact comparative data.
+Here we distinguish four different modes of forms of character correlation (described in more details in \citealt{wilkinson1992ordered,wilkinson1995comparison,wilkinson1995coping}):
 
 \begin{itemize}
-    \item \textbf{Intra-organismal dependence:} this is the result of an intrinsic biological link between two characters.
-    For example the occlusal surface of two opposing molars will be expected to directly covary.
+
+    \item \textcolor{blue}{\textbf{Intra-organismal dependence:}} this is the result of a biological link between two characters that exists within the organism itself.
+The main cause of this would be pleiotropic genetic effects.
+This kind of dependence is often described in terms of integration and modularity, which can range from genetic to functional dependencies.
 
@@
     For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
+This type of dependence can be thought of in terms of functional suites of characters coevolving.
 
@@
 using inapplicable tokens; ][]{Brazeau2011} and that these choices affect results more or less depending on the estimation method (parsimony methods contrast with probabilistic methods were ancestral conditions are respectively fixed or not).}
+
+    \textcolor{blue}{\item \textbf{Phenomenological dependence:} This occurs where two morphological characters are expressed (often semantically) as two distinct things, but refer ultimately to the same phenomenon within an organism.
+For instance, occlusal patterns of mammalian dentitions would be expected to co-vary precisely in opposing mammal dentitions \cite{sansom2017dental}.
+    A similar example is the relative position of the mandible in early gnathostomes and the course of its innervation.
+As the mandibular arch moves position relative to the rest of the braincase, the innervation patterns are expected to follow suit.
+Treating these as separate character may create spurious support for placoderm paraphyly \citep[e.g.][]{brazeau2009braincase,king2017bayesian}.}
+
 \end{itemize}
 
-\noindent Of course, the three sources of dependence can have an interaction: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding.
+\noindent Of course, the four sources of dependence are not necessarily exclusive: characters describing the left and right lower/upper molars will have \textcolor{blue}{phenomenological} dependence due to the modularity of the molars, their shared history and the duplicated coding.
 Logical dependence, however, is easily distinguished prior to phylogenetic inference, while intra-organismal and evolutionary dependence are much harder to parse.
@@
 linear measurements or morphometric geometric measurements) or non-morphological (e.g.
-behaviour, or - much more commonly used - molecular data).}
+behaviour, or - much more commonly used - molecular data, \citealt{sansom2017dental}).}
 
@@
     \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
-    We drew the number of character states with a probability of $0.85$ for binary characters \textcolor{blue}{(using a $2\times2$ Q matrix)} and $0.15$ for three state characters (\textcolor{blue}{using a $3\times3$ Q matrix);} \citealt{Guillerme2016146,ZouConvergence}).
-%TG: this Q matrix bit is very jargony and unsincere...
+    We drew the number of character states with a probability of $0.85$ for binary characters \textcolor{blue}{(using a $2\times2$ Q matrix)} and $0.15$ for three state characters (\textcolor{blue}{using a $3\times3$ Q matrix;} \citealt{Guillerme2016146,ZouConvergence}).
 \end{itemize}
@@
 
-We standardised each character by arbitrarily modifying their character state tokens (or symbols) by \textcolor{blue}{order of appearance in the matrix - for example for the first character (1st column) in the matrix, if the state coded for the first OTU (1st row) was \texttt{3}, it was translated into \texttt{1} for the whole character (across the 1st column); if the state coded for the second OTU (2nd row) was \texttt{0}, it was translated into \texttt{2} for the whole character, etc.}
-We replaced all the occurrences of the first token to be \texttt{1}, the second to be \texttt{2}, etc.
+We standardised each character by arbitrarily modifying their character state tokens (or symbols) by \textcolor{blue}{order of appearance in the matrix - for example for the first character (1st column) in the matrix, if the state coded for the first OTU (1st row) was \texttt{3}, it was translated into \texttt{1} for the whole character (across the 1st column); if the state coded for the second OTU (2nd row) was \texttt{0}, it was translated into \texttt{2} for the whole character, etc.
+In other words, we} replaced all the occurrences of the first token to be \texttt{1}, the second to be \texttt{2}, etc.
 This procedure allows comparison of characters regardless of the significance given to their tokens (following the \textit{xyz} notation in \citealt{felsenstein2004inferring}; as used in \citealt{Davalos01072014}).
@@
 Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$) with an variable rate prior an exponential (0.5) shape.
-\textcolor{blue}{We’ve used the Mk model with ascertainment bias (Mkv) to correct for the fact that no invariant characters were present in the matrix (as is often the case with morphological data) allowing the model to estimate the correct evolutionary rates (see above).
+\textcolor{blue}{We used the Mk model with ascertainment bias (Mkv) to correct for the fact that no invariant characters were present in the matrix (as is often the case with morphological data) allowing the model to estimate the correct evolutionary rates (see above).
 We used four discrete character rates categories making the assumption that each character has an independent $\frac{1}{4}$ chance to be in the $1^{st}$, $2^{nd}$, $3^{rd}$ or $4^{th}$ quartile of the rate distribution.
@@
 This is especially true in large matrices where we advise to reduce character difference by using bootstrapping with ``bias'' whereas this may not be major problem in modest size matrices (25 taxa; 100 to 350 characters).
-However, in empirical datasets, if character difference is not driven by selection (e.g.
+% However, in empirical datasets, if character difference is not driven by selection (e.g.
 pleiotropy) correlation is likely be cancelled out if the correlated characters are randomly distributed with respect to traits.

### 2023-03-30 Writing/Palaeontology/Reviews/response_to_reviewers.tex
sync
Update response_to_reviewers.tex

@@
 
-We are very grateful to the three referees for their helpful and constructive comments, which we believe have helped me to significantly improve the manuscript.
-Unfortunately, due to circumstances out of our control (pandemic, moving continent and new job), it took us way longer to reply to the reviewers requests as we could have done.
-Fortunately, however, we have taken all of their comments on board, and respond to their points in details below.
+We are very grateful to the three referees for their helpful and constructive comments, which we believe have helped significantly improve the manuscript.
+Unfortunately, due to circumstances out of our control (pandemic, moving continent and new job), it took us far longer to reply to the reviewers requests than anticipated.
+Fortunately, however, we have taken all of their comments on board, and respond to their points in detail below.
 
-For improving clarity in this document, reviewers' comments are displayed in blue, our responses in normal text and the changes to the manuscript in italic.
+For improving clarity in this document, reviewer comments are displayed in blue, our responses in normal text and the changes to the manuscript in italic.
 Additionally, we've uploaded two versions of the revised manuscript: both have the exact same content but one has all the changes and additions to the main text highlighted in yellow to help with reviewing.
@@
 
-%@@@TODO-1: add a line in the intro explaining this? 
-%TG: something like that "Note that investigating the shared evolutionary history is often part of the goal of phylogenetic analyses."
+This is a very helpful point.
+We have adopted terminological consistency between the abstract and the introduction.
+We have now standardised the language to "evolutionary dependence" throughout the text and abstract.
 
@@
 
+This manuscript was written in \LaTeX\ and no proprietary softwares (e.g.
+Word) where used.
+We have added an .odt format version of the manuscript as well as all the figures as separated files with the figure captions at the end of the manuscript.
+
 \item{\textcolor{blue}{Please respond directly to all referee comments, including the technical comments below.
@@
 
+We have now added our ORCiDs to the author list.
+
 \end{itemize}
@@
 
+We have now added a proper reference for the data and code.
+
 \item{\textcolor{blue}{We no longer upload any supplementary information on our publisher's website.
@@
 Dryad encourages uploading data in multiple formats if styling is important.}}
+
+We have now added all the supplementary documents as a FigShare item with it's own DOI.
 
@@
 
+We have now fixed this issue.
+
 \subsection{Figures}
+
+TODO: @@@ I'll do that later
+
 \begin{itemize}
@@
 \subsection{Tables}
+
+TODO: @@@ I'll do that later
+
 \begin{itemize}
@@
 
-% @@@TODO-6: Check if that works with comments: https://docs.google.com/document/d/1ED7b-uxtEzJ5fUUXugBYMIq88zx2ySY9fHcJDcUekAs/edit?disco=AAAAMGcMrqU
-%TG: still to do. But I think it's kind of in there...
+We agree with these two broad categories of correlation types, but we didn’t find them adequate to characterise the range of possible correlations that could be encountered.
+Given the nature of our work, we thought ‘atomizing’ the types of correlations possible would be useful in preventing misinterpretations of our results.
+Nevertheless, we think the reviewer has raised a really important point that we do not wish to dismiss.
+We have modified our discussion of the types of correlation and made specific reference to the two types mentioned by the referee.
 
@@
 
-%@@@ TODO:
+In this sense, this metric is used to compare \textit{characters} with each other to create matrices with more or less induced correlation.
+Our character correlation metric (character difference) is used as an optimisable variable to increase or decrease correlation in our simulated matrices.
+We believe that the reviewer here suggests comparing character distances (phenetic distance) between \textit{taxa} (cf.
+between characters).
+We are not sure what this would bring to the present manuscript since this was not the focal point of this study.
+In fact, one could argue that this phenetic distance is somehow recovered when inferring the topologies based on the modified and un-modified matrices.
+But again, this is not what we where trying to explore here.
 
@@
 
-\textcolor{blue}{Sadly however, I am skeptical that this paper really addresses the central issue with character correlation and inferring phylogenies.}
+\textcolor{blue}{Sadly however, I am skeptical that this paper really addresses the central issue with character correlation and inferring phylogenies.
+Particularly, their main conclusion is that character correlation is a 'good' thing, i.e.
+having increased character correlation actually results in more accurate topologies.}
 
-We are also sad that we did not address the central issue(s) with character correlation although that was never our intention.
-We have now specified more clearly that what this paper ``offers is an idea of what happens when there is some amount of correlated change, but we just do not know how much.'' (as correctly worded by referee 2).
-We believe that the issue(s!) with character correlation and inferring phylogenies have to be addressed in an incremental way and hope that this manuscript will be a step forward in that direction.
-We have now clarified that throughout the manuscript:
-
-% @@@TODO-7.
-
-\textcolor{blue}{Particularly, their main conclusion is that character correlation is a 'good' thing, i.e. having increased character correlation actually results in more accurate topologies.}
-
-We do not claim this in the manuscript but rather that character correlation can, for specific datasets and methods, more or less bias topological inference. We have now clarified this
-
-% @@@TODO-8.
+This isn't quite a correct interpretation of the paper.
+We appreciate therefore that our explanation may not have been entirely clear and must have caused some confusion.
+We think, however, it would be an unreasonable standard to expect one paper to solve a monumental problem such as character correlation effects on phylogeny.
+We don't propose to solve problem in its totality.
+We have now revised the introduction and discussion to clarify the message of the paper; if character correlation is highly stochastic within a dataset, the individual correlations effectively cancel each other out.
+Whereas, if correlation is intentionally decreased by the investigator, this can seriously bias the results.
+Similarly, if it is maximized by the investigator, as expected, it will also bias the result.
 
@@
 We do agree that these are valid limitations from this study and have acknowledged them more clearly:
-Third, by not focusing on modeling the sources of character correlation specifically (see section “Distinction between different character correlations” above), we do not allow to distinguish the specific effect of each correlation for downstream analyses (see Sansom et al.
+
+\textit{Third, by not focusing on modeling the sources of character correlation specifically (see section “Distinction between different character correlations” above), we do not allow to distinguish the specific effect of each correlation for downstream analyses (see Sansom et al.
 2017 for such an application on dental morphological data).
 However, these sources are often not available to authors, either because of the data availability or the resources available to the workers (i.e.
-time and money) and thus we here propose an approach showing the effect of character correlation when the sources of correlation are effectively unknown.
-Lines @@@
-% @@@TODO-9
+time and money) and thus we here propose an approach showing the effect of character correlation when the sources of correlation are effectively unknown.} Lines @@@
 
@@
 linear measurements or morphometric geometric measurements) or non-morphological (e.g.
-behaviour, or - much more commonly used - molecular data).} Lines @@@
-% @@@TODO-10: add ref to Sansom somewhere in the discussion.
+behaviour, or - much more commonly used - molecular data, Sansom et al.
+2017.} Lines @@@
 
@@
 
-We do agree that these are valid limitations from this study and have acknowledged them more clearly:
+We do agree that these are valid limitations from this study and have acknowledged them more clearly on lines @@@ as highlighted above.
 
-\textit{Third, by not focusing on modeling the sources of character correlation specifically (see section “Distinction between different character correlations” above), we do not allow to distinguish the specific effect of each correlation for downstream analyses (see Sansom et al.
+% This is the text it should refer to: \textit{Third, by not focusing on modeling the sources of character correlation specifically (see section “Distinction between different character correlations” above), we do not allow to distinguish the specific effect of each correlation for downstream analyses (see Sansom et al.
 2017 for such an application on dental morphological data).
@@
 
-% @@@TODO-11: specify this in the conclusion
+We have removed this sentence altogether since it was not impactful for the last part of the conclusion.
+We agree though that it would have been interested to look at how correlation arise between traits in organisms (e.g.
+through drift or/and selection) and how that would translate into discrete morphological characters but that was outside of the scope of this project.
 
@@
 
-We’ve added the comparisons between the trees from the modified matrices and the "starting tree" in the supplementary materials. %TG: is that done?
+We’ve added the comparisons between the trees from the modified matrices and the "starting tree" in the supplementary materials.
 However, as stated above, what these results actually show is the effect of A: changing the character correlation + B: the effect of phylogenetic inference + C: the effect of simulating discrete morphological characters.
@@
 
-% @@@TODO-13 we have changed the font size for the figures/tables captions and the main text.
+We have changed this in this revised version.
 
@@
 We have fixed the typo in figure 1 and checked the rest of the manuscript more thoroughly.
-%TG: TODO
 
